### 2020-06-10 analysis.Rmd
update

---
title: "analysis"
output: html_document
---

### Prepare environment
```{r setup, include=FALSE}
knitr::opts_chunk$set(include = FALSE)
library(bipartite)
library(mgcv)
library(mgcViz)
#library(lme4)
#library(lmerTest)
library(taxize)
library(tidyverse)
library(lubridate)
library(GGally)
```

### Define functions
#### Colony phase breakpoint functions
```{r include=FALSE}
# Find breakpoints on GDD scale
breakpoints_gdd <- function(x, year, t1, t2) {

  data <- filter(x, year == !!year)

  # Logistic regression for first break point (founding to buildup)
  glm1.low <- glm(q.prop ~ gdd.cum,
                  family = quasibinomial,
                  data = filter(data, elev.class2 == "low"))

  glm1.mid <- glm(q.prop ~ gdd.cum,
                  family = quasibinomial,
                  data = filter(data, elev.class2 == "mid"))

  glm1.high <- glm(q.prop ~ gdd.cum,
                   family = quasibinomial,
                   data = filter(data, elev.class2 == "high"))

  # Logistic regression for second break point (buildup to reproductive)
  glm2.low <- glm(r.prop ~ gdd.cum,
                  family = quasibinomial,
                  data = filter(data, elev.class2 == "low"))

  glm2.mid <- glm(r.prop ~ gdd.cum,
                  family = quasibinomial,
                  data = filter(data, elev.class2 == "mid"))

  glm2.high <- glm(r.prop ~ gdd.cum,
                   family = quasibinomial,
                   data = filter(data, elev.class2 == "high"))

  ### Extract gdd corresponding to desired proportion; based on https://stackoverflow.com/questions/32040504/regression-logistic-in-r-finding-x-value-predictor-for-a-particular-y-value

  getX <- function(x, y) {
    (log(y/(1-y)) - coef(x)[1])/coef(x)[2]
    }

  break1.low <- getX(glm1.low, t1) %>%
    data.frame() %>%
    select(gdd = 1) %>%
    mutate(elev.class2 = "low",
           breakID = "break1")

  break1.mid <- getX(glm1.mid, t1) %>%
    data.frame() %>%
    select(gdd = 1) %>%
    mutate(elev.class2 = "mid",
           breakID = "break1")

  break1.high <- getX(glm1.high, t1) %>%
    data.frame() %>%
    select(gdd = 1) %>%
    mutate(elev.class2 = "high",
           breakID = "break1")

  break2.low <- getX(glm2.low, t2) %>%
    data.frame() %>%
    select(gdd = 1) %>%
    mutate(elev.class2 = "low",
           breakID = "break2")

  break2.mid <- getX(glm2.mid, t2) %>%
    data.frame() %>%
    select(gdd = 1) %>%
    mutate(elev.class2 = "mid",
           breakID = "break2")

  break2.high <- getX(glm2.high, t2) %>%
    data.frame() %>%
    select(gdd = 1) %>%
    mutate(elev.class2 = "high",
           breakID = "break2")

  # Join into data frame, convert elev.class2 to ordered factor, spread
  breaks <- full_join(break1.low, break1.mid) %>%
    full_join(break1.high) %>%
    full_join(break2.low) %>%
    full_join(break2.mid) %>%
    full_join(break2.high) %>%
    mutate(elev.class2 = factor(elev.class2, levels = c("high", "mid", "low"))) %>%
    spread(breakID, gdd)

  #
  out <- data %>%
    full_join(breaks) %>%
    gather(caste, count, -c(site, date, year, q.prop, r.prop,
                          gdd.cum, yday, elev.class, elev.class2, 
                          management, transect, elev.mean, 
                          lat, lon, break1, break2)) %>%
  group_by(site, date) %>%
  mutate(prop = count/sum(count))
}

# Find breakpoints on Julian scale
breakpoints_yday <- function(x, year, t1, t2) {

  data <- filter(x, year == !!year)

  # Logistic regression for first break point (founding to buildup)
  glm1.low <- glm(q.prop ~ yday,
                  family = quasibinomial,
                  data = filter(data, elev.class2 == "low"))

  glm1.mid <- glm(q.prop ~ yday,
                  family = quasibinomial,
                  data = filter(data, elev.class2 == "mid"))

  glm1.high <- glm(q.prop ~ yday,
                   family = quasibinomial,
                   data = filter(data, elev.class2 == "high"))

  # Logistic regression for second break point (buildup to reproductive)
  glm2.low <- glm(r.prop ~ yday,
                  family = quasibinomial,
                  data = filter(data, elev.class2 == "low"))

  glm2.mid <- glm(r.prop ~ yday,
                  family = quasibinomial,
                  data = filter(data, elev.class2 == "mid"))

  glm2.high <- glm(r.prop ~ yday,
                   family = quasibinomial,
                   data = filter(data, elev.class2 == "high"))

  ### Extract gdd corresponding to desired proportion; based on https://stackoverflow.com/questions/32040504/regression-logistic-in-r-finding-x-value-predictor-for-a-particular-y-value

  getX <- function(x, y) {
    (log(y/(1-y)) - coef(x)[1])/coef(x)[2]
    }

  break1.low <- getX(glm1.low, t1) %>%
    data.frame() %>%
    select(day = 1) %>%
    mutate(elev.class2 = "low",
           breakID = "break1")

  break1.mid <- getX(glm1.mid, t1) %>%
    data.frame() %>%
    select(day = 1) %>%
    mutate(elev.class2 = "mid",
           breakID = "break1")

  break1.high <- getX(glm1.high, t1) %>%
    data.frame() %>%
    select(day = 1) %>%
    mutate(elev.class2 = "high",
           breakID = "break1")

  break2.low <- getX(glm2.low, t2) %>%
    data.frame() %>%
    select(day = 1) %>%
    mutate(elev.class2 = "low",
           breakID = "break2")

  break2.mid <- getX(glm2.mid, t2) %>%
    data.frame() %>%
    select(day = 1) %>%
    mutate(elev.class2 = "mid",
           breakID = "break2")

  break2.high <- getX(glm2.high, t2) %>%
    data.frame() %>%
    select(day = 1) %>%
    mutate(elev.class2 = "high",
           breakID = "break2")

  # Join into data frame, convert elev.class2 to ordered factor, spread
  breaks <- full_join(break1.low, break1.mid) %>%
    full_join(break1.high) %>%
    full_join(break2.low) %>%
    full_join(break2.mid) %>%
    full_join(break2.high) %>%
    mutate(elev.class2 = factor(elev.class2, levels = c("high", "mid", "low"))) %>%
    spread(breakID, day)

  #
  out <- data %>%
    full_join(breaks) %>%
    gather(caste, count, -c(site, date, year, q.prop, r.prop,
                          gdd.cum, yday, elev.class, elev.class2, 
                          management, transect, elev.mean, 
                          lat, lon, break1, break2)) %>%
  group_by(site, date) %>%
  mutate(prop = count/sum(count))
}

# Find Juian breakpoints across all elevations
breakpoints_yday_all <- function(x, year, t1, t2) {

  data <- filter(x, year == !!year)

  # Logistic regression for first break point (founding to buildup)
  glm1 <- glm(q.prop ~ yday,
                  family = quasibinomial,
                  data = data)

  # Logistic regression for second break point (buildup to reproductive)
  glm2 <- glm(r.prop ~ yday,
                  family = quasibinomial,
                  data = data)

  ### Extract day corresponding to desired proportion; based on https://stackoverflow.com/questions/32040504/regression-logistic-in-r-finding-x-value-predictor-for-a-particular-y-value

  getX <- function(x, y) {
    (log(y/(1-y)) - coef(x)[1])/coef(x)[2]
    }

  break1 <- getX(glm1, t1) %>%
    data.frame() %>%
    select(day = 1) %>%
    mutate(elev.class2 = "low",
           breakID = "break1")

  break2 <- getX(glm2, t2) %>%
    data.frame() %>%
    select(day = 1) %>%
    mutate(elev.class2 = "low",
           breakID = "break2")

  # Join into data frame, convert elev.class2 to ordered factor, spread
  breaks <- full_join(break1, break2) %>%
    spread(breakID, day)

  #
  out <- data %>%
    full_join(breaks) %>%
    gather(caste, count, -c(site, date, year, q.prop, r.prop,
                          gdd.cum, yday, elev.class, elev.class2, 
                          management, transect, elev.mean, 
                          lat, lon, break1, break2)) %>%
  group_by(site, date) %>%
  mutate(prop = count/sum(count))
}

# Binomial ggplot smoother
binomial_smooth <- function(...) {
  geom_smooth(method = "glm", method.args = list(family = "quasibinomial"), se = FALSE, ...)
}
```

#### Rarefied network analysis
```{r, include=FALSE}
# Rare_net function: rarefies networks by iteratively subsampling more frequently sampled networks to match the transect count of the least frequently sampled one
# species level
rare_net_sp <- function(net, n, iter) {

  out <- data.frame()

  for(i in 1:iter) {
    temp <- net %>%
      select(site, date) %>%
      unique() %>%
      group_by(site) %>%
      sample_n(n) %>%
      left_join(net) %>%
      group_by(site, bb.sp, plant.sp) %>% # group by site, date, bee species, plant species
      summarize(freq = n()) %>% # calculate interaction frequency per species pair within each site
      group_by(site) %>%
      filter(length(unique(bb.sp)) > 1 & length(unique(plant.sp)) > 1) %>%
      ungroup() %>%
      filter(freq > 1) %>% # remove singletons
      dplyr::select(higher = bb.sp, lower = plant.sp, webID = site, freq) %>% # rename columns to match bipartite's expectations
      data.frame() %>% # convert to data frame
      frame2webs() %>% # convert to to bipartite web
      map(specieslevel, level = "higher", index = c("d", "species strength", "proportional generality")) %>%
      map(rownames_to_column) %>%
      map(as_tibble) %>%
      bind_rows(.id = "site") %>% # Collapse list into big data frame
      rename(sp = rowname) %>%
      mutate(iteration = rep(i, n()))

    out <- bind_rows(out, temp)
  } 
  out <- out %>%
    gather(metric, value, -c(sp, site, iteration)) %>% # get mean value across iterations for each metric
    group_by(sp, site, metric) %>%
    summarize(value = mean(value)) %>%
    left_join(site_data) %>%
    left_join(bb_traits, by = c("sp" = "bb.sp"))
}

# network level
rare_net_nt <- function(net, n, iter) {

  out <- data.frame()

  for(i in 1:iter) {
    temp <- net %>%
      select(site, date) %>%
      unique() %>%
      group_by(site) %>%
      sample_n(n) %>%
      left_join(net) %>%
      group_by(site, bb.sp, plant.sp) %>% # group by site, date, bee species, plant species
      summarize(freq = n()) %>% # calculate interaction frequency per species pair within each site
      group_by(site) %>%
      filter(length(unique(bb.sp)) > 1 & length(unique(plant.sp)) > 1) %>%
      ungroup() %>%
      dplyr::select(higher = bb.sp, lower = plant.sp, webID = site, freq) %>% # rename columns to match bipartite's expectations
      data.frame() %>% # convert to data frame
      frame2webs() %>% # convert to to bipartite web
      map(networklevel, weighted = TRUE, level = "higher", index = c("H2", "weighted NODF", "niche overlap", "C score", "generality", "web asymmetry", "weighted connectance")) %>% # We probably care mainly about the BB level of the network
      data.frame() %>%
      t() %>%
      data.frame() %>%
      rownames_to_column() %>%
      dplyr::select(site = rowname, everything()) %>%
      as_tibble() %>%
      mutate(iteration = rep(i, n()))

    out <- bind_rows(out, temp)
  } 
  out <- out  %>%
      gather(metric, value, -c(site, iteration)) %>% # get mean value across iterations for each metric
      group_by(site, metric) %>%
      summarize(value = mean(value)) %>%
      left_join(site_data)
}

```

### Rarefied network analysis by floral k.type instead of floral species
```{r}
# Rare_net function: rarefies networks by iteratively subsampling more frequently sampled networks to match the transect count of the least frequently sampled one
# species level
rare_net_sp_k.type.s <- function(net, n, iter) {

  out <- data.frame()

  for(i in 1:iter) {
    temp <- net %>%
      select(site, date) %>%
      unique() %>%
      group_by(site) %>%
      sample_n(n) %>%
      left_join(net) %>%
      group_by(site, bb.sp, k.type.s) %>% # group by site, date, bee species, plant species
      summarize(freq = n()) %>% # calculate interaction frequency per species pair within each site
      na.omit() %>%
      group_by(site) %>%
      filter(length(unique(bb.sp)) > 1 & length(unique(k.type.s)) > 1) %>%
      ungroup() %>%
      filter(freq > 1) %>% # remove singletons
      dplyr::select(higher = bb.sp, lower = k.type.s, webID = site, freq) %>% # rename columns to match bipartite's expectations
      data.frame() %>% # convert to data frame
      frame2webs() %>% # convert to to bipartite web
      map(specieslevel, level = "higher", index = c("d", "species strength", "proportional generality")) %>%
      map(rownames_to_column) %>%
      map(as_tibble) %>%
      bind_rows(.id = "site") %>% # Collapse list into big data frame
      rename(sp = rowname) %>%
      mutate(iteration = rep(i, n()))

    out <- bind_rows(out, temp)
  } 
  out <- out %>%
    gather(metric, value, -c(sp, site, iteration)) %>% # get mean value across iterations for each metric
    group_by(sp, site, metric) %>%
    summarize(value = mean(value)) %>%
    left_join(site_data) %>%
    left_join(bb_traits, by = c("sp" = "bb.sp"))
}

# network level
rare_net_nt_k.type.s <- function(net, n, iter) {

  out <- data.frame()

  for(i in 1:iter) {
    temp <- net %>%
      select(site, date) %>%
      unique() %>%
      group_by(site) %>%
      sample_n(n) %>%
      left_join(net) %>%
      group_by(site, bb.sp, k.type.s) %>% # group by site, date, bee species, plant species
      summarize(freq = n()) %>% # calculate interaction frequency per species pair within each site
      na.omit() %>%
      group_by(site) %>%
      filter(length(unique(bb.sp)) > 1 & length(unique(k.type.s)) > 1) %>%
      ungroup() %>%
      dplyr::select(higher = bb.sp, lower = k.type.s, webID = site, freq) %>% # rename columns to match bipartite's expectations
      data.frame() %>% # convert to data frame
      frame2webs() %>% # convert to to bipartite web
      map(networklevel, weighted = TRUE, level = "higher", index = c("H2", "weighted NODF", "niche overlap", "C score", "generality", "web asymmetry", "weighted connectance")) %>% # We probably care mainly about the BB level of the network
      data.frame() %>%
      t() %>%
      data.frame() %>%
      rownames_to_column() %>%
      dplyr::select(site = rowname, everything()) %>%
      as_tibble() %>%
      mutate(iteration = rep(i, n()))

    out <- bind_rows(out, temp)
  } 
  out <- out  %>%
      gather(metric, value, -c(site, iteration)) %>% # get mean value across iterations for each metric
      group_by(site, metric) %>%
      summarize(value = mean(value)) %>%
      left_join(site_data)
}
```

### Load site data
```{r, include=FALSE}
site_data <- read_csv("./processed_data/site_data.csv") %>%
  dplyr::select(-c(slope.calc, slope.est, elev.min, elev.max, temp.mean)) %>% # drop these variables
  mutate(elev.class = factor(elev.class, levels = c("oben", "mitte", "unten")),
         elev.class2 = factor(elev.class2, levels = c("high", "mid", "low"))) # turn elav.class into an ordered factor
```

### Climate data
```{r, include=FALSE}
gdd <- read_csv("./processed_data/climate.csv") %>%
  mutate(year = factor(year)) %>%
  select(site, date, gdd.cum)
```

### Colony phase breakpoint analysis
```{r, include=FALSE}
castes <- read_csv("./processed_data/network.csv") %>%
  filter(bb.sp != "psit") %>% # We do not want to consider Psithyrus, since these parasitic species do not have a worker caste
  group_by(site, date, year, caste) %>%
  summarize(caste.count = n()) %>%
  spread(caste, caste.count) %>%
  replace(is.na(.), 0) %>%
  mutate(queen = oldqueen + queen,
         reproductive = male + youngqueen,
         q.prop = queen/(queen+worker+reproductive),
         r.prop = reproductive/(queen+worker+reproductive)) %>%
  select(-oldqueen, -youngqueen, -male) %>%
  left_join(gdd, by = c("site", "date")) %>%
  left_join(site_data, by = c("site")) %>%
  arrange(gdd.cum) %>%
  mutate(yday = yday(date))

breaks_2012_gdd <- breakpoints_gdd(castes, 2012, t1 = 0.5, t2 = 0.15)
breaks_2011_gdd <- breakpoints_gdd(castes, 2011, t1 = 0.5, t2 = 0.15)
breaks_2010_gdd <- breakpoints_gdd(castes, 2010, t1 = 0.5, t2 = 0.15)

breaks_2012_yday <- breakpoints_yday(castes, 2012, t1 = 0.75, t2 = 0.15)
breaks_2011_yday <- breakpoints_yday(castes, 2011, t1 = 0.75, t2 = 0.15)
breaks_2010_yday <- breakpoints_yday(castes, 2010, t1 = 0.75, t2 = 0.15)

breaks_gdd_summary <- full_join(breaks_2012_gdd, 
                                breaks_2011_gdd) %>%
  full_join(breaks_2010_gdd)

write_csv(breaks_gdd_summary, "./output/breaks_gdd.csv")

breaks_yday_summary <- full_join(breaks_2012_yday, 
                                 breaks_2011_yday) %>%
  full_join(breaks_2010_yday)

write_csv(breaks_yday_summary, "./output/breaks_yday.csv")

breaks_yday <- bind_rows(breaks_2012_yday, breaks_2011_yday, breaks_2010_yday) %>%
  ungroup() %>%
  select(site, year, break1.yday = break1, break2.yday = break2) %>%
  unique()

breaks_gdd <- bind_rows(breaks_2012_gdd, breaks_2011_gdd, breaks_2010_gdd) %>%
  ungroup() %>%
  select(site, year, break1.gdd = break1, break2.gdd = break2) %>%
  unique()

breaks <- full_join(breaks_yday, breaks_gdd, by = c("site", "year")) %>%
  mutate(year = factor(year))
```

### Plot phase breakpoints
```{r, include=FALSE}
breaks_summary <- full_join(breaks_gdd_summary, breaks_yday_summary) %>%
  select(site, date, year, elev.class2, yday, gdd.cum, break1, break2, caste, prop) %>%
  gather(time.metric, value, -c(site, date, year, elev.class2, break1, break2, caste, prop))

ggplot(breaks_yday_summary, aes(yday, prop, color = caste)) +
  geom_point(alpha = 0.5) +
  binomial_smooth(formula = y ~ splines::ns(x, 2)) +
  geom_vline(aes(xintercept = break1), linetype = "dashed") +
  geom_vline(aes(xintercept = break2), linetype = "dashed") +
  facet_grid(elev.class2 ~ year) +
  xlab("Julian day") +
  ylab("Proportional abundance")

ggplot(breaks_gdd_summary, aes(gdd.cum, prop, color = caste)) +
  geom_point(alpha = 0.5) +
  binomial_smooth(formula = y ~ splines::ns(x, 2)) +
  geom_vline(aes(xintercept = break1), linetype = "dashed") +
  geom_vline(aes(xintercept = break2), linetype = "dashed") +
  facet_grid(elev.class2 ~ year) +
  xlab("GDD") +
  ylab("Proportional abundance")

ggplot(breaks_summary, aes(value, prop, color = caste)) +
  geom_point(alpha = 0.25) +
  binomial_smooth(formula = y ~ splines::ns(x, 2)) +
  facet_grid(elev.class2~ time.metric, scales = "free") +
  xlab("Julian day") +
  ylab("Proportional abundance") 
```

### Load network and survey data
```{r, include=FALSE}
bb_traits <- read_csv("./processed_data/bb_traits.csv") %>%
  select(subgenus, bb.sp, pbl.w, pbl.w.class)

fl_traits <- read.csv("./processed_data/floral_k_type_nom_fix.csv") %>%
  select(plant.sp, k.type) %>%
  mutate(k.type.s = str_extract(k.type, "[-+]?[0-9]*\\.?[0-9]+"),
         k.type.ss = str_extract(k.type, "[-+]?[0-9]*"))

fl_tax <- read_csv("./processed_data/floral_tax.csv")

net <- read_csv("./processed_data/network.csv") %>%
  unite(site.year, site, year, sep = "_", remove = FALSE) %>% # create site.year field that can later be split back up
  unite(site.date, site, date, sep = "_", remove = FALSE) %>%
  mutate(year = factor(year)) %>% # convert year to factor
  full_join(bb_traits, by = c("bb.sp" = "bb.sp")) %>%
  left_join(site_data, by = "site") %>%
  left_join(gdd, by = c("site", "date")) %>%
  left_join(breaks, by = c("site", "year")) %>%
  left_join(fl_traits) %>%
  left_join(fl_tax, by = "plant.genus") %>%
  ungroup() %>%
  mutate(phase = case_when(
    gdd.cum < break1.gdd ~ "founding",
    gdd.cum >= break1.gdd & gdd.cum < break2.gdd ~ "buildup",
    gdd.cum > break2.gdd ~ "reproductive"
  )) %>%
  mutate(phase = factor(phase, levels = c("founding", "buildup", "reproductive"))) %>%
  arrange(site, date)

survey <- read_csv("./processed_data/floral_survey.csv") %>%
  mutate(year = factor(year)) %>% # convert year to factor
  semi_join(net, by = "plant.sp") %>% # Consider only plants visited at a given site-date
  left_join(gdd, by = c("site", "date")) %>%
  left_join(breaks, by = c("site", "year")) %>%
  left_join(site_data, by = "site") %>%
  left_join(fl_traits) %>%
  left_join(fl_tax) %>%
  ungroup() %>%
  mutate(phase = case_when(
    gdd.cum < break1.gdd ~ "founding",
    gdd.cum >= break1.gdd & gdd.cum < break2.gdd ~ "buildup",
    gdd.cum > break2.gdd ~ "reproductive"
  )) %>%
  mutate(phase = factor(phase, levels = c("founding", "buildup", "reproductive"))) %>%
  arrange(site, date)

```
### BB and floral turnover ~ elevation
```{r, include=FALSE}
bb_range <- net %>%
  group_by(site, elev.mean, bb.sp) %>%
  summarize(abund = n()) %>%
  group_by(bb.sp) %>%
  mutate(elev.floor = min(elev.mean),
         elev.ceiling = max(elev.mean),
         elev.range = elev.ceiling - elev.floor,
         elev.med = elev.floor + ((elev.ceiling - elev.floor)/2))

fl_survey_elev <- survey %>%
  filter(flower.cover > 0) %>%
  group_by(site, plant.sp) %>%
  summarize(abund = sum(flower.cover)) %>%
  left_join(site_data) %>%
  group_by(plant.sp) %>%
  mutate(elev.floor = min(elev.mean),
         elev.ceiling = max(elev.mean),
         elev.range = elev.ceiling - elev.floor,
         elev.med = elev.floor + ((elev.ceiling - elev.floor)/2),
         sites = length(unique(site)))

fl_survey_elev_family <- survey %>%
  filter(flower.cover > 0) %>%
  group_by(site, plant.family) %>%
  summarize(abund = sum(flower.cover)) %>%
  left_join(site_data) %>%
  group_by(plant.family) %>%
  mutate(elev.floor = min(elev.mean),
         elev.ceiling = max(elev.mean),
         elev.range = elev.ceiling - elev.floor,
         elev.med = elev.floor + ((elev.ceiling - elev.floor)/2),
         sites = length(unique(site)))

fl_net_elev <- net %>%
  group_by(site, elev.mean, plant.sp) %>%
  summarize(abund = n()) %>%
  group_by(plant.sp) %>%
  mutate(elev.floor = min(elev.mean),
         elev.ceiling = max(elev.mean),
         elev.range = elev.ceiling - elev.floor,
         elev.med = elev.floor + ((elev.ceiling - elev.floor)/2),
         sites = length(unique(site)))

ggplot(filter(fl_survey_elev, sites > 0), 
       aes(reorder(plant.sp, elev.ceiling), elev.mean)) +
  geom_line(size = 1, color = "gray60") +
  geom_point(alpha = 0.5) +
  theme_light(14) +
  theme(axis.text.x = element_blank()) +
  xlab("Plant species") +
  ylab("Elevation")

ggplot(filter(fl_survey_elev_family, sites > 0), 
       aes(reorder(plant.family, elev.med), elev.mean)) +
  geom_line(size = 1, color = "gray60") +
  geom_point(alpha = 0.5) +
  theme_light(10) +
  #theme(axis.text.x = element_blank()) +
  xlab("Plant family") +
  ylab("Elevation") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

ggplot(filter(fl_net_elev, sites > 0), 
       aes(reorder(plant.sp, elev.med), elev.mean)) +
  geom_line(size = 1, color = "gray60") +
  geom_point(alpha = 0.5) +
  theme_light(14) +
  theme(axis.text.x = element_blank()) +
  xlab("Plant species") +
  ylab("Elevation")

ggplot(bb_range, aes(reorder(bb.sp, elev.med), elev.mean)) +
  geom_line(size = 3, color = "gray60") +
  geom_point(aes(size = abund), alpha = 0.5) +
  theme_light(14) +
  xlab("BB species") +
  ylab("Elevation") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

### Plot only the plants that drop out at 1750 m
ggplot(filter(fl_net_elev, sites > 0 & elev.ceiling < 1750), 
       aes(reorder(plant.sp, elev.med), elev.mean)) +
  geom_line(size = 1, color = "gray60") +
  geom_point(alpha = 0.5, aes(size = abund)) +
  theme_light(8) +
  xlab("Plant species") +
  ylab("Elevation") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

ggplot(filter(fl_survey_elev, sites > 0 & elev.ceiling < 1750), 
       aes(reorder(plant.sp, elev.med), elev.mean)) +
  geom_line(size = 1, color = "gray60") +
  geom_point(alpha = 0.5, aes(size = abund)) +
  theme_light(8) +
  xlab("Plant species") +
  ylab("Elevation") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

### FL traits by elevation -- do we see a sharp change at the tree line?
```{r}
survey_trait_elev <- survey %>%
  group_by(site, year, elev.mean,  k.type.s) %>%
  summarize(abund = sum(flower.cover))

survey_trait_elev2 <- survey %>%
  group_by(site, year, elev.mean,  k.type.ss) %>%
  summarize(abund = sum(flower.cover))

net_trait_elev <- net %>%
  group_by(site, year, elev.mean,  k.type.s) %>%
  summarize(abund = n())

net_trait_elev2 <- net %>%
  group_by(site, year, elev.mean,  k.type.ss) %>%
  summarize(abund = n())

ggplot(survey_trait_elev, aes(elev.mean, sqrt(abund))) +
  geom_point() +
  geom_smooth() +
  geom_vline(xintercept = 1750, color = "red") +
  facet_wrap(~k.type.s, scales = "free_y")

ggplot(survey_trait_elev2, aes(elev.mean, sqrt(abund))) +
  geom_point() +
  geom_smooth() +
  geom_vline(xintercept = 1750, color = "red") +
  facet_wrap(~k.type.ss, scales = "free_y")

ggplot(net_trait_elev, aes(elev.mean, sqrt(abund))) +
  geom_point() +
  geom_smooth() +
  geom_vline(xintercept = 1750, color = "red") +
  facet_wrap(~k.type.s, scales = "free_y")

ggplot(net_trait_elev2, aes(elev.mean, sqrt(abund))) +
  geom_point() +
  geom_smooth() +
  geom_vline(xintercept = 1750, color = "red") +
  facet_wrap(~k.type.ss, scales = "free_y")

net_trait_genus_elev <- net %>%
  group_by(site, year, elev.mean, plant.genus, k.type.s) %>%
  summarize(abund = n())

survey_trait_genus_elev <- survey %>%
  group_by(site, year, elev.mean, plant.genus, k.type.s) %>%
  summarize(abund = sum(flower.cover))

ggplot(filter(net_trait_genus_elev, k.type.s == 5.1), 
       aes(elev.mean, sqrt(abund))) +
  geom_point() +
  geom_smooth() +
  facet_wrap(~plant.genus)

ggplot(filter(survey_trait_genus_elev, k.type.s %in% c(5.1)), 
       aes(elev.mean, sqrt(abund))) +
  geom_point() +
  geom_smooth() +
  geom_vline(xintercept = 1750, color = "red") +
  facet_wrap(~plant.genus, ncol = 5) +
  ylab("sqrt(Abundance)") +
  xlab("Elevation")

ggplot(filter(net_trait_genus_elev, k.type.s %in% c(5.1)), 
       aes(elev.mean, sqrt(abund))) +
  geom_point() +
  geom_smooth() +
  geom_vline(xintercept = 1750, color = "red") +
  facet_wrap(~plant.genus, ncol = 5) +
  ylab("sqrt(Visitation frequency)") +
  xlab("Elevation")
```

### Genera and families by elevation
```{r}

survey_genus_elev <- survey %>%
  group_by(site, year, elev.mean, plant.genus) %>%
  summarize(abund = sum(flower.cover)) %>%
  group_by(site, elev.mean, year) %>%
  mutate(div = length(unique(plant.genus)))

survey_family_elev <- survey %>%
  group_by(site, year, elev.mean,  plant.family) %>%
  summarize(abund = sum(flower.cover)) %>%
  group_by(site, elev.mean, year) %>%
  mutate(div = length(unique(plant.family)))

net_genus_elev <- survey %>%
  group_by(site, year, elev.mean,  plant.genus) %>%
  summarize(abund = n()) %>%
  group_by(site, elev.mean, year) %>%
  mutate(div = length(unique(plant.genus)))

net_family_elev <- survey %>%
  group_by(site, year, elev.mean,  plant.family) %>%
  summarize(abund = n()) %>%
  group_by(site, elev.mean, year) %>%
  mutate(div = length(unique(plant.family)))

#### Plot
ggplot(survey_genus_elev, aes(elev.mean, log(abund))) +
  geom_point() +
  stat_smooth() +
  facet_wrap(~plant.genus, scales = "free_y")

ggplot(survey_family_elev, aes(elev.mean, log(abund))) +
  geom_point() +
  stat_smooth() +
  facet_wrap(~plant.family, scales = "free_y")

ggplot(survey_genus_elev, aes(elev.mean, div)) +
  geom_point() +
  stat_smooth()

ggplot(survey_family_elev, aes(elev.mean, div)) +
  geom_point() +
  stat_smooth()

ggplot(net_genus_elev, aes(elev.mean, log(abund))) +
  geom_point() +
  stat_smooth() +
  facet_wrap(~plant.genus, scales = "free_y")

ggplot(net_family_elev, aes(elev.mean, log(abund))) +
  geom_point() +
  stat_smooth() +
  facet_wrap(~plant.family, scales = "free_y")

ggplot(net_genus_elev, aes(elev.mean, log(abund))) +
  geom_point() +
  stat_smooth(se = FALSE) +
  facet_wrap(~plant.genus)

ggplot(net_genus_elev, aes(elev.mean, log(abund))) +
  geom_point() +
  stat_smooth() +
  facet_wrap(~plant.genus, scales = "free_y")
ggplot(net_genus_elev, aes(elev.mean, abund)) +
  geom_point() +
  stat_smooth(se = FALSE) +
  facet_wrap(~plant.genus)

ggplot(net_genus_elev, aes(elev.mean, div)) +
  geom_point() +
  stat_smooth()
```

### BB abundance, diversity, and per-capita resource availability
```{r eval=FALSE, echo=FALSE, include=FALSE}
### Per-transect Bombus density
bb_abund <- net %>%
  group_by(site, date, dayofyear, gdd.cum, year, elev.class2, 
           elev.mean, break1.gdd, break2.gdd) %>%
  summarize(bb.abund = n())

### Per-transect Bombus richness
bb_rich <- net %>%
  group_by(site, date, dayofyear, gdd.cum, year, elev.class2,
           elev.mean, break1.gdd, break2.gdd) %>%
  summarize(bb.rich = length(unique(bb.sp)))

### Per-transect floral abund
fl_abund <- survey %>%
  group_by(site, date, dayofyear, gdd.cum, year, elev.class2, 
           elev.mean, break1.gdd, break2.gdd) %>%
  summarize(fl.abund = sum(flower.cover))

### Per-transect floral richness
fl_rich <- survey %>%
  filter(flower.cover > 0) %>%
  group_by(site, date, dayofyear, gdd.cum, year, elev.class2, 
           elev.mean, break1.gdd, break2.gdd) %>%
  summarize(fl.rich = n())

### Join BB density and floral cover
abund_div <- bb_abund %>%
  full_join(bb_rich) %>%
  full_join(fl_abund) %>% # Note that there is not perfect alignment in site-dates between the network observations and the floral surveying, so this join will cause some NAs that will need to be removed, causing some transects to drop out of the dataset
  full_join(fl_rich) %>%
  #na.omit() %>%
  ungroup %>%
  mutate(fl.per.bb = fl.abund/bb.abund,
         site = factor(site),
         yday = dayofyear,
         week = week(date),
         month = month(date)) %>%
  ungroup() %>%
  mutate(phase = case_when(
    gdd.cum < break1.gdd ~ "founding",
    gdd.cum >= break1.gdd & gdd.cum < break2.gdd ~ "buildup",
    gdd.cum > break2.gdd ~ "reproductive"
  )) %>%
  mutate(phase = factor(phase, levels = c("founding", "buildup", "reproductive")),
         total.rich = bb.rich + fl.rich,
         web.asymmetry = (bb.rich - fl.rich) / (bb.rich + fl.rich))

abund_div_mean <- abund_div %>%
  select(-dayofyear, -week, -month, -break1.gdd, -break2.gdd) %>%
  gather(metric, value, -c(site, date, gdd.cum, year, elev.class2, elev.mean, phase)) %>%
  group_by(site, year, elev.class2, elev.mean, metric) %>%
  summarize(value = mean(value, na.rm = TRUE))
```

### Plotting abundance and diversity patterns
```{r}
### BB abundance ~ floral abundance
ggplot(abund_div, aes(log(fl.abund), bb.abund, color = year)) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = "lm", se = TRUE) +
  facet_grid(elev.class2 ~ phase, scales = "free") +
  ylab("BB abundance") +
  xlab("log(floral cover)") +
  theme_bw()

ggplot(abund_div, aes(log(fl.abund), bb.abund, color = year)) +
  geom_point(alpha = 0.5) +
  geom_smooth() +
  ylab("BB abundance") +
  xlab("log(floral cover)") +
  theme_bw()

### BB richness ~ floral abundance
ggplot(abund_div, aes(log(fl.abund), bb.rich, color = year)) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = "lm", se = TRUE) +
  facet_grid(elev.class2 ~ phase, scales = "free") +
  ylab("BB richness") +
  xlab("log(floral cover)") +
  theme_bw()

### BB abundance ~ floral richness
ggplot(abund_div, aes(fl.rich, bb.abund, color = year)) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = "lm", se = TRUE) +
  facet_grid(elev.class2 ~ phase, scales = "free") +
  ylab("BB abundance") +
  xlab("Floral richness")

### BB richness ~ floral richness
ggplot(abund_div, aes(fl.rich, bb.rich, color = year)) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = "lm", se = TRUE) +
  facet_grid(elev.class2 ~ phase, scales = "free") +
  ylab("BB richness") +
  xlab("Floral richness")

### Floral richness ~ floral abundance
ggplot(abund_div, aes(log(fl.abund), fl.rich, color = year)) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = "lm", se = TRUE) +
  facet_grid(elev.class2 ~ phase, scales = "free") +
  ylab("Floral richness") +
  xlab("log(floral cover)")

### Per-capita floral resource abundance ~ elevation
ggplot(abund_div, aes(elev.mean, log(fl.per.bb))) +
  geom_point(alpha = 0.5) +
  #stat_summary(aes(y = log(fl.per.bb), group=1), fun.y=mean, colour="black", geom="point",group=1) +
  geom_smooth() +
  facet_grid(year~phase, scales = "free") +
  ylab("log(fl-per-bb)") +
  xlab("Elevation") +
  theme_light()

ggplot(abund_div, aes(elev.mean, log(fl.per.bb))) +
  geom_point(alpha = 0.5) +
  #stat_summary(aes(y = log(fl.per.bb), group=1), fun.y=mean, colour="black", geom="point",group=1) +
  geom_smooth() +
  facet_grid(.~year, scales = "free") +
  ylab("log(fl-per-bb)") +
  xlab("Elevation") +
  theme_light()

ggplot(abund_div, aes(elev.mean, log(fl.per.bb))) +
  geom_point(alpha = 0.5, aes(color = phase)) +
  #stat_summary(aes(y = log(fl.per.bb), group=1), fun.y=mean, colour="black", geom="point",group=1) +
  geom_smooth(color = "black", linetype = "dashed", size = 0.5) +
  facet_grid(year~., scales = "free") +
  ylab("log(fl-per-bb)") +
  xlab("Elevation") +
  theme_light()

ggplot(abund_div, aes(elev.mean, log(fl.per.bb))) +
  geom_point(alpha = 0.5, aes(color = phase)) +
  #stat_summary(aes(y = log(fl.per.bb), group=1), fun.y=mean, colour="black", geom="point",group=1) +
  geom_smooth(color = "black", linetype = "dashed", size = 0.5) +
  ylab("log(fl-per-bb)") +
  xlab("Elevation") +
  theme_light()

ggplot(abund_div, aes(elev.mean, fl.per.bb)) +
  geom_point(alpha = 0.5, aes(color = phase)) +
  #stat_summary(aes(y = log(fl.per.bb), group=1), fun.y=mean, colour="black", geom="point",group=1) +
  geom_smooth(color = "black", linetype = "dashed", size = 0.5) +
  ylab("log(fl-per-bb)") +
  xlab("Elevation") +
  theme_light()

ggplot(filter(abund_div_mean, metric == "fl.per.bb"), aes(elev.mean, log(value))) +
  geom_point(alpha = 1) +
  #geom_smooth(color = "black", linetype = "dashed", size = 0.5) +
  geom_smooth(method = "lm", formula = y ~ poly(x, 2)) +
  ylab("log(fl-per-bb)") +
  xlab("Elevation") +
  facet_wrap(~year) +
  theme_light()

lm1 <- lm(value ~ elev.mean * year, data = filter(abund_div_mean, metric == "fl.per.bb"))
summary(lm1)

ggplot(filter(abund_div_mean, metric == "fl.per.bb"), aes(elev.mean, value)) +
  geom_point(alpha = 1) +
  #geom_smooth(color = "black", linetype = "dashed", size = 0.5) +
  geom_smooth() +
  ylab("fl-per-bb") +
  xlab("Elevation") +
  theme_light()

ggplot(filter(abund_div_mean, metric == "fl.per.bb"), aes(elev.mean, log(value), color = year)) +
  geom_point(alpha = 1) +
  #geom_smooth(color = "black", linetype = "dashed", size = 0.5) +
  geom_smooth() +
  ylab("FL per BB") +
  xlab("Elevation") +
  theme_light()

ggplot(filter(abund_div_mean, metric == "fl.per.bb"), aes(elev.mean, log(value))) +
  geom_point(alpha = 1, aes(color = year)) +
  geom_smooth(color = "black", linetype = "dashed", size = 0.5) +
  #geom_smooth() +
  ylab("FL per BB") +
  xlab("Elevation") +
  theme_light(16)

### Per-capita floral resource abundance ~ GDD
ggplot(abund_div, aes(gdd.cum, log(fl.per.bb))) +
  geom_point(alpha = 0.5, aes(color = phase)) +
  geom_smooth(color = "black", linetype = "dashed", size = 0.5) +
  facet_grid(elev.class2 ~ year, scales = "free") +
  ylab("log(fl-per-bb)") +
  xlab("GDD") +
  theme_light()

ggplot(abund_div, aes(gdd.cum, log(fl.per.bb))) +
  geom_point(alpha = 0.5, aes(color = phase)) +
  geom_smooth(color = "black", linetype = "dashed", size = 0.5) +
  ylab("log(fl-per-bb)") +
  xlab("GDD") +
  theme_light()

### Per-capita floral resource abundance ~ Julian day
ggplot(abund_div, aes(yday, log(fl.per.bb))) +
  geom_point(alpha = 0.5, aes(color = phase)) +
  geom_smooth(color = "black", linetype = "dashed", size = 0.5) +
  facet_grid(elev.class2 ~ year, scales = "free") +
  ylab("log(fl-per-bb)") +
  xlab("Julian day") +
  theme_light()

### BB richness ~ elevation
ggplot(abund_div, aes(elev.mean, bb.rich)) +
  geom_point() +
  geom_smooth() +
  facet_grid(.~year)

### BB richness ~ GDD
ggplot(abund_div, aes(gdd.cum, bb.rich)) +
  geom_point() +
  geom_smooth() +
  facet_grid(elev.class2 ~ year)

### BB richness ~ Julian day
ggplot(abund_div, aes(yday, bb.rich)) +
  geom_point() +
  geom_smooth() +
  facet_grid(elev.class2 ~ year)

### BB abundance ~ elevation
ggplot(abund_div, aes(elev.mean, bb.abund)) +
  geom_point() +
  geom_smooth() +
  facet_grid(year ~ phase, scales = "free")

ggplot(abund_div, aes(elev.mean, bb.abund)) +
  geom_point() +
  geom_smooth() +
  facet_grid(.~year, scales = "free")

ggplot(abund_div, aes(elev.mean, bb.abund, color = year)) +
  geom_point(alpha = 0.5) +
  geom_smooth() +
  xlab("Elevation") +
  ylab("BB abundance")  +
  theme_light(16)

ggplot(abund_div, aes(elev.mean, sqrt(bb.abund), color = year)) +
  geom_point(alpha = 0.5) +
  geom_smooth() +
  xlab("Elevation") +
  ylab("sqrt(BB abundance)")  +
  theme_light(16)

### BB abundance ~ GDD
ggplot(abund_div, aes(gdd.cum, bb.abund)) +
  geom_point() +
  geom_smooth() +
  facet_grid(elev.class2 ~ year)

### BB abundance ~ Julian day
ggplot(abund_div, aes(yday, bb.abund)) +
  geom_point() +
  geom_smooth() +
  facet_grid(elev.class2 ~ year)

### Floral richness ~ elevation
ggplot(abund_div, aes(elev.mean, fl.rich)) +
  geom_point() +
  geom_smooth() +
  facet_grid(year ~ phase)

ggplot(abund_div, aes(elev.mean, fl.rich)) +
  geom_point(alpha = 0.5, aes(color = year)) +
  geom_smooth(color = "black", linetype = "dashed", size = 0.75) +
  theme_light(16) +
  xlab("Elevation") +
  ylab("Floral richness")

### Floral richness ~ GDD
ggplot(abund_div, aes(gdd.cum, fl.rich)) +
  geom_point() +
  geom_smooth() +
  facet_grid(elev.class2 ~ year)

### Floral richness ~ Julian day
ggplot(abund_div, aes(yday, fl.rich)) +
  geom_point() +
  geom_smooth() +
  facet_grid(elev.class2 ~ year)

### Floral abundance ~ elevation
ggplot(abund_div, aes(elev.mean, log(fl.abund))) +
  geom_point(alpha = 0.5) +
  geom_smooth() +
  facet_grid(year ~ phase, scales = "free") +
  theme_light()

ggplot(abund_div, aes(elev.mean, log(fl.abund), color = year)) +
  geom_point(alpha = 0.5) +
  geom_smooth() +
  theme_light()

ggplot(abund_div, aes(elev.mean, fl.abund)) +
  geom_point(alpha = 0.5, aes(color = year)) +
  geom_smooth(color = "black", linetype = "dashed", size = 0.75) +
  theme_light(16) +
  xlab("Elevation") +
  ylab("Floral abundance")

ggplot(abund_div, aes(elev.mean, log(fl.abund))) +
  geom_point(alpha = 0.5, aes(color = year)) +
  geom_smooth(color = "black", linetype = "dashed", size = 0.75) +
  theme_light(16) +
  xlab("Elevation") +
  ylab("log(Floral abundance)")

ggplot(abund_div, aes(elev.mean, fl.abund, color = year)) +
  geom_point(alpha = 0.5) +
  theme_light(16) +
  xlab("Elevation") +
  ylab("Floral abundance")

ggplot(abund_div, aes(elev.mean, log(fl.abund), color = year)) +
  geom_point(alpha = 0.5) +
  geom_smooth() +
  theme_light(16) +
  xlab("Elevation") +
  ylab("log(Floral abundance)")

ggplot(abund_div, aes(elev.mean, sqrt(bb.abund), color = year)) +
  geom_point(alpha = 0.5) +
  geom_smooth() +
  xlab("Elevation") +
  ylab("sqrt(BB abundance)")  +
  theme_light(16)

### Floral abundance ~ GDD
ggplot(abund_div, aes(gdd.cum, log(fl.abund))) +
  geom_point(aes(color = phase), alpha =  0.5) +
  geom_smooth(color = "black", linetype = "dashed", size = 0.5) +
  facet_grid(elev.class2 ~ year, scales = "free") +
  theme_light()

### Floral abundance ~ Julian day
ggplot(abund_div, aes(yday, log(fl.abund))) +
  geom_point() +
  geom_smooth() +
  facet_grid(elev.class2 ~ year)

### Total (BB = FL) richness ~ elevation
ggplot(abund_div, aes(elev.mean, total.rich)) +
  geom_point() +
  geom_smooth() +
  facet_grid(year ~ phase, scales = "free")

### Web asymmetry ~ elevation
ggplot(abund_div, aes(elev.mean, web.asymmetry)) +
  geom_point(alpha = 0.5) +
  geom_smooth() +
  facet_grid(year ~ phase, scales = "free")

ggplot(filter(abund_div_mean, metric == "web.asymmetry"), aes(elev.mean, value)) +
  geom_point() +
  geom_smooth() +
  facet_grid(year ~ ., scales = "free")

### Web asymmetry ~ gdd
ggplot(abund_div, aes(gdd.cum, web.asymmetry)) +
  geom_point(alpha = 0.5) +
  geom_smooth() +
  facet_grid(year ~.)

### Web asymmetry ~ yday
ggplot(abund_div, aes(yday, web.asymmetry)) +
  geom_point(alpha = 0.5) +
  geom_smooth() +
  facet_grid(year ~.)

### Web asymmetry ~ floral abundance
ggplot(abund_div, aes(log(fl.abund), web.asymmetry)) +
  geom_point(alpha = 0.5) +
  geom_smooth() +
  facet_grid(year ~ phase, scales = "free")

### Web asymmetry ~ BB abundance
ggplot(abund_div, aes(bb.abund, web.asymmetry)) +
  geom_point(alpha = 0.5) +
  geom_smooth() +
  facet_grid(year ~ phase, scales = "free")

### Web asymmetry ~ Per-capita floral resource abundance
ggplot(abund_div, aes(log(fl.per.bb), web.asymmetry)) +
  geom_point(alpha = 0.5) +
  geom_smooth() +
  facet_grid(year ~ phase, scales = "free")

ggplot(abund_div, aes(log(fl.per.bb), web.asymmetry)) +
  geom_point(alpha = 0.5) +
  geom_smooth() +
  facet_grid(year ~ .)
```

### GAM surface plots of abundance and diversity metrics
#### HGAM elevation x time x year (I would really like someone -- maybe my future self -- to check that I'm doing this right)
##### Floral abundance
```{r eval=FALSE, include=FALSE}
fl_abund_yday <- gam(fl.abund ~ year +
                    te(yday, elev.mean, by=year, bs= c("cc", "tp"), k=c(10, 10), m=1),
                    data = filter(abund_div),
                    family = "tw",
                    method = "REML")

fl_abund_yday.vis <- getViz(fl_abund_yday)
check.gamViz(fl_abund_yday.vis)
summary(fl_abund_yday.vis)

plot(fl_abund_yday.vis, allTerms = TRUE) +
  ggtitle("FL abundance (yday)") +
  xlab("Day of year") +
  ylab("Elevation")

fl_abund_gdd <- gam(fl.abund ~ year +
                    te(gdd.cum, elev.mean, by=year, bs= c("cc", "tp"), k=c(10, 10), m=1),
                    data = filter(abund_div),
                    family = "tw",
                    method = "REML")

fl_abund_gdd.vis <- getViz(fl_abund_gdd)
check.gamViz(fl_abund_gdd.vis)
summary(fl_abund_gdd.vis)

plot(fl_abund_gdd.vis, allTerms = TRUE) +
  ggtitle("FL abundance (gdd)") +
  xlab("GDD") +
  ylab("Elevation")
```

##### Floral richness -- since year does not seem to change the overall pattern (though perhaps the magnitude), I could use a simpler model, perhaps with year as a random intercept effect; some explanat
```{r eval=FALSE, include=FALSE}
### By year
fl_rich_yday <- gam(fl.rich ~ year +
                    te(yday, elev.mean, by=year, bs= c("cc", "tp"), k=c(10, 10), m=1),
                    data = filter(abund_div),
                    method = "REML")

fl_rich_yday.vis <- getViz(fl_rich_yday)
check.gamViz(fl_rich_yday.vis)
summary(fl_rich_yday.vis)

plot(fl_rich_yday.vis, allTerms = TRUE) +
  ggtitle("FL abundance (gdd)") +
  xlab("Day of year") +
  ylab("Elevation")

### With year as parametric term
fl_rich_yday2 <- gam(fl.rich ~ year +
                    te(yday, elev.mean, bs= "cc", k= 10, m=1),
                    data = filter(abund_div),
                    method = "REML")

fl_rich_yday2.vis <- getViz(fl_rich_yday2)
check.gamViz(fl_rich_yday2.vis)
summary(fl_rich_yday2.vis)

plot(sm(fl_rich_yday2.vis, 1)) + 
  l_fitRaster() + 
  l_fitContour() + 
  l_points(size = 1) +
  ggtitle("FL richness (yday)") +
  xlab("Day of year") +
  ylab("Elevation")

plot(pterm(fl_rich_yday2.vis, 1))

### By year
fl_rich_gdd <- gam(fl.rich ~ year +
                    te(gdd.cum, elev.mean, by=year, bs= c("cc", "tp"), k=c(10, 10), m=1),
                    data = filter(abund_div),
                    method = "REML")

fl_rich_gdd.vis <- getViz(fl_rich_gdd)
check.gamViz(fl_rich_gdd.vis)
summary(fl_rich_gdd.vis)

plot(fl_rich_gdd.vis, allTerms = TRUE) +
  ggtitle("FL richness (gdd)") +
  xlab("GDD") +
  ylab("Elevation")

### With year as parametric term
fl_rich_gdd2 <- gam(fl.rich ~ year +
                    te(gdd.cum, elev.mean, bs= "cc", k= 10, m=1),
                    data = filter(abund_div),
                    method = "REML")

fl_rich_gdd2.vis <- getViz(fl_rich_gdd2)
check.gamViz(fl_rich_gdd2.vis)
summary(fl_rich_gdd2.vis)

plot(sm(fl_rich_gdd2.vis, 1)) + 
  l_fitRaster() + 
  l_fitContour() + 
  l_points(size = 1) +
  ggtitle("FL richness (gdd)") +
  xlab("Day of year") +
  ylab("Elevation")

plot(pterm(fl_rich_gdd2.vis, 1))
```

##### BB abundance
```{r eval=FALSE, include=FALSE}
bb_abund_yday <- gam(bb.abund ~ year +
                    te(yday, elev.mean, by=year, bs= c("cc", "tp"), k=c(10, 10), m=1),
                    data = filter(abund_div),
                    method = "REML")

bb_abund_yday.vis <- getViz(bb_abund_yday)
check.gamViz(bb_abund_yday.vis)
summary(bb_abund_yday.vis)

plot(bb_abund_yday.vis, allTerms = TRUE) +
  ggtitle("BB abundance (yday)") +
  xlab("Day of year") +
  ylab("Elevation")

bb_abund_gdd <- gam(bb.abund ~ year +
                    te(gdd.cum, elev.mean, by=year, bs= c("cc", "tp"), k=c(10, 10), m=1),
                    data = filter(abund_div),
                    method = "REML")

bb_abund_gdd.vis <- getViz(bb_abund_gdd)
check.gamViz(bb_abund_gdd.vis)
summary(bb_abund_gdd.vis)

plot(bb_abund_gdd.vis, allTerms = TRUE) +
  ggtitle("BB abundance (gdd)") +
  xlab("GDD") +
  ylab("Elevation")
```

##### BB richness
```{r eval=FALSE, include=FALSE}
bb_rich_yday <- gam(bb.rich ~ year +
                    te(yday, elev.mean, by=year, bs= c("cc", "tp"), k=c(10, 10), m=1),
                    data = filter(abund_div),
                    method = "REML")

bb_rich_yday.vis <- getViz(bb_rich_yday)
check.gamViz(bb_rich_yday.vis)
summary(bb_rich_yday.vis)
plot(sm(bb_rich_yday.vis, 1)) +
  l_fitRaster() +
  l_fitContour() +
  l_points() +
  ggtitle("bb richness 2010 (yday)") +
  xlab("Day of year") +
  ylab("Elevation")
plot(sm(bb_rich_yday.vis, 2)) +
  l_fitRaster() +
  l_fitContour() +
  l_points() +
  ggtitle("bb richness 2011 (yday)") +
  xlab("Day of year") +
  ylab("Elevation")
plot(sm(bb_rich_yday.vis, 3)) +
  l_fitRaster() +
  l_fitContour() +
  l_points() +
  ggtitle("bb richness 2012 (yday)") +
  xlab("Day of year") +
  ylab("Elevation")

bb_rich_gdd <- gam(bb.rich ~ year +
                    te(gdd.cum, elev.mean, by=year, bs= c("cc", "tp"), k=c(10, 10), m=1),
                    data = filter(abund_div),
                    method = "REML")

bb_rich_gdd.vis <- getViz(bb_rich_gdd)
check.gamViz(bb_rich_gdd.vis)
summary(bb_rich_gdd.vis)

plot(bb_rich_gdd.vis, allTerms = TRUE) +
  ggtitle("BB richness (gdd)") +
  xlab("Day of year") +
  ylab("Elevation")
```

### Rarefied network analysis (not broken down by phase)
```{r, include = FALSE}
### Determine the minimum transect count per site*year to set rarefaction sampling
transects <- net %>%
  select(site, year, date) %>%
  unique()

transects_per_siteyear <- transects %>%
  group_by(site, year) %>%
  summarize(transects = n())

min_transects_per_year <- transects_per_siteyear %>%
  group_by(year) %>%
  summarize(min_transects = min(transects))

### Create yearly subsets to feed to rare_net
net_2012 <- filter(net, year == "2012")
net_2011 <- filter(net, year == "2011")
net_2010 <- filter(net, year == "2010")

### Demonstrate that the distribution of interaction frequencies does not differ across elevation classes; in other words, no systematic bias is introduced to the network analysis due to uneven frequency of species with few observations
net_freq <- net %>%
  group_by(site, dayofyear, year, bb.sp, elev.mean, elev.class2) %>%
  summarize(freq = n())

ggplot(net_freq, aes(elev.class2, log(freq))) +
  geom_boxplot() +
  facet_wrap(~year)

### Species level
spmet_2012 <- rare_net_sp(net_2012, 9, 20) %>%
  mutate(year = factor(rep("2012", n())))
spmet_2011 <- rare_net_sp(net_2011, 4, 20) %>%
  mutate(year = factor(rep("2011", n())))  
spmet_2010 <- rare_net_sp(net_2010, 3, 20) %>%
  mutate(year = factor(rep("2010", n())))

spmet <- bind_rows(spmet_2012, spmet_2011, spmet_2010)

spmet_mean <- spmet %>%
  group_by(site, year, elev.mean, metric) %>%
  summarize(value = mean(value))

### Network- and group-level
netmet_rare_2012 <- rare_net_nt(net_2012, n = 9, iter = 20) %>%
  mutate(year = factor(rep("2012", n())))
netmet_rare_2011 <- rare_net_nt(net_2011, n = 4, iter = 20)  %>%
  mutate(year = factor(rep("2011", n())))
netmet_rare_2010 <- rare_net_nt(net_2010, n = 3, iter = 20) %>%
  mutate(year = factor(rep("2010", n())))

netmet <- bind_rows(netmet_rare_2012, 
                    netmet_rare_2011, 
                    netmet_rare_2010) %>%
  mutate(metric = factor(metric, levels = c("H2",
                                            "weighted.NODF",
                                            "niche.overlap.HL", 
                                            "C.score.HL",
                                            "generality.HL",
                                            "web.asymmetry",
                                            "weighted.connectance")))

### Plot
#### Per species values
ggplot(filter(spmet_2012,
              sp %in% c("bss", "hort", "pasc", "prat", "soro", "wurf")), 
       aes(elev.mean, value)) +
  geom_point() +
  #geom_smooth(method = "lm", formula = y ~ poly(x, 3)) +
  geom_smooth() +
  facet_grid(metric~sp, scales = "free")

ggplot(filter(spmet_2011,
              sp %in% c("bss", "hort", "pasc", "prat", "soro", "wurf")), 
       aes(elev.mean, value)) +
  geom_point() +
  #geom_smooth(method = "lm", formula = y ~ poly(x, 3)) +
  geom_smooth() +
  facet_grid(metric~sp, scales = "free")

ggplot(filter(spmet_2010,
              sp %in% c("bss", "hort", "pasc", "prat", "soro", "wurf")), 
       aes(elev.mean, value)) +
  geom_point() +
  #geom_smooth(method = "lm", formula = y ~ poly(x, 3)) +
  geom_smooth() +
  facet_grid(metric~sp, scales = "free")

#### Community mean d
ggplot(filter(spmet_mean, metric == "d"), aes(elev.mean, value)) +
  geom_point(aes(color = year)) +
  geom_smooth(method = "gam", formula = y ~ s(x, k = 10), color = "black", linetype = "dashed", size = 0.75) +
  xlab("Elevation") +
  ylab("Community mean d'") +
  theme_light(16)

gam_d <- gam(value ~ year + s(elev.mean, k = 10), 
             data = filter(spmet, metric == "d"),
             method = "REML")

gam_d.vis <- getViz(gam_d)
check.gamViz(gam_d.vis)
summary(gam_d.vis)

plot(gam_d.vis, allTerms = TRUE) +
  xlab("Elevation")

#### Network level
ggplot(netmet, aes(elev.mean, value)) +
  geom_point() +
  #geom_smooth(method = "lm", formula = y ~ poly(x, 3)) +
  geom_smooth() +
  facet_grid(metric~year, scales = "free")

##### H2
ggplot(filter(netmet, metric == "H2"), aes(elev.mean, value)) +
  geom_point(aes(color = year)) +
  geom_smooth(method = "gam", formula = y ~ s(x), color = "black", linetype = "dashed", size = 0.75) +
  xlab("Elevation") +
  ylab("H2'") +
  theme_light(16)

gam_H2 <- gam(value ~ year + s(elev.mean), 
              data = filter(netmet, metric == "H2"),
              method = "REML")

gam_H2.vis <- getViz(gam_H2)
check.gamViz(gam_H2.vis)
summary(gam_H2.vis)

plot(gam_H2.vis, allTerms = TRUE) +
  xlab("Elevation")

##### niche overlap
ggplot(filter(netmet, metric == "niche.overlap.HL"), aes(elev.mean, value)) +
  geom_point(aes(color = year)) +
  #geom_smooth(method = "lm", formula = y ~ poly(x, 3)) +
  geom_smooth(method = "gam", formula = y ~ s(x), color = "black", linetype = "dashed", size = 0.75) +
  #geom_smooth() +
  xlab("Elevation") +
  ylab("Niche overlap") +
  theme_light(16)

gam_niche.overlap <- gam(value ~ year + s(elev.mean), 
                         data = filter(netmet, metric == "niche.overlap.HL"),
                         method = "REML")

gam_niche.overlap.vis <- getViz(gam_niche.overlap)
check.gamViz(gam_niche.overlap.vis)
summary(gam_niche.overlap.vis)

plot(gam_niche.overlap.vis, allTerms = TRUE) +
  xlab("Elevation")

##### C.score
ggplot(filter(netmet, metric == "C.score.HL"), aes(elev.mean, value)) +
  geom_point(aes(color = year)) +
  #geom_smooth(method = "lm", formula = y ~ poly(x, 3)) +
  geom_smooth(method = "gam", formula = y ~ s(x), color = "black", linetype = "dashed", size = 0.75) +
  #geom_smooth() +
  xlab("Elevation") +
  ylab("C score") +
  theme_light(16)

gam_cscore <- gam(value ~ year + s(elev.mean), 
                  data = filter(netmet, metric == "C.score.HL"),
                  method = "REML")

gam_cscore.vis <- getViz(gam_cscore)
check.gamViz(gam_cscore.vis)
summary(gam_cscore.vis)

plot(gam_cscore.vis, allTerms = TRUE) +
  xlab("Elevation")

##### Connectance
ggplot(filter(netmet, metric == "weighted.connectance"), aes(elev.mean, value)) +
  geom_point(aes(color = year)) +
  #geom_smooth(method = "lm", formula = y ~ poly(x, 3)) +
  geom_smooth(method = "gam", formula = y ~ s(x), color = "black", linetype = "dashed", size = 0.75) +
  #geom_smooth() +
  xlab("Elevation") +
  ylab("Connectance (weighted)") +
  theme_light(16)

gam_connectance <- gam(value ~ year + s(elev.mean), 
                  data = filter(netmet, metric == "weighted.connectance"),
                  method = "REML")

gam_connectance.vis <- getViz(gam_connectance)
check.gamViz(gam_connectance.vis)
summary(gam_connectance.vis)

plot(gam_connectance.vis, allTerms = TRUE) +
  xlab("Elevation")

##### NODF
ggplot(filter(netmet, metric == "weighted.NODF"), aes(elev.mean, value)) +
  geom_point(aes(color = year)) +
  #geom_smooth(method = "lm", formula = y ~ poly(x, 3)) +
  geom_smooth(method = "gam", formula = y ~ s(x), color = "black", linetype = "dashed", size = 0.75) +
  #geom_smooth() +
  xlab("Elevation") +
  ylab("NODF (weighted)") +
  theme_light(16)

gam_NODF <- gam(value ~ year + s(elev.mean), 
                data = filter(netmet, metric == "weighted.NODF"),
                method = "REML")

gam_NODF.vis <- getViz(gam_NODF)
check.gamViz(gam_NODF.vis)
summary(gam_NODF.vis)

plot(gam_NODF.vis, allTerms = TRUE) +
  xlab("Elevation")

##### Web asymmetry
ggplot(filter(netmet, metric == "web.asymmetry"), aes(elev.mean, value)) +
  geom_point(aes(color = year)) +
  #geom_smooth(method = "lm", formula = y ~ poly(x, 3)) +
  geom_smooth(method = "gam", formula = y ~ s(x), color = "black", linetype = "dashed", size = 0.75) +
  #geom_smooth() +
  xlab("Elevation") +
  ylab("Web asymmetry") +
  theme_light(16)

gam_webass <- gam(value ~ year + s(elev.mean), 
                data = filter(netmet, metric == "web.asymmetry"),
                method = "REML")

gam_webass.vis <- getViz(gam_webass)
check.gamViz(gam_webass.vis)
summary(gam_webass.vis)

plot(gam_webass.vis, allTerms = TRUE) +
  xlab("Elevation")
```

### Rarefied network analysis by k.type instead of by plant sp
```{r, include = FALSE}
### Create yearly subsets to feed to rare_net
net_2012 <- filter(net, year == "2012")
net_2011 <- filter(net, year == "2011")
net_2010 <- filter(net, year == "2010")

### Species level
spmet_2012 <- rare_net_sp_k.type(net_2012, 9, 20) %>%
  mutate(year = factor(rep("2012", n())))
spmet_2011 <- rare_net_sp(net_2011, 4, 20) %>%
  mutate(year = factor(rep("2011", n())))  
spmet_2010 <- rare_net_sp(net_2010, 3, 20) %>%
  mutate(year = factor(rep("2010", n())))

spmet <- bind_rows(spmet_2012, spmet_2011, spmet_2010)

spmet_mean <- spmet %>%
  group_by(site, year, elev.mean, metric) %>%
  summarize(value = mean(value))

### Network- and group-level
netmet_rare_2012 <- rare_net_nt_k.type(net_2012, n = 9, iter = 20) %>%
  mutate(year = factor(rep("2012", n())))
netmet_rare_2011 <- rare_net_nt_k.type(net_2011, n = 4, iter = 20)  %>%
  mutate(year = factor(rep("2011", n())))
netmet_rare_2010 <- rare_net_nt_k.type(net_2010, n = 3, iter = 20) %>%
  mutate(year = factor(rep("2010", n())))

netmet <- bind_rows(netmet_rare_2012, 
                    netmet_rare_2011, 
                    netmet_rare_2010) %>%
  mutate(metric = factor(metric, levels = c("H2",
                                            "weighted.NODF",
                                            "niche.overlap.HL", 
                                            "C.score.HL",
                                            "generality.HL",
                                            "web.asymmetry",
                                            "weighted.connectance")))

### Plot
#### Per species values
ggplot(filter(spmet_2012,
              sp %in% c("bss", "hort", "pasc", "prat", "soro", "wurf")), 
       aes(elev.mean, value)) +
  geom_point() +
  #geom_smooth(method = "lm", formula = y ~ poly(x, 3)) +
  geom_smooth() +
  facet_grid(metric~sp, scales = "free")

ggplot(filter(spmet_2011,
              sp %in% c("bss", "hort", "pasc", "prat", "soro", "wurf")), 
       aes(elev.mean, value)) +
  geom_point() +
  #geom_smooth(method = "lm", formula = y ~ poly(x, 3)) +
  geom_smooth() +
  facet_grid(metric~sp, scales = "free")

ggplot(filter(spmet_2010,
              sp %in% c("bss", "hort", "pasc", "prat", "soro", "wurf")), 
       aes(elev.mean, value)) +
  geom_point() +
  #geom_smooth(method = "lm", formula = y ~ poly(x, 3)) +
  geom_smooth() +
  facet_grid(metric~sp, scales = "free")

#### Community mean d
ggplot(filter(spmet_mean, metric == "d"), aes(elev.mean, value)) +
  geom_point(aes(color = year)) +
  geom_smooth(method = "gam", formula = y ~ s(x, k = 10), color = "black", linetype = "dashed", size = 0.75) +
  xlab("Elevation") +
  ylab("Community mean d'") +
  theme_light(16)

gam_d <- gam(value ~ year + s(elev.mean, k = 10), 
             data = filter(spmet, metric == "d"),
             method = "REML")

gam_d.vis <- getViz(gam_d)
check.gamViz(gam_d.vis)
summary(gam_d.vis)

plot(gam_d.vis, allTerms = TRUE) +
  xlab("Elevation")

#### Network level
ggplot(netmet, aes(elev.mean, value)) +
  geom_point() +
  #geom_smooth(method = "lm", formula = y ~ poly(x, 3)) +
  geom_smooth() +
  facet_grid(metric~year, scales = "free")

##### H2
ggplot(filter(netmet, metric == "H2"), aes(elev.mean, value)) +
  geom_point(aes(color = year)) +
  geom_smooth(method = "gam", formula = y ~ s(x), color = "black", linetype = "dashed", size = 0.75) +
  xlab("Elevation") +
  ylab("H2'") +
  theme_light(16)

gam_H2 <- gam(value ~ year + s(elev.mean), 
              data = filter(netmet, metric == "H2"),
              method = "REML")

gam_H2.vis <- getViz(gam_H2)
check.gamViz(gam_H2.vis)
summary(gam_H2.vis)

plot(gam_H2.vis, allTerms = TRUE) +
  xlab("Elevation")

##### niche overlap
ggplot(filter(netmet, metric == "niche.overlap.HL"), aes(elev.mean, value)) +
  geom_point(aes(color = year)) +
  #geom_smooth(method = "lm", formula = y ~ poly(x, 3)) +
  geom_smooth(method = "gam", formula = y ~ s(x), color = "black", linetype = "dashed", size = 0.75) +
  #geom_smooth() +
  xlab("Elevation") +
  ylab("Niche overlap") +
  theme_light(16)

gam_niche.overlap <- gam(value ~ year + s(elev.mean), 
                         data = filter(netmet, metric == "niche.overlap.HL"),
                         method = "REML")

gam_niche.overlap.vis <- getViz(gam_niche.overlap)
check.gamViz(gam_niche.overlap.vis)
summary(gam_niche.overlap.vis)

plot(gam_niche.overlap.vis, allTerms = TRUE) +
  xlab("Elevation")

##### C.score
ggplot(filter(netmet, metric == "C.score.HL"), aes(elev.mean, value)) +
  geom_point(aes(color = year)) +
  #geom_smooth(method = "lm", formula = y ~ poly(x, 3)) +
  geom_smooth(method = "gam", formula = y ~ s(x), color = "black", linetype = "dashed", size = 0.75) +
  #geom_smooth() +
  xlab("Elevation") +
  ylab("C score") +
  theme_light(16)

gam_cscore <- gam(value ~ year + s(elev.mean), 
                  data = filter(netmet, metric == "C.score.HL"),
                  method = "REML")

gam_cscore.vis <- getViz(gam_cscore)
check.gamViz(gam_cscore.vis)
summary(gam_cscore.vis)

plot(gam_cscore.vis, allTerms = TRUE) +
  xlab("Elevation")

##### Connectance
ggplot(filter(netmet, metric == "weighted.connectance"), aes(elev.mean, value)) +
  geom_point(aes(color = year)) +
  #geom_smooth(method = "lm", formula = y ~ poly(x, 3)) +
  geom_smooth(method = "gam", formula = y ~ s(x), color = "black", linetype = "dashed", size = 0.75) +
  #geom_smooth() +
  xlab("Elevation") +
  ylab("Connectance (weighted)") +
  theme_light(16)

gam_connectance <- gam(value ~ year + s(elev.mean), 
                  data = filter(netmet, metric == "weighted.connectance"),
                  method = "REML")

gam_connectance.vis <- getViz(gam_connectance)
check.gamViz(gam_connectance.vis)
summary(gam_connectance.vis)

plot(gam_connectance.vis, allTerms = TRUE) +
  xlab("Elevation")

##### NODF
ggplot(filter(netmet, metric == "weighted.NODF"), aes(elev.mean, value)) +
  geom_point(aes(color = year)) +
  #geom_smooth(method = "lm", formula = y ~ poly(x, 3)) +
  geom_smooth(method = "gam", formula = y ~ s(x), color = "black", linetype = "dashed", size = 0.75) +
  #geom_smooth() +
  xlab("Elevation") +
  ylab("NODF (weighted)") +
  theme_light(16)

gam_NODF <- gam(value ~ year + s(elev.mean), 
                data = filter(netmet, metric == "weighted.NODF"),
                method = "REML")

gam_NODF.vis <- getViz(gam_NODF)
check.gamViz(gam_NODF.vis)
summary(gam_NODF.vis)

plot(gam_NODF.vis, allTerms = TRUE) +
  xlab("Elevation")

##### Web asymmetry
ggplot(filter(netmet, metric == "web.asymmetry"), aes(elev.mean, value)) +
  geom_point(aes(color = year)) +
  #geom_smooth(method = "lm", formula = y ~ poly(x, 3)) +
  geom_smooth(method = "gam", formula = y ~ s(x), color = "black", linetype = "dashed", size = 0.75) +
  #geom_smooth() +
  xlab("Elevation") +
  ylab("Web asymmetry") +
  theme_light(16)

gam_webass <- gam(value ~ year + s(elev.mean), 
                data = filter(netmet, metric == "web.asymmetry"),
                method = "REML")

gam_webass.vis <- getViz(gam_webass)
check.gamViz(gam_webass.vis)
summary(gam_webass.vis)

plot(gam_webass.vis, allTerms = TRUE) +
  xlab("Elevation")
```

### Do BB species differ in their k.type associations?
```{r}
web_by_site <- net %>%
  group_by(site, bb.sp, k.type.s) %>% # group by site, date, bee species, plant species
  summarize(freq = n()) %>% # calculate interaction frequency per species pair within each site
  na.omit() %>%
  group_by(site) %>%
  filter(length(unique(bb.sp)) > 1 & length(unique(k.type.s)) > 1) %>%
  ungroup() %>%
  dplyr::select(higher = bb.sp, lower = k.type.s, webID = site, freq) %>% # rename columns to match bipartite's expectations
  data.frame() %>% # convert to data frame
  frame2webs()

visweb(web[[1]])
plotweb(web[[1]])

web <- net %>%
  group_by(year, elev.class2, bb.sp, k.type.s) %>% # group by site, date, bee species, plant species
  summarize(freq = n()) %>% # calculate interaction frequency per species pair within each site
  na.omit() %>%
  unite(webID, c(year, elev.class2), sep = "_") %>%
  dplyr::select(higher = bb.sp, lower = k.type.s, webID, freq) %>% # rename columns to match bipartite's expectations
  data.frame() %>% # convert to data frame
  frame2webs()

# 2010 high
visweb(web[[1]])
plotweb(web[[1]])

# 2010 low
visweb(web[[2]])
plotweb(web[[2]])

# 2010 med
visweb(web[[3]])
plotweb(web[[3]])

# 2011 high
visweb(web[[4]])
plotweb(web[[4]])

# 2011 low
visweb(web[[5]])
plotweb(web[[5]])

# 2011 med
visweb(web[[6]])
plotweb(web[[6]])

# 2012 high
visweb(web[[7]])
plotweb(web[[7]])

# 2012 low
visweb(web[[8]])
plotweb(web[[8]])

# 2012 med
visweb(web[[9]])
plotweb(web[[9]])

```

### Does tongue length class predict k.type association? Remember that B. wurflenii is primarily a nectar robber.
```{r}
library(infer)

kugler_desc <- read_delim("./kugler.tsv", delim = "\t") %>%
  select(1,2)

ggplot(bb_traits, aes(pbl.w)) +
  geom_histogram()

net_cont <- net %>%
  group_by(year, elev.class2, pbl.w.class, k.type.ss) %>%
  summarize(freq = n()) %>%
  group_by(elev.class2, pbl.w.class) %>%
  mutate(prop = freq/sum(freq))

ggplot(net_cont, aes(pbl.w.class, freq, fill = k.type.ss)) +
  geom_col() +
  facet_grid(.~elev.class2)

ggplot(net_cont, aes(pbl.w.class, prop, fill = k.type.ss)) +
  geom_col() +
  facet_grid(.~elev.class2)

chisq_mod <- chisq.test(net$pbl.w.class, net$k.type.ss)
chisq_mod

  c <- net %>%
  select(k.type.s, pbl.w.class) %>%
  mutate(k.type.s = factor(k.type.s),
         pbl.w.class = factor(pbl.w.class)) %>%
  chisq_test(k.type.s ~ pbl.w.class)

```

### Let's look at how abundance varies on a per-species basis
```{r eval=FALSE, include=FALSE}
species <- net %>%
  group_by(site, date, year, dayofyear, gdd.cum, phase, elev.mean, elev.class2, bb.sp, subgenus, pbl.w, pbl.w.class) %>%
  summarize(abund = n()) %>%
  group_by(site, date) %>%
  mutate(prop = abund/sum(abund))

subgenera <- net %>%
  group_by(site, date, year, dayofyear, gdd.cum, phase, elev.mean, elev.class2, subgenus, pbl.w, pbl.w.class) %>%
  summarize(abund = n()) %>%
  group_by(site, date) %>%
  mutate(prop = abund/sum(abund))

ggplot(species, aes(dayofyear, abund)) +
  geom_point(alpha = 0.1) +
  geom_smooth() +
  facet_grid(year ~ bb.sp)

ggplot(species, aes(dayofyear, prop)) +
  geom_point(alpha = 0.1) +
  geom_smooth() +
  facet_grid(year ~ bb.sp)

ggplot(species, aes(elev.mean, abund)) +
  geom_point(alpha = 0.1) +
  geom_smooth(se = FALSE, span = 1) +
  facet_grid(year ~ bb.sp, scales = "free")

ggplot(species, aes(elev.mean, prop)) +
  geom_point(alpha = 0.1) +
  geom_smooth(se = FALSE, span = 1) +
  facet_grid(year ~ bb.sp, scales = "free")

ggplot(subgenera, aes(dayofyear, abund)) +
  geom_point(alpha = 0.1) +
  geom_smooth() +
  facet_grid(year ~ subgenus)

ggplot(subgenera, aes(dayofyear, prop)) +
  geom_point(alpha = 0.1) +
  geom_smooth() +
  facet_grid(year ~ subgenus)

ggplot(subgenera, aes(elev.mean, abund)) +
  geom_point(alpha = 0.1) +
  geom_smooth() +
  facet_grid(year ~ subgenus)

ggplot(subgenera, aes(elev.mean, prop)) +
  geom_point(alpha = 0.1) +
  geom_smooth(se = FALSE, span = 1) +
  facet_grid(year ~ subgenus)

### Pairs analysis
species_pairs <- species %>%
  select(site, year, date, bb.sp, abund) %>%
  ungroup() %>%
  spread(bb.sp, abund) %>%
  replace(is.na(.), 0)

ggplot(species_pairs, aes(pasc, prat)) +
  geom_point() +
  geom_smooth(method = "lm") +
  facet_wrap(~year)

ggpairs(species_pairs, columns = c("bss", "camp", "flav", 
                                   "gers", "hort", "humi",
                                   "hypn", "jone", "lapi",
                                   "mend", "mont", "muci",
                                   "pasc", "prat", "psyt",
                                   "pyre", "soro", "wurf"))

ggpairs(species, columns = c("bb.sp", "abund"),
        cardinality_threshold = 20)

```

### Plot some dang networks
```{r eval=FALSE, include=FALSE}
web <- net %>%
  group_by(year, elev.class2, phase, bb.sp, plant.sp) %>% # group by site, date, bee species, plant species
  summarize(freq = n()) %>%
  unite(webID, c(year, elev.class2, phase), sep = "_") %>% # calculate interaction frequency per species pair within each site
  dplyr::select(higher = bb.sp, lower = plant.sp, webID, freq) %>% # rename columns to match bipartite's expectations
  data.frame() %>% # convert to data frame
  frame2webs()  # convert to to bipartite web

visweb(web[[1]])
visweb(web[[2]])
```

######## Old analyses that I probably don't care to use but don't want to delete yet #########

### Per-transect network analysis -- is it worth it? I would have to filter out the transects for which the networks are too small.
#### Network level
```{r eval=FALSE, include=FALSE}
per.transect_netmet <- net %>%
  group_by(site, dayofyear, gdd.cum, year, bb.sp, plant.sp) %>% # group by site, date, bee species, plant species
  summarize(freq = n()) %>% # calculate interaction frequency per species pair within each site
  group_by(site, dayofyear, gdd.cum, year) %>%
  mutate(size = sum(freq),
         bb.rich = length(unique(bb.sp)),
         fl.rich = length(unique(plant.sp)),
         total.rich = bb.rich + fl.rich) %>%
  filter(bb.rich > 1 & fl.rich > 1) %>%
  unite(webID, c(site, dayofyear, gdd.cum, year, size), sep = "_") %>%
  dplyr::select(higher = bb.sp, lower = plant.sp, webID, freq) %>% # rename columns to match bipartite's expectations
  data.frame() %>% # convert to data frame
  frame2webs() %>% # convert to to bipartite web
  map(networklevel, weighted = TRUE, level = "higher", index = c("web asymmetry", "weighted NODF", "niche overlap", "C score", "H2")) %>% # We probably care mainly about the BB level of the network
  data.frame() %>%
  t() %>%
  data.frame() %>%
  rownames_to_column() %>%
  dplyr::select(id = rowname, everything()) %>%
  as_tibble() %>%
  gather(metric, value, -id) %>%
  separate(id, into = c("site", "yday", "gdd.cum", "year", "size"), sep = "_") %>%
  mutate(size = as.numeric(size),
         year = factor(year),
         gdd.cum = as.numeric(gdd.cum),
         yday = as.numeric(yday),
         site = as.factor(site)) %>%
  left_join(site_data) %>%
  left_join(abund_div)

### Can we explain network metrics with abundance and diversity data?
# BB rich
ggplot(per.transect_netmet, aes(bb.rich, value)) +
  geom_point() +
  geom_smooth(method = "lm") +
  facet_grid(metric ~ year, scales = "free")
# FL rich
ggplot(per.transect_netmet, aes(fl.rich, value)) +
  geom_point() +
  geom_smooth(method = "lm") +
  facet_grid(metric ~ year, scales = "free")
# BB abund
ggplot(per.transect_netmet, aes(log(bb.abund), value)) +
  geom_point() +
  geom_smooth(method = "lm") +
  facet_grid(metric ~ year, scales = "free")
# FL abund
ggplot(per.transect_netmet, aes(log(fl.abund), value)) +
  geom_point() +
  geom_smooth(method = "lm") +
  facet_grid(metric ~ year, scales = "free")
# Per-capita FL abund
ggplot(per.transect_netmet, aes(log(fl.per.bb), value)) +
  geom_point() +
  geom_smooth(method = "lm") +
  facet_grid(metric ~ year, scales = "free")

### Can we explain network metrics with elevation and/or time?
ggplot(per.transect_netmet, aes(elev.mean, value, color = log(size))) +
  geom_point() +
  geom_smooth() +
  facet_grid(metric ~ year, scales = "free")

ggplot(per.transect_netmet, aes(log(size), value, color = log(size))) +
  geom_point() +
  geom_smooth() +
  facet_grid(metric ~ year, scales = "free")

ggplot(filter(per.transect_netmet, size > 1), 
       aes(yday, value)) +
  geom_point() +
  geom_smooth() +
  facet_grid(metric ~ year, scales = "free")

ggplot(filter(per.transect_netmet, size > 10), 
       aes(gdd.cum, value)) +
  geom_point() +
  geom_smooth() +
  facet_grid(metric ~ year, scales = "free")

ggplot(filter(per.transect_netmet, size > 10), 
       aes(yday, elev.mean)) +
  geom_point(aes(color = value)) +
  geom_smooth() +
  facet_grid(metric ~ year, scales = "free")
```

#### Species level
```{r eval=FALSE, include=FALSE}
per.transect_spmet <- net %>%
  group_by(site, dayofyear, gdd.cum, year, bb.sp, plant.sp) %>% # group by site, date, bee species, plant species
  summarize(freq = n()) %>% # calculate interaction frequency per species pair within each site
  group_by(site, dayofyear, gdd.cum, year) %>%
  mutate(size = sum(freq),
         bb.rich = length(unique(bb.sp)),
         fl.rich = length(unique(plant.sp)),
         total.rich = bb.rich + fl.rich) %>%
  filter(bb.rich > 1 & fl.rich > 1) %>%
  unite(webID, c(site, dayofyear, gdd.cum, year, size), sep = "_") %>%
  dplyr::select(higher = bb.sp, lower = plant.sp, webID, freq) %>% # rename columns to match bipartite's expectations
  data.frame() %>% # convert to data frame
  frame2webs() %>% # convert to to bipartite web
  map(specieslevel, level = "higher", index = "d") %>% # We probably care mainly about the BB level of the network
  map(rownames_to_column) %>%
  map(as_tibble) %>%
  bind_rows(.id = "id") %>% # Collapse list into big data frame
  rename(sp = rowname) %>%
  gather(metric, value, -c(sp, id)) %>% # get mean value across iterations for each metric
  separate(id, sep = "_", into = c("site", "yday", "gdd.cum", "year", "size")) %>%
  mutate(size = as.numeric(size),
         year = factor(year),
         gdd.cum = as.numeric(gdd.cum),
         yday = as.numeric(yday)) %>%
  left_join(site_data) %>%
  left_join(bb_traits, by = c("sp" = "bb.sp")) %>%
  left_join(abund_div)

per.transect_spmet.mean <- per.transect_spmet %>%
  group_by(site, yday, gdd.cum, year) %>%
  summarize(value = mean(value)) %>%
  left_join(site_data)

ggplot(filter(per.transect_spmet), 
       aes(log(fl.abund), value)) +
  geom_point() +
  geom_smooth() +
  facet_grid(. ~ year, scales = "free")

ggplot(filter(per.transect_spmet), 
       aes(elev.mean, size)) +
  geom_point() +
  geom_smooth() +
  facet_grid(. ~ year, scales = "free")

ggplot(filter(per.transect_spmet.mean), 
       aes(yday, value)) +
  geom_point() +
  geom_smooth() +
  facet_grid(. ~ year, scales = "free")

ggplot(filter(per.transect_spmet), 
       aes(elev.mean, value)) +
  geom_point() +
  geom_smooth() +
  facet_grid(. ~ year, scales = "free")

ggplot(filter(per.transect_spmet, size >= 10), 
       aes(gdd.cum, value, color = log(size))) +
  geom_point() +
  geom_smooth() +
  facet_grid(metric ~ year, scales = "free")

ggplot(filter(per.transect_spmet, size >= 10), 
       aes(yday, value, color = log(size))) +
  geom_point() +
  geom_smooth() +
  facet_grid(metric ~ year, scales = "free")
```

### Phase x elev network analysis
```{r eval=FALSE, include=FALSE}
### Prepare pxe abundance and diversity table
pxe_abund_div <- abund_div %>%
  select(-gdd.cum, -yday, -week, -month, -date, -elev.class2) %>%
  gather(ad.metric, ad.value, -c(site, year, phase, elev.mean)) %>%
  group_by(site, phase, year, ad.metric) %>%
  summarize(ad.value = mean(ad.value, na.rm = TRUE))

pxe_net <- rare_net_nt2(net, 4, 20) %>%
  full_join((pxe_abund_div))

ggplot(filter(pxe_net, 
              metric == "C.score.HL" & ad.metric == "fl.abund"),
       aes(log(ad.value), value)) +
  geom_point() +
  geom_smooth(method = "lm")  +
  facet_grid(year ~ phase, scales = "free") +
  xlab("log(floral abundance)") +
  ylab("C.score")

ggplot(filter(pxe_net, 
              metric == "C.score.HL" & ad.metric == "fl.per.bb"),
       aes(log(ad.value), value)) +
  geom_point() +
  geom_smooth(method = "lm") +
  facet_grid(year ~ phase, scales = "free") +
  xlab("log(per-capita floral abundance)") +
  ylab("C.score")

ggplot(filter(pxe_net, 
              metric == "H2" & ad.metric == "fl.rich"),
       aes(ad.value, value)) +
  geom_point() +
  geom_smooth(method = "lm") +
  facet_grid(year ~ phase, scales = "free") +
  xlab("Floral richness") +
  ylab("H2")

ggplot(filter(pxe_net, 
              metric == "H2" & ad.metric == "bb.abund"),
       aes(log(ad.value), value)) +
  geom_point() +
  geom_smooth(method = "lm") +
  facet_grid(year ~ phase, scales = "free") +
  xlab("BB abund") +
  ylab("H2")

ggplot(filter(pxe_net, 
              metric == "H2" & ad.metric == "bb.rich"),
       aes(ad.value, value)) +
  geom_point() +
  geom_smooth(method = "lm") +
  facet_grid(year ~ phase, scales = "free") +
  xlab("BB richness") +
  ylab("H2")

ggplot(filter(pxe_net, metric == "weighted.NODF"), aes(elev.mean, value)) +
  geom_point() +
  geom_smooth(method = "lm", formula = y ~ poly(x, 3)) +
  facet_grid(year ~ phase, scales = "free") +
  xlab("Elevation") +
  ylab("Weighted NODF")

ggplot(filter(pxe_net, metric == "H2"), aes(elev.mean, value)) +
  geom_point() +
  geom_smooth(method = "lm", formula = y ~ poly(x, 3)) +
  facet_grid(year ~ phase, scales = "free") +
  xlab("Elevation") +
  ylab("H2")

ggplot(filter(pxe_net, metric == "C.score.HL"), aes(elev.mean, value)) +
  geom_point() +
  geom_smooth(method = "lm", formula = y ~ poly(x, 3)) +
  facet_grid(year ~ phase, scales = "free") +
  xlab("Elevation") +
  ylab("C.score")

ggplot(filter(pxe_net, metric == "niche.overlap.HL"), aes(elev.mean, value)) +
  geom_point() +
  geom_smooth(method = "lm", formula = y ~ poly(x, 3)) +
  facet_grid(year ~ phase, scales = "free") +
  xlab("Elevation") +
  ylab("Niche overlap")

ggplot(filter(pxe_net, metric == "generality.HL"), aes(elev.mean, value)) +
  geom_point() +
  geom_smooth(method = "lm", formula = y ~ poly(x, 3)) +
  facet_grid(year ~ phase, scales = "free") +
  xlab("Elevation") +
  ylab("Generality")

ggplot(filter(pxe_net, metric == "web.asymmetry"), aes(elev.mean, value)) +
  geom_point() +
  geom_smooth(method = "lm", formula = y ~ poly(x, 3)) +
  facet_grid(year ~ phase, scales = "free") +
  xlab("Elevation") +
  ylab("Web asymmetry")

pxe_sp <- rare_net_sp2(net, 4, 20)  %>%
  full_join((pxe_abund_div))

ggplot(filter(pxe_sp, 
              metric == "d" &
                sp %in% c("bss", "hort", "pasc", "prat", "soro", "wurf", "psit") &
                ad.metric == "fl.rich"), 
       aes(ad.value, value, color = year)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  facet_grid(sp ~ phase, scales = "free") +
  ylim(c(0,1)) +
  xlab("Floral richness") +
  ylab("d'")

pxe_sp_mean <- pxe_sp %>%
  filter(sp %in% c("bss", "hort", "pasc", "prat", "soro", "wurf", "psit")) %>%
  group_by(site, year, phase, elev.mean, metric) %>%
  summarize(value = mean(value, na.rm = TRUE))

ggplot(filter(pxe_sp_mean, metric == "d"), aes(elev.mean, value)) +
  geom_point() +
  geom_smooth(method = "lm", formula = y ~ poly(x, 3)) +
  facet_grid(year ~ phase, scales = "free") +
  xlab("Elevation") +
  ylab("Mean d'")

ggplot(filter(pxe_sp_mean, metric == "species.strength"), aes(elev.mean, value)) +
  geom_point() +
  geom_smooth(method = "lm", formula = y ~ poly(x, 3)) +
  facet_grid(year ~ phase, scales = "free") +
  xlab("Elevation") +
  ylab("Mean species strength'")

ggplot(filter(pxe_sp, 
              metric == "species.strength",
              sp %in% c("bss", "hort", "pasc", "prat", "soro", "wurf", "psit")), 
       aes(elev.mean, value, color = year)) +
  geom_point() +
  geom_smooth(method = "lm", formula = y ~ poly(x, 3)) +
  facet_grid(sp ~ phase, scales = "free") +
  xlab("Elevation") +
  ylab("Species strength")

ggplot(filter(pxe_sp, 
              metric == "d",
              sp %in% c("bss", "hort", "pasc", "prat", "soro", "wurf", "psit")), 
       aes(elev.mean, value, color = year)) +
  geom_point() +
  geom_smooth(method = "lm", formula = y ~ poly(x, 3)) +
  facet_grid(sp ~ phase, scales = "free") +
  ylim(c(0,1)) +
  xlab("Elevation") +
  ylab("d'")

ggplot(filter(pxe_sp, 
              metric == "d",
              sp %in% c("bss", "hort", "pasc", "prat", "soro", "wurf", "psit")), 
       aes(elev.mean, value)) +
  geom_point(aes(color = year)) +
  geom_smooth(method = "lm", formula = y ~ poly(x, 3)) +
  facet_grid(sp ~ phase, scales = "free") +
  ylim(c(0,1)) +
  xlab("Elevation") +
  ylab("d'")

ggplot(filter(pxe_sp, 
              metric == "d",
              sp %in% c("bss", "hort", "pasc", "prat", "soro", "wurf", "psit")), 
       aes(elev.mean, value, color = phase)) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = "lm", formula = y ~ poly(x, 3)) +
  facet_grid(year ~ sp, scales = "free") +
  theme_light() +
  xlab("Elevation") +
  ylab("d'")
```

```{r}

# network level PxE
rare_net_nt_pxe <- function(net, n, iter) {

  out <- data.frame()

  t_filter <- net %>%
    select(site, date, phase, year) %>%
    unique() %>%
    group_by(site, phase, year) %>%
    mutate(transects = n()) %>%
    filter(transects >= n)

  for(i in 1:iter) {
    temp <- net %>%
      semi_join(t_filter) %>%
      select(site, date, phase, year) %>%
      unique() %>%
      group_by(site, phase, year) %>%
      sample_n(n) %>%
      left_join(net) %>%
      group_by(site, phase, year, bb.sp, plant.sp) %>% # group by site, date, bee species, plant species
      summarize(freq = n()) %>% # calculate interaction frequency per species pair within each site
      group_by(site, phase, year) %>%
      filter(length(unique(bb.sp)) > 1 & length(unique(plant.sp)) > 1) %>%
      ungroup() %>%
      unite(webID, c(site, phase, year), sep = "_") %>%
      dplyr::select(higher = bb.sp, lower = plant.sp, webID, freq) %>% # rename columns to match bipartite's expectations
      data.frame() %>% # convert to data frame
      frame2webs() %>% # convert to to bipartite web
      map(networklevel, weighted = TRUE, level = "higher",
          index = c("web asymmetry", "weighted NODF", "H2", "C score", "niche overlap", "generality")) %>% 
      data.frame() %>%
      t() %>%
      data.frame() %>%
      rownames_to_column() %>%
      dplyr::select(id = rowname, everything()) %>%
      as_tibble() %>%
      mutate(iteration = rep(i, n()))

    out <- bind_rows(out, temp)
  } 
  out <- out  %>%
      gather(metric, value, -c(id, iteration)) %>% # get mean value across iterations for each metric
      group_by(id, metric) %>%
      summarize(value = mean(value)) %>%
      separate(id, sep = "_", into = c("site", "phase", "year")) %>%
      mutate(phase = factor(phase, levels = c("founding", "buildup", "reproductive"))) %>%
      left_join(site_data)
}

# species level version 2
rare_net_sp2 <- function(net, n, iter) {

  out <- data.frame()

  t_filter <- net %>%
    select(site, date, phase, year) %>%
    unique() %>%
    group_by(site, phase, year) %>%
    mutate(transects = n()) %>%
    filter(transects >= n)

  for(i in 1:iter) {
    temp <- net %>%
      semi_join(t_filter) %>%
      select(site, date, phase, year) %>%
      unique() %>%
      group_by(site, phase, year) %>%
      sample_n(n) %>%
      left_join(net) %>%
      group_by(site, phase, year, bb.sp, plant.sp) %>% # group by site, date, bee species, plant species
      summarize(freq = n()) %>% # calculate interaction frequency per species pair within each site
      group_by(site, phase, year) %>%
      filter(length(unique(bb.sp)) > 1 & length(unique(plant.sp)) > 1) %>%
      ungroup() %>%
      unite(webID, c(site, phase, year), sep = "_") %>%
      dplyr::select(higher = bb.sp, lower = plant.sp, webID, freq) %>% # rename columns to match bipartite's expectations
      data.frame() %>% # convert to data frame
      frame2webs() %>% # convert to to bipartite web
      map(specieslevel, level = "higher", index = c("d", "species strength")) %>%
      map(rownames_to_column) %>%
      map(as_tibble) %>%
      bind_rows(.id = "id") %>% # Collapse list into big data frame
      rename(sp = rowname) %>%
      mutate(iteration = rep(i, n()))

    out <- bind_rows(out, temp)
  } 
  out <- out %>%
    gather(metric, value, -c(sp, id, iteration)) %>% # get mean value across iterations for each metric
    group_by(sp, id, metric) %>%
    summarize(value = mean(value)) %>%
    separate(id, sep = "_", into = c("site", "phase", "year")) %>%
    mutate(phase = factor(phase, levels = c("founding", "buildup", "reproductive"))) %>%
    left_join(site_data) %>%
    left_join(bb_traits, by = c("sp" = "bb.sp"))
}

```