### 2020-07-02 analyses/climate.Rmd
broke down analyses into separate scripts; did some other things

---
title: "climate"
output: html_document
---

### Prepare environment
```{r}
library(tidyverse)
```

### Climate data
```{r, include=FALSE}
samples <- read_csv("../processed_data/network.csv")

climate <- read_csv("../processed_data/climate.csv") %>%
  mutate(year = factor(year),
         yday = yday(date)) %>%
  dplyr::select(site, date, yday, year, gdd.cum)
```

### Plot
```{r}
ggplot(climate, aes(yday, gdd.cum, color = year)) +
  geom_point(alpha = 0.01) +
  geom_smooth()
  #facet_wrap(~site) +
  xlim(c(83, 262))

```