### 2020-07-02 analyses/community_GAM_analysis.Rmd

---
title: "community_GAM_analysis"
output:
  pdf_document: default
  html_document: default
---

### Prepare environment
```{r setup, include=FALSE, echo=FALSE}
library(mgcv)
library(mgcViz)
library(tidymv)
#library(lme4)
#library(lmerTest)
library(tidyverse)
library(lubridate)
```

### Climate and weather data
```{r, include=FALSE, echo=FALSE}
gdd <- read_csv("../processed_data/climate.csv") %>%
  mutate(year = factor(year)) %>%
  dplyr::select(site, year, date, gdd.cum) %>%
  semi_join(read_csv("../processed_data/network.csv"), by = "site")
# 
# temperature <- read_csv("../processed_data/temperature.csv")
# 
# gdd_total <- gdd %>%
#   group_by(site, year) %>%
#   summarize(gdd.total = max(gdd.cum))
```

### Load site data
```{r, include=FALSE, echo=FALSE}
site_data <- read_csv("../processed_data/site_data.csv") %>%
  dplyr::select(-c(slope.calc, slope.est, elev.min, elev.max, temp.mean)) %>% # drop these variables
  mutate(elev.class = factor(elev.class, levels = c("oben", "mitte", "unten")),
         elev.class2 = factor(elev.class2, levels = c("high", "mid", "low")))

# %>% # turn elav.class into an ordered factor
#   left_join(gdd_total, by = "site")
```

### Load network and survey data
```{r, echo=FALSE, include=FALSE}
bb_traits <- read_csv("../processed_data/bb_traits.csv") %>%
  dplyr::select(subgenus, bb.sp, pbl.w, pbl.w.class)

fl_traits <- read.csv("../processed_data/floral_k_type_nom_fix.csv") %>%
  dplyr::select(plant.sp, k.type) %>%
  mutate(k.type.s = str_extract(k.type, "[-+]?[0-9]*\\.?[0-9]+"),
         k.type.ss = str_extract(k.type, "[-+]?[0-9]*"))

fl_corolla_depth <- read.csv("../processed_data/flower_morphology_GitaBenadi.csv") %>%
  group_by(plant.sp = Species) %>%
  summarize(corolla.depth = mean(LN))

fl_tax <- read_csv("../processed_data/floral_tax.csv")

breaks <- read_csv("./output/breaks.csv") %>%
  mutate(year = factor(year))

net <- read_csv("../processed_data/network.csv") %>%
  unite(site.year, site, year, sep = "_", remove = FALSE) %>% # create site.year field that can later be split back up
  unite(site.date, site, date, sep = "_", remove = FALSE) %>%
  mutate(year = factor(year)) %>% # convert year to factor
  full_join(bb_traits, by = c("bb.sp" = "bb.sp")) %>%
  left_join(site_data, by = c("site")) %>%
  left_join(gdd, by = c("site", "date", "year")) %>%
  left_join(breaks, by = c("site", "year")) %>%
  left_join(fl_traits) %>%
  left_join(fl_corolla_depth) %>%
  left_join(fl_tax, by = "plant.genus") %>%
  #left_join(temperature, by = c("site", "date")) %>%
  ungroup() %>%
  mutate(phase = case_when(
    gdd.cum < break1.gdd ~ "founding",
    gdd.cum >= break1.gdd & gdd.cum < break2.gdd ~ "buildup",
    gdd.cum > break2.gdd ~ "reproductive"
  )) %>%
  mutate(phase = factor(phase, levels = c("founding", "buildup", "reproductive"))) %>%
  mutate(caste = case_when(
    caste == "worker" ~ "worker",
    caste == "male" ~ "reproductive",
    caste == "youngqueen" ~ "reproductive",
    caste == "queen" ~ "foundress",
    caste == "oldqueen" ~ "foundress"
  )) %>%
  rename(yday = dayofyear) %>%
  arrange(site, date)

survey <- read_csv("../processed_data/floral_survey.csv") %>%
  mutate(year = factor(year)) %>% # convert year to factor
  semi_join(net, by = "plant.sp") %>% # Consider only plants visited at a given site-date
  left_join(gdd, by = c("site", "date", "year")) %>%
  left_join(breaks, by = c("site", "year")) %>%
  left_join(site_data, by = c("site")) %>%
  left_join(fl_traits) %>%
  left_join(fl_traits) %>%
  left_join(fl_tax) %>%
  #left_join(temperature, by = c("site", "date")) %>%
  ungroup() %>%
  mutate(phase = case_when(
    gdd.cum < break1.gdd ~ "founding",
    gdd.cum >= break1.gdd & gdd.cum < break2.gdd ~ "buildup",
    gdd.cum > break2.gdd ~ "reproductive"
  )) %>%
  mutate(phase = factor(phase, levels = c("founding", "buildup", "reproductive"))) %>%
  rename(yday = dayofyear) %>%
  arrange(site, date)
```

### FL and BB abundance / diversity
```{r echo=FALSE, include=FALSE}
### Per-transect Bombus abundance
bb_abund <- net %>%
  group_by(site, date, yday = dayofyear, gdd.cum, year, elev.class2, 
           elev.mean, break1.gdd, break2.gdd) %>%
  summarize(bb.abund = n())

ggplot(bb_abund, aes(log(top.fl.abund + 0.01), bb.abund, color = year)) +
  geom_point() +
  geom_smooth(method = "lm") +
  facet_grid(elev.class2~phase, scales = "free")

ggplot(bb_abund, aes(log(top.fl.abund + 0.01), bb.abund, color = year)) +
  geom_point(alpha = 0.5) +
  #geom_smooth(method = "lm") +
  geom_smooth() +
  facet_wrap(~bb.sp, scales = "free")

#### Species-level 
bb_abund_sp <- net %>%
  group_by(site, date, yday, gdd.cum, year, elev.class2, 
           elev.mean, break1.gdd, break2.gdd, bb.sp) %>%
  summarize(bb.abund = n()) %>%
  ungroup() %>%
  mutate(bb.sp = factor(bb.sp))

bb_abund_sp_workers <- net %>%
  filter(caste == "worker") %>%
  group_by(site, date, yday = dayofyear, gdd.cum, year, elev.class2, 
           elev.mean, break1.gdd, break2.gdd, bb.sp) %>%
  summarize(bb.abund = n()) %>%
  ungroup() %>%
  mutate(bb.sp = factor(bb.sp))

bb_abund_caste <- net %>%
  group_by(site, date, yday = dayofyear, gdd.cum, year, elev.class2, 
           elev.mean, break1.gdd, break2.gdd, caste) %>%
  summarize(bb.abund = n()) %>%
  ungroup() %>%
  mutate(caste = factor(caste))

bb_abund_caste_sp <- net %>%
  group_by(site, date, yday = dayofyear, gdd.cum, year, elev.class2, 
           elev.mean, break1.gdd, break2.gdd, caste, bb.sp) %>%
  summarize(bb.abund = n()) %>%
  ungroup() %>%
  mutate(caste = factor(caste),
         bb.sp = factor(bb.sp))

### Per-transect Bombus richness
bb_rich <- net %>%
  group_by(site, date, yday = dayofyear, gdd.cum, year, elev.class2,
           elev.mean, break1.gdd, break2.gdd) %>%
  summarize(bb.rich = length(unique(bb.sp))) %>%
  mutate(year2010 = ifelse(year == 2010, 1, 0),
         year2011 = ifelse(year == 2011, 1, 0),
         year2012 = ifelse(year == 2012, 1, 0))

ggplot(bb_rich, aes(yday, bb.rich)) +
  geom_point() +
  facet_wrap(~year)

### Per-transect floral abund
fl_abund <- survey %>%
  group_by(site, date, yday = dayofyear, gdd.cum, year, elev.class2, 
           elev.mean, break1.gdd, break2.gdd) %>%
  summarize(fl.abund = sum(flower.cover)) %>%
  mutate(year2010 = ifelse(year == 2010, 1, 0),
         year2011 = ifelse(year == 2011, 1, 0),
         year2012 = ifelse(year == 2012, 1, 0))

### Per-transect floral richness
fl_rich <- survey %>%
  filter(flower.cover > 0) %>%
  group_by(site, date, yday = dayofyear, gdd.cum, year, elev.class2, 
           elev.mean, break1.gdd, break2.gdd) %>%
  summarize(fl.rich = n())  %>%
  mutate(year2010 = ifelse(year == 2010, 1, 0),
         year2011 = ifelse(year == 2011, 1, 0),
         year2012 = ifelse(year == 2012, 1, 0))
```

### Get most important plants for each BB species
```{r}
# Per transect visitation rates of flower species per BB species
visit_rates <- net %>%
  group_by(bb.sp, plant.sp, year, site, date, yday, gdd.cum, elev.class2, 
           elev.mean, break1.gdd, break2.gdd, phase) %>%
  summarize(visits = n()) %>%
  group_by(year, site, date, bb.sp, yday, gdd.cum, elev.class2, 
           elev.mean, break1.gdd, break2.gdd) %>%
  mutate(prop.visits = visits/sum(visits)) %>%
  arrange(site, year, date, bb.sp, -prop.visits) %>%
  mutate(cum.prop.visits = cumsum(prop.visits))

### Per-transect floral abund
fl_abund_sp <- survey %>%
  group_by(site, date, year, plant.sp, yday, gdd.cum, elev.class2, 
           elev.mean, break1.gdd, break2.gdd, phase) %>%
  summarize(fl.abund = sum(flower.cover))

### Preference per transect per species
prop_fl_abund <- visit_rates %>%
  left_join(fl_abund_sp) %>%
  na.omit() %>% # drop the samples without floral survey complement
  group_by(site, date, year, bb.sp, yday, gdd.cum, elev.class2, 
           elev.mean, break1.gdd, break2.gdd, phase) %>%
  mutate(prop.fl.abund = fl.abund/sum(fl.abund))

### Median preference per bb.sp per plant.sp
prop_fl_abund.med <- prop_fl_abund %>%
  group_by(bb.sp, plant.sp, year) %>%
  summarize(visits.median = median(visits),
            prop.visits.median = median(prop.visits),
            prop.fl.abund.median = median(prop.fl.abund)) %>%
  filter(prop.fl.abund.median < 1)

prop_fl_abund_median.lm <- lm(prop.visits.median ~ poly(prop.fl.abund.median, 2) + year, 
                       data = prop_fl_abund.med)
summary(prop_fl_abund_median.lm)
plot(prop_fl_abund_median.lm)

### Extract residuals
prop_fl_abund_mean.resid <- residuals(prop_fl_abund_median.lm) %>%
  data.frame() %>%
  as_tibble() %>%
  rename(preference = 1)

### Join residuals back into data
preference <- prop_fl_abund.mean %>%
  filter(prop.fl.abund.mean != 1) %>%
  bind_cols(prop_fl_abund_mean.resid)

### Join preference data to floral abundance data
fl_abund_sp_pref <- fl_abund_sp %>%
  inner_join(preference) %>%
  mutate(fl.abund.weighted = fl.abund^(1 + preference)) # Is this the right way to model preference?

ggplot(fl_abund_sp_pref, aes(preference)) +
  geom_histogram()

### Summarize weighted fl.abund by site*date (i.e. transect)
fl_abund_pref <- fl_abund_sp_pref %>%
  group_by(site, date, year, bb.sp, yday, gdd.cum, elev.class2, 
           elev.mean, break1.gdd, break2.gdd, phase) %>%
  summarize(fl.abund = sum(fl.abund),
            fl.abund.weighted = sum(fl.abund.weighted)) %>%
  left_join(bb_abund_sp) %>%
  replace_na(list(bb.abund = 0))

ggplot(fl_abund_pref, aes(log(fl.abund.weighted + 0.01), bb.abund, color = year)) +
  geom_point() +
  geom_smooth() +
  facet_wrap(~bb.sp, scales = "free")

ggplot(fl_abund_pref, aes(log(fl.abund + 0.01), bb.abund, color = year)) +
  geom_point() +
  geom_smooth() +
  facet_wrap(~bb.sp, scales = "free")

fl_abund_pref.sum <- fl_abund_pref %>%
  group_by(site, year, bb.sp) %>%
  summarize(tot.fl.abund.weighted = sum(fl.abund.weighted),
            tot.fl.abund = sum(fl.abund),
            tot.bb.abund = sum(bb.abund))

ggplot(fl_abund_pref.sum, aes(log(tot.fl.abund.weighted + 0.01), tot.bb.abund, color = year)) +
  geom_point() +
  geom_smooth() +
  facet_wrap(~bb.sp, scales = "free")

ggplot(fl_abund_pref.sum, aes(log(tot.fl.abund + 0.01), tot.bb.abund, color = year)) +
  geom_point() +
  geom_smooth() +
  facet_wrap(~bb.sp, scales = "free")

ggplot(filter(preference, plant.sp == "Trifolium pratense"), 
       aes(bb.sp, preference, color = year)) +
  geom_point()

ggplot(filter(preference, plant.sp == "Acinos alpinus"), 
       aes(bb.sp, preference, color = year)) +
  geom_point()

ggplot(filter(prop_fl_abund, prop.fl.abund != 1), aes(prop.fl.abund, prop.visits)) +
  geom_point(alpha = 0.05) +
  geom_smooth(method = "lm", formula = y ~ poly(x, 2)) 
  #geom_smooth()

ggplot(filter(prop_fl_abund.mean, prop.fl.abund.mean != 1), 
       aes(prop.fl.abund.mean, prop.visits.mean)) +
  geom_point(alpha = 0.25) +
  geom_smooth(method = "lm")

```

**The shape of floral resource variation in time, space, and space-time** \

If we consider floral resource abundance with respect to time alone, we basically get a monotonic and unimodal response
```{r}
fl_abund_time <- bam(log(fl.abund + 0.01) ~
                       year +
                       s(yday, 
                         bs = "tp", 
                         k = 10, m = 2),
                     data = fl_abund,
                     discrete = TRUE,
                     method = "fREML") %>% getViz()

check.gamViz(fl_abund_time)
summary(fl_abund_time)
print(plot(fl_abund_time , allTerms = TRUE), pages = 1)
plot(sm(fl_abund_time, 1)) + 
  l_points() + 
  l_fitLine() + 
  l_ciLine()

fl_abund_time_2010 <- bam(log(fl.abund + 0.01) ~
                            s(yday, bs = "tp", k = 10, m = 2),
                          data = filter(fl_abund, year == 2010),
                          discrete = TRUE,
                          method = "fREML") %>% getViz()

check.gamViz(fl_abund_time_2010)
summary(fl_abund_time_2010)
print(plot(fl_abund_time_2010 , allTerms = TRUE), pages = 1)

fl_abund_time_2011 <- bam(log(fl.abund + 0.01) ~
                            s(yday, bs = "tp", k = 10, m = 2),
                          data = filter(fl_abund, year == 2011),
                          discrete = TRUE,
                          method = "fREML") %>% getViz()

check.gamViz(fl_abund_time_2011)
summary(fl_abund_time_2011)
print(plot(fl_abund_time_2011 , allTerms = TRUE), pages = 1)

fl_abund_time_2012 <- bam(log(fl.abund + 0.01) ~
                            s(yday, bs = "tp", k = 10, m = 2),
                          data = filter(fl_abund, year == 2012),
                          discrete = TRUE,
                          method = "fREML") %>% getViz()

check.gamViz(fl_abund_time_2012)
summary(fl_abund_time_2012)
print(plot(fl_abund_time_2012 , allTerms = TRUE), pages = 1)

ggplot(fl_abund, aes(yday, fl.abund, color = year)) +
  geom_point() +
  geom_smooth() +
  facet_wrap(~site, scales = "free")
```

If we look at floral abundance solely in terms of elevation, we get a more wiggly picture, with peaks ~1000 m and ~1600 m.
```{r}
fl_abund_elev <- bam(log(fl.abund + 0.01) ~
                       year +
                       s(elev.mean, 
                         bs = "tp", 
                         k = 10, m = 2),
                     data = fl_abund,
                     discrete = TRUE,
                     method = "fREML") %>% getViz()

check.gamViz(fl_abund_elev)
summary(fl_abund_elev)
print(plot(fl_abund_elev, allTerms = TRUE), pages = 1)
plot(sm(fl_abund_elev, 1)) + 
  l_points() + 
  l_fitLine() + 
  l_ciLine()

fl_abund_elev_2010 <- bam(log(fl.abund + 0.01) ~
                            s(elev.mean, bs = "tp", k = 10, m = 2),
                          data = filter(fl_abund, year == 2010),
                          discrete = TRUE,
                          method = "fREML") %>% getViz()

check.gamViz(fl_abund_elev_2010)
summary(fl_abund_elev_2010)
print(plot(fl_abund_elev_2010 , allTerms = TRUE), pages = 1)

fl_abund_elev_2011 <- bam(log(fl.abund + 0.01) ~
                            s(elev.mean, bs = "tp", k = 10, m = 2),
                          data = filter(fl_abund, year == 2011),
                          discrete = TRUE,
                          method = "fREML") %>% getViz()

check.gamViz(fl_abund_elev_2011)
summary(fl_abund_elev_2011)
print(plot(fl_abund_elev_2011 , allTerms = TRUE), pages = 1)

fl_abund_elev_2012 <- bam(log(fl.abund + 0.01) ~
                            s(elev.mean, bs = "tp", k = 10, m = 2),
                          data = filter(fl_abund, year == 2012),
                          discrete = TRUE,
                          method = "fREML") %>% getViz()

check.gamViz(fl_abund_elev_2012)
summary(fl_abund_elev_2012)
print(plot(fl_abund_elev_2012 , allTerms = TRUE), pages = 1)

```

But now let's look at space and time together in a 2D GAM.
```{r}
# Call models
fl_abund_2D <- bam(log(fl.abund + 0.01) ~ year +
                            te(yday, elev.mean, 
                               bs= c("tp", "tp"), 
                               k=c(10, 10), m=2),
                    data = fl_abund,
                    discrete = TRUE,
                    method = "fREML") %>% getViz()

check.gamViz(fl_abund_2D)
summary(fl_abund_2D)
plot(sm(fl_abund_2D, 1), too.far = 0.05) + l_fitRaster() + l_fitContour() + l_points()
plot(pterm(fl_abund_2D, 1))

fl_abund_2D_2010 <- bam(log(fl.abund + 0.01) ~ te(yday, elev.mean, 
                                                  bs= c("tp", "tp"), 
                                                  k=c(10, 10), m=2),
                        data = filter(fl_abund, year == 2010),
                        discrete = TRUE,
                        method = "fREML") %>% getViz()

check.gamViz(fl_abund_2D_2010)
summary(fl_abund_2D_2010)
plot(sm(fl_abund_2D_2010, 1)) + 
  l_fitRaster() + 
  l_fitContour() + 
  l_points()

fl_abund_2D_2011 <- bam(log(fl.abund + 0.01) ~ te(yday, elev.mean, 
                                                  bs= c("tp", "tp"), 
                                                  k=c(10, 10), m=2),
                        data = filter(fl_abund, year == 2011),
                        discrete = TRUE,
                        method = "fREML") %>% getViz()

check.gamViz(fl_abund_2D_2011)
summary(fl_abund_2D_2011)
plot(sm(fl_abund_2D_2011, 1)) + 
  l_fitRaster() + 
  l_fitContour() + 
  l_points()

fl_abund_2D_2012 <- bam(log(fl.abund + 0.01) ~ te(yday, elev.mean, 
                                                  bs= c("tp", "tp"), 
                                                  k=c(10, 10), m=2),
                        data = filter(fl_abund, year == 2012),
                        discrete = TRUE,
                        method = "fREML") %>% getViz()

check.gamViz(fl_abund_2D_2012)
summary(fl_abund_2D_2012)
plot(sm(fl_abund_2D_2012, 1)) + 
  l_fitRaster() + 
  l_fitContour() + 
  l_points()
```

How does the pooled abundance of each bumble bee species' top 5 floral hosts vary in time and elevation?
```{r}
fl_abund_2D_2010_top <- bam(log(top.fl.abund + 0.01) ~ te(yday, elev.mean, 
                                                  by = factor(bb.sp),
                                                  bs= c("tp", "tp"), 
                                                  k=c(10, 10), m=1),
                        data = filter(fl_abund_top_plants_pooled, year == 2010),
                        discrete = TRUE,
                        method = "fREML") %>% getViz()

check.gamViz(fl_abund_2D_2010_top)
summary(fl_abund_2D_2010_top)
print(plot(fl_abund_2D_2010_top , allTerms = TRUE), pages = 1)

fl_abund_2D_2011_top <- bam(log(top.fl.abund + 0.01) ~ te(yday, elev.mean, 
                                                  by = factor(bb.sp),
                                                  bs= c("tp", "tp"), 
                                                  k=c(10, 10), m=1),
                        data = filter(fl_abund_top_plants_pooled, year == 2011),
                        discrete = TRUE,
                        method = "fREML") %>% getViz()

check.gamViz(fl_abund_2D_2011_top)
summary(fl_abund_2D_2011_top)
print(plot(fl_abund_2D_2011_top , allTerms = TRUE), pages = 1)

fl_abund_2D_2012_top <- bam(log(top.fl.abund + 0.01) ~ te(yday, elev.mean, 
                                                  by = factor(bb.sp),
                                                  bs= c("tp", "tp"), 
                                                  k=c(10, 10), m=1),
                        data = filter(fl_abund_top_plants_pooled, year == 2012),
                        discrete = TRUE,
                        method = "fREML") %>% getViz()

check.gamViz(fl_abund_2D_2012_top)
summary(fl_abund_2D_2012_top)
print(plot(fl_abund_2D_2012_top , allTerms = TRUE), pages = 1)

```

Now let's do the same with BB abundance. First by time.
```{r}
bb_abund_time_2010 <- bam(bb.abund ~ s(yday, bs = "tp", k = 20, m = 2),
                          data = filter(bb_abund, year == 2010),
                          discrete = TRUE,
                          family = "poisson",
                          method = "fREML") %>% getViz()

bb_abund_time_2011 <- bam(bb.abund ~ s(yday, bs = "tp", k = 20, m = 2),
                          data = filter(bb_abund, year == 2011),
                          discrete = TRUE,
                          family = "poisson",
                          method = "fREML") %>% getViz()

bb_abund_time_2012 <- bam(bb.abund ~ s(yday, bs = "tp", k = 20, m = 2),
                          data = filter(bb_abund, year == 2012),
                          discrete = TRUE,
                          family = "poisson",
                          method = "fREML") %>% getViz()

check.gamViz(bb_abund_time_2010)
summary(bb_abund_time_2010)
print(plot(bb_abund_time_2010 , allTerms = TRUE), pages = 1)

check.gamViz(bb_abund_time_2011)
summary(bb_abund_time_2011)
print(plot(bb_abund_time_2011 , allTerms = TRUE), pages = 1)

check.gamViz(bb_abund_time_2012)
summary(bb_abund_time_2012)
print(plot(bb_abund_time_2012 , allTerms = TRUE), pages = 1)

ggplot(bb_abund_sp, aes(yday, bb.abund, color = bb.sp)) +
  geom_point() +
  geom_smooth(se = FALSE) +
  facet_grid(year~site, scales = "free")

ggplot(filter(bb_abund_sp, bb.sp == "prat"), aes(yday, bb.abund)) +
  geom_point() +
  geom_smooth(se = FALSE) +
  facet_grid(year~site, scales = "free")
```

Then by elevation...this kinda sucks
```{r}
bb_abund_elev_2010 <- bam(bb.abund ~ s(elev.mean, bs = "tp", k = 10, m = 2),
                          data = filter(bb_abund, year == 2010),
                          discrete = TRUE,
                          family = "poisson",
                          method = "fREML") %>% getViz()

bb_abund_elev_2011 <- bam(bb.abund ~ s(elev.mean, bs = "tp", k = 10, m = 2),
                          data = filter(bb_abund, year == 2011),
                          discrete = TRUE,
                          family = "poisson",
                          method = "fREML") %>% getViz()

bb_abund_elev_2012 <- bam(bb.abund ~ s(elev.mean, bs = "tp", k = 10, m = 2),
                          data = filter(bb_abund, year == 2012),
                          discrete = TRUE,
                          family = "poisson",
                          method = "fREML") %>% getViz()

check.gamViz(bb_abund_elev_2010)
summary(bb_abund_elev_2010)
print(plot(bb_abund_elev_2010 , allTerms = TRUE), pages = 1)

check.gamViz(bb_abund_elev_2011)
summary(bb_abund_elev_2011)
print(plot(bb_abund_elev_2011 , allTerms = TRUE), pages = 1)

check.gamViz(bb_abund_elev_2012)
summary(bb_abund_elev_2012)
print(plot(bb_abund_elev_2012 , allTerms = TRUE), pages = 1)
```

But now let's look at space and time together in a 2D GAM.
```{r}
### Pooled across species
bb_abund_2010 <- bam(bb.abund ~ te(yday, elev.mean, 
                            bs = c("tp", "tp"), k = c(10, 10)),
                       data = filter(bb_abund, year == 2010),
                       discrete = TRUE,
                       method = "fREML",
                       family = "poisson",
                       select = TRUE) %>% getViz()

check.gamViz(bb_abund_2010)
summary(bb_abund_2010)
print(plot(bb_abund_2010, allTerms = TRUE), pages = 1)

bb_abund_2011 <- bam(bb.abund ~ te(yday, elev.mean, 
                            bs = c("tp", "tp"), k = c(10, 10)),
                       data = filter(bb_abund, year == 2011),
                       discrete = TRUE,
                       method = "fREML",
                       family = "poisson",
                       select = TRUE) %>% getViz()

check.gamViz(bb_abund_2011)
summary(bb_abund_2011)
print(plot(bb_abund_2011, allTerms = TRUE), pages = 1)

bb_abund_2012 <- bam(bb.abund ~ te(yday, elev.mean, 
                            bs = c("tp", "tp"), k = c(10, 10)),
                       data = filter(bb_abund, year == 2012),
                       discrete = TRUE,
                       method = "fREML",
                       family = "poisson",
                       select = TRUE) %>% getViz()

check.gamViz(bb_abund_2012)
summary(bb_abund_2012)
print(plot(bb_abund_2012, allTerms = TRUE), pages = 1)
# 
# 
# bb_abund_2010.caste <- bam(bb.abund ~ caste + 
#                              te(yday, elev.mean, by = caste,
#                                 bs = c("tp", "tp"), k = c(10, 10)),
#                            data = filter(bb_abund_caste, year == 2010),
#                            discrete = TRUE,
#                            method = "fREML",
#                            family = "poisson",
#                            select = TRUE) %>% getViz()
#                          
# check.gamViz(bb_abund_2010.caste)
# summary(bb_abund_2010.caste)
# print(plot(bb_abund_2010.caste, allTerms = TRUE), pages = 1)
# 
# bb_abund_2011.caste <- bam(bb.abund ~ caste + 
#                              te(yday, elev.mean, by = caste,
#                                 bs = c("tp", "tp"), k = c(10, 10)),
#                            data = filter(bb_abund_caste, year == 2011),
#                            discrete = TRUE,
#                            method = "fREML",
#                            family = "poisson",
#                            select = TRUE) %>% getViz()
#                          
# check.gamViz(bb_abund_2011.caste)
# summary(bb_abund_2011.caste)
# print(plot(bb_abund_2011.caste, allTerms = TRUE), pages = 1)
# 
# 
# bb_abund_2012.caste <- bam(bb.abund ~ caste + 
#                              te(yday, elev.mean, by = caste,
#                                 bs = c("tp", "tp"), k = c(10, 10)),
#                            data = filter(bb_abund_caste, year == 2012),
#                            discrete = TRUE,
#                            method = "fREML",
#                            family = "poisson",
#                            select = TRUE) %>% getViz()
#                          
# check.gamViz(bb_abund_2012.caste)
# summary(bb_abund_2012.caste)
# print(plot(bb_abund_2012.caste, allTerms = TRUE), pages = 1)
#

### Species level
bb_abund_sp <- net %>%
  group_by(site, date, yday = dayofyear,  gdd.cum, year, elev.class2, 
           elev.mean, break1.gdd, break2.gdd, bb.sp) %>%
  summarize(bb.abund = n()) %>%
  ungroup() %>%
  mutate(bb.sp = factor(bb.sp)) %>%
  spread(bb.sp, bb.abund) %>%
  gather(bb.sp, bb.abund, -c(site, date, yday,  
                             gdd.cum, year, elev.class2, elev.mean, 
                             break1.gdd, break2.gdd)) %>%
  replace_na(list(bb.abund = 0)) %>%
  mutate(bb.sp = factor(bb.sp)) %>%
  group_by(bb.sp, year) %>%
  filter(sum(bb.abund) > 5) # remove rare species (on a per-year basis)

# bb_fl_abund_sp <- left_join(bb_abund_sp, fl_abund)
# bb_fl_abund_sp_2010 <- left_join(bb_abund_sp, fl_abund) %>% filter(year == 2010)
# bb_fl_abund_sp_2011 <- left_join(bb_abund_sp, fl_abund) %>% filter(year == 2011)
# bb_fl_abund_sp_2012 <- left_join(bb_abund_sp, fl_abund) %>% filter(year == 2012)

bb_abund_sp_2010 <- bam(bb.abund ~ bb.sp +
                                 te(yday, elev.mean, by = bb.sp, 
                                    bs = c("tp", "tp"), k = c(10, 10), m = 1),
                               data = filter(bb_abund_sp, year == 2010),
                               discrete = TRUE,
                               method = "fREML",
                               family = "poisson") %>% getViz()

check.gamViz(bb_abund_sp_2010)
summary(bb_abund_sp_2010)
print(plot(bb_abund_sp_2010, allTerms = TRUE), pages =4)

bb_abund_sp_2011 <- bam(bb.abund ~ bb.sp +
                                 te(yday, elev.mean, by = bb.sp, 
                                    bs = c("tp", "tp"), k = c(10, 10), m = 1),
                               data = filter(bb_abund_sp, year == 2011),
                               discrete = TRUE,
                               method = "fREML",
                               family = "poisson") %>% getViz()

check.gamViz(bb_abund_sp_2011)
summary(bb_abund_sp_2011)
print(plot(bb_abund_sp_2011, allTerms = TRUE), pages = 4)

bb_abund_sp_2012 <- bam(bb.abund ~ bb.sp +
                                 te(yday, elev.mean, by = bb.sp, 
                                    bs = c("tp", "tp"), k = c(10, 10), m = 1),
                               data = filter(bb_abund_sp, year == 2012),
                               discrete = TRUE,
                               method = "fREML",
                               family = "poisson") %>% getViz()

check.gamViz(bb_abund_sp_2012)
summary(bb_abund_sp_2012)
print(plot(bb_abund_sp_2012, allTerms = TRUE), pages = 4)
```

And BB richness
```{r}
bb_rich_2010 <- bam(bb.rich ~ te(yday, elev.mean, 
                            bs = c("tp", "tp"), k = c(10, 10), m = 1),
                       data = filter(bb_rich, year == 2010),
                       discrete = TRUE,
                       method = "fREML",
                       family = "poisson",
                       select = TRUE) %>% getViz()

check.gamViz(bb_rich_2010)
summary(bb_rich_2010)
print(plot(bb_rich_2010, allTerms = TRUE), pages = 1)

bb_rich_2011 <- bam(bb.rich ~ te(yday, elev.mean, 
                            bs = c("tp", "tp"), k = c(10, 10), m = 1),
                       data = filter(bb_rich, year == 2011),
                       discrete = TRUE,
                       method = "fREML",
                       family = "poisson",
                       select = TRUE) %>% getViz()

check.gamViz(bb_rich_2011)
summary(bb_rich_2011)
print(plot(bb_rich_2011, allTerms = TRUE), pages = 1)

bb_rich_2012 <- bam(bb.rich ~ te(yday, elev.mean, 
                            bs = c("tp", "tp"), k = c(10, 10), m = 1),
                       data = filter(bb_rich, year == 2012),
                       discrete = TRUE,
                       method = "fREML",
                       family = "poisson",
                       select = TRUE) %>% getViz()

check.gamViz(bb_rich_2012)
summary(bb_rich_2012)
print(plot(bb_rich_2012, allTerms = TRUE), pages = 1)

```

To what extent is the relationship between BB abundance and elev*time reflect a relationship between BB abundance and floral abundance?
```{r}
bb_fl_abund <- left_join(bb_abund, fl_abund)
bb_fl_abund_sp <- left_join(bb_abund_sp, fl_abund)
bb_fl_abund.ktype <- left_join(bb_abund, fl_abund_ktype)
bb_fl_abund_sp.ktype <- left_join(bb_abund_sp, fl_abund_ktype)
bb_fl_abund.family <- left_join(bb_abund, fl_abund_family)
bb_fl_abund_sp.family <- left_join(bb_abund_sp, fl_abund_family)
bb_fl_abund_sp.genus <- left_join(bb_abund_sp, fl_abund_genus)

ggplot(bb_fl_abund, aes(log(fl.abund), bb.abund)) +
  geom_point() +
  geom_smooth() +
  facet_wrap(~year)

ggplot(bb_fl_abund_sp, aes(log(fl.abund), bb.abund)) +
  geom_point() +
  geom_smooth() +
  facet_wrap(~bb.sp)

ggplot(bb_fl_abund.ktype, aes(log(fl.abund), bb.abund)) +
  geom_point() +
  geom_smooth() +
  facet_wrap(~k.type.ss)

ggplot(bb_fl_abund_sp.ktype, aes(log(fl.abund), bb.abund)) +
  geom_point() +
  geom_smooth() +
  facet_grid(k.type.ss~bb.sp)

ggplot(bb_fl_abund.family, aes(log(fl.abund), bb.abund, color = year)) +
  #geom_point() +
  geom_smooth(se = FALSE, method = "lm") +
  facet_wrap(~plant.family)

ggplot(filter(bb_fl_abund_sp.family,
              bb.sp %in% c("pasc", "prat", "bss", "mont", "hort", "mont", "soro", "wurf", "psit") &
                plant.family %in% c("Rosaceae", "Lamiaceae", "Asteraceae")), 
       aes(log(fl.abund), bb.abund, color = year)) +
  geom_point(alpha = 0.1) +
  geom_smooth() +
  facet_grid(bb.sp~plant.family, scales = "free") +
  theme_gray(8)

### Let's see how tightly BB abundance is correlated with FL abundance within each site*year
ggplot(filter(bb_fl_abund, year == 2012), aes(fl.abund, bb.abund)) +
  geom_point() +
  geom_smooth() +
  facet_wrap(~site, scales = "free")

ggplot(filter(bb_fl_abund, year == 2011), aes(fl.abund, bb.abund)) +
  geom_point() +
  geom_smooth() +
  facet_wrap(~site, scales = "free")
```
Hmm...so, bumble bee abundance is correlated with floral abundance, but with a ton of noise. Perhaps that is due to the temporal patterns of colony growth.
```{r}
bb_fl_abund <- left_join(bb_abund, fl_abund)

bb_abund_2010_2 <- gam(bb.abund ~ te(yday, log(fl.abund + 0.01), bs = c("tp", "tp"), k = c(10, 10)),
                     data = filter(bb_fl_abund, year == 2010),
                     method = "REML",
                     family = "gaussian") %>% getViz()

check.gamViz(bb_abund_2010_2)
summary(bb_abund_2010_2)
print(plot(bb_abund_2010_2, allTerms = TRUE), pages = 1)

bb_abund_2012_2 <- gam(bb.abund ~ te(yday, log(fl.abund + 0.01), bs = c("tp", "tp"), k = c(10, 10)),
                     data = filter(bb_fl_abund, year == 2012),
                     method = "REML",
                     family = "gaussian") %>% getViz()

check.gamViz(bb_abund_2012_2)
summary(bb_abund_2012_2)
  plot <- print(plot(bb_abund_2012_2, allTerms = TRUE), pages = 1)

```

And dare we attempt to model the ratio of FL to BB abundance?
```{r}
fl_per_bb <- bb_fl_abund %>%
  ungroup() %>%
  mutate(fl.per.bb = fl.abund/bb.abund)

fl_per_bb_2010 <- bam(log(fl.abund/bb.abund + 0.01) ~ te(yday, elev.mean, 
                                             bs = c("tp", "tp"), 
                                             k = c(10, 10)),
                      data = filter(fl_per_bb, year == 2010),
                      discrete = FALSE,
                      method = "REML",
                      #family = "poisson",
                      select = TRUE) %>% getViz()

check.gamViz(fl_per_bb_2010)
summary(fl_per_bb_2010)
print(plot(fl_per_bb_2010, allTerms = TRUE), pages = 1)

fl_per_bb_2011 <- bam(log(fl.abund/bb.abund + 0.01) ~ te(yday, elev.mean, 
                                             bs = c("tp", "tp"), 
                                             k = c(10, 10)),
                      data = filter(fl_per_bb, year == 2011),
                      discrete = FALSE,
                      method = "REML",
                      #family = "poisson",
                      select = TRUE) %>% getViz()

check.gamViz(fl_per_bb_2011)
summary(fl_per_bb_2011)
print(plot(fl_per_bb_2011, allTerms = TRUE), pages = 1)

fl_per_bb_2012 <- bam(log(fl.abund/bb.abund + 0.01) ~ te(yday, elev.mean, 
                                             bs = c("tp", "tp"), 
                                             k = c(10, 10)),
                      data = filter(fl_per_bb, year == 2012),
                      discrete = FALSE,
                      method = "REML",
                      #family = "poisson",
                      select = TRUE) %>% getViz()

check.gamViz(fl_per_bb_2012)
summary(fl_per_bb_2012)
print(plot(fl_per_bb_2012, allTerms = TRUE), pages = 1)

```

What if we split up flora by...stuff?
```{r}
# fl_abund_family <- survey %>%
#   group_by(site, date, yday = dayofyear,  gdd.cum, year, elev.class2, 
#            elev.mean, break1.gdd, break2.gdd, gdd.total, plant.family) %>%
#   summarize(fl.abund = sum(flower.cover)) %>%
#   mutate(year2010 = ifelse(year == 2010, 1, 0),
#          year2011 = ifelse(year == 2011, 1, 0),
#          year2012 = ifelse(year == 2012, 1, 0))
# 
# fl_abund_fam_2010 <- bam(log(fl.abund + 0.01) ~ plant.family +
#                            te(yday, elev.mean, by = factor(plant.family), bs = c("tp", "tp"), k = c(10, 10)),
#                          data = filter(fl_abund_family, year == 2010),
#                          discrete = TRUE,
#                          method = "fREML") %>% getViz()
#                          
# check.gamViz(fl_abund_fam_2010)
# summary(fl_abund_fam_2010)
# print(plot(fl_abund_fam_2010, allTerms = TRUE), pages = 1)

### Prepare data
# fl_abund_ktype.ss <- survey %>%
#   group_by(site, date, yday = dayofyear,  gdd.cum, year, elev.mean, k.type.ss) %>%
#   summarize(fl.abund = sum(flower.cover)) %>%
#   ungroup() %>%
#   spread(k.type.ss, fl.abund) %>%
#   gather(k.type.ss, fl.abund, -c(site, date, yday, gdd.cum, year, elev.mean))
# %>%
#   gather()
#   mutate(year2010 = ifelse(year == 2010, 1, 0),
#          year2011 = ifelse(year == 2011, 1, 0),
#          year2012 = ifelse(year == 2012, 1, 0),
#          k.type.ss = factor(k.type.ss))
# 
# fl_abund_ktype.s <- survey %>%
#   group_by(site, date, yday = dayofyear,  gdd.cum, year, elev.class2, 
#            elev.mean, break1.gdd, break2.gdd, k.type.s) %>%
#   summarize(fl.abund = sum(flower.cover)) %>%
#   ungroup() %>%
#   mutate(year2010 = ifelse(year == 2010, 1, 0),
#          year2011 = ifelse(year == 2011, 1, 0),
#          year2012 = ifelse(year == 2012, 1, 0),
#          k.type.s = factor(k.type.s))

top_plants <- net %>%
  group_by(year, bb.sp, plant.sp) %>%
  summarize(visits = n()) %>%
  group_by(year, bb.sp) %>%
  top_n(3, visits) %>%
  arrange(bb.sp, year, -visits)

ggplot(top_plants, aes(year, visits, fill = plant.sp)) +
  geom_col(position = "fill") +
  facet_wrap(~bb.sp)

fl_abund_sp <- survey %>%
  group_by(site, date, yday = dayofyear,  gdd.cum, year, 
           elev.mean, plant.sp) %>%
  summarize(fl.abund = sum(flower.cover)) %>%
  ungroup() %>%
  spread(plant.sp, fl.abund) %>%
  gather(plant.sp, fl.abund, -c(site, date, yday, gdd.cum, year, elev.mean)) %>%
  replace_na(list(fl.abund = 0)) %>%
  semi_join(top_plants, by = c("year", "plant.sp")) %>%
  mutate(year2010 = ifelse(year == 2010, 1, 0),
         year2011 = ifelse(year == 2011, 1, 0),
         year2012 = ifelse(year == 2012, 1, 0),
         plant.sp = factor(plant.sp))

### As an example, let's see how some BB species are predicted by their #1 plant
Stachys_alopecuros_2012 <- fl_abund_sp %>%
  filter(plant.sp == "Stachys alopecuros", year == 2012) %>%
  filter(fl.abund > 0)

prat_2012 <- bb_abund_sp %>% 
  filter(bb.sp == "prat", year == 2012) %>%
  left_join(Stachys_alopecuros_2012)

ggplot(prat_2012, aes(log(fl.abund), bb.abund)) +
  geom_point()

Rhododendron_hirsutum_2011 <- fl_abund_sp %>%
  filter(plant.sp == "Rhododendron hirsutum", year == 2011) %>%
  filter(fl.abund > 0)

prat_2011 <- bb_abund_sp %>% 
  filter(bb.sp == "prat", year == 2011) %>%
  left_join(Rhododendron_hirsutum_2011)

ggplot(prat_2011, aes(log(fl.abund), bb.abund)) +
  geom_point()

Campanula_scheuchzeri_2012 <-  fl_abund_sp %>%
  filter(plant.sp == "Campanula scheuchzeri", year == 2012) %>%
  filter(fl.abund > 0)

soro_2012 <- bb_abund_sp %>% 
  filter(bb.sp == "soro", year == 2012) %>%
  left_join(Campanula_scheuchzeri_2012)

ggplot(soro_2012, aes(log(fl.abund), bb.abund)) +
  geom_point()

# ### Call ktype.ss models
# fl_abund_ktype.ss_2010 <- bam(log(fl.abund + 0.01) ~ k.type.ss +
#                            te(yday, elev.mean, by = k.type.ss, bs = c("tp", "tp"), k = c(10, 10), m = 1),
#                          data = filter(fl_abund_ktype.ss, year == 2010),
#                          discrete = TRUE,
#                          method = "fREML") %>% getViz()
#                          
# check.gamViz(fl_abund_ktype.ss_2010)
# summary(fl_abund_ktype.ss_2010)
# print(plot(fl_abund_ktype.ss_2010, allTerms = TRUE), pages = 1)
# 
# fl_abund_ktype.ss_2011 <- bam(log(fl.abund + 0.01) ~ k.type.ss +
#                            te(yday, elev.mean, by = k.type.ss, bs = c("tp", "tp"), k = c(10, 10), m = 1),
#                          data = filter(fl_abund_ktype.ss, year == 2011),
#                          discrete = TRUE,
#                          method = "fREML") %>% getViz()
#                          
# check.gamViz(fl_abund_ktype.ss_2011)
# summary(fl_abund_ktype.ss_2011)
# print(plot(fl_abund_ktype.ss_2011, allTerms = TRUE), pages = 1)
# 
# fl_abund_ktype.ss_2012 <- bam(log(fl.abund + 0.01) ~ k.type.ss +
#                            te(yday, elev.mean, by = k.type.ss, bs = c("tp", "tp"), k = c(15, 15)),
#                          data = filter(fl_abund_ktype.ss, year == 2012),
#                          discrete = TRUE,
#                          method = "fREML") %>% getViz()
#                          
# check.gamViz(fl_abund_ktype.ss_2012)
# summary(fl_abund_ktype.ss_2012)
# print(plot(fl_abund_ktype.ss_2012, allTerms = TRUE), pages = 1)
# 
# ### Call ktype.s models
# fl_abund_ktype.s_2010 <- bam(log(fl.abund + 0.01) ~ k.type.s +
#                            te(yday, elev.mean, by = k.type.s, bs = c("tp", "tp"), k = c(10, 10), m = 1),
#                          data = filter(fl_abund_ktype.s, year == 2010),
#                          discrete = TRUE,
#                          method = "fREML") %>% getViz()
#                          
# check.gamViz(fl_abund_ktype.s_2010)
# summary(fl_abund_ktype.s_2010)
# print(plot(fl_abund_ktype.s_2010, allTerms = TRUE), pages = 1)
# 
# fl_abund_ktype.s_2011 <- bam(log(fl.abund + 0.01) ~ k.type.s +
#                            te(yday, elev.mean, by = k.type.s, bs = c("tp", "tp"), k = c(10, 10), m = 1),
#                          data = filter(fl_abund_ktype.s, year == 2011),
#                          discrete = TRUE,
#                          method = "fREML") %>% getViz()
#                          
# check.gamViz(fl_abund_ktype.s_2011)
# summary(fl_abund_ktype.s_2011)
# print(plot(fl_abund_ktype.s_2011, allTerms = TRUE), pages = 1)
# 
# fl_abund_ktype.s_2012 <- bam(log(fl.abund + 0.01) ~ k.type.s +
#                            te(yday, elev.mean, by = k.type.s, bs = c("tp", "tp"), k = c(15, 15)),
#                          data = filter(fl_abund_ktype.s, year == 2012),
#                          discrete = TRUE,
#                          method = "fREML") %>% getViz()
#                          
# check.gamViz(fl_abund_ktype.s_2012)
# summary(fl_abund_ktype.s_2012)
# print(plot(fl_abund_ktype.s_2012, allTerms = TRUE), pages = 1)

### Call species models
### What are the most visited species?
fl_abund_sp_2010 <- bam(log(fl.abund + 0.01) ~ plant.sp +
                           te(yday, elev.mean, by = plant.sp, bs = c("tp", "tp"), k = c(10, 10), m = 1),
                         data = filter(fl_abund_sp, year == 2010),
                         discrete = TRUE,
                         method = "fREML") %>% getViz()

check.gamViz(fl_abund_sp_2010)
summary(fl_abund_sp_2010)
print(plot(fl_abund_sp_2010, allTerms = TRUE), pages = 1)

fl_abund_sp_2011 <- bam(log(fl.abund + 0.01) ~ plant.sp +
                           te(yday, elev.mean, by = plant.sp, bs = c("tp", "tp"), k = c(10, 10), m = 1),
                         data = filter(fl_abund_sp, year == 2011),
                         discrete = TRUE,
                         method = "fREML") %>% getViz()

check.gamViz(fl_abund_sp_2011)
summary(fl_abund_sp_2011)
print(plot(fl_abund_sp_2011, allTerms = TRUE), pages = 1)

fl_abund_sp_2012 <- bam(log(fl.abund + 0.01) ~ plant.sp +
                           te(yday, elev.mean, by = plant.sp, bs = c("tp", "tp"), k = c(10, 10), m = 1),
                         data = filter(fl_abund_sp, year == 2012),
                         discrete = TRUE,
                         method = "fREML") %>% getViz()

check.gamViz(fl_abund_sp_2012)
summary(fl_abund_sp_2012)
print(plot(fl_abund_sp_2012, allTerms = TRUE), pages = 1)
```

Alright -- how does the visitation rate of each floral species differ (or not) from its surveyed abundance?
```{r}
fl_abund_net_sp <- net %>%
  group_by(site, date, yday = dayofyear, gdd.cum, year, 
           elev.mean, plant.sp) %>%
  summarize(fl.abund = n()) %>%
  ungroup() %>%
  spread(plant.sp, fl.abund) %>%
  gather(plant.sp, fl.abund, -c(site, date, yday, gdd.cum, year, elev.mean)) %>%
  replace_na(list(fl.abund = 0)) %>%
  semi_join(top_plants, by = c("year", "plant.sp")) %>%
  mutate(year2010 = ifelse(year == 2010, 1, 0),
         year2011 = ifelse(year == 2011, 1, 0),
         year2012 = ifelse(year == 2012, 1, 0),
         plant.sp = factor(plant.sp))

fl_abund_net_sp_2012 <- bam(fl.abund ~ plant.sp +
                           te(yday, elev.mean, by = plant.sp, bs = c("tp", "tp"), k = c(10, 10), m = 1),
                         data = filter(fl_abund_net_sp, year == 2012 & plant.sp %in% c("Trifolium pratense", "Origanum vulgare", "Prunella vulgaris", "Stachys alopecuros")),
                         discrete = TRUE,
                         family = "poisson",
                         method = "fREML") %>% getViz()

check.gamViz(fl_abund_net_sp_2012)
summary(fl_abund_net_sp_2012)
print(plot(fl_abund_net_sp_2012, allTerms = TRUE), pages = 1)
```

Now let's get really granular
```{r}
sp_by_sp <- net %>%
  group_by(site, date, yday = dayofyear, gdd.cum, year, 
           elev.mean, bb.sp, plant.sp) %>%
  summarize(fl.abund = n()) %>%
  ungroup() %>%
  spread(plant.sp, fl.abund) %>%
  gather(plant.sp, fl.abund, -c(site, date, yday, gdd.cum, year, elev.mean, bb.sp)) %>%
  replace_na(list(fl.abund = 0)) %>%
  semi_join(top_plants, by = c("year", "plant.sp")) %>%
  mutate(year2010 = ifelse(year == 2010, 1, 0),
         year2011 = ifelse(year == 2011, 1, 0),
         year2012 = ifelse(year == 2012, 1, 0),
         plant.sp = factor(plant.sp))

Stachys_alopecuros_2012 <-  bam(fl.abund ~ bb.sp +
                           te(yday, elev.mean, by = factor(bb.sp), bs = c("tp", "tp"), k = c(10, 10), m = 1),
                         data = filter(sp_by_sp, 
                                       year == 2012 & 
                                         bb.sp %in% c("prat", "pasc", "bss", "soro", "mont", "wurf") &
                                         plant.sp == "Stachys alopecuros"),
                         discrete = TRUE,
                         family = "poisson",
                         method = "fREML") %>% getViz()

check.gamViz(Stachys_alopecuros_2012)
summary(Stachys_alopecuros_2012)
print(plot(Stachys_alopecuros_2012, allTerms = TRUE), pages = 1)

Trifolium_pratense_2011 <-  bam(fl.abund ~ bb.sp +
                           te(yday, elev.mean, by = factor(bb.sp), bs = c("tp", "tp"), k = c(10, 10), m = 1),
                         data = filter(sp_by_sp, 
                                       year == 2011 & 
                                         bb.sp %in% c("prat", "pasc", "bss", "soro", "wurf") &
                                         plant.sp == "Trifolium pratense"),
                         discrete = TRUE,
                         family = "poisson",
                         method = "fREML") %>% getViz()

check.gamViz(Trifolium_pratense_2011)
summary(Trifolium_pratense_2011)
print(plot(Trifolium_pratense_2011, allTerms = TRUE), pages = 1)
```

### GAM surface plots of abundance and diversity metrics
#### After puzzling over whether to model the three years together, using year as a "by" factor, or to set up an independent model for each year, I've settled (for now) on the latter. Since the yday range of each year is determined largely by when the snow melted, I do not want to extrapolate beyond my data.
##### Floral abundance
```{r eval=FALSE, include=FALSE}
### yday ranges
tapply(fl_abund$yday, fl_abund$year, summary)
tapply(fl_abund$gdd.cum, fl_abund$year, summary)

fl_abund_yday_2010 <- bam(log(fl.abund + 0.001) ~ te(yday, elev.mean, bs= c("tp", "tp"), k=c(12, 12), m=1),
                    data = filter(fl_abund, year == "2010"),
                    discrete = TRUE,
                    method = "fREML") %>% getViz()

check.gamViz(fl_abund_yday_2010)
summary(fl_abund_yday_2010)
print(plot(fl_abund_yday_2010, allTerms = TRUE), pages = 1)

fl_abund_yday_2011 <- bam(log(fl.abund + 0.001) ~ te(yday, elev.mean, bs= c("tp", "tp"), k=c(12, 12), m=1),
                    data = filter(fl_abund, year == "2011"),
                    discrete = TRUE,
                    method = "fREML") %>% getViz()

check.gamViz(fl_abund_yday_2011)
summary(fl_abund_yday_2011)
print(plot(fl_abund_yday_2011, allTerms = TRUE), pages = 1)

fl_abund_yday_2012 <- bam(log(fl.abund + 0.001) ~ te(yday, elev.mean, bs= c("tp", "tp"), k=c(12, 12), m=1),
                    data = filter(fl_abund, year == "2012"),
                    discrete = TRUE,
                    method = "fREML") %>% getViz()

check.gamViz(fl_abund_yday_2012)
summary(fl_abund_yday_2012)
print(plot(fl_abund_yday_2012, allTerms = TRUE), pages = 1)
plot(predict(fl_abund_yday_2012))

### HGAM as per Gavin's recommendation, but based on GDD instead of yday
fl_abund_gdd <- bam(log(fl.abund + 0.001) ~ year +
                      te(gdd.cum, elev.mean, k=c(10, 10)) +
                      te(gdd.cum, elev.mean, by = year2010, bs= c("tp", "tp"), k=c(12, 12), m=1) +
                      te(gdd.cum, elev.mean, by = year2011, bs= c("tp", "tp"), k=c(12, 12), m=1) +
                      te(gdd.cum, elev.mean, by = year2012, bs= c("tp", "tp"), k=c(12, 12), m=1),
                    data = fl_abund,
                    discrete = TRUE,
                    method = "fREML") %>% getViz()

fl_abund_yday <- bam(log(fl.abund + 0.001) ~ year +
                      te(yday, elev.mean, by = year, bs= c("tp", "tp"), k=c(15, 15), m=1),
                    data = fl_abund,
                    discrete = TRUE,
                    method = "fREML") %>% getViz()

fl_abund_yday <- bam(log(fl.abund + 0.001) ~ year +
                      te(yday, elev.mean, by = year2010, bs= c("tp", "tp"), k=c(15, 15), m=1) +
                       te(yday, elev.mean, by = year2011, bs= c("tp", "tp"), k=c(15, 15), m=1) +
                       te(yday, elev.mean, by = year2012, bs= c("tp", "tp"), k=c(15, 15), m=1),
                    data = fl_abund,
                    discrete = TRUE,
                    method = "fREML") %>% getViz()

check.gamViz(fl_abund_yday)
summary(fl_abund_yday)
print(plot(fl_abund_yday, allTerms = TRUE), pages = 1)
```

### Let's refine our plotting with tidymv
```{r, eval=FALSE, include=FALSE}
p2010 <- predict_gam(fl_abund_yday_2010)
p2011 <- predict_gam(fl_abund_yday_2011)
p2012 <- predict_gam(fl_abund_yday_2012)
p <- predict_gam(fl_abund_gdd)

ggplot(p2010, aes(yday, elev.mean, fit)) +
  geom_raster(aes(fill = exp(fit))) +
  #geom_contour(aes(z = fit), color = "black") +
  geom_point(data = filter(fl_abund, year == 2010), 
             aes(yday, elev.mean, size = fl.abund),
             color = "white", alpha = 0.25) +
  scale_fill_viridis_c()

ggplot(p2011, aes(yday, elev.mean, fit)) +
  geom_raster(aes(fill = exp(fit))) +
  #geom_contour(aes(z = fit), color = "black") +
  geom_point(data = filter(fl_abund, year == 2011), 
             aes(yday, elev.mean, size = fl.abund),
             color = "white", alpha = 0.25) +
  scale_fill_viridis_c()

ggplot(p2012, aes(yday, elev.mean, fit)) +
  geom_raster(aes(fill = exp(fit))) +
  #geom_contour(aes(z = fit), color = "black") +
  geom_point(data = filter(fl_abund, year == 2012), 
             aes(yday, elev.mean, size = fl.abund),
             color = "white", alpha = 0.25) +
  scale_fill_viridis_c()

ggplot(p, aes(yday, elev.mean, fit)) +
  geom_raster(aes(fill = exp(fit))) +
  geom_contour(aes(z = fit), color = "black") +
  geom_point(data = fl_abund,
  aes(yday, elev.mean, size = fl.abund),
  color = "white", alpha = 0.25) +
  scale_fill_viridis_c() +
  facet_wrap(~year)

```

##### Floral richness 
```{r eval=FALSE, include=FALSE}
fl_rich_yday_2010 <- bam(fl.rich ~ te(yday, elev.mean, bs= c("tp", "tp"), k=c(10, 10), m=1),
                    data = filter(fl_rich, year == "2010"),
                    discrete = TRUE,
                    family = "poisson",
                    method = "fREML") %>% getViz()

check.gamViz(fl_rich_yday_2010)
summary(fl_rich_yday_2010)
print(plot(fl_rich_yday_2010, allTerms = TRUE), pages = 1)

fl_rich_yday_2011 <- bam(fl.rich ~ te(yday, elev.mean, bs= c("tp", "tp"), k=c(10, 10), m=1),
                    data = filter(fl_rich, year == "2011"),
                    discrete = TRUE,
                    family = "poisson",
                    method = "fREML") %>% getViz()

check.gamViz(fl_rich_yday_2011)
summary(fl_rich_yday_2011)
print(plot(fl_rich_yday_2011, allTerms = TRUE), pages = 1)

fl_rich_yday_2012 <- bam(fl.rich ~ te(yday, elev.mean, bs= c("tp", "tp"), k=c(10, 10), m=1),
                    data = filter(fl_rich, year == "2012"),
                    discrete = TRUE,
                    family = "poisson",
                    method = "fREML") %>% getViz()

check.gamViz(fl_rich_yday_2012)
summary(fl_rich_yday_2012)
print(plot(fl_rich_yday_2012, allTerms = TRUE), pages = 1)
```

##### BB abundance
```{r eval=FALSE, include=FALSE}
bb_abund_yday_2012 <- bam(bb.abund ~ te(yday, elev.mean, bs= c("tp", "tp"), k=c(10, 10), m=1),
                    data = filter(bb_abund, year == "2012"),
                    family = "poisson",
                    discrete = TRUE,
                    method = "fREML") %>% getViz()

check.gamViz(bb_abund_yday_2012)
summary(bb_abund_yday_2012)
print(plot(bb_abund_yday_2012, allTerms = TRUE), pages = 1)

bb_abund_yday_2011 <- bam(bb.abund ~ te(yday, elev.mean, bs= c("tp", "tp"), k=c(10, 10), m=1),
                    data = filter(bb_abund, year == "2011"),
                    family = "poisson",
                    discrete = TRUE,
                    method = "fREML") %>% getViz()

check.gamViz(bb_abund_yday_2011)
summary(bb_abund_yday_2011)
print(plot(bb_abund_yday_2011, allTerms = TRUE), pages = 1)

bb_abund_yday_2010 <- bam(bb.abund ~ te(yday, elev.mean, bs= c("tp", "tp"), k=c(10, 10), m=1),
                    data = filter(bb_abund, year == "2010"),
                    family = "poisson",
                    discrete = TRUE,
                    method = "fREML") %>% getViz()

check.gamViz(bb_abund_yday_2010)
summary(bb_abund_yday_2010)
print(plot(bb_abund_yday_2010, allTerms = TRUE), pages = 1)
```

##### BB abundance with species as a "by" factor smooth
##### I am using family = "poisson"; this results in somewhat overdispersed models, but with a higher variance explained than when I use quasipoisson or negative binomial distributions, and the latter two distributions seem to introduce skew. I am more concerned with maximizing the explanatory power of my models than I am with having accurate p-values. We'll see what reviewers think, I guess. 
```{r eval=FALSE, include=FALSE}
### Species-level abundance
bb_abund_sp <- net %>%
  group_by(site, date, dayofyear, gdd.cum, year, elev.class2, 
           elev.mean, break1.gdd, break2.gdd, bb.sp) %>%
  summarize(bb.abund = n()) %>%
  rename(yday = dayofyear) %>%
  ungroup() %>%
  filter(bb.sp %in% c("bss", "hort", "pasc", "prat", "soro", "wurf")) %>% # we only want the common species I think
  mutate(bb.sp = factor(bb.sp))

### 2010
bb_abund_sp_yday_2010 <- bam(bb.abund ~ te(yday, elev.mean, by= bb.sp, bs= c("tp", "tp"), k=c(12, 12), m=2),
                             data = filter(bb_abund_sp, year == "2010"),
                             family = "poisson",
                             discrete = TRUE,
                             method = "fREML") %>% getViz()

check.gamViz(bb_abund_sp_yday_2010)
summary(bb_abund_sp_yday_2010)

### 2011
bb_abund_sp_yday_2011 <- bam(bb.abund ~ te(yday, elev.mean, by= bb.sp, bs= c("tp", "tp"), k=c(12, 12), m=2),
                             data = filter(bb_abund_sp, year == "2011"),
                             family = "poisson",
                             discrete = TRUE,
                             method = "fREML") %>% getViz()

check.gamViz(bb_abund_sp_yday_2011)
summary(bb_abund_sp_yday_2011)

### 2012
bb_abund_sp_yday_2012 <- bam(bb.abund ~ te(yday, elev.mean, by= bb.sp, bs= c("tp", "tp"), k=c(15, 15), m=2),
                             data = filter(bb_abund_sp, year == "2012"),
                             family = "poisson",
                             discrete = TRUE,
                             method = "fREML") %>% getViz()

check.gamViz(bb_abund_sp_yday_2012)
summary(bb_abund_sp_yday_2012)

### Global HGAM
bb_abund_sp_yday <- bam(bb.abund ~ year +
                          te(elev.mean, yday, bs= c("tp", "tp"), k=c(10, 10), m=2) +
                          te(yday, elev.mean, by= bb.sp, bs= c("tp", "tp"), k=c(15, 15), m=1),
                             data = bb_abund_sp,
                             family = "nb",
                             discrete = TRUE,
                             method = "fREML") %>% getViz()

check.gamViz(bb_abund_sp_yday)
summary(bb_abund_sp_yday)

### Extract fits
bb2010 <- predict_gam(bb_abund_sp_yday_2010)
bb2011 <- predict_gam(bb_abund_sp_yday_2011)
bb2012 <- predict_gam(bb_abund_sp_yday_2012)
bb <- predict_gam(bb_abund_sp_yday)

### Plot
ggplot(bb2010, aes(yday, elev.mean, fit)) +
  geom_raster(aes(fill = fit)) +
  geom_contour(aes(z = fit), color = "black") +
  geom_point(data = filter(bb_abund_sp, year == 2010), 
             aes(yday, elev.mean, size = bb.abund),
             color = "white", alpha = 0.5) +
  scale_fill_viridis_c() +
  facet_wrap(~bb.sp)

ggplot(bb2011, aes(yday, elev.mean, fit)) +
  geom_raster(aes(fill = fit)) +
  geom_contour(aes(z = fit), color = "black") +
  geom_point(data = filter(bb_abund_sp, year == 2011), 
             aes(yday, elev.mean, size = bb.abund),
             color = "white", alpha = 0.5) +
  scale_fill_viridis_c() +
  facet_wrap(~bb.sp)

ggplot(bb2012, aes(yday, elev.mean, fit)) +
  geom_raster(aes(fill = fit)) +
  geom_contour(aes(z = fit), color = "black") +
  geom_point(data = filter(bb_abund_sp, year == 2012), 
             aes(yday, elev.mean, size = bb.abund),
             color = "white", alpha = 0.5) +
  scale_fill_viridis_c() +
  facet_wrap(~bb.sp)
```

##### FL abundance by k.type
##### Need to go back through to simplify, include only ktypes that are used heavily
```{r, eval=FALSE, include=FALSE}
k.type_key <- read_delim("../processed_data/kugler_key.tsv", delim = "\t", col_names = FALSE)

fl_abund_ktype_yday_2010 <- bam(log(fl.abund + 0.001) ~ te(yday, elev.mean, 
                                                           by= k.type.ss, 
                                                           bs= c("tp", "tp"), 
                                                           k=c(12, 12), m=1),
                    data = filter(fl_abund_ktype, year == "2010"),
                    discrete = TRUE,
                    method = "fREML") %>% getViz()

check.gamViz(fl_abund_ktype_yday_2010)
summary(fl_abund_ktype_yday_2010)
print(plot(fl_abund_ktype_yday_2010, allTerms = TRUE), pages = 10)

fl_abund_ktype_yday_2011 <- bam(log(fl.abund + 0.001) ~ te(yday, elev.mean, 
                                                           by= k.type.ss, 
                                                           bs= c("tp", "tp"), 
                                                           k=c(12, 12), m=1),
                    data = filter(fl_abund_ktype, year == "2011"),
                    discrete = TRUE,
                    method = "fREML") %>% getViz()

check.gamViz(fl_abund_ktype_yday_2011)
summary(fl_abund_ktype_yday_2011)
print(plot(fl_abund_ktype_yday_2011, allTerms = TRUE), pages = 10)

fl_abund_ktype_yday_2012 <- bam(log(fl.abund + 0.001) ~ te(yday, elev.mean, 
                                                           by= k.type.ss, 
                                                           bs= c("tp", "tp"), 
                                                           k=c(12, 12), m=1),
                    data = filter(fl_abund_ktype, year == "2012"),
                    discrete = TRUE,
                    method = "fREML") %>% getViz()

check.gamViz(fl_abund_ktype_yday_2012)
summary(fl_abund_ktype_yday_2012)
print(plot(fl_abund_ktype_yday_2012, allTerms = TRUE), pages = 10)
```

##### FL abundance by genus
##### Need to go back through to simplify, include only ktypes that are used heavily
```{r, eval=FALSE, include=FALSE}
fl_abund_genus <- survey %>%
  group_by(site, date, yday = dayofyear, gdd.cum, year, elev.class2, 
           elev.mean, break1.gdd, break2.gdd, plant.genus) %>%
  summarize(fl.abund = sum(flower.cover)) %>%
  ungroup() %>%
  mutate(plant.genus = factor(plant.genus))

fl_abund_genus_yday_2010 <- bam(log(fl.abund + 0.001) ~ te(yday, elev.mean, 
                                                           by= plant.genus, 
                                                           bs= c("tp", "tp"), 
                                                           k=c(8, 8), m=1),
                    data = filter(fl_abund_genus, year == "2010"),
                    discrete = TRUE,
                    method = "fREML") %>% getViz()

check.gamViz(fl_abund_genus_yday_2010)
summary(fl_abund_genus_yday_2010)
print(plot(fl_abund_genus_yday_2010, allTerms = TRUE), pages = 10)

```

##### BB richness
```{r eval=FALSE, include=FALSE}
bb_rich_yday_2010 <- gamV(bb.rich ~ te(yday, elev.mean, bs= c("tp", "tp"), k=c(10, 10), m=1),
                    data = filter(bb_rich, year == "2010"),
                    family = "poisson",
                    method = "REML")

check.gamViz(bb_rich_yday_2010)
summary(bb_rich_yday_2010)
print(plot(bb_rich_yday_2010, allTerms = TRUE), pages = 1)

bb_rich_yday_2011 <- gamV(bb.rich ~ te(yday, elev.mean, bs= c("tp", "tp"), k=c(10, 10), m=1),
                    data = filter(bb_rich, year == "2011"),
                    family = "poisson",
                    method = "REML")

check.gamViz(bb_rich_yday_2011)
summary(bb_rich_yday_2011)
print(plot(bb_rich_yday_2011, allTerms = TRUE), pages = 1)

bb_rich_yday_2012 <- gamV(bb.rich ~ te(yday, elev.mean, bs= c("tp", "tp"), k=c(10, 10), m=1),
                    data = filter(bb_rich, year == "2012"),
                    family = "poisson",
                    method = "REML")

check.gamViz(bb_rich_yday_2012)
summary(bb_rich_yday_2012)
print(plot(bb_rich_yday_2012, allTerms = TRUE), pages = 1)

```

##### FL.per.BB
```{r, eval=FALSE, include=FALSE}
fl.per.bb_yday <- gamV(log(fl.per.bb) ~ year +
                    te(yday, elev.mean, by=year, bs= c("tp", "tp"), k=c(10, 10), m=1),
                    data = abund_div,
                    method = "REML")

check.gamViz(fl.per.bb_yday)
summary(fl.per.bb_yday)
print(plot(fl.per.bb_yday, allTerms = TRUE), pages = 1)

fl.per.bb_gdd <- gamV(log(fl.per.bb) ~ year +
                    te(gdd.cum, elev.mean, by=year, bs= c("tp", "tp"), k=c(10, 10), m=1),
                    data = abund_div,
                    method = "REML")

check.gamViz(fl.per.bb_gdd)
summary(fl.per.bb_gdd)
print(plot(fl.per.bb_gdd, allTerms = TRUE), pages = 1)
```