### 2020-07-08 analyses/BB_abund_GAM.Rmd
update

---
title: "preference"
output: html_document
---
### Prepare environment
```{r setup, include=FALSE, echo=FALSE}
library(mgcv)
library(mgcViz)
library(tidyverse)
library(lubridate)
```

### Climate data
```{r, include=FALSE, echo=FALSE}
gdd <- read_csv("../processed_data/climate.csv") %>%
  mutate(year = factor(year)) %>%
  dplyr::select(site, year, date, gdd.cum) %>%
  semi_join(read_csv("../processed_data/network.csv"), by = "site")
```

### Load site data
```{r, include=FALSE, echo=FALSE}
site_data <- read_csv("../processed_data/site_data.csv") %>%
  mutate(elev.class = factor(elev.class, levels = c("oben", "mitte", "unten")),
         elev.class2 = factor(elev.class2, levels = c("high", "mid", "low")))
```

### Load network data
```{r}
### BB-FL visitation data
net <- read_csv("../processed_data/network.csv") %>%
  dplyr::select(site, date, bb.sp, plant.sp)
```

### FL and BB abundance / diversity
```{r echo=FALSE, include=FALSE}
### Per-transect BB abundance: species-level
bb_abund_sp <- net %>%
  dplyr::select(site, date, bb.sp) %>%
  group_by(site, date, bb.sp) %>% # group by bb.sp*transect (and other variables we want to retain)
  summarize(bb.abund = n()) %>% # get bb.sp*transect abundance
  ungroup %>%
  complete(site, date, bb.sp, fill = list(bb.abund = 0)) %>% # complete implicit absences
  mutate(year = factor(year(date))) %>%
  left_join(site_data) %>%
  mutate(yday = yday(date),
         bb.sp = factor(bb.sp),
         management = factor(management)) %>%
  semi_join(net, by = c("site", "date")) %>%
  group_by(year, bb.sp) %>%
  mutate(bb.total.abund = sum(bb.abund))
```

GAM modeling of abundance per bb species
```{r}
### Call models
bb_abund.2010 <- bam(bb.abund ~ s(bb.sp, bs = "re") + management + slope.est + te(yday, elev.mean, 
                                                            by = bb.sp,
                                                            bs= c("tp", "tp"), 
                                                            k=c(10, 10), m=1),
                        data = filter(bb_abund_sp, year == 2010 & bb.total.abund >= 5),
                        discrete = TRUE,
                        family = "poisson",
                        select = TRUE,
                        method = "fREML") %>% getViz()

check.gamViz(bb_abund.2010)
bb_abund.2010.sum <- summary(bb_abund.2010)
print(plot(bb_abund.2010, allTerms = TRUE), pages = 4)

bb_abund.2011 <- bam(bb.abund ~ s(bb.sp, bs = "re") + management + slope.est + te(yday, elev.mean, 
                                                            by = bb.sp,
                                                            bs= c("tp", "tp"), 
                                                            k=c(10, 10), m=1),
                        data = filter(bb_abund_sp, year == 2011 & bb.total.abund >= 5),
                        discrete = TRUE,
                        family = "poisson",
                        select = TRUE,
                        method = "fREML") %>% getViz()

check.gamViz(bb_abund.2011)
bb_abund.2011.sum <- summary(bb_abund.2011)
print(plot(bb_abund.2011 , allTerms = TRUE), pages = 4)

bb_abund.2012 <- bam(bb.abund ~ s(bb.sp, bs = "re") + management + slope.est + te(yday, elev.mean, 
                                                            by = bb.sp,
                                                            bs= c("tp", "tp"), 
                                                            k=c(10, 10), m=1),
                        data = filter(bb_abund_sp, year == 2012 & bb.total.abund >= 5),
                        discrete = TRUE,
                        family = "poisson",
                        select = TRUE,
                        method = "fREML") %>% getViz()

check.gamViz(bb_abund.2012)
bb_abund.2012.sum <- summary(bb_abund.2012)
print(plot(bb_abund.2012 , allTerms = TRUE), pages = 4)
plot(bb_abund.2012)

```

```{r}
bb_abund.2012.pred <- predict(bb_abund.2012)

ggplot(bb_abund.2012.pred, aes(yday, elev.mean, fit)) +
  geom_raster(aes(fill = exp(fit))) +
  #geom_contour(aes(z = fit), color = "black") +
  geom_point(data = filter(fl_abund, year == 2012), 
             aes(yday, elev.mean, size = fl.abund),
             color = "white", alpha = 0.25) +
  scale_fill_viridis_c()
```