### 2020-07-16 analyses/FL_abund_GAM_sp.Rmd
preparing to try SEM

---
title: "FL_abund_GAM_sp"
output: html_document
---

### Prepare environment
```{r setup, include=FALSE, echo=FALSE}
library(mgcv)
library(mgcViz)
library(tidyverse)
library(lubridate)
```

### Climate data
```{r, include=FALSE, echo=FALSE}
gdd <- read_csv("../processed_data/climate.csv") %>%
  mutate(year = factor(year)) %>%
  dplyr::select(site, year, date, gdd.cum) %>%
  semi_join(read_csv("../processed_data/network.csv"), by = "site")
```

### Load site data
```{r, include=FALSE, echo=FALSE}
site_data <- read_csv("../processed_data/site_data.csv") %>%
  #dplyr::select(-c(slope.calc, slope.est, elev.min, elev.max, temp.mean)) %>% # drop these variables
  mutate(elev.class = factor(elev.class, levels = c("oben", "mitte", "unten")),
         elev.class2 = factor(elev.class2, levels = c("high", "mid", "low")))
```

### Floral survey data
```{r}
survey <- read_csv("../processed_data/floral_survey.csv") %>%
  dplyr::select(site, date, plant.sp, flower.cover) %>%
  group_by(site, date, plant.sp) %>%
  summarize(flower.cover = sum(flower.cover)) %>%
  mutate(yday = yday(date),
       year = year(date)) %>%
  left_join(site_data) %>%
  ungroup() %>%
  mutate(site = factor(site),
         plant.sp = factor(plant.sp),
         year = factor(year),
         management = factor(management))

preference <- read_csv("../analyses/output/preference.csv")

top_plants <- net %>%
  mutate(year = year(date)) %>%
  group_by(bb.sp, year, plant.sp) %>%
  summarize(visits = n()) %>%
  mutate(prop.visits = visits/sum(visits)) %>%
  group_by(bb.sp, year) %>%
  arrange(bb.sp, year, -prop.visits) %>%
  top_n(3)

top_3 <- read_csv("../analyses/output/visitation_by_abundance_mean.csv") %>%
  group_by(bb.sp) %>%
  arrange(-prop.visits.wm) %>%
  top_n(3)

```

```{r}
ggplot(filter(survey, plant.sp == "Trifolium pratense"), aes(yday, elev.mean, size = flower.cover)) +
  geom_point(alpha = 0.25) +
  facet_wrap(~year, scales = "free")

ggplot(filter(survey, plant.sp == "Erica carnea"), aes(yday, elev.mean, size = flower.cover)) +
  geom_point(alpha = 0.25) +
  facet_wrap(~year, scales = "free")

ggplot(filter(survey, plant.sp == "Stachys alopecuros"), aes(yday, elev.mean, size = flower.cover)) +
  geom_point(alpha = 0.25) +
  facet_wrap(~year, scales = "free")

t.pratense_GAM <- bam(flower.cover ~ s(management, bs = "re") + year + 
                                  s(site, bs = "re") +
                                  te(yday, elev.mean, 
                                     by = year,
                                     bs= c("tp", "tp"), 
                                     k=c(10, 10), m=1),
                                data = filter(survey, plant.sp == "Trifolium pratense"),
                                discrete = TRUE,
                                family = "tw",
                                select = TRUE,
                                method = "fREML") %>% getViz()

check.gamViz(t.pratense_GAM)
t.pratense_GAM.sum <- summary(t.pratense_GAM)
print(plot(t.pratense_GAM, allTerms = TRUE), pages = 1)

e.carnea_GAM <- bam(flower.cover ~ s(management, bs = "re") + year + 
                                  s(site, bs = "re") +
                                  te(yday, elev.mean, 
                                     by = year,
                                     bs= c("tp", "tp"), 
                                     k=c(3, 3), m=1),
                                data = filter(survey, plant.sp == "Erica carnea"),
                                discrete = TRUE,
                                family = "tw",
                                select = TRUE,
                                method = "fREML") %>% getViz()

check.gamViz(e.carnea_GAM)
e.carnea_GAM.sum <- summary(e.carnea_GAM)
print(plot(e.carnea_GAM, allTerms = TRUE), pages = 1)
```