### 2020-07-16 analyses/aggregate_network_analysis.Rmd

---
title: "aggregate_network_analysis"
output: html_document
---

### Prepare environment
```{r setup, include=FALSE}
library(bipartite)
library(tidyverse)
library(lubridate)
```

### Load network and survey data
```{r, include=FALSE}
net <- read_csv("../processed_data/network.csv") %>%
  unite(site.year, site, year, sep = "_", remove = FALSE) %>% # create site.year field that can later be split back up
  mutate(year = factor(year)) %>% # convert year to factor
  ungroup()

web <- net %>%
  group_by(bb.sp, plant.sp, year) %>%
  summarize(freq = n()) %>%
  select(higher = bb.sp, lower = plant.sp, freq, webID = year) %>%
  as.data.frame() %>%
  frame2webs()

web2010 <- net %>%
  filter(year == 2010) %>%
  group_by(bb.sp, plant.sp, year) %>%
  summarize(freq = n()) %>%
  select(higher = bb.sp, lower = plant.sp, freq, webID = year) %>%
  as.data.frame() %>%
  frame2webs()

network_metric <- web %>%
  map(networklevel, level = "higher", weighted = TRUE) %>% 
  data.frame() %>%
  t() %>%
  data.frame() %>%
  rownames_to_column() %>%
  dplyr::select(site_year = rowname, everything()) %>%
  as_tibble()

modules2010 <- web2010 %>%
  map(computeModules)

web2010 <- web[[1]]
web2011 <- web[[2]]
web2012 <- web[[3]]

plotweb(web2010)
plotweb(web2011)
plotweb(web2012)

visweb(web2010)
visweb(web2011)
visweb(web2012)
```