### 2020-07-16 analyses/beta_diversity.Rmd

---
title: "Beta Diversity through Time and Space"
output: html_document
---

### Prepare environment
```{r message=FALSE, warning=FALSE}
library(vegan)
library(lubridate)
library(gdm)
library(bipartite)
library(sf)
library(mgcv)
library(mgcViz)
library(tidyverse)
```

### Climate data
```{r, include=FALSE, echo=FALSE}
gdd <- read_csv("../processed_data/climate.csv") %>%
  mutate(year = factor(year),
         yday = yday(date)) %>%
  dplyr::select(site, year, date, yday, gdd.cum) %>%
  semi_join(read_csv("../processed_data/network.csv"), by = "site")
```

### Load site data
```{r, include=FALSE, echo=FALSE}
site_data <- read_csv("../processed_data/site_data.csv") %>%
  dplyr::select(-c(slope.calc, slope.est, elev.min, elev.max, temp.mean)) %>% # drop these variables
  mutate(elev.class = factor(elev.class, levels = c("oben", "mitte", "unten")),
         elev.class2 = factor(elev.class2, levels = c("high", "mid", "low"))) %>% # turn elav.class into an ordered factor
  semi_join(read_csv("../processed_data/network.csv"), by = "site")
```

### Load network and survey data and generate site.year and site.month.year webs
```{r}
net <- read_csv("../processed_data/network.csv")%>%
  unite(site.year, site, year, sep = "_", remove = FALSE) %>% # create site.year field that can later be split back up
  unite(site.year.day, site, year, dayofyear, sep = "_", remove = FALSE) %>%
  mutate(year = factor(year)) %>% # convert year to factor
  group_by(site, date) %>%
  arrange(site, date)

survey <- read_csv("../processed_data/floral_survey.csv") %>%
  semi_join(net, by = "plant.sp") %>% # Consider only plants visited (anytime anywhere, not specifically for a given site*date)
  dplyr::select(site, date, plant.sp, flower.cover)

site.year_transects <- net %>%
  group_by(site, year) %>%
  summarize(transects = length(unique(date)))

site_transects <- net %>%
  group_by(site) %>%
  summarize(transects = length(unique(date)))

web_site.year <- net %>%
  group_by(site.year, bb.sp, plant.sp) %>%
  summarize(freq = n()) %>%
  dplyr::select(higher = bb.sp, lower = plant.sp, webID = site.year, freq) %>% # rename columns to match bipartite's expectations
  data.frame() %>% # convert to data frame
  frame2webs() # convert to to bipartite web

web_site <- net %>%
  group_by(site, bb.sp, plant.sp) %>%
  summarize(freq = n()) %>% 
  dplyr::select(higher = bb.sp, lower = plant.sp, webID = site, freq) %>% # rename columns to match bipartite's expectations
  data.frame() %>% # convert to data frame
  frame2webs() # convert to to bipartite web

web_site.date.2010 <- net %>%
  filter(year == 2010) %>%
  group_by(site.year.day, bb.sp, plant.sp) %>%
  summarize(freq = n()) %>%
  dplyr::select(higher = bb.sp, lower = plant.sp, webID = site.year.day, freq) %>% # rename columns to match bipartite's expectations
  data.frame() %>% # convert to data frame
  frame2webs() # convert to to bipartite web

web_site.date.2011 <- net %>%
  filter(year == 2011) %>%
  group_by(site.year.day, bb.sp, plant.sp) %>%
  summarize(freq = n()) %>%
  dplyr::select(higher = bb.sp, lower = plant.sp, webID = site.year.day, freq) %>% # rename columns to match bipartite's expectations
  data.frame() %>% # convert to data frame
  frame2webs() # convert to to bipartite web

web_site.date.2012 <- net %>%
  filter(year == 2012) %>%
  group_by(site.year.day, bb.sp, plant.sp) %>%
  summarize(freq = n()) %>%
  dplyr::select(higher = bb.sp, lower = plant.sp, webID = site.year.day, freq) %>% # rename columns to match bipartite's expectations
  data.frame() %>% # convert to data frame
  frame2webs() # convert to to bipartite web

array_site <- webs2array(web_site)
array_site.year <- webs2array(web_site.year)
array_site.date.2010 <- webs2array(web_site.date.2010)
array_site.date.2011 <- webs2array(web_site.date.2011)
array_site.date.2012 <- webs2array(web_site.date.2012)
```

### Tabulate webs
S	= beta_S, the dissimilarity in species composition
OS = beta_OS, the dissimilarity (component) explained by "rewiring" among shared species
WN = beta_WN, the dissimilarity between the two networks
ST = beta_ST, the dissimilarity (component) explained by difference in species community composition 
```{r}
beta_site <- betalinkr_multi(array_site, partition.st = TRUE, partition.rr = TRUE) %>%
  as_tibble() %>%
  rename(site1 = i) %>%
  rename(site2 = j)%>%
  left_join(site_data, by = c("site1" = "site")) %>%
  dplyr::select(site1, site2, S, OS, WN, ST, 
         ST.l, ST.h, ST.lh, WN.repl, OS.repl, WN.rich, OS.rich, 
         elev1 = elev.mean, man1 = management, lon1 = lon, lat1 = lat) %>%
  left_join(site_data, by = c("site2" = "site")) %>%
  dplyr::select(site1, site2, S, OS, WN, ST, 
         ST.l, ST.h, ST.lh, WN.repl, OS.repl, WN.rich, OS.rich, 
         elev1, elev2 = elev.mean, man1, man2 = management, lon1, lat1, lon2 = lon, lat2 = lat) %>%
  left_join(site_transects, by = c("site1" = "site")) %>%
  dplyr::select(site1, site2, S, OS, WN, ST, 
                ST.l, ST.h, ST.lh, WN.repl, OS.repl, WN.rich, OS.rich, 
                elev1, elev2, man1, man2, lon1, lat1, lon2, lat2, transects1 = transects) %>%
  left_join(site_transects, by = c("site2" = "site")) %>%
  dplyr::select(site1, site2, S, OS, WN, ST, 
                ST.l, ST.h, ST.lh, WN.repl, OS.repl, WN.rich, OS.rich, 
                elev1, elev2, man1, man2, lon1, lat1, lon2, lat2, transects1, transects2 = transects) %>%
  mutate(elev.diff = abs(elev1 - elev2),
         mean.elev = (elev1 + elev2)/2,
         transects.diff = abs(transects1 - transects2),
         man.diff = case_when(
           man1 == man2 ~ FALSE,
           man1 != man2 ~ TRUE
         )) %>%
  gather(metric, value, -c(site1, site2, elev1, elev2, elev.diff, 
                           man1, man2, man.diff, mean.elev, lon1, 
                           lat1, lon2, lat2, transects1, transects2, transects.diff)) %>%
  mutate(metric.class = case_when(
    metric %in% c("WN", "ST", "OS", "S") ~ "aggregate",
    metric %in% c("WN.repl", "WN.rich") ~ "WN.partition",
    metric %in% c("ST.h", "ST.l", "ST.lh") ~ "ST.partition",
    metric %in% c("OS.repl", "OS.rich") ~ "OS.partition"
  ))

#write_csv(beta_site, "../analyses/output/beta_site.csv")

beta_site.year <- betalinkr_multi(array_site.year, partition.st = TRUE, partition.rr = TRUE, binary = TRUE) %>%
  as_tibble() %>%
  separate(i, into = c("site1", "year1"), sep = "_") %>%
  separate(j, into = c("site2", "year2"), sep = "_") %>%
  left_join(site_data, by = c("site1" = "site")) %>%
  dplyr::select(site1, year1, site2, year2, S, OS, WN, ST, 
         ST.l, ST.h, ST.lh, WN.repl, OS.repl, WN.rich, OS.rich, 
         elev1 = elev.mean, man1 = management, lon1 = lon, lat1 = lat) %>%
  left_join(site_data, by = c("site2" = "site")) %>%
  dplyr::select(site1, year1, site2, year2, S, OS, WN, ST, 
         ST.l, ST.h, ST.lh, WN.repl, OS.repl, WN.rich, OS.rich, 
         elev1, elev2 = elev.mean, man1, man2 = management, lon1, lat1, lon2 = lon, lat2 = lat) %>%
  left_join(site.year_transects, by = c("site1" = "site", "year1" = "year")) %>%
  dplyr::select(site1, site2, year1, year2, S, OS, WN, ST, 
                ST.l, ST.h, ST.lh, WN.repl, OS.repl, WN.rich, OS.rich, 
                elev1, elev2, man1, man2, lon1, lat1, lon2, lat2, transects1 = transects) %>%
  left_join(site.year_transects, by = c("site2" = "site", "year2" = "year")) %>%
  dplyr::select(site1, site2, year1, year2, S, OS, WN, ST, 
                ST.l, ST.h, ST.lh, WN.repl, OS.repl, WN.rich, OS.rich, 
                elev1, elev2, man1, man2, lon1, lat1, lon2, lat2, transects1, transects2 = transects) %>%
  mutate(elev.diff = abs(elev1 - elev2),
         mean.elev = (elev1 + elev2)/2,
         transects.diff = abs(transects1 - transects2),
         man.diff = case_when(
           man1 == man2 ~ FALSE,
           man1 != man2 ~ TRUE
         )) %>%
  gather(metric, value, -c(site1, year1, site2, year2, elev1, elev2, elev.diff, 
                           man1, man2, man.diff, mean.elev, lon1, lat1, lon2, lat2,
                           transects1, transects2, transects.diff)) %>%
  mutate(metric.class = case_when(
    metric %in% c("WN", "ST", "OS", "S") ~ "aggregate",
    metric %in% c("WN.repl", "WN.rich") ~ "WN.partition",
    metric %in% c("ST.h", "ST.l", "ST.lh") ~ "ST.partition",
    metric %in% c("OS.repl", "OS.rich") ~ "OS.partition"
  ))

### By individual transect

beta_site.date.2010 <- betalinkr_multi(array_site.date.2010, 
                                       partition.st = TRUE, 
                                       partition.rr = TRUE) %>%
  as_tibble() %>%
  separate(i, into = c("site1", "year1", "yday1"), sep = "_") %>%
  separate(j, into = c("site2", "year2", "yday2"), sep = "_") %>%
  left_join(site_data, by = c("site1" = "site"))%>%
  dplyr::select(site1, year1, yday1, site2, year2, yday2, S, OS, WN, ST, 
         ST.l, ST.h, ST.lh, WN.repl, OS.repl, WN.rich, OS.rich, 
         elev1 = elev.mean, man1 = management, lon1 = lon, lat1 = lat) %>%
  left_join(site_data, by = c("site2" = "site")) %>%
  dplyr::select(site1, year1, yday1, site2, year2, yday2, S, OS, WN, ST, 
         ST.l, ST.h, ST.lh, WN.repl, OS.repl, WN.rich, OS.rich, 
         elev1, elev2 = elev.mean, man1, man2 = management, lon1, lat1, lon2 = lon, lat2 = lat) %>%
  mutate(site1 = factor(site1),
         site2 = factor(site2),
         year1 = factor(year1),
         year2 = factor(year2),
         yday1 = as.numeric(yday1),
         yday2 = as.numeric(yday2),
         elev.diff = abs(elev1 - elev2),
         mean.elev = (elev1 + elev2)/2,
         yday.diff = abs(yday1 - yday2),
         man.diff = case_when(
           man1 == man2 ~ FALSE,
           man1 != man2 ~ TRUE
         )) %>%
  gather(metric, value, -c(site1, year1, yday1, site2, year2, yday2,
                           elev1, elev2, elev.diff, yday.diff, 
                           man1, man2, man.diff, mean.elev, 
                           lon1, lat1, lon2, lat2)) %>%
  mutate(metric.class = case_when(
    metric %in% c("WN", "ST", "OS", "S") ~ "aggregate",
    metric %in% c("WN.repl", "WN.rich") ~ "WN.partition",
    metric %in% c("ST.h", "ST.l", "ST.lh") ~ "ST.partition",
    metric %in% c("OS.repl", "OS.rich") ~ "OS.partition"
  ))

beta_site.date.2011 <- betalinkr_multi(array_site.date.2011, 
                                       partition.st = TRUE, 
                                       partition.rr = TRUE) %>%
  as_tibble() %>%
  separate(i, into = c("site1", "year1", "yday1"), sep = "_") %>%
  separate(j, into = c("site2", "year2", "yday2"), sep = "_") %>%
  left_join(site_data, by = c("site1" = "site"))%>%
  dplyr::select(site1, year1, yday1, site2, year2, yday2, S, OS, WN, ST, 
         ST.l, ST.h, ST.lh, WN.repl, OS.repl, WN.rich, OS.rich, 
         elev1 = elev.mean, man1 = management, lon1 = lon, lat1 = lat) %>%
  left_join(site_data, by = c("site2" = "site")) %>%
  dplyr::select(site1, year1, yday1, site2, year2, yday2, S, OS, WN, ST, 
         ST.l, ST.h, ST.lh, WN.repl, OS.repl, WN.rich, OS.rich, 
         elev1, elev2 = elev.mean, man1, man2 = management, lon1, lat1, lon2 = lon, lat2 = lat) %>%
  mutate(site1 = factor(site1),
         site2 = factor(site2),
         year1 = factor(year1),
         year2 = factor(year2),
         yday1 = as.numeric(yday1),
         yday2 = as.numeric(yday2),
         elev.diff = abs(elev1 - elev2),
         mean.elev = (elev1 + elev2)/2,
         yday.diff = abs(yday1 - yday2),
         man.diff = case_when(
           man1 == man2 ~ FALSE,
           man1 != man2 ~ TRUE
         )) %>%
  gather(metric, value, -c(site1, year1, yday1, site2, year2, yday2,
                           elev1, elev2, elev.diff, yday.diff, 
                           man1, man2, man.diff, mean.elev, 
                           lon1, lat1, lon2, lat2)) %>%
  mutate(metric.class = case_when(
    metric %in% c("WN", "ST", "OS", "S") ~ "aggregate",
    metric %in% c("WN.repl", "WN.rich") ~ "WN.partition",
    metric %in% c("ST.h", "ST.l", "ST.lh") ~ "ST.partition",
    metric %in% c("OS.repl", "OS.rich") ~ "OS.partition"
  ))

beta_site.date.2012 <- betalinkr_multi(array_site.date.2012, 
                                       partition.st = TRUE, 
                                       partition.rr = TRUE) %>%
  as_tibble() %>%
  separate(i, into = c("site1", "year1", "yday1"), sep = "_") %>%
  separate(j, into = c("site2", "year2", "yday2"), sep = "_") %>%
  left_join(site_data, by = c("site1" = "site"))%>%
  dplyr::select(site1, year1, yday1, site2, year2, yday2, S, OS, WN, ST, 
         ST.l, ST.h, ST.lh, WN.repl, OS.repl, WN.rich, OS.rich, 
         elev1 = elev.mean, man1 = management, lon1 = lon, lat1 = lat) %>%
  left_join(site_data, by = c("site2" = "site")) %>%
  dplyr::select(site1, year1, yday1, site2, year2, yday2, S, OS, WN, ST, 
         ST.l, ST.h, ST.lh, WN.repl, OS.repl, WN.rich, OS.rich, 
         elev1, elev2 = elev.mean, man1, man2 = management, lon1, lat1, lon2 = lon, lat2 = lat) %>%
  mutate(site1 = factor(site1),
         site2 = factor(site2),
         year1 = factor(year1),
         year2 = factor(year2),
         yday1 = as.numeric(yday1),
         yday2 = as.numeric(yday2),
         elev.diff = abs(elev1 - elev2),
         mean.elev = (elev1 + elev2)/2,
         yday.diff = abs(yday1 - yday2),
         man.diff = case_when(
           man1 == man2 ~ FALSE,
           man1 != man2 ~ TRUE
         )) %>%
  gather(metric, value, -c(site1, year1, yday1, site2, year2, yday2,
                           elev1, elev2, elev.diff, yday.diff, 
                           man1, man2, man.diff, mean.elev, 
                           lon1, lat1, lon2, lat2)) %>%
  mutate(metric.class = case_when(
    metric %in% c("WN", "ST", "OS", "S") ~ "aggregate",
    metric %in% c("WN.repl", "WN.rich") ~ "WN.partition",
    metric %in% c("ST.h", "ST.l", "ST.lh") ~ "ST.partition",
    metric %in% c("OS.repl", "OS.rich") ~ "OS.partition"
  ))

beta_site.date <- bind_rows(beta_site.date.2010, beta_site.date.2011, beta_site.date.2012)
beta_site.date_gdm.prep <- beta_site.date %>%
  filter(metric == "WN")

write_csv(beta_site.date, "../analyses/output/beta_site.date.csv")
```

### Visualize transect-level beta-diversity
```{r}
ggplot(filter(beta_site.date.2010, metric != "S" & 
                metric.class != "OS.partition"), 
       aes(elev.diff, value, color = metric)) +
  geom_point(alpha = 0.05) +
  geom_smooth() +
  facet_wrap(~metric.class)

ggplot(filter(beta_site.date.2010, yday.diff <= 7 &
                metric != "S" & 
                metric.class != "OS.partition"), 
       aes(elev.diff, value, color = metric)) +
  geom_point(alpha = 0.05) +
  geom_smooth() +
  facet_wrap(~metric.class)

ggplot(filter(beta_site.date.2011, yday.diff <= 7 &
                metric != "S" & 
                metric.class != "OS.partition"), 
       aes(elev.diff, value, color = metric)) +
  geom_point(alpha = 0.05) +
  geom_smooth() +
  facet_wrap(~metric.class)

ggplot(filter(beta_site.date.2012, yday.diff <= 7 &
                metric != "S" & 
                metric.class != "OS.partition"), 
       aes(elev.diff, value, color = metric)) +
  geom_point(alpha = 0.05) +
  geom_smooth() +
  facet_wrap(~metric.class)

ggplot(filter(beta_site.date.2012, yday.diff == 0 &
                metric != "S" & 
                metric.class != "OS.partition"),
       aes(value)) +
  geom_histogram()  +
  facet_wrap(~metric)

ggplot(filter(beta_site.date.2012, yday.diff <= 10 &
                metric != "S" & 
                metric.class != "OS.partition"),
       aes(value)) +
  geom_histogram()  +
  facet_wrap(~metric)

ggplot(filter(beta_site.date.2012, yday.diff <= 7 &
                metric != "S" & 
                metric.class != "OS.partition" &
                site1 == site2),
       aes(value)) +
  geom_histogram()  +
  facet_wrap(~metric)

```

# Beta diversity by elevation difference
```{r}
# Interaction beta diversity by elevation difference (by year)
## This shows that the patterns are extremely consistent across years, which is not surprising, since they are driven overwhelmingly by floral turnover, and plants tend to be in the same place from year to year.
ggplot(filter(beta_site.year, year1 == year2 & 
                metric != "S" & 
                metric.class != "OS.partition"), 
       aes(elev.diff, value, color = metric)) +
  geom_point(alpha = 0.25) +
  geom_smooth() +
  facet_grid(metric.class ~ year1)

ggplot(filter(beta_site.year, year1 == year2 & 
                metric != "S" & 
                metric.class != "OS.partition"), 
       aes(transects.diff, value, color = metric)) +
  geom_point(alpha = 0.25) +
  geom_smooth() +
  facet_grid(metric.class ~ year1, scales = "free_x")

# Interaction beta diversity by elevation difference, this time pooling across years
ggplot(filter(beta_site, metric != "S" & 
                metric.class != "OS.partition"), 
       aes(elev.diff, value, color = metric)) +
  geom_point(alpha = 0.05) +
  geom_smooth() +
  facet_wrap(~metric.class)

# How important is the difference in transects? Not as important as the difference in elevation.
ggplot(filter(beta_site, metric != "S" & 
                metric.class != "OS.partition"), 
       aes(transects.diff, value, color = metric)) +
  geom_point(alpha = 0.05) +
  geom_smooth() +
  facet_wrap(~metric.class)

ggplot(beta_site, 
       aes(elev.diff, transects.diff)) +
  geom_point(alpha = 0.05) +
  geom_smooth()

### Okay, we'll just do partial mantel tests to partial out the effect of transects.diff; and we'll only do that for the aggregate measures because it seems a bit silly to doi it for the partitions  
WN_diff <- beta_site %>%
  filter(metric == "WN") %>%
  dplyr::select(site1, site2, value) %>%
  spread(site2, value) %>%
  dist()

WN_diff <- beta_site %>%
  filter(metric == "WN") %>%
  dplyr::select(site1, site2, value) %>%
  spread(site2, value) %>%
  dist()

ST_diff <- beta_site %>%
  filter(metric == "ST") %>%
  dplyr::select(site1, site2, value) %>%
  spread(site2, value) %>%
  dist()

OS_diff <- beta_site %>%
  filter(metric == "OS") %>%
  dplyr::select(site1, site2, value) %>%
  spread(site2, value) %>%
  dist()

elev_diff <- beta_site %>%
  dplyr::select(site1, site2, elev.diff) %>%
  unique() %>%
  spread(site2, elev.diff) %>%
  dist()

transects_diff <- beta_site %>%
  dplyr::select(site1, site2, transects.diff) %>%
  unique() %>%
  spread(site2, transects.diff) %>%
  dist()

WN_mantel <- mantel.partial(elev_diff, WN_diff, transects_diff)
ST_mantel <- mantel.partial(elev_diff, ST_diff, transects_diff)
OS_mantel <- mantel.partial(elev_diff, OS_diff, transects_diff)
```

### Let's look at floral survey diversity patterns
```{r}
### 2012
survey_bioData.2012 <- survey %>%
  mutate(year = year(date),
         yday = yday(date)) %>%
  filter(flower.cover > 0 & year == "2012") %>%
  mutate(plant.sp = str_replace_all(plant.sp, " ", "_")) %>%
  group_by(site, yday, plant.sp) %>%
  summarize(cover = sum(flower.cover)) %>%
  spread(plant.sp, cover) %>%
  replace(is.na(.), 0) %>%
  arrange(site, yday) %>%
  unite(site.day, site, yday, sep = "_")

predData.2012 <- survey %>%
  mutate(year = year(date),
         yday = yday(date)) %>%
  filter(flower.cover > 0 & year == "2012") %>%
  select(site, yday) %>%
  left_join(site_data) %>%
  ungroup() %>%
  dplyr::select(site, elev.mean, lat, lon, yday) %>%
  unique() %>%
  arrange(site, yday) %>%
  unite(site.day, site, yday, sep = "_", remove = "FALSE") %>%
  select(-site)

survey_sitepair.2012 <- formatsitepair(survey_bioData.2012, bioFormat = 1, siteColumn = "site.day", dist = "jaccard",
                              XColumn = "lon", YColumn = "lat", predData = predData.2012)

### Call gdm()
gdm_survey.2012 <- gdm(survey_sitepair.2012, geo = TRUE)
#gdm.varImp(survey_sitepair, geo = TRUE)

summary(gdm_survey.2012)
plot(gdm_survey.2012)
#plotUncertainty(survey_sitepair, sampleSites = 0.7, bsIters = 50, geo  = TRUE)

### 2011
survey_bioData.2011 <- survey %>%
  mutate(year = year(date),
         yday = yday(date)) %>%
  filter(flower.cover > 0 & year == "2011") %>%
  mutate(plant.sp = str_replace_all(plant.sp, " ", "_")) %>%
  group_by(site, yday, plant.sp) %>%
  summarize(cover = sum(flower.cover)) %>%
  spread(plant.sp, cover) %>%
  replace(is.na(.), 0) %>%
  arrange(site, yday) %>%
  unite(site.day, site, yday, sep = "_")

predData.2011 <- survey %>%
  mutate(year = year(date),
         yday = yday(date)) %>%
  filter(flower.cover > 0 & year == "2011") %>%
  select(site, yday) %>%
  left_join(site_data) %>%
  ungroup() %>%
  dplyr::select(site, elev.mean, lat, lon, yday) %>%
  unique() %>%
  arrange(site, yday) %>%
  unite(site.day, site, yday, sep = "_", remove = "FALSE") %>%
  select(-site)

survey_sitepair.2011 <- formatsitepair(survey_bioData.2011, bioFormat = 1, siteColumn = "site.day", dist = "jaccard",
                              XColumn = "lon", YColumn = "lat", predData = predData.2011)

### Call gdm()
gdm_survey.2011 <- gdm(survey_sitepair.2011, geo = TRUE)
#gdm.varImp(survey_sitepair, geo = TRUE)

summary(gdm_survey.2011)
plot(gdm_survey.2011)
#plotUncertainty(survey_sitepair, sampleSites = 0.7, bsIters = 50, geo  = TRUE)

### 2010
survey_bioData.2010 <- survey %>%
  mutate(year = year(date),
         yday = yday(date)) %>%
  filter(flower.cover > 0 & year == "2010") %>%
  mutate(plant.sp = str_replace_all(plant.sp, " ", "_")) %>%
  group_by(site, yday, plant.sp) %>%
  summarize(cover = sum(flower.cover)) %>%
  spread(plant.sp, cover) %>%
  replace(is.na(.), 0) %>%
  arrange(site, yday) %>%
  unite(site.day, site, yday, sep = "_")

predData.2010 <- survey %>%
  mutate(year = year(date),
         yday = yday(date)) %>%
  filter(flower.cover > 0 & year == "2010") %>%
  select(site, yday) %>%
  left_join(site_data) %>%
  ungroup() %>%
  dplyr::select(site, elev.mean, lat, lon, yday) %>%
  unique() %>%
  arrange(site, yday) %>%
  unite(site.day, site, yday, sep = "_", remove = "FALSE") %>%
  select(-site)

survey_sitepair.2010 <- formatsitepair(survey_bioData.2010, bioFormat = 1, siteColumn = "site.day", dist = "jaccard",
                              XColumn = "lon", YColumn = "lat", predData = predData.2010)

### Call gdm()
gdm_survey.2010 <- gdm(survey_sitepair.2010, geo = TRUE)
#gdm.varImp(survey_sitepair, geo = TRUE)

summary(gdm_survey.2010)
plot(gdm_survey.2010)
#plotUncertainty(survey_sitepair, sampleSites = 0.7, bsIters = 50, geo  = TRUE)
```

### And now Bombus diversity patterns
```{r}
BB_bioData <- net %>%
  group_by(site, date, bb.sp) %>%
  summarize(count = n()) %>%
  group_by(site, date) %>%
  mutate(prop = count/sum(count)) %>%
  group_by(site, bb.sp) %>%
  summarize(abund = sum(prop)) %>%
  spread(bb.sp, abund) %>%
  replace(is.na(.), 0)

predData <- site_data %>%
  ungroup() %>%
  dplyr::select(site, elev.mean, lat, lon) %>%
  unique()

bb_sitepair <- formatsitepair(BB_bioData, bioFormat = 1, siteColumn = "site", dist = "jaccard",
                              XColumn = "lon", YColumn = "lat", predData = predData)

### Call gdm()
gdm_bb <- gdm(bb_sitepair, geo = TRUE)
gdm.varImp(bb_sitepair, geo = TRUE)

summary(gdm_bb)
plot(gdm_bb)
plotUncertainty(bb_sitepair, sampleSites = 0.7, bsIters = 50, geo  = TRUE)
```

### Now let's look at interaction beta diversity
```{r}
net_sitepair <- beta_site %>%
  filter(metric == "WN") %>%
  mutate(weights = rep(1, n())) %>%
  dplyr::select(distance = value, weights,
         s1.xCoord = lon1, s1.yCoord = lat1, 
         s2.xCoord = lon2, s2.yCoord = lat2, 
         s1.elev.mean = elev1, s2.elev.mean = elev2) %>%
  data.frame()

### Call gdm()
gdm_net <- gdm(net_sitepair, geo = TRUE)
gdm.varImp(net_sitepair, geo = TRUE)

summary(gdm_net)
plot(gdm_net)
plotUncertainty(net_sitepair, sampleSites = 0.7, bsIters = 50, geo  = TRUE)
```

### At the transect level
```{r}
beta_site.date_gdm.prep

transects_2010 <- beta_site.date.2010 %>%
  filter(metric == "WN") %>%
  mutate(weights = rep(1, n())) %>%
  dplyr::select(distance = value, weights,
         s1.xCoord = lon1, s1.yCoord = lat1, 
         s2.xCoord = lon2, s2.yCoord = lat2, 
         s1.elev = elev1, s2.elev = elev2,
         s1.yday = yday1, s2.yday = yday2) %>%
  data.frame()

### Call gdm()
gdm_net <- gdm(transects_2010, geo = TRUE)
gdm.varImp(transects_2010, geo = TRUE)

summary(gdm_net)
plot(gdm_net)
plotUncertainty(net_sitepair, sampleSites = 0.7, bsIters = 50, geo  = TRUE)
```