### 2020-09-15 analyses/abund_div_GAM.Rmd
refined GDM analysis

---
title: "Abundance/diversity GAM"
output: html_document
---

### Prepare environment
```{r setup, include=FALSE, echo=FALSE}
library(mgcv)
library(mgcViz)
library(tidyverse)
library(lubridate)
```

### Load site data
```{r, include=FALSE, echo=FALSE}
site_data <- read_csv("../processed_data/site_data.csv") %>%
  mutate(elev.class = factor(elev.class, levels = c("oben", "mitte", "unten")), # rename in English
         elev.class2 = factor(elev.class2, levels = c("high", "mid", "low")))
```

### Load network and survey data
```{r message=FALSE}
### BB-FL visitation data
net <- read_csv("../processed_data/network.csv") %>%
  dplyr::select(site, date, bb.sp, plant.sp) # drop unused columns

### Floral survey data
survey <- read_csv("../processed_data/floral_survey.csv") %>%
  semi_join(net, by = "plant.sp") %>% # Consider only visited plant species (anytime anywhere, not specifically for a given site*date)
  dplyr::select(site, date, plant.sp, flower.cover) %>%
  group_by(plant.sp) %>%
  mutate(present = case_when(
    max(flower.cover) > 0 ~ TRUE,
    max(flower.cover) == 0 ~ FALSE
  )) %>%
  filter(present == TRUE) # Ignore plants that never had any flower cover

### Identify duplicate observations; these may be errors or intentionally duplicated (see correspondence with Kalli)
survey_dups <- survey %>%
  group_by(site, date, plant.sp) %>%
  mutate(dups = n()) %>%
  filter(dups > 1)

### Shared transects
shared_transects <- inner_join(net, survey, by = c("site", "date")) %>%
  dplyr::select(site, date) %>%
  unique() 
```

### FL and BB abundance / diversity
```{r echo=FALSE, include=TRUE, message=FALSE}
# Per-transect BB abundance
## full dataset
bb_abund_sp <- net %>%
  dplyr::select(site, date, bb.sp) %>%
  group_by(site, date, bb.sp) %>% # 
  summarize(bb.abund = n()) %>% # get per transect bb abundance
  ungroup() %>%
  complete(site, date, bb.sp, fill = list(bb.abund = 0)) %>% # complete implicit absences
  semi_join(net, by = c("site", "date")) %>% # drop site-dates that did not exist in the data set
  left_join(site_data) %>% # add site data
  mutate(year = factor(year(date)),
         yday = yday(date),
         bb.sp = factor(bb.sp),
         management = factor(management),
         site = factor(site),
         transect = factor(transect)) %>%
  group_by(year, bb.sp) %>%
  mutate(bb.annual.abund = sum(bb.abund))

## filtered to include only samples with corresponding floral survey data
bb_abund_sp.sub <- bb_abund_sp %>%
  ungroup() %>%
  semi_join(shared_transects) %>%
  select(site, date, yday, bb.sp, bb.abund, bb.annual.abund, elev.mean)

### Per-transect floral abund: by species
fl_abund_sp <- survey %>%
  group_by(site, date, plant.sp) %>% # group by plant.sp*transect
  summarize(fl.abund = sum(flower.cover)) %>% # get plant.sp*transect total flower cover; this step really shouldn't be required, but in a few cases, there is more than one abundance entry for the same species on a given transect; not sure how that happened, but it's in the original data that Fabrice sent me; alternatively, I could drop the dups or average them. We'll see what Kalli and Fabrice have to say.
  group_by(site, date) %>%
  mutate(prop.fl.abund = fl.abund/sum(fl.abund)) %>% # get the proportional flower cover for each species.
  ungroup() %>%
  #full_join(dplyr::select(net, site, date, plant.sp), by = c("site", "date", "plant.sp")) %>% # add the handful of species occurring in the visitation network but missing from the survey data
  complete(site, date, plant.sp, fill = list(fl.abund = 0, prop.fl.abund = 0)) %>%
  semi_join(survey, by = c("site", "date")) %>%  # only include site-dates that actually occurred
  unique() %>%
  mutate(year = factor(year(date)))
```

### Preference analysis
```{r message=FALSE}
observed_visits <- net %>%
  semi_join(shared_transects) %>% # include only site*date combos where floral surveying was done
  group_by(site, date, bb.sp, plant.sp) %>% # group by bb*plant*transect
  summarize(visits = n()) %>% # total visits per bb-fl pair
  group_by(site, date, bb.sp) %>%
  mutate(prop.visits = visits/sum(visits))

# All possible bumble-plant pairings *site * date, filtered to include only those when the given bb.sp was present and the given plant.sp. Note that in about 10% of the bb.sp*plant.sp*site*year combos, the visited plant was not recorded in the floral survey data. There are a few different ways we could handle this, but for now I am simply ignoring plant spp. for which we have no abundance data.
visitation_by_abundance <- fl_abund_sp %>%
  inner_join(bb_abund_sp.sub) %>%
  filter(bb.abund > 0 & fl.abund > 0) %>%
  left_join(observed_visits) %>%
  replace_na(list(visits = 0, prop.visits = 0))

### Get mean abundance/visitation for each plant*bb combo
visitation_by_abundance_mean <- visitation_by_abundance %>%
  group_by(plant.sp, bb.sp) %>%
  summarize(prop.fl.abund.m = mean(prop.fl.abund),
            prop.visits.wm = weighted.mean(prop.visits, bb.abund))

# Visualize
ggplot(filter(visitation_by_abundance_mean, prop.visits.wm > 0), 
       aes(log(prop.fl.abund.m + 0.01), log(prop.visits.wm))) +
  geom_point(alpha = 0.5, aes(color = bb.sp)) +
  geom_smooth(method = "lm")

ggplot(filter(visitation_by_abundance_mean), 
       aes(prop.fl.abund.m, prop.visits.wm)) +
  geom_point(alpha = 0.5, aes(color = bb.sp)) +
  geom_smooth(method = "lm")

### Okay, here's what we're going to do. Preference has a strong bimodal pattern, with lots of plant species totally ignored while the ones that are foraged upon are visited in proportion to their abundance. We can't model that latter pattern well if we retain all the totally ignored plant species. So, let's model the visited plants separately from the unvisited ones.

### Call linear model on the subset of the floral species that were visited at least once
# log on log regression is common for proportional datam (log ratio regression)
pref.lm <- lm(log(prop.visits.wm) ~ log(prop.fl.abund.m + 0.01), 
              data = filter(visitation_by_abundance_mean, prop.visits.wm > 0))
summary(pref.lm)
plot(pref.lm)

### Extract residuals
pref.lm.resid <- residuals(pref.lm) %>%
  data.frame() %>%
  as_tibble() %>%
  rename(preference = 1)

#################################
### Build preference data set ###
#################################

# Start with the mean prop.fl.abund and weighted mean prop.visits
preference <- visitation_by_abundance_mean %>%  
  ungroup() %>%

# Remove bb*fl combos that had 0 visits
  filter(prop.visits.wm > 0) %>%

# With the zeroes removed, what's left can be joined 1:1 with the 
# residuals of the model above, i.e. the raw preference data
  bind_cols(pref.lm.resid) %>%

# Exponentiate the residuals to back-transform the data and ensure 
# that preference is > 0 for all visited plant species.
  mutate(preference = exp(preference)) %>%

# Now we need to add back all the bb*fl*year combos that had 0 visits
# Then we have NAs for preference and floral abundance columns
  full_join(unique(dplyr::select(visitation_by_abundance_mean, bb.sp, plant.sp))) %>%

# Replace the NAs of the preference column with 0s; 
  replace_na(list(preference = 0)) %>%

# Drop the columns used to calculate preference so that they don't confuse us when we add site-date level abundance data
  dplyr::select(bb.sp, plant.sp, preference) %>%
  mutate(sqrt.preference = sqrt(preference),
         bin.preference = case_when(
           preference > 0 ~ 1,
           preference == 0 ~ 0
         ))

write_csv(preference, "../analyses/output/preference.csv")

ggplot(preference, aes(preference)) +
  geom_histogram() +
  facet_wrap(~bb.sp)

ggplot(preference, aes(sqrt.preference)) +
  geom_histogram() +
  facet_wrap(~bb.sp)

#################################
### Join with FL and BB abund ###
#################################

#After reconsidering this approach, I think it is better to assume preference = 0  for non-co-occuring spp.

# # For bb-fl combos that never co-occurred, set preference = 0
# pref_forbidden_link_fill <- inner_join(fl_abund_sp, bb_abund_sp.sub) %>%
#   mutate(co = case_when(
#     fl.abund == 0 | bb.abund == 0 ~ 0, # bb.sp and fl.sp that did not co-occur on given site date
#     fl.abund != 0 & bb.abund != 0 ~ 1) # bb.sp and fl.sp that did co-occur on a given site date
#   ) %>%
#   group_by(plant.sp, bb.sp) %>%
#   summarize(co.max = max(co)) %>% # if max(co) = 0, a given bb.sp*fl.sp combo *never* co-occurred, and therefore will be assigned aneutral preference
#   filter(co.max == 0) %>%
#   mutate(preference.update = rep(1, n())) %>%
#   dplyr::select(-co.max) %>%
#   inner_join(fl_abund_sp) %>%
#   select(-year)

fl_abund_wt <- fl_abund_sp %>%
  select(-year) %>%
  left_join(preference) %>%

# Add the forbidden links with preference set to 1 (neutral) (not doing this anymore)
  # full_join(pref_forbidden_link_fill) %>%
  # mutate(preference = case_when(
  #   is.na(preference) ~ preference.update,
  #   !is.na(preference) ~ preference)
  #   ) %>%
  # dplyr::select(-preference.update) %>%
  mutate(fl.abund.wt.bin = fl.abund*bin.preference,
         fl.abund.wt.sqrt = fl.abund*sqrt.preference,
         fl.abund.wt = fl.abund*preference) %>% # using sqrt-transformed preference allows me to avoid log-transforming when trying to fit GAMs
  mutate(yday = yday(date)) %>%
  group_by(site, date, yday, bb.sp) %>%
  summarize(fl.abund.wt.bin = sum(fl.abund.wt.bin, na.rm = TRUE),
            fl.abund.wt.sqrt = sum(fl.abund.wt.sqrt, na.rm = TRUE),
            fl.abund.wt = sum(fl.abund.wt, na.rm = TRUE)) %>%
  ungroup() %>%
  semi_join(survey, by = c("site", "date")) %>% # This gets rid of site*date combos that don't exist
  left_join(site_data, by = "site") %>%
  mutate(year = factor(year(date)), # add in the years missing from the bb_abund join
         bb.sp = factor(bb.sp), # convert bb.sp to a factor
         management = factor(management),
         site = factor(site),
         transect = factor(transect)) %>%
  group_by(year) %>%
  mutate(total.transects.year = n()) %>%
  ungroup()

write_csv(fl_abund_wt, "../analyses/output/fl_abund_wt.csv")
```

### Call GAMs
##### My argument for not including site as a random effect is that it would create identifiability issues with the elevation variable that I am actally interested in. This will probably cause the p-values to be underestimated, maybe to the point of being uninformative. But I don't think I care. My goal is not to establish inferentially that time and elevation affect BB abundance; I want to say how they do. And I don't want to obscure the "how" for the sake of inferential scruples.
#### Floral resources
##### After trying a lot of things, I'm pretty sure the only option I have for fitting a good GAM is using log-transformation. The question that remains is how to treat the concept of preference. My "raw" preference-weighted abundance gives our measure of preference a lot of influence. If we trust it, this may be most informative, and it allows the greatest variation in floral resources between BB species. But if there is anything misleading in our preference estimates, using the "raw" measures will do nothing to mitigate this. An approach that gives us a "softer" notion of preference is to take the sqrt of the raw prference before multiplying floral abundance by it. An argument for this approach is that it retains most of the meaning of the raw preference data while dampening its effect, which may help avoid spurious hotspots based on overestimated preference. Finally, an approach that says "we don't really trust our data very much" is to treat preference as binary: visited species get a 1, non-visited species get a 0, and all the "weighting" of floral resource abundance means is dropping floral species that a given BB species did not visit. In all these cases, though, the distribution of weighted floral abundance is strongly logarithmic (or exponential or something), so a log transformation is necessary. I tried softer methods, e.g. sqrt-transformation, Tweedie family, etc., but the problem with all these is that the variance fans out like crazy as you move up to the higher end of the distribution. Happily, it makes little difference which of these options we choose, as the resulting plots lead to the same interpretation.
```{r message=FALSE, warning=FALSE}
# These are the distributions of my response variable options 
ggplot(fl_abund_wt, aes(fl.abund.wt.bin)) +
  geom_histogram()

ggplot(fl_abund_wt, aes(fl.abund.wt.sqrt)) +
  geom_histogram()

ggplot(fl_abund_wt, aes(fl.abund.wt)) +
  geom_histogram()

### 2010
weighted_fl_abundance.2010 <- bam(log(fl.abund.wt + 0.01) ~
                                      s(bb.sp, bs = "re") +
                                      #s(management, bs = "re") +
                                      #s(site, bs = "re") +
                                      te(yday, elev.mean, 
                                                by = bb.sp,
                                                #id = 0,
                                                bs= c("gp", "tp"), 
                                                k=c(8, 8), 
                                                m=2),
                                    data = filter(fl_abund_wt, year == 2010),
                                    discrete = TRUE,
                                    family = "gaussian",
                                    select = TRUE,
                                    method = "fREML") %>% getViz(nsim = 500)

#check.gamViz(weighted_fl_abundance.2010)
#weighted_fl_abundance.2010.sum <- summary(weighted_fl_abundance.2010)
#print(plot(weighted_fl_abundance.2010, allTerms = TRUE), pages = 4)
#weighted_fl_abundance.2010.ck <- check2D(weighted_fl_abundance.2010, x1 = "yday", x2 = "elev.mean")

weighted_fl_abundance.2010.sqrt <- bam(log(fl.abund.wt.sqrt + 0.01) ~
                                      s(bb.sp, bs = "re") +
                                      #s(management, bs = "re") +
                                      #s(site, bs = "re") +
                                      te(yday, elev.mean, 
                                                by = bb.sp,
                                                #id = 0,
                                                bs= c("gp", "tp"), 
                                                k=c(8, 8), 
                                                m=2),
                                    data = filter(fl_abund_wt, year == 2010),
                                    discrete = TRUE,
                                    family = "gaussian",
                                    select = TRUE,
                                    method = "fREML") %>% getViz(nsim = 500)

#check.gamViz(weighted_fl_abundance.2010.sqrt)
#weighted_fl_abundance.2010.sum.sqrt <- summary(weighted_fl_abundance.2010.sqrt)
#print(plot(weighted_fl_abundance.2010.sqrt, allTerms = TRUE), pages = 4)
#weighted_fl_abundance.2010.ck.sqrt <- check2D(weighted_fl_abundance.2010.sqrt, x1 = "yday", x2 = "elev.mean")

weighted_fl_abundance.2010.bin <- bam(log(fl.abund.wt.bin + 0.01) ~
                                      s(bb.sp, bs = "re") +
                                      #s(management, bs = "re") +
                                      #s(site, bs = "re") +
                                      te(yday, elev.mean, 
                                                by = bb.sp,
                                                #id = 0,
                                                bs= c("gp", "tp"), 
                                                k=c(8, 8), 
                                                m=2),
                                    data = filter(fl_abund_wt, year == 2010),
                                    discrete = TRUE,
                                    family = "gaussian",
                                    select = TRUE,
                                    method = "fREML") %>% getViz(nsim = 500)

#check.gamViz(weighted_fl_abundance.2010.bin)
#weighted_fl_abundance.2010.sum.bin <- summary(weighted_fl_abundance.2010.bin)
#print(plot(weighted_fl_abundance.2010.bin, allTerms = TRUE), pages = 4)
#weighted_fl_abundance.2010.ck.bin <- check2D(weighted_fl_abundance.2010.bin, x1 = "yday", x2 = "elev.mean")

### 2011
weighted_fl_abundance.2011 <- bam(log(fl.abund.wt + 0.01) ~
                                      s(bb.sp, bs = "re") +
                                      #s(management, bs = "re") +
                                      #s(site, bs = "re") +
                                      te(yday, elev.mean, 
                                                by = bb.sp,
                                                #id = 0,
                                                bs= c("gp", "tp"), 
                                                k=c(8, 8), 
                                                m=2),
                                    data = filter(fl_abund_wt, year == 2011),
                                    discrete = TRUE,
                                    family = "gaussian",
                                    select = TRUE,
                                    method = "fREML") %>% getViz(nsim = 500)

#check.gamViz(weighted_fl_abundance.2011)
#weighted_fl_abundance.2011.sum <- summary(weighted_fl_abundance.2011)
#print(plot(weighted_fl_abundance.2011, allTerms = TRUE), pages = 4)
#weighted_fl_abundance.2011.ck <- check2D(weighted_fl_abundance.2011, x1 = "yday", x2 = "elev.mean")

weighted_fl_abundance.2011.sqrt <- bam(log(fl.abund.wt.sqrt + 0.01) ~
                                      s(bb.sp, bs = "re") +
                                      #s(management, bs = "re") +
                                      #s(site, bs = "re") +
                                      te(yday, elev.mean, 
                                                by = bb.sp,
                                                #id = 0,
                                                bs= c("gp", "tp"), 
                                                k=c(8, 8), 
                                                m=2),
                                    data = filter(fl_abund_wt, year == 2011),
                                    discrete = TRUE,
                                    family = "gaussian",
                                    select = TRUE,
                                    method = "fREML") %>% getViz(nsim = 500)

#check.gamViz(weighted_fl_abundance.2011.sqrt)
#weighted_fl_abundance.2011.sum.sqrt <- summary(weighted_fl_abundance.2011.sqrt)
#print(plot(weighted_fl_abundance.2011.sqrt, allTerms = TRUE), pages = 4)
#weighted_fl_abundance.2011.ck.sqrt <- check2D(weighted_fl_abundance.2011.sqrt, x1 = "yday", x2 = "elev.mean")

weighted_fl_abundance.2011.bin <- bam(log(fl.abund.wt.bin + 0.01) ~
                                      s(bb.sp, bs = "re") +
                                      #s(management, bs = "re") +
                                      #s(site, bs = "re") +
                                      te(yday, elev.mean, 
                                                by = bb.sp,
                                                #id = 0,
                                                bs= c("gp", "tp"), 
                                                k=c(8, 8), 
                                                m=2),
                                    data = filter(fl_abund_wt, year == 2011),
                                    discrete = TRUE,
                                    family = "gaussian",
                                    select = TRUE,
                                    method = "fREML") %>% getViz(nsim = 500)

#check.gamViz(weighted_fl_abundance.2011.bin)
#weighted_fl_abundance.2011.sum.bin <- summary(weighted_fl_abundance.2011.bin)
#print(plot(weighted_fl_abundance.2011.bin, allTerms = TRUE), pages = 4)
#weighted_fl_abundance.2011.ck.bin <- check2D(weighted_fl_abundance.2011.bin, x1 = "yday", x2 = "elev.mean")

### 2012
weighted_fl_abundance.2012 <- bam(log(fl.abund.wt + 0.01) ~
                                      s(bb.sp, bs = "re") +
                                      #s(management, bs = "re") +
                                      #s(site, bs = "re") +
                                      te(yday, elev.mean, 
                                                by = bb.sp,
                                                #id = 0,
                                                bs= c("gp", "tp"), 
                                                k=c(8, 8), 
                                                m=2),
                                    data = filter(fl_abund_wt, year == 2012),
                                    discrete = TRUE,
                                    family = "gaussian",
                                    select = TRUE,
                                    method = "fREML") %>% getViz(nsim = 500)

#check.gamViz(weighted_fl_abundance.2012)
#weighted_fl_abundance.2012.sum <- summary(weighted_fl_abundance.2012)
#print(plot(weighted_fl_abundance.2012, allTerms = TRUE), pages = 4)
#weighted_fl_abundance.2012.ck <- check2D(weighted_fl_abundance.2012, x1 = "yday", x2 = "elev.mean")

weighted_fl_abundance.2012.sqrt <- bam(log(fl.abund.wt.sqrt + 0.01) ~
                                      s(bb.sp, bs = "re") +
                                      #s(management, bs = "re") +
                                      #s(site, bs = "re") +
                                      te(yday, elev.mean, 
                                                by = bb.sp,
                                                #id = 0,
                                                bs= c("gp", "tp"), 
                                                k=c(8, 8), 
                                                m=2),
                                    data = filter(fl_abund_wt, year == 2012),
                                    discrete = TRUE,
                                    family = "gaussian",
                                    select = TRUE,
                                    method = "fREML") %>% getViz(nsim = 500)

#check.gamViz(weighted_fl_abundance.2012.sqrt)
#weighted_fl_abundance.2012.sum.sqrt <- summary(weighted_fl_abundance.2012.sqrt)
#print(plot(weighted_fl_abundance.2012.sqrt, allTerms = TRUE), pages = 4)
#weighted_fl_abundance.2012.ck.sqrt <- check2D(weighted_fl_abundance.2012.sqrt, x1 = "yday", x2 = "elev.mean")

weighted_fl_abundance.2012.bin <- bam(log(fl.abund.wt.bin + 0.01) ~
                                      s(bb.sp, bs = "re") +
                                      #s(management, bs = "re") +
                                      #s(site, bs = "re") +
                                      te(yday, elev.mean, 
                                                by = bb.sp,
                                                #id = 0,
                                                bs= c("gp", "tp"), 
                                                k=c(8, 8), 
                                                m=2),
                                    data = filter(fl_abund_wt, year == 2012),
                                    discrete = TRUE,
                                    family = "gaussian",
                                    select = TRUE,
                                    method = "fREML") %>% getViz(nsim = 500)

#check.gamViz(weighted_fl_abundance.2012.bin)
#weighted_fl_abundance.2012.sum.bin <- summary(weighted_fl_abundance.2012.bin)
#print(plot(weighted_fl_abundance.2012.bin, allTerms = TRUE), pages = 4)
#weighted_fl_abundance.2012.ck.bin <- check2D(weighted_fl_abundance.2012.bin, x1 = "yday", x2 = "elev.mean")
```

#### Bumbles
```{r message=FALSE}
### 2010
bb_abund.2010 <- bam(bb.abund ~ 
                       s(bb.sp, bs = "re") + 
                       te(yday, elev.mean, 
                          by = bb.sp,
                          bs= c("gp", "cr"), 
                          k=c(8, 8), 
                          m=2),
                     data = filter(bb_abund_sp, year == 2010 & bb.annual.abund >= 5),
                     discrete = TRUE,
                     family = "tw",
                     select = TRUE,
                     method = "fREML") %>% getViz(nsim = 500)

#check.gamViz(bb_abund.2010)
#bb_abund.2010.sum <- summary(bb_abund.2010)
#print(plot(bb_abund.2010, allTerms = TRUE), pages = 1)
#bb_abund.2010.ck <- check2D(bb_abund.2010, x1 = "yday", x2 = "elev.mean") # See if there is any spatiotemporal pattern to the residuals

### 2011
bb_abund.2011 <- bam(bb.abund ~ 
                       s(bb.sp, bs = "re") + 
                       te(yday, elev.mean, 
                          by = bb.sp,
                          bs= c("gp", "cr"), 
                          k=c(8, 8), 
                          m=2),
                     data = filter(bb_abund_sp, year == 2011 & bb.annual.abund >= 5),
                     discrete = TRUE,
                     family = "tw",
                     select = TRUE,
                     method = "fREML") %>% getViz(nsim = 500)

#check.gamViz(bb_abund.2011)
#bb_abund.2011.sum <- summary(bb_abund.2011)
#print(plot(bb_abund.2011, allTerms = TRUE), pages = 1)
#bb_abund.2011.ck <- check2D(bb_abund.2011, x1 = "yday", x2 = "elev.mean") # See if there is any spatiotemporal pattern to the residuals

### 2012
bb_abund.2012 <- bam(bb.abund ~ 
                       s(bb.sp, bs = "re") + 
                       te(yday, elev.mean, 
                          by = bb.sp,
                          bs= c("gp", "cr"), 
                          k=c(8, 8), 
                          m=2),
                     data = filter(bb_abund_sp, year == 2012 & bb.annual.abund >= 5),
                     discrete = TRUE,
                     family = "tw",
                     select = TRUE,
                     method = "fREML") %>% getViz(nsim = 500)

#check.gamViz(bb_abund.2012)
#bb_abund.2012.sum <- summary(bb_abund.2012)
#print(plot(bb_abund.2012, allTerms = TRUE), pages = 1)
#bb_abund.2012.ck <- check2D(bb_abund.2012, x1 = "yday", x2 = "elev.mean") # See if there is any spatiotemporal pattern to the residuals
```

### Extract predictions
```{r message=FALSE}
# Function to build data set over which to generate predictions
newdata_build <- function(dat, y) {

  ## yday min/max
  min_yday <- min(filter(dat, year == y)$yday)
  max_yday <- max(filter(dat, year == y)$yday)

  ## elev min.max
  min_elev <- min(filter(dat, year == y)$elev.mean)
  max_elev <- max(filter(dat, year == y)$elev.mean)

  ## species list
  spp <- dat %>%
    filter(year == y) %>%
    select(bb.sp) %>%
    unique()

  # sequence of 50 points along the yday and elev ranges, respectively
  seq_yday <- seq(min_yday, max_yday, length.out = 50)
  seq_elev <- seq(min_elev, max_elev, length.out = 50)

  newdata <- tibble(seq_yday, seq_elev) %>% # combine seqs into data frame
    complete(seq_yday, seq_elev) %>% # get all possible values
    crossing(spp) %>% # get all possible combos with species
    select(yday = seq_yday, elev.mean = seq_elev, bb.sp) # rename columns to match the original model inputs
}

newdata2010.FL <- newdata_build(fl_abund_wt, 2010)
newdata2011.FL <- newdata_build(fl_abund_wt, 2011)
newdata2012.FL <- newdata_build(fl_abund_wt, 2012)

newdata2010.BB <- newdata_build(filter(bb_abund_sp, bb.annual.abund >= 5), 2010)
newdata2011.BB <- newdata_build(filter(bb_abund_sp, bb.annual.abund >= 5), 2011)
newdata2012.BB <- newdata_build(filter(bb_abund_sp, bb.annual.abund >= 5), 2012)

# A little wrapper to make predictions and return a tibble with desired columns 
predict_func <- function(mod, new, y) {
  predict.bam(mod, newdata = new, type = "response") %>%
    data.frame() %>%
    tibble() %>%
    bind_cols(new) %>%
    select(yday, elev.mean, bb.sp, pred = 1) %>%
    mutate(year = rep(y, n()))
}

# Generate predictions
predict2010.FL <- predict_func(weighted_fl_abundance.2010, newdata2010.FL, 2010)
predict2010.FL.sqrt <- predict_func(weighted_fl_abundance.2010.sqrt, newdata2010.FL, 2010)
predict2010.FL.bin <- predict_func(weighted_fl_abundance.2010.bin, newdata2010.FL, 2010)

predict2011.FL <- predict_func(weighted_fl_abundance.2011, newdata2011.FL, 2011)
predict2011.FL.sqrt <- predict_func(weighted_fl_abundance.2011.sqrt, newdata2011.FL, 2011)
predict2011.FL.bin <- predict_func(weighted_fl_abundance.2011.bin, newdata2011.FL, 2011)

predict2012.FL <- predict_func(weighted_fl_abundance.2012, newdata2012.FL, 2012)
predict2012.FL.sqrt <- predict_func(weighted_fl_abundance.2012.sqrt, newdata2012.FL, 2012)
predict2012.FL.bin <- predict_func(weighted_fl_abundance.2012.bin, newdata2012.FL, 2012)

predict_pool.FL <- bind_rows(predict2010.FL, predict2011.FL, predict2012.FL) %>%
  mutate(date = as_date(yday),
         subject = factor(rep("FL", n())))
predict_pool.FL.sqrt <- bind_rows(predict2010.FL.sqrt, predict2011.FL.sqrt, predict2012.FL.sqrt) %>%
  mutate(date = as_date(yday),
         subject = factor(rep("FL", n())))
predict_pool.FL.bin <- bind_rows(predict2010.FL.bin, predict2011.FL.bin, predict2012.FL.bin) %>%
  mutate(date = as_date(yday),
         subject = factor(rep("FL", n())))

predict2010.BB <- predict_func(bb_abund.2010, newdata2010.BB, 2010)
predict2011.BB <- predict_func(bb_abund.2011, newdata2011.BB, 2011)
predict2012.BB <- predict_func(bb_abund.2012, newdata2012.BB, 2012)
predict_pool.BB <- bind_rows(predict2010.BB, predict2011.BB, predict2012.BB) %>%
  mutate(date = as_date(yday),
         subject = factor(rep("BB", n())),
         pred = log(pred))

predict_pool <- bind_rows(predict_pool.FL, predict_pool.BB)
predict_pool.sqrt <- bind_rows(predict_pool.FL.sqrt, predict_pool.BB)
predict_pool.bin <- bind_rows(predict_pool.FL.bin, predict_pool.BB)

# Subset predictions
predict_major <- filter(predict_pool, 
                           bb.sp %in% c("bss", "pasc", "prat", "soro", "wurf")) %>%
  mutate(bb.sp = case_when(
    bb.sp == "bss" ~ "B. ss.",
    bb.sp == "pasc" ~ "B. pasc",
    bb.sp == "prat" ~ "B. prat",
    bb.sp == "soro" ~ "B. soro",
    bb.sp == "wurf" ~ "B. wurf"
  ))

predict_major.sqrt <- filter(predict_pool.sqrt, 
                           bb.sp %in% c("bss", "pasc", "prat", "soro", "wurf")) %>%
  mutate(bb.sp = case_when(
    bb.sp == "bss" ~ "B. ss.",
    bb.sp == "pasc" ~ "B. pasc",
    bb.sp == "prat" ~ "B. prat",
    bb.sp == "soro" ~ "B. soro",
    bb.sp == "wurf" ~ "B. wurf"
  ))

predict_major.bin <- filter(predict_pool.bin, 
                           bb.sp %in% c("bss", "pasc", "prat", "soro", "wurf")) %>%
  mutate(bb.sp = case_when(
    bb.sp == "bss" ~ "B. ss.",
    bb.sp == "pasc" ~ "B. pasc",
    bb.sp == "prat" ~ "B. prat",
    bb.sp == "soro" ~ "B. soro",
    bb.sp == "wurf" ~ "B. wurf"
  ))

predict_highelev <- filter(predict_pool, 
                              bb.sp %in% c("pyre", "mend", "mont")) %>%
  mutate(bb.sp = case_when(
    bb.sp == "pyre" ~ "B. pyre",
    bb.sp == "mend" ~ "B. mend",
    bb.sp == "mont" ~ "B. mont"
  ))

predict_highelev.sqrt <- filter(predict_pool.sqrt, 
                                   bb.sp %in% c("pyre", "mend", "mont")) %>%
  mutate(bb.sp = case_when(
    bb.sp == "pyre" ~ "B. pyre",
    bb.sp == "mend" ~ "B. mend",
    bb.sp == "mont" ~ "B. mont"
  ))

predict_highelev.bin <- filter(predict_pool.bin, 
                                  bb.sp %in% c("pyre", "mend", "mont")) %>%
  mutate(bb.sp = case_when(
    bb.sp == "pyre" ~ "B. pyre",
    bb.sp == "mend" ~ "B. mend",
    bb.sp == "mont" ~ "B. mont"
  ))

predict_minor <- filter(predict_pool, 
                           bb.sp %in% c("gers", "hort", "hypn","jone", "lapi", "muci")) %>%
  mutate(bb.sp = case_when(
    bb.sp == "gers" ~ "B. gers",
    bb.sp == "hort" ~ "B. hort",
    bb.sp == "hypn" ~ "B. hypn",
    bb.sp == "jone" ~ "B. jone",
    bb.sp == "lapi" ~ "B. lapi",
    bb.sp == "muci" ~ "B. muci"
  ))

predict_minor.sqrt <- filter(predict_pool.sqrt, 
                                bb.sp %in% c("gers", "hort", "hypn","jone", "lapi", "muci")) %>% 
  mutate(bb.sp = case_when(
    bb.sp == "gers" ~ "B. gers",
    bb.sp == "hort" ~ "B. hort",
    bb.sp == "hypn" ~ "B. hypn",
    bb.sp == "jone" ~ "B. jone",
    bb.sp == "lapi" ~ "B. lapi",
    bb.sp == "muci" ~ "B. muci"
  ))

predict_minor.bin <- filter(predict_pool.bin, 
                               bb.sp %in% c("gers", "hort", "hypn","jone", "lapi", "muci")) %>% 
  mutate(bb.sp = case_when(
    bb.sp == "gers" ~ "B. gers",
    bb.sp == "hort" ~ "B. hort",
    bb.sp == "hypn" ~ "B. hypn",
    bb.sp == "jone" ~ "B. jone",
    bb.sp == "lapi" ~ "B. lapi",
    bb.sp == "muci" ~ "B. muci"
  ))
```

### Visualize
```{r}
#### Major
ggplot(predict_major, aes(date, elev.mean, fill = pred)) +
  geom_raster() +
  geom_contour(aes(z = pred), color = "black", size = 0.25) +
  scale_fill_viridis_c(name = "log(abund.)") +
  #facet_grid(bb.sp + subject ~ year) +
  facet_grid(year ~ bb.sp + subject) +
  xlab("Date") +
  ylab("Elevation (m)") +
  theme(axis.text.x = element_text(angle = 90, size = 8, vjust = 0.5))
ggsave("./output/GAM_major.png", height = 5, width = 11, units = "in")

ggplot(predict_major.sqrt, aes(date, elev.mean, fill = pred)) +
  geom_raster() +
  geom_contour(aes(z = pred), color = "black", size = 0.25) +
  scale_fill_viridis_c(name = "log(abund.)") +
  #facet_grid(bb.sp + subject ~ year) +
  facet_grid(year ~ bb.sp + subject) +
  xlab("Date") +
  ylab("Elevation (m)") +
  theme(axis.text.x = element_text(angle = 90, vjust = 1, hjust = 1))
ggsave("./output/GAM_major_sqrt.png")

ggplot(predict_major.bin, aes(date, elev.mean, fill = pred)) +
  geom_raster() +
  geom_contour(aes(z = pred), color = "black", size = 0.25) +
  scale_fill_viridis_c(name = "log(abund.)") +
  #facet_grid(bb.sp + subject ~ year) +
  facet_grid(year ~ bb.sp + subject) +
  xlab("Date") +
  ylab("Elevation (m)") +
  theme(axis.text.x = element_text(angle = 90, vjust = 1, hjust = 1))
ggsave("./output/GAM_major_bin.png")

#### High Elev
ggplot(predict_highelev, aes(date, elev.mean, fill = pred)) +
  geom_raster() +
  geom_contour(aes(z = pred), color = "black", size = 0.25) +
  scale_fill_viridis_c(name = "log(abund.)") +
  #facet_grid(bb.sp + subject ~ year) +
  facet_grid(year ~ bb.sp + subject) +
  xlab("Date") +
  ylab("Elevation (m)") +
  theme(axis.text.x = element_text(angle = 90, size = 8, vjust = 0.5))
ggsave("./output/GAM_highelev.png", height = 5, width = 11, units = "in")

ggplot(predict_highelev.sqrt, aes(date, elev.mean, fill = pred)) +
  geom_raster() +
  geom_contour(aes(z = pred), color = "black", size = 0.25) +
  scale_fill_viridis_c(name = "log(abund.)") +
  #facet_grid(bb.sp + subject ~ year) +
  facet_grid(year ~ bb.sp + subject) +
  xlab("Date") +
  ylab("Elevation (m)") +
  theme(axis.text.x = element_text(angle = 90, size = 8, vjust = 0.5))
ggsave("./output/GAM_highelev_sqrt.png", height = 5, width = 11, units = "in")

ggplot(predict_highelev.bin, aes(date, elev.mean, fill = pred)) +
  geom_raster() +
  geom_contour(aes(z = pred), color = "black", size = 0.25) +
  scale_fill_viridis_c(name = "log(abund.)") +
  #facet_grid(bb.sp + subject ~ year) +
  facet_grid(year ~ bb.sp + subject) +
  xlab("Date") +
  ylab("Elevation (m)") +
  theme(axis.text.x = element_text(angle = 90, size = 8, vjust = 0.5))
ggsave("./output/GAM_highelev_bin.png", height = 5, width = 11, units = "in")

#### Minor
ggplot(predict_minor, aes(date, elev.mean, fill = pred)) +
  geom_raster() +
  geom_contour(aes(z = pred), color = "black", size = 0.25) +
  scale_fill_viridis_c(name = "log(abund.)") +
  #facet_grid(bb.sp + subject ~ year) +
  facet_grid(year ~ bb.sp + subject) +
  xlab("Date") +
  ylab("Elevation (m)") +
  theme(axis.text.x = element_text(angle = 90, size = 8, vjust = 0.5))
ggsave("./output/GAM_minor.png", height = 5, width = 11, units = "in")

ggplot(predict_minor.sqrt, aes(date, elev.mean, fill = pred)) +
  geom_raster() +
  geom_contour(aes(z = pred), color = "black", size = 0.25) +
  scale_fill_viridis_c(name = "log(abund.)") +
  #facet_grid(bb.sp + subject ~ year) +
  facet_grid(year ~ bb.sp + subject) +
  xlab("Date") +
  ylab("Elevation (m)") +
  theme(axis.text.x = element_text(angle = 90, size = 8, vjust = 0.5))
ggsave("./output/GAM_minor_sqrt.png", height = 5, width = 11, units = "in")

ggplot(predict_minor.bin, aes(date, elev.mean, fill = pred)) +
  geom_raster() +
  geom_contour(aes(z = pred), color = "black", size = 0.25) +
  scale_fill_viridis_c(name = "log(abund.)") +
  #facet_grid(bb.sp + subject ~ year) +
  facet_grid(year ~ bb.sp + subject) +
  xlab("Date") +
  ylab("Elevation (m)") +
  theme(axis.text.x = element_text(angle = 90, size = 8, vjust = 0.5))
ggsave("./output/GAM_minor_bin.png", height = 5, width = 11, units = "in")
```