### 2020-11-26 analyses/old_scripts/resource_phenology.Rmd

---
title: "floral_resources"
output: html_document
---

### Prepare environment
```{r setup, message=FALSE, warning=FALSE, include=FALSE}
library(ggforce)
library(vegan)
library(tidyverse)
library(lubridate)
library(tidymodels)
```

```{r}
# Network
net <- read_csv("../processed_data/network.csv") %>%
  unite(site.year, site, year, sep = "_", remove = FALSE) %>% # create site.year field that can later be split back up
  mutate(year = factor(year)) %>% # convert year to factor
  left_join(site_data, by = "site") %>%
  left_join(gdd, by = c("site", "date")) %>%
  ungroup() %>%
  arrange(site, date)

# Floral survey
# FL traits
fl_tax <- read_csv("../processed_data/floral_tax.csv")

fl_ktype <- read_csv("../processed_data/floral_k_type.csv") %>%
  select(plant.sp, plant.genus = genus, k.type) %>%
  mutate(k.type.s = str_extract(k.type, "[-+]?[0-9]*\\.?[0-9]+"),
         k.type.ss = str_extract(k.type, "[-+]?[0-9]*"))

fl_morph <- read_csv("../processed_data/flower_morphology_GitaBenadi.csv") %>%
  select(SampleID, plant.family = Family, plant.sp = Species, 
         Unit, Flowers_per_unit, diam = Diameter, LN) %>%
  group_by(plant.sp) %>%
  summarize(diam = mean(diam),
            LN = mean(LN))

fl_ktype_key <- read_delim("../processed_data/kugler_key.tsv", delim = "\t", col_names = FALSE) %>%
  select(k.type = 1, description = 2)

fl_traits <- fl_tax %>%
  full_join(fl_ktype) %>%
  full_join(fl_morph) %>%
  select(plant.sp, plant.genus, plant.family, 
         k.type, k.type.s, k.type.ss, diam, LN)

survey <- read_csv("../processed_data/floral_survey.csv") %>%
  group_by(site, date, plant.sp) %>%
  summarize(flower.cover = sum(flower.cover)) %>%
  group_by(site, date) %>%
  mutate(prop = flower.cover/sum(flower.cover)) %>%
  semi_join(net, by = "plant.sp") %>% # Consider only plants visited (anytime anywhere, not specifically for a given site*date)
  dplyr::select(site, date, plant.sp, flower.cover, prop) %>%
  mutate(yday = yday(date),
         year = year(date)) %>%
  left_join(fl_traits)

top_plants <- survey %>%
  select(site, date, plant.sp, prop) %>%
  complete(fill = list(prop = 0)) %>%
  group_by(plant.sp) %>%
  summarize(mean.prop = mean(prop))

```

```{r}
ggplot(survey, aes(month(date), flower.cover, fill = k.type.ss)) +
  geom_col(position = "fill") +
  facet_grid(year ~ site)
+
  theme(legend.position = "none")
```