### 2020-11-26 analyses/old_scripts/treemap.Rmd

---
title: "trees"
output: html_document
---

```{r}
library(treemapify)
library(tidyverse)
```

# Survey data
```{r}
taxa <- read_csv("../processed_data/floral_tax.csv")

survey <- read_csv("../processed_data/floral_survey.csv") %>%
  filter(flower.cover > 0) %>%
  dplyr::select(site, date, plant.sp, flower.cover) %>%
  separate(plant.sp, c("plant.genus", "plant.sp"), sep = " ") %>%
  left_join(taxa) %>%
  select(site, date, plant.family, plant.genus, plant.sp, flower.cover) %>%
  group_by(site, plant.family, plant.genus, plant.sp) %>%
  summarize(flower.cover = sum(flower.cover))
```

# Tree plot
```{r}
ggplot(survey, aes(area = flower.cover, fill = plant.family)) +
  geom_treemap() +
  theme(legend.position = "none") +
  facet_wrap(~site)

ggplot(survey, aes(area = flower.cover, fill = plant.family, label = plant.family,
                subgroup = plant.family)) +
  geom_treemap()
```