### 2020-11-30 analyses/niche_relations_v6.Rmd
went ordi surfing

---
title: "Floral resource choice by bumble bee species in the Berchtesgadener Alps"
output: html_document
---

# Prepare workspace
## Load packages 
```{r message=FALSE, warning=FALSE, include=FALSE, echo = FALSE}
library(bipartite)
library(mgcv)
library(mgcViz)
library(vegan)
library(ggvegan)
library(ggrepel)
library(treemapify)
#library(RColorBrewer)
library(tidyverse)
library(lubridate)
library(patchwork)
library(tidymodels)
```

## Load data
```{r message=FALSE, warning=FALSE, include=FALSE, echo = FALSE}
# Site data
site_data <- read_csv("../processed_data/site_data.csv") %>%
  dplyr::select(-c(slope.calc, slope.est, elev.min, elev.max, temp.mean)) %>% # drop these variables
  mutate(elev.class2 = factor(elev.class2, levels = c("low", "mid", "high"))) # turn elav.class into an ordered factor

# BB traits
bb_traits <- read_csv("../processed_data/bb_traits.csv")

# FL traits
## family-level taxonomy
fl_tax <- read_csv("../processed_data/floral_tax.csv")

## Kugler morphotypes (accessed through Bioflor)
fl_ktype <- read_csv("../processed_data/floral_k_type.csv")

## Flower colors (accessed through BioFlor)
fl_color <- read_csv("../processed_data/floral_color.csv")

## diameter and corolla depth data from Gita Benadi
fl_morph <- read_csv("../processed_data/flower_morphology_GitaBenadi.csv") %>% 
  dplyr::select(SampleID, plant.family = Family, plant.sp = Species, 
                Unit, Flowers_per_unit, diam = Diameter, LN) %>%
  group_by(plant.sp) %>%
  summarize(diam = mean(diam), LN = mean(LN))

## verbal description of Kugler morphotypes
fl_ktype_key <- read_delim("../processed_data/kugler_key.tsv", delim = "\t", col_names = FALSE) %>%
  dplyr::select(k.type = 1, description = 2)

## join all floral trait tables
fl_traits <- fl_tax %>%
  full_join(fl_ktype) %>%
  full_join(fl_color) %>%
  full_join(fl_morph) %>%
  dplyr::select(plant.sp, plant.genus, plant.family, k.type, k.type.s, k.type.ss, color, diam, LN)

# Network
net <- read_csv("../processed_data/network.csv") %>%
  mutate(year = factor(year)) %>% # convert year to factor
  left_join(site_data, by = "site") %>% # add site data
  ungroup() %>%
  arrange(site, date) %>%
  left_join(fl_traits) %>% # add floral traits
  group_by(site) %>%
  mutate(period = cut_interval(dayofyear, 4, labels = c("early", "mid1", "mid2", "late"))) # divide each elev.class2 into 4 equal time slices

# Floral survey
survey <- read_csv("../processed_data/floral_survey.csv") %>%
  filter(flower.cover > 0) %>% # only count plants in bloom
  semi_join(net, by = "plant.sp") %>% # Consider only plants visited (anytime anywhere, not specifically for a given site*date)
  dplyr::select(site, date, plant.sp, flower.cover) %>%
  left_join(fl_traits) # add floral traits
```

Core species:  
  1. *B. sensu-strictu*
  2. *B. pascuorum*
  3. *B. pratorum*
  4. *B. hortorum*
  5. *B. soroensis*
  6. *B. wurflenii*

```{r fig.cap="**Figure 1: Focal species**", message=FALSE, warning=FALSE, echo=FALSE}
bb_range.year <- net %>%
  left_join(site_data) %>%
  group_by(site, elev.mean, bb.sp, year) %>%
  summarize(abund = n()) %>%
  group_by(bb.sp) %>%
  mutate(elev.floor = min(elev.mean),
         elev.ceiling = max(elev.mean),
         elev.range = elev.ceiling - elev.floor,
         elev.med = elev.floor + ((elev.ceiling - elev.floor)/2)) %>%
  mutate(focal = if_else(
    bb.sp %in% c("bss", "pasc", "prat", "hort", "soro", "wurf"), TRUE, FALSE)
    )

ggplot(bb_range.year, aes(reorder(bb.sp, elev.med), elev.mean, color = focal)) +
  geom_line(size = 2, alpha = 0.25) +
  geom_point(aes(size = abund), alpha = 0.5) +
  facet_wrap(~year, ncol = 1) +
  theme_light(12) +
  scale_size_continuous(name = "Abundance") +
  scale_color_discrete(name = "Focal species") +
  xlab("BB species") +
  ylab("Elevation") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
ggsave("./output/bb_range_year.pdf", height = 6, width = 6)

# Do we see the bimodal abundance pattern predicted by the core-satellite hypothesis? Well, sort of.
bb_abund <- net %>%
  group_by(bb.sp) %>%
  summarize(abund = n())

ggplot(bb_abund, aes(abund)) +
  geom_histogram(binwidth = 250)
```

We will parse the samples by elevation class and season, dividing the samples from each site into equal-interval quarters
```{r}
samples <- net %>%
  filter(bb.sp %in% c("bss", "pasc", "prat", "hort", "soro", "wurf")) %>% # focal species
  #group_by(elev.class2) %>%
  group_by(site) %>%
  mutate(period = cut_interval(dayofyear, 4, labels = c("early", "mid1", "mid2", "late"))) %>%
  group_by(site, period, dayofyear) %>%
  summarize(obs = n()) %>%
  left_join(site_data) %>%
  mutate(date = as_date(dayofyear))

ggplot(samples, aes(as_date(yday(date)), elev.mean, color = period)) +
  #geom_point(size = 1, aes(alpha = obs)) +
  geom_point(size = 2, alpha = 0.75) +
  #facet_wrap(~year) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_alpha_continuous(name = "Observations") +
  scale_x_date(date_labels = "%b", breaks = as_date(c("1970-04-01", "1970-05-01", "1970-06-01",
                                                    "1970-07-01", "1970-08-01", "1970-09-01"))) +
  geom_hline(yintercept = c(900, 1500), linetype = "dashed") +
  theme_light(12) +
  labs(x = "Time",
       y = "Elevation (m x 10)") 
```

## Define functions
```{r, echo = FALSE}
# Fortify envfit() object fitted to BB tongue length data
envfit_fort <- function(x){
  fortify(x) %>%
  select(Score = Type, Label, MDS1, MDS2) %>%
  mutate(Label = case_when(Label == "pbl.w" ~ "T.len"),
         Score = case_when(Score == "Vector" ~ "T.len"))
}

# Fortify PCOA and add the results of an envfit() to BB tongue length
pcoa_fort <- function(x, y){
  fortify(x, axes = 1:2) %>%
  bind_rows(y) %>%
  mutate(Score = case_when(
    Score == "species" ~ "FL.sp",
    Score == "sites" ~ "BB.sp",
    Score == "T.len" ~ "T.len"))
}

# Plot PCoA ordination with arrow for tongue length envfit()
pcoa_plot <- function(x, y, min_load = 0.25) {
  ggplot(filter(x, (abs(MDS1) > min_load | abs(MDS2) > min_load) | (Score %in% c("BB.sp", "T.len")))) + # to avoid overplotting, we can filter to include only floral species > loading threshold
    geom_point(aes(MDS1, MDS2, color = Score)) +
    geom_text_repel(aes(MDS1, MDS2, color = Score, label = Label), key_glyph = "rect") +
    geom_segment(data = y, 
               aes(x = 0, y = 0, xend = MDS1, yend = MDS2, colour = Label),
               arrow = arrow(), size = 2, alpha = 0.5, key_glyph = "rect") +
    theme_light() +
    coord_fixed() # Gavin Simpson says this is **really** important
}

# Generates PCoAs, ADONIS tests, and ordination plots (fully aggregated)
## The ... represents the variables we will use to parse our analysis. For aggregate analysis, just leave the ...
## unpopulated, e.g. niche_analysis(net, bb_traits, plant.sp). 
## To parse by elev.class2: niche_analysis(net, bb_traits, plant.sp)
## To parse by  elev.class2, period, and pollen: niche_analysis(net, bb_traits, plant.sp, period, pollen)
## You get the idea. You should anyway -- you wrote it. https://tidyeval.tidyverse.org/multiple.html
niche_analysis <- function(x, y, criterion, ...) {

  criterion <- enquo(criterion) # this argument toggles taxonomic (plant.sp) vs. morphological (k.type.s) analysis

  x %>%
    filter(caste != "male") %>%
    filter(!bb.sp %in% c("psit", "gers", "humi")) %>% # omit parasitic species, gers because it's a specialist, humi because it's so rare
    group_by(..., bb.sp, !!criterion) %>% # pooling all sites and years
    summarize(freq = n()) %>% # get interaction frequency for each BB-FL pair
    group_by(..., bb.sp) %>%
    mutate(prop = freq/sum(freq)) %>% # get proportional interaction frequency for each BB-FL pair
    dplyr::select(-freq) %>% # drop freq variable
    group_by(...) %>%
    nest() %>% # nest into list-column
    mutate(comm1 = map(data, # spread into the kind of array vegan likes
                       function(x) pivot_wider(x, names_from = !!criterion, values_from = prop) %>%
                         replace(is.na(.), 0)
                       ), 
           comm2 = map(comm1, # column to rowname
                       function(x) column_to_rownames(x, var = "bb.sp")
                       ),
           pcoa = map(comm2, # call PCoA
                      function(x) capscale(x ~ 1, distance = "horn")
                      ),
           traits = map(comm1, # join BB traits
                        function(x) x %>% 
                          left_join(y) %>% # y is the trait table, an argument in the main function call
                          column_to_rownames(var = "bb.sp") %>% 
                          select(pbl.w) # pbl.w is the variable I'm interested in
                        ),
           adonis = map2(comm2, traits, # run adonis permuatation test for distance ~ tongue class
                         function(x, y) adonis(x ~ pbl.w, data = y, method = "horn")
                         ),
           envfit = map2(pcoa, traits, function(x, y) envfit(x, y)),
           envfit.fort = map(envfit, envfit_fort),
           pcoa.fort = map2(pcoa, envfit.fort, pcoa_fort),
           pcoa.plot = map2(pcoa.fort, envfit.fort, pcoa_plot), 
           envfit.r2 = map(envfit, function(x) x$vectors$r),
           envfit.p = map(envfit, function(x) x$vectors$pvals),
           adonis.r2 = map(adonis, # extract adonis r2
                           function(x) x$aov.tab$R2[[1]]
                           ),
           adonis.p = map(adonis, # extract adonis p-value
                          function(x) x$aov.tab$'Pr(>F)'[[1]]
                          ),
           ordisurf = map2(pcoa, traits, function(x, y) ordisurf(x ~ pbl.w, data = y, k = 8)),
           ordisurf.dev_exp = map(ordisurf, function(x) summary(x)$dev.expl),
           ordisurf.p = map(ordisurf, function(x) summary(x)$s.pv),
           MDS1 = map(pcoa, # extract variance explained by MDS1
                      function(x) x$CA$eig[[1]] / sum(x$CA$eig)
                      ), 
           MDS2 = map(pcoa, # extract variance explained by MDS2
                      function(x) x$CA$eig[[2]] / sum(x$CA$eig)
                      )) %>%
    unnest(cols = c(ordisurf.dev_exp, ordisurf.p, envfit.r2, envfit.p, adonis.r2, adonis.p, MDS1, MDS2)) %>% # these are lists of length = 1, so unnesting them simply displays the value
    select(..., comm2, traits, pcoa, ordisurf, ordisurf.dev_exp, ordisurf.p, envfit, envfit.fort, pcoa.fort, pcoa.plot, envfit.r2, envfit.p, adonis.r2, adonis.p, MDS1, MDS2)
}

ordi1 <- ordisurf(niche.tax$pcoa[[1]] ~ pbl.w, data = niche.tax$traits[[1]])

# Calculate site-wise beta diversity from visitation data
beta_func <- function(x, traits, sitedata, level, criterion, binary, ...) {

  criterion <- enquo(criterion) # plant.sp or k.type.s
  level <- enquo(level) # bb.sp or pbl.fac (i.e. tongue length  guild)

  x %>% 
    filter(caste != "male") %>%
    filter(bb.sp %in% c("bss", "pasc", "prat", "hort", "soro", "wurf")) %>% # focal species
    left_join(traits) %>% # add BB traits
    group_by(site, !!level, !!criterion, ...) %>% # in addition to grouping by level (bb.sp/pbl.fac) and criterion (plant.sp/k.type.s), the ... argument will also group by any variables supplied in the function call, e.g. elev.class2, period, year.
    summarize(freq = n()) %>% # get visitation frequency
    group_by(site, !!level, ...) %>%
    mutate(prop = freq/sum(freq)) %>% # get visitation proportions
    group_by(!!level, ...) %>%
    select(-freq) %>%
    nest() %>% # nest into list-column
    mutate(comm = map(data, function(x) pivot_wider(x, names_from = !!criterion, values_from = prop, values_fill = 0)), # convert to community matrix
           comm2 = map(comm, function(x) column_to_rownames(x, var = "site")), # column to rowname
           dist = map(comm2, function(x) vegdist(x, method = "horn", binary = binary)), # conver to Horn's distance matrix
           tidy.dist = map(dist, tidy), # convert matrix to tidy array
           elevation = map(comm, function(x) left_join(x, sitedata, by = "site") %>% # add elevation data
                             select(site, elev.mean) %>% 
                             column_to_rownames(var = "site")),
           adonis = map2(dist, elevation, function(x, y) adonis(x ~ elev.mean, data = y)), # run adonis test
           adonis.r2 = map(adonis, function(x) x$aov.tab$R2[[1]]), # extract adonis r2
           adonis.p = map(adonis, function(x) x$aov.tab$'Pr(>F)'[[1]])) %>% # extract adonis p-val
    unnest(c(tidy.dist, adonis.r2, adonis.p)) %>% 
    select(..., !!level, site1 = item1, site2 = item2, distance, adonis.r2, adonis.p) %>%
    left_join(select(site_data, site, elev.mean), by = c("site1" = "site")) %>%
    rename(elev1 = elev.mean) %>%
    left_join(select(site_data, site, elev.mean), by = c("site2" = "site")) %>%
    rename(elev2 = elev.mean) %>%
    mutate(elev.diff = abs(elev1 - elev2)) 
}

# Calculate site-wise beta diversity from survey data
beta_func_fl <- function(x, sitedata, criterion, binary, ...) {

  criterion <- enquo(criterion) # plant.sp or k.type.s

  x %>% 
    group_by(site, !!criterion, ...) %>% # 
    summarize(abund = sum(flower.cover)) %>% # get visitation frequency
    group_by(site, ...) %>%
    mutate(prop = abund/sum(abund)) %>% # get visitation proportions
    select(-abund) %>%
    pivot_wider(names_from = !!criterion, values_from = prop, values_fill = 0) %>%
    column_to_rownames(var = "site") %>%
    vegdist(method = "horn", binary = binary) %>%
    tidy() %>%
    select(..., site1 = item1, site2 = item2, distance) %>%
    left_join(select(site_data, site, elev.mean), by = c("site1" = "site")) %>%
    rename(elev1 = elev.mean) %>%
    left_join(select(site_data, site, elev.mean), by = c("site2" = "site")) %>%
    rename(elev2 = elev.mean) %>%
    mutate(elev.diff = abs(elev1 - elev2)) 
}

# Niche overlap by sp.pair * site * year
niche_overlapper <- function(x, criterion) {

  criterion <- enquo(criterion) # criterion is set to either plant.sp (for taxonomic) or k.type.s (for morphological)

  x %>%
    filter(caste != "male") %>%
    filter(bb.sp %in% c("bss", "pasc", "prat", "hort", "soro", "wurf")) %>% # focal species
    group_by(bb.sp, !!criterion, site, year) %>% # group by site-year and BB-FL pair
    summarize(freq = n()) %>% # get interaction frequency for each BB-FL pair per site-year
    group_by(bb.sp, site, year) %>% # group by site-year and BB species
    mutate(prop = freq/sum(freq)) %>% # get proportional interaction frequency for each BB-FL pair
    dplyr::select(-freq) %>% # drop freq variable
    pivot_wider(names_from = !!criterion, values_from = prop, values_fill = 0) %>% # spread into the kind of array vegan likes
    group_by(site, year) %>%
    nest() %>% # nest into list-column
    mutate(data2 = map(data, function(x) column_to_rownames(x, var = "bb.sp")), # column to rownames
           dist = map(data2, function(x) vegdist(x, method = "horn")), # convert to Horn's distance matrix
           tidy_dist = map(dist, tidy)) %>% # convert to tidy array of Horn's distances
    dplyr::select(site, year, tidy_dist) %>% # drop intermediate list-columns
    unnest(cols = c(tidy_dist)) %>% # unnest back into a single data frame
    left_join(site_data) %>% # add site data
    dplyr::select(site, year, bb.sp1 = item1, bb.sp2 = item2, distance, elev.mean) %>% # dplyr::select and rename vars
    left_join(bb_traits, by = c("bb.sp1" = "bb.sp")) %>% # add traits for bb.sp1
    rename(subgenus1 = subgenus, pbl.w1 = pbl.w, pbl.w.class1 = pbl.w.class, pbl.fac1 = pbl.fac) %>%
    left_join(bb_traits, by = c("bb.sp2" = "bb.sp")) %>% # add traits for bb.sp2
    rename(subgenus2 = subgenus, pbl.w2 = pbl.w, pbl.w.class2 = pbl.w.class, pbl.fac2 = pbl.fac) %>%
    dplyr::select(site, year, bb.sp1, bb.sp2, subgenus1, subgenus2, distance, 
           pbl.w1, pbl.w2, pbl.w.class1, pbl.w.class2, pbl.fac1, pbl.fac2, elev.mean) %>%
    mutate(pbl.diff = abs(pbl.w1 - pbl.w2), # tongue length differences
           pbl.fac.diff = if_else(pbl.fac1 == pbl.fac2, "same", "different"),
           subgenus.diff = if_else(subgenus1 == subgenus2, "same", "different"), # binary: same or different subgenus

           # Logical columns indicating whether a given species belongs to a given species pair
           # This is handy for extracting data for individual species across all their counterparts
           bss = if_else(bb.sp1 == "bss" | bb.sp2 == "bss", TRUE, FALSE),
           pasc = if_else(bb.sp1 == "pasc" | bb.sp2 == "pasc", TRUE, FALSE),
           wurf = if_else(bb.sp1 == "wurf" | bb.sp2 == "wurf", TRUE, FALSE),
           prat = if_else(bb.sp1 == "prat" | bb.sp2 == "prat", TRUE, FALSE),
           soro = if_else(bb.sp1 == "soro" | bb.sp2 == "soro", TRUE, FALSE),
           hort = if_else(bb.sp1 == "hort" | bb.sp2 == "hort", TRUE, FALSE)) %>%

    unite(sp.pair, c(bb.sp1, bb.sp2), sep = ":") %>% # create a single species pair column...
    mutate(sp.pair = factor(sp.pair)) %>% # ...and make it a factor so that mgcv can handle it
    mutate(niche.overlap = 1-distance) # express niche overlap so that 1 = perfect overlap and 0 = no overlap
}

# Diet breadth analysis
diet_breit <- function(x, y, z, criterion, indices = c("normalised degree")) {

  criterion <- enquo(criterion) # plant.sp or k.type.s

  a <- x %>%
    filter(caste != "male") %>%
    filter(bb.sp %in% c("bss", "pasc", "prat", "hort", "soro", "wurf")) %>% # focal species
    left_join(y) %>%
    #group_by(elev.class2) %>%
    group_by(site) %>%
    mutate(period = cut_interval(dayofyear, 4, labels = c("early", "mid1", "mid2", "late"))) %>%
    group_by(site, period, pbl.fac, !!criterion) %>% # group by site, date, bee species, plant species
    summarize(freq = n()) %>% # calculate interaction frequency per species pair within each site
    unite(webID, c(site, period), sep = "_") %>%
    ungroup() %>%
    dplyr::select(higher = pbl.fac, lower = !!criterion, webID, freq) %>% # rename columns to match bipartite's expectations
    data.frame() %>% # convert to data frame
    frame2webs() %>% # convert to to bipartite web
    map(specieslevel, level = "higher", index = indices) %>%
    map(rownames_to_column) %>%
    map(as_tibble) %>%
    bind_rows(.id = "webID") %>% # Collapse list into big data frame
    rename(pbl.fac = rowname) %>%
    separate(webID, c("site", "period"), sep =  "_") %>%
    left_join(z) %>%
    filter(pbl.fac %in% c("long", "short")) %>% # for some reason, a few rows have a "1" for pbl.fac, and I don't know where this comes from; this gets rid of it, but I would still like to know why
    mutate(site = factor(site),
           period = factor(period, levels = c("early", "mid", "late")))

  b <- x %>%
    filter(caste != "male") %>%
    filter(bb.sp %in% c("bss", "pasc", "prat", "hort", "soro", "wurf")) %>% # focal species
    #group_by(elev.class2) %>%
    group_by(site) %>%
    mutate(period = cut_interval(dayofyear, 4, labels = c("early", "mid1", "mid2", "late"))) %>%
    left_join(bb_traits) %>%
    group_by(site, period, pbl.fac) %>% # group by site, date, bee species, plant species
    summarize(abund = n()) %>%
    group_by(site, period) %>%
    mutate(prop = abund/sum(abund))

  out <- left_join(a, b)
}

diet_breit2 <- function(x, y, z, criterion, indices = c("normalised degree")) {

  criterion <- enquo(criterion) # plant.sp or k.type.s

  a <- x %>%
    filter(caste != "male") %>%
    filter(bb.sp %in% c("bss", "pasc", "prat", "hort", "soro", "wurf")) %>% # focal species
    left_join(y) %>%
    # group_by(elev.class2) %>%
    # mutate(period = cut_interval(dayofyear, 4, labels = c("early", "mid1", "mid2", "late"))) %>%
    #group_by(site, period, pbl.fac, !!criterion) %>% # group by site, date, bee species, plant species
    group_by(site, pbl.fac, !!criterion) %>% # group by site, date, bee species, plant species
    summarize(freq = n()) %>% # calculate interaction frequency per species pair within each site
    #unite(webID, c(site, period), sep = "_") %>%
    ungroup() %>%
    #dplyr::select(higher = pbl.fac, lower = !!criterion, webID, freq) %>% # rename columns to match bipartite's expectations
    dplyr::select(higher = pbl.fac, lower = !!criterion, webID = site, freq) %>% # rename columns to match bipartite's expectations
    data.frame() %>% # convert to data frame
    frame2webs() %>% # convert to to bipartite web
    map(specieslevel, level = "higher", index = indices) %>%
    map(rownames_to_column) %>%
    map(as_tibble) %>%
    bind_rows(.id = "webID") %>% # Collapse list into big data frame
    #rename(pbl.fac = rowname) %>%
    rename(pbl.fac = rowname,
           site = webID) %>%
    #separate(webID, c("site", "period"), sep =  "_") %>%
    left_join(z) %>%
    filter(pbl.fac %in% c("long", "short")) %>% # for some reason, a few rows have a "1" for pbl.fac, and I don't know where this comes from; this gets rid of it, but I would still like to know why
    # mutate(site = factor(site),
    #        period = factor(period, levels = c("early", "mid", "late")))
    mutate(site = factor(site))

  b <- x %>%
    filter(caste != "male") %>%
    filter(bb.sp %in% c("bss", "pasc", "prat", "hort", "soro", "wurf")) %>% # focal species
    # group_by(elev.class2) %>%
    # mutate(period = cut_interval(dayofyear, 4, labels = c("early", "mid1", "mid2", "late"))) %>%
    left_join(bb_traits) %>%
   # group_by(site, period, pbl.fac) %>% # group by site, date, bee species, plant species
    group_by(site, pbl.fac) %>% # group by site, date, bee species, plant species
    summarize(abund = n()) %>%
    #group_by(site, period) %>%
    group_by(site) %>%
    mutate(prop = abund/sum(abund))

  filter(caste != "male") %>%out <- left_join(a, b)
}

# Null model generator
nullify1 <- function(x, n, criterion){

  criterion <- enquo(criterion) # plant.sp or k.type.s

  x %>%
    filter(caste != "male") %>%
    filter(bb.sp %in% c("bss", "pasc", "prat", "hort", "soro", "wurf")) %>% # focal species
    select(site, bb.sp, !!criterion) %>% 
    group_by(site, bb.sp, !!criterion) %>%
    summarize(freq = n()) %>% # get interaction frequency for each BB-FL pair
    group_by(site) %>%
    nest() %>% # nest data into list-column grouped by pollen, site, date
    mutate(data = map(data, function(x) pivot_wider(x, 
                                                    names_from = !!criterion, 
                                                    values_from = freq, 
                                                    values_fill = 0) %>% column_to_rownames("bb.sp")), # turn table into community matrix
           data = map(data, function(x) vegan::permatfull(x, 
                                                          fixedmar = "rows", 
                                                          shuffle = "ind", 
                                                          times = n)), # permute community matrices
           data = map(data, function(x) return(x$perm))) %>% # extract permutations from permat objects
    unnest(data) %>% # unnest
    mutate(data = map(data, function(x) vegdist(x, method = "horn")), # convert each permutation into distance matrix
           data = map(data, tidy)) %>% # convert distance matrices to tidy arrays
    group_by(site) %>%
    mutate(iter = row_number()) %>% # add index to denote iterations
    unnest(cols = data) %>% # unnest again
    rename(bb.sp1 = item1, bb.sp2 = item2) %>%
    unite(sp.pair, c(bb.sp1, bb.sp2), sep = ":") %>% # create a single species pair column...
    mutate(sp.pair = factor(sp.pair), # ...and make it a factor so that mgcv can handle it
           niche.overlap = 1-distance) %>% # express niche overlap so that 1 = perfect overlap and 0 = no overlap
    select(site, sp.pair, iter, niche.overlap.sim = niche.overlap)
}

# Null model generator
nullify2 <- function(x, n, criterion){

  criterion <- enquo(criterion) # plant.sp or k.type.s

  x %>%
    filter(caste != "male") %>%
    filter(bb.sp %in% c("bss", "pasc", "prat", "hort", "soro", "wurf")) %>% # focal species
    select(site, bb.sp, !!criterion) %>% 
    group_by(site, bb.sp, !!criterion) %>%
    summarize(freq = n()) %>% # get interaction frequency for each BB-FL pair
    group_by(site) %>%
    nest() %>% # nest data into list-column grouped by pollen, site, date
    mutate(data = map(data, function(x) pivot_wider(x, 
                                                    names_from = !!criterion, 
                                                    values_from = freq, 
                                                    values_fill = 0) %>% column_to_rownames("bb.sp")), # turn table into community matrix
           data = map(data, function(x) vegan::permatfull(x, 
                                                          fixedmar = "both", 
                                                          shuffle = "ind", 
                                                          times = n)), # permute community matrices
           data = map(data, function(x) return(x$perm))) %>% # extract permutations from permat objects
    unnest(data) %>% # unnest
    mutate(data = map(data, function(x) vegdist(x, method = "horn")), # convert each permutation into distance matrix
           data = map(data, tidy)) %>% # convert distance matrices to tidy arrays
    group_by(site) %>%
    mutate(iter = row_number()) %>% # add index to denote iterations
    unnest(cols = data) %>% # unnest again
    rename(bb.sp1 = item1, bb.sp2 = item2) %>%
    unite(sp.pair, c(bb.sp1, bb.sp2), sep = ":") %>% # create a single species pair column...
    mutate(sp.pair = factor(sp.pair), # ...and make it a factor so that mgcv can handle it
           niche.overlap = 1-distance) %>% # express niche overlap so that 1 = perfect overlap and 0 = no overlap
    select(site, sp.pair, iter, niche.overlap.sim = niche.overlap)
}
```

# 1. What is the logic of niche partitioning between core species, and does it change through elevation and time?

For ordination-based analyses, I have opted to pool across years to emphasize species relationships rather than interannual differences. To aid interpretation, I have classified BB species by tongue length. The two "true" long-tongued species are *B. pascuorum* and *B. hortorum*, with the latter having a considerably longer tongue than the former, but I have opted to consider *B. wurflenii* as "behaviorally long-tongued" because it characteristically forages as a robber, and evidently converges in flower choice with *B. pascuorum* and *B. hortorum*. This decision was made *post-hoc* after noting the clustering patterns.

I have also opted to not distinguish between putative pollen and putative nectar foragers. First, the classification of nectar vs. pollen foragers based on whether they are carrying pollen at the time of observation is imperfect (though probably still informative). Second, and more importantly in my opinion, the more I break the data up, the more difficult it becomes to visualize and interpret the data. Finally, when I ran the analysis parsing by pollen vs. nectar foraging, the patterns tended to be similar between the two.

Note that in this version of the analysis, I am not excluding observations that were unclassified with respect to pollen carrying. A closer inspection of the data revealed that this is simply how Kalli handled males and Psithyrus queens, since they never collect pollen. I am omitting Psithyrus from my analysis, but I am not including the males of the focal specis. I suppose an argument could be made that parsing by caste would be informative, but again, there are so many ways these data could be parsed, and I can't use all of them without whittling the data down to something intractable.

Kugler classes are coded in the form of number.decimal.letter.letter. I have chosen to simplify to the level of number.decimal.

To avoid overplotting plant species / Kugler classes, I have only plotted plant species scoring above a given threshold on either the first or second axis. Thus, the plotted species are those that contribute most to the separation, not necessarily those that contribute most to diet. A species that is visited heavily by all BB species will fall close to (0,0) and will therefore not be plotted. This is because they are added based on weighted averaging. i.e. they get "pulled" toward BB species in proportion to their (relative) abundance (i.e. visitation rate) for each species.

## 1.1 Ordination by species with post-hoc fitting of tongue length  
### Script
```{r message=FALSE, warning=FALSE, echo = FALSE}
# By plant.sp
niche.tax <- niche_analysis(net, bb_traits, criterion = plant.sp.abb)
niche.tax.plot <- niche.tax$pcoa.plot[[1]]

# By ktype
niche.morph <- niche_analysis(net, bb_traits, criterion = k.type.s)
niche.morph.plot <- niche.morph$pcoa.plot[[1]]

## Taxonomic and morphological together
pdf("../analyses/output/niche_pcoa.pdf", height = 8.5, width = 11)
(niche.tax.plot + ggtitle("Taxonomic") | niche.morph.plot + ggtitle("Morphological")) + plot_layout(guides = 'collect') 
dev.off()

# Playing with ordisurf as alternative to envfit()
library(ggordiplots)

## Based on https://oliviarata.wordpress.com/2014/07/17/ordinations-in-ggplot2-v2-ordisurf/
# Get ordination object
ordi1 <- niche.tax$ordisurf[[1]] 
ordi1.grid <- ordi1$grid #extracts the ordisurf object
ordi1.mite <- expand.grid(x = ordi1.grid$x, y = ordi1.grid$y) #get x and ys
ordi1.mite$z <- as.vector(ordi1.grid$z) #unravel the matrix for the z scores
ordi1.mite.na <- data.frame(na.omit(ordi1.mite)) %>% #gets rid of the nas
  as_tibble() %>%
  rename(MDS1 = x, MDS2 = y)

# Get scores
scores <- fortify(niche.tax$pcoa[[1]], axes = 1:2) %>%
  mutate(Score = case_when(
    Score == "species" ~ "FL.sp",
    Score == "sites" ~ "BB.sp")
    ) %>%
  as_tibble()

# Plot
dat <- filter(scores, (abs(MDS1) > 0.25 | abs(MDS2) > 0.25) | Score == "BB.sp")

ggplot(dat, aes(MDS1, MDS2, pch = Score)) +
  geom_contour(data = ordi1.mite.na, aes(MDS1, MDS2, z = z, color = stat(level)), size = 1.5, alpha = 0.75, inherit.aes = FALSE) +
  geom_point(size = 2) +
  geom_text_repel(aes(MDS1, MDS2, label = Label), key_glyph = "rect") +
  theme_light() +
  scale_color_viridis_c(name = "Tongue length") +
  scale_shape_manual(values = c(16, 1), name = "Level") +
  coord_fixed() # Gavin Simpson says this is **really** important
```

### Results

In parentheses following the initial mention of each MDS axis, the variance explained is shown for the taxonomic (T:) and morphological (M:) ordinations.

1. MDS1 (T:48%, M:86%) separates long-tongued from short--tongued species. This mirrors the findings of Goulson et al. (2008). MDS2 (T:27%, M:12%) mainly accounts for the divergent foraging behavior of *B. sensu-stricto*. I wonder if this has been described before? A caveat to the interpretation of this pattern, though, is that *B. sensu-stricto* exhibited some weird temporal abundance patterns, particularly in 2012. It's possible that this biased its foraging in a way that could account for some of its apparent deviance.

2. LT species cluster around the lip flowers *Salvia glutinosa* and *Rhinanthus glacialis* (Kugler 5.1) and the flag flower *Anthyllis vulneraria* (Kugler 6.1). ST species *B. soroensis* and *B. pratorum* cluster around the bell flower *Campanula scheuchzeri* (Kugler 3.2) and flower heads *Carduus defloratus* and *Centaurea jacea* (Kugler 7.2). The short-tongued *B. sensu-stricto* seems to be especially associated with the bell flower *Erica carnea* (Kugler 3.1), which scores near 0 on MDS1 (i.e. not accounting for the separation of tongue-legnth guilds) but strongly negative on MDS2, mirroring the scoring *B. sensu-stricto*. The lip flower *Stachys alopecuros* (Kugler 5.1) also scores near 0 on MDS1 but high on MDS2, and together with *Erica carnea* seems to contribute to the separation of *B. sensu-strictu* from the other species of both tongue classes.

## 1.2 Ordination by species with post-hoc fitting of tongue length, broken down by elevation class
### Script
```{r message=FALSE, warning=FALSE, echo = FALSE}
# Taxonomic
niche_x_elev.tax <- niche_analysis(net, bb_traits, criterion = plant.sp.abb, elev.class2)

## Extract and name plots
niche_high.tax.plot <- filter(niche_x_elev.tax, elev.class2 == "high")$pcoa.plot[[1]] + ggtitle("high/Tax")
niche_mid.tax.plot <- filter(niche_x_elev.tax, elev.class2 == "mid")$pcoa.plot[[1]] + ggtitle("mid/Tax")
niche_low.tax.plot <- filter(niche_x_elev.tax, elev.class2 == "low")$pcoa.plot[[1]] + ggtitle("low/Tax")

# Morphological
niche_x_elev.morph <- niche_analysis(net, bb_traits, criterion = k.type.s, elev.class2)

## Extract and name plots
niche_high.morph.plot <- filter(niche_x_elev.morph, elev.class2 == "high")$pcoa.plot[[1]] + ggtitle("high/morph")
niche_mid.morph.plot <- filter(niche_x_elev.morph, elev.class2 == "mid")$pcoa.plot[[1]] + ggtitle("mid/morph")
niche_low.morph.plot <- filter(niche_x_elev.morph, elev.class2 == "low")$pcoa.plot[[1]] + ggtitle("low/morph")

## Display plots on grid
pdf("../analyses/output/niche_x_elev_pcoa.pdf", height = 8.5, width = 11)
(niche_high.tax.plot | niche_high.morph.plot) /
(niche_mid.tax.plot | niche_mid.morph.plot) /
(niche_low.tax.plot | niche_low.morph.plot) + plot_layout(guides = 'collect') 
dev.off()
```

### Results  
1. At low elevation, MDS1 (T:63%, M:64%) does *not* clearly separate LT from ST species; instead, it seems to be driven by the strong association between *B. soroensis* and the flower head *Centaurea jacea* (Kugler 7.2), which score high on MDS1. MDS2 (T:34%. M:35%) seems to be driven by the associated of *B. pratorum* and *B. hortorum* with the stalk-disc flower *Primula elatior* (Kugler 4.1). In general, though, the BB species at low elevation do not seem to cluster into clear groups.

2. At mid elevation, MDS1 (T:48%, M:79%) seems to be strongly driven by tongue-length. LT species are associated with the lip flowers *Salvia glutinosa* and *Rhinanthus glacialis* (Kugler 5.1) and the flag flower *Anthyllis vulneraria* (Kugler 6.1). ST species are associated with the flower head *Centaurea jacea* (Kugler 7.2). MDS2 (T:25%, M:16%) seems to be driven by the divergent foraging of *B. sensu-stricto*, particularly its association with the bell flower *Erica carnea* (Kugler 3.1), though *E. carnea* plots centrally because it was also visited by other BB species.

3. At high elevation, MDS1(T:52%, M:79%) again separates LT and ST species (though not as strongly as at mid elevation) while MDS2 (T:38%, M:18%) separates *B. sensu-stricto*. In this case, though, *B. sensu-stricto* seems to be especially associated with the disk flower *Helianthemum nummularium* (Kugler 1.1). LT species are associated with the lip flower *Stachys alopecuros* (Kugler 5.1) and the flag flower *Anthyllis vulneraria* (Kugler 6.1), while the ST species *B. pratorum* and *B. soroensis* are associated with the bell flower *Campanula scheuchzeri* (Kugler 3.2) and flower head *Carduus defloratus* (Kugler 7.2).

## 1.3 Ordination by BB species * elev.class2 * period, pooling across years
### Script
```{r message=FALSE, warning=FALSE, echo = FALSE}
# Taxonomic
niche_x_elevperiod.tax <- niche_analysis(net, bb_traits, criterion = plant.sp.abb, elev.class2, period)

## Extract and name plots
niche_high_early.tax.plot <- filter(niche_x_elevperiod.tax, elev.class2 == "high" & period == "early")$pcoa.plot[[1]] 
niche_high_mid1.tax.plot <- filter(niche_x_elevperiod.tax, elev.class2 == "high" & period == "mid1")$pcoa.plot[[1]] 
niche_high_mid2.tax.plot <- filter(niche_x_elevperiod.tax, elev.class2 == "high" & period == "mid2")$pcoa.plot[[1]] 
niche_high_late.tax.plot <- filter(niche_x_elevperiod.tax, elev.class2 == "high" & period == "late")$pcoa.plot[[1]]

niche_mid_early.tax.plot <- filter(niche_x_elevperiod.tax, elev.class2 == "mid" & period == "early")$pcoa.plot[[1]]
niche_mid_mid1.tax.plot <- filter(niche_x_elevperiod.tax, elev.class2 == "mid" & period == "mid1")$pcoa.plot[[1]] 
niche_mid_mid2.tax.plot <- filter(niche_x_elevperiod.tax, elev.class2 == "mid" & period == "mid2")$pcoa.plot[[1]]
niche_mid_late.tax.plot <- filter(niche_x_elevperiod.tax, elev.class2 == "mid" & period == "late")$pcoa.plot[[1]]

niche_low_early.tax.plot <- filter(niche_x_elevperiod.tax, elev.class2 == "low" & period == "early")$pcoa.plot[[1]]
niche_low_mid1.tax.plot <- filter(niche_x_elevperiod.tax, elev.class2 == "low" & period == "mid1")$pcoa.plot[[1]]
niche_low_mid2.tax.plot <- filter(niche_x_elevperiod.tax, elev.class2 == "low" & period == "mid2")$pcoa.plot[[1]]
niche_low_late.tax.plot <- filter(niche_x_elevperiod.tax, elev.class2 == "low" & period == "late")$pcoa.plot[[1]]

# Morphological
niche_x_elevperiod.morph <- niche_analysis(net, bb_traits, criterion = k.type.s, elev.class2, period)

## Extract and name plots
niche_high_early.morph.plot <- filter(niche_x_elevperiod.morph, elev.class2 == "high" & period == "early")$pcoa.plot[[1]]
niche_high_mid1.morph.plot <- filter(niche_x_elevperiod.morph, elev.class2 == "high" & period == "mid1")$pcoa.plot[[1]] 
niche_high_mid2.morph.plot <- filter(niche_x_elevperiod.morph, elev.class2 == "high" & period == "mid2")$pcoa.plot[[1]] 
niche_high_late.morph.plot <- filter(niche_x_elevperiod.morph, elev.class2 == "high" & period == "late")$pcoa.plot[[1]]

niche_mid_early.morph.plot <- filter(niche_x_elevperiod.morph, elev.class2 == "mid" & period == "early")$pcoa.plot[[1]]
niche_mid_mid1.morph.plot <- filter(niche_x_elevperiod.morph, elev.class2 == "mid" & period == "mid1")$pcoa.plot[[1]]
niche_mid_mid2.morph.plot <- filter(niche_x_elevperiod.morph, elev.class2 == "mid" & period == "mid2")$pcoa.plot[[1]] 
niche_mid_late.morph.plot <- filter(niche_x_elevperiod.morph, elev.class2 == "mid" & period == "late")$pcoa.plot[[1]]

niche_low_early.morph.plot <- filter(niche_x_elevperiod.morph, elev.class2 == "low" & period == "early")$pcoa.plot[[1]]
niche_low_mid1.morph.plot <- filter(niche_x_elevperiod.morph, elev.class2 == "low" & period == "mid1")$pcoa.plot[[1]] 
niche_low_mid2.morph.plot <- filter(niche_x_elevperiod.morph, elev.class2 == "low" & period == "mid2")$pcoa.plot[[1]] 
niche_low_late.morph.plot <- filter(niche_x_elevperiod.morph, elev.class2 == "low" & period == "late")$pcoa.plot[[1]]

## Display plots on grid
#pdf("../analyses/output/niche_x_elevperiod_pcoa_tax.pdf", height = 8.5, width = 11)
(niche_high_early.tax.plot | niche_high_mid1.tax.plot  | niche_high_mid2.tax.plot | niche_high_late.tax.plot) /
(niche_mid_early.tax.plot | niche_mid_mid1.tax.plot | niche_mid_mid2.tax.plot | niche_mid_late.tax.plot) /
(niche_low_early.tax.plot | niche_low_mid1.tax.plot  | niche_low_mid2.tax.plot  | niche_low_late.tax.plot)  + 
  plot_layout(guides = 'collect') 
#dev.off()

#pdf("../analyses/output/niche_x_elevperiod_pcoa_morph.pdf", height = 8.5, width = 11)
(niche_high_early.morph.plot | niche_high_mid1.morph.plot  | niche_high_mid2.morph.plot | niche_high_late.morph.plot) /
(niche_mid_early.morph.plot | niche_mid_mid1.morph.plot | niche_mid_mid2.morph.plot | niche_mid_late.morph.plot) /
(niche_low_early.morph.plot | niche_low_mid1.morph.plot  | niche_low_mid2.morph.plot  | niche_low_late.morph.plot)  + 
  plot_layout(guides = 'collect') 
#dev.off()
```

### Results

This is probably the most interpretable analysis. One thing I like about it is how it reveals important floral species that didn't score highly in the aggregate analysis. But there is also a lot to say.

1. *Low elevation*: The most salient pattern is a lack of one -- tongue length does not seem to be important at low elevation *until late in the season*, when all the short ST bees agree that *Centaurea jacea* is the only game in town.

  - *Early*: MDS1 (T:98%, M:97%) dominates the ordination and mainly separates *B. soroensis*, associated with the bell flower *Geum rivale* (Kugler 3.2) from the other BB species, associated with the stalk-disc flower *Primula elatior* (Kugler 4.1). MDS2 (T:2%, M:2%) accounted  for almost no variation.

  - *Mid*: MDS1 (T:90%, M:94%) again dominates the ordination, apparently representing a gradient of association ranging from the lip flower *Rhinanthus alectorolophus* (Kugler 5.1) and the flag blossom *Trifolium pratense* (Kugler 6.1) on the  low end of the axis to the flower head *Knautia arvensis* (Kugler 7.1). BB species are spread fairly evenly along this gradient, with *B. sensu-stricto* and *B. wurflenii* leaning hardest in the *Rhinanthus/Trifolium* direction and *B. pratorum* and *B. pascuorum* leaning hardest in the *Knautia* direction. Interestingly, the little variance MDS2 (T:9%, M:6%) explains seems to be associated with tongue length, though because MDS2 accounts for so little variance, we must conclude that tongue length is of little importance in structuring the visitation network in this elevation-time slice.

  - *Late*: MDS1 (T:51%, M:70%) mainly separates *B. wurflenii* from the rest of the BB species, apparently due to its strong association with the flag blossom *Lotus corniculatus* (Kugler 6.1) and the funnel flower *Gentiana aspera* (Kugler 2.1). MDS2 (T:41%, M:27%), which matters again, seems to separate LT ans ST species, though the pattern is not so much one of separation as of dispersion. All the ST species are huddled tightly around the flower head *Centaurea jacea* (Kugler 7.2), while the LT species are rather spread out, with *B. wurflenii* as described, *B. hortorum* expressing a distinct taste for the lip flower *Stachys officinalis* (Kugler 5.1), and *B. pascuorum* middling as usual.

2. *Mid elevation*:

  - *Early*: MDS1 (T:52%, M:68%) is clearly the tongue-length axis. ST species are all clustered around the bell flower *Erica carnea* (Kugler 3.1), while the LT species are more spread out, with *B. wurflenii* commidd to the flag blossom *Polygala chamaebuxus*, *B. hortorum* more interested in the stalk-disk flower *Gentiana verna* (Kugler 4.1), and *B. pascuorum* falls in between (though a lot closer to *B. wurflenii*), evidentally pulled a bit toward the the stalk-disc flower *Primula elatior* (Kugler 4.1). MDS2 (T:32%, M:29%) seems to be driven strongly by *B. hortorum* divergence from the other two LT species.

  - *Mid*: MDS1 (T:49%, M:68%) separates *B. soroensis* and *B. pratorum* from the other BB species. The former are associated with the flower head *Phyteuma orbiculare* (Kugler 7.1) and the bell flower *Geum rivale* (Kugler 3.2). The latter consist of a fairly tight cluster of *B. hortorum*, *B. pascuorum*, and *B. wurflenii* around the flag flowers *Anthyllis vulneraria* and *Lotus corniculatus* (Kugler 6.1) and the lip flower *Ajuga reptans* (Kugler 5.1). MDS2 (T:26%, M:17%), together with MDS1, accounts for the uniqueness of *B. sensu-strictu* foraging, with that species plotting both very low on MDS1 and very high on MDS2. Interestingly, none of the plant species plot close to it, so it must exceptional not because it forages on something unique but because it forages a little bit on everything. That's the  only explanation I can come up with, anyway, for why the weighted-average plotting pulls plants away from *B. sensu-stricto*. The apparent exceptionality of *B. sensu-stricto* is a theme in this analysis.

  - *Late*: MDS1 (T:67%, M:98%) sort of separates LT and ST species, except that *B. pratorum* and *B. pascuorum* fall close to each other on this axis (but are separated widely on MDS2). The LT species are fairly well-clustered around the funnel flower *Gentiana aspera* (Kugler 2.1), the lip flowers *Salvia glutinosa* and *Rhinanthus glacialis*. The ST species are spreadwidely along MDS2(T:18%, M:1%), with *B. sensu-stricto* and *B. pratorum* associated with the lip flower *Origanum vulgare* and *B. soroensis* (Kugler 5.1) favoring the flower head *Centaurea jacea* (Kugler 7.2).

3. *High elevation*:
  - *Early*: MDS1 (T:54%, M:74%) is sort of tongue-length axis, but not perfectly, again because *B. sensu-stricto* is so abberant. In general, the species are fairly widespread with distincntfloral associations -- consistent with the observed decline in overall niche overlap at high elevation -- with the exception of *B. soroensis* and *B. pratorum*, which cluster tightly around the bell flower *Rhododendron hirsutum* (Kugler 3.1). *B. hortorum* likes the flag flower *Anthyllis vulneraria* (Kugler 6.1), *B. wurflenii* like the flag flower *Lotus corniculatus* (Kugler 6.1), *B. pascuorum* (middling as usual) seems to be associated with both the flag flower *Trifolium pratense* (Kugler 6.1) and the bell flower *Rhododendron hirsutum* (Kugler 3.1), and *B. sensu-stricto* seems slightly partial to disk flower *Trollius europaeus* (Kugler 1.2), though again its position suggests that it is exceptional in its generalism. MDS2 (T:28%, M:23%) is mostly about accounting for the deviance of *B. sensu-stricto*.

  - *Mid*: By mid-season at high elevation, the LT species are tightly clustered with each other and associated with the flag blossom *Trifolium pratense* (Kugler 6.1) and the lip flower *Stachys alopecuros* (Kugler 5.1). The ST species are widely spread, with *B. soroensis* and *B. pratorum* suggetsing some affinity for the flower head *Carduus defloratus* (Kugler 7.2) and the bell flower *Campanula scheuchzeri* (Kugler 3.2). *B. pratorum* plots similarly close to *T. pratense*. Again, *B. sensu-stricto* is off on its own with no defining floral association. MDS1 (T:64%, M:91%), MDS2(T:35%, M:9%).

  - *Late*: Finally, late in the season at high elevation, LT and ST species are separated diagonally with respect to MDS1 (T:75%, M:79%) and MDS2(T:20%, M:18%), with LT species falling high on one or both axes and ST species falling low on one or both axes. LT species are loosely associated with the flag blossom *Trifolium pratense* (Kugler 6.1), the funnel flowers *Gentiana asclepiadea* and congener *Gentiana aspera* (Kugler 2.1) and the flower head **(unusual for LT to be associated with flower head!)** *Carlina acaulus* (Kugler 7.2). *B. soroensis* and *B. pratorum* are associated with the flower heads *Carduus defloratus* (Kugler 7.2) and *Scabiosa lucida* (Kugler 7.1) and the bell flower *Campanula scheuchzeri* (Kugler 3.2). *B. sensu-stricto*, ever off on its own, was most closely associated with the lip flower *Thymus pulegioides* (Kugler 5.1).

## 1.4 A follow up to the question of diet breadth in long- vs. short-tongued BB species
The question is which guild -- long-tongued or short-tongued species -- uses greater proportion of available floral species. We will keep this really, really simple, and calculate the normalized degree for each guild. That is just the number of floral species visited divided by the total number of floral species in the network. It ranges between 0 and 1, and 0 and 1 mean exactly what you would expect them to mean. I guess we'll go ahead and calculate the raw degree, too, just so that we can see the absolute diet breadth.

For starters, I'm going to use diet_breit2, which does not parse by period.

**This dovetails with Goulson's hypothesis that long-tongued bees are more specialized than short-tongued bees.** I can test this.

### Script
```{r}
# Calculate die breadth metrics
diet_breadth.tax <- diet_breit2(net, bb_traits, site_data, criterion = plant.sp,
                               indices = c("degree", "normalised degree")) %>%
  mutate(normalised.degree.abundnorm = normalised.degree/abund,
         normalised.degree.propnorm = normalised.degree/prop,
         degree.abundnorm = degree/abund,
         degree.propnorm = degree/prop)

diet_breadth.morph <- diet_breit2(net, bb_traits, site_data, criterion = k.type.s,
                               indices = c("degree", "normalised degree")) %>%
  mutate(normalised.degree.abundnorm = normalised.degree/abund,
         normalised.degree.propnorm = normalised.degree/prop,
         degree.abundnorm = degree/abund,
         degree.propnorm = degree/prop)

# Relative diet breadth ~ elevation
ggplot(diet_breadth.tax, aes(elev.mean, normalised.degree, color = pbl.fac)) +
  geom_point() +
  geom_smooth(method = "gam", formula = y ~ s(x, k = 5))

ggplot(diet_breadth.morph, aes(elev.mean, normalised.degree, color = pbl.fac)) +
  geom_point() +
  geom_smooth(method = "gam", formula = y ~ s(x, k = 5))

# Absolute diet breadth ~ elevation
ggplot(diet_breadth.tax, aes(elev.mean, degree, color = pbl.fac)) +
  geom_point() +
  geom_smooth(method = "gam", formula = y ~ s(x, k = 5))

ggplot(diet_breadth.morph, aes(elev.mean, degree, color = pbl.fac)) +
  geom_point() +
  geom_smooth(method = "gam", formula = y ~ s(x, k = 5))

# Relative diet breadth / abundance
ggplot(diet_breadth.tax, aes(elev.mean, normalised.degree.abundnorm, color = pbl.fac)) +
  geom_point() +
  geom_smooth(method = "gam", formula = y ~ s(x, k = 5))

ggplot(diet_breadth.morph, aes(elev.mean, normalised.degree.abundnorm, color = pbl.fac)) +
  geom_point() +
  geom_smooth(method = "gam", formula = y ~ s(x, k = 5))

# Relative diet breadth / proportional abundance
ggplot(diet_breadth.tax, aes(elev.mean, normalised.degree.propnorm, color = pbl.fac)) +
  geom_point() +
  geom_smooth(method = "gam", formula = y ~ s(x, k = 5))

ggplot(diet_breadth.morph, aes(elev.mean, normalised.degree.propnorm, color = pbl.fac)) +
  geom_point() +
  geom_smooth(method = "gam", formula = y ~ s(x, k = 5))

# Absolute diet breadth / abundance
ggplot(diet_breadth.tax, aes(elev.mean, degree.abundnorm, color = pbl.fac)) +
  geom_point() +
  geom_smooth(method = "gam", formula = y ~ s(x, k = 5))

ggplot(diet_breadth.morph, aes(elev.mean, degree.abundnorm, color = pbl.fac)) +
  geom_point() +
  geom_smooth(method = "gam", formula = y ~ s(x, k = 5))

# Absolute diet breadth / proportional abundance
ggplot(diet_breadth.tax, aes(elev.mean, degree.propnorm, color = pbl.fac)) +
  geom_point() +
  geom_smooth(method = "gam", formula = y ~ s(x, k = 5))

ggplot(diet_breadth.morph, aes(elev.mean, degree.propnorm, color = pbl.fac)) +
  geom_point() +
  geom_smooth(method = "gam", formula = y ~ s(x, k = 5))

# BB abundance ~ elevation
ggplot(diet_breadth.tax, aes(elev.mean, abund, color = pbl.fac)) +
  geom_point() +
  geom_smooth(method = "gam", formula = y ~ s(x, k = 5))

# BB proportional abundance ~ elevation
ggplot(diet_breadth.tax, aes(elev.mean, prop, color = pbl.fac)) +
  geom_point() +
  geom_smooth(method = "gam", formula = y ~ s(x, k = 5))

# Absolute diet breadth ~ abundance
ggplot(diet_breadth.tax, aes(abund, degree, color = pbl.fac)) +
  geom_point() +
  geom_smooth(method = "lm")

# Relative diet breadth ~ abundance
ggplot(diet_breadth.tax, aes(abund, normalised.degree, color = pbl.fac)) +
  geom_point() +
  geom_smooth(method = "lm")

# Relative diet breadth ~ proportional abundance
ggplot(diet_breadth.tax, aes(prop, normalised.degree, color = pbl.fac)) +
  geom_point() +
  geom_smooth(method = "lm")
```
### Results
In a nutshell, proportional diet breadth is extremely tightly correlated with proportional anbundance. The story here is that LT species are more abundant than ST species at lower elevations, and ST species are more abundant than LT species at higher elevations. This relationship is strong and basically linear, with the relative abundances crossing at mid-elevation.

## 1.5 Another follow-up. How does the morphological composition of the plant community change with elevation? 
To make the graphs readable, we'll simplify Kugler classes to integers.

### Script
```{r}
# Define periods based on the network data
periods <- net %>%
  filter(bb.sp %in% c("bss", "pasc", "prat", "hort", "soro", "wurf")) %>% # only focal species
  #group_by(elev.class2) %>%
  group_by(site) %>%
  mutate(period = cut_interval(dayofyear, 4, labels = c("early", "mid1", "mid2", "late"))) %>%
  ungroup() %>%
  select(site, date, period) %>%
  distinct()

# Calculate the absolute and relative abundance of each Kugler class ~ site * period

## Floral survey (constrained to visited species)
survey_morph_abund <- survey %>%
  left_join(periods) %>%
  fill(period, .direction = "downup") %>%
  mutate(k.type.ss = factor(k.type.ss)) %>%
  group_by(site, period, k.type.ss) %>%
  summarize(abund = sum(flower.cover)) %>%
  group_by(site, period) %>%
  mutate(prop = abund/sum(abund)) %>%
  left_join(site_data)

## Visitation network
### Aggregate
visitation_morph_abund <- net %>%
  filter(bb.sp %in% c("bss", "pasc", "prat", "hort", "soro", "wurf")) %>% # only focal species
  #group_by(elev.class2) %>%
  group_by(site) %>%
  mutate(period = cut_interval(dayofyear, 4, labels = c("early", "mid1", "mid2", "late"))) %>%
  mutate(k.type.ss = factor(k.type.ss)) %>%
  group_by(site, period, k.type.ss) %>%
  summarize(abund = n()) %>%
  group_by(site, period) %>%
  mutate(prop = abund/sum(abund)) %>%
  left_join(site_data)

### By BB species
visitation_morph_abund.sp <- net %>%
  filter(bb.sp %in% c("bss", "pasc", "prat", "hort", "soro", "wurf")) %>% # only focal species
  #group_by(elev.class2) %>%
  group_by(site) %>%
  mutate(period = cut_interval(dayofyear, 4, labels = c("early", "mid1", "mid2", "late"))) %>%
  mutate(k.type.ss = factor(k.type.ss)) %>%
  group_by(site, period, k.type.ss, bb.sp) %>%
  summarize(abund = n()) %>%
  group_by(site, period, bb.sp) %>%
  mutate(prop = abund/sum(abund)) %>%
  left_join(site_data)

### Taxonomic by BB species
visitation_tax_abund.sp <- net %>%
  filter(bb.sp %in% c("bss", "pasc", "prat", "hort", "soro", "wurf")) %>% # only focal species
  #group_by(elev.class2) %>%
  group_by(site) %>%
  mutate(period = cut_interval(dayofyear, 4, labels = c("early", "mid1", "mid2", "late"))) %>%
  mutate(k.type.ss = factor(k.type.ss),
         bb.sp = factor(bb.sp, levels = c("hort", "pasc", "wurf", "bss", "prat", "soro"))) %>%
  group_by(period, elev.class2, plant.genus, k.type.ss, plant.family, bb.sp) %>%
  summarize(abund = n()) %>%
  group_by(elev.class2, period, bb.sp) %>%
  mutate(prop = abund/sum(abund)) %>%
  arrange(elev.class2, period, bb.sp, -prop)

# Morphological composition ~ elevation (survey)
## Relative abundance
ggplot(survey_morph_abund, aes(factor(elev.mean), prop, fill = k.type.ss)) +
  geom_col() +
  facet_wrap(~period)

## Absolute abundance
ggplot(survey_morph_abund, aes(factor(elev.mean), abund, fill = k.type.ss)) +
  geom_col() +
  facet_wrap(~period)

# Morphological composition ~ elevation (visitation)
## Relative abundance 
ggplot(visitation_morph_abund, aes(factor(elev.mean), prop, fill = k.type.ss)) +
  geom_col() +
  facet_wrap(~period)

## Relative abundance ~ species
ggplot(visitation_morph_abund.sp, aes(factor(elev.mean), prop, fill = k.type.ss)) +
  geom_col() +
  facet_grid(bb.sp ~ period)

## Absolute abundance ~ species
ggplot(visitation_morph_abund.sp, aes(factor(elev.mean), abund, fill = k.type.ss)) +
  geom_col() +
  facet_grid(bb.sp ~ period)

## Relative abundance ~ species
ggplot(filter(visitation_tax_abund.sp, period == "early"), aes(factor(elev.mean), prop, fill = names)) +
  geom_col() +
  facet_wrap(~bb.sp)

ggplot(filter(visitation_tax_abund.sp, period == "mid"), aes(factor(elev.mean), prop, fill = names)) +
  geom_col() +
  facet_wrap(~bb.sp)

ggplot(filter(visitation_tax_abund.sp, period == "late"), aes(factor(elev.mean), prop, fill = names)) +
  geom_col() +
  facet_wrap(~bb.sp)

## Absolute abundance ~ species
ggplot(visitation_tax_abund.sp, aes(factor(elev.mean), abund, fill = names)) +
  geom_col() +
  facet_grid(bb.sp ~ period)
```

### Results

## 1.6 Tree plots
This visualization deserves to be pulled out as a stand-alone chunk. It's so good.
```{r}
### Visitation by BB species, parsed by elev.class2 and period
visitation_abund.sp <- net %>%
  filter(caste != "male") %>%
  filter(!bb.sp %in% c("psit", "gers", "humi")) %>% # only focal species
  #group_by(elev.class2) %>%
  group_by(site) %>%
  mutate(period = cut_interval(dayofyear, 4, labels = c("early", "mid1", "mid2", "late"))) %>%
  # mutate(k.type.ss = factor(k.type.ss),
  #        color = factor(color),
  #        bb.sp = factor(bb.sp, levels = c("hort", "pasc", "wurf", "bss", "prat", "soro"))) %>%
  mutate(k.type.ss = factor(k.type.ss),
         color = factor(color),
         bb.sp = factor(bb.sp)) %>%
  group_by(period, elev.class2, plant.sp.abb, plant.genus, k.type.ss, color, plant.family, bb.sp) %>%
  summarize(abund = n()) %>%
  group_by(elev.class2, period, bb.sp) %>%
  mutate(prop = abund/sum(abund)) %>% 
  left_join(bb_traits) %>%
  arrange(elev.class2, period, bb.sp, -prop)

# Color palette

cols = c("0" = "gray60", "1" = "#8DD3C7", "2" = "#FFFFB3", "3" = "#BEBADA",
         "4" = "#FB8072", "5" = "#80B1D3", "6" = "#FDB462", "7" = "#B3DE69",
         "8" = "#FCCDE5", "9" = "#D9D9D9")

cols2 = c("blue" = "cornflowerblue", 
          "yellow" = "gold1", 
          "pink" = "orchid2", 
          "red" = "firebrick1",
         "violet" = "mediumpurple2", 
         "purple" = "purple", 
         "white" = "white", 
         "various colors" = "black",
         "brown" = "sienna",
         "undet" = "gray40", 
         "green" = "olivedrab1", 
         "orange" = "orange", 
         "lilac" = "mediumorchid")

# Early
tp.early <- ggplot(filter(visitation_abund.sp, period == "early"), 
                aes(area = prop, fill = k.type.ss, label = plant.sp.abb)) +
  geom_treemap(color = "white") +
  geom_treemap_text(fontface = "italic", color = "black", place = "centre", "grow" = TRUE, alpha = 0.5) +
  scale_fill_manual(values = cols) + 
  labs(fill = "Morphotype") +
  facet_grid(fct_rev(elev.class2)~reorder(bb.sp, pbl.w)) +
  theme(strip.text = element_text(size = 14))

tp.early.col <- ggplot(filter(visitation_abund.sp, period == "early"), 
                aes(area = prop, fill = color, label = plant.sp.abb)) +
  geom_treemap(color = "white") +
  geom_treemap_text(fontface = "italic", color = "black", place = "centre", "grow" = TRUE, alpha = 0.5) +
  scale_fill_manual(values = cols2) + 
  labs(fill = "Morphotype") +
  facet_grid(fct_rev(elev.class2)~reorder(bb.sp, pbl.w)) +
  theme(strip.text = element_text(size = 14))

# Mid
tp.mid1 <- ggplot(filter(visitation_abund.sp, period == "mid1"), 
                 aes(area = prop, fill = k.type.ss, label = plant.sp.abb)) +
  geom_treemap(color = "white") +
  geom_treemap_text(fontface = "italic", color = "black", place = "centre", "grow" = TRUE, alpha = 0.5) +
  scale_fill_manual(values = cols) + 
  labs(fill = "Morphotype") +
  facet_grid(fct_rev(elev.class2)~reorder(bb.sp, pbl.w)) +
  theme(strip.text = element_text(size = 14))

tp.mid1.col <- ggplot(filter(visitation_abund.sp, period == "mid1"), 
                 aes(area = prop, fill = color, label = plant.sp.abb)) +
  geom_treemap(color = "white") +
  geom_treemap_text(fontface = "italic", color = "black", place = "centre", "grow" = TRUE, alpha = 0.5) +
  scale_fill_manual(values = cols2) + 
  labs(fill = "Morphotype") +
  facet_grid(fct_rev(elev.class2)~reorder(bb.sp, pbl.w)) +
  theme(strip.text = element_text(size = 14))

tp.mid2 <- ggplot(filter(visitation_abund.sp, period == "mid2"), 
                 aes(area = prop, fill = k.type.ss, label = plant.sp.abb)) +
  geom_treemap(color = "white") +
  geom_treemap_text(fontface = "italic", color = "black", place = "centre", "grow" = TRUE, alpha = 0.5) +
  scale_fill_manual(values = cols) + 
  labs(fill = "Morphotype") +
  facet_grid(fct_rev(elev.class2)~reorder(bb.sp, pbl.w)) +
  theme(strip.text = element_text(size = 14))

tp.mid2.col <- ggplot(filter(visitation_abund.sp, period == "mid2"), 
                 aes(area = prop, fill = color, label = plant.sp.abb)) +
  geom_treemap(color = "white") +
  geom_treemap_text(fontface = "italic", color = "black", place = "centre", "grow" = TRUE, alpha = 0.5) +
  scale_fill_manual(values = cols2) + 
  labs(fill = "Morphotype") +
  facet_grid(fct_rev(elev.class2)~reorder(bb.sp, pbl.w)) +
  theme(strip.text = element_text(size = 14))

# Late
tp.late <- ggplot(filter(visitation_abund.sp, period == "late"), 
                  aes(area = prop, fill = k.type.ss, label = plant.sp.abb)) +
  geom_treemap(color = "white") +
  geom_treemap_text(fontface = "italic", color = "black", place = "centre", "grow" = TRUE, alpha = 0.5) +
  scale_fill_manual(values = cols) + 
  labs(fill = "Morphotype") +
  facet_grid(fct_rev(elev.class2)~reorder(bb.sp, pbl.w)) +
  theme(strip.text = element_text(size = 14))

tp.late.col <- ggplot(filter(visitation_abund.sp, period == "late"), 
                  aes(area = prop, fill = color, label = plant.sp.abb)) +
  geom_treemap(color = "white") +
  geom_treemap_text(fontface = "italic", color = "black", place = "centre", "grow" = TRUE, alpha = 0.5) +
  scale_fill_manual(values = cols2) + 
  labs(fill = "Morphotype") +
  facet_grid(fct_rev(elev.class2)~reorder(bb.sp, pbl.w)) +
  theme(strip.text = element_text(size = 14))

# Pair with ordination plots
tp_ord1 <- (
  tp.early / (
    niche_low_early.tax.plot + ggtitle("low") | 
    niche_mid_early.tax.plot + ggtitle("mid") | 
    niche_high_early.tax.plot + ggtitle("high")
    ) + 
    plot_layout(guides = 'collect')
  ) + 
  plot_layout(heights = c(2, 1))
ggsave("./output/tree_ord_early.png", tp_ord1, width = 16, height = 10)

tp_ord2 <- (
  tp.mid1 | (
    niche_high_mid1.tax.plot / 
    niche_mid_mid1.tax.plot / 
    niche_low_mid1.tax.plot
    ) + 
    plot_layout(guides = 'collect')
  ) + 
  plot_layout(widths = c(5, 1))
ggsave("./output/tree_ord_mid1.png", tp_ord2, width = 14, height = 8)

tp_ord3 <- (
  tp.mid2 | (
    niche_high_mid2.tax.plot / 
    niche_mid_mid2.tax.plot / 
    niche_low_mid2.tax.plot
    ) + 
    plot_layout(guides = 'collect')
  ) + 
  plot_layout(widths = c(5, 1))
ggsave("./output/tree_ord_mid2.png", tp_ord3, width = 14, height = 8)

tp_ord4 <- (
  tp.late | (
    niche_high_late.tax.plot / 
    niche_mid_late.tax.plot / 
    niche_low_late.tax.plot
    ) + 
    plot_layout(guides = 'collect')
  ) + 
  plot_layout(widths = c(5, 1))
ggsave("./output/tree_ord_late.png", tp_ord4, width = 14, height = 8)
```

## 1.7 How do abundant generalist bumble bees deal with a dynamic floral community? Based on the ordination analysis, it looks like they maintain fairly steady morphological preferences, but they are taxonomically flexible. Let's take a closer look at that.

A question to consider is whether the shallower slope of morphological turnover is caused by the "reluctance" of BBs to switch morphotypes or lag in the floristic turnover of morphotypes relative to species. Morphotypes might be less "sensitive" than species to elevational change, so you first get species turnover within morphotypes, then with enough elevational chnage you start to see the morphotypes turning over. **I can freaking test this with my floral survey data!**

### Aggregate analysis
```{r}
## Binary
### Taxonomic
#### BB species
beta.tax.sp.bin <- beta_func(net, traits = bb_traits, sitedata = site_data, level = bb.sp, 
                             criterion = plant.sp, binary = TRUE) %>%
  mutate(criterion = rep("tax", n()))

#### Tongnue length guild
beta.tax.guild.bin <- beta_func(net, traits = bb_traits, sitedata = site_data, level = pbl.fac,
                                criterion = plant.sp, binary = TRUE) %>%
  mutate(criterion = rep("tax", n()))

#### Plants
betaFL_tax.bin <- beta_func_fl(survey, sitedata = site_data, criterion = plant.sp, binary = TRUE) %>%
  mutate(criterion = rep("tax", n()))

### Morphological
#### BB species
beta.morph.sp.bin <- beta_func(net, traits = bb_traits, sitedata = site_data, level = bb.sp,
                               criterion = k.type.s, binary = TRUE) %>%
  mutate(criterion = rep("morph", n()))

#### Tongue length guild
beta.morph.guild.bin <- beta_func(net, traits = bb_traits, sitedata = site_data, level = pbl.fac,
                                  criterion = k.type.s, binary = TRUE) %>%
  mutate(criterion = rep("morph", n()))

#### Plants
betaFL_morph.bin <- beta_func_fl(survey, sitedata = site_data, criterion = k.type.s, binary = TRUE) %>%
  mutate(criterion = rep("morph", n()))

#### Collated
beta.guild.bin <- bind_rows(beta.tax.guild.bin, beta.morph.guild.bin)
beta.sp.bin <- bind_rows(beta.tax.sp.bin, beta.morph.sp.bin)
beta.fl.bin <- bind_rows(betaFL_tax.bin, betaFL_morph.bin)

## Not binary (i.e. abundance-weighted)
### Taxonomic
#### BB species
beta.tax.sp.wt <- beta_func(net,  traits = bb_traits, sitedata = site_data, level = bb.sp,
                         criterion = plant.sp, binary = FALSE) %>%
  mutate(criterion = rep("tax", n()))

beta.tax.guild.wt <- beta_func(net,  traits = bb_traits, sitedata = site_data, level = pbl.fac,
                            criterion = plant.sp, binary = FALSE) %>%
  mutate(criterion = rep("tax", n()))

beta.morph.sp.wt <- beta_func(net,  traits = bb_traits, sitedata = site_data, level = bb.sp,
                           criterion = k.type.s, binary = FALSE) %>%
  mutate(criterion = rep("morph", n()))

beta.morph.guild.wt <- beta_func(net,  traits = bb_traits, sitedata = site_data, level = pbl.fac, 
                              criterion = k.type.s, binary = FALSE) %>%
  mutate(criterion = rep("morph", n()))

beta.sp.wt <- bind_rows(beta.tax.sp.wt, beta.morph.sp.wt)
beta.guild.wt <- bind_rows(beta.tax.guild.wt, beta.morph.guild.wt)

#### Plants
betaFL_tax.wt <- beta_func_fl(survey, sitedata = site_data, criterion = plant.sp, binary = FALSE) %>%
  mutate(criterion = rep("tax", n()))
betaFL_morph.wt <- beta_func_fl(survey, sitedata = site_data, criterion = k.type.s, binary = FALSE) %>%
  mutate(criterion = rep("morph", n()))
beta.fl.wt <- bind_rows(betaFL_tax.wt, betaFL_morph.wt)

# Pull out ADONIS results
## Binary
beta.tax.sp.adonis.bin <- beta.tax.sp.bin %>%
  select(bb.sp, adonis.r2, adonis.p) %>%
  distinct()

beta.morph.sp.adonis.bin <- beta.morph.sp.bin %>%
  select(bb.sp, adonis.r2, adonis.p) %>%
  distinct()

beta.tax.guild.adonis.bin <- beta.tax.guild.bin %>%
  select(pbl.fac, adonis.r2, adonis.p) %>%
  distinct()

beta.morph.guild.adonis.bin <- beta.morph.guild.bin %>%
  select(pbl.fac, adonis.r2, adonis.p) %>%
  distinct()

## Not binary
beta.tax.sp.adonis <- beta.tax.sp.wt %>%
  select(bb.sp, adonis.r2, adonis.p) %>%
  distinct()

beta.morph.sp.adonis <- beta.morph.sp.wt %>%
  select(bb.sp, adonis.r2, adonis.p) %>%
  distinct()

beta.tax.guild.adonis <- beta.tax.guild.wt %>%
  select(pbl.fac, adonis.r2, adonis.p) %>%
  distinct()

beta.morph.guild.adonis <- beta.morph.guild.wt %>%
  select(pbl.fac, adonis.r2, adonis.p) %>%
  distinct()

# Plot
## Binary
beta.sp.plot.bin <- ggplot(beta.sp.bin, aes(elev.diff, distance, color = criterion)) +
  geom_point(alpha = 0.05) +
  geom_smooth(method = "gam", formula = y ~ s(x, k = 4), se = FALSE) +
  facet_wrap(~bb.sp) + 
  ggtitle("Species")

beta.guild.plot.bin <- ggplot(beta.guild.bin, aes(elev.diff, distance, color = criterion)) +
  geom_point(alpha = 0.05) +
  geom_smooth(method = "gam", formula = y ~ s(x, k = 4), se = FALSE) +
  facet_wrap(~pbl.fac) + 
  ggtitle("Tongue guild")

(beta.sp.plot.bin | beta.guild.plot.bin) + plot_layout(guides = 'collect')

## Not binary
beta.sp.plot.wt <- ggplot(beta.sp.wt, aes(elev.diff, distance, color = criterion)) +
  geom_point(alpha = 0.05) +
  geom_smooth(method = "gam", formula = y ~ s(x, k = 4), se = FALSE) +
  facet_wrap(~bb.sp) + 
  ggtitle("Species")

beta.guild.plot.wt <- ggplot(beta.guild.wt, aes(elev.diff, distance, color = criterion)) +
  geom_point(alpha = 0.125) +
  geom_smooth(method = "gam", formula = y ~ s(x, k = 4), se = FALSE) +
  facet_wrap(~pbl.fac) + 
  ggtitle("Tongue guild")

(beta.sp.plot.wt | beta.guild.plot.wt) + plot_layout(guides = 'collect')

# Plants
ggplot(beta.fl.bin, aes(elev.diff, distance, color = criterion)) +
  geom_point(alpha = 0.125) +
  geom_smooth(method = "gam", formula = y ~ s(x, k = 4), se = FALSE)

ggplot(beta.fl.wt, aes(elev.diff, distance, color = criterion)) +
  geom_point(alpha = 0.125) +
  geom_smooth(method = "gam", formula = y ~ s(x, k = 4), se = FALSE)

# Plotting visitation beta-diversity with floral survey beta-diversity

beta.sp.plot.bin2 <- ggplot(beta.sp.bin, aes(elev.diff, distance, color = criterion)) +
  #geom_point(alpha = 0.05) +
  geom_smooth(method = "gam", formula = y ~ s(x, k = 4), se = TRUE) +
  geom_smooth(inherit.aes = FALSE, data = beta.fl.wt, 
              aes(elev.diff, distance, color = criterion), 
              linetype = "dashed", se = TRUE,
              method = "gam", formula = y ~ s(x, k = 4)) +
  facet_wrap(~bb.sp) + 
  ggtitle("Species") +
  theme_light()

beta.guild.plot.bin2 <- ggplot(beta.guild.bin, aes(elev.diff, distance, color = criterion)) +
  #geom_point(alpha = 0.125) +
  geom_smooth(method = "gam", formula = y ~ s(x, k = 4), se = TRUE) +
  geom_smooth(inherit.aes = FALSE, data = beta.fl.wt, 
              aes(elev.diff, distance, color = criterion), 
              linetype = "dashed", se = TRUE,
              method = "gam", formula = y ~ s(x, k = 4)) +
  facet_wrap(~pbl.fac) + 
  ggtitle("Tongue guild") +
  theme_light()

(beta.sp.plot.bin2 | beta.guild.plot.bin2) + plot_layout(guides = 'collect')

beta.sp.plot.wt2 <- ggplot(beta.sp.wt, aes(elev.diff, distance, color = criterion)) +
  #geom_point(alpha = 0.05) +
  geom_smooth(method = "gam", formula = y ~ s(x, k = 4), se = TRUE) +
  geom_smooth(inherit.aes = FALSE, data = beta.fl.wt, 
              aes(elev.diff, distance, color = criterion), 
              linetype = "dashed", se = TRUE,
              method = "gam", formula = y ~ s(x, k = 4)) +
  facet_wrap(~bb.sp) + 
  ggtitle("Species") +
  theme_light()

beta.guild.plot.wt2 <- ggplot(beta.guild.wt, aes(elev.diff, distance, color = criterion)) +
  #geom_point(alpha = 0.125) +
  geom_smooth(method = "gam", formula = y ~ s(x, k = 4), se = TRUE) +
  geom_smooth(inherit.aes = FALSE, data = beta.fl.wt, 
              aes(elev.diff, distance, color = criterion), 
              linetype = "dashed", se = TRUE,
              method = "gam", formula = y ~ s(x, k = 4)) +
  facet_wrap(~pbl.fac) + 
  ggtitle("Tongue guild") +
  theme_light()

(beta.sp.plot.wt2 | beta.guild.plot.wt2) + plot_layout(guides = 'collect')

# Plotting visitation beta-diversity against floral survey beta-diversity
beta.sp.bin2 <- beta.sp.bin %>%
  full_join(beta.fl.bin, by = c("site1", "site2", "elev1", "elev2", "elev.diff")) %>%
  rename(distance.vis = distance.x, distance.surv = distance.y, 
         criterion.vis = criterion.x, criterion.surv = criterion.y)

beta.guild.bin2 <- beta.guild.bin %>%
  full_join(beta.fl.bin, by = c("site1", "site2", "elev1", "elev2", "elev.diff")) %>%
  rename(distance.vis = distance.x, distance.surv = distance.y, 
         criterion.vis = criterion.x, criterion.surv = criterion.y)

beta.sp.wt2 <- beta.sp.wt %>%
  full_join(beta.fl.wt, by = c("site1", "site2", "elev1", "elev2", "elev.diff")) %>%
  rename(distance.vis = distance.x, distance.surv = distance.y, 
         criterion.vis = criterion.x, criterion.surv = criterion.y)

beta.guild.wt2 <- beta.guild.wt %>%
  full_join(beta.fl.wt, by = c("site1", "site2", "elev1", "elev2", "elev.diff")) %>%
  rename(distance.vis = distance.x, distance.surv = distance.y, 
         criterion.vis = criterion.x, criterion.surv = criterion.y)

ggplot(beta.sp.bin2, aes(distance.surv, distance.vis, color = criterion.vis)) +
  geom_point(alpha = 0.25) +
  geom_smooth(method = "gam", formula = y ~ s(x, k = 4)) +
  facet_wrap(~bb.sp)

ggplot(beta.sp.wt2, aes(distance.surv, distance.vis, color = criterion.vis)) +
  geom_point(alpha = 0.25) +
  geom_smooth(method = "gam", formula = y ~ s(x, k = 4)) +
  facet_wrap(~bb.sp)

ggplot(beta.guild.bin2, aes(distance.surv, distance.vis, color = criterion.vis)) +
  geom_point(alpha = 0.25) +
  geom_smooth(method = "gam", formula = y ~ s(x, k = 4)) +
  facet_wrap(~pbl.fac)

ggplot(beta.guild.wt2, aes(distance.surv, distance.vis, color = criterion.vis)) +
  geom_point(alpha = 0.25) +
  geom_smooth(method = "gam", formula = y ~ s(x, k = 4)) +
  facet_wrap(~bb.guild)

```

### Parsed by period
```{r}
# Binary
## Taxonomic
### BB species
beta_x_period.tax.sp.bin <- beta_func(net, traits = bb_traits, sitedata = site_data, level = bb.sp, criterion = plant.sp, 
                                    binary = TRUE, period) %>%
  mutate(criterion = rep("tax", n()))

### Tongue length guild
beta_x_period.tax.guild.bin <- beta_func(net, traits = bb_traits, sitedata = site_data, level = pbl.fac, criterion = plant.sp, 
                                       binary = TRUE, period) %>%
  mutate(criterion = rep("tax", n()))

## Morphological
### BB species
beta_x_period.morph.sp.bin <- beta_func(net, traits = bb_traits, sitedata = site_data, level = bb.sp, criterion = k.type.s, 
                                    binary = TRUE, period) %>%
  mutate(criterion = rep("morph", n()))

### Tongue length guild
beta_x_period.morph.guild.bin <- beta_func(net, traits = bb_traits, sitedata = site_data, level = pbl.fac, criterion = k.type.s, 
                                       binary = TRUE, period) %>%
  mutate(criterion = rep("morph", n()))

### Collated
beta_x_period.sp.bin <- bind_rows(beta_x_period.tax.sp.bin, beta_x_period.morph.sp.bin)
beta_x_period.guild.bin <- bind_rows(beta_x_period.tax.guild.bin, beta_x_period.morph.guild.bin)

# Not binary
## Taxonomic
### BB species
beta_x_period.tax.sp.wt <- beta_func(net, traits = bb_traits, sitedata = site_data, level = bb.sp, criterion = plant.sp, 
                                    binary = FALSE, period) %>%
  mutate(criterion = rep("tax", n()))

### Tongue length guild
beta_x_period.tax.guild.wt <- beta_func(net, traits = bb_traits, sitedata = site_data, level = pbl.fac, criterion = plant.sp, 
                                       binary = FALSE, period) %>%
  mutate(criterion = rep("tax", n()))

## Morphological
### BB species
beta_x_period.morph.sp.wt <- beta_func(net, traits = bb_traits, sitedata = site_data, level = bb.sp, criterion = k.type.s, 
                                    binary = FALSE, period) %>%
  mutate(criterion = rep("morph", n()))

### Tongue length guild
beta_x_period.morph.guild.wt <- beta_func(net, traits = bb_traits, sitedata = site_data, level = pbl.fac, criterion = k.type.s, 
                                       binary = FALSE, period) %>%
  mutate(criterion = rep("morph", n()))

### Collated
beta_x_period.sp.wt <- bind_rows(beta_x_period.tax.sp.wt, beta_x_period.morph.sp.wt)
beta_x_period.guild.wt <- bind_rows(beta_x_period.tax.guild.wt, beta_x_period.morph.guild.wt)

# Pull out ADONIS results
## Binary
beta_x_period.tax.sp.adonis.bin <- beta_x_period.tax.sp.bin %>%
  select(bb.sp, period, adonis.r2, adonis.p) %>%
  distinct()

beta_x_period.morph.sp.adonis.bin <- beta_x_period.morph.sp.bin %>%
  select(bb.sp, period, adonis.r2, adonis.p) %>%
  distinct()

beta_x_period.tax.guild.adonis.bin <- beta_x_period.tax.guild.bin %>%
  select(pbl.fac, period, adonis.r2, adonis.p) %>%
  distinct()

beta_x_period.morph.guild.adonis.bin <- beta_x_period.morph.guild.bin %>%
  select(pbl.fac, period, adonis.r2, adonis.p) %>%
  distinct()

## Not binary
beta_x_period.tax.sp.adonis <- beta_x_period.tax.sp.wt %>%
  select(bb.sp, adonis.r2, adonis.p) %>%
  distinct()

beta_x_period.morph.sp.adonis <- beta_x_period.morph.sp.wt %>%
  select(bb.sp, adonis.r2, adonis.p) %>%
  distinct()

beta_x_period.tax.guild.adonis <- beta_x_period.tax.guild.wt %>%
  select(pbl.fac, adonis.r2, adonis.p) %>%
  distinct()

beta_x_period.morph.guild.adonis <- beta_x_period.morph.guild.wt %>%
  select(pbl.fac, adonis.r2, adonis.p) %>%
  distinct()

# Plot
## Binary
beta_x_period.sp.plot.bin <- ggplot(beta_x_period.sp.bin, aes(elev.diff, distance, color = criterion)) +
  geom_point(alpha = 0.1) +
  geom_smooth(method = "gam", formula = y ~ s(x, k = 4), se = FALSE) +
  facet_grid(bb.sp ~ period) + 
  ggtitle("Species")

beta_x_period.guild.plot.bin <- ggplot(beta_x_period.guild.bin, aes(elev.diff, distance, color = criterion)) +
  geom_point(alpha = 0.1) +
  geom_smooth(method = "gam", formula = y ~ s(x, k = 4), se = FALSE) +
  facet_grid(pbl.fac ~ period) + 
  ggtitle("Tongue guild")

(beta_x_period.sp.plot.bin | beta_x_period.guild.plot.bin) + plot_layout(guides = 'collect')

## Not binary
beta_x_period.sp.plot.wt <- ggplot(beta_x_period.sp.wt, aes(elev.diff, distance, color = criterion)) +
  geom_point(alpha = 0.1) +
  geom_smooth(method = "gam", formula = y ~ s(x, k = 4), se = FALSE) +
  facet_grid(bb.sp ~ period) + 
  ggtitle("Species")

beta_x_period.guild.plot.wt <- ggplot(beta_x_period.guild.wt, aes(elev.diff, distance, color = criterion)) +
  geom_point(alpha = 0.1) +
  geom_smooth(method = "gam", formula = y ~ s(x, k = 4), se = FALSE) +
  facet_grid(pbl.fac ~ period) + 
  ggtitle("Tongue guild")

(beta_x_period.sp.plot.wt | beta_x_period.guild.plot.wt) + plot_layout(guides = 'collect')

########## A different way to look at this
ggplot(beta_x_period.guild.wt, aes(pbl.fac, distance, color = criterion)) +
  geom_boxplot() +
  facet_wrap( ~ period)

ggplot(beta_x_period.sp.wt, aes(bb.sp, distance, color = criterion)) +
  geom_boxplot() +
  facet_wrap( ~ period) 
```

## 1.8 Does the influence of tongue length on niche partitioning vary with elevation or time?
```{r, echo = FALSE}
# By plant.sp
ggplot(niche_x_elevperiod.tax, aes(elev.class2, adonis.r2)) +
  geom_boxplot()

ggplot(niche_x_elevperiod.tax, aes(period, adonis.r2)) +
  geom_boxplot()

# By ktype
ggplot(niche_x_elevperiod.morph, aes(elev.class2, adonis.r2)) +
  geom_boxplot()

ggplot(niche_x_elevperiod.morph, aes(period, adonis.r2)) +
  geom_boxplot()
```

# 2. Testing the stress-gradient hypothesis and the Ranta and Vepsäläinen heterogeneity hypothesis

**Stress-gradient hypothesis:** Is there any evidence that niche partitioning amongst generalist core bumble bee species is governed by an inverse relationship between elevation and competition, broadly consistent with the stress-gradient hypothesis (Bertness and Callaway 1994). This mechanism has been inferred from trait patterns in the whole bee community of the Berchtesgadener Alps (Hoiss et al. 2012) and from interaction specialization patterns in the plant-pollinator community of Mt. Kilimanjaro (Classen et al. 2020). It predicts that niche overlap will increase with elevation, reflecting competitive pressures at low elevations and reduced competition but increased need for opportunism at high elevations.

**Ranta/Vepsäläinen heterogeneity hypothesis :** Ranta and Vepsäläinen (1981) proposed that bumble bees communities are governed by a spatiotemporally heterogeneous mosaic of competitive conditions, such that competitive inequalities between species are sufficiently variable to allow more species to coexist than predicted by trait-based competition models.

## 2.1: Testing the stress-gradient hypothesis
### 2.1.1 Face-value analysis

*Results:*

```{r message=FALSE, warning=FALSE, echo = FALSE}
# Pairwise (bb.sp:bb.sp) niche overlap by site, pooled across years

niche_overlap.tax <- niche_overlapper(net, criterion = plant.sp.abb)
niche_overlap.morph <- niche_overlapper(net, criterion = k.type.s)

# All spp pairs
sg1 <- ggplot(niche_overlap.tax, aes(elev.mean, niche.overlap)) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = "gam", formula = y ~ s(x, k = 10))

sg2 <- ggplot(niche_overlap.morph, aes(elev.mean, niche.overlap)) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = "gam", formula = y ~ s(x, k = 10))

(sg1 + ggtitle("Taxonomic") | sg2 + ggtitle("Morphological")) + plot_layout(guides = 'collect')

## Within LT spp
sg3 <- ggplot(filter(niche_overlap.tax, pbl.fac1 == "long" & pbl.fac2 == "long"), 
       aes(elev.mean, niche.overlap)) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = "gam", formula = y ~ s(x, k = 10))

sg4 <- ggplot(filter(niche_overlap.morph, pbl.fac1 == "long" & pbl.fac2 == "long"), 
       aes(elev.mean, niche.overlap)) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = "gam", formula = y ~ s(x, k = 10))

(sg3 + ggtitle("Taxonomic") | sg4 + ggtitle("Morphological")) + plot_layout(guides = 'collect')

## Within ST spp
sg5 <- ggplot(filter(niche_overlap.tax, pbl.fac1 == "short" & pbl.fac2 == "short"), 
       aes(elev.mean, niche.overlap)) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = "gam", formula = y ~ s(x, k = 10))

sg6 <- ggplot(filter(niche_overlap.morph, pbl.fac1 == "short" & pbl.fac2 == "short"), 
       aes(elev.mean, niche.overlap)) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = "gam", formula = y ~ s(x, k = 10))

(sg5 + ggtitle("Taxonomic") | sg6 + ggtitle("Morphological")) + plot_layout(guides = 'collect')

## Between ST and LT
sg7 <- ggplot(filter(niche_overlap.tax, pbl.fac.diff == "different"), 
       aes(elev.mean, niche.overlap)) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = "gam", formula = y ~ s(x, k = 10))

sg8 <- ggplot(filter(niche_overlap.morph, pbl.fac.diff == "different"), 
       aes(elev.mean, niche.overlap)) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = "gam", formula = y ~ s(x, k = 8))

(sg7 + ggtitle("Taxonomic") | sg8 + ggtitle("Morphological")) + plot_layout(guides = 'collect')

# GAMs
glob_GAM.tax <- gam(niche.overlap ~ 
                      s(year, bs  = "re") + # random intercept effect for year
                      s(elev.mean, k = 8, m = 2) + # global smooth ~ elevation
                      s(elev.mean, sp.pair, bs = "fs", k = 8, m = 1), # group-level smooth ~ elevation*sp.pair
                    method = "REML",
                    family = "gaussian",
                    data = niche_overlap.tax) %>% getViz()

check.gamViz(glob_GAM.tax)
summary(glob_GAM.tax)
print(plot(glob_GAM.tax, allTerms = TRUE), pages = 1)

glob_GAM.morph <- gam(niche.overlap ~ 
                      s(year, bs  = "re") + # random intercept effect for year
                      s(elev.mean, k = 8, m = 2) + # global smooth ~ elevation
                      s(elev.mean, sp.pair, bs = "fs", k = 8, m = 1), # group-level smooth ~ elevation*sp.pair
                    method = "REML",
                    family = "gaussian",
                    data = niche_overlap.morph) %>% getViz()

check.gamViz(glob_GAM.morph)
summary(glob_GAM.morph)
print(plot(glob_GAM.morph, allTerms = TRUE), pages = 1)

# Within LT class
LT_GAM.tax <- gam(niche.overlap ~ 
                      s(year, bs  = "re") + # random intercept effect for year
                      s(elev.mean, k = 8, m = 2) + # global smooth ~ elevation
                      s(elev.mean, sp.pair, bs = "fs", k = 8, m = 1), # group-level smooth ~ elevation*sp.pair
                  method = "REML",
                  family = "gaussian",
                  data = filter(niche_overlap.tax, pbl.fac1 == "long" & pbl.fac2 == "long")) %>% getViz()

check.gamViz(LT_GAM.tax)
summary(LT_GAM.tax)
print(plot(LT_GAM.tax, allTerms = TRUE), pages = 1)

LT_GAM.morph <- gam(niche.overlap ~ 
                      s(year, bs  = "re") + # random intercept effect for year
                      s(elev.mean, k = 8, m = 2) + # global smooth ~ elevation
                      s(elev.mean, sp.pair, bs = "fs", k = 8, m = 1), # group-level smooth ~ elevation*sp.pair
                  method = "REML",
                  family = "gaussian",
                  data = filter(niche_overlap.morph, pbl.fac1 == "long" & pbl.fac2 == "long")) %>% getViz()

check.gamViz(LT_GAM.morph)
summary(LT_GAM.morph)
print(plot(LT_GAM.morph, allTerms = TRUE), pages = 1)

# Within ST class
ST_GAM.tax <- gam(niche.overlap ~ 
                      s(year, bs  = "re") + # random intercept effect for year
                      s(elev.mean, k = 8, m = 2) + # global smooth ~ elevation
                      s(elev.mean, sp.pair, bs = "fs", k = 8, m = 1), # group-level smooth ~ elevation*sp.pair
                  method = "REML",
                  family = "gaussian",
                  data = filter(niche_overlap.tax, pbl.fac1 == "short" & pbl.fac2 == "short")) %>% getViz()

check.gamViz(ST_GAM.tax)
summary(ST_GAM.tax)
print(plot(ST_GAM.tax, allTerms = TRUE), pages = 1)

ST_GAM.morph <- gam(niche.overlap ~ 
                      s(year, bs  = "re") + # random intercept effect for year
                      s(elev.mean, k = 8, m = 2) + # global smooth ~ elevation
                      s(elev.mean, sp.pair, bs = "fs", k = 8, m = 1), # group-level smooth ~ elevation*sp.pair
                  method = "REML",
                  family = "gaussian",
                  data = filter(niche_overlap.morph, pbl.fac1 == "short" & pbl.fac2 == "short")) %>% getViz()

check.gamViz(ST_GAM.morph)
summary(ST_GAM.morph)
print(plot(ST_GAM.morph, allTerms = TRUE), pages = 1)

# Between ST and LT
STxLT_GAM.tax <- gam(niche.overlap ~ 
                      s(year, bs  = "re") + # random intercept effect for year
                      s(elev.mean, k = 8, m = 2) + # global smooth ~ elevation
                      s(elev.mean, sp.pair, bs = "fs", k = 8, m = 1), # group-level smooth ~ elevation*sp.pair
                  method = "REML",
                  family = "gaussian",
                  data = filter(niche_overlap.tax, pbl.fac.diff == "different")) %>% getViz()

check.gamViz(STxLT_GAM.tax)
summary(STxLT_GAM.tax)
print(plot(STxLT_GAM.tax, allTerms = TRUE), pages = 1)

STxLT_GAM.morph <- gam(niche.overlap ~ 
                      s(year, bs  = "re") + # random intercept effect for year
                      s(elev.mean, k = 8, m = 2) + # global smooth ~ elevation
                      s(elev.mean, sp.pair, bs = "fs", k = 8, m = 1), # group-level smooth ~ elevation*sp.pair
                  method = "REML",
                  family = "gaussian",
                  data = filter(niche_overlap.morph, pbl.fac.diff == "different")) %>% getViz()

check.gamViz(STxLT_GAM.morph)
summary(STxLT_GAM.morph)
print(plot(STxLT_GAM.morph, allTerms = TRUE), pages = 1)
```

### 2.1.2 Null-model residual analysis

Arguably the most rigorous way to tease out the effect of elevation while controlling for community composition (abundance and richness patterns) is to compare the observed niche overlap at each site to that predicted by a random null model, then see how the residuals vary with elevation.

I'll be the first to admit that I'm new to null models, so I want to be extremely clear about the logic of how I'm setting this up. Step-by-step:  
1. Generate community matrices by site and pollen/nectar
2. Permute each matrix *n* times by randomly redistributing the visits of each bumble bee species while holding the total visits for each bumble bee species constant. This is, therefore, a random selection model constrained to the set of flora visited for a given site-pollen/nectar matrix. Importantly, I am opting not to hold the floral marginal totals constant, i.e. the total number of visits that each plant receives will vary with each permutation. My gut feeling is that this is the right place to start, but this is a decision that requires scrutiny.
3. Calculate niche overlap for each permuted matrix and store as tidy arrays; this mirrors the workflow for calculating empirical niche overlap, the only difference being that I had to deal with some tricky nested lists; I'm pretty sure I did this correctly.
4. Take the mean niche overlap for each permuted community.
5. Join null model mean niche overlap table to empirical niche overlap table, joining by site and pollen/nectar. Thus, for each site-pollen/nectar combo in the empirical niche overlap table, we have *n* iterations of random niche overlap.
6. Calculate the difference between empirical niche overlap and null niche overlap for each iteration; we'll call this "residual niche overlap".
7. Calculate mean residual niche overlap for each site-pollen/nectar combo. This can then be regressed against elevation to see how the deviation from null expectations varies along the elevation gradient, with the interpretation that this represents the "pure" influence of elevation on niche overlap whle controlling for community differences. We could also use the raw differences without collapsing them to a mean. But that's not what I did this first time through.

There are surely a lot of other ways a null model could be constructed and analyzed, but this seems to me a good place to start.

*Results*

```{r message=FALSE, warning=TRUE, echo = FALSE}
# Holding BB marginal totals constant
random_niche.tax <- nullify1(net, 50, plant.sp)
random_niche.morph <- nullify1(net, 50, k.type.s)

# Holding both BB and plant marginal totals constant
# random_niche.tax <- nullify2(net, 50, plant.sp)
# random_niche.morph <- nullify2(net, 50, k.type.s)

## Join null model to empirical niche overlap
# Taxonomic niche overlap
comparison.tax <- niche_overlap.tax %>%
  left_join(random_niche.tax, by = c("site", "sp.pair")) %>%
  mutate(niche.overlap.resid = niche.overlap - niche.overlap.sim) %>%
  group_by(site, elev.mean, sp.pair, pbl.fac1, pbl.fac2, pbl.fac.diff) %>%
  summarize(niche.overlap.resid = mean(niche.overlap.resid))

# Morphological niche overlap
comparison.morph <- niche_overlap.morph %>%
  left_join(random_niche.morph, by = c("site", "sp.pair")) %>%
  mutate(niche.overlap.resid = niche.overlap - niche.overlap.sim) %>%
  group_by(site, elev.mean, sp.pair, pbl.fac1, pbl.fac2, pbl.fac.diff) %>%
  summarize(niche.overlap.resid = mean(niche.overlap.resid))

# Plot
# All spp pairs
sgn1 <- ggplot(comparison.tax, aes(elev.mean, niche.overlap.resid)) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = "gam", formula = y ~ s(x, k = 10))

sgn2 <- ggplot(comparison.morph, aes(elev.mean, niche.overlap.resid)) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = "gam", formula = y ~ s(x, k = 10))

(sgn1 + ggtitle("Taxonomic") | sgn2 + ggtitle("Morphological")) + plot_layout(guides = 'collect')

## Within LT spp
sgn3 <- ggplot(filter(comparison.tax, pbl.fac1 == "long" & pbl.fac2 == "long"), 
       aes(elev.mean, niche.overlap.resid)) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = "gam", formula = y ~ s(x, k = 10))

sgn4 <- ggplot(filter(comparison.morph, pbl.fac1 == "long" & pbl.fac2 == "long"), 
       aes(elev.mean, niche.overlap.resid)) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = "gam", formula = y ~ s(x, k = 10))

(sgn3 + ggtitle("Taxonomic") | sgn4 + ggtitle("Morphological")) + plot_layout(guides = 'collect')

## Within ST spp
sgn5 <- ggplot(filter(comparison.tax, pbl.fac1 == "short" & pbl.fac2 == "short"), 
       aes(elev.mean, niche.overlap.resid)) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = "gam", formula = y ~ s(x, k = 10))

sgn6 <- ggplot(filter(comparison.morph, pbl.fac1 == "short" & pbl.fac2 == "short"), 
       aes(elev.mean, niche.overlap.resid)) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = "gam", formula = y ~ s(x, k = 10))

(sgn5 + ggtitle("Taxonomic") | sgn6 + ggtitle("Morphological")) + plot_layout(guides = 'collect')

## Between ST and LT
sgn7 <- ggplot(filter(comparison.tax, pbl.fac.diff == "different"), 
       aes(elev.mean, niche.overlap.resid)) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = "gam", formula = y ~ s(x, k = 10))

sgn8 <- ggplot(filter(comparison.morph, pbl.fac.diff == "different"), 
       aes(elev.mean, niche.overlap.resid)) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = "gam", formula = y ~ s(x, k = 8))

(sgn7 + ggtitle("Taxonomic") | sgn8 + ggtitle("Morphological")) + plot_layout(guides = 'collect')

# GAMs
glob_nullmodel_GAM.tax <- gam(niche.overlap.resid ~ 
                      s(elev.mean, k = 10, m = 2),
                    method = "REML",
                    family = "gaussian",
                    data = comparison.tax) %>% getViz()

check.gamViz(glob_nullmodel_GAM.tax)
summary(glob_nullmodel_GAM.tax)
print(plot(glob_nullmodel_GAM.tax, allTerms = TRUE), pages = 1)

glob_nullmodel_GAM.morph <- gam(niche.overlap.resid ~ 
                      s(elev.mean, k = 10, m = 2),
                    method = "REML",
                    family = "gaussian",
                    data = comparison.morph) %>% getViz()

check.gamViz(glob_nullmodel_GAM.morph)
summary(glob_nullmodel_GAM.morph)
print(plot(glob_nullmodel_GAM.morph, allTerms = TRUE), pages = 1)

# Within LT class
LT_nullmodel_GAM.tax <- gam(niche.overlap.resid ~ 
                    s(elev.mean, k = 10, m = 2),
                  method = "REML",
                  family = "gaussian",
                  data = filter(comparison.tax, pbl.fac1 == "long" & pbl.fac2 == "long")) %>% getViz()

check.gamViz(LT_nullmodel_GAM.tax)
summary(LT_nullmodel_GAM.tax)
print(plot(LT_nullmodel_GAM.tax, allTerms = TRUE), pages = 1)

LT_nullmodel_GAM.morph <- gam(niche.overlap.resid ~ 
                    s(elev.mean, k = 10, m = 2),
                  method = "REML",
                  family = "gaussian",
                  data = filter(comparison.morph, pbl.fac1 == "long" & pbl.fac2 == "long")) %>% getViz()

check.gamViz(LT_nullmodel_GAM.morph)
summary(LT_nullmodel_GAM.morph)
print(plot(LT_nullmodel_GAM.morph, allTerms = TRUE), pages = 1)

# Within ST class
ST_nullmodel_GAM.tax <- gam(niche.overlap.resid ~ 
                    s(elev.mean, k = 10, m = 2),
                  method = "REML",
                  family = "gaussian",
                  data = filter(comparison.tax, pbl.fac1 == "short" & pbl.fac2 == "short")) %>% getViz()

check.gamViz(ST_nullmodel_GAM.tax)
summary(ST_nullmodel_GAM.tax)
print(plot(ST_nullmodel_GAM.tax, allTerms = TRUE), pages = 1)

ST_nullmodel_GAM.morph <- gam(niche.overlap.resid ~ 
                    s(elev.mean, k = 10, m = 2),
                  method = "REML",
                  family = "gaussian",
                  data = filter(comparison.morph, pbl.fac1 == "short" & pbl.fac2 == "short")) %>% getViz()

check.gamViz(ST_nullmodel_GAM.morph)
summary(ST_nullmodel_GAM.morph)
print(plot(ST_nullmodel_GAM.morph, allTerms = TRUE), pages = 1)

# Between ST and LT
STxLT_nullmodel_GAM.tax <- gam(niche.overlap.resid ~ 
                    s(elev.mean, k = 10, m = 2),
                  method = "REML",
                  family = "gaussian",
                  data = filter(comparison.tax, pbl.fac.diff == "different")) %>% getViz()

check.gamViz(STxLT_nullmodel_GAM.tax)
summary(STxLT_nullmodel_GAM.tax)
print(plot(STxLT_nullmodel_GAM.tax, allTerms = TRUE), pages = 1)

STxLT_nullmodel_GAM.morph <- gam(niche.overlap.resid ~ 
                    s(elev.mean, k = 10, m = 2),
                  method = "REML",
                  family = "gaussian",
                  data = filter(comparison.morph, pbl.fac.diff == "different")) %>% getViz()

check.gamViz(STxLT_nullmodel_GAM.morph)
summary(STxLT_nullmodel_GAM.morph)
print(plot(STxLT_nullmodel_GAM.morph, allTerms = TRUE), pages = 1)
```

### 2.1.3. Parsing niche contraction and niche separation

Contrary to the predictions of H1, niche overlap decreases sharply at high elevation. This is true for both taxonomic and morphological niche overlap, and it holds in both the face-value and null-residual approaches. This decrease in niche overlap might be explained by two modest, proximate hypotheses:

1.  **H1:** The decrease of niche overlap at high elevations is due to *niche contraction*, i.e. BB spp. visiting a smaller proportion of available (and "eligible", i.e. visited at least once by a bumble bee) flora. 
  *Prediction:* "resource range" (Poisot et al. 2012) will decrease with elevation (as Dormann notes, resource range is a "backward" metric insofar as a large resource range means a small proportional niche breadth; I will calculate the inverse of resource range so that it makes more sense)
2.  **H2:** The decrease of niche overlap at high elevations is due to *niche separation*, i.e. BB spp. avoiding one another's niches but not contracting their own.
  *Predictions:* C score will increase with elevation

*Results:* I haven't run formal stats on this analysis yet, but from the visualization, it looks like the overall story is that niche overlap is driven mainly by niche separation, not so much but niche contraction. My first impression is that this suggests a behavior response to competition.

```{r, echo = FALSE}
resource_range.sp <- net %>%
  filter(bb.sp %in% c("bss", "pasc", "prat", "hort", "soro", "wurf")) %>% # focal species
  filter(!is.na(pollen)) %>%
  group_by(site, year, pollen, bb.sp, plant.sp) %>% # group by site, date, bee species, plant species
  summarize(freq = n()) %>% # calculate interaction frequency per species pair within each site
  unite(webID, c(site, year, pollen), sep = "_") %>%
  ungroup() %>%
  dplyr::select(higher = bb.sp, lower = plant.sp, webID, freq) %>% # rename columns to match bipartite's expectations
  data.frame() %>% # convert to data frame
  frame2webs() %>% # convert to to bipartite web
  map(specieslevel, level = "higher", index = "resource range") %>%
  map(rownames_to_column) %>%
  map(as_tibble) %>%
  bind_rows(.id = "webID") %>% # Collapse list into big data frame
  rename(bb.sp = rowname) %>%
  separate(webID, c("site", "year", "pollen"), sep =  "_") %>%
  left_join(site_data) %>%
  mutate(site = factor(site),
         year = factor(year),
         pollen = factor(pollen)) %>%
  filter(bb.sp %in% c("bss", "pasc", "prat", "hort", "soro", "wurf")) %>% # for some reason, one row has a "1" for bb.sp, and I don't know where this comes from; this gets rid of it, but I would still like to know why
  mutate(resource.range = 1-resource.range) %>%
  left_join(BB_covars.splevel) %>%
  filter(bb.sp.abund > 1)

resource_range.sp.mean <- resource_range.sp %>%
  group_by(site, year, elev.mean, pollen) %>%
  summarize(mean.resource.range = mean(resource.range))

ggplot(resource_range.sp, aes(elev.mean, resource.range, color = pollen)) +
  geom_point() +
  geom_smooth() +
  facet_grid(bb.sp ~ year) +
  ylim(c(0,1))

ggplot(resource_range.sp.mean, aes(elev.mean, mean.resource.range, color = pollen)) +
  geom_point() +
  geom_smooth() +
  facet_wrap(~year) +
  ylim(c(0,1))

resource_range.ktype <- net %>%
  filter(bb.sp %in% c("bss", "pasc", "prat", "hort", "soro", "wurf")) %>% # focal species
  filter(!is.na(pollen)) %>%
  group_by(site, year, pollen, bb.sp, k.type.s) %>% # group by site, date, bee species, plant species
  summarize(freq = n()) %>% # calculate interaction frequency per species pair within each site
  unite(webID, c(site, year, pollen), sep = "_") %>%
  ungroup() %>%
  dplyr::select(higher = bb.sp, lower = k.type.s, webID, freq) %>% # rename columns to match bipartite's expectations
  data.frame() %>% # convert to data frame
  frame2webs() %>% # convert to to bipartite web
  map(specieslevel, level = "higher", index = "resource range") %>%
  map(rownames_to_column) %>%
  map(as_tibble) %>%
  bind_rows(.id = "webID") %>% # Collapse list into big data frame
  rename(bb.sp = rowname) %>%
  separate(webID, c("site", "year", "pollen"), sep =  "_") %>%
  left_join(site_data) %>%
  mutate(site = factor(site),
         year = factor(year),
         pollen = factor(pollen)) %>%
  filter(bb.sp %in% c("bss", "pasc", "prat", "hort", "soro", "wurf")) %>% # for some reason, one row has a "1" for bb.sp, and I don't know where this comes from; this gets rid of it, but I would still like to know why
  mutate(resource.range = 1-resource.range) %>%
  left_join(BB_covars.splevel) %>%
  filter(bb.sp.abund > 1)

resource_range.ktype.mean <- resource_range.ktype %>%
  group_by(site, year, elev.mean, pollen) %>%
  summarize(mean.resource.range = mean(resource.range))

ggplot(resource_range.ktype, aes(elev.mean, resource.range, color = pollen)) +
  geom_point() +
  geom_smooth() +
  facet_grid(bb.sp ~ year) +
  ylim(c(0,1))

ggplot(resource_range.ktype.mean, aes(elev.mean, mean.resource.range, color = pollen)) +
  geom_point() +
  geom_smooth() +
  facet_wrap(~year) +
  ylim(c(0,1))

cscore.sp <- net %>%
    filter(bb.sp %in% c("bss", "pasc", "prat", "hort", "soro", "wurf")) %>% # focal species
    filter(!is.na(pollen)) %>%
    group_by(site, year, pollen, bb.sp, plant.sp) %>% # group by site, date, bee species, plant species
    summarize(freq = n()) %>% # calculate interaction frequency per species pair within each site
    group_by(site, year, pollen) %>%
    unite(webID, c(site, year, pollen), sep = "_") %>%
    ungroup() %>%
    dplyr::select(higher = bb.sp, lower = plant.sp, webID, freq) %>% # rename columns to match bipartite's expectations
    data.frame() %>% # convert to data frame
    frame2webs() %>% # convert to to bipartite web
    map(grouplevel, level = "higher", weighted = TRUE, index = c("C score")) %>% 
    data.frame() %>%
    t() %>%
    data.frame() %>%
    rownames_to_column() %>%
    dplyr::select(webID = rowname, everything()) %>%
    as_tibble() %>%
    separate(webID, c("site", "year", "pollen"), sep =  "_") %>%
    left_join(site_data) %>%
    mutate(site = factor(site),
           year = factor(year),
           pollen = factor(pollen))

ggplot(cscore.sp, aes(elev.mean, C.score.HL, color = pollen)) +
  geom_point() +
  geom_smooth() +
  facet_wrap(~year)

cscore.ktype <- net %>%
    filter(bb.sp %in% c("bss", "pasc", "prat", "hort", "soro", "wurf")) %>% # focal species
    filter(!is.na(pollen)) %>%
    group_by(site, year, pollen, bb.sp, k.type.s) %>% # group by site, date, bee species, plant species
    summarize(freq = n()) %>% # calculate interaction frequency per species pair within each site
    group_by(site, year, pollen) %>%
    unite(webID, c(site, year, pollen), sep = "_") %>%
    ungroup() %>%
    dplyr::select(higher = bb.sp, lower = k.type.s, webID, freq) %>% # rename columns to match bipartite's expectations
    data.frame() %>% # convert to data frame
    frame2webs() %>% # convert to to bipartite web
    map(grouplevel, level = "higher", weighted = TRUE, index = c("C score")) %>% 
    data.frame() %>%
    t() %>%
    data.frame() %>%
    rownames_to_column() %>%
    dplyr::select(webID = rowname, everything()) %>%
    as_tibble() %>%
    separate(webID, c("site", "year", "pollen"), sep =  "_") %>%
    left_join(site_data) %>%
    mutate(site = factor(site),
           year = factor(year),
           pollen = factor(pollen))

ggplot(cscore.ktype, aes(elev.mean, C.score.HL, color = pollen)) +
  geom_point() +
  geom_smooth() +
  facet_wrap(~year)
```