### 2020-12-18 analyses/visitation_bias.Rmd

---
title: "visitation_bias"
author: "Doug Sponsler"
date: "12/14/2020"
output: html_document
---
# Preparations
## Packages 
```{r, warning=FALSE, include=FALSE, echo = FALSE}
library(bipartite)
library(mgcv)
library(mgcViz)
library(brms)
library(bayesplot)
#library(tidybayes)
library(tidyverse)
library(lubridate)
```

## Load data
```{r message=FALSE, warning=FALSE, include=FALSE, echo = FALSE}
# Site data
site_data <- read_csv("../processed_data/site_data.csv") %>%
  dplyr::select(-c(slope.calc, slope.est, elev.min, elev.max, temp.mean)) %>% # drop these variables
  mutate(elev.class2 = factor(elev.class2, levels = c("low", "mid", "high")),
         site = factor(site)) # turn elav.class into an ordered factor and site into a regular factor

# BB traits
bb_traits <- read_csv("../processed_data/bb_traits.csv")

# FL traits
## family-level taxonomy
fl_tax <- read_csv("../processed_data/floral_tax.csv")

## Kugler morphotypes (accessed through Bioflor)
fl_ktype <- read_csv("../processed_data/floral_k_type.csv")

## Flower colors (accessed through BioFlor)
fl_color <- read_csv("../processed_data/floral_color.csv")

## diameter and corolla depth data from Gita Benadi
fl_morph <- read_csv("../processed_data/flower_morphology_GitaBenadi.csv") %>% 
  dplyr::select(SampleID, plant.family = Family, plant.sp = Species, 
                Unit, Flowers_per_unit, diam = Diameter, LN) %>%
  group_by(plant.sp) %>%
  summarize(diam = mean(diam), LN = mean(LN))

## verbal description of Kugler morphotypes
fl_ktype_key <- read_delim("../processed_data/kugler_key.tsv", delim = "\t", col_names = FALSE) %>%
  dplyr::select(k.type = 1, description = 2)

## join all floral trait tables
fl_traits <- fl_tax %>%
  full_join(fl_ktype) %>%
  full_join(fl_color) %>%
  full_join(fl_morph) %>%
  dplyr::select(plant.sp, plant.genus, plant.family, k.type, k.type.s, k.type.ss, color, diam, LN)
```

## Align survey and visitation data with pseudodates
```{r}
# Site-dates where BB visitation was recorded (excluding dates where only males were recorded)
visitation_samples <- read_csv("../processed_data/network.csv") %>%
  filter(caste != "male") %>%
  select(site, date) %>%
  distinct()

# Site dates where floral surveying was conducted
survey_samples <- read_csv("../processed_data/floral_survey.csv") %>%
  select(site, date) %>%
  distinct()

# Site-dates where visitation sampling and floral sampling coincided
shared_dates <- inner_join(visitation_samples, survey_samples)

# Site-dates where visitation was recorded WITHOUT corresponding survey data
visitation_only <- anti_join(visitation_samples, survey_samples) %>%
  mutate(date.v = date) %>%
  select(-date)

# This creates a pseudo date column for unmatched visitation site-dates that will match each visitation date
# to its nearest survey data. For site t4, there is a 2-year mismatch; we will drop these site-dates later.
pseudodates <- full_join(visitation_only, survey_samples) %>%
  mutate(mismatch = abs(date.v - date)) %>%
  group_by(site, date.v) %>%
  filter(mismatch == min(mismatch)) %>%
  rename(date = date.v, pseudo.date = date)
```

## Generate a data frame that represents visitation and survey abundance
```{r}
# BB visitation data
visitation <- read_csv("../processed_data/network.csv") %>%
  filter(caste != "male") %>% # we will ignore males 
  group_by(site, date, bb.sp, plant.sp) %>%
  summarize(visits = n()) %>% # visits per BB-FL pair
  ungroup() %>%
  left_join(pseudodates) %>% # Now we will replace unmatched dates with their pseudo dates
  replace_na(replace = list(mismatch = 0)) %>% # set mismatch to zero for perfectly aligned dates
  filter(mismatch <= 7) %>% # this gets rid of irreconcilable dates (like the t4 dates mentioned above)
  mutate(date2 = if_else(is.na(pseudo.date), date, pseudo.date)) %>% # replace date with pseudodate for all site-dates lacking perfect alignments
  select(-c(date, pseudo.date, mismatch)) %>%
  rename(date = date2) %>%
  group_by(site, date) %>%
  complete(bb.sp, plant.sp, fill = list(visits = 0)) # generate all possible BB-FL combos for each site date and fill in zeros for visits

# Floral cover data
flower_cover <- read_csv("../processed_data/floral_survey.csv") %>%
  filter(flower.cover > 0) %>% # ignore plants that were not blooming
  group_by(site, date, plant.sp) %>%
  summarize(flower.cover = sum(flower.cover))

# Get shared dates after adding pseudodates. This will let us filter out survey dates that don't
# have corresponding visitation sampling dates
shared_dates_with_pseudo <- inner_join(visitation, flower_cover, by = c("site", "date")) %>%
  select(site, date) %>%
  distinct()

# Join visitation and flower cover data

## First we need a list of all BB spp. by site and date
sp_by_site <- visitation %>% 
  ungroup() %>%
  select(site, date, bb.sp) %>%
  distinct()

## Then we need a list of unvisited plant.sp by site and date
unvisited <- flower_cover %>% 
  semi_join(shared_dates_with_pseudo) %>% # filter flower cover data to include only site-dates with visitation complement (real or pseudodates)
  full_join(visitation) %>% # join to visitation data
  filter(is.na(bb.sp)) %>% # select only the site-date-plants that were totally unvisited (and therefore have no BB spp to join to)
  select(-bb.sp) %>% # drop bb.sp variable because I just want the plants
  left_join(sp_by_site) %>% # join in the species list by site and date, thus populating the unvisited plants with all the BB spp recorded for each site-date
  replace_na(replace = list(visits = 0)) # replace NAs (created in the line above) with zero

## Now we can combine everything into a single data frame
visit_cover <- flower_cover %>% 
  semi_join(shared_dates_with_pseudo) %>% # filter flower cover data to include only site-dates with visitation complement (real or pseudodates)
  full_join(visitation) %>% # join to visitation data
  replace_na(replace = list(flower.cover = 0.01)) %>% # If a plant was visited but not surveyed, we will assign it the minimum non-zero flower cover value of 0.01.
  filter(!is.na(bb.sp)) %>% # drop the univisited plants
  bind_rows(unvisited) %>% # add them back, with BB spp and zeros
  select(site, date, bb.sp, plant.sp, visits, flower.cover) %>%
  group_by(site, date, bb.sp, plant.sp, flower.cover) %>% # In a few cases, the use of pseudodates causes two sampling period to converge on the same date, which creates apparent duplicate visitation data. We will handle this by averaging the visits.
  summarize(visits = mean(visits)) %>%
  left_join(fl_traits) %>% # add floral traits
  ungroup()

## Get total visits by BB sp and site-date; this will be a measure of BB abundance that we can use after we've calculated d'. I could, of course, work this into the biaseval function, but I'll put that off for now.
bb_abundance <- visit_cover %>%
  group_by(site, date, bb.sp) %>%
  summarize(bb.sp.abund = sum(visits)) %>%
  group_by(site, date) %>%
  mutate(bb.sp.prop.abund = bb.sp.abund/sum(bb.sp.abund))
```

## A handy function for calculating Blüthgen's d'
```{r}
# A function for calculating Blüthgen's d', parsed in various ways
biaseval <- function(x, criterion, ...) { # criterion = plant grouping variable, e.g. plant.sp, k.type.ss, color

  criterion <- enquo(criterion) # tidy eval

  x %>%

# Select variables of interest; the ... represents the variables we will use to parse our analysis
    select(..., bb.sp, !!criterion, flower.cover, visits) %>% 
    group_by(..., bb.sp, !!criterion) %>%

# If criterion = plant.sp, this next step does nothing. But if criterion is some other grouping variable (e.g. k.type.ss), this sums flower cover and visits over that grouping variable.
    summarize(flower.cover = sum(flower.cover), 
              visits = sum(visits)) %>%

# Nest into list column
    group_by(...) %>%
    nest() %>%

# Pivot each nested data frame into the kind of web that bipartite likes
    mutate(webs = map(data, function(x) select(x, bb.sp, !!criterion, visits) %>%
                        pivot_wider(names_from = !!criterion, values_from = visits, values_fill = 0) %>%
                        column_to_rownames("bb.sp"))) %>%

# Pull out the flower.cover variable to use in the "abuns" argument of dfun below
    mutate(low.abund = map(data, function(x) select(x, !!criterion, flower.cover) %>%
                            unique() %>%
                            column_to_rownames(paste(criterion)[2]))) %>% # Tell me why I need to do this, Hadley.

# Call the dfun command from bipartite to calculate Blüthgen's d'
    mutate(d = map2(webs, low.abund, function(x, y) bipartite::dfun(x, abuns = y$flower.cover) %>%
                      data.frame() %>%
                      rownames_to_column(var = "bb.sp"))) %>%

    unnest(cols = d) %>%
    na.omit() %>% # in a few cases where the sampling is very sparse, d' cannot be calculated; these are safe to drop
    ungroup() %>%
    mutate(site = factor(site),
         bb.sp = factor(bb.sp),
         year = factor(lubridate::year(date)))
}
```

# 1. Is bumble bee foraging biased? How do taxonomic, phylogenetic, and morphological bias compare? Are some BB species more biased than others?
## 1.1 Calculate Blüthgen's d' (i.e. visitation bias)
```{r}
# No grouping -- just plant.sp
bias_plantsp.sitedate <- biaseval(visit_cover, plant.sp, site, date) %>%
  left_join(site_data) %>% # add site data
  left_join(bb_abundance) %>% # add BB abundance data
  mutate(criterion = factor(rep("plant.sp", n())))

# Grouped by k.type.ss
bias_ktype.sitedate <- biaseval(visit_cover, k.type.ss, site, date) %>%
  left_join(site_data) %>%
  left_join(bb_abundance) %>% # add BB abundance data
  mutate(criterion = factor(rep("ktype", n())))

# Grouped by plant.family
bias_family.sitedate <- biaseval(visit_cover, plant.family, site, date) %>%
  left_join(site_data) %>%
  left_join(bb_abundance) %>% # add BB abundance data
  mutate(criterion = factor(rep("plant.family", n())))

# Collated
bias_sitedate <- bind_rows(bias_plantsp.sitedate, 
                           bias_ktype.sitedate, 
                           bias_family.sitedate)
```

## 1.2 Visualize
```{r}
# Compare mean bias by criterion
ggplot(bias_sitedate, aes(criterion, dprime)) +
  geom_boxplot()

# Bias ~ elevation * criterion
ggplot(bias_sitedate, aes(elev.mean, dprime, color = criterion)) +
  geom_point(alpha = 0.25) +
  geom_smooth(se = TRUE)

ggplot(bias_sitedate, aes(elev.mean, dprime, color = criterion)) +
  geom_point(alpha = 0.25) +
  geom_smooth(se = FALSE, method = "gam", formula = y ~ s(x, k = 5)) +
  facet_wrap(~bb.sp)

# Bias ~ year; Blüthgen's d' seems very robust to sampling effects, since it does not vary across years
ggplot(bias_sitedate, aes(year, dprime, color = criterion)) +
  geom_boxplot()

ggplot(bias_sitedate, aes(year, dprime, color = criterion)) +
  geom_boxplot() +
  facet_wrap(~bb.sp)

# Does bias vary with abundance?
ggplot(bias_sitedate, aes(log(bb.sp.abund), dprime, color = criterion)) +
  #geom_point() +
  geom_smooth() +
  facet_wrap(~bb.sp)

ggplot(bias_plantsp.sitedate, aes(log(bb.sp.prop.abund), dprime)) +
  geom_point() +
  geom_smooth() +
  facet_wrap(~bb.sp)
```

## 1.3 Are some BB species more biased than others? (brms models)
```{r}
# Do species vary in d' (simple intercept model)
brm.plantsp <- brm(dprime ~ (1 | bb.sp), data = filter(bias.sitedate, criterion ==  "plant.sp"))
summary(brm.plantsp)
mcmc_plot(brm.plantsp, pars = c("^r_"))
mcmc_plot(brm.plantsp, pars = c("^b_"))
parnames(brm.plantsp)

brm.ktype <- brm(dprime ~ (1 | bb.sp), data = filter(bias.sitedate, criterion ==  "ktype"))
summary(brm.ktype)
mcmc_plot(brm.ktype, pars = c("^r_"))
mcmc_plot(brm.ktype, pars = c("^b_"))
parnames(brm.ktype)

brm.family <- brm(dprime ~ (1 | bb.sp), data = filter(bias.sitedate, criterion ==  "plant.family"))
summary(brm.family)
mcmc_plot(brm.family, pars = c("^r_"))
mcmc_plot(brm.family, pars = c("^b_"))
parnames(brm.family)
```

## 1.4 Does bias vary with criterion? (brms models)
```{r}
brm.criterion <- brm(dprime ~ (1 | criterion), data = bias.sitedate)
summary(brm.criterion)
mcmc_plot(brm.criterion, pars = c("^r_"))
mcmc_plot(brm.criterion, pars = c("^b_"))
parnames(brm.criterion)
```

## 1.5 Does bias vary with elevation? (brms models)
```{r}
# Does d' vary with elevation
brm2.plantsp <- brm(dprime ~ elev.mean + (1 | bb.sp), 
                    control = list(max_treedepth = 12),
                    data = filter(bias.sitedate, criterion ==  "plant.sp"))
summary(brm2.plantsp)
mcmc_plot(brm2.plantsp, pars = c("^r_"))
mcmc_plot(brm2.plantsp, pars = c("^b_"))
conditional_effects(brm2.plantsp)
parnames(brm2.plantsp)

brm2.ktype <- brm(dprime ~ elev.mean + (1 | bb.sp), 
                  control = list(max_treedepth = 12),
                  data = filter(bias.sitedate, criterion ==  "ktype"))
summary(brm2.ktype)
mcmc_plot(brm2.ktype, pars = c("^r_"))
mcmc_plot(brm2.ktype, pars = c("^b_"))
conditional_effects(brm2.ktype)
parnames(brm2.ktype)

brm2.family <- brm(dprime ~ elev.mean + (1 + elev.mean | bb.sp), 
                  control = list(max_treedepth = 10),
                  data = filter(bias.sitedate, criterion ==  "plant.family"))
summary(brm2.family)
mcmc_plot(brm2.family, pars = c("^r_"))
mcmc_plot(brm2.family, pars = c("^b_"))
conditional_effects(brm2.family)
parnames(brm2.family)

brm2.family <- brm(dprime ~ elev.mean, 
                  control = list(max_treedepth = 10),
                  data = filter(bias.sitedate, criterion ==  "plant.family"))
summary(brm2.family)
mcmc_plot(brm2.family, pars = c("^r_"))
mcmc_plot(brm2.family, pars = c("^b_"))
conditional_effects(brm2.family)
parnames(brm2.family)
```

## 1.6 Define hypotheses
```{r}
# Is d' different from 0?
H1 <- paste("Intercept = 0")

# Are some species more biased in their foraging than others? 
H.bss_gers <- paste("r_bb.sp[bss,Intercept] = r_bb.sp[gers,Intercept]")
H.bss_hort <- paste("r_bb.sp[bss,Intercept] = r_bb.sp[hort,Intercept]")
H.bss_humi <- paste("r_bb.sp[bss,Intercept] = r_bb.sp[humi,Intercept]")
H.bss_hypn <- paste("r_bb.sp[bss,Intercept] = r_bb.sp[hypn,Intercept]")
H.bss_jone <- paste("r_bb.sp[bss,Intercept] = r_bb.sp[jone,Intercept]")
H.bss_lapi <- paste("r_bb.sp[bss,Intercept] = r_bb.sp[lapi,Intercept]")
H.bss_mend <- paste("r_bb.sp[bss,Intercept] = r_bb.sp[mend,Intercept]")
H.bss_mont <- paste("r_bb.sp[bss,Intercept] = r_bb.sp[mont,Intercept]")
H.bss_muci <- paste("r_bb.sp[bss,Intercept] = r_bb.sp[muci,Intercept]")
H.bss_pasc <- paste("r_bb.sp[bss,Intercept] = r_bb.sp[pasc,Intercept]")
H.bss_prat <- paste("r_bb.sp[bss,Intercept] = r_bb.sp[prat,Intercept]")
H.bss_psit <- paste("r_bb.sp[bss,Intercept] = r_bb.sp[psit,Intercept]")
H.bss_pyre <- paste("r_bb.sp[bss,Intercept] = r_bb.sp[pyre,Intercept]")
H.bss_soro <- paste("r_bb.sp[bss,Intercept] = r_bb.sp[soro,Intercept]")
H.bss_wurf <- paste("r_bb.sp[bss,Intercept] = r_bb.sp[wurf,Intercept]")

H.gers_hort <- paste("r_bb.sp[gers,Intercept] = r_bb.sp[hort,Intercept]")
H.gers_humi <- paste("r_bb.sp[gers,Intercept] = r_bb.sp[humi,Intercept]")
H.gers_hypn <- paste("r_bb.sp[gers,Intercept] = r_bb.sp[hypn,Intercept]")
H.gers_jone <- paste("r_bb.sp[gers,Intercept] = r_bb.sp[jone,Intercept]")
H.gers_lapi <- paste("r_bb.sp[gers,Intercept] = r_bb.sp[lapi,Intercept]")
H.gers_mend <- paste("r_bb.sp[gers,Intercept] = r_bb.sp[mend,Intercept]")
H.gers_mont <- paste("r_bb.sp[gers,Intercept] = r_bb.sp[mont,Intercept]")
H.gers_muci <- paste("r_bb.sp[gers,Intercept] = r_bb.sp[muci,Intercept]")
H.gers_pasc <- paste("r_bb.sp[gers,Intercept] = r_bb.sp[pasc,Intercept]")
H.gers_prat <- paste("r_bb.sp[gers,Intercept] = r_bb.sp[prat,Intercept]")
H.gers_psit <- paste("r_bb.sp[gers,Intercept] = r_bb.sp[psit,Intercept]")
H.gers_pyre <- paste("r_bb.sp[gers,Intercept] = r_bb.sp[pyre,Intercept]")
H.gers_soro <- paste("r_bb.sp[gers,Intercept] = r_bb.sp[soro,Intercept]")
H.gers_wurf <- paste("r_bb.sp[gers,Intercept] = r_bb.sp[wurf,Intercept]")

H.hort_humi <- paste("r_bb.sp[hort,Intercept] = r_bb.sp[humi,Intercept]")
H.hort_hypn <- paste("r_bb.sp[hort,Intercept] = r_bb.sp[hypn,Intercept]")
H.hort_jone <- paste("r_bb.sp[hort,Intercept] = r_bb.sp[jone,Intercept]")
H.hort_lapi <- paste("r_bb.sp[hort,Intercept] = r_bb.sp[lapi,Intercept]")
H.hort_mend <- paste("r_bb.sp[hort,Intercept] = r_bb.sp[mend,Intercept]")
H.hort_mont <- paste("r_bb.sp[hort,Intercept] = r_bb.sp[mont,Intercept]")
H.hort_muci <- paste("r_bb.sp[hort,Intercept] = r_bb.sp[muci,Intercept]")
H.hort_pasc <- paste("r_bb.sp[hort,Intercept] = r_bb.sp[pasc,Intercept]")
H.hort_prat <- paste("r_bb.sp[hort,Intercept] = r_bb.sp[prat,Intercept]")
H.hort_psit <- paste("r_bb.sp[hort,Intercept] = r_bb.sp[psit,Intercept]")
H.hort_pyre <- paste("r_bb.sp[hort,Intercept] = r_bb.sp[pyre,Intercept]")
H.hort_soro <- paste("r_bb.sp[hort,Intercept] = r_bb.sp[soro,Intercept]")
H.hort_wurf <- paste("r_bb.sp[hort,Intercept] = r_bb.sp[wurf,Intercept]")

H.humi_hypn <- paste("r_bb.sp[humi,Intercept] = r_bb.sp[hypn,Intercept]")
H.humi_jone <- paste("r_bb.sp[humi,Intercept] = r_bb.sp[jone,Intercept]")
H.humi_lapi <- paste("r_bb.sp[humi,Intercept] = r_bb.sp[lapi,Intercept]")
H.humi_mend <- paste("r_bb.sp[humi,Intercept] = r_bb.sp[mend,Intercept]")
H.humi_mont <- paste("r_bb.sp[humi,Intercept] = r_bb.sp[mont,Intercept]")
H.humi_muci <- paste("r_bb.sp[humi,Intercept] = r_bb.sp[muci,Intercept]")
H.humi_pasc <- paste("r_bb.sp[humi,Intercept] = r_bb.sp[pasc,Intercept]")
H.humi_prat <- paste("r_bb.sp[humi,Intercept] = r_bb.sp[prat,Intercept]")
H.humi_psit <- paste("r_bb.sp[humi,Intercept] = r_bb.sp[psit,Intercept]")
H.humi_pyre <- paste("r_bb.sp[humi,Intercept] = r_bb.sp[pyre,Intercept]")
H.humi_soro <- paste("r_bb.sp[humi,Intercept] = r_bb.sp[soro,Intercept]")
H.humi_wurf <- paste("r_bb.sp[humi,Intercept] = r_bb.sp[wurf,Intercept]")

H.hypn_jone <- paste("r_bb.sp[hypn,Intercept] = r_bb.sp[jone,Intercept]")
H.hypn_lapi <- paste("r_bb.sp[hypn,Intercept] = r_bb.sp[lapi,Intercept]")
H.hypn_mend <- paste("r_bb.sp[hypn,Intercept] = r_bb.sp[mend,Intercept]")
H.hypn_mont <- paste("r_bb.sp[hypn,Intercept] = r_bb.sp[mont,Intercept]")
H.hypn_muci <- paste("r_bb.sp[hypn,Intercept] = r_bb.sp[muci,Intercept]")
H.hypn_pasc <- paste("r_bb.sp[hypn,Intercept] = r_bb.sp[pasc,Intercept]")
H.hypn_prat <- paste("r_bb.sp[hypn,Intercept] = r_bb.sp[prat,Intercept]")
H.hypn_psit <- paste("r_bb.sp[hypn,Intercept] = r_bb.sp[psit,Intercept]")
H.hypn_pyre <- paste("r_bb.sp[hypn,Intercept] = r_bb.sp[pyre,Intercept]")
H.hypn_soro <- paste("r_bb.sp[hypn,Intercept] = r_bb.sp[soro,Intercept]")
H.hypn_wurf <- paste("r_bb.sp[hypn,Intercept] = r_bb.sp[wurf,Intercept]")

H.jone_lapi <- paste("r_bb.sp[jone,Intercept] = r_bb.sp[lapi,Intercept]")
H.jone_mend <- paste("r_bb.sp[jone,Intercept] = r_bb.sp[mend,Intercept]")
H.jone_mont <- paste("r_bb.sp[jone,Intercept] = r_bb.sp[mont,Intercept]")
H.jone_muci <- paste("r_bb.sp[jone,Intercept] = r_bb.sp[muci,Intercept]")
H.jone_pasc <- paste("r_bb.sp[jone,Intercept] = r_bb.sp[pasc,Intercept]")
H.jone_prat <- paste("r_bb.sp[jone,Intercept] = r_bb.sp[prat,Intercept]")
H.jone_psit <- paste("r_bb.sp[jone,Intercept] = r_bb.sp[psit,Intercept]")
H.jone_pyre <- paste("r_bb.sp[jone,Intercept] = r_bb.sp[pyre,Intercept]")
H.jone_soro <- paste("r_bb.sp[jone,Intercept] = r_bb.sp[soro,Intercept]")
H.jone_wurf <- paste("r_bb.sp[jone,Intercept] = r_bb.sp[wurf,Intercept]")

H.lapi_mend <- paste("r_bb.sp[lapi,Intercept] = r_bb.sp[mend,Intercept]")
H.lapi_mont <- paste("r_bb.sp[lapi,Intercept] = r_bb.sp[mont,Intercept]")
H.lapi_muci <- paste("r_bb.sp[lapi,Intercept] = r_bb.sp[muci,Intercept]")
H.lapi_pasc <- paste("r_bb.sp[lapi,Intercept] = r_bb.sp[pasc,Intercept]")
H.lapi_prat <- paste("r_bb.sp[lapi,Intercept] = r_bb.sp[prat,Intercept]")
H.lapi_psit <- paste("r_bb.sp[lapi,Intercept] = r_bb.sp[psit,Intercept]")
H.lapi_pyre <- paste("r_bb.sp[lapi,Intercept] = r_bb.sp[pyre,Intercept]")
H.lapi_soro <- paste("r_bb.sp[lapi,Intercept] = r_bb.sp[soro,Intercept]")
H.lapi_wurf <- paste("r_bb.sp[lapi,Intercept] = r_bb.sp[wurf,Intercept]")

H.mend_mont <- paste("r_bb.sp[mend,Intercept] = r_bb.sp[mont,Intercept]")
H.mend_muci <- paste("r_bb.sp[mend,Intercept] = r_bb.sp[muci,Intercept]")
H.mend_pasc <- paste("r_bb.sp[mend,Intercept] = r_bb.sp[pasc,Intercept]")
H.mend_prat <- paste("r_bb.sp[mend,Intercept] = r_bb.sp[prat,Intercept]")
H.mend_psit <- paste("r_bb.sp[mend,Intercept] = r_bb.sp[psit,Intercept]")
H.mend_pyre <- paste("r_bb.sp[mend,Intercept] = r_bb.sp[pyre,Intercept]")
H.mend_soro <- paste("r_bb.sp[mend,Intercept] = r_bb.sp[soro,Intercept]")
H.mend_wurf <- paste("r_bb.sp[mend,Intercept] = r_bb.sp[wurf,Intercept]")

H.mont_muci <- paste("r_bb.sp[mont,Intercept] = r_bb.sp[muci,Intercept]")
H.mont_pasc <- paste("r_bb.sp[mont,Intercept] = r_bb.sp[pasc,Intercept]")
H.mont_prat <- paste("r_bb.sp[mont,Intercept] = r_bb.sp[prat,Intercept]")
H.mont_psit <- paste("r_bb.sp[mont,Intercept] = r_bb.sp[psit,Intercept]")
H.mont_pyre <- paste("r_bb.sp[mont,Intercept] = r_bb.sp[pyre,Intercept]")
H.mont_soro <- paste("r_bb.sp[mont,Intercept] = r_bb.sp[soro,Intercept]")
H.mont_wurf <- paste("r_bb.sp[mont,Intercept] = r_bb.sp[wurf,Intercept]")

H.muci_pasc <- paste("r_bb.sp[muci,Intercept] = r_bb.sp[pasc,Intercept]")
H.muci_prat <- paste("r_bb.sp[muci,Intercept] = r_bb.sp[prat,Intercept]")
H.muci_psit <- paste("r_bb.sp[muci,Intercept] = r_bb.sp[psit,Intercept]")
H.muci_pyre <- paste("r_bb.sp[muci,Intercept] = r_bb.sp[pyre,Intercept]")
H.muci_soro <- paste("r_bb.sp[muci,Intercept] = r_bb.sp[soro,Intercept]")
H.muci_wurf <- paste("r_bb.sp[muci,Intercept] = r_bb.sp[wurf,Intercept]")

H.pasc_prat <- paste("r_bb.sp[pasc,Intercept] = r_bb.sp[prat,Intercept]")
H.pasc_psit <- paste("r_bb.sp[pasc,Intercept] = r_bb.sp[psit,Intercept]")
H.pasc_pyre <- paste("r_bb.sp[pasc,Intercept] = r_bb.sp[pyre,Intercept]")
H.pasc_soro <- paste("r_bb.sp[pasc,Intercept] = r_bb.sp[soro,Intercept]")
H.pasc_wurf <- paste("r_bb.sp[pasc,Intercept] = r_bb.sp[wurf,Intercept]")

H.prat_psit <- paste("r_bb.sp[prat,Intercept] = r_bb.sp[psit,Intercept]")
H.prat_pyre <- paste("r_bb.sp[prat,Intercept] = r_bb.sp[pyre,Intercept]")
H.prat_soro <- paste("r_bb.sp[prat,Intercept] = r_bb.sp[soro,Intercept]")
H.prat_wurf <- paste("r_bb.sp[prat,Intercept] = r_bb.sp[wurf,Intercept]")

H.psit_pyre <- paste("r_bb.sp[psit,Intercept] = r_bb.sp[pyre,Intercept]")
H.psit_soro <- paste("r_bb.sp[psit,Intercept] = r_bb.sp[soro,Intercept]")
H.psit_wurf <- paste("r_bb.sp[psit,Intercept] = r_bb.sp[wurf,Intercept]")

H.pyre_soro <- paste("r_bb.sp[pyre,Intercept] = r_bb.sp[soro,Intercept]")
H.pyre_wurf <- paste("r_bb.sp[pyre,Intercept] = r_bb.sp[wurf,Intercept]")

H.soro_wurf <- paste("r_bb.sp[soro,Intercept] = r_bb.sp[wurf,Intercept]")
```

## 1.7 Test contrasts
```{r}
### Plant.sp
hypothesis(brm.plantsp, H1)

hypothesis(brm.plantsp, H.bss_gers, class = "")
hypothesis(brm.plantsp, H.bss_hort, class = "")
hypothesis(brm.plantsp, H.bss_humi, class = "")
hypothesis(brm.plantsp, H.bss_hypn, class = "")
hypothesis(brm.plantsp, H.bss_jone, class = "")
hypothesis(brm.plantsp, H.bss_lapi, class = "")
hypothesis(brm.plantsp, H.bss_mend, class = "")
hypothesis(brm.plantsp, H.bss_mont, class = "")
hypothesis(brm.plantsp, H.bss_muci, class = "")
hypothesis(brm.plantsp, H.bss_pasc, class = "")
hypothesis(brm.plantsp, H.bss_prat, class = "")
hypothesis(brm.plantsp, H.bss_psit, class = "")
hypothesis(brm.plantsp, H.bss_pyre, class = "")
hypothesis(brm.plantsp, H.bss_soro, class = "")
hypothesis(brm.plantsp, H.bss_wurf, class = "")

hypothesis(brm.plantsp, H.gers_hort, class = "")
hypothesis(brm.plantsp, H.gers_humi, class = "")
hypothesis(brm.plantsp, H.gers_hypn, class = "")
hypothesis(brm.plantsp, H.gers_jone, class = "")
hypothesis(brm.plantsp, H.gers_lapi, class = "")
hypothesis(brm.plantsp, H.gers_mend, class = "")
hypothesis(brm.plantsp, H.gers_mont, class = "")
hypothesis(brm.plantsp, H.gers_muci, class = "")
hypothesis(brm.plantsp, H.gers_pasc, class = "")
hypothesis(brm.plantsp, H.gers_prat, class = "")
hypothesis(brm.plantsp, H.gers_psit, class = "")
hypothesis(brm.plantsp, H.gers_pyre, class = "")
hypothesis(brm.plantsp, H.gers_soro, class = "")
hypothesis(brm.plantsp, H.gers_wurf, class = "")

hypothesis(brm.plantsp, H.hort_humi, class = "")
hypothesis(brm.plantsp, H.hort_hypn, class = "")
hypothesis(brm.plantsp, H.hort_jone, class = "")
hypothesis(brm.plantsp, H.hort_lapi, class = "")
hypothesis(brm.plantsp, H.hort_mend, class = "")
hypothesis(brm.plantsp, H.hort_mont, class = "")
hypothesis(brm.plantsp, H.hort_muci, class = "")
hypothesis(brm.plantsp, H.hort_pasc, class = "")
hypothesis(brm.plantsp, H.hort_prat, class = "")
hypothesis(brm.plantsp, H.hort_psit, class = "")
hypothesis(brm.plantsp, H.hort_pyre, class = "")
hypothesis(brm.plantsp, H.hort_soro, class = "")
hypothesis(brm.plantsp, H.hort_wurf, class = "")

hypothesis(brm.plantsp, H.humi_hypn, class = "")
hypothesis(brm.plantsp, H.humi_jone, class = "")
hypothesis(brm.plantsp, H.humi_lapi, class = "")
hypothesis(brm.plantsp, H.humi_mend, class = "")
hypothesis(brm.plantsp, H.humi_mont, class = "")
hypothesis(brm.plantsp, H.humi_muci, class = "")
hypothesis(brm.plantsp, H.humi_pasc, class = "")
hypothesis(brm.plantsp, H.humi_prat, class = "")
hypothesis(brm.plantsp, H.humi_psit, class = "")
hypothesis(brm.plantsp, H.humi_pyre, class = "")
hypothesis(brm.plantsp, H.humi_soro, class = "")
hypothesis(brm.plantsp, H.humi_wurf, class = "")

hypothesis(brm.plantsp, H.hypn_jone, class = "")
hypothesis(brm.plantsp, H.hypn_lapi, class = "")
hypothesis(brm.plantsp, H.hypn_mend, class = "")
hypothesis(brm.plantsp, H.hypn_mont, class = "")
hypothesis(brm.plantsp, H.hypn_muci, class = "")
hypothesis(brm.plantsp, H.hypn_pasc, class = "")
hypothesis(brm.plantsp, H.hypn_prat, class = "")
hypothesis(brm.plantsp, H.hypn_psit, class = "")
hypothesis(brm.plantsp, H.hypn_pyre, class = "")
hypothesis(brm.plantsp, H.hypn_soro, class = "")
hypothesis(brm.plantsp, H.hypn_wurf, class = "")

hypothesis(brm.plantsp, H.jone_lapi, class = "")
hypothesis(brm.plantsp, H.jone_mend, class = "")
hypothesis(brm.plantsp, H.jone_mont, class = "")
hypothesis(brm.plantsp, H.jone_muci, class = "")
hypothesis(brm.plantsp, H.jone_pasc, class = "")
hypothesis(brm.plantsp, H.jone_prat, class = "")
hypothesis(brm.plantsp, H.jone_psit, class = "")
hypothesis(brm.plantsp, H.jone_pyre, class = "")
hypothesis(brm.plantsp, H.jone_soro, class = "")
hypothesis(brm.plantsp, H.jone_wurf, class = "")

hypothesis(brm.plantsp, H.lapi_mend, class = "")
hypothesis(brm.plantsp, H.lapi_mont, class = "")
hypothesis(brm.plantsp, H.lapi_muci, class = "")
hypothesis(brm.plantsp, H.lapi_pasc, class = "")
hypothesis(brm.plantsp, H.lapi_prat, class = "")
hypothesis(brm.plantsp, H.lapi_psit, class = "")
hypothesis(brm.plantsp, H.lapi_pyre, class = "")
hypothesis(brm.plantsp, H.lapi_soro, class = "")
hypothesis(brm.plantsp, H.lapi_wurf, class = "")

hypothesis(brm.plantsp, H.mend_mont, class = "")
hypothesis(brm.plantsp, H.mend_muci, class = "")
hypothesis(brm.plantsp, H.mend_pasc, class = "")
hypothesis(brm.plantsp, H.mend_prat, class = "")
hypothesis(brm.plantsp, H.mend_psit, class = "")
hypothesis(brm.plantsp, H.mend_pyre, class = "")
hypothesis(brm.plantsp, H.mend_soro, class = "")
hypothesis(brm.plantsp, H.mend_wurf, class = "")

hypothesis(brm.plantsp, H.mont_muci, class = "")
hypothesis(brm.plantsp, H.mont_pasc, class = "")
hypothesis(brm.plantsp, H.mont_prat, class = "")
hypothesis(brm.plantsp, H.mont_psit, class = "")
hypothesis(brm.plantsp, H.mont_pyre, class = "")
hypothesis(brm.plantsp, H.mont_soro, class = "")
hypothesis(brm.plantsp, H.mont_wurf, class = "")

hypothesis(brm.plantsp, H.muci_pasc, class = "")
hypothesis(brm.plantsp, H.muci_prat, class = "")
hypothesis(brm.plantsp, H.muci_psit, class = "")
hypothesis(brm.plantsp, H.muci_pyre, class = "")
hypothesis(brm.plantsp, H.muci_soro, class = "")
hypothesis(brm.plantsp, H.muci_wurf, class = "")

hypothesis(brm.plantsp, H.pasc_prat, class = "")
hypothesis(brm.plantsp, H.pasc_psit, class = "")
hypothesis(brm.plantsp, H.pasc_pyre, class = "")
hypothesis(brm.plantsp, H.pasc_soro, class = "")
hypothesis(brm.plantsp, H.pasc_wurf, class = "")

hypothesis(brm.plantsp, H.prat_psit, class = "")
hypothesis(brm.plantsp, H.prat_pyre, class = "")
hypothesis(brm.plantsp, H.prat_soro, class = "")
hypothesis(brm.plantsp, H.prat_wurf, class = "")

hypothesis(brm.plantsp, H.psit_pyre, class = "")
hypothesis(brm.plantsp, H.psit_soro, class = "")
hypothesis(brm.plantsp, H.psit_wurf, class = "")

hypothesis(brm.plantsp, H.pyre_soro, class = "")
hypothesis(brm.plantsp, H.pyre_wurf, class = "")

hypothesis(brm.plantsp, H.soro_wurf, class = "")

### Kugler type
hypothesis(brm.ktype, H1)

hypothesis(brm.ktype, H.bss_gers, class = "")
hypothesis(brm.ktype, H.bss_hort, class = "")
hypothesis(brm.ktype, H.bss_humi, class = "")
hypothesis(brm.ktype, H.bss_hypn, class = "")
hypothesis(brm.ktype, H.bss_jone, class = "")
hypothesis(brm.ktype, H.bss_lapi, class = "")
hypothesis(brm.ktype, H.bss_mend, class = "")
hypothesis(brm.ktype, H.bss_mont, class = "")
hypothesis(brm.ktype, H.bss_muci, class = "")
hypothesis(brm.ktype, H.bss_pasc, class = "")
hypothesis(brm.ktype, H.bss_prat, class = "")
hypothesis(brm.ktype, H.bss_psit, class = "")
hypothesis(brm.ktype, H.bss_pyre, class = "")
hypothesis(brm.ktype, H.bss_soro, class = "")
hypothesis(brm.ktype, H.bss_wurf, class = "")

hypothesis(brm.ktype, H.gers_hort, class = "")
hypothesis(brm.ktype, H.gers_humi, class = "")
hypothesis(brm.ktype, H.gers_hypn, class = "")
hypothesis(brm.ktype, H.gers_jone, class = "")
hypothesis(brm.ktype, H.gers_lapi, class = "")
hypothesis(brm.ktype, H.gers_mend, class = "")
hypothesis(brm.ktype, H.gers_mont, class = "")
hypothesis(brm.ktype, H.gers_muci, class = "")
hypothesis(brm.ktype, H.gers_pasc, class = "")
hypothesis(brm.ktype, H.gers_prat, class = "")
hypothesis(brm.ktype, H.gers_psit, class = "")
hypothesis(brm.ktype, H.gers_pyre, class = "")
hypothesis(brm.ktype, H.gers_soro, class = "")
hypothesis(brm.ktype, H.gers_wurf, class = "")

hypothesis(brm.ktype, H.hort_humi, class = "")
hypothesis(brm.ktype, H.hort_hypn, class = "")
hypothesis(brm.ktype, H.hort_jone, class = "")
hypothesis(brm.ktype, H.hort_lapi, class = "")
hypothesis(brm.ktype, H.hort_mend, class = "")
hypothesis(brm.ktype, H.hort_mont, class = "")
hypothesis(brm.ktype, H.hort_muci, class = "")
hypothesis(brm.ktype, H.hort_pasc, class = "")
hypothesis(brm.ktype, H.hort_prat, class = "")
hypothesis(brm.ktype, H.hort_psit, class = "")
hypothesis(brm.ktype, H.hort_pyre, class = "")
hypothesis(brm.ktype, H.hort_soro, class = "")
hypothesis(brm.ktype, H.hort_wurf, class = "")

hypothesis(brm.ktype, H.humi_hypn, class = "")
hypothesis(brm.ktype, H.humi_jone, class = "")
hypothesis(brm.ktype, H.humi_lapi, class = "")
hypothesis(brm.ktype, H.humi_mend, class = "")
hypothesis(brm.ktype, H.humi_mont, class = "")
hypothesis(brm.ktype, H.humi_muci, class = "")
hypothesis(brm.ktype, H.humi_pasc, class = "")
hypothesis(brm.ktype, H.humi_prat, class = "")
hypothesis(brm.ktype, H.humi_psit, class = "")
hypothesis(brm.ktype, H.humi_pyre, class = "")
hypothesis(brm.ktype, H.humi_soro, class = "")
hypothesis(brm.ktype, H.humi_wurf, class = "")

hypothesis(brm.ktype, H.hypn_jone, class = "")
hypothesis(brm.ktype, H.hypn_lapi, class = "")
hypothesis(brm.ktype, H.hypn_mend, class = "")
hypothesis(brm.ktype, H.hypn_mont, class = "")
hypothesis(brm.ktype, H.hypn_muci, class = "")
hypothesis(brm.ktype, H.hypn_pasc, class = "")
hypothesis(brm.ktype, H.hypn_prat, class = "")
hypothesis(brm.ktype, H.hypn_psit, class = "")
hypothesis(brm.ktype, H.hypn_pyre, class = "")
hypothesis(brm.ktype, H.hypn_soro, class = "")
hypothesis(brm.ktype, H.hypn_wurf, class = "")

hypothesis(brm.ktype, H.jone_lapi, class = "")
hypothesis(brm.ktype, H.jone_mend, class = "")
hypothesis(brm.ktype, H.jone_mont, class = "")
hypothesis(brm.ktype, H.jone_muci, class = "")
hypothesis(brm.ktype, H.jone_pasc, class = "")
hypothesis(brm.ktype, H.jone_prat, class = "")
hypothesis(brm.ktype, H.jone_psit, class = "")
hypothesis(brm.ktype, H.jone_pyre, class = "")
hypothesis(brm.ktype, H.jone_soro, class = "")
hypothesis(brm.ktype, H.jone_wurf, class = "")

hypothesis(brm.ktype, H.lapi_mend, class = "")
hypothesis(brm.ktype, H.lapi_mont, class = "")
hypothesis(brm.ktype, H.lapi_muci, class = "")
hypothesis(brm.ktype, H.lapi_pasc, class = "")
hypothesis(brm.ktype, H.lapi_prat, class = "")
hypothesis(brm.ktype, H.lapi_psit, class = "")
hypothesis(brm.ktype, H.lapi_pyre, class = "")
hypothesis(brm.ktype, H.lapi_soro, class = "")
hypothesis(brm.ktype, H.lapi_wurf, class = "")

hypothesis(brm.ktype, H.mend_mont, class = "")
hypothesis(brm.ktype, H.mend_muci, class = "")
hypothesis(brm.ktype, H.mend_pasc, class = "")
hypothesis(brm.ktype, H.mend_prat, class = "")
hypothesis(brm.ktype, H.mend_psit, class = "")
hypothesis(brm.ktype, H.mend_pyre, class = "")
hypothesis(brm.ktype, H.mend_soro, class = "")
hypothesis(brm.ktype, H.mend_wurf, class = "")

hypothesis(brm.ktype, H.mont_muci, class = "")
hypothesis(brm.ktype, H.mont_pasc, class = "")
hypothesis(brm.ktype, H.mont_prat, class = "")
hypothesis(brm.ktype, H.mont_psit, class = "")
hypothesis(brm.ktype, H.mont_pyre, class = "")
hypothesis(brm.ktype, H.mont_soro, class = "")
hypothesis(brm.ktype, H.mont_wurf, class = "")

hypothesis(brm.ktype, H.muci_pasc, class = "")
hypothesis(brm.ktype, H.muci_prat, class = "")
hypothesis(brm.ktype, H.muci_psit, class = "")
hypothesis(brm.ktype, H.muci_pyre, class = "")
hypothesis(brm.ktype, H.muci_soro, class = "")
hypothesis(brm.ktype, H.muci_wurf, class = "")

hypothesis(brm.ktype, H.pasc_prat, class = "")
hypothesis(brm.ktype, H.pasc_psit, class = "")
hypothesis(brm.ktype, H.pasc_pyre, class = "")
hypothesis(brm.ktype, H.pasc_soro, class = "")
hypothesis(brm.ktype, H.pasc_wurf, class = "")

hypothesis(brm.ktype, H.prat_psit, class = "")
hypothesis(brm.ktype, H.prat_pyre, class = "")
hypothesis(brm.ktype, H.prat_soro, class = "")
hypothesis(brm.ktype, H.prat_wurf, class = "")

hypothesis(brm.ktype, H.psit_pyre, class = "")
hypothesis(brm.ktype, H.psit_soro, class = "")
hypothesis(brm.ktype, H.psit_wurf, class = "")

hypothesis(brm.ktype, H.pyre_soro, class = "")
hypothesis(brm.ktype, H.pyre_wurf, class = "")

hypothesis(brm.ktype, H.soro_wurf, class = "")

### Plant family
hypothesis(brm.family, H1)

hypothesis(brm.family, H.bss_gers, class = "")
hypothesis(brm.family, H.bss_hort, class = "")
hypothesis(brm.family, H.bss_humi, class = "")
hypothesis(brm.family, H.bss_hypn, class = "")
hypothesis(brm.family, H.bss_jone, class = "")
hypothesis(brm.family, H.bss_lapi, class = "")
hypothesis(brm.family, H.bss_mend, class = "")
hypothesis(brm.family, H.bss_mont, class = "")
hypothesis(brm.family, H.bss_muci, class = "")
hypothesis(brm.family, H.bss_pasc, class = "")
hypothesis(brm.family, H.bss_prat, class = "")
hypothesis(brm.family, H.bss_psit, class = "")
hypothesis(brm.family, H.bss_pyre, class = "")
hypothesis(brm.family, H.bss_soro, class = "")
hypothesis(brm.family, H.bss_wurf, class = "")

hypothesis(brm.family, H.gers_hort, class = "")
hypothesis(brm.family, H.gers_humi, class = "")
hypothesis(brm.family, H.gers_hypn, class = "")
hypothesis(brm.family, H.gers_jone, class = "")
hypothesis(brm.family, H.gers_lapi, class = "")
hypothesis(brm.family, H.gers_mend, class = "")
hypothesis(brm.family, H.gers_mont, class = "")
hypothesis(brm.family, H.gers_muci, class = "")
hypothesis(brm.family, H.gers_pasc, class = "")
hypothesis(brm.family, H.gers_prat, class = "")
hypothesis(brm.family, H.gers_psit, class = "")
hypothesis(brm.family, H.gers_pyre, class = "")
hypothesis(brm.family, H.gers_soro, class = "")
hypothesis(brm.family, H.gers_wurf, class = "")

hypothesis(brm.family, H.hort_humi, class = "")
hypothesis(brm.family, H.hort_hypn, class = "")
hypothesis(brm.family, H.hort_jone, class = "")
hypothesis(brm.family, H.hort_lapi, class = "")
hypothesis(brm.family, H.hort_mend, class = "")
hypothesis(brm.family, H.hort_mont, class = "")
hypothesis(brm.family, H.hort_muci, class = "")
hypothesis(brm.family, H.hort_pasc, class = "")
hypothesis(brm.family, H.hort_prat, class = "")
hypothesis(brm.family, H.hort_psit, class = "")
hypothesis(brm.family, H.hort_pyre, class = "")
hypothesis(brm.family, H.hort_soro, class = "")
hypothesis(brm.family, H.hort_wurf, class = "")

hypothesis(brm.family, H.humi_hypn, class = "")
hypothesis(brm.family, H.humi_jone, class = "")
hypothesis(brm.family, H.humi_lapi, class = "")
hypothesis(brm.family, H.humi_mend, class = "")
hypothesis(brm.family, H.humi_mont, class = "")
hypothesis(brm.family, H.humi_muci, class = "")
hypothesis(brm.family, H.humi_pasc, class = "")
hypothesis(brm.family, H.humi_prat, class = "")
hypothesis(brm.family, H.humi_psit, class = "")
hypothesis(brm.family, H.humi_pyre, class = "")
hypothesis(brm.family, H.humi_soro, class = "")
hypothesis(brm.family, H.humi_wurf, class = "")

hypothesis(brm.family, H.hypn_jone, class = "")
hypothesis(brm.family, H.hypn_lapi, class = "")
hypothesis(brm.family, H.hypn_mend, class = "")
hypothesis(brm.family, H.hypn_mont, class = "")
hypothesis(brm.family, H.hypn_muci, class = "")
hypothesis(brm.family, H.hypn_pasc, class = "")
hypothesis(brm.family, H.hypn_prat, class = "")
hypothesis(brm.family, H.hypn_psit, class = "")
hypothesis(brm.family, H.hypn_pyre, class = "")
hypothesis(brm.family, H.hypn_soro, class = "")
hypothesis(brm.family, H.hypn_wurf, class = "")

hypothesis(brm.family, H.jone_lapi, class = "")
hypothesis(brm.family, H.jone_mend, class = "")
hypothesis(brm.family, H.jone_mont, class = "")
hypothesis(brm.family, H.jone_muci, class = "")
hypothesis(brm.family, H.jone_pasc, class = "")
hypothesis(brm.family, H.jone_prat, class = "")
hypothesis(brm.family, H.jone_psit, class = "")
hypothesis(brm.family, H.jone_pyre, class = "")
hypothesis(brm.family, H.jone_soro, class = "")
hypothesis(brm.family, H.jone_wurf, class = "")

hypothesis(brm.family, H.lapi_mend, class = "")
hypothesis(brm.family, H.lapi_mont, class = "")
hypothesis(brm.family, H.lapi_muci, class = "")
hypothesis(brm.family, H.lapi_pasc, class = "")
hypothesis(brm.family, H.lapi_prat, class = "")
hypothesis(brm.family, H.lapi_psit, class = "")
hypothesis(brm.family, H.lapi_pyre, class = "")
hypothesis(brm.family, H.lapi_soro, class = "")
hypothesis(brm.family, H.lapi_wurf, class = "")

hypothesis(brm.family, H.mend_mont, class = "")
hypothesis(brm.family, H.mend_muci, class = "")
hypothesis(brm.family, H.mend_pasc, class = "")
hypothesis(brm.family, H.mend_prat, class = "")
hypothesis(brm.family, H.mend_psit, class = "")
hypothesis(brm.family, H.mend_pyre, class = "")
hypothesis(brm.family, H.mend_soro, class = "")
hypothesis(brm.family, H.mend_wurf, class = "")

hypothesis(brm.family, H.mont_muci, class = "")
hypothesis(brm.family, H.mont_pasc, class = "")
hypothesis(brm.family, H.mont_prat, class = "")
hypothesis(brm.family, H.mont_psit, class = "")
hypothesis(brm.family, H.mont_pyre, class = "")
hypothesis(brm.family, H.mont_soro, class = "")
hypothesis(brm.family, H.mont_wurf, class = "")

hypothesis(brm.family, H.muci_pasc, class = "")
hypothesis(brm.family, H.muci_prat, class = "")
hypothesis(brm.family, H.muci_psit, class = "")
hypothesis(brm.family, H.muci_pyre, class = "")
hypothesis(brm.family, H.muci_soro, class = "")
hypothesis(brm.family, H.muci_wurf, class = "")

hypothesis(brm.family, H.pasc_prat, class = "")
hypothesis(brm.family, H.pasc_psit, class = "")
hypothesis(brm.family, H.pasc_pyre, class = "")
hypothesis(brm.family, H.pasc_soro, class = "")
hypothesis(brm.family, H.pasc_wurf, class = "")

hypothesis(brm.family, H.prat_psit, class = "")
hypothesis(brm.family, H.prat_pyre, class = "")
hypothesis(brm.family, H.prat_soro, class = "")
hypothesis(brm.family, H.prat_wurf, class = "")

hypothesis(brm.family, H.psit_pyre, class = "")
hypothesis(brm.family, H.psit_soro, class = "")
hypothesis(brm.family, H.psit_wurf, class = "")

hypothesis(brm.family, H.pyre_soro, class = "")
hypothesis(brm.family, H.pyre_wurf, class = "")

hypothesis(brm.family, H.soro_wurf, class = "")
```

# 2. Having established the bias of BB foraging, let's zoom in on which morphotypes and families were over/under-represented
## 2.1 Calculate representation
```{r}
representation.ktype <- visit_cover %>%
  mutate(k.type.ss = factor(k.type.ss)) %>%
  group_by(site, date, bb.sp, k.type.ss) %>%
  summarize(flower.cover = sum(flower.cover),
            visits = sum(visits)) %>%
  group_by(site, date, bb.sp) %>%
  mutate(prop.flower.cover = flower.cover/sum(flower.cover),
         prop.visits = visits/sum(visits)) %>%
  mutate(representation = prop.visits - prop.flower.cover,
         representation.exp = if_else(visits == 0, 0, exp(prop.visits - prop.flower.cover)),
         representation.bin = if_else(representation > 0, 1, 0)) %>%
  left_join(site_data) %>%
  mutate(yday = yday(date)) %>%
  select(site, elev.mean, date, yday, bb.sp, k.type.ss, representation, representation.bin)

representation.family <- visit_cover %>%
  mutate(plant.family = factor(plant.family)) %>%
  group_by(site, date, bb.sp, plant.family) %>%
  summarize(flower.cover = sum(flower.cover),
            visits = sum(visits)) %>%
  group_by(site, date, bb.sp) %>%
  mutate(prop.flower.cover = flower.cover/sum(flower.cover),
         prop.visits = visits/sum(visits)) %>%
  mutate(representation = prop.visits - prop.flower.cover,
         representation.exp = if_else(visits == 0, 0, exp(prop.visits - prop.flower.cover)),
         representation.bin = if_else(representation > 0, 1, 0)) %>%
  left_join(site_data) %>%
  mutate(yday = yday(date)) %>%
  select(site, elev.mean, date, yday, bb.sp, plant.family, representation, representation.bin)
```

## 2.2 Visualize
```{r}
ggplot(representation.ktype, aes(representation)) +
  geom_histogram()

ggplot(representation.ktype, aes(representation.bin)) +
  geom_bar()

ggplot(representation.ktype, aes(k.type.ss, representation)) +
  geom_boxplot() +
  #geom_hline(yintercept = 1, linetype = "dashed")+
  facet_wrap(~bb.sp)

ggplot(representation.ktype, aes(k.type.ss, fill = representation.bin)) +
  geom_bar() +
  facet_wrap(~bb.sp)

ggplot(representation.ktype, aes(elev.mean, representation)) +
  geom_point() +
  geom_smooth(method = "gam", formula = y ~ s(x, k = 5)) +
  facet_wrap(~k.type.ss)

ggplot(representation.family, aes(plant.family, representation)) +
  geom_boxplot() +
  facet_wrap(~bb.sp)

ggplot(representation.family, aes(elev.mean, representation)) +
  geom_point() +
  geom_smooth(method = "gam", formula = y ~ s(x, k = 5)) +
  facet_wrap(~plant.family)

ggplot(representation.family, aes(yday, exp(representation))) +
  geom_point(alpha = 0.1) +
  geom_smooth(method = "gam", formula = y ~ s(x, k = 3), se = FALSE) +
  facet_wrap(~plant.family)
```

## 2.3 Explaining representation by floral traits, BB spp, and elevation (and maybe time)
```{r}
repmod.ktype <- brm(representation ~ (1 | k.type.ss), 
                    data = representation.ktype)
mcmc_plot(repmod.ktype)
conditional_effects(repmod.ktype)

repmod.ktype2 <- update(repmod.ktype, formula. = representation ~ elev.mean + (1 + elev.mean | k.type.ss))
```

# 3. Can we explain bias by BB-Fl trait associations? (RLQ + fourth corner analysis)
## 3.1
## 3.2 Do BB-FL trait associations vary with elevation?

# 4. Can we explain BB-FL trait associations in terms of floral reward or handling time? We can use Neumayer and Paulus' (1999) data. 
## 4.1 
## 4.2 Do there appear to be suboptimalities in BB-FL trait associations? That is, are trait classes over/under-represented relative to their reward/efficiency? If so, how might explain such suboptimalities?  
  - Innate bias trumps learning?
  - Ecological modulators are involved, e.g. competition or distribution (see Sowig 1989)