### 2021-01-22 code/visitation_bias.Rmd
backup

---
title: "visitation_bias"
author: "Doug Sponsler"
date: "12/14/2020"
output: html_document
---
# Preparations
## Packages 
```{r, warning=FALSE, include=FALSE, echo = FALSE}
library(bipartite)
library(brms)
library(lme4)
library(sjPlot)
library(DHARMa)
library(bayesplot)
library(tidybayes)
library(modelr)
library(ggdist)
library(vegan)
library(ggvegan)
library(ggrepel)
library(treemapify)
library(patchwork)
library(tidymodels)
library(tidyverse)
library(lubridate)

rstan::rstan_options(auto_write = TRUE)
options(mc.cores = parallel::detectCores())
```

## Load data
```{r message=FALSE, warning=FALSE, include=FALSE, echo = FALSE}
# Site data
site_data <- read_csv("../data/processed_data/site_data.csv") %>%
  dplyr::select(-c(slope.calc, slope.est, elev.min, elev.max, temp.mean)) %>% # drop these variables
  mutate(elev.class2 = factor(elev.class2, levels = c("low", "mid", "high")),
         site = factor(site)) # turn elav.class into an ordered factor and site into a regular factor

# BB traits
bb_traits <- read_csv("../data/processed_data/bb_traits.csv")

ggplot(bb_traits, aes(pbl.w, pbl.w.durieaux,
                      color = bb.sp, pch = factor(pbl.w.class2))) +
  geom_point()

# FL traits
## family-level taxonomy
fl_tax <- read_csv("../data/processed_data/floral_tax.csv")

## Kugler morphotypes (accessed through Bioflor)
fl_ktype <- read_csv("../data/processed_data/floral_k_type.csv")

## Flower colors (accessed through BioFlor)
fl_color <- read_csv("../data/processed_data/floral_color.csv")

## diameter and corolla depth data from Gita Benadi
fl_morph <- read_csv("../data/processed_data/flower_morphology_GitaBenadi.csv") %>% 
  dplyr::select(SampleID, plant.family = Family, plant.sp = Species, 
                Unit, Flowers_per_unit, diam = Diameter, LN) %>%
  group_by(plant.sp) %>%
  summarize(diam = mean(diam), LN = mean(LN))

## verbal description of Kugler morphotypes
fl_ktype_key <- read_delim("../data/processed_data/kugler_key.tsv", delim = "\t", col_names = FALSE) %>%
  dplyr::select(k.type = 1, description = 2)

## join all floral trait tables
fl_traits <- fl_tax %>%
  full_join(fl_ktype) %>%
  full_join(fl_color) %>%
  full_join(fl_morph) %>%
  dplyr::select(plant.sp, plant.genus, plant.family, k.type, k.type.s, k.type.ss, color, diam, LN)
```

## Align survey and visitation data with pseudodates
```{r}
# Site-dates where BB visitation was recorded (excluding dates where only males were recorded)
visitation_samples <- read_csv("../data/processed_data/network.csv") %>%
  filter(caste != "male") %>%
  select(site, date) %>%
  distinct()

# Site dates where floral surveying was conducted
survey_samples <- read_csv("../data/processed_data/floral_survey.csv") %>%
  select(site, date) %>%
  distinct()

# Site-dates where visitation sampling and floral sampling coincided
shared_dates <- inner_join(visitation_samples, survey_samples)

# Site-dates where visitation was recorded WITHOUT corresponding survey data
visitation_only <- anti_join(visitation_samples, survey_samples) %>%
  mutate(date.v = date) %>%
  select(-date)

# This creates a pseudo date column for unmatched visitation site-dates that will match each visitation date
# to its nearest survey data. For site t4, there is a 2-year mismatch; we will drop these site-dates later.
pseudodates <- full_join(visitation_only, survey_samples) %>%
  mutate(mismatch = abs(date.v - date)) %>%
  group_by(site, date.v) %>%
  filter(mismatch == min(mismatch)) %>%
  rename(date = date.v, pseudo.date = date)
```

## Generate a data frame that represents visitation and survey abundance
```{r}
# BB visitation data
visitation <- read_csv("../data/processed_data/network.csv") %>%
  filter(caste != "male") %>% # we will ignore males 
  group_by(site, date, bb.sp, plant.sp) %>%
  summarize(visits = n()) %>% # visits per BB-FL pair
  ungroup() %>%
  left_join(pseudodates) %>% # Now we will replace unmatched dates with their pseudo dates
  replace_na(replace = list(mismatch = 0)) %>% # set mismatch to zero for perfectly aligned dates
  filter(mismatch <= 7) %>% # this gets rid of irreconcilable dates (like the t4 dates mentioned above)
  mutate(date2 = if_else(is.na(pseudo.date), date, pseudo.date)) %>% # replace date with pseudodate for all site-dates lacking perfect alignments
  select(-c(date, pseudo.date, mismatch)) %>%
  rename(date = date2) %>%
  group_by(site, date) %>%
  complete(bb.sp, plant.sp, fill = list(visits = 0)) # generate all possible BB-FL combos for each site date and fill in zeros for visits

# Floral cover data
flower_cover <- read_csv("../data/processed_data/floral_survey.csv") %>%
  filter(flower.cover > 0) %>% # ignore plants that were not blooming
  group_by(site, date, plant.sp) %>%
  summarize(flower.cover = sum(flower.cover))

# Get shared dates after adding pseudodates. This will let us filter out survey dates that don't
# have corresponding visitation sampling dates
shared_dates_with_pseudo <- inner_join(visitation, flower_cover, by = c("site", "date")) %>%
  select(site, date) %>%
  distinct()

# Join visitation and flower cover data

## First we need a list of all BB spp. by site and date
sp_by_site <- visitation %>% 
  ungroup() %>%
  select(site, date, bb.sp) %>%
  distinct()

## Then we need a list of unvisited plant.sp by site and date
unvisited <- flower_cover %>% 
  semi_join(shared_dates_with_pseudo) %>% # filter flower cover data to include only site-dates with visitation complement (real or pseudodates)
  full_join(visitation) %>% # join to visitation data
  filter(is.na(bb.sp)) %>% # select only the site-date-plants that were totally unvisited (and therefore have no BB spp to join to)
  select(-bb.sp) %>% # drop bb.sp variable because I just want the plants
  left_join(sp_by_site) %>% # join in the species list by site and date, thus populating the unvisited plants with all the BB spp recorded for each site-date
  replace_na(replace = list(visits = 0)) # replace NAs (created in the line above) with zero

## Now we can combine everything into a single data frame
visit_cover <- flower_cover %>% 
  semi_join(shared_dates_with_pseudo) %>% # filter flower cover data to include only site-dates with visitation complement (real or pseudodates)
  full_join(visitation) %>% # join to visitation data
  replace_na(replace = list(flower.cover = 0.01)) %>% # If a plant was visited but not surveyed, we will assign it the minimum non-zero flower cover value of 0.01.
  filter(!is.na(bb.sp)) %>% # drop the univisited plants
  bind_rows(unvisited) %>% # add them back, with BB spp and zeros
  select(site, date, bb.sp, plant.sp, visits, flower.cover) %>%
  group_by(site, date, bb.sp, plant.sp, flower.cover) %>% # In a few cases, the use of pseudodates causes two sampling period to converge on the same date, which creates apparent duplicate visitation data. We will handle this by averaging the visits.
  summarize(visits = mean(visits)) %>%
  left_join(fl_traits) %>% # add floral traits
  ungroup() %>%
  mutate(k.type.ss = factor(k.type.ss))

ktype_families <- visit_cover %>%
  select(plant.sp, plant.family, k.type.ss) %>%
  distinct()

ggplot(ktype_families, aes(plant.family, fill = k.type.ss)) +
  geom_bar(position = "fill") +
  theme(axis.text.x = element_text(angle = 90))

## Get total visits by BB sp and site-date; this will be a measure of BB abundance that we can use after we've calculated d'. I could, of course, work this into the biaseval function, but I'll put that off for now.
bb_abundance <- visit_cover %>%
  group_by(site, date, bb.sp) %>%
  summarize(bb.sp.abund = sum(visits)) %>%
  group_by(site, date) %>%
  mutate(bb.sp.prop.abund = bb.sp.abund/sum(bb.sp.abund))

fl_abundance <- visit_cover %>%
  group_by(site, date) %>%
  summarize(total.flower.cover = sum(flower.cover))
```

## A handy function for calculating Blüthgen's d'
```{r}
# A function for calculating Blüthgen's d', parsed in various ways
biaseval <- function(x, criterion, ...) { # criterion = plant grouping variable, e.g. plant.sp, k.type.ss, color

  criterion <- enquo(criterion) # tidy eval

  x %>%

# Select variables of interest; the ... represents the variables we will use to parse our analysis
    select(..., bb.sp, !!criterion, flower.cover, visits) %>% 
    group_by(..., bb.sp, !!criterion) %>%

# If criterion = plant.sp, this next step does nothing. But if criterion is some other grouping variable (e.g. k.type.ss), this sums flower cover and visits over that grouping variable.
    summarize(flower.cover = sum(flower.cover), 
              visits = sum(visits)) %>%

# Nest into list column
    group_by(...) %>%
    nest() %>%

# Pivot each nested data frame into the kind of web that bipartite likes
    mutate(webs = map(data, function(x) select(x, bb.sp, !!criterion, visits) %>%
                        pivot_wider(names_from = !!criterion, values_from = visits, values_fill = 0) %>%
                        column_to_rownames("bb.sp"))) %>%

# Pull out the flower.cover variable to use in the "abuns" argument of dfun below
    mutate(low.abund = map(data, function(x) select(x, !!criterion, flower.cover) %>%
                            unique() %>%
                            column_to_rownames(paste(criterion)[2]))) %>% # Tell me why I need to do this, Hadley.

# Call the dfun command from bipartite to calculate Blüthgen's d'
    mutate(d = map2(webs, low.abund, function(x, y) bipartite::dfun(x, abuns = y$flower.cover) %>%
                      data.frame() %>%
                      rownames_to_column(var = "bb.sp"))) %>%

    unnest(cols = d) %>%
    na.omit() %>% # in a few cases where the sampling is very sparse, d' cannot be calculated; these are safe to drop
    ungroup() %>%
    mutate(site = factor(site),
         bb.sp = factor(bb.sp),
         year = factor(lubridate::year(date)))
}
```

# 1. Quantifying bias
How does overall bias (Blüthgen's d') vary across species and through elev*time? For now, we will focus on bias at the level of plant species. We'll start considering traits later on when our goal is to explain patterns of bias.
## 1.1 Calculate Blüthgen's d' (i.e. visitation bias)
```{r}
# No grouping -- just plant.sp
bias_plantsp <- biaseval(visit_cover, plant.sp, site, date) %>%
  rename(dprime.sp = dprime) %>%
  select(site, date, bb.sp, dprime.sp)

# Grouped by k.type.ss
bias_ktype <- biaseval(visit_cover, k.type.ss, site, date) %>%
  rename(dprime.ktype = dprime) %>%
  select(site, date, bb.sp, dprime.ktype)

# Grouped by plant.family
bias_family <- biaseval(visit_cover, plant.family, site, date) %>%
  rename(dprime.fam = dprime) %>%
  select(site, date, bb.sp, dprime.fam)

# Collated
bias <- full_join(bias_plantsp, bias_ktype) %>%
  full_join(bias_family) %>%
  left_join(site_data) %>% # add site data
  left_join(bb_abundance) %>% # add BB abundance data
  left_join(fl_abundance) %>% # add fl abundance
  filter(bb.sp.abund >= 3) %>% # dprime tends to become binary when n is very low; this threshold removes more than half the data points, but almost all those were ones and zeros. These ones and zeros probably don't mean much, and they make the models hard to fit.
  mutate(bb.sp = factor(bb.sp), 
         yday = yday(date),
         elev = elev.mean/10) %>% #  scale this down for convenience
  mutate(yday0 = scale(yday),  # normalized versions of variables
         elev0 = scale(elev),
         dprime.sp0 = scale(dprime.sp),
         dprime.ktype0 = scale(dprime.ktype),
         dprime.fam0 = scale(dprime.fam)) %>%
  select(site, date, bb.sp, dprime.sp, dprime.sp0, dprime.ktype, 
         dprime.ktype0, dprime.fam, dprime.fam0, elev, elev0, 
         yday, yday0, bb.sp.abund, bb.sp.prop.abund, total.flower.cover)
```

## 1.3 Multi-level model: bias ~ yday * elevation + (1 + elevation | bb.sp)
### 1.3.1 Preliminary visualization
```{r}
# I am interested both in the variation of bias through elevation and time and in the intercept effect of each bumble bee species

ggplot(bias, aes(dprime.sp)) +
  geom_histogram()

# First, let's look at the distribution of bias per BB sp
ggplot(bias, aes(dprime.sp)) +
  geom_histogram() +
  facet_wrap(~bb.sp)

ggplot(bias, aes(dprime.ktype)) +
  geom_histogram() +
  facet_wrap(~bb.sp)

ggplot(bias, aes(dprime.fam)) +
  geom_histogram() +
  facet_wrap(~bb.sp)

# Now, let's try to get a sense for the shape of the response to elevation and time; we might expect to find nonlinearities here
ggplot(bias, aes(elev, dprime.sp)) +
  geom_point() +
  geom_smooth(method = "gam", formula = y ~ s(x, k = 5)) +
  facet_wrap(~bb.sp)

ggplot(bias, aes(yday, dprime.sp)) +
  geom_point() +
  geom_smooth(method = "gam", formula = y ~ s(x, k = 5)) +
  facet_wrap(~bb.sp)

ggplot(bias, aes(elev, dprime.ktype)) +
  geom_point() +
  geom_smooth(method = "gam", formula = y ~ s(x, k = 5)) +
  facet_wrap(~bb.sp)

ggplot(bias, aes(yday, dprime.ktype)) +
  geom_point() +
  geom_smooth(method = "gam", formula = y ~ s(x, k = 5)) +
  facet_wrap(~bb.sp)

ggplot(bias, aes(elev, dprime.fam)) +
  geom_point() +
  geom_smooth(method = "gam", formula = y ~ s(x, k = 5)) +
  facet_wrap(~bb.sp)

ggplot(bias, aes(yday, dprime.fam)) +
  geom_point() +
  geom_smooth(method = "gam", formula = y ~ s(x, k = 5)) +
  facet_wrap(~bb.sp)

# After inspecting these plots, I see no reason to indulge in wiggly models.
```

### 1.3.3 brms models
There's a lot I don't understand about the standardization of variables in the context of Bayesian MLM, but for now I have opted to scale and center the predictor variables but not the response. My reason for scaling and centering the predictors is that is massively speeds up estimation and enables the interaction model to converge (it would not converge with raw variables). My reason for not scaling the response is that it doesn't change the outcome of the model (I've checked) and I am interested in seeing the global intercept for d' on its native scale so that I can compare it with previously reported values. Right now, I am using all default priors. This means flat priors for global effects and student priors for varying effects (I think). My understanding is that brms internally scales priors to match the scale of the data. If I want to define my own priors, which I know I really should, particularly to replace the flat priors on the global effects, then I need to understand how the scaling of the variables affects the specification of the prior.
```{r}
# elev*yday interaction model
brm_dprime.sp00 <- brm(dprime.sp ~ elev0 * yday0 + (1 | bb.sp),
                       #+ (1 | site),
                       cores = 4,
                       iter = 2000,
                       family = "skew_normal", # this accounts for a slight left-skew in the posterior
                       file = "../output/brm_dprime.sp00",
                       data = bias)

summary(brm_dprime.sp00)
prior_summary(brm_dprime.sp00)
parnames(brm_dprime.sp00)
pp_check(brm_dprime.sp00)
bayes_R2(brm_dprime.sp00)

# Intercept-only model
brm_dprime.sp01 <- brm(dprime.sp ~ 1 + (1 | bb.sp) + (1 | site),
                       cores = 4,
                       iter = 2000,
                       family = "skew_normal",
                       file = "../output/brm_dprime.sp01",
                       data = bias)

summary(brm_dprime.sp01)
prior_summary(brm_dprime.sp01)
parnames(brm_dprime.sp01)
pp_check(brm_dprime.sp01)
bayes_R2(brm_dprime.sp01)

# Compare interaction model and intercept-only model. This favors the intercept-only model.
loo(brm_dprime.sp00, brm_dprime.sp01)
```

#### 1.3.3.1 Plot
```{r}
get_variables(brm_dprime.sp01)

brm_dprime.sp01_plotframe <- brm_dprime.sp01 %>%
  spread_draws(b_Intercept, r_bb.sp[condition, term]) %>% 
  mutate(condition_mean = b_Intercept + r_bb.sp)

brm_dprime.sp01_plotframe2 <- brm_dprime.sp01 %>%
  spread_draws(b_Intercept, r_site[condition, term]) %>% 
  mutate(condition_mean = b_Intercept + r_site)

brm_dprime.sp01_subplot1 <- ggplot(brm_dprime.sp01_plotframe, aes(condition_mean, condition)) +
  stat_halfeye() +
  ylab("Bumble bee species") +
  xlab("Discrimination (d')") +
  theme_light()

brm_dprime.sp01_subplot2 <- ggplot(brm_dprime.sp01_plotframe2, aes(condition_mean, condition)) +
  stat_halfeye() +
  ylab("Site") +
  xlab("Discrimination (d')") +
  theme_light()

brm_dprime.sp01_plot1 <- brm_dprime.sp01_subplot1 | brm_dprime.sp01_subplot2

ggsave(brm_dprime.sp01_plot1, file = "../output/brm_dprime.sp01_plot1.png")
```

# 2. Explaining bias
## 2.1 Calculate visitation probability per plant speces/ktype/family
```{r}
pvis_ktype <- visit_cover %>%
  group_by(site, date, bb.sp, plant.sp) %>%
  summarize(flower.cover = sum(flower.cover), # get raw flower cover and visits by site*date*bb.sp*plant.sp
            visits = sum(visits)) %>%
  group_by(site, date) %>%
  mutate(prop.flower.cover = flower.cover/sum(flower.cover), # proportional flower cover
         prop.visits = visits/sum(visits), # proportional visits
         visits.bin = factor(if_else(visits > 0, 1, 0))) %>% # binary visits
  group_by(site, date, bb.sp) %>%
  mutate(bb.sp.abund = sum(visits)) %>% # abundance of each BB sp. per site-date
  ungroup() %>%
  left_join(fl_traits) %>% # add floral traits
  left_join(bb_traits) %>% # add BB traits
  mutate(k.type.ss = factor(k.type.ss), # recode variables
         pbl.w.class2 = factor(pbl.w.class2),
         visits = as.integer(visits)) %>%
  filter(!k.type.ss %in% c("9", "10")) %>% # Drop ktypes 9 and 10 since both are extremely rare
  left_join(site_data)  %>%
  mutate(prop.flower.cover0 = scale(prop.flower.cover), # normalized predictors
         bb.sp.abund0 = scale(bb.sp.abund),
         elev.mean0 = scale(elev.mean),
         yday = yday(date), # and add yday variable
         yday0 = scale(yday)) %>%
  select(site, elev.mean0, elev.mean, elev.class2, date, yday, yday0, bb.sp, plant.sp, visits, 
         visits.bin, prop.visits, prop.flower.cover, prop.flower.cover0,
         k.type.ss, bb.sp.abund, bb.sp.abund0, pbl.w.class2) 
```

## 2.2 Multi-level models

### 2.2.1 Preliminary visualization
The tons of zeros are due to the lots of plant species that were present at very low abundance and received few or zero visits. This might not be ideal. We could opt for one of our alternative bias measure, like fold change, which all give something like a poisson distribution, but I don't know if this would be an improvement. We could also filter our data to include only plants that were above some threshold of relative abundance, say, 1%.
```{r}
# Visit.prop
ggplot(filter(pvis_ktype, prop.flower.cover >= 0.05), aes(prop.visits)) +
  geom_histogram()

# Visits.bin
ggplot(pvis_ktype, aes(k.type.ss, fill = visits.bin)) +
  geom_bar()

ggplot(pvis_ktype, aes(elev.mean, prop.visits)) +
  geom_point() +
  geom_smooth() +
  facet_grid(bb.sp~k.type.ss)

# Visits.bin filtered
ggplot(filter(pvis_ktype, prop.flower.cover >= 0.05), aes(k.type.ss, fill = visits.bin)) +
  geom_bar()
```

### 2.2.2 brms models
```{r}
# Poisson models of raw visit counts
brm_pvis00 <- brm(visits ~ prop.flower.cover0 + 
                    bb.sp.abund0 + 
                    (1 | k.type.ss) + 
                    (1 | site) + 
                    (1 | bb.sp),
                  #control = list(adapt_delta = 0.99),
                  family = "poisson",
                  file = "./output/brm_pvis00b",
                  chains = 4,
                  #cores = 4,
                  iter = 2000,
                  data = pvis_ktype)

summary(brm_pvis00)
prior_summary(brm_pvis00)
parnames(brm_pvis00)
mcmc_plot(brm_pvis00, pars = c("^r_k"))
mcmc_plot(brm_pvis00, pars = c("^r_s"))
mcmc_plot(brm_pvis00, pars = c("^b_"))
mcmc_plot(brm_pvis00)
conditional_effects(brm_pvis00)
pp_check(brm_pvis00)
bayes_R2(brm_pvis00)

# Perhaps I should add the bb.sp varying intercept effect; otherwise, I think I could be accused of pseudo replication.
brm_pvis11 <- brm(visits ~ prop.flower.cover0 + 
                    bb.sp.abund0 + 
                    (1 | k.type.ss:pbl.w.class2) + 
                    (1 | site) +
                    (1 | bb.sp),
                  family = "poisson",
                  file = "../output/brm_pvis11",
                  chains = 4,
                  #cores = 4,
                  iter = 2000,
                  data = pvis_ktype)

summary(brm_pvis11)
prior_summary(brm_pvis11)
parnames(brm_pvis11)
mcmc_plot(brm_pvis11, pars = c("^r_k"))
mcmc_plot(brm_pvis11, pars = c("^r_b"))
mcmc_plot(brm_pvis11, pars = c("^b_"))
mcmc_plot(brm_pvis11)
conditional_effects(brm_pvis11)
pp_check(brm_pvis11)
bayes_R2(brm_pvis11)
```

#### 2.2.2.1 Plot
```{r}
# Note that the intercept is on the log scale, so if you want to transform it back to predicted # visits, you have to exponentiate it. I think.

## Interaction model
brm_pvis11_plotframe <- brm_pvis11 %>%
  spread_draws(b_Intercept, `r_k.type.ss:pbl.w.class2`[condition, term]) %>% # the back ticks are necessary to escape the :
  rename(r_ktype.pbl = `r_k.type.ss:pbl.w.class2`) %>%
  mutate(condition_mean = b_Intercept + r_ktype.pbl) %>%
  separate(condition, c("ktype", "pbl.class"), sep = "_", remove = FALSE) %>%
  mutate(pbl.class = case_when(
    pbl.class == "1" ~ "short",
    pbl.class == "2" ~ "med",
    pbl.class == "3" ~ "long")
    ) %>%
  mutate(ktype = case_when(
    ktype == "0" ~ "NA (0)",
    ktype == "1" ~ "Disc (1)",
    ktype == "2" ~ "Funnel (2)",
    ktype == "3" ~ "Bell (3)",
    ktype == "4" ~ "Stalk-disc (4)",
    ktype == "5" ~ "Lip (5)",
    ktype == "6" ~ "Flag (6)",
    ktype == "7" ~ "Head (7)")
  ) %>%
  mutate(ktype = factor(ktype,
                        levels = c("Disc (1)", "Funnel (2)", "Bell (3)", "Stalk-disc (4)",
                                   "Lip (5)", "Flag (6)", "Head (7)", "NA (0)")))

# ggplot(brm_pvis11_plotframe, aes(condition_mean, pbl.class, color = pbl.class, fill = pbl.class)) +
#   stat_pointinterval() +
#   #stat_halfeye(alpha = 0.5, interval_alpha = 1, point_alpha = 1, size = 1, .width = 0.95) +
#   facet_grid(ktype ~ .) +
#   theme_light() +
#   labs(color = "Tongue length", x = "Visits (log-scale)") +
#   guides(fill = FALSE) +
#   theme(axis.title.y=element_blank(),
#         axis.text.y=element_blank(),
#         axis.ticks.y=element_blank())

ggplot(brm_pvis11_plotframe, aes(condition_mean, pbl.class, color = pbl.class, fill = pbl.class)) +
  #stat_pointinterval() +
  stat_halfeye(alpha = 0.4, interval_alpha = 1, point_alpha = 1) +
  facet_wrap(~ktype, scales = "free") +
  theme_light() +
  labs(color = "Tongue length", x = "Visits (log-scale)") +
  guides(fill = FALSE) +
  theme(axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank())

ggsave(file = "../output/brm_pvis11_plot1.png", width = 7, height = 5.5)

# Just ktype model
brm_pvis00_plotframe <- brm_pvis00 %>%
  spread_draws(b_Intercept, r_k.type.ss[condition, term]) %>% # the back ticks are necessary to escape the :
  mutate(condition_mean = b_Intercept + r_k.type.ss) %>%
  mutate(ktype = case_when(
    condition == "0" ~ "NA",
    condition == "1" ~ "Disc",
    condition == "2" ~ "Funnel",
    condition == "3" ~ "Bell",
    condition == "4" ~ "Stalk-disc",
    condition == "5" ~ "Lip",
    condition == "6" ~ "Flag",
    condition == "7" ~ "Head")
  ) %>%
  mutate(ktype = factor(ktype,
                        levels = c("Disc", "Funnel", "Bell", "Stalk-disc",
                                   "Lip", "Flag", "Head", "NA")))

ggplot(brm_pvis00_plotframe, aes(condition_mean, ktype)) +
  stat_halfeye() +
  theme_light() +
  guides(fill = FALSE) +
  xlab("Visits (log-scale)") +
  ylab("Morphotype")

ggsave(file = "../output/brm_pvis00_plot1.png", width = 7, height = 5.5)
```

### 2.2.3 The lability of bias: How does the foraging bias of each species with respect to each morphotype vary through elevation and time?
```{r}
ggplot(pvis_ktype, aes(elev.mean, prop.visits)) +
  geom_point() +
  geom_smooth(method = "gam", formula = y ~ s(x, k = 5)) +
  facet_grid(bb.sp ~ k.type.ss)

ggplot(pvis_ktype, aes(elev.mean, prop.visits)) +
  geom_point() +
  geom_smooth(method = "gam", formula = y ~ s(x, k = 5)) +
  facet_grid(bb.sp ~ k.type.ss)

ggplot(pvis_ktype, aes(yday0, prop.visits)) +
  geom_point() +
  geom_smooth(method = "gam", formula = y ~ s(x, k = 5)) +
  facet_grid(bb.sp ~ k.type.ss)

brm_pviselev00 <- brm(visits ~ prop.flower.cover0 + 
                        bb.sp.abund0 + 
                        (1 + elev.mean0:yday0 | k.type.ss:bb.sp),
                      family = "poisson",
                      file = "../output/brm_pviselev00",
                      chains = 4,
                      cores = 4,
                      iter = 2000,
                      data = pvis_ktype)

summary(brm_pviselev00)
prior_summary(brm_pviselev00)
parnames(brm_pviselev00)
mcmc_plot(brm_pviselev00, pars = c("^r_k"))
mcmc_plot(brm_pviselev00, pars = c("^r_s"))
mcmc_plot(brm_pviselev00, pars = c("^b_"))
mcmc_plot(brm_pviselev00)
conditional_effects(brm_pviselev00)
pp_check(brm_pviselev00)
bayes_R2(brm_pviselev00)
```

# 3. Effects of bias

## 3.1 Relationship between bias(d') and niche overlap
### 3.1.1 A function for calculating niche overlap
Niche overlap (Horn's distance) is first calculated for each species pair and each site-date. Naturally, site-dates in which only one BB species was observed return empty niche overlap matrices and so are excluded from further analysis.  
```{r}
# Niche overlap by sp.pair * site * year
niche_overlapper <- function(x, criterion) {

  criterion <- enquo(criterion) # criterion is set to either plant.sp (for taxonomic) or k.type.s (for morphological)

  no1 <- x %>%
    group_by(bb.sp, !!criterion, site, date) %>% # group by site-year and BB-FL pair
    summarize(visits = sum(visits)) %>%
    group_by(bb.sp, site, date) %>% # group by site-year and BB species
    mutate(prop.visits = visits/sum(visits)) %>% # get proportional interaction frequency for each BB-FL pair
    select(site, date, bb.sp, !!criterion, prop.visits) %>%
    pivot_wider(names_from = !!criterion, values_from = prop.visits, values_fill = 0) %>% # spread into the kind of array vegan likes
    group_by(site, date) %>%
    nest() %>% # nest into list-column
    mutate(data2 = map(data, function(x) column_to_rownames(x, var = "bb.sp")), # column to rownames
           richness = map(data, function(x) length(x$bb.sp)),
           dist = map(data2, function(x) vegdist(x, method = "horn")), # convert to Horn's distance matrix
           tidy_dist = map(dist, tidy)) %>% # convert to tidy array of Horn's distances
    #dplyr::select(site, date, tidy_dist) %>% # drop intermediate list-columns
    unnest(cols = c(tidy_dist)) %>%
    rename(bb.sp1 = item1, bb.sp2 = item2) %>%
    mutate(niche.overlap = 1-distance,  # express niche overlap so that 1 = perfect overlap and 0 = no overlap
           telu = if_else(bb.sp1 == "telu" | bb.sp2 == "telu", TRUE, FALSE),
           pasc = if_else(bb.sp1 == "pasc" | bb.sp2 == "pasc", TRUE, FALSE),
           wurf = if_else(bb.sp1 == "wurf" | bb.sp2 == "wurf", TRUE, FALSE),
           prat = if_else(bb.sp1 == "prat" | bb.sp2 == "prat", TRUE, FALSE),
           soro = if_else(bb.sp1 == "soro" | bb.sp2 == "soro", TRUE, FALSE),
           humi = if_else(bb.sp1 == "humi" | bb.sp2 == "humi", TRUE, FALSE),
           muci = if_else(bb.sp1 == "muci" | bb.sp2 == "muci", TRUE, FALSE),
           pyre = if_else(bb.sp1 == "pyre" | bb.sp2 == "pyre", TRUE, FALSE),
           mont = if_else(bb.sp1 == "mont" | bb.sp2 == "mont", TRUE, FALSE),
           mend = if_else(bb.sp1 == "mend" | bb.sp2 == "mend", TRUE, FALSE),
           psit = if_else(bb.sp1 == "psit" | bb.sp2 == "psit", TRUE, FALSE),
           gers = if_else(bb.sp1 == "gers" | bb.sp2 == "gers", TRUE, FALSE),
           jone = if_else(bb.sp1 == "jone" | bb.sp2 == "jone", TRUE, FALSE),
           hypn = if_else(bb.sp1 == "hypn" | bb.sp2 == "hypn", TRUE, FALSE),
           lapi = if_else(bb.sp1 == "lapi" | bb.sp2 == "lapi", TRUE, FALSE),
           hort = if_else(bb.sp1 == "hort" | bb.sp2 == "hort", TRUE, FALSE))
}
```

### 3.1.2 My first question is how niche overlap compares across species pairs
```{r}
# Unaggregated niche overlap per species pair; not that this drops a lot of data because of the inner join with bias
niche_overlap <- read_csv("../data/processed_data/network.csv") %>%
  filter(caste != "male") %>% # we will ignore males 
  group_by(site, date, bb.sp, plant.sp) %>%
  summarize(visits = n()) %>%
  niche_overlapper(plant.sp) %>%
  mutate(yday = yday(date)) %>%
  left_join(bb_traits, by = c("bb.sp1" = "bb.sp")) %>%
  rename(pbl.w.class2_sp1 = pbl.w.class2) %>%
  left_join(bb_traits, by = c("bb.sp2" = "bb.sp")) %>%
  rename(pbl.w.class2_sp2 = pbl.w.class2) %>%
  mutate(same.class = case_when(
    pbl.w.class2_sp1 == "1" & pbl.w.class2_sp2 == "1" ~ "short:short",
    pbl.w.class2_sp1 == "2" & pbl.w.class2_sp2 == "2" ~ "med:med",
    pbl.w.class2_sp1 == "3" & pbl.w.class2_sp2 == "3" ~ "long:long",
    pbl.w.class2_sp1 == "1" & pbl.w.class2_sp2 == "2" ~ "short:med",
    pbl.w.class2_sp1 == "2" & pbl.w.class2_sp2 == "1" ~ "short:med",
    pbl.w.class2_sp1 == "1" & pbl.w.class2_sp2 == "3" ~ "short:long",
    pbl.w.class2_sp1 == "3" & pbl.w.class2_sp2 == "1" ~ "short:long",
    pbl.w.class2_sp1 == "2" & pbl.w.class2_sp2 == "3" ~ "med:long",
    pbl.w.class2_sp1 == "3" & pbl.w.class2_sp2 == "2" ~ "med:long")
    ) %>%
  left_join(bb_abundance, by = c("site", "date", "bb.sp1" = "bb.sp")) %>%
  rename(bb.sp1.abund = bb.sp.abund) %>%
  left_join(bb_abundance, by = c("site", "date", "bb.sp2" = "bb.sp")) %>%
  rename(bb.sp2.abund = bb.sp.abund) %>%
  unite(sp.pair, c(bb.sp1, bb.sp2), sep = "_", remove = FALSE) %>%
  left_join(site_data) %>%
  mutate(yday = yday(date)) %>%
  select(site, date, yday, elev.mean, sp.pair, bb.sp1, bb.sp2, bb.sp1.abund, 
         bb.sp2.abund, niche.overlap, same.class)

# Niche overlap becomes rather binary when one or both species is rare. This is driven especially by singletons.
ggplot(niche_overlap, aes(sp.pair, niche.overlap)) +
  geom_violin() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

# Let's see if we can improve matters by dropping singletons
ggplot(filter(niche_overlap, bb.sp1.abund >= 2 & bb.sp1.abund >= 2), 
       aes(sp.pair, niche.overlap)) +
  geom_violin() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

# Now a threshold of 3
ggplot(filter(niche_overlap, bb.sp1.abund >= 3 & bb.sp1.abund >= 3), 
       aes(sp.pair, niche.overlap)) +
  geom_violin() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

ggplot(filter(niche_overlap, bb.sp1.abund >= 3 & bb.sp1.abund >= 3), 
       aes(same.class, niche.overlap)) +
  geom_violin() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

# Now a threshold of 5...bye, all the species pairs...
ggplot(filter(niche_overlap, bb.sp1.abund >= 5 & bb.sp1.abund >= 5), 
       aes(sp.pair, niche.overlap)) +
  geom_violin() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

ggplot(filter(niche_overlap, bb.sp1.abund >= 10 & bb.sp1.abund >= 10), 
       aes(sp.pair, niche.overlap)) +
  geom_violin() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

# Let's see if there is any influence of site, elevation, yday
ggplot(filter(niche_overlap, bb.sp1.abund >= 3 & bb.sp1.abund >= 3), 
       aes(site, niche.overlap)) +
  geom_boxplot() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

ggplot(filter(niche_overlap, bb.sp1.abund >= 3 & bb.sp1.abund >= 3), 
       aes(elev.mean, niche.overlap)) +
  geom_point() +
  geom_smooth() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

ggplot(filter(niche_overlap, bb.sp1.abund >= 3 & bb.sp1.abund >= 3), 
       aes(yday, niche.overlap)) +
  geom_point() +
  geom_smooth() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

# Abundance relatonships
ggplot(niche_overlap, aes(log(bb.sp1.abund), log(bb.sp2.abund))) +
  geom_point() +
  geom_smooth(se = FALSE) +
  facet_wrap(~sp.pair)

#
ggplot(niche_overlap, aes(same.class, niche.overlap)) +
  geom_boxplot()

# Now for what is probably a pretty naive model

brm_niche00 <- brm(niche.overlap ~ bb.sp1.abund*bb.sp2.abund + 
                     (1 | sp.pair) + 
                     (1 | same.class) + 
                     (1 | site),
                   chains = 4,
                   cores = 4,
                   iter = 2000,
                   file = "../output/brm_niche00",
                   data = filter(niche_overlap, bb.sp1.abund >= 1 & bb.sp1.abund >= 1))

parnames(brm_niche02)
conditional_effects(brm_niche02)
mcmc_plot(brm_niche02, pars = c("^r_sp"))
mcmc_plot(brm_niche02, pars = c("^r_sa"))
mcmc_plot(brm_niche02, pars = c("^b_"))
hypothesis(brm_niche02, "dprime = 0")
pp_check(brm_niche02)
bayes_R2(brm_niche02)

```

### 3.1.3 Now I'd be interested to know how niche overlap relates to dprime
```{r}
niche_overlap_bias <- niche_overlap %>%
  inner_join(bias, by = c("site", "date", c("bb.sp1" = "bb.sp"))) %>%
  rename(dprime.sp1 = dprime.sp) %>%
  inner_join(bias, by = c("site", "date", c("bb.sp2" = "bb.sp"))) %>%
  rename(dprime.sp2 = dprime.sp) %>%
  select(site, date, bb.sp1, bb.sp2, bb.sp1.abund, bb.sp2.abund, 
         dprime.sp1, dprime.sp2, niche.overlap, same.class)

# When bias is high, it could mean that there are few good options on which all (or many) BB species converge. But apparently it doesn't. When bias is high, BB spp are biased toward DIFFERENT flora, and thus their niches diverge. 
ggplot(niche_overlap_bias, aes(dprime.sp, mean.niche.overlap)) +
  geom_point(alpha = 0.1) +
  geom_smooth(method = "lm") +
  #geom_smooth(method = "gam", formula = y ~ s(x, k = 5), se = TRUE) +
  facet_wrap(~bb.sp) +
  ylim(c(0,1))

brm.bias_niche <- brm(mean.niche.overlap ~ dprime.sp + (1 + dprime.sp | bb.sp) + (1 + dprime.sp | site),
                     cores = 4,
                     chains = 4,
                     iter = 2000,
                     file = "../output/brm_bias_nicheoverlap",
                     data = bias_nicheoverlap)

parnames(brm.bias_niche)
conditional_effects(brm.bias_niche)
mcmc_plot(brm.bias_niche, pars = c("^r_"))
hypothesis(brm.bias_niche, "dprime = 0")
pp_check(brm.bias_niche)
bayes_R2(brm.bias_niche)

brm.dprime_niche00 <- brm(mean.niche.overlap ~ dprime.sp + (1 | bb.sp) + (1 | site),
                     cores = 4,
                     chains = 4,
                     iter = 2000,
                     file = "../output/brm_bias_niche00",
                     data = bias_nicheoverlap)

parnames(brm.dprime_niche00)
conditional_effects(brm.dprime_niche00)
mcmc_plot(brm.dprime_niche00, pars = c("^r_bb"))
mcmc_plot(brm.dprime_niche00, pars = c("^b_"))
hypothesis(brm.dprime_niche00, "dprime = 0")
pp_check(brm.dprime_niche00)
bayes_R2(brm.dprime_niche00)
```

## 3.2 Ordination of dietary niche
```{r}
# Aggregated by mean proportional visitation

# We will use the whole dataset this time, since there is no need to align visitation to floral surveying
visitation1 <- read_csv("../data/processed_data/network.csv") %>%
  filter(caste != "male") %>% # we will ignore males 
  group_by(site, date, bb.sp, plant.sp) %>%
  summarize(visits = n()) %>%
  group_by(site, date, bb.sp) %>%
  mutate(prop.visits = visits/sum(visits)) %>%
  group_by(bb.sp, plant.sp) %>%
  summarize(mean.prop.visits = mean(prop.visits)) %>%
  pivot_wider(names_from = plant.sp, values_from = mean.prop.visits, values_fill = 0) %>%
  column_to_rownames("bb.sp")

visitation2 <- read_csv("../data/processed_data/network.csv") %>%
  filter(caste != "male") %>% # we will ignore males 
  group_by(site, date, bb.sp, plant.sp) %>%
  summarize(visits = n()) %>%
  group_by(site, date, bb.sp) %>%
  mutate(prop.visits = visits/sum(visits)) %>%
  unite(id, c(site, date, bb.sp)) %>%
  select(-visits) %>%
  pivot_wider(names_from = plant.sp, values_from = prop.visits, values_fill = 0) %>%
  column_to_rownames("id")

# Aggregated
pcoa_horn1 <- capscale(visitation1 ~ 1, distance = "horn")
pcoa_horn1.fort1 <- fortify(pcoa_horn1, axes = 1:2) %>%
  filter(Score == "species") %>%
  rename(plant.sp = Label) %>%
  left_join(fl_traits)
pcoa_horn1.fort2 <- fortify(pcoa_horn1, axes = 1:2) %>%
  filter(Score == "sites") %>%
  rename(bb.sp = Label) %>%
  left_join(bb_traits)

ggplot(pcoa_horn1.fort2, aes(MDS1, MDS2)) +
  geom_text(aes(label = bb.sp)) +
  coord_fixed()

# Disaggregated
pcoa_horn <- capscale(visitation2 ~ 1, distance = "horn")
pcoa_horn.fort <- fortify(pcoa_horn, axes = 1:2)
pcoa_horn.fort1 <- pcoa_horn.fort %>%
  filter(Score == "species") %>%
  rename(plant.sp = Label) %>%
  left_join(fl_traits)
pcoa_horn.fort2 <- pcoa_horn.fort %>%
  filter(Score == "sites") %>%
  separate(Label, c("site", "date", "bb.sp"), sep = "_") %>%
  left_join(bb_traits)

ggplot(pcoa_horn.fort1, aes(MDS1, MDS2, color = factor(k.type.ss))) +
  geom_point(alpha = 0.25) +
  coord_fixed()

ggplot(pcoa_horn.fort2, aes(MDS1, MDS2, color = factor(pbl.w.class2))) +
  geom_point(alpha = 0.5) +
  coord_fixed()

ggplot(pcoa_horn.fort2, aes(MDS1, MDS2, color = factor(pbl.w.class2))) +
  geom_density2d() +
  coord_fixed()

visitation3 <- read_csv("../data/processed_data/network.csv") %>%
  filter(caste != "male") %>% # we will ignore males 
  group_by(bb.sp, plant.sp) %>%
  summarize(visits = n()) %>%
  group_by(bb.sp) %>%
  mutate(prop.visits = visits/sum(visits))

visitation_aggregate2 <- visitation2 %>%
  group_by(bb.sp, plant.sp) %>%
  summarize(mean.prop.visits = mean(prop.visits)) %>%
  pivot_wider(names_from = plant.sp, values_from = mean.prop.visits, values_fill = 0) %>%
  column_to_rownames("bb.sp")

visitation_aggregate3 <- visitation3 %>%
  pivot_wider(names_from = plant.sp, values_from = prop.visits, values_fill = 0) %>%
  column_to_rownames("bb.sp")

bb_traits2 <- bb_traits %>%
  select(bb.sp, pbl.w.class2) %>%
  arrange(bb.sp)

pcoa_horn <- capscale(visitation_aggregate ~ 1, distance = "horn")
envfit_horn <- envfit(pcoa_horn ~ pbl.w.class2, data = bb_traits2)
plot(pcoa_horn, type = "t")
plot(envfit_horn, add = TRUE)
pcoa_horn.fort <- fortify(pcoa_horn, axes = 1:2)

# What if I did not aggregate the data?
visit_cover_prop.plantsp2 <- visit_cover %>%
  filter(bb.sp %in% c("pasc", "prat", "telu", "soro", "wurf", "psit")) %>%
  select(site, date, bb.sp, plant.sp, k.type.ss, visits, flower.cover) %>%
  group_by(site, date, bb.sp) %>%
  mutate(prop.visits = visits/sum(visits)) %>%
  group_by(site, date, plant.sp) %>%
  mutate(prop.flower.cover = flower.cover/sum(flower.cover)) %>%
  unite(id, c(site, date, bb.sp), remove = TRUE) %>%
  select(id, plant.sp, prop.visits) %>%
  pivot_wider(names_from = plant.sp, values_from = prop.visits, values_fill = 0) %>%
  column_to_rownames("id")

traits <- bb_traits %>%
  filter(bb.sp %in% c("pasc", "prat", "telu", "soro", "wurf", "psit")) %>%
  select(bb.sp, pbl.w.class2) %>%
  arrange(bb.sp)

pcoa_horn2 <- capscale(visit_cover_prop.plantsp2 ~ 1, distance = "horn")
envfit_horn2 <- envfit(pcoa_horn2 ~ pbl.w.class2, data = traits)
plot(pcoa_horn2, type = "t")
plot(envfit_horn2, add = TRUE)

pcoa_horn2_fort <- fortify(pcoa_horn2)

pcoa_horn2_plot <- pcoa_horn2_fort %>%
  filter(Score == "sites") %>%
  separate(Label, c("site", "date", "bb.sp"), sep = "_") %>%
  mutate(date = ymd(date)) %>%
  mutate(yday = yday(date))

ggplot(pcoa_horn2_plot, aes(MDS1, MDS2, color = bb.sp)) +
  geom_point() +
  coord_fixed()

ggplot(pcoa_horn2_plot, aes(MDS1, MDS2, color = yday)) +
  geom_point() +
  coord_fixed() 
```

# Trash can

```{r}
niche_overlapper <- function(x, criterion) {

  criterion <- enquo(criterion) # criterion is set to either plant.sp (for taxonomic) or k.type.s (for morphological)

  no1 <- x %>%
    group_by(bb.sp, !!criterion, site, date) %>% # group by site-year and BB-FL pair
    summarize(visits = sum(visits)) %>%
    group_by(bb.sp, site, date) %>% # group by site-year and BB species
    mutate(prop.visits = visits/sum(visits)) %>% # get proportional interaction frequency for each BB-FL pair
    select(site, date, bb.sp, !!criterion, prop.visits) %>%
    pivot_wider(names_from = !!criterion, values_from = prop.visits, values_fill = 0) %>% # spread into the kind of array vegan likes
    group_by(site, date) %>%
    nest() %>% # nest into list-column
    mutate(data2 = map(data, function(x) column_to_rownames(x, var = "bb.sp")), # column to rownames
           richness = map(data, function(x) length(x$bb.sp)),
           dist = map(data2, function(x) vegdist(x, method = "horn")), # convert to Horn's distance matrix
           tidy_dist = map(dist, tidy)) %>% # convert to tidy array of Horn's distances
    #dplyr::select(site, date, tidy_dist) %>% # drop intermediate list-columns
    unnest(cols = c(tidy_dist)) %>%
    rename(bb.sp1 = item1, bb.sp2 = item2) %>%
    mutate(niche.overlap = 1-distance,  # express niche overlap so that 1 = perfect overlap and 0 = no overlap
           telu = if_else(bb.sp1 == "telu" | bb.sp2 == "telu", TRUE, FALSE),
           pasc = if_else(bb.sp1 == "pasc" | bb.sp2 == "pasc", TRUE, FALSE),
           wurf = if_else(bb.sp1 == "wurf" | bb.sp2 == "wurf", TRUE, FALSE),
           prat = if_else(bb.sp1 == "prat" | bb.sp2 == "prat", TRUE, FALSE),
           soro = if_else(bb.sp1 == "soro" | bb.sp2 == "soro", TRUE, FALSE),
           humi = if_else(bb.sp1 == "humi" | bb.sp2 == "humi", TRUE, FALSE),
           muci = if_else(bb.sp1 == "muci" | bb.sp2 == "muci", TRUE, FALSE),
           pyre = if_else(bb.sp1 == "pyre" | bb.sp2 == "pyre", TRUE, FALSE),
           mont = if_else(bb.sp1 == "mont" | bb.sp2 == "mont", TRUE, FALSE),
           mend = if_else(bb.sp1 == "mend" | bb.sp2 == "mend", TRUE, FALSE),
           psit = if_else(bb.sp1 == "psit" | bb.sp2 == "psit", TRUE, FALSE),
           gers = if_else(bb.sp1 == "gers" | bb.sp2 == "gers", TRUE, FALSE),
           jone = if_else(bb.sp1 == "jone" | bb.sp2 == "jone", TRUE, FALSE),
           hypn = if_else(bb.sp1 == "hypn" | bb.sp2 == "hypn", TRUE, FALSE),
           lapi = if_else(bb.sp1 == "lapi" | bb.sp2 == "lapi", TRUE, FALSE),
           hort = if_else(bb.sp1 == "hort" | bb.sp2 == "hort", TRUE, FALSE))

  no <- list(

  no_telu <- no1 %>%
    group_by(site, date, telu) %>% 
    summarize(mean.niche.overlap = mean(niche.overlap),
              max.niche.overlap = max(niche.overlap)) %>%
    mutate(bb.sp = rep("telu", n())) %>%
    filter(telu == TRUE),

  no_pasc <- no1 %>%
    group_by(site, date, pasc) %>%
    summarize(mean.niche.overlap = mean(niche.overlap),
              max.niche.overlap = max(niche.overlap)) %>%    
    mutate(bb.sp = rep("pasc", n())) %>%
    filter(pasc == TRUE),

  no_wurf <- no1 %>%
    group_by(site, date, wurf) %>%
    summarize(mean.niche.overlap = mean(niche.overlap),
              max.niche.overlap = max(niche.overlap)) %>%    
    mutate(bb.sp = rep("wurf", n())) %>%
    filter(wurf == TRUE),

  no_prat <- no1 %>%
    group_by(site, date, prat) %>%
    summarize(mean.niche.overlap = mean(niche.overlap),
              max.niche.overlap = max(niche.overlap)) %>%   
    mutate(bb.sp = rep("prat", n())) %>%
    filter(prat == TRUE),

  no_soro <- no1 %>%
    group_by(site, date, soro) %>% 
    summarize(mean.niche.overlap = mean(niche.overlap),
              max.niche.overlap = max(niche.overlap)) %>%    
    mutate(bb.sp = rep("soro", n())) %>%
    filter(soro == TRUE),

  no_humi <- no1 %>%
    group_by(site, date, humi) %>%
    summarize(mean.niche.overlap = mean(niche.overlap),
              max.niche.overlap = max(niche.overlap)) %>%    
    mutate(bb.sp = rep("humi", n())) %>%
    filter(humi == TRUE),

  no_muci <- no1 %>%
    group_by(site, date, muci) %>%
    summarize(mean.niche.overlap = mean(niche.overlap),
              max.niche.overlap = max(niche.overlap)) %>%    
    mutate(bb.sp = rep("muci", n())) %>%
    filter(muci == TRUE),

  no_pyre <- no1 %>%
    group_by(site, date, pyre) %>%
    summarize(mean.niche.overlap = mean(niche.overlap),
              max.niche.overlap = max(niche.overlap)) %>%   
    mutate(bb.sp = rep("pyre", n())) %>%
    filter(pyre == TRUE),

  no_mont <- no1 %>%
    group_by(site, date, mont) %>% 
    summarize(mean.niche.overlap = mean(niche.overlap),
              max.niche.overlap = max(niche.overlap)) %>%    
    mutate(bb.sp = rep("mont", n())) %>%
    filter(mont == TRUE),

  no_mend <- no1 %>%
    group_by(site, date, mend) %>%
    summarize(mean.niche.overlap = mean(niche.overlap),
              max.niche.overlap = max(niche.overlap)) %>%    
    mutate(bb.sp = rep("mend", n())) %>%
    filter(mend == TRUE),

  no_psit <- no1 %>%
    group_by(site, date, psit) %>%
    summarize(mean.niche.overlap = mean(niche.overlap),
              max.niche.overlap = max(niche.overlap)) %>%    
    mutate(bb.sp = rep("psit", n())) %>%
    filter(psit == TRUE),

  no_gers <- no1 %>%
    group_by(site, date, gers) %>%
    summarize(mean.niche.overlap = mean(niche.overlap),
              max.niche.overlap = max(niche.overlap)) %>%    
    mutate(bb.sp = rep("gers", n())) %>%
    filter(gers == TRUE),

  no_jone <- no1 %>%
    group_by(site, date, jone) %>% 
    summarize(mean.niche.overlap = mean(niche.overlap),
              max.niche.overlap = max(niche.overlap)) %>%    
    mutate(bb.sp = rep("jone", n())) %>%
    filter(jone == TRUE),

  no_hypn <- no1 %>%
    group_by(site, date, hypn) %>%
    summarize(mean.niche.overlap = mean(niche.overlap),
              max.niche.overlap = max(niche.overlap)) %>%    
    mutate(bb.sp = rep("hypn", n())) %>%
    filter(hypn == TRUE),

  no_lapi <- no1 %>%
    group_by(site, date, lapi) %>%
    summarize(mean.niche.overlap = mean(niche.overlap),
              max.niche.overlap = max(niche.overlap)) %>%    
    mutate(bb.sp = rep("lapi", n())) %>%
    filter(lapi == TRUE),

  no_hort <- no1 %>%
    group_by(site, date, hort) %>%
    summarize(mean.niche.overlap = mean(niche.overlap),
              max.niche.overlap = max(niche.overlap)) %>%    
    mutate(bb.sp = rep("hort", n())) %>%
    filter(hort == TRUE)
  ) %>%

    bind_rows() %>%
    select(site, date, bb.sp, mean.niche.overlap, max.niche.overlap)
}
```

## 2.1 Defining bias

Deciding exactly how to represent bias turns out to be rather tricky. I am trying out a few options right now.

1. Bias could  be treated simply as the Euclidean difference between prop.flower.cover and prop.visits. This has the effect of weighting bias by flower.cover, particularly negative bias, since an abundant plant can have a greater negative bias (because it is farther from 0) than a rare plant. Euclidean bias is bounded between -1 and 1. Its symmetry is a nice property that sets it apart from proportional or fold-change bias. If, for whatever reason, I decided that a log-transformation of this measure were desirable, I think I could simply add a constant of 1 to all values, which would move the boundary up to (0, 2). I think the biggest problem with measuring bias in this way is that it makes no distinction between a plant that was visited a little and one that was visited not at all, and that could be an important biological difference. For example, consider two plants, one with a proportional abundance = 0.01 and proportional visitation = 0.01, and the second with proportional abundance = 0.01 and proportional visitation = 0.00. These will look identical in our analysis, but it is possible that the unvisited plant would not have been visited no matter how high its abundance was. **For starters, though, I am going to proceed with this metric because it is the simplest. And because DJ thought it was a good idea.**

2. Alternatively, bias could be calculated as a simple proportion: prop.visits/prop.flower.cover. If anything, this might tend to reverse the weighting, since a small Euclidean change in a rare species would be a large proportional change. Proportional change has a lower boundary of 0 and no upper boundary. This approach highlights cases of positive bias, but it is even worse than euclidean bias in terms of obscuring the distinction between plants visited slightly and those visited not at all. Worse in a sense. In another sense, it is better than euclidean bias, because with proportional visitation a plant visited 0 times is at least numerical distinguishable from one visited 1 time (visited plants will always have non-zero proportional bias).

3. Similarly to proportional bias, I could calculate bias as fold change. This converges to be virtually identical to proportional change as bias becomes more positive, but fold change has a lower bound of -1 instead of 0.

4. If the distinction between plants that were visited and those that were not is what we're most interested in, we could simply treat bias as binary. This, of course, comes at the price of not being able to distinguish a plant that received 0.01 proportional visitation and one that recevied  0.99 proportional visitation, and this distinction is arguably important.

5. An interesting option is to first separate visited from unvisited plants, set the bias of unvisited plants to 0, and then exponentiate the euclidean bias of all visited plant species. This has some desirable properties. First, it clearly distinguished visited plants from unvisited ones. Second, it assigns plants visited perfectly in proportion to their abundance a sensible bias of 1 (just like proportional bias). Thus, plants with a psotive bias are > 1 (bounded at 2.718282 -- okay, so that's a bit inelegant), those with a negative bias are < 1 (bounded at 0.367895), and those not visited at all = 0.

6. The problem with the option above is that is adds weight to positive bias. Another approach is to separate the unvisited plants from the visited, set unvisited bias = 0, and visited bias = prop.visits - prop.flower.cover + 1. The +1 serves to make the distribution positive, with 1 represening perfect proportionality between visitation and floral abundance. This creates a symmetrically bounded distribution between 0 (not inclusive) and 2 (inclusive), plus the true zeros from the unvisited plants. It feels a bit...slapped together, though.

*We will adopt the term representation to avoid confusion with the object defined above.*

## 2.2 Function for calculating bias per group (= criterion)
```{r}
biaseval_group <- function(x, criterion) {

  criterion <- enquo(criterion)

  x %>%
    mutate(criterion = factor(!!criterion)) %>% # having this as a factor will help with modeling later on
    group_by(site, date, bb.sp, criterion) %>%
    summarize(flower.cover = sum(flower.cover), # get raw floer cover and visits by group
              visits = sum(visits)) %>%
    group_by(site, date, bb.sp) %>%
    mutate(prop.flower.cover = flower.cover/sum(flower.cover), # get proportional flower cover and visits by group
           prop.visits = visits/sum(visits)) %>%
    mutate(bias = prop.visits - prop.flower.cover, # bias as difference
           bias.prop = prop.visits/prop.flower.cover, # bias as proportion
           bias.fold = bias/prop.flower.cover, # bias as fold change
           bias.bin = if_else(bias > 0, 1, 0), # bias as binary
           bias.exp = if_else(visits == 0, 0, exp(prop.visits - prop.flower.cover)),  # bias as exponentiated difference
           bias.split = if_else(visits == 0, 0, (prop.visits - prop.flower.cover + 1))) %>%
    mutate(yday = yday(date)) %>%
    select(site, date, yday, bb.sp, criterion, 
           bias, bias.prop, bias.fold, bias.bin, bias.exp, bias.split,
           visits, flower.cover, prop.visits, prop.flower.cover)
}
```
```