### 2021-01-22 writing/manuscript_v4.Rmd

---
title: "Bumble bee foraging bias"
author:
- Douglas B. Sponsler
- Fabrice Requier
- Katharina Kallnik
- Alice Claßen
- A. Fabienne Maihoff
- Ingolf Steffan-Dewenter
bibliography: refs.bib
output:
  pdf_document:
    fig_caption: yes
---

# 1. Introduction

Plant-pollinator networks are characterized by unequal interaction strengths among partners. While the distribution of interaction strengths can be partially explained by the neutral effects of abundance, pollinators exhibit positive and negative biases toward available flora even when interaction strength is conditioned on floral abundance. Taken as a focal phenomenon, the floral biases of pollinators can be interposed between their underlying causal mechanisms and their higher-order significance **(Figure 1)**. From the causal perspective, floral biases can be partitioned into innate and learned components. Both these components depend on interactions between the functional traits of a given pollinator and those of its potential floral hosts, manifest over ecological time scales with respect to learned biases and evolutionary time scales with respect to innate biases. From the vantage of higher-order effects, the floral biases of pollinators contribute to the complex structures observed in plant-pollinator networks, which have consequences for both the ecological function of the network and the fitness of its constituent species. These effects can, in turn, feedback on the causes of bias via their modulation of community composition and biotic interactions.

Bumble bees (Hymenoptera: *Bombus* spp.) are an especially interesting group in which to study floral biases together with their causal mechanisms and downstream effects. On the one hand, bumble bees are exceptionally versatile foragers, and their ability to solve the puzzles posed by complex flower morphologies [@Plowright1984-lc], or indeed by human investigators [@Alem2016-rn; @Loukola2017-bt], renders almost any flower potentially accessible. Even when a gross mismatch between tongue length and corolla depth makes the nectar of a given flower inaccessible by ordinary means, some bumble bee species will resort readily to biting through the base of the flower to “rob” nectar, incidentally making the nectar available to other prospective foragers as well. Flower handling skills are, moreover, socially transmitted, amplifying the plasticity of individual behavior [@Leadbeater2008-rh; @Loukola2017-bt]. Nevertheless, the flexibility of bumble bee foraging is tuned to floral reward, which depends on both the intrinsic reward of a given flower and the searching/learning/handling costs associated with exploiting it. At an evolutionary level, the innate preferences of bumble bees for certain floral traits likely reflect historic associations between floral traits and floral rewards [@Raine2007-nq]. These innate preferences, however, can be modulated or even nullified by patterns of reward learned during foraging [@Kugler1943-uy; Maharaj2019-wm], which depend not only on intrinsic floral properties but also ecological modulators, such as the spatial distribution of floral patches or competition with other bumble bees [@Ishii2013-sm; @Plowright1984-lc; @Sowig1989-ld]. Thus, realized biases in bumble bee foraging are the labile product of bumble bee traits, floral traits, and ecological context. The influence of these biases on interaction network structure has consequences both for plant pollination and bumble bee species coexistence, and these consequences potentially feed back upon the causes of bumble bee foraging bias via the modulation of competition between bumble bees and ultimately via the regulation of community assembly.  <!-- strengthen this sentence with some refs? -->

The topic of floral preference in bumble bees has been approached via both laboratory experimentation [e.g. @Kugler1943-uy; @Raine2007-nq] and observational field study [e.g. Heinrich1976-wc; @Goulson2008-zy]. The former approach enables stronger causal inference with respect to the mechanisms of floral choice, but it suffers all the usual limitations of biological realism associated with experimental control, and the higher-order effects of floral choices on network structure cannot be observed meaningfully under laboratory conditions. Observational field studies, however, tend to be anecdotal in scope, typically describing bumble bee foraging patterns in particular localities, over short time frames, and with limited floristic data to enable a distinction between true foraging bias and the neutral effects of floral abundance [though see @Neumayer1999-il]. The steep elevation gradients of mountain slopes provide an opportunity for more powerful observational study of bumble foraging because they generate broad gradients of floristic and abiotic variation while controlling for confounding effects of geographic variation, such as genotypic differentiation within bumble bee species. When bumble bee visitation in such environments is sampled with sufficient spatial and temporal replication and complemented with quantitative floral surveying, inference can be made regarding both the trait interactions that underpin floral biases and the network structures that result from them. Mountain habitats are also arguably the most relevant places to study bumble bee ecology, since bumble bees are adapted to cool climates, and their global distribution is concentrated in areas of high elevation or high latitude. In such environments, bumble bees tend to dominate the pollinator community, and local flora often depend primarily on bumble bee visitation for pollination. <!-- This paragraph needs to be beefed up a bit, but it at least sketches out the points I want to make. I remember a reference somewhere -- I think in one of the Nordic papers, and I think a recent one -- that suggested that the morphology of alpine flowers has been shaped by selection pressure to attract bumble bees -->

# 2. Methods

## 2.1 Field sampling

### Study sites

The study was conducted in Berchtesgaden National Park (47.55°N, 12.92°E), located in the Northern Limestone Alps of southeast Germany. The landscape is characterized by mountain pastures mainly surrounded by coniferous forests. We selected 25 study sites (60 x 60 m) on mountain pastures at elevations ranging from 641-2032 m above sea level. Fourteen of these pastures are extensively grazed by cattle or sheep, three are mowed for hay production, and eight have been abandoned throughout the last century and are no longer subject to any human management. The sites sampled in this study are the same as those used in a series of previous studies [@Hoiss2012-sn; @Hoiss2012-kr; Hoiss2015-gv], but the data reported here are otherwise independent.

### Bumble bee visitation

Sampling consisted of repeated visits to each study site at approximately weekly intervals. Samples were only collected on dry days when the air temperature was at least 6°C. During each visit, bumble bees and their activity were recorded during a 50-minute transect walk. Bumblebees observed on or a given flower were counted as floral visitors. Bumble bee queens were identified to species level in the field, whereas workers and males were stored in individually labeled tubes in the freezer for later identification in the laboratory after @Amiet1996-jz.

visitation sample = site-date of bumble bee visitation observations

While floral visitation by male bumble bees was recorded during visitation sampling, we chose to analyze only visitation by queens and workers.

### Flora surveying

floral survey sample = site-date of floral surveying

## 2.2. Data analysis

Prior to analysis, we aligned each visitation sample with a corresponding floral survey sample. In most cases, floral surveying was conducted jointly with bumble bee observation for each site and date. Occasionally, though, floral surveying could not be completed on the same day as bumble bee observation, resulting in a small temporal offset between visitation samples and floral survey samples. In these cases, visitation samples were paired with the nearest available floral survey data within a maximum distance of 7 days. Visitation samples lacking corresponding floral survey data within 7 days were omitted from further analysis. In a few cases, two consecutive visitation samples at a given site shared the same nearest floral survey sample; in these cases, the visitation samples were first pooled, then aligned to their shared nearest floral survey sample. In rare cases, a given floral species was recorded during visitation sampling but missed during floral surveying. These species were retroactively assigned the minimum floral area value of 0.01 m^2.

### 2.2.1 Modeling cumulative bias

To quantify cumulative bias per bumble bee species, we used the discrimination (*d’*) metric of @Bluthgen2006-if, which represents the net deviation of a species’ visitation patterns from the predictions of a null model in which the probability of visitation is purely a function of floral abundance. We calculated *d'* values separately for each visitation sample, with floral abundances supplied by its corresponding floral survey sample.

Preliminary visualization showed that *d'* values tended to become binary (0 or 1) when calculated for bumble bee species for which fewer than 3 individuals were observed in a given visitation sample. This generated a tri-modal distribution of *d'* values, with peaks at 0 and 1 flanking an otherwise approximately normal distribution. To ease model estimation and avoid the influence of the highly uncertain *d'* values calculated based on one or two observations, we chose to filter our data to include for each visitation sample only species that were observed at least 3 times.

The variation of *d'* was analyzed using Bayesian multi-level modeling. We modeled the response variable *d'* as a function of the interaction between elevation and day-of-year, with bumble bee species as a varying intercept term. The continuous predictors elevation and day-of-year were scaled and centered to optimize software performance. <!-- Explain choice of priors (or lack thereof). Maybe also explain the choice of modeling bb.sp as a varying intercept term rather than a fixed effect --> We then specified an alternative model in which the elevation and day of year terms were replaced with a varying intercept term for site. This simpler model was compared to the interaction model via leave-one-out (LOO) cross-validation using the *loo()* function from *brms*, which is a wrapper for a function of the same name from the R package *loo* [@Yao2017aa].

### 2.2.2 Modeling visitation rate as a function of floral morphotype and bumble bee tongue length

The discrimination metric *d'* expresses the overall disproportionality between relative abundance and visitation, but it does not reveal the causes of bias, i.e. the functional relationships between bumble bees and flowers that lead to differential visitation. To explore these functional underpinnings of bumble foraging bias, we analyzed visitation rate as a function of the interaction between floral morphology and bumble bee tongue length, conditioned on the abundance of each flower and bumble bee species.

Bumble bee tongue-length data were obtained from the literature values compiled and summarized by @Arbetman2017-xj. Values for *B. gerstaeckeri* and *B. mendax*, which were not included in @Arbetman2017-xj, were obtained from @Obeso1992-yp and @Durieux2000-uk, respectively. We then binned bumble bee species into three tongue-length classes: short (< 7.5 mm), medium (7.5---10.0 mm), and long (>= 10.0 mm). Short-tongued species included *B. jonellus*, *B. lapidarius*, *B. pratorum*, *B. psithyrus-group*, *B. pyrenaeus*, and *B. terrestris/lucorum*. Medium-tongued species included *B. humilis*, *B. hypnorum*, *B. monticola*, *B. mucidus*, *B. pascuorum*, and *B. wurflenii*. Long-tongued species included *B. gerstaeckeri*, *B. hortorum*, and *B. mendax*.

Floral morphology was summarized using floral morphotypes of @Kugler1970-gb, accessed via the BIOFLOR database [@Klotz2002-uc]. Morphotypes were simplified to the basic categories numbered 0-10: 0 = not applicable (includes wind-pollinated flora such as grasses), 1 = disc- and bowl-shaped flowers, 2 = funnel flowers, 3 = bell-shaped flowers, 4 = stalk-disc flowers, 5 = lip flowers, 6 = flag flowers, 7 = flower heads, 8 = spike flowers, 9 = brush flowers, 10 = trap flowers **(Figure X)**. <!-- I think I would like to include a figure with photographs of flowers exemplifying each morphotype --> In a few cases, a plant species in our data set was not found in the BIOFLOR database. In most cases, morphotypes are constant at the genus level, and missing morphotypes were added based on the morphotypes found for congeners. The genus *Gentiana* consisted of both group 2 and group 4 representatives, so the missing species *Gentiana aspera* and *Gentiana ciliata* were classified based on visual comparison with their classified congeners. In the case of the genus *Phyteuma*, we followed @Neumayer1999-il and classified it as a group 7 flower head.

We modeled visitation using multilevel Bayesian Poisson regression. First, we specified a model focused on comparing visitation rates across morphotypes. Visitation was modeled using varying intercept terms for morphotype, site, and bumble bee species, and fixed terms for the abundances of floral and bumble bee species. Then, we specified a more complex model in which the varying intercept term for morphotype was replaced with one for the interaction of morphotype and bumble bee tongue length class. In both models, the continuous predictors were scaled and centered to optimize software performance. <!-- Should I perhaps have included a fixed interaction of bb abundance and floral abundance? -->

### 2.2.3 Modeling the relationship between bias and niche relationships 
<!-- I might try to include a plant-centric analysis, too, to see how bias relates to potential competition between plants for BB pollinators. I think it is true to say that if floral resources are not limiting for bees, bee visitation is limiting for flora. -->

#### Model 5: niche_overlap ~ dprime

#### Model 6: Ordination of diet

### 2.2.4 Software and code

All analyses were conducted in R v. 3.6 . Discrimination (*d'*) was calculated using the function *dfun()* [@Dormann2011-aa] from the R package *bipartite* [@Dormann2008-ab].Bayesian multilevel models were implemented in *Stan* [@Stan2020aa] via the R package *brms* [@Buerkner2018aa], and visualizations were constructed with the R packages *tidybayes* [@Kay2020aa] and *ggdist* [Kay2020ab]. Ordination analyses were performed with the R package *vegan* [@Oksanen2019aa] and visualized using the R packages *ggplot2* [@Wickham2016aa] and *ggvegan* [@Simpson2019aa]. Data handling was performed the *tidyverse* suite of R packages [Wickham2019aa]. Code and model diagnostics are available in Appendix X.

# 3. Results
## 3.1 Cumulative bias (d')

LOO model comparison favored the simpler model that used a varying intercept effect for site in lieu of the elevation:day-of-year interaction term (elpd_diff = -104). The estimated global mean of *d'* was 0.38 (0.33---0.43 95% CI). Variation of bumble bee species from the global mean of *d'* was modest, with an estimated standard deviation of 0.05 (0.02---0.08 95% CI) and substantial overlap between species in their posterior distributions **(Figure XA)**. Spatial variation in *d'* was somewhat more pronounced, with an estimated standard deviation of 0.08 (0.06---0.11 95% CI) **(Figure XB)**.

See supplemental material for the details of model specification and evaluation.

```{r fig1, echo = FALSE,  fig.cap = paste("Posterior conditional mean discrimination (d') by bumble bee species and site. Half-eye intervals depict 66\\% and 95\\% CIs.")}
knitr::include_graphics("../output/brm_dprime.sp01_plot1.png")
```

## 3.2 Explaining bias via analysis of visitation rate by flower morphotype and bee tongue length

Visitation rates varied markedly across morphotypes **(Figure X)**. Flag, lip, and bell flowers had the highest visitation rates, with broadly overlapping posterior distributions. Funnel flowers and bell flowers had the next highest visitation rates, followed by stalk-disc flowers and then disc flowers. Unclassified flowers, which are mainly wind-pollinated species, had the lowest visitation rates.

```{r fig2, echo = FALSE,  fig.cap = paste("Posterior conditional mean visits (log-scale) by Kugler morphotype. Half-eye plots depict 66\\% and 95\\% CIs.")}
knitr::include_graphics("../output/brm_pvis00_plot1.png")
```

Visitation also varied strongly within morphotypes across bumble bee tongue length classes **(Figure X)**. Disc flowers (group 1), bell flowers (group 3), and unclassified flowers (group 0) were visited chiefly by short-tongued bumble bee species; visitation of these morphotypes by medium- and long-tongued species was much lower, and there was little difference in visitation rates between medium- and long-tongued species. Funnel flowers (group 2) were visited mostly by long-tongued species, with medium- and short-tongued species following in rank order. The opposite pattern appears in head flowers (group 7), which were visited at the highest rate by short-tongued species, with medium- and long-tongued species following in reverse rank-order. Lip flowers (group 5) were visited at high rates by all tongue-length classes, with little differentiation between them. Flag flowers (group 6) were received similarly high rates of visitation from long- and medium-tongued species, with lower visitation from short-tongued species.

```{r fig3, echo = FALSE,  fig.cap = paste("Posterior conditional mean visits (log-scale) by Kugler morphotype and bumble bee tongue length class. Half-eye plots depict 66\\% and 95\\% CIs.")}
knitr::include_graphics("../output/brm_pvis11_plot1.png")
```

## 3.3. Consequences of bias for niche relationships

# 4. Discussion

## 4.1 The phenomenon of bias: Cumulative bias and variation across species

## 4.2 The mechanism of bias: Explaining bias by the interaction of flower and bumble bee morphology

## 4.3 The outcome of bias: Linking bias to niche relationships

# Sandbox

constrained by the learning and handling costs associated with available flower morphologies, and these costs are determined principally by the morphological fittedness between bumble bee tongue length and floral corolla depth. Short-tongued bumble bee species exploit deep flowers less efficiently than their long-tongued counterparts (robbing notwithstanding), and long-tongued species are less efficient than short-tongued species when foraging on shallow flowers [@Inouye1980-mv].

This phenomenon is formalized by the discrimination (d’) metric of @Bluthgen2006-if, but to avoid conflation with the metric *per se*, we use the term bias in a general sense to describe disproportionality between pollinator visitation and floral abundance.

# 5. References