### 2021-02-05 writing/manuscript_v5.Rmd

---
title: "Bumble bees exhibit strong but flexible foraging biases through temporal and elevational gradients of floristic turnover"
author:
- Douglas B. Sponsler
- Fabrice Requier
- Katharina Kallnik
- Alice Claßen
- A. Fabienne Maihoff
- Ingolf Steffan-Dewenter
bibliography: refs.bib
output:
  html_document:
    df_print: paged
  pdf_document:
    fig_caption: yes
---

# 1. Introduction

Bumble bees are exceptionally versatile foragers whose ability to solve the puzzles posed by complex flower morphologies [@Plowright1984-lc], or indeed by human investigators [@Alem2016-rn; @Loukola2017-bt], renders almost any flower potentially accessible. Even when a gross mismatch between tongue length and corolla depth makes the nectar of a given flower inaccessible by ordinary means, some bumble bee species will resort readily to biting through the base of the flower to “rob” nectar, incidentally making the nectar available to other prospective foragers as well. Flower handling skills are, moreover, socially transmitted, amplifying the plasticity of individual behavior [@Leadbeater2008-rh; @Loukola2017-bt]. Nevertheless, the flexibility of bumble bee foraging is tuned to floral reward, which depends on both the intrinsic reward of a given flower and the searching/learning/handling costs associated with exploiting it. At an evolutionary level, the innate biases of bumble bees for certain floral traits likely reflect historic associations between floral traits and floral rewards [@Raine2007-nq]. These innate biases, however, can be modulated or even nullified by patterns of reward learned during foraging [@Kugler1943-uy; Maharaj2019-wm], which depend not only on intrinsic floral properties but also ecological modulators, such as the spatial distribution of floral patches or competition with other bumble bees [@Ishii2013-sm; @Plowright1984-lc; @Sowig1989-ld].

The topic of floral bias in bumble bees has been approached via both laboratory experimentation [e.g. @Kugler1943-uy; @Raine2007-nq] and observational field study [e.g. Heinrich1976-wc; @Goulson2008-zy]. The former approach enables stronger causal inference with respect to the mechanisms of floral choice, but it suffers from all the usual limitations of biological realism associated with laboratory conditions. Observational field studies, however, tend to be anecdotal in scope, typically describing bumble bee foraging patterns in particular localities, over short time frames, and with limited floristic data to enable a distinction between true foraging bias and the neutral effects of floral abundance. The steep elevation gradients of mountain slopes provide an opportunity for more powerful observational study of bumble foraging because they generate broad gradients of environmental variation while controlling for confounding effects of geographic variation, such as genotypic differentiation within bumble bee species. When bumble bee visitation in such environments is sampled with sufficient spatial and temporal replication and complemented with quantitative floral surveying, true bias can be distinguished from the neutral effects of abundance, and variation in bias can be explored along elevational and temporal dimensions of floristic turnover.

In the present study, we focus on the question of bumble bee foraging bias with respect to floral morphology, and the lability of morphological biases across gradients of floristic variation. At the outset, we emphasize the use of the term "bias" in favor of the term "preference", since the latter is laden with more assumptions and its inference under field conditions is dubious []. By "bias", we mean simply an empirical disproportionality between the relative abundance of a given floral taxon or group and its proportional visitation. The exact mechanism of this disproportionality --- whether an arbitrary "preference" on the part of the foraging bee, an unmeasured decoupling of relative abundance and rate of reward, an exogenous effect of competition, etc. --- is beyond the scope of our discussion, and probably best studied under laboratory conditions where strict factorial control is possible. We focus instead on the empirical phenomenon of bias and its morphological correlations.

The meaning of "morphology", however, is hardly more clear-cut than that of bias. Bumble bees and flowers vary in an array of interrelated morphological parameters, both within and between species, and the selection of relevant and measurable morphological metrics is nontrivial. For bumble bees, morphological variation consists principally (though not exclusively) of variation in tongue length. Tongue length in bumble bees corresponds functionally to corolla depth in flowers, and there is a well-established (though imperfect) association between tongue length and floral choice [e.g. @Inouye1980-mv]. Importantly, though, accessing floral reward involves more than just a match between tongue length and corolla depth. Such matching is entirely bypassed by nectar robbers, but even in the case of "legitimate" foraging, access to reward involves learning to manipulate floral structures that, in some floral morphotypes, are challenging irrespective of tongue length [@Laverty1994-hm]. For example, the characteristic Papilionaceous floral morphology of legumes requires a bee to separate tightly appressed keel petals to access anthers and nectaries [@Westerkamp1997-hs], and the narrowly campanulate flowers of many Ericaceous flora pose a stringent barrier to large-bodied visitors like bumble bees [@Wallace1977-kz]. Rather than considering various floral morphological traits independently, we will rely on the synthesis of @Kugler1970-gb, who developed a typology consisting of ten primary morphotypes and nested subdivision of each. Kugler's typology was informed by laboratory studies of bumble bee flower choice [@Kugler1943-uy], and there is a precedent for analyzing bumble bee visitation in terms of Kugler's morphotypes [@Neumayer1999-il; @Schneller2014-ry].

First, we compare the overall rate of visitation across floral morphotypes. Then, we parse visitation rates by bumble bee species and explore the functional relationship between floral morphotype and bumble bee tongue length. We then compare overall diet similarity across bumble bee species as an emergent consequence of their respective biases. Lastly, we analyze the flexibility of the morphological preferences of individual bumble bee species through elevation and time.

# 2. Methods

## 2.1 Field system

### Study sites

The study was conducted in Berchtesgaden National Park (47.55°N, 12.92°E), located in the Northern Limestone Alps of southeast Germany. The landscape is composed of mountain pastures mainly surrounded by coniferous forests. We selected 25 study sites (60 x 60 m) on mountain pastures at elevations ranging from 641-2032 m above sea level. Fourteen of these pastures are extensively grazed by cattle or sheep, three are mowed for hay production, and eight have been abandoned throughout the last century and are no longer subject to any human management. The sites sampled in this study are the same as those used in a series of previous studies [@Hoiss2012-sn; @Hoiss2012-kr; Hoiss2015-gv], but the data reported here are otherwise independent.

### Bumble bee visitation

Sampling consisted of repeated visits (hereafter, "visitation samples") to each study site at approximately weekly intervals. Samples were only collected on dry days when the air temperature was at least 6°C. During each visit, bumble bees and their activity were recorded during a 50-minute transect walk. Bumblebees observed on or a given flower were counted as floral visitors. Bumble bee queens were identified to species level in the field, whereas workers and males were stored in individually labeled tubes in the freezer for later identification in the laboratory after @Amiet1996-jz. While floral visitation by male bumble bees was recorded during visitation sampling, we chose to analyze only visitation by queens and workers. During observation, it was noted whether each bee was visibly carrying pollen at the time of its visit, but this provides at best an imperfect distinction between pollen and nectar foragers, and we opted not to differentiate between bees with and without pollen in our analyses.

### Floral surveying

floral survey sample = site-date of floral surveying

## 2.2. Data analysis

### 2.2.1 Bumble bee and floral morphology data

Bumble bee tongue-length data were obtained from the literature values compiled and summarized by @Arbetman2017-xj. Values for *B. gerstaeckeri* and *B. mendax*, which were not included in @Arbetman2017-xj, were obtained from @Obeso1992-yp and @Durieux2000-uk, respectively. We then binned bumble bee species into three tongue-length classes: short (< 7.5 mm), medium (7.5---10.0 mm), and long (>= 10.0 mm). Short-tongued species included *B. jonellus*, *B. lapidarius*, *B. pratorum*, *B. psithyrus-group*, *B. pyrenaeus*, and *B. terrestris/lucorum*. Medium-tongued species included *B. humilis*, *B. hypnorum*, *B. monticola*, *B. mucidus*, *B. pascuorum*, and *B. wurflenii*. Long-tongued species included *B. gerstaeckeri*, *B. hortorum*, and *B. mendax*.

Floral morphology was summarized using floral morphotypes of @Kugler1970-gb, accessed via the BIOFLOR database [@Klotz2002-uc]. Morphotypes were simplified to the basic categories numbered 0-10: 0 = not applicable (includes wind-pollinated flora such as grasses), 1 = disc- and bowl-shaped flowers, 2 = funnel flowers, 3 = bell-shaped flowers, 4 = stalk-disc flowers, 5 = lip flowers, 6 = flag flowers, 7 = flower heads, 8 = spike flowers, 9 = brush flowers, 10 = trap flowers **(Figure X)**. <!-- I think I would like to include a figure with photographs of flowers exemplifying each morphotype --> In a few cases, a plant species in our data set was not found in the BIOFLOR database. In most cases, morphotypes are constant at the genus level, and missing morphotypes were added based on the morphotypes found for congeners. The genus *Gentiana* consisted of both group 2 and group 4 representatives, so the missing species *Gentiana aspera* and *Gentiana ciliata* were classified based on visual comparison with their classified congeners. In the case of the genus *Phyteuma*, we followed @Neumayer1999-il and classified it as a group 7 flower head.

### 2.2.2 Processing visitation and survey data

Prior to analysis, we aligned each visitation sample with a corresponding floral survey sample. In most cases, floral surveying was conducted jointly with bumble bee observation for each site and date. Occasionally, though, floral surveying could not be completed on the same day as bumble bee observation, resulting in a small temporal offset between visitation samples and floral survey samples. In these cases, visitation samples were paired with the nearest available floral survey data within a maximum distance of 7 days. Visitation samples lacking corresponding floral survey data within 7 days were omitted from further analysis. In a few cases, two consecutive visitation samples at a given site shared the same nearest floral survey sample; in these cases, the visitation samples were first pooled, then aligned to their shared nearest floral survey sample. In rare cases, a given floral species was recorded during visitation sampling but missed during floral surveying. These species were retroactively assigned the minimum floral area value of 0.01 m^2.

Visitation data were extremely right-skewed because the majority of available flora received zero visits during a given sampling period. To facilitate model fitting, we created a binary version of visitation rate in which visited flora were assigned to 1 and unvisited flora to 0. We then modeled binary visitation using generalized models with binomial or Bernoulli distribution families. The conditional effects inferred from these models should, therefore, be interpreted as effects on *visitation probability* rather than visitation rate.

### 2.2.3 Model 1: Comparing overall visitation probability across floral morphotypes

We first compared overall visitation probability across floral morphotypes. In this model, visitation for each sample was pooled across bumble bee species and within floral morphotypes prior to binarization. Visitation probability was then modeled using multilevel Bayesian regression with a Bernoulli distribution family (logit link function). *Bumble bee abundance* and *percent flower cover* (per morphotype) were included as a global interaction term, thus conditioning visitation probability on abundance and enabling the inference of bias. *Floral morphotype* was specified as a varying effect term with both a varying intercept and a varying slope with respect to *percent flower cover*, and a varying intercept term was also included for *site*. See Appendix X for the details of model specification and evaluation.

### 2.2.4 Model 2: Visitation probability as a function of floral morphotype and bumble bee species

To explore the functional underpinnings of bumble foraging bias, we analyzed visitation probability as a function of the interaction between floral morphology and bumble bee tongue length class. For this analysis, visitation was pooled within floral morphotype but not across bumble bee species. We then specified a Bayesian model of the same form as Model 1, but this time included a varying slope and intercept term for *floral morphotype : bumble bee tongue length class* and *floral morphotype : bumble bee species*. In this model, the relationship between floral morphotype and bumble bee tongue length class was the focus of inference, and the bumble bee species effect was added to avoid pseudoreplication arising from the grouping of observation within species. See Appendix X for the details of model specification and evaluation.

### 2.2.4 Model 3: Visitation probability as a function of floral morphotype and bumble bee tongue length

To focus on species-level morphotype biases, we removed the tongue length effect from Model 2 and re-fit the model considering only bumble bee species. See Appendix X for the details of model specification and evaluation.

### 2.2.5 Model 4: Aggregate diet ordination

To visualize the net diet similarity between bumble bee species, we pooled visitation data across all samples within bumble bee species and plant species, then performed a principal coordinates analysis based on the Jaccard distance metric. To test the significance of tongue length class as a predictor of diet similarity, we performed a permutation test.

### 2.2.6 Model 5: The variation of morphological bias through elevation and time

To explore the plasticity of floral biases, we first subsetted our data to include only the dominant bumble bee species (*B. hortorum*, *B. pascuorum*, *B. pratorum*, *B. soroensis*, *B. terrestris/lucorum*, and *B. wurflenii*) and the most visited floral morphotypes (groups 3, 5, 6, and 7). We then constructed a generalized additive model in which binary visitation was modeled as a tensor-product smooth of *elevation* and *day-of-year*, grouped by floral morphotype. This response was conditioned on linear interaction terms for *bumble bee abundance* : *percent flower cover* and *percent flower cover* : *floral morphotype* and a random intercept term for *bumble bee species* : *floral morphotype*. See Appendix X for the details of model specification and evaluation.

### 2.2.7 Software

All analyses were conducted in R v. 3.6. Bayesian multilevel models were implemented in *Stan* [@Stan2020aa] via the R package *brms* [@Buerkner2018aa], and visualizations were constructed with the R packages *tidybayes* [@Kay2020aa] and *ggdist* [Kay2020ab]. Ordination analysis and permutation testing were performed with the R package *vegan* [@Oksanen2019aa] and visualized using the R packages *ggplot2* [@Wickham2016aa] and *ggvegan* [@Simpson2019aa]. GAMs were implemented with the R package *mgcv* [@Wood2017aa] and visualized with the R package *mgcViz* [@Fasiolo2018aa]. Data handling was performed the *tidyverse* suite of R packages [Wickham2019aa]. Input data and code are available in Appendix X.

# 3. Results
## 3.1 Model 1: Overall visitation probability by floral morphotype

Since we report the results of a Bernoulli regression model, the term "visitation rate" used here should be interpreted as conditional visitation *probability*, not conditional visitation frequency.

Lip (group 5) and flag (group 6) flowers had the highest overall visitation probability, with a conditional mean visitation probabilities of 0.36 (95% CI = 0.30--0.42) and 0.33 (95% CI = 0.27--0.39), respectively **(Figure X)**. In second place, head flowers (group 7) had a mean conditional visitation probability of 0.19 (95% CI = 0.16--0.24). Bell (group 3) flowers were a distant third with a conditional mean visitation probability of 0.092 (95% CI = 0.07--0.12), followed by stalk-disc (group 4), funnel (group 2), and disc flowers (group 1) with conditional visitation probabilities of 0.07 (95% CI = 0.05--0.09), 0.05 (95% CI = 0.04--0.07), and 0.05 (95% CI = 0.03--0.07), respectively. Unclassified (group 0) flowers, consisting mainly of wind-pollinated grasses, were the least-visited, with a conditional mean visitation probability of 0.3 (95% CI = 0.02--0.04). Our model explained approximately 35% of the total variance (97.5% CI = 34--36%).

```{r fig1, echo = FALSE,  fig.cap = paste("Posterior conditional mean visitation probability by floral morphotype. Point-interval plots depict 66\\% and 95\\% CIs.")}
knitr::include_graphics("../output/brm_pvis01_plot1.png")
```

## 3.2 Models 2 and 3: Visitation probability by flower morphotype, bumble bee species, and bumble bee tongue length class

Descriptions of tongue classes are based on Model 2 and refer to **(Figure X)**, and descriptions of species are based on Model 3 and refer to **(Figure X)**. Both models explained approximately 28% of total variance (97.5% CI = 27--29%). As in Model 1, visitation rates refer to conditional visitation probabilities.

Disc (group 1) flowers were visited at similar rates across tongue length classes. At the species level, disc flowers were visited at relatively high rates by the short-tongued *B. terrestris/lucorum* and the medium-tongued *Bombus monticola*. Visitation to funnel (group 2) flowers was dominated by long-tongued species, an effect driven entirely by the *Aconitum* (group 2) specialist *B. gerstaeckeri*. Bell (group 3) flowers were visited at a higher rate by short-tongued species than by long-tongued species, with medium-tongued species exhibiting intermediate visitation rates. At the species-level, bell flowers were visited at highest probability by the short-tongued species *B. jonellus*, *B. pyrenaeus*, and *B. soroensis*, and by the medium-tongued species *B. hypnorum* and *B. monticola*. Stalk-disc (group 4) flowers were visited at similar rates across all tongue length classes, and there was also broad overlap in visitation rates at the species level. Lip (group 5) flowers were also visited at similar rates across tongue length classes, but at the species level *B. pascuorum* exhibited an uniquely high preference for lip flowers. Flag (group 6) flowers were visited at a higher rate by medium-tongued species than by short-tongued species, with long-tongued species  exhibiting intermediate visitation rates. Flag flowers generated strong separation in visitation rate at the species level, receiving very high visitation from the medium-tongued *B. pascuorum* and *B. mucidus*, moderately high visitation from the short-tongued *B. terrestris/lucorum* and *B. lapidarius*, the medium-tongued *B. monticola*, and the long-tongued *B. mendax* and *B. hortorum*, and low visitation by the short-tongued *B. soroensis*,  *B. pyrenaeus*, *B. psithyrus*, *B. pratorum*, and *B. jonellus*, the medium-tongued *B. hypnorum*, and the long-tongued *B. gerstaeckeri*.

```{r fig2, echo = FALSE,  fig.cap = paste("Posterior conditional mean visitation probability by Kugler morphotype and bumble bee tongue length class. Point-interval plots depict 66\\% and 95\\% CIs. Letters indicate group differences based on evidence ratio at alpha = 0.05. Note that the range of the x-axis has been rescaled for each panel to emphasize the comparison of visitation probabilities across tongue length classes.")}
knitr::include_graphics("../output/brm_pvis00_plot1_mod.png")
```
```{r fig3, echo = FALSE,  fig.cap = paste("Posterior conditional mean visitation probability by Kugler morphotype and bumble bee species. The tongue length class of each species is color-coded. Point-interval plots depict 66\\% and 95\\% CIs. Note that the range of the x-axis has been rescaled for each panel to emphasize the comparison of visitation probabilities across bumble bee species.")}
knitr::include_graphics("../output/brm_pvis02_plot1.png")
```

## 3.3. Model 4: Diet ordination

Axis 1 accounted for 24% of total diet variance and primarily separated long- and medium-tongued species from short-tongued species. Long- and medium-tongued species scored low on Axis one and were associated with the lip flowers (group 5) *Ajuga reptans*, *Rhinanthus glacialis*, and *Stachys alopecuros*) and flag flowers (group 6) *Anthyllis vulneraria*, *Lotus corniculatus*, and *Trifolium pratense*. Short tongued species scored high on Axis one and were associated with the bell flowers (group 3) *Campanula scheuchzeri*, *Erica carnea*, and *Geum rivale* and the head flowers (group 7) *Carduus defloratus*, *Centaurea jacea*, and *Phyteuma orbiculare*. Axis 2 accounted for 19% of total variance and primarily explained variation within short-tongued species. Short-tongued species scoring high on Axis 2 were associated with *Campanula scheuchzeri* while those scoring low on Axis 2 were associated with *Erica carnea*. Short-tongued species scoring close to zero on Axis 2 also scored lower on Axis 1 (closer to medium- and long-tongued species) and were associated with head flowers. Tongue length class explained 22% of diet variance (Adonis permutation test, F = 1.68, p  = 0.006). While short-tongued bees appeared most broadly spread in ordination space, there were no significant differences in dispersion across tongue length classes (F = 0.62, p = 0.55).

```{r fig5, echo = FALSE,  fig.cap = paste("Principal coordinates ordination of bumble bee visitation pooled across all samples.")}
knitr::include_graphics("../output/pcoa_aggregate.png")
```

## 3.3 Model 5:

Model 5 explained 33.8% of total variance (adjusted R2 = 0.37). All tensor product smooths were significant at p < 0.001 except for the one representing *B. pratorum* and flag (group 6), which was more weakly significant at p = 0.02. As in Models 1-3, the results of Model 5 should be interpreted in terms of conditional visitation probability.

Bell (group 3) flowers were visited at highest rates early in the season and at low elevation by all species except *B. hortorum*, whose visitation rate to bell flowers peaked in mid-season and at mid-elevation. Notably, the medium-tongued *B. pascuorum* and *B. wurflenii* also exhibited relatively high visitation to bell flowers at mid-season and mid-elevation, while the visitation rates of the short-tongued *B. pratorum*, *B. soroensis*. and *B. terrestris lucorum* were more concentrated at the early-season and low-elevation peak.

Lip (group 5) flowers exhibited roughly the mirror image of the pattern of visitation rate seen in bell flowers, with peak visitation late in the season and at low elevation, but relatively high visitation extending toward mid-season and mid-elevation. This pattern was consistent for all bumble bee species.

The pattern of visitation rates to flag (group 6) flowers fell generally along the lines of tongue length class. The short-tongued *B. pratorum* and *B. soroensis* exhibited peak visitation early in the season and at low elevation. The long-tongued *B. hortorum* and the medium-tongued *B. pascuorum* and *B. wurflenii* exhibited peak visitation at mid-season, though *B. pascuorum* exhibited a bimodal pattern in elevation, with peaks around 1000 m and 1750 m, while the visitation rate of *B. hortorum* declined monotonically from high elevation to low. The short-tongued *B. terrestis/lucorum* broke the tongue-length pattern, exhibiting a diffuse pattern of visitation rate more similar to that of the long- and medium-tongued species than the other short-tongued species.

Visitation rates for head (group 7) flowers were more or less bimodal for all species, with a first peak in mid-season and low elevation and a second peak in late-season and high elevation.⎄

```{r fig6, echo = FALSE,  fig.cap = paste("GAM analysis of bias as a function of the tensor-product interaction of elevation and day of year, parsed by bumble bee species and floral morphotype and conditioned on the abundance of each bumble bee species and the relative abundance of each floral morphotype. The color ramp of each smooth has been scaled to equal color range, thereby obscuring the absolute magnitude of effect. To avoid cluttering the figure, we have opted not to include a unique color ramp key for each plot, since the focus of our inference is on relative rather than absolute patterns.")}
knitr::include_graphics("../output/gam_plot1_mod.png")
```

# 4. Discussion

## 4.1 Foraging bias and floral morphology

Our study is not the first to compare bumble bee visitation rates across Kugler's (1970) floral morphotypes [see @Neumayer1999-il; @Schneller2014-ry], but to our knowledge it is the first to enable the inference of bias across morphotypes by controlling statistically for the neutral effects of abundance.

When bumble bee visitation is pooled across all species, it is clear that lip, flag, and head flowers dominate the rest of the floral community in terms of conditional visitation probability. Patterns of pooled visitation are, of course, driven disproportionately by the behavior of the most abundant bumble bee species, which in our system were *B. hortorum*, *B. pascuorum*, *B. pratorum*, *B. psithyrus*, *B. soroensis*, *B. terrestris/lucorum*, and *B. wurflenii*. Visitation to lip and flag flowers was driven especially by the medium-tongued *B. pascuorum*, the most abundant bumble bee in our study system, and secondarily by the medium-tongued *B. wurflenii*. Visitation to head flowers was driven mainly by the short-tongued *B. soroensis* and *B. psithyrus*.

The influence of tongue length on visitation bias was pronounced for some morphotypes but not for others. In the cases of disc, and stalk-disc, and unclassified flowers, visitation rates were uniformly low across all tongue length classes. The opposite was true of lip flowers, which were visited at high rates by all tongue length classes. Bell flowers and head flowers were visited chiefly by short-tongued bees, consistent with the shallow corollas of these morphotypes. In contrast, funnel flowers, with their deep corollas were visited chiefly by long-tongued bees, though this effect was driven disproportionately by *B. gerstaeckeri*, which specializes on the funnel flowers of the genus *Aconitum*. Flag flowers were visited at the highest rate by medium-tongued species, and especially by *B. pascuorum* and *B. mucidus*.

While variation in floral bias across tongue length classes largely comports with the established relationship between tongue length and corolla depth, variation in bias *within* tongue length classes demonstrates the insufficiency of simple morphological matching to explain bumble bee foraging patterns. Nectar robbing is one mechanism by which the relationship between tongue length and corolla depth might be relaxed, but its influence in our data is unclear. Of the bumble bees in our study system, *B. terrestris/lucorum* and especially *B. wurflenii* are known to engage readily in nectar robbing. In the case of *B. wurflenii*, however, its morphotype affinities did not differ markedly from those of other medium-tongued species, and it may be that the condition of medium-tongue length alone confers enough to foraging flexibility in our study system to dampen the divergent effects of nectar robbing. Nectar robbing might, however, account for the deviation of *B. terrestris/lucorum* from most other short-tongued species with respect to flag flowers. With the exception of *B. lapidarius*, all short-tongued species besides *B. terrestris/lucorum* eschewed flag flowers, while *B. terrestris/lucorum* visited them at rates comparable to medium- and long-tongued species. Indeed, *B. terrestris/lucorum* was, on the whole, an idiosyncratic forager, demonstrating a relatively high affinity for disc flowers and unclassified flowers (particularly *Plantago lanceolata*) that were largely ignored by other species. Perhaps the most striking divergence of floral bias within tongue length class, though, was the divisive effect of bell flowers on short-tongued bumble bees. *B. pyrenaeus*, *B. soroensis*, *B. pratorum*, and *B. jonellus* visited bell flowers at high rates, while *B. terrestris/lucorum*, *B. psithyrus*, and *B. lapidarius* visited them relatively rarely. There is no obvious morphological distinction between these groups of short-tongued bees that would account for their divergent affinities for bell flowers; ecological drivers such as competition or differences in spatial or temporal activity patterns are more likely explanations.

Marked differences in the relative visitation rates of floral morphotypes compared to [@Schneller2014-ry], but also some consistencies. We reproduce their finding that *B. soroensis* is partial to bell flowers. There data collection was limited to July and August, which probably explains their finding that head (group 7) flowers were most visited, followed by lip (group 5) flowers. Bell flowers were strongly underrepresented in their data, but that makes sense given what I see in my smooth plots. They are most attractive early in the year.

## 4.2 The emergent phenomenon of diet similarity

Our finding that bumble bee diet ordination produces a first axis corresponding to tongue length and a second axis describing the dispersion of short-tongued species is strikingly consistent with past studies [@Goulson2008-zy; Durieux2000-uk].

## 4.3 The lability of bias through elevation and time

Species differed not just in their overall bias toward given morphotypes, but also in *when* and *where* they were biased toward a morphotype.

Ours is, to our knowledge, the first study to characterize formally the variation of bumble bee foraging bias through elevation an time.

# 5. References