### 2021-02-13 code/abund_div_GAM.Rmd
update

---
title: "Abundance/diversity GAM"
output: html_document
---

### Prepare environment
```{r setup, include=FALSE, echo=FALSE}
library(mgcv)
library(mgcViz)
library(tidyverse)
library(lubridate)
```

### Load site and trait data
```{r, include=FALSE, echo=FALSE}
site_data <- read_csv("../data/processed_data/site_data.csv") %>%
  dplyr::select(-c(slope.calc, slope.est, elev.min, elev.max, temp.mean)) %>% # drop these variables
  mutate(elev.class2 = factor(elev.class2, levels = c("low", "mid", "high")),
         site = factor(site)) # turn elav.class into an ordered factor and site into a regular factor

# BB traits
bb_traits <- read_csv("../data/processed_data/bb_traits.csv")

# FL traits
## family-level taxonomy
fl_tax <- read_csv("../data/processed_data/floral_tax.csv")

## Kugler morphotypes (accessed through Bioflor)
fl_ktype <- read_csv("../data/processed_data/floral_k_type.csv")

## Flower colors (accessed through BioFlor)
fl_color <- read_csv("../data/processed_data/floral_color.csv")

## diameter and corolla depth data from Gita Benadi
fl_morph <- read_csv("../data/processed_data/flower_morphology_GitaBenadi.csv") %>% 
  dplyr::select(SampleID, plant.family = Family, plant.sp = Species, 
                Unit, Flowers_per_unit, diam = Diameter, LN) %>%
  group_by(plant.sp) %>%
  summarize(diam = mean(diam), LN = mean(LN))

## verbal description of Kugler morphotypes
fl_ktype_key <- read_delim("../data/processed_data/kugler_key.tsv", delim = "\t", col_names = FALSE) %>%
  dplyr::select(k.type = 1, description = 2)

## join all floral trait tables
fl_traits <- fl_tax %>%
  full_join(fl_ktype) %>%
  full_join(fl_color) %>%
  full_join(fl_morph) %>%
  dplyr::select(plant.sp, plant.genus, plant.family, k.type, k.type.s, k.type.ss, color, diam, LN)
```

### Load network and survey data
```{r message=FALSE}
### BB-FL visitation data
net <- read_csv("../data/processed_data/network.csv") %>%
  filter(caste != "male") %>%
  dplyr::select(site, date, bb.sp, plant.sp) # drop unused columns

### Floral survey data
survey <- read_csv("../data/processed_data/floral_survey.csv") %>%
  semi_join(net, by = "plant.sp") %>% # Consider only visited plant species (anytime anywhere, not specifically for a given site*date)
  dplyr::select(site, date, plant.sp, flower.cover) %>%
  group_by(plant.sp) %>%
  mutate(present = case_when(
    max(flower.cover) > 0 ~ TRUE,
    max(flower.cover) == 0 ~ FALSE
  )) %>%
  filter(present == TRUE) # Ignore plants that never had any flower cover

vis_sp <- net %>%
  left_join(fl_tax) %>%
  select(plant.sp, plant.genus, plant.family) %>%
  unique()

surv_sp <- survey %>%
  left_join(fl_tax) %>%
  select(plant.sp, plant.genus, plant.family) %>%
  unique()

missing <- anti_join(vis_sp, surv_sp)
```

### FL and BB abundance / diversity
```{r echo=FALSE, include=TRUE, message=FALSE}
# Per-transect BB abundance
## full dataset
bb_abund_sp <- net %>%
  dplyr::select(site, date, bb.sp) %>%
  group_by(site, date, bb.sp) %>% # 
  summarize(bb.abund = n()) %>% # get per transect bb abundance
  ungroup() %>%
  complete(site, date, bb.sp, fill = list(bb.abund = 0)) %>% # complete implicit absences
  semi_join(net, by = c("site", "date")) %>% # drop site-dates that did not exist in the data set
  left_join(site_data) %>% # add site data
  mutate(year = factor(year(date)),
         yday = yday(date),
         bb.sp = factor(bb.sp),
         management = factor(management),
         site = factor(site),
         transect = factor(transect)) %>%
  group_by(year, bb.sp) %>%
  mutate(bb.annual.abund = sum(bb.abund),
         abund.class = case_when(
           bb.sp %in% c("telu", "pasc", "prat", "soro", "wurf") ~ "dominant",
           bb.sp %in% c("jone", "psit", "hort") ~ "common",
           bb.sp %in% c("pyre", "mend", "mont", "hypn", "lapi") ~ "high elev.",
           bb.sp %in% c("gers", "humi", "muci") ~ "minor"
         ),
         abund.class = factor(abund.class, levels = c("dominant", "common", "high elev.", "minor")),
         bb.sp.ord = factor(bb.sp, levels = c("telu", "pasc", "prat", "soro", "wurf", 
                                              "hort", "jone", "psit",
                                              "pyre", "mend", "mont",
                                              "gers", "humi", "muci", "hypn", "lapi"))) %>%
  left_join(bb_traits) %>%
  mutate(tongue = factor(pbl.w.class2))

# ## filtered to include only samples with corresponding floral survey data
# bb_abund_sp.sub <- bb_abund_sp %>%
#   ungroup() %>%
#   semi_join(shared_transects) %>%
#   select(site, date, yday, bb.sp, bb.abund, bb.annual.abund, elev.mean)

### Per-transect floral abund by species
fl_abund_sp <- survey %>%
  group_by(site, date, plant.sp) %>% # group by plant.sp*transect
  summarize(fl.abund = sum(flower.cover)) %>% # get plant.sp*transect total flower cover; this step really shouldn't be required, but in a few cases, there is more than one abundance entry for the same species on a given transect; not sure how that happened, but it's in the original data that Fabrice sent me; alternatively, I could drop the dups or average them. We'll see what Kalli and Fabrice have to say.
  group_by(site, date) %>%
  mutate(prop.fl.abund = fl.abund/sum(fl.abund)) %>% # get the proportional flower cover for each species.
  ungroup() %>%
  #full_join(dplyr::select(net, site, date, plant.sp), by = c("site", "date", "plant.sp")) %>% # add the handful of species occurring in the visitation network but missing from the survey data
  complete(site, date, plant.sp, fill = list(fl.abund = 0, prop.fl.abund = 0)) %>%
  semi_join(survey, by = c("site", "date")) %>%  # only include site-dates that actually occurred
  unique() %>%
  mutate(year = factor(year(date))) %>%
  left_join(fl_traits)

### Visualize raw abundance/occurrence
ggplot(filter(bb_abund_sp, bb.abund > 0), aes(yday, elev.mean, color = abund.class)) +
  geom_point(aes(alpha = bb.abund)) +
  #geom_point() +
  #facet_grid(year~ bb.sp) 
  facet_wrap(~bb.sp.ord)

ggplot(bb_abund_sp, aes(bb.sp, log(bb.abund + 1))) +
  geom_boxplot()

ggplot(bb_abund_sp, aes(log(bb.abund + 1))) +
  geom_histogram() +
  facet_wrap(~bb.sp)
```

### Binary "preference" for each plant genus by each bumble bee species; score species 0 or 1 based on whether they were visited
```{r message=FALSE}
weights <- fl_abund_sp %>%
  select(plant.sp, plant.genus) %>%
  unique() %>%
  left_join(select(net, bb.sp, plant.sp)) %>%
  select(plant.genus, bb.sp) %>%
  distinct() %>%
  mutate(visited = rep(1, n())) %>%
  pivot_wider(values_from = visited, names_from = bb.sp, values_fill = 0) %>%
  pivot_longer(cols = 2:17, values_to = "visited", names_to = "bb.sp")

weighted_fl_abund <- fl_abund_sp %>%
  group_by(site, date, plant.genus) %>%
  summarize(flower.cover = sum(fl.abund)) %>%
  left_join(weights) %>%
  mutate(weighted.flower.cover = flower.cover*visited) %>%
  group_by(site, date, bb.sp) %>%
  summarize(total.weighted.flower.cover = sum(weighted.flower.cover)) %>%
  left_join(site_data) %>%
  select(site, date, elev.mean, bb.sp, total.weighted.flower.cover) %>%
  mutate(yday = yday(date),
         bb.sp = factor(bb.sp),
         site = factor(site),
         year = factor(year(date)))

ggplot(weighted_fl_abund, aes(total.weighted.flower.cover)) +
  geom_histogram() +
  facet_wrap(~bb.sp)

ggplot(weighted_fl_abund, aes(bb.sp, total.weighted.flower.cover)) +
  geom_boxplot()

ggplot(weighted_fl_abund, aes(bb.sp, log(total.weighted.flower.cover))) +
  geom_boxplot()
```

### Call GAMs
#### Floral abundance
```{r message=FALSE, warning=FALSE}
# First, let's model overall floral abundance without parsing it out by which species were visited by which bumble bee species
fl_abund_pool <- fl_abund_sp %>%
  group_by(site, date, year) %>%
  summarize(fl.abund = sum(fl.abund)) %>%
  mutate(site = factor(site),
         yday = yday(date)) %>%
  left_join(site_data)

ggplot(fl_abund_pool, aes(fl.abund)) +
  geom_histogram()

fl_abund.qp <- bam(fl.abund ~ 
                      s(site, bs = "re") +
                      s(year, bs = "re") +
                      te(yday, elev.mean,
                         bs = c("gp", "tp"),
                         k = c(8, 8)),
                    data = fl_abund_pool,
                    family = "quasipoisson",
                    discrete = TRUE,
                    select = TRUE,
                    method = "fREML") %>% getViz()

check.gamViz(fl_abund.qp)
summary(fl_abund.qp)
plot(fl_abund.qp)

# Now, by bumble bee species
weighted_fl_abund.qp <- bam(total.weighted.flower.cover ~ 
                              s(bb.sp, bs = "re") +
                              s(site, bs = "re") +
                              s(year, bs = "re") +
                              te(yday, elev.mean,
                                 by = bb.sp,
                                 bs = c("gp", "tp"),
                                 k = c(5, 5)),
                            data = filter(weighted_fl_abund, bb.sp != "humi"),
                            family = "quasipoisson",
                            discrete = TRUE,
                            select = TRUE,
                            method = "fREML") %>% getViz(nsim = 500)

check.gamViz(weighted_fl_abund.qp)
sum_fl.qp <- summary(weighted_fl_abund.qp)
print(plot(weighted_fl_abund.qp, allTerms = TRUE), pages = 4)
check2D(weighted_fl_abund.qp, x1 = "yday", x2 = "elev.mean")

gers_fl_abund <- plot(sm(weighted_fl_abund.qp, 4))
hort_fl_abund <- plot(sm(weighted_fl_abund.qp, 5))
hypn_fl_abund <- plot(sm(weighted_fl_abund.qp, 6))
jone_fl_abund <- plot(sm(weighted_fl_abund.qp, 7))
lapi_fl_abund <- plot(sm(weighted_fl_abund.qp, 8))
mend_fl_abund <- plot(sm(weighted_fl_abund.qp, 9))
mont_fl_abund <- plot(sm(weighted_fl_abund.qp, 10))
muci_fl_abund <- plot(sm(weighted_fl_abund.qp, 11))
pasc_fl_abund <- plot(sm(weighted_fl_abund.qp, 12))
prat_fl_abund <- plot(sm(weighted_fl_abund.qp, 13))
psit_fl_abund <- plot(sm(weighted_fl_abund.qp, 14))
pyre_fl_abund <- plot(sm(weighted_fl_abund.qp, 15))
soro_fl_abund <- plot(sm(weighted_fl_abund.qp, 16))
telu_fl_abund <- plot(sm(weighted_fl_abund.qp, 17))
wurf_fl_abund <- plot(sm(weighted_fl_abund.qp, 18))

pdf("../output/gam_FL.pdf")
grid_plot_fl_abun <- gridPrint(gers_fl_abund + 
                                  l_fitRaster() + 
                                  l_fitContour() + 
                                  labs(title = "gers", x = NULL, y = NULL) +
                                  guides(fill=FALSE),
                               hort_fl_abund + 
                                 l_fitRaster() + 
                                 l_fitContour() + 
                                 labs(title = "hort", x = NULL, y = NULL) +
                                 guides(fill=FALSE),
                               hypn_fl_abund + 
                                 l_fitRaster() + 
                                 l_fitContour() + 
                                 labs(title = "hypn", x = NULL, y = NULL) +
                                 guides(fill=FALSE),
                               jone_fl_abund + 
                                 l_fitRaster() + 
                                 l_fitContour() + 
                                 labs(title = "jone", x = NULL, y = NULL) +
                                 guides(fill=FALSE),
                               lapi_fl_abund + 
                                 l_fitRaster() + 
                                 l_fitContour() + 
                                 labs(title = "lapi", x = NULL, y = NULL) +
                                 guides(fill=FALSE),
                               mend_fl_abund + 
                                 l_fitRaster() + 
                                 l_fitContour() + 
                                 labs(title = "mend", x = NULL, y = NULL) +
                                 guides(fill=FALSE),
                               mont_fl_abund + 
                                 l_fitRaster() + 
                                 l_fitContour() + 
                                 labs(title = "mont", x = NULL, y = NULL) +
                                 guides(fill=FALSE),
                               muci_fl_abund + 
                                 l_fitRaster() + 
                                 l_fitContour() + 
                                 labs(title = "muci", x = NULL, y = NULL) +
                                 guides(fill=FALSE),
                               pasc_fl_abund + 
                                 l_fitRaster() + 
                                 l_fitContour() + 
                                 labs(title = "pasc", x = NULL, y = NULL) +
                                 guides(fill=FALSE),
                               prat_fl_abund + 
                                 l_fitRaster() + 
                                 l_fitContour() + 
                                 labs(title = "prat", x = NULL, y = NULL) +
                                 guides(fill=FALSE),
                               psit_fl_abund + 
                                 l_fitRaster() + 
                                 l_fitContour() + 
                                 labs(title = "psit", x = NULL, y = NULL) +
                                 guides(fill=FALSE),
                               pyre_fl_abund + 
                                 l_fitRaster() + 
                                 l_fitContour() + 
                                 labs(title = "pyre", x = NULL, y = NULL) +
                                 guides(fill=FALSE),
                               soro_fl_abund + 
                                 l_fitRaster() + 
                                 l_fitContour() + 
                                 labs(title = "soro", x = NULL, y = NULL) +
                                 guides(fill=FALSE),
                               telu_fl_abund + 
                                 l_fitRaster() + 
                                 l_fitContour() + 
                                 labs(title = "telu", x = NULL, y = NULL) +
                                 guides(fill=FALSE),
                               wurf_fl_abund + 
                                 l_fitRaster() + 
                                 l_fitContour() + 
                                 labs(title = "wurf", x = NULL, y = NULL) +
                                 guides(fill=FALSE),
                               ncol = 4)
dev.off()
```

#### Bumbles
```{r message=FALSE}
# Zero-inflation may be an issue; a simple Poisson model is badly overdispersed
ggplot(bb_abund_sp, aes(bb.abund)) +
  geom_histogram() +
  facet_wrap(~bb.sp)

# Quasipoisson (winner)
bb_abund.qp <- bam(bb.abund ~ 
                     s(bb.sp.ord, bs = "re") + # center species
                     s(year, bs = "re") + # random intercept for year
                     s(site, bs = "re") + # random intercept to avoid pseudoreplicating elevation
                     te(yday, elev.mean, # tensor product smooth of elevation and time
                        by = bb.sp.ord, # by bb.sp
                        bs= c("gp", "tp"), # gaussian procees for time, thinplate spline for elevation
                        k=c(5, 5), 
                        m=2),
                   data = filter(bb_abund_sp, bb.sp != "humi"), # humi is just too rare
                   discrete = TRUE,
                   family = "quasipoisson",
                   select = TRUE,
                   method = "fREML") %>% getViz(nsim = 500)

check.gamViz(bb_abund.qp)
sum.qp <- summary(bb_abund.qp)
print(plot(bb_abund.qp, allTerms = TRUE), pages = 4)
check2D(bb_abund.qp, x1 = "yday", x2 = "elev.mean")

telu_abund <- plot(sm(bb_abund.qp, 4))
pasc_abund <- plot(sm(bb_abund.qp, 5))
prat_abund <- plot(sm(bb_abund.qp, 6))
soro_abund <- plot(sm(bb_abund.qp, 7))
wurf_abund <- plot(sm(bb_abund.qp, 8))
hort_abund <- plot(sm(bb_abund.qp, 9))
jone_abund <- plot(sm(bb_abund.qp, 10))
psit_abund <- plot(sm(bb_abund.qp, 11))
pyre_abund <- plot(sm(bb_abund.qp, 12))
mend_abund <- plot(sm(bb_abund.qp, 13))
mont_abund <- plot(sm(bb_abund.qp, 14))
gers_abund <- plot(sm(bb_abund.qp, 15))
muci_abund <- plot(sm(bb_abund.qp, 16))
hypn_abund <- plot(sm(bb_abund.qp, 17))
lapi_abund <- plot(sm(bb_abund.qp, 18))

pdf("../output/gam_BB.pdf")
grid_plot_fl_abun <- gridPrint(gers_abund + 
                                  l_fitRaster() + 
                                  l_fitContour() + 
                                  labs(title = "gers", x = NULL, y = NULL) +
                                  guides(fill=FALSE),
                               hort_abund + 
                                 l_fitRaster() + 
                                 l_fitContour() + 
                                 labs(title = "hort", x = NULL, y = NULL) +
                                 guides(fill=FALSE),
                               hypn_abund + 
                                 l_fitRaster() + 
                                 l_fitContour() + 
                                 labs(title = "hypn", x = NULL, y = NULL) +
                                 guides(fill=FALSE),
                               jone_abund + 
                                 l_fitRaster() + 
                                 l_fitContour() + 
                                 labs(title = "jone", x = NULL, y = NULL) +
                                 guides(fill=FALSE),
                               lapi_abund + 
                                 l_fitRaster() + 
                                 l_fitContour() + 
                                 labs(title = "lapi", x = NULL, y = NULL) +
                                 guides(fill=FALSE),
                               mend_abund + 
                                 l_fitRaster() + 
                                 l_fitContour() + 
                                 labs(title = "mend", x = NULL, y = NULL) +
                                 guides(fill=FALSE),
                               mont_abund + 
                                 l_fitRaster() + 
                                 l_fitContour() + 
                                 labs(title = "mont", x = NULL, y = NULL) +
                                 guides(fill=FALSE),
                               muci_abund + 
                                 l_fitRaster() + 
                                 l_fitContour() + 
                                 labs(title = "muci", x = NULL, y = NULL) +
                                 guides(fill=FALSE),
                               pasc_abund + 
                                 l_fitRaster() + 
                                 l_fitContour() + 
                                 labs(title = "pasc", x = NULL, y = NULL) +
                                 guides(fill=FALSE),
                               prat_abund + 
                                 l_fitRaster() + 
                                 l_fitContour() + 
                                 labs(title = "prat", x = NULL, y = NULL) +
                                 guides(fill=FALSE),
                               psit_abund + 
                                 l_fitRaster() + 
                                 l_fitContour() + 
                                 labs(title = "psit", x = NULL, y = NULL) +
                                 guides(fill=FALSE),
                               pyre_abund + 
                                 l_fitRaster() + 
                                 l_fitContour() + 
                                 labs(title = "pyre", x = NULL, y = NULL) +
                                 guides(fill=FALSE),
                               soro_abund + 
                                 l_fitRaster() + 
                                 l_fitContour() + 
                                 labs(title = "soro", x = NULL, y = NULL) +
                                 guides(fill=FALSE),
                               telu_abund + 
                                 l_fitRaster() + 
                                 l_fitContour() + 
                                 labs(title = "telu", x = NULL, y = NULL) +
                                 guides(fill=FALSE),
                               wurf_abund + 
                                 l_fitRaster() + 
                                 l_fitContour() + 
                                 labs(title = "wurf", x = NULL, y = NULL) +
                                 guides(fill=FALSE),
                               ncol = 4)
dev.off()
```

#### Functional bumbles
```{r}
bb_abund_sp_tongues <- bb_abund_sp %>%
  group_by(site, date, year, pbl.w.class2, elev.mean, yday) %>%
  summarize(bb.abund = sum(bb.abund)) %>%
  mutate(tongue = factor(pbl.w.class2)) %>%
  dplyr::select(-pbl.w.class2)

bb_abund_tongue.qp <- bam(bb.abund ~ 
                             s(bb.sp.ord, bs = "re") + # center species
                             s(tongue, bs = "re") + # center species
                             s(year, bs = "re") + # random intercept for year
                             s(site, bs = "re") + # random intercept to avoid pseudoreplicating elevation
                             te(yday, elev.mean, # tensor product smooth of elevation and time
                                by = tongue, # by bb.sp
                                bs= c("gp", "tp"), # gaussian procees for time, thinplate spline for elevation
                                k=c(8, 8), 
                                m=2),
                           data = bb_abund_sp, # humi is just too rare
                           discrete = TRUE,
                           family = "quasipoisson",
                           select = TRUE,
                           method = "fREML") %>% getViz()

check.gamViz(bb_abund_tongue.qp)
sum.tongue_qp <- summary(bb_abund_tongue.qp)
print(plot(bb_abund_tongue.qp, allTerms = TRUE), pages = 1)
plot(bb_abund_tongue.qp)
```

#### Occurrence clustering
```{r}
library(vegan)
library(ggvegan)
library(ggrepel)

# Unpooled data
bb_rel_abund <- bb_abund_sp %>%
  filter(bb.sp != "humi") %>%
  ungroup() %>%
  group_by(site, date, bb.sp) %>%
  summarize(bb.abund = sum(bb.abund)) %>%
  unite(id, c(site, date)) %>%
  pivot_wider(names_from = id, values_from = bb.abund) %>%
  column_to_rownames("bb.sp") %>%
  decostand(method = "hellinger")

dist <- vegdist(bb_rel_abund, method = "horn")
clust <- hclust(dist, method = "ward.D2")

pdf("../output/clust.pdf", height = 4.5, width = 9)
plot(clust)
dev.off()

pcoa <- capscale(bb_rel_abund ~ 1, method = "horn")
plot(pcoa, type = "t")

pcoa_mds1 <- round(pcoa$CA$eig[[1]] / sum(pcoa$CA$eig)*100)
pcoa_mds2 <- round(pcoa$CA$eig[[2]] / sum(pcoa$CA$eig)*100)

pcoa.fort <- fortify(pcoa)

ggplot(filter(pcoa.fort, Score == "sites"), aes(MDS1, MDS2)) +
  geom_point() +
  geom_text_repel(aes(label = Label), size = 6) +
  xlab(paste("MDS1 ", "(", pcoa_mds1, "%", ")", sep = "")) +
  ylab(paste("MDS2 ", "(", pcoa_mds2, "%", ")", sep = "")) +
  coord_fixed() +
  theme_gray(16)

ggsave("../output/pcoa_abund.pdf")

# Pooled data
bb_rel_abund2 <- bb_abund_sp %>%
  filter(bb.sp != "humi") %>%
  ungroup() %>%
  group_by(site, date, bb.sp) %>%
  summarize(bb.abund = sum(bb.abund)) %>%
  group_by(site, bb.sp) %>%
  summarize(bb.abund = mean(bb.abund)) %>%
  pivot_wider(names_from = site, values_from = bb.abund) %>%
  column_to_rownames("bb.sp") %>%
  decostand(method = "hellinger")

dist2 <- vegdist(bb_rel_abund2, method = "horn")
clust2 <- hclust(dist2, method = "ward.D2")

pdf("../output/clust2.pdf", height = 4.5, width = 9)
plot(clust2)
dev.off()

pcoa2 <- capscale(bb_rel_abund2 ~ 1, method = "horn")
plot(pcoa2, type = "t")

pcoa2_mds1 <- round(pcoa2$CA$eig[[1]] / sum(pcoa2$CA$eig)*100)
pcoa2_mds2 <- round(pcoa2$CA$eig[[2]] / sum(pcoa2$CA$eig)*100)

pcoa2.fort <- fortify(pcoa2)

ggplot(filter(pcoa2.fort, Score == "sites"), aes(MDS1, MDS2)) +
  geom_point() +
  geom_text_repel(aes(label = Label), size = 6) +
  xlab(paste("MDS1 ", "(", pcoa2_mds1, "%", ")", sep = "")) +
  ylab(paste("MDS2 ", "(", pcoa2_mds2, "%", ")", sep = "")) +
  coord_fixed() +
  theme_gray(16)

ggsave("../output/pcoa2_abund2.pdf")
```

### BB range and abundance
```{r}
species_names <- cbind(c("telu", "pasc", "prat", 
                         "soro", "wurf", "hort", 
                         "pyre", "mend", "mont", 
                         "gers", "hypn","jone", 
                         "lapi", "muci", "psit"),
                       c("B. terrestris/lucorum", "B. pascuorum", "B. pratorum", 
                         "B. soroensis", "B. wurflenii", "B. hortorum",
                         "B. pyrenaeus", "B. mendax", "B. monticola",
                         "B. gerstaeckeri", "B. hypnorum", "B. jonellus", 
                         "B. lapidarius", "B. mucidus", "B. psithyrus"),
                       c("B. telu.", "B. pasc.", "B. prat.", 
                         "B. soro.", "B. wurf.", "B. hort.",
                         "B. pyre.", "B. mend.", "B. mont.",
                         "B. gers.", "B. hypn.", "B. jone.", 
                         "B. lapi.", "B. muci.", "B. psit."),
                       c("major", "major", "major", 
                         "major", "major", "common", 
                         "high elev.", "high elev.", "high elev.",
                         "minor", "minor", "common", 
                         "minor", "high elev.", "common")
                       ) %>%
  as.data.frame() %>%
  as_tibble() %>%
  dplyr::select(bb.sp = 1, species.name = 2, species.abb = 3, abund.class = 4)

bb_range <- net %>%
  #filter(bb.sp != "humi") %>%
  left_join(site_data) %>%
  group_by(site, date, elev.mean, bb.sp) %>%
  summarize(abund = n()) %>%
  group_by(site, elev.mean, bb.sp) %>%
  summarize(abund = mean(abund)) %>%
  group_by(bb.sp) %>%
  mutate(elev.floor = min(elev.mean),
         elev.ceiling = max(elev.mean),
         elev.range = elev.ceiling - elev.floor,
         elev.med = elev.floor + ((elev.ceiling - elev.floor)/2))
# %>%
#   left_join(species_names) %>%
#   mutate(abund.class = factor(abund.class, levels = c("major", "common", "high elev.", "minor")))

ggplot(bb_range, aes(reorder(bb.sp, elev.med), elev.mean)) +
  geom_line(size = 2, alpha = 0.25) +
  geom_point(aes(size = abund), alpha = 0.5) +
  #facet_wrap(~year, ncol = 1) +
  theme_light(12) +
  scale_size_continuous(name = "Abundance") +
  #scale_color_discrete(name = "Abund. class") +
  xlab("BB species") +
  ylab("Elevation") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
ggsave("../output/bb_range.png", width = 6, height = 4, units = "in")
ggsave("../output/bb_range.pdf", width = 6, height = 4)
```