### 2021-03-03 code/bias_niche.Rmd
oops

---
title: "visitation_bias"
author: "Doug Sponsler"
date: "12/14/2020"
output: html_document
---
# Preparations
## Packages 
```{r, warning=FALSE, include=FALSE, echo = FALSE}
library(bipartite)
library(brms)
library(bayesplot)
library(tidybayes)
library(mgcv)
library(mgcViz)
library(DHARMa)
library(modelr)
library(ggdist)
library(vegan)
library(ggvegan)
library(ggrepel)
library(treemapify)
library(patchwork)
library(tidymodels)
library(tidyverse)
library(lubridate)

rstan::rstan_options(auto_write = TRUE)
options(mc.cores = parallel::detectCores())
```

## Load data
```{r message=FALSE, warning=FALSE, include=FALSE, echo = FALSE}
# Site data
site_data <- read_csv("../data/processed_data/site_data.csv") %>%
  dplyr::select(-c(slope.calc, slope.est, elev.min, elev.max, temp.mean)) %>% # drop these variables
  mutate(elev.class2 = factor(elev.class2, levels = c("low", "mid", "high")),
         site = factor(site)) # turn elav.class into an ordered factor and site into a regular factor

# BB traits
bb_traits <- read_csv("../data/processed_data/bb_traits.csv")

# FL traits
## family-level taxonomy
fl_tax <- read_csv("../data/processed_data/floral_tax.csv")

## Kugler morphotypes (accessed through Bioflor)
fl_ktype <- read_csv("../data/processed_data/floral_k_type.csv")

## Flower colors (accessed through BioFlor)
fl_color <- read_csv("../data/processed_data/floral_color.csv")

## diameter and corolla depth data from Gita Benadi
fl_morph <- read_csv("../data/processed_data/flower_morphology_GitaBenadi.csv") %>% 
  dplyr::select(SampleID, plant.family = Family, plant.sp = Species, 
                Unit, Flowers_per_unit, diam = Diameter, LN) %>%
  group_by(plant.sp) %>%
  summarize(diam = mean(diam), LN = mean(LN))

## verbal description of Kugler morphotypes
fl_ktype_key <- read_delim("../data/processed_data/kugler_key.tsv", delim = "\t", col_names = FALSE) %>%
  dplyr::select(k.type = 1, description = 2)

## join all floral trait tables
fl_traits <- fl_tax %>%
  full_join(fl_ktype) %>%
  full_join(fl_color) %>%
  full_join(fl_morph) %>%
  dplyr::select(plant.sp, plant.genus, plant.family, k.type, k.type.s, k.type.ss, color, diam, LN)
```

## Align survey and visitation data with pseudodates
```{r}
# Site-dates where BB visitation was recorded (excluding dates where only males were recorded)
visitation_samples <- read_csv("../data/processed_data/network.csv") %>%
  filter(caste != "male") %>%
  select(site, date) %>%
  distinct()

# Site dates where floral surveying was conducted
survey_samples <- read_csv("../data/processed_data/floral_survey.csv") %>%
  select(site, date) %>%
  distinct()

# Site-dates where visitation sampling and floral sampling coincided
shared_dates <- inner_join(visitation_samples, survey_samples)

# Site-dates where visitation was recorded WITHOUT corresponding survey data
visitation_only <- anti_join(visitation_samples, survey_samples) %>%
  mutate(date.v = date) %>%
  select(-date)

# This creates a pseudo date column for unmatched visitation site-dates that will match each visitation date
# to its nearest survey data. For site t4, there is a 2-year mismatch; we will drop these site-dates later.
pseudodates <- full_join(visitation_only, survey_samples) %>%
  mutate(mismatch = abs(date.v - date)) %>%
  group_by(site, date.v) %>%
  filter(mismatch == min(mismatch)) %>%
  rename(date = date.v, pseudo.date = date)
```

## Generate a data frame that represents visitation and survey abundance
```{r}
# BB visitation data
visitation <- read_csv("../data/processed_data/network.csv") %>%
  filter(caste != "male") %>% # we will ignore males 
  group_by(site, date, bb.sp, plant.sp) %>%
  summarize(visits = n()) %>% # visits per BB-FL pair
  ungroup() %>%
  left_join(pseudodates) %>% # Now we will replace unmatched dates with their pseudo dates
  replace_na(replace = list(mismatch = 0)) %>% # set mismatch to zero for perfectly aligned dates
  filter(mismatch <= 7) %>% # this gets rid of irreconcilable dates (like the t4 dates mentioned above)
  mutate(date2 = if_else(is.na(pseudo.date), date, pseudo.date)) %>% # replace date with pseudodate for all site-dates lacking perfect alignments
  select(-c(date, pseudo.date, mismatch)) %>%
  rename(date = date2) %>%
  group_by(site, date) %>%
  complete(bb.sp, plant.sp, fill = list(visits = 0)) # generate all possible BB-FL combos for each site date and fill in zeros for visits

# Floral cover data
flower_cover <- read_csv("../data/processed_data/floral_survey.csv") %>%
  filter(flower.cover > 0) %>% # ignore plants that were not blooming
  group_by(site, date, plant.sp) %>%
  summarize(flower.cover = sum(flower.cover))

# Get shared dates after adding pseudodates. This will let us filter out survey dates that don't
# have corresponding visitation sampling dates
shared_dates_with_pseudo <- inner_join(visitation, flower_cover, by = c("site", "date")) %>%
  select(site, date) %>%
  distinct()

# Join visitation and flower cover data

## First we need a list of all BB spp. by site and date
sp_by_site <- visitation %>% 
  ungroup() %>%
  select(site, date, bb.sp) %>%
  distinct()

## Then we need a list of unvisited plant.sp by site and date
unvisited <- flower_cover %>% 
  semi_join(shared_dates_with_pseudo) %>% # filter flower cover data to include only site-dates with visitation complement (real or pseudodates)
  full_join(visitation) %>% # join to visitation data
  filter(is.na(bb.sp)) %>% # select only the site-date-plants that were totally unvisited (and therefore have no BB spp to join to)
  select(-bb.sp) %>% # drop bb.sp variable because I just want the plants
  left_join(sp_by_site) %>% # join in the species list by site and date, thus populating the unvisited plants with all the BB spp recorded for each site-date
  replace_na(replace = list(visits = 0)) # replace NAs (created in the line above) with zero

## Now we can combine everything into a single data frame
visit_cover <- flower_cover %>% 
  semi_join(shared_dates_with_pseudo) %>% # filter flower cover data to include only site-dates with visitation complement (real or pseudodates)
  full_join(visitation) %>% # join to visitation data
  replace_na(replace = list(flower.cover = 0.01)) %>% # If a plant was visited but not surveyed, we will assign it the minimum non-zero flower cover value of 0.01.
  filter(!is.na(bb.sp)) %>% # drop the univisited plants
  bind_rows(unvisited) %>% # add them back, with BB spp and zeros
  select(site, date, bb.sp, plant.sp, visits, flower.cover) %>%
  group_by(site, date, bb.sp, plant.sp, flower.cover) %>% # In a few cases, the use of pseudodates causes two sampling period to converge on the same date, which creates apparent duplicate visitation data. We will handle this by averaging the visits.
  summarize(visits = mean(visits)) %>%
  left_join(fl_traits) %>% # add floral traits
  ungroup() %>%
  mutate(k.type.ss = factor(k.type.ss))

## Get total visits by BB sp and site-date; this will be a measure of BB abundance that we can use after we've calculated d'. I could, of course, work this into the biaseval function, but I'll put that off for now.
bb_abundance <- visit_cover %>%
  group_by(site, date, bb.sp) %>%
  summarize(bb.sp.abund = sum(visits)) %>%
  group_by(site, date) %>%
  mutate(bb.sp.prop.abund = bb.sp.abund/sum(bb.sp.abund))

fl_abundance <- visit_cover %>%
  group_by(site, date) %>%
  summarize(total.flower.cover = sum(flower.cover))
```

# 1. Overall bias

## 1.1 Calculate visitation probability per plant morphotype. Instead of keeping the plant species resolved, we will pool visitation by morphotype. The problem with not pooling by morphotype is that more species-rich morphotypes will be penalized simply because their visitation is diluted across more species.
```{r}
prop_visits_ktype <- visit_cover %>%
  group_by(site, date, bb.sp, k.type.ss) %>%
  summarize(flower.cover = sum(flower.cover), # get raw flower cover and visits by site*date*bb.sp*plant.sp
            visits = sum(visits)) %>%
  group_by(site, date) %>%
  mutate(prop.flower.cover = flower.cover/sum(flower.cover))%>% # binary visits
  group_by(site, date, bb.sp) %>%
  mutate(prop.visits = visits/sum(visits),
         bb.sp.abund = sum(visits)) %>% # abundance of each BB sp. per site-date
  ungroup() %>%
  left_join(bb_traits) %>% # add BB traits
  mutate(ktype = factor(k.type.ss), # recode variables
         tongue = factor(pbl.w.class2),
         visits = as.integer(visits)) %>% # so we can run a Poisson model
  left_join(site_data) %>% # add site data
  ungroup() %>%
  mutate(yday = yday(date),
         elev.mean0 = scale(elev.mean),
         yday0 = scale(yday),
         prop.flower.cover0 = scale(prop.flower.cover),
         perc.flower.cover = prop.flower.cover*100,
         bb.sp.abund0 = scale(bb.sp.abund)) %>%
  select(site, elev.mean, date, yday, bb.sp, ktype, visits, 
         prop.visits, flower.cover, perc.flower.cover, bb.sp.abund, tongue,
         elev.mean0, yday0, prop.flower.cover0, bb.sp.abund0) %>%
  mutate(visits.bin = if_else(visits > 0, 1, 0))

write_csv(prop_visits_ktype, "../hpc_packets/prop_visits_ktype.csv")

prop_visits_ktype_pooled <- visit_cover %>%
  group_by(site, date, k.type.ss) %>%
  summarize(flower.cover = sum(flower.cover), # get raw flower cover and visits by site*date*bb.sp*plant.sp
            visits = sum(visits)) %>%
  group_by(site, date) %>%
  mutate(prop.flower.cover = flower.cover/sum(flower.cover),
         prop.visits = visits/sum(visits),
         bb.abund = sum(visits),
         visits = as.integer(visits)) %>%
  mutate(ktype = factor(k.type.ss)) %>%
  left_join(site_data) %>% # add site data
  ungroup() %>%
  mutate(yday = yday(date),
         elev.mean0 = scale(elev.mean),
         yday0 = scale(yday),
         perc.flower.cover = prop.flower.cover*100,
         prop.flower.cover0 = scale(prop.flower.cover),
         bb.abund0 = scale(bb.abund)) %>%
  select(site, elev.mean, date, yday, ktype, visits, 
         prop.visits, bb.abund, perc.flower.cover,
         elev.mean0, yday0, prop.flower.cover0, bb.abund0) %>%
  mutate(visits.bin = if_else(visits > 0, 1, 0),
         year = factor(year(date)))

write_csv(prop_visits_ktype_pooled, "../hpc_packets/prop_visits_ktype_pool.csv")

```

## 1.2 Binomial multi-level regression models
```{r}
## Exploring the differential bias of bumble bee species and tongue length classes towards floral morphotypes
### Bernoulli with binarized visitation
brm_pvis00 <- brm(as.integer(visits.bin) ~
                    bb.sp.abund*perc.flower.cover + # fixed term: because more abundant flowers will tend to be more visited,and more abundance bees visit more
                    (1 + perc.flower.cover | ktype:bb.sp) + # a random intercept and random slope for ktype:species; I think I need to model this before I can make conclusions about ktype:pbl.class
                    (1 + perc.flower.cover | ktype:tongue) + # a random intercept and slope for ktype:tongue
                    (1 | ktype:site), # a varying intercept and slope for ktype:site
                  family = "bernoulli", 
                  file = "../output/brm_pvis_ktype_00",  # for running locally
                  #file = "brm_pvis_ktype_00", # for running on the HPC
                  chains = 4,
                  cores = 12,
                  iter = 2000,
                  data = filter(prop_visits_ktype, !ktype %in% c("0", "9", "10")))

brm_pvis01 <- brm(as.integer(visits.bin) ~
                    bb.abund*perc.flower.cover + # fixed term: because more abundant flowers will tend to be more visited,and more abundance bees visit more
                    (1 + perc.flower.cover | ktype) + # a random intercept and slope for ktype
                    (1 | ktype:site),  # a varying intercept term for ktype:site
                  family = "bernoulli", 
                  file = "../output/brm_pvis_ktype_01",  # for running locally
                  #file = "brm_pvis_ktype_01", # for running on the HPC
                  chains = 4,
                  cores = 12,
                  iter = 2000,
                  data = filter(prop_visits_ktype_pooled, !ktype %in% c("0", "9", "10")))

# If I want to make inference about species visitation rates, I don't need the tongue-length class effect; I just need the site effect. But, when I want to make inference about the tongue length classes, I do need the species-level effect.
brm_pvis02 <- brm(as.integer(visits.bin) ~
                    bb.sp.abund*perc.flower.cover + # fixed term: because more abundant flowers will tend to be more visited,and more abundance bees visit more
                    (1 + perc.flower.cover | ktype:bb.sp) + # a random intercept and random slope for ktype:species; I think I need to model this before I can make conclusions about ktype:pbl.class
                    (1 | ktype:site), # a varying intercept and slope for ktype:site
                  family = "bernoulli", 
                  file = "../output/brm_pvis_ktype_02",  # for running locally
                  #file = "brm_pvis_ktype_02", # for running on the HPC
                  chains = 4,
                  cores = 12,
                  iter = 5000, # Bulk ESS was a bit low wit 2000 iter
                  data = filter(prop_visits_ktype, !ktype %in% c("0", "9", "10")))
```

## 1.3 Evaluate models
```{r}
# Summary
summary(brm_pvis00)
summary(brm_pvis01)
summary(brm_pvis02)
summary(brm_pvis11)
summary(brm_pvis21)

# Posterior predictive checks
pp_check(brm_pvis00)
pp_check(brm_pvis01)
pp_check(brm_pvis02)
pp_check(brm_pvis11)
pp_check(brm_pvis21)

# Bayes R2
bayes_R2(brm_pvis00)
bayes_R2(brm_pvis01)
bayes_R2(brm_pvis02)
bayes_R2(brm_pvis11)
bayes_R2(brm_pvis21)

```

## 1.4 Preliminary visualization
```{r}
get_variables(brm_pvis00)
conditional_effects(brm_pvis00)
mcmc_plot(brm_pvis00, pars = c("^r_ktype:tongue.*Intercept]$"))
mcmc_plot(brm_pvis00, pars = c("^r_ktype:tongue.*perc.flower.cover]$"))
mcmc_plot(brm_pvis00, pars = c("^r_ktype:bb.sp.*Intercept]$"))
mcmc_plot(brm_pvis00, pars = c("^r_ktype:bb.sp.*perc.flower.cover]$"))
mcmc_plot(brm_pvis00, pars = c("^r_ktype:site.*Intercept]$"))
mcmc_plot(brm_pvis00, pars = c("^r_ktype:site.*perc.flower.cover]$"))

get_variables(brm_pvis01)
conditional_effects(brm_pvis01)
mcmc_plot(brm_pvis01, pars = c("^r_ktype"))
mcmc_plot(brm_pvis01, pars = c("^r_ktype:tongue.*perc.flower.cover]$"))
mcmc_plot(brm_pvis01, pars = c("^r_ktype:bb.sp.*Intercept]$"))
mcmc_plot(brm_pvis01, pars = c("^r_ktype:site.*Intercept]$"))
mcmc_plot(brm_pvis01, pars = c("^r_ktype:bb.sp.*perc.flower.cover]$"))
mcmc_plot(brm_pvis01, pars = c("^r_ktype:site.*Intercept]$"))
mcmc_plot(brm_pvis01, pars = c("^r_ktype:site.*perc.flower.cover]$"))

get_variables(brm_pvis02)
conditional_effects(brm_pvis02)
mcmc_plot(brm_pvis02, pars = c("^r_ktype"))
mcmc_plot(brm_pvis02, pars = c("^r_ktype:bb.sp.*Intercept]$"))
mcmc_plot(brm_pvis01, pars = c("^r_ktype:site.*Intercept]$"))
mcmc_plot(brm_pvis02, pars = c("^r_ktype:bb.sp.*perc.flower.cover]$"))
mcmc_plot(brm_pvis02, pars = c("^r_ktype:site.*Intercept]$"))
mcmc_plot(brm_pvis02, pars = c("^r_ktype:site.*perc.flower.cover]$"))

get_variables(brm_pvis11)
conditional_effects(brm_pvis11)
mcmc_plot(brm_pvis11, pars = c("^r_ktype:year"))

get_variables(brm_pvis21)
conditional_effects(brm_pvis21)
mcmc_plot(brm_pvis21, pars = c("^r_ktype"))
```

## 1.5 Final visualization
Note that the intercept is on the log scale, so if you want to transform it back to predicted # visits, you have to exponentiate it. I think. Here is a helplful conversation involving the esteemed Ben Bolker:
### 1.5.1 morphotype:tongue intercept
```{r}
brm_pvis00_plotframe1 <- brm_pvis00 %>%
  spread_draws(b_Intercept, `r_ktype:tongue`[condition, term]) %>% # the back ticks are necessary to escape the :
  rename(r_ktype.tongue = `r_ktype:tongue`) %>% # get rid of the problematic :
  filter(term == "Intercept") %>% # drop slope terms for now
  mutate(condition_mean = b_Intercept + r_ktype.tongue) %>% # get conditional mean
  group_by(condition) %>% # group by tongue:morpotype pair
  point_interval(condition_mean, .width = c(0.95, 0.66)) %>% # calculate credible intervals
  mutate(condition_mean_trans = exp(condition_mean), # exponentiating the estimates gets us back to the response scale (negbinomial uses log link function)
         .lower_trans = exp(.lower),
         .upper_trans = exp(.upper)) %>% 
separate(condition, c("ktype", "tongue"), sep = "_", remove = FALSE) %>% # separate condition term into ktype and tongue
  mutate(ktype = case_when( # rename
    ktype == "1" ~ "Disc (1)",
    ktype == "2" ~ "Funnel (2)",
    ktype == "3" ~ "Bell (3)",
    ktype == "4" ~ "Stalk-disc (4)",
    ktype == "5" ~ "Lip (5)",
    ktype == "6" ~ "Flag (6)",
    ktype == "7" ~ "Head (7)")
  ) %>%
  mutate(ktype = factor(ktype, levels = c("Disc (1)", "Funnel (2)", # relevel
                                          "Bell (3)", "Stalk-disc (4)", "Lip (5)", 
                                          "Flag (6)", "Head (7)")))

ggplot(brm_pvis00_plotframe1, aes(condition_mean_trans, tongue, 
                                 xmin = .lower_trans, xmax = .upper_trans,
                                 color = tongue, fill = tongue)) +
  geom_pointinterval() +
  facet_wrap(~ktype, scales = "free") +
  theme_light() +
  labs(color = "Tongue length", x = "Visitation probability") +
  guides(fill = FALSE) +
  theme(axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank())

ggsave(file = "../output/brm_pvis00_plot1.pdf", width = 6, height = 5)
```

### 1.5.2 morphotype:bb.sp intercept
```{r}
brm_pvis02_plotframe2 <- brm_pvis02 %>%
  spread_draws(b_Intercept, `r_ktype:bb.sp`[condition, term]) %>% # the back ticks are necessary to escape the :
  rename(r_ktype.bb.sp = `r_ktype:bb.sp`) %>% # get rid of the problematic :
  filter(term == "Intercept") %>% # drop slope terms for now
  mutate(condition_mean = b_Intercept + r_ktype.bb.sp) %>% # get conditional mean
  group_by(condition) %>% # group by bb.sp:morpotype pair
  point_interval(condition_mean, .width = c(0.95, 0.66)) %>% # calculate credible intervals
  mutate(condition_mean_trans = inv_logit_scaled(condition_mean), # exponentiating the estimates gets us back to the response scale (negbinomial uses log link function)
         .lower_trans = inv_logit_scaled(.lower),
         .upper_trans = inv_logit_scaled(.upper)) %>% 
separate(condition, c("ktype", "bb.sp"), sep = "_", remove = FALSE) %>% # separate condition term into ktype and bb.sp
  mutate(ktype = case_when( # rename
    ktype == "1" ~ "Disc (1)",
    ktype == "2" ~ "Funnel (2)",
    ktype == "3" ~ "Bell (3)",
    ktype == "4" ~ "Stalk-disc (4)",
    ktype == "5" ~ "Lip (5)",
    ktype == "6" ~ "Flag (6)",
    ktype == "7" ~ "Head (7)")
  ) %>%
   mutate(ktype = factor(ktype, levels = c("Disc (1)", "Funnel (2)", # relevel
                                          "Bell (3)", "Stalk-disc (4)", "Lip (5)", 
                                          "Flag (6)", "Head (7)"))) %>%
  left_join(bb_traits) %>%
  mutate(bb.sp = factor(bb.sp, levels = c("telu", "soro", "pyre", "psit", "prat", "lapi", "jone",
                                           "wurf", "pasc", "muci", "mont", "hypn",
                                           "mend", "hort", "gers")))

ggplot(brm_pvis02_plotframe2, aes(condition_mean_trans, bb.sp, 
                                 xmin = .lower_trans, xmax = .upper_trans,
                                 color = pbl.w.class2, fill = pbl.w.class2)) +
  geom_pointinterval() +
  facet_wrap(~ktype, scales = "free") +
  theme_light() +
  labs(color = "Tongue length", x = "Visitation probability", y = "Morphotype") +
  guides(fill = FALSE)

ggsave(file = "../output/brm_pvis02_plot1.pdf", width = 8, height = 8)
```

### 1.5.3 morphotype:tongue slope
```{r}
brm_pvis00_plotframe3 <- brm_pvis00 %>%
  spread_draws(b_perc.flower.cover, `r_ktype:tongue`[condition, term]) %>% # the back ticks are necessary to escape the :
  rename(r_ktype.tongue = `r_ktype:tongue`) %>%
  filter(term == "perc.flower.cover") %>%
  mutate(condition_slope = b_perc.flower.cover + r_ktype.tongue) %>%
  separate(condition, c("ktype", "tongue"), sep = "_", remove = FALSE) %>%
  mutate(ktype = case_when(
    ktype == "1" ~ "Disc (1)",
    ktype == "2" ~ "Funnel (2)",
    ktype == "3" ~ "Bell (3)",
    ktype == "4" ~ "Stalk-disc (4)",
    ktype == "5" ~ "Lip (5)",
    ktype == "6" ~ "Flag (6)",
    ktype == "7" ~ "Head (7)")
  ) %>%
  mutate(ktype = factor(ktype,
                        levels = c("Disc (1)", "Funnel (2)", "Bell (3)", "Stalk-disc (4)",
                                   "Lip (5)", "Flag (6)", "Head (7)")))

ggplot(brm_pvis00_plotframe3, aes(condition_slope, tongue, color = tongue)) +
  stat_pointinterval() +
  facet_grid(ktype~.) +
  theme_light() +
  labs(color = "Tongue length", x = "Slope of logit-likelihood") +
  guides(fill = FALSE) +
  theme(axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank())

ggsave(file = "../output/brm_pvis00_plot3.png", width = 7, height = 5.5)
```

### 1.5.3 morphotype intercept
```{r}
brm_pvis01_plotframe1 <- brm_pvis01 %>%
  spread_draws(b_Intercept, r_ktype[condition, term]) %>% # the back ticks are necessary to escape the :
  filter(term == "Intercept") %>% # drop slope terms for now
  mutate(condition_mean = b_Intercept + r_ktype) %>% # get conditional mean
  group_by(condition) %>% # group by tongue:morpotype pair
  point_interval(condition_mean, .width = c(0.95, 0.66)) %>% # calculate credible intervals
  mutate(condition_mean_trans = inv_logit_scaled(condition_mean), # exponentiating the estimates gets us back to the response scale (negbinomial uses log link function)
         .lower_trans = inv_logit_scaled(.lower),
         .upper_trans = inv_logit_scaled(.upper)) %>% 
  mutate(ktype = case_when( # rename
    condition == "1" ~ "Disc (1)",
    condition == "2" ~ "Funnel (2)",
    condition == "3" ~ "Bell (3)",
    condition == "4" ~ "Stalk-disc (4)",
    condition == "5" ~ "Lip (5)",
    condition == "6" ~ "Flag (6)",
    condition == "7" ~ "Head (7)")
  ) %>%
  mutate(ktype = factor(ktype, levels = c("Head (7)", "Flag (6)", # relevel
                                          "Lip (5)", "Stalk-disc (4)", "Bell (3)", 
                                          "Funnel (2)", "Disc (1)")))

ggplot(brm_pvis01_plotframe1, aes(condition_mean_trans, ktype, 
                                 xmin = .lower_trans, xmax = .upper_trans)) +
  geom_pointinterval() +
  theme_light(14) +
  labs(x = "Visitation probability", y= "Morphotype") +
  guides(fill = FALSE)

ggsave(file = "../output/brm_pvis01_plot1.pdf", width = 5, height = 5)

```

### 1.5.3 morphotype slope
```{r}
brm_pvis01_plotframe2 <- brm_pvis01 %>%
  spread_draws(b_perc.flower.cover, r_ktype[condition, term]) %>% # the back ticks are necessary to escape the :
  filter(term == "perc.flower.cover") %>%
  mutate(condition_slope = b_perc.flower.cover + r_ktype) %>%
  mutate(ktype = case_when(
    condition == "0" ~ "NA (0)",
    condition == "1" ~ "Disc (1)",
    condition == "2" ~ "Funnel (2)",
    condition == "3" ~ "Bell (3)",
    condition == "4" ~ "Stalk-disc (4)",
    condition == "5" ~ "Lip (5)",
    condition == "6" ~ "Flag (6)",
    condition == "7" ~ "Head (7)")
  ) %>%
  mutate(ktype = factor(ktype,
                        levels = c("Disc (1)", "Funnel (2)", "Bell (3)", "Stalk-disc (4)",
                                   "Lip (5)", "Flag (6)", "Head (7)", "NA (0)")))

ggplot(brm_pvis01_plotframe2, aes(condition_slope, ktype)) +
  stat_pointinterval() +
  theme_light() +
  labs(color = "Tongue length", x = "Slope of visitation probability ~ percent flower cover") +
  guides(fill = FALSE)

ggsave(file = "../output/brm_pvis01_plot2.png", width = 7, height = 5.5)
```

## 1.6 Model 1 hypothesis tests
I'm not doing every possible comparison. Just the ones where it is possible to have some doubt. It would be stupid to compare group whose 95% CIs are miles apart.
```{r}
get_variables(brm_pvis01)

hypothesis(brm_pvis01, "r_ktype[0,Intercept] = r_ktype[1,Intercept]", class = "") # this is lower than its closest neighbor; no need to make any other comparisons

hypothesis(brm_pvis01, "r_ktype[1,Intercept] = r_ktype[4,Intercept]", class = "") 
hypothesis(brm_pvis01, "r_ktype[2,Intercept] = r_ktype[4,Intercept]", class = "")

hypothesis(brm_pvis01, "r_ktype[2,Intercept] = r_ktype[1,Intercept]", class = "")
hypothesis(brm_pvis01, "r_ktype[1,Intercept] = r_ktype[3,Intercept]", class = "") 
hypothesis(brm_pvis01, "r_ktype[2,Intercept] = r_ktype[3,Intercept]", class = "")
hypothesis(brm_pvis01, "r_ktype[4,Intercept] = r_ktype[3,Intercept]", class = "")

hypothesis(brm_pvis01, "r_ktype[5,Intercept] = r_ktype[6,Intercept]", class = "") 
hypothesis(brm_pvis01, "r_ktype[5,Intercept] = r_ktype[7,Intercept]", class = "") 
```

## 1.7 Model 2 hypothesis tests
```{r}
get_variables(brm_pvis00)

# Disc (group  1)
hypothesis(brm_pvis00, "r_ktype:tongue[1_short,Intercept] = r_ktype:tongue[1_med,Intercept]", class = "")
hypothesis(brm_pvis00, "r_ktype:tongue[1_short,Intercept] = r_ktype:tongue[1_long,Intercept]", class = "")
hypothesis(brm_pvis00, "r_ktype:tongue[1_med,Intercept] = r_ktype:tongue[1_long,Intercept]", class = "")

# Funnel (group  2)
hypothesis(brm_pvis00, "r_ktype:tongue[2_short,Intercept] = r_ktype:tongue[2_med,Intercept]", class = "")
hypothesis(brm_pvis00, "r_ktype:tongue[2_short,Intercept] = r_ktype:tongue[2_long,Intercept]", class = "")
hypothesis(brm_pvis00, "r_ktype:tongue[2_med,Intercept] = r_ktype:tongue[2_long,Intercept]", class = "")

# Bell (group  3)
hypothesis(brm_pvis00, "r_ktype:tongue[3_short,Intercept] = r_ktype:tongue[3_med,Intercept]", class = "")
hypothesis(brm_pvis00, "r_ktype:tongue[3_short,Intercept] = r_ktype:tongue[3_long,Intercept]", class = "")
hypothesis(brm_pvis00, "r_ktype:tongue[3_med,Intercept] = r_ktype:tongue[3_long,Intercept]", class = "")

# Stalk-disc (group  4)
hypothesis(brm_pvis00, "r_ktype:tongue[4_short,Intercept] = r_ktype:tongue[4_med,Intercept]", class = "")
hypothesis(brm_pvis00, "r_ktype:tongue[4_short,Intercept] = r_ktype:tongue[4_long,Intercept]", class = "")
hypothesis(brm_pvis00, "r_ktype:tongue[4_med,Intercept] = r_ktype:tongue[4_long,Intercept]", class = "")

# Lip (group  5)
hypothesis(brm_pvis00, "r_ktype:tongue[5_short,Intercept] = r_ktype:tongue[5_med,Intercept]", class = "")
hypothesis(brm_pvis00, "r_ktype:tongue[5_short,Intercept] = r_ktype:tongue[5_long,Intercept]", class = "")
hypothesis(brm_pvis00, "r_ktype:tongue[5_med,Intercept] = r_ktype:tongue[5_long,Intercept]", class = "")

# Flag (group  6)
hypothesis(brm_pvis00, "r_ktype:tongue[6_short,Intercept] = r_ktype:tongue[6_med,Intercept]", class = "")
hypothesis(brm_pvis00, "r_ktype:tongue[6_short,Intercept] = r_ktype:tongue[6_long,Intercept]", class = "")
hypothesis(brm_pvis00, "r_ktype:tongue[6_med,Intercept] = r_ktype:tongue[6_long,Intercept]", class = "")

# Head (group  7)
hypothesis(brm_pvis00, "r_ktype:tongue[7_short,Intercept] = r_ktype:tongue[7_med,Intercept]", class = "")
hypothesis(brm_pvis00, "r_ktype:tongue[7_short,Intercept] = r_ktype:tongue[7_long,Intercept]", class = "")
hypothesis(brm_pvis00, "r_ktype:tongue[7_med,Intercept] = r_ktype:tongue[7_long,Intercept]", class = "")

# NA (group  0)
hypothesis(brm_pvis00, "r_ktype:tongue[0_short,Intercept] = r_ktype:tongue[0_med,Intercept]", class = "")
hypothesis(brm_pvis00, "r_ktype:tongue[0_short,Intercept] = r_ktype:tongue[0_long,Intercept]", class = "")
hypothesis(brm_pvis00, "r_ktype:tongue[0_med,Intercept] = r_ktype:tongue[0_long,Intercept]", class = "")
```

# 2. Ordination of dietary niche
Aggregation changes meaning. Don't forget that. When I pool visitation across dates, I am introducing the effects of relative abundance and floral phenology. Species in bloom when BB abundance is high will be over-represented, and vice-versa. And species with a long bloom time will be over-represented, and vice-versa. This isn't necessarily a problem. I just need to make sure I interpret the results the right way. When I pool data like this, I am comparing the "net" diet of species, which includes the effects of abundance and phenology in addition to the effects of floral preferences. I think that is appropriate for this stage of the analysis, but I need to be careful not to assume that what I'm doing here is the same sort of analysis as what I did above without pooling samples. 
```{r}
# Fl traits with abbreviated names
fl_traits2 <- fl_traits %>%
  left_join(read_csv("../data/processed_data/network.csv")) %>%
  select(plant.sp, plant.sp.abb, plant.genus, plant.family, k.type.ss, color) %>%
  mutate(k.type.ss = factor(k.type.ss)) %>%
  distinct()

# Totally aggregated; note I'm using the full data, not the aligned and trimmed data. No need for complementary floral survey data.
visitation2 <- read_csv("../data/processed_data/network.csv") %>%
  filter(caste != "male") %>% # we will ignore males 
  filter(!bb.sp %in% c("humi")) %>% # humi is too rare
  group_by(bb.sp, plant.sp.abb) %>%
  summarize(visits = n()) %>%
  group_by(bb.sp) %>%
  mutate(prop.visits = visits/sum(visits)) %>%
  select(-visits) %>%
  pivot_wider(names_from = plant.sp.abb, values_from = prop.visits, values_fill = 0) %>%
  column_to_rownames("bb.sp")

bb_traits2 <- bb_traits %>%
  filter(bb.sp != "humi") %>%
  select(bb.sp, pbl.w.class2) %>%
  arrange(bb.sp) %>%
  mutate(pbl.w.class2.pool = case_when(
    pbl.w.class2 == "short" ~ "short",
    pbl.w.class2 == "med" ~ "med/long",
    pbl.w.class2 == "long" ~ "med/long"
  ))

# Generate ordination
pcoa_horn2 <- capscale(visitation2 ~ 1, distance = "horn")
pcoa_horn2_adonis <- adonis(visitation2 ~ pbl.w.class2, data = bb_traits2)
pcoa_horn2_betadisper <- betadisper(vegdist(visitation2, method = "horn"), 
                                    group = bb_traits2$pbl.w.class2,
                                    type = "median")
anova(pcoa_horn2_betadisper) # I think this is not significant because short tongued species are disperse on axis 2, but axis 2 does not explain nearly as much variation.
plot(pcoa_horn2_betadisper)
# I'm suprised that the betadisper test is not significant, but I think it might be driven by the fact that replication is so low for long and medium tongue species. Let's see what happens if we pool long and medium tongue species, since they overlap so much anyway.
pcoa_horn2_betadisper_pool <- betadisper(vegdist(visitation2, method = "horn"), 
                                    group = bb_traits2$pbl.w.class2.pool,
                                    type = "median")
anova(pcoa_horn2_betadisper_pool) # I think this is not significant because short tongued species are disperse on axis 2, but axis 2 does not explain nearly as much variation.
plot(pcoa_horn2_betadisper_pool)

# Get variance explained
pcoa_horn2_mds1 <- round(pcoa_horn2$CA$eig[[1]] / sum(pcoa_horn2$CA$eig)*100)
pcoa_horn2_mds2 <- round(pcoa_horn2$CA$eig[[2]] / sum(pcoa_horn2$CA$eig)*100)

# Fortify plants
pcoa_horn2.fort1 <- fortify(pcoa_horn2, axes = 1:2) %>%
  filter(Score == "species") %>%
  rename(plant.sp.abb = Label) %>%
  left_join(fl_traits2)

# Fortify bees
pcoa_horn2.fort2 <- fortify(pcoa_horn2, axes = 1:2) %>%
  filter(Score == "sites") %>%
  rename(bb.sp = Label) %>%
  left_join(bb_traits) %>%
  mutate(pbl.w.class2 = factor(pbl.w.class2))

# Generate convex hulls around tongue length classes
pcoa_horn2_hulls <- pcoa_horn2.fort2 %>%
  group_by(pbl.w.class2) %>%
  slice(chull(MDS1, MDS2))

# Plot
ggplot(pcoa_horn2.fort2, aes(MDS1, MDS2, color = pbl.w.class2)) +
  geom_polygon(data = pcoa_horn2_hulls, 
               aes(fill = pbl.w.class2, color = NULL),
               alpha = 0.25) +
  geom_text(aes(label = bb.sp), key_glyph = "rect") +
  geom_text_repel(data = filter(pcoa_horn2.fort1, abs(MDS1) >= 0.2 | abs(MDS2) >= 0.2), 
            aes(MDS1, MDS2, label = plant.sp.abb), inherit.aes = FALSE, color = "gray40") +
  geom_point(data = filter(pcoa_horn2.fort1, abs(MDS1) >= 0.2 | abs(MDS2) >= 0.2),
            aes(MDS1, MDS2, shape = k.type.ss), inherit.aes = FALSE, color = "gray40", size = 3) +
  scale_shape_discrete() +
  xlab(paste("MDS1 ", "(", pcoa_horn2_mds1, "%", ")", sep = "")) +
  ylab(paste("MDS2 ", "(", pcoa_horn2_mds2, "%", ")", sep = "")) +
  labs(fill = "Tongue length", 
       shape = "Morphotype") +
  guides(color = FALSE) +
  coord_fixed() +
  theme_light(10)

ggsave("../output/pcoa_aggregate.png")

# a quick peak at diet breadth...quick, I promise
taxon_diet_breadth <- read_csv("../data/processed_data/network.csv") %>%
  filter(caste != "male") %>% # we will ignore males 
  filter(!bb.sp %in% c("humi")) %>% # humi is too rare
  group_by(site, date, bb.sp, plant.sp) %>%
  summarize(visits = n()) %>%
  group_by(site, date, bb.sp) %>%
  summarize(bb.sp.abund = sum(visits),
            tax.diet.breadth = length(plant.sp)) %>%
  mutate(log.bb.sp.abund = log(bb.sp.abund),
         prop.tax.diet.breadth = tax.diet.breadth/bb.sp.abund)

morph_diet_breadth <- read_csv("../data/processed_data/network.csv") %>%
  filter(caste != "male") %>% # we will ignore males 
  filter(!bb.sp %in% c("humi")) %>% # humi is too rare
  left_join(fl_traits) %>%
  group_by(site, date, bb.sp, k.type.ss) %>%
  summarize(visits = n()) %>%
  group_by(site, date, bb.sp) %>%
  summarize(bb.sp.abund = sum(visits),
            morph.diet.breadth = length(k.type.ss)) %>%
  mutate(log.bb.sp.abund = log(bb.sp.abund),
         prop.morph.diet.breadth = morph.diet.breadth/bb.sp.abund)

diet_breadth <- full_join(taxon_diet_breadth, morph_diet_breadth) %>%
  left_join(site_data)

ggplot(diet_breadth, aes(tax.diet.breadth, morph.diet.breadth, color = bb.sp)) +
  geom_point() +
  geom_smooth(method = "lm")

ggplot(diet_breadth, aes(morph.diet.breadth, tax.diet.breadth, color = bb.sp)) +
  geom_point() +
  geom_smooth(method = "lm")

ggplot(diet_breadth, aes(elev.mean, tax.diet.breadth)) +
  geom_point() +
  geom_smooth() +
  facet_wrap(~bb.sp)

ggplot(diet_breadth, aes(elev.mean, morph.diet.breadth)) +
  geom_point() +
  geom_smooth() +
  facet_wrap(~bb.sp)
```

# 3. Dietary niche overlap (i.e. potential competition)

## 3.1 Check BB distribution patterns to decide whom to retain in the analysis
```{r}
bb_distribution <- visitation %>%
  mutate(year = year(date)) %>%
  left_join(site_data) %>%
  group_by(site, date, elev.mean, bb.sp, year) %>%
  summarize(abund = n()) %>%
  group_by(site, elev.mean, bb.sp, year) %>%
  summarize(abund = mean(abund)) %>%
  group_by(bb.sp) %>%
  mutate(elev.floor = min(elev.mean),
         elev.ceiling = max(elev.mean),
         elev.range = elev.ceiling - elev.floor,
         elev.med = elev.floor + ((elev.ceiling - elev.floor)/2))

ggplot(bb_distribution, aes(reorder(bb.sp, elev.med), elev.mean)) +
  geom_line(size = 2, alpha = 0.25) +
  geom_point(aes(size = abund), alpha = 0.5) +
  facet_wrap(~year, ncol = 1) +
  theme_light(12) +
  scale_size_continuous(name = "Abundance") +
  xlab("BB species") +
  ylab("Elevation") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

## 3.2 Prep data for calculating niche overlap
```{r}
niche_data <- visitation %>%
  #filter(!bb.sp %in% c("humi", "muci", "gers")) %>%
  left_join(fl_traits) %>%
  mutate(year = factor(year(date)))
```

## 3.3 Function for calculating niche overlap
```{r}
niche_overlapper <- function(x, criterion) {

  criterion <- enquo(criterion) # criterion is set to either plant.sp (for taxonomic) or k.type.s (for morphological)

  x %>%
    #group_by(bb.sp, !!criterion, site, date, year) %>% # group by site-year and BB-FL pair
    group_by(bb.sp, !!criterion, site, date) %>% # group by site-year and BB-FL pair
    summarize(freq = sum(visits)) %>% # get interaction frequency for each BB-FL pair per site-year
    group_by(bb.sp, site, date) %>% # group by site-year and BB species
    mutate(prop = freq/sum(freq)) %>% # get proportional interaction frequency for each BB-FL pair
    #group_by(bb.sp, site, year, !!criterion) %>%
    group_by(bb.sp, site, !!criterion) %>%
    summarize(mean.prop = mean(prop)) %>%
    pivot_wider(names_from = !!criterion, values_from = mean.prop, values_fill = 0) %>% # spread into the kind of array vegan likes
    #group_by(site, year) %>%
    group_by(site) %>%
    nest() %>% # nest into list-column
    mutate(data2 = map(data, function(x) column_to_rownames(x, var = "bb.sp")), # column to rownames
           dist = map(data2, function(x) vegdist(x, method = "horn")), # convert to Horn's distance matrix
           tidy_dist = map(dist, tidy)) %>% # convert to tidy array of Horn's distances
    #dplyr::select(site, year, tidy_dist) %>% # drop intermediate list-columns
    dplyr::select(site, tidy_dist) %>% # drop intermediate list-columns
    unnest(cols = c(tidy_dist)) %>% # unnest back into a single data frame
    left_join(site_data) %>% # add site data
    # dplyr::select(site, year, bb.sp1 = item1, bb.sp2 = item2, distance, elev.mean) %>% # dplyr::select and rename vars
    dplyr::select(site, bb.sp1 = item1, bb.sp2 = item2, distance, elev.mean) %>% # dplyr::select and rename vars
    left_join(bb_traits, by = c("bb.sp1" = "bb.sp")) %>% # add traits for bb.sp1
    rename(subgenus1 = subgenus, tongue1 = pbl.w.class2) %>%
    left_join(bb_traits, by = c("bb.sp2" = "bb.sp")) %>% # add traits for bb.sp2
    rename(subgenus2 = subgenus, tongue2 = pbl.w.class2) %>%
    # dplyr::select(site, year, bb.sp1, bb.sp2, subgenus1, subgenus2, distance, 
    #        tongue1, tongue2, elev.mean) %>%
    dplyr::select(site, bb.sp1, bb.sp2, subgenus1, subgenus2, distance, 
       tongue1, tongue2, elev.mean) %>%
    mutate(tongue.diff = as.logical(if_else(tongue1 == tongue2, FALSE, TRUE)),
           tongue.pair = factor(case_when(
             tongue1 == "med" & tongue2 == "med" ~ "med:med",
             tongue1 == "med" & tongue2 == "long" ~ "med:long",
             tongue1 == "med" & tongue2 == "short" ~ "short:med",
             tongue1 == "long" & tongue2 == "med" ~ "med:long",
             tongue1 == "short" & tongue2 == "med" ~ "short:med",
             tongue1 == "long" & tongue2 == "short" ~ "short:long",
             tongue1 == "short" & tongue2 == "long" ~ "short:long",
             tongue1 == "long" & tongue2 == "long" ~ "long:long",
             tongue1 == "short" & tongue2 == "short" ~ "short:short",
           )),
           subgenus.diff = if_else(subgenus1 == subgenus2, "same", "different")) %>%
    unite(sp.pair, c(bb.sp1, bb.sp2), sep = ":", remove = FALSE) %>% # create a single species pair column...
    mutate(sp.pair = factor(sp.pair)) %>% # ...and make it a factor so that mgcv can handle it
    mutate(niche.overlap = 1-distance) %>% # express niche overlap so that 1 = perfect overlap and 0 = no overlap
    select(-distance)
}
```

## 3.4 Calculate niche overlap pooled within site
```{r}
# Pooled within site
niche_overlap_gen <- niche_overlapper(niche_data, plant.genus) %>%
  mutate(level = rep("genus", n()),
         niche.overlap.genus = niche.overlap)

niche_overlap_fam <- niche_overlapper(niche_data, plant.family) %>%
  mutate(level = rep("family", n()),
         niche.overlap.family = niche.overlap)

niche_overlap_ktype <- niche_overlapper(niche_data, k.type.ss) %>%
  mutate(level = rep("morphotype", n()),
         niche.overlap.ktype = niche.overlap)

niche_pool <- bind_rows(niche_overlap_gen,
                        niche_overlap_fam, 
                        niche_overlap_ktype,) %>%
  mutate(level = factor(level, levels = c("morphotype", "family", "genus")),
         tongue.diff = tongue.diff,
         site = factor(site))

write_csv(niche_pool, "../hpc_packets/niche_pool.csv")

morph_genus_difference <- niche_overlap_gen %>%
  select(-niche.overlap, -level) %>%
  full_join(niche_overlap_fam) %>%
  select(-niche.overlap, -level) %>%
  full_join(niche_overlap_ktype) %>%
  select(-niche.overlap, -level) %>%
  mutate(tongue.diff = tongue.diff,
         site = factor(site),
         morph.genus.diff = niche.overlap.ktype - niche.overlap.genus)

ggplot(morph_genus_difference, aes(elev.mean, morph.genus.diff)) +
  geom_point() +
  geom_smooth() +
  facet_wrap(~tongue.pair)

ggplot(morph_genus_difference, aes(tongue.pair, morph.genus.diff)) +
  geom_boxplot()
```

## 3.5 Niche overlap ~ level
```{r}
brm_niche00 <- brm(niche.overlap ~ 1 +
                     level + # fixed effect of level 
                     (1 | site) + # varying intercept for site
                     (1 | sp.pair), # varying intercept for sp.pair
                   family = "zero_one_inflated_beta", # for prop data with 0s and 1s
                   file = "../output/brm_niche_00", # local
                   #file = "./brm_niche_00", # HPC
                   iter = 2000,
                   chains = 4,
                   data = niche_pool)

pp_check(brm_niche00)
bayes_R2(brm_niche00)

get_variables(brm_niche00)
conditional_effects(brm_niche00)
mcmc_plot(brm_niche00, pars = c("^r_sp.pair\\[mend*"))
mcmc_plot(brm_niche00, pars = c("^r_site*"))
```

## 3.6 Genus niche overlap ~ tongue
```{r}
brm_niche10 <- brm(niche.overlap ~ 1 +
                     (1 | site) + # varying intercept for site
                     (1 | tongue.pair) +
                     (1 | sp.pair), # varying intercept for sp.pair
                   family = "zero_one_inflated_beta", # for prop data with 0s and 1s
                   file = "../output/brm_niche_10", # local
                   #file = "./brm_niche_10", # HPC
                   iter = 2000,
                   chains = 4,
                   data = filter(niche_pool, level == "genus"))

pp_check(brm_niche10)
bayes_R2(brm_niche10)

get_variables(brm_niche10)
conditional_effects(brm_niche10)
mcmc_plot(brm_niche10, pars = c("^r_tongue*"))
mcmc_plot(brm_niche10, pars = c("^r_site*"))
```

## 3.7 Genus niche overlap ~ sp.pair
```{r}
brm_niche20 <- brm(niche.overlap ~ 1 +
                     (1 | site) + # varying intercept for site
                     (1 | sp.pair), # varying intercept for sp.pair
                   family = "zero_one_inflated_beta", # for prop data with 0s and 1s
                   file = "../output/brm_niche_20", # local
                   #file = "./brm_niche_20", # HPC
                   iter = 2000,
                   chains = 4,
                   data = filter(niche_pool, level == "genus"))

pp_check(brm_niche20)
bayes_R2(brm_niche20)

get_variables(brm_niche20)
conditional_effects(brm_niche20)
mcmc_plot(brm_niche20, pars = c("^r_sp.pair"))
mcmc_plot(brm_niche20, pars = c("^r_site*"))
```

## 3.8 Niche overlap ~ elevation
```{r}
# Beta models
gen_gam_niche.Ib <- bam(niche.overlap ~
                            s(sp.pair, bs = "re") +
                            s(elev.mean, by = sp.pair, k = 6, m = 2),
                          method = "fREML",
                          family = betar(eps = 0.001),
                          discrete = TRUE,
                          select = TRUE,
                          data = filter(niche_pool, level == "genus")) %>% getViz()

check.gamViz(gen_gam_niche.Ib)
summary(gen_gam_niche.Ib)
print(plot.gamViz(gen_gam_niche.Ib, allTerms = TRUE), pages = 1)

gen_gam_niche.GI <- bam(niche.overlap ~
                            s(sp.pair, bs = "re") +
                            s(elev.mean, k = 6, m = 2) +
                            s(elev.mean, by = sp.pair, k = 6, m = 1),
                          method = "fREML",
                          family = betar(eps = 0.001),
                          discrete = TRUE,
                          select = TRUE,
                          data = filter(niche_pool, level == "genus")) %>% getViz()

check.gamViz(gen_gam_niche.GI)
summary(gen_gam_niche.GI)
print(plot.gamViz(gen_gam_niche.GI, allTerms = TRUE), pages = 1)

brm_niche01 <- brm(niche.overlap ~ 1 +
                     s(elev.mean, by = sp.pair, k = 10) +
                     (1 | sp.pair),
                   family = "zero_one_inflated_beta",
                   #file = "../output/brm_niche_01",
                   file = "./brm_niche_01",
                   iter = 2000,
                   chains = 4,
                   data = filter(niche_pool, level == "genus"))
```