### 2021-03-09 ms_2/ms_2_v2.Rmd
did some writing and added some references

---
title: "Potential competition in alpine bumble bees is mitigated by imperfect trait-matching and subpartitioning within floral morphotypes"

author: Douglas B. Sponsler^1^*, Fabrice Requier^2^, Katharina Kallnik^1^, Alice Claßen^1^, A. Fabienne Maihoff^1^, Ingolf Steffan-Dewenter^1^

# date: `r Sys.Date()`

output:
  bookdown::word_document2:
    fig_caption: yes
    fig_width: 7
    fig_height: 7
    df_print: kable
    keep_md: yes
    number_sections: no
    toc: no
    reference_docx: Word_template.docx

bibliography: 
  - references.bib
#  - knitcitations.bib
csl: ecology.csl
# csl: https://www.zotero.org/styles/ecology

---

^1^ Department of Animal Ecology and Tropical Biology, Biocenter, University of Würzburg, Würzburg, Germany

^2^ Université Paris-Saclay, CNRS, IRD, UMR Évolution, Génomes, Comportement et Écologie, 91198, Gif-sur-Yvette, France

`*` Corresponding author: douglas.sponsler@uni-wuerzburg.de

# ABSTRACT

Species coexistence is constrained in part by competition over shared food sources. Bumble bees are classic model organisms for the study of competition and coexistence, and this topic has taken on new urgency as climate-induced range shifts are expected to disrupt competitive relationships between bumble bee species, particularly along the climatic gradients of the mountain slopes that host the world's richest bumble bee communities. We surveyed floral abundance and recorded bumble bee floral visitation patterns along a ~1400 m elevation gradient in German Alps. Using these data, we first ask how floral preferences compare across bumble bee species and whether these patterns can be explained by trait-matching between bumble bee tongue length and floral morphotype. We then investigate the patterns of diet overlap between bumble bee species at the nested levels of floral morphotype, floral family, and floral genus, and model diet overlap as a function bumble bee tongue length and elevation. We find that while tongue length is a strong predictor of bumble bees' floral morphotype preferences, this pattern is far from perfect, and species with similar tongue lengths exhibit divergent preferences with respect to certain flora morphotypes. This indicates that factors other than tongue length are involved in flower choice, which might explain the puzzling coexistence of bumble bee species with similar tongue lengths. Moreover, genus-level diet overlap was approximately one half that of morphotype-level diet overlap, indicating that even when bumble bees exhibit similar biases at the level of floral morphotype, partly due to the constraints of tongue length, floral diversity enables the mitigation of competition via the genus- or species-level subpartitioning of floral resources within morphotypes. This suggests that floral species richness may be as important as overall floral abundance in avoiding the competitive exclusion of bumble bee species. Diet overlap was highest at the low end of our elevation gradient, declining sharply over the transition from valley floor to mid-slope, then leveling off or declining weakly for the remainder of the gradient. If diet overlap is interpreted as an index of potential competition, upslope shifts in bumble bees and flora could generate more competitive conditions at high elevations if and when floral resources become limiting.

*Keywords*: foraging, trait matching, pollinator, *Bombus*, alpine

```{r setup, include=FALSE, cache=FALSE, message = FALSE}

library("knitr")

### Chunk options: see http://yihui.name/knitr/options/ ###
## Text results
opts_chunk$set(echo = FALSE, warning = TRUE, message = TRUE)

## Code decoration
opts_chunk$set(tidy = TRUE, comment = NA, highlight = TRUE)

## Cache
# opts_chunk$set(cache = 2, cache.path = "output/cache/")

## Plots
opts_chunk$set(fig.path = "../output/")

```

```{r knitcitations, cache = FALSE}
library(knitcitations)
cleanbib()   
cite_options(citation_format = "pandoc")
```

# INTRODUCTION

The role of competition in regulating community assembly and species coexistence has long been a central question in ecology. It has taken on new urgency, however, on account of recent climate change, which threatens to disrupt existing patterns of species interactions [@Tylianakis2008-fi], including competition [@Ettinger2017-pk], with consequences for community composition and function. These effects are likely to be most pronounced along the steep climate gradients of mountain slopes, where rapid climate-induced range shifts are expected to introduce novel competitive interactions within trophic levels and cause spatial or phenological mismatches between organisms and their food resources.

Bumble bees (Hymenoptera: *Bombus* spp.) have become classic model organisms for the study of competition and species coexistence due to their broad and overlapping distributions and their gross similarity in morphology, behavior, and dietary niche [@Plowright1984-lc; @Goulson2008-zy]. In this role, however, they have proven somewhat enigmatic. On the one hand, there is evidence that resource limitation can be a significant constraint on local bumble communities [@Heinrich1976-wc; @Bowers1985-df], and certain bumble bee communities have been shown to be over-dispersed in terms of tongue length (a matching trait corresponding to floral corolla depth) or phylogeny (understood to encompass unmeasured trait variation), suggesting that competition functions as a community assembly process [@Inouye1977-kl; @Pellissier2013-pk]. Nevertheless, bumble bee communities with an apparent excess of species richness and deficit of trait (or phylogenetic) dispersion are commonly observed [@Ranta1981-vx; @Goulson2008-zy]. To account for this inconsistency, @Ranta1981-vx proposed that bumble bees communities are governed by a spatiotemporally heterogeneous mosaic of competitive conditions, such that competitive inequalities between species are sufficiently variable to allow more species to coexist than predicted by trait-based competition models. A complementary explanation, based on the core/satellite species hypothesis [@Hanski1982-hs], is that competitive forces are significant only among regionally abundant and widespread “core species”, and the effects of competition are more pronounced in the structure of regional species pools than in that of local assemblages [@Hanski1982-ie]. In some cases, competition may simply be precluded by a saturating abundance of floral resources [@Williams1989-vr].

Teasing apart the still-obscure mechanisms of bumble bee coexistence and niche partitioning remains a salient problem from a basic ecological perspective, but it has also become an urgent question in the context of applied conservation. Bumble bees are cold-adapted species [@Heinrich1994-cv] associated mainly with mountain habitats, where they serve as the principal pollinators of many alpine plants [@Williams1998-xi; @Goulson2010-lo]. Warming climates have already caused measurable upslope shifts in bumble bees and their floral hosts [@Ploquin2013-da; @Kerr2015-dg; @Pyke2016-ll; @Fourcade2019-ct; @Soroye2020-zb; @Marshall2020-pc], and these are predicted to cause declines in many bumble bees species [@Soroye2020-zb]. Moreover, the introduction or extinction of species from bumble bee communities, with corresponding addition and release of competitive pressure, can alter both the floral visitation patterns of bumble bees and, consequently, the pollination biology of their host plants [@Brosi2013-mu; @Ishii2013-sm]. Thus, the speculative question of how competition has acted in the historic assembly of bumble bee communities has become an unfolding process observable over ecological time scales and instrumental in the future viability of bumble bee species and their floral mutualists.

All else held equal, species coexistence is favored when intraspecific competition exceeds interspecific competition, and this stabilizing inequality is favored to the extent that species differ in dietary niche [@Abrams1977-zp; @Chesson2000-aw]. Bumble bees are exceptionally versatile foragers whose ability to solve the puzzles posed by complex flower morphologies [Plowright1984-lc], or indeed by human investigators [@Alem2016-rn; @Loukola2017-bt], renders almost any flower potentially accessible. Nevertheless, the flexibility of bumble bee foraging is tuned to floral reward, which depends on both the intrinsic reward of a given flower and the costs associated with exploiting it. These costs are closely related to the degree of morphological fittedness between bumble bees and their floral hosts, which consists principally in the matching of bumble bee tongue length to floral corolla depth. Bumble bee species can differ markedly in tongue length, and floral handling time increases both for short-tongued bumble bees foraging on deep flowers and long-tongued bees foraging on shallow flowers [@Inouye1980-mv], which leads to an association between tongue length and floral niche in bumble bee communities [e.g. @Inouye1978-qi; @Inouye1980-mv]. This association between tongue length and floral morphology, though, is imperfect [e.g. @Ranta1980-ok; @Goulson2008-oe], perhaps in part because access to floral rewards in some flowers involves learning to manipulate floral structures that are challenging irrespective of tongue length [@Laverty1994-hm].

Empirical measurements of niche overlap are, however, inherently ambiguous with respect to competition. It has been well-demonstrated that niche overlap can never serve as a metric of competition *per se* [@Abrams1980-am], but it is often used as an index of *potential* competition, particularly in the context of dietary studies [e.g. @French1985-eo; @Steffan-Dewenter2000-ll; @Jung2015-cw], based on the reasoning that if and when food becomes limiting, competition can be expected to vary in proportion to the dietary overlap of co-occurring species [@Pianka1972-oi]. Focusing on *potential* competition does not, however, escape the equivocity of niche overlap, since an observed pattern of niche overlap could just as plausibly be interpreted as adaptive niche partitioning in response to *realized* competition [@Pianka1974-qk]. This creates a predicament in which niche overlap can only be interpreted as an index of potential competition when realized competition --- which cannot be inferred from niche overlap --- is assumed to be negligible, and this assumption can only be justified on the basis of additional knowledge of the study system under consideration.

Our study is set in the Berchtesgadener Alps of southeast Germany, a landscape of flower-rich mountain meadows --- generated by past or ongoing grazing management --- embedded in a matrix of coniferous forest. This system is one of rich floral resources but harsh abiotic conditions, so it is likely that competition for floral resources is low, or at least transient, and that bumble bees are limited instead by abiotic conditions during the critical period of spring colony founding. On this basis, we argue for the interpretation of niche overlap as an index of *potential* competition rather than a response to *realized* competition, with the caveats this interpretation may misrepresent certain temporal intervals (e.g. early spring colony founding) or elevational bands (e.g. the sparse alpine grasslands above the tree line ecotone) [Sponsler et al., submitted] and that the residual effects of past competitive pressure on innate floral biases [@Connell1980-dn] are beyond the scope of our inference.

The aim of our study is to synthesize the phenomena of floral preference, trait-matching, and dietary niche overlap in a way that accounts for the coexistence of bumble bee species with similar tongue lengths and predicts the effects of climate-driven elevational range shifts on the future of potential competition between bumble bee species in mountain habitats. We first ask how floral preferences compare across bumble bee species and to what extent these patterns can be explained by trait-matching between bumble bee tongue length and floral morphotype. We then investigate the patterns of dietary niche overlap between bumble bee species that emerge from their respective floral preferences, comparing niche overlap at the nested levels of floral morphotype, floral family, and floral genus. Finally, we ask how niche overlap varies with elevation and discuss the implications of this variation for bumble bee conservation in a changing climate.

# METHODS

## Field system

<!-- I guess I'll need to tweak the wording of this section so that it is not a verbatim copy of the other manuscript. But it's okay if it recycles a lot of language because it is, after all, the same thing being described. -->

<!-- Describe general floristic and temperature patterns. Describe sampling in more detail. -->

Field sampling was conducted 2010-2012 in Berchtesgaden National Park (47.55°N, 12.92°E), located in the Northern Limestone Alps of southeast Germany. The landscape is composed of mountain pastures in a coniferous forest matrix. We selected 25 study sites (60 x 60 m) on pastures at elevations ranging from 641-2032 m above sea level. Fourteen of these pastures are extensively grazed by cattle or sheep, three are mowed for hay production, and eight have been abandoned throughout the last century and are no longer subject to any human management.

Sampling consisted of repeated visits (hereafter, “visitation samples”) to each study site at approximately weekly intervals. Samples were only collected on dry days when the air temperature was at least 6°C. During each visit, bumble bees and their activity were recorded during a 50-minute transect walk. Bumblebees observed on or a given flower were counted as floral visitors. Bumble bee queens were identified to species level in the field, whereas workers and males were stored in individually labeled tubes in the freezer for later identification in the laboratory after Amiet (1996). While floral visitation by male bumble bees was recorded during visitation sampling, we chose to analyze only visitation by queens and workers. During observation, it was noted whether each bee was visibly carrying pollen at the time of its visit, but this provides at best an imperfect distinction between pollen and nectar foragers, and we opted not to differentiate between bees with and without pollen in our analyses. In conjunction with visitation observations, we estimated the flower cover of each herbaceous or shrubby plant species within each 60 x 60 m study plot to the nearest 0.1 m^2^. Species identification followed @Lauber2007-ha and @Oberndorfer2001-ro.

While per-sample effort was standardized across all years, the total number of samples (i.e. site-dates) varied markedly,  with 153 in 2010, 212 in 2011, and 539 in 2012. The increased total sampling effort in 2012 was driven in part by the overall scarcity of bumble bees that year. Since the start of field sampling each year was determined by snowmelt, lower sites were sampled as early as April while higher sites could not be sampled until June or July. This resulted in a temporally staggered pattern of sampling in 2010 and 2012, while in 2011 a combination of early snowmelt and a late start to field activities resulted in sampling that was approximately aligned in time across the whole elevation gradient. See Appendix X for details of sampling patterns.

## Data sources

### Floral morphology

Rather than considering various floral morphological traits independently, we relied on the synthesis of @Kugler1970-gb, who developed a typology consisting of ten primary morphotypes and nested subdivision of each. Kugler’s typology was informed by laboratory studies of bumble bee flower choice [@Kugler1943-uy], and there is a precedent for analyzing bumble bee visitation in terms of Kugler’s morphotypes [@Neumayer1999-il; @Schneller2014-ry].

Kugler morphotype classifications of the plants in our study system were accessed via the BIOFLOR database [@Klotz2002-uc]. Morphotypes were simplified to the primary classes numbered 0-10: 0 = not applicable (includes wind-pollinated flora such as grasses), 1 = disc- and bowl-shaped flowers, 2 = funnel flowers, 3 = bell-shaped flowers, 4 = stalk-disc flowers, 5 = lip flowers, 6 = flag flowers, 7 = flower heads, 8 = spike flowers, 9 = brush flowers, 10 = trap flowers **(Figure 1)**. In a few cases, a plant species in our data set was not found in the BIOFLOR database. In most cases, morphotypes are constant at the genus level, and missing morphotypes were added based on the morphotypes found for congeners. The genus *Gentiana* consisted of both group 2 and group 4 representatives, so the missing species *Gentiana aspera* and *Gentiana ciliata* were classified based on visual comparison with their classified congeners. In the case of the genus *Phyteuma*, we followed @Neumayer1999-il and classified it as a group 7 flower head. Morphotypes 9 and 10 were extremely rare in our study system and were therefore omitted from all analyses. Plants belonging to morphotype 0 were typically not recorded during floral surveys due to their lack of conspicuous floral displays, and they were rarely visited by bumble bees, so this morphotype was also omitted from all analyses.

### Bumble bee tongue length

Bumble bee tongue-length data were obtained from the literature values compiled and summarized by @Arbetman2017-xj. Values for *B. gerstaeckeri* and *B. mendax*, which were not included in @Arbetman2017-xj, were obtained from @Obeso1992-yp and @Durieux2000-uk, respectively. We then binned bumble bee species into three tongue-length classes: short (< 7.5 mm), medium (7.5—10.0 mm), and long (>= 10.0 mm). Short-tongued species included *B. jonellus*, *B. lapidarius*, *B. pratorum*, *B. psithyrus*-group, *B. pyrenaeus*, and *B. terrestris/lucorum*. Medium-tongued species included *B. humilis*, *B. hypnorum*, *B. monticola*, *B. mucidus*, *B. pascuorum*, and *B. wurflenii*. Long-tongued species included *B. gerstaeckeri*, *B. hortorum*, and *B. mendax*.

## Data analysis

### Software

All analyses were conducted in `R` `r citep(citation())`. Data handling and visualization was performed with the `tidyverse` suite of R packages [@Wickham2019-aa]. Bayesian multilevel models were implemented in Stan [@Stan2020aa] via the package `brms` [@Burkner2017-aa] using default priors, and visualizations were prepared with the packages `tidybayes` `r citep(citation("tidybayes"))` and `ggdist` `r citep(citation("ggdist"))`. Ordination analysis and permutation testing were performed with the package `vegan` `r citep(citation("vegan"))` and visualized using the packages `ggplot2` [@Wickham2016-aa] and `ggvegan` [@Simpson2019aa]. Hierarchical generalized additive models (HGAMs) were implemented with the package `mgcv` [@Wood_2017] and visualized with the packages `mgcViz` [@Fasiolo2018aa] and `tidymv` [@Coretta-aa2021]. All R code used in this analysis, together with full model output and diagnostics, is available in Appendix X.

### Data processing

Prior to analysis, we aligned each visitation sample with a corresponding floral survey sample. In most cases, floral surveying was conducted jointly with bumble bee observation for each site and date. Occasionally, though, floral surveying could not be completed on the same day as bumble bee observation, resulting in a small temporal offset between visitation samples and floral survey samples. In these cases, visitation samples were paired with the nearest available floral survey data within a maximum distance of 7 days. Visitation samples lacking corresponding floral survey data within 7 days were omitted from further analysis, which resulted in dropping 40 site-dates out of a total of 895, leaving 855 site-dates included in our analyses. In a few cases, two consecutive visitation samples at a given site shared the same nearest floral survey sample; in these cases, the visitation samples were first pooled, then aligned to their shared nearest floral survey sample. In rare cases, a given floral species was recorded during visitation sampling but missed during floral surveying. These species were retroactively assigned the minimum floral area value of 0.01 m^2^.

We defined floral preference as the relative rate of visitation between floral morphotypes after conditioning on the neutral effects of abundance (see model descriptions below for details). Visitation data were extremely right-skewed because the majority of available flora received zero visits during any given sampling period. To facilitate model fitting, we created a binary version of visitation rate in which visited flora were assigned to 1 and unvisited flora to 0. We then modeled binary visitation using generalized models with binomial or Bernoulli distribution families. The conditional effects inferred from these models must, therefore, be interpreted as effects on visitation probability rather than visitation frequency.

We calculated niche overlap with the goal of focusing on elevational patterns, sacrificing temporal resolution to pool observations within site. First, we calculated each bumble bee species' proportional visitation frequency at the nested levels of floral morphotype, plant family, and plant genus. Then, we calculated, for each bumble bee species, the mean proportional visitation rate to each morphotype/family/genus across all sampling dates within each site. Using these species-level visitation matrices, we calculated pairwise niche overlap as 1 - Horn-Morisita distance (calculated separately for morphotype-, family-, and genus-level matrices) for all species pairs within each site. This approach was chosen for its normalizing effect on the highly variable "snapshot" visitation data generated by our individual transect walks, but it has the disadvantage of conflating niche overlap due to similar floral preference and niche overlap due to similar temporal (i.e. seasonal) activity patterns. We opted to analyze niche overlap only down to the level of floral genus in order to dampen the possible effects of false non-detection of floral species due to undersampling of the diets of rare bumble bee species [@Williams2005-yp].

While our study spanned three years, we chose not to model year effects. Differences in sampling intensity and duration across years (described above) would make comparisons across years difficult to interpret. This decision entails the caveats that our analysis cannot detect year-to-year variation in floral biases and niche overlap, and main effects cannot be distinguished from annual anomalies. With only three years of data, though, such annual inferences would be of little value, and we deemed them not worth the cost of added model complexity.

<!-- State more explicityl how the models should be interpreted with respect to bias -->

<!-- Mor edescriptive fgure captions -->

<!-- Describe the differences in models in the Figure2 and 3 captions -->

<!-- Open each results paragraph with a summary sentence. -->

### Model 1: Comparing overall visitation probability across floral morphotypes

We first compared overall visitation probability across floral morphotypes. In this model, visitation for each sample was pooled across bumble bee species and within floral morphotypes prior to binarization. Visitation probability was then modeled using multilevel Bayesian regression with a Bernoulli distribution family (logit link function). *Bumble bee abundance* and *percent flower cover* (per morphotype) were included as a global interaction term, thus conditioning visitation probability on abundance and enabling the inference of preference independent of the neutral effects of abundance. *Floral morphotype* was specified as a both a varying intercept term and a varying slope term with respect to *percent flower cover*. *Site* was included as a varying intercept term to avoid pseudoreplication owing to the repeated observations within each site.

### Model 2: Visitation probability as a function of floral morphotype and bumble bee tongue length

To explore the functional underpinnings of bumble foraging preference, we analyzed visitation probability as a function of the interaction between floral morphology and bumble bee tongue length class. For this analysis, visitation was pooled within floral morphotype but not across bumble bee species. We then specified a Bayesian model of the same form as Model 1, but this time included a varying slope and intercept term for floral morphotype : bumble bee tongue length class and floral morphotype : bumble bee species. In this model, the relationship between floral morphotype and bumble bee tongue length class was the focus of inference, and the bumble bee species effect was added to account for the grouping of observations within species.

### Model 3: Visitation probability as a function of floral morphotype and bumble bee species

To focus on species-level morphotype preferences, we removed the tongue length effect from Model 2 and re-fit the model considering only bumble bee species.

### Model 4: Aggregate diet ordination

To visualize the overall diet similarity between bumble bee species, we pooled visitation data across all samples within bumble bee species and plant genus, then performed a principal coordinates analysis based on Horn-Morisita distance metric. To test the significance of tongue length class as a predictor of diet similarity, we performed a permutation test. Importantly, this approach treats the whole study system in aggregate, thus conflating the causal effects of floral preference, seasonal activity patterns, and site occupancy patterns in the generation of overall diet similarity. In this sense, it can be understood as a model of potential diet overlap given the aggregate foraging patterns of the bumble bee species in our system, with the understanding that realized niche overlap at particular locations or times may vary.

### Model 5: Niche overlap by level and elevation

To investigate elevational patterns of *within-site* niche overlap, we modeled community-mean niche overlap as a function of level (morphotype, family, genus) and elevation using HGAMs. Community mean niche overlap was calculated by taking the mean niche overlap (within level) across all species pairs for each site. Level was included as both a fixed intercept effect and a smoothing factor in conjunction with elevation, following the type "I" model form described by @Pedersen2019-kf. To account for the proportional nature of our data, we used a beta distribution family with a logit link function.

### Model 6: Genus-level niche overlap by tongue length and elevation

To focus on genus-level niche overlap as a function of tongue length class, we generated a binary tongue-difference variable for each species pair: "same" for species belonging to the same tongue length class and "different" for species belonging to different tongue length classes. We then took the mean genus-level niche overlap across all species pairs within tongue-difference class and site. Then, we modeled genus-level niche overlap using tongue-difference as both a fixed intercept effect and a smoothing factor in conjunction with elevation, again following the type "I" model form @Pedersen2019-kf.

# RESULTS

## Summary of visitation and floral survey data

We recorded a total of 12,918 bumble- bee-flower interactions (excluding males) over the three years of our study. The metaweb across all sites and dates consisted of 16 bumble bee species (with *Bombus terrestric/lucorum* and *Bombus psithyrus* species groups lumped into species groups), 163 plant species (110 genera, 37 families), and 736 unique bumble- bee-plant interaction pairs. Five species --- *B. pascuorum*, *B. pratorum*, *B. soroeensis*, *B. terrestris-lucorum*, and *B. wurflenii* --- accounted for the bulk of overall bumble bee abundance and were present over the entire elevation range of our sites. *B. hortorum*, *B. jonellus*, and *B. psithyrus* (species-group) were widespread but at lower abundance. *B. monticola* occurred at moderate abundance but was rare or absent at sites below 1000 m. The remaining species were relatively rare in our study system. *B. mendax*, *B. mucidus*, and *B. pyrenaeus* occurred primarily above 1500 m, consistent with their known affinity for high elevation habitats [@Rasmont2010-ci]. *B. gerstaeckeri*, *B. hypnorum*, *B. lapidarius* were sparsely distributed across mid-elevation sites. *B. humilis* was recorded only three times and only in one year, so we omitted it from all analyses.

Floral surveying yielded a total of 352 plant species, representing 191 genera and 52 families. Of these, 155 species from 103 genera and 35 families were observed to be visited by bumble bees. Eight species --- *Rubus idaeus*, *Rosa canina*, *Juniperus communis*, *Larix decidua*, *Salix* sp., *Caltha palustris*, *Pulmonaria officinalis*, and *Rheum barbarum* --- were recorded in visitation data but not recorded during floral surveying. Each accounted for no more than 4 visits in total over the three years of our study, and they were omitted from the analysis of floral abundance.

## Floral preference (Models 1-3)

Our visitation models (Models 1-3) revealed strong morphotype preferences across the aggregate bumble bee community and marked differences in floral preference between bumble bee species. These differences in preference were partly explained by differences in tongue length, but certain floral morphotypes generated divergent preference patterns among bumble bee species with similar tongue lengths.

Lip (group 5) and flag (group 6) flowers had the highest overall visitation probability, with a conditional mean visitation probabilities of 0.35 (95% CI = 0.27–0.4) and 0.31 (95% CI = 0.24–0.40), respectively (Figure 1). In second place, head flowers (group 7) had a mean conditional visitation probability of 0.18 (95% CI = 0.13–0.25). Bell (group 3) flowers were a distant third with a conditional mean visitation probability of 0.09 (95% CI = 0.06–0.13), followed by stalk-disc (group 4), funnel (group 2), and disc flowers (group 1) with conditional visitation probabilities of 0.06 (95% CI = 0.04–0.08), 0.04 (95% CI = 0.03–0.07), and 0.04 (95% CI = 0.02–0.06), respectively. Model 1 explained approximately 37% of the total variance (97.5% CI = 35–38%).

```{r fig1, echo = FALSE,  fig.cap = paste("Posterior conditional mean visitation probability by floral morphotype. Point-interval plots depict 66\\% and 95\\% CIs. Each floral morphotype is exemplified by a representative species common in our study system. Illustrations are modifications of public domain works obtained from www.plantillustrations.org. For details of original sources, see Appendix X.")}
knitr::include_graphics("../output/brm_pvis01_plot1_mod.png")
```

Disc (group 1) flowers were visited at similar rates across tongue length classes **(Figure 2)**. At the species level, disc flowers were visited at relatively high rates by the short-tongued *B. terrestris/lucorum* and the medium-tongued *Bombus monticola* **(Figure 3)**. Visitation to funnel (group 2) flowers was dominated by long-tongued species, an effect driven entirely by the *Aconitum* (group 2) specialist *B. gerstaeckeri*. Bell (group 3) flowers were visited at a higher rate by short-tongued species than by long-tongued species, with medium-tongued species exhibiting intermediate visitation rates. At the species-level, bell flowers were visited at highest probability by the short-tongued species *B. jonellus*, *B. pyrenaeus*, and *B. soroeensis*, and by the medium-tongued species *B. hypnorum* and *B. monticola*. Stalk-disc (group 4) flowers were visited at similar rates across all tongue length classes, and there was also broad overlap in visitation rates at the species level. Lip (group 5) flowers were also visited at similar rates across tongue length classes, but at the species level *B. pascuorum* exhibited an uniquely high bias toward lip flowers. Flag (group 6) flowers were visited at a higher rate by medium-tongued species than by short-tongued species, with long-tongued species exhibiting intermediate visitation rates. Flag flowers generated strong separation in visitation rate at the species level, receiving very high visitation from the medium-tongued *B. pascuorum* and *B. mucidus*, moderately high visitation from the short-tongued *B. terrestris/lucorum* and *B. lapidarius*, the medium-tongued *B. monticola*, and the long-tongued *B. mendax* and *B. hortorum*, and low visitation by the short-tongued *B. soroeensis*, *B. pyrenaeus*, *B. psithyrus*, *B. pratorum*, and *B. jonellus*, the medium-tongued *B. hypnorum*, and the long-tongued *B. gerstaeckeri*. Models 2 and 3 both explained approximately 31% of total variance (97.5% CI = 30–32%).

```{r fig3, echo = FALSE,  fig.cap = paste("Posterior conditional mean visitation probability by Kugler morphotype and bumble bee tongue length class. Point-interval plots depict 66\\% and 95\\% CIs. Letters indicate group differences based on evidence ratio at alpha = 0.05. Note that the range of the x-axis has been rescaled for each panel to emphasize the comparison of visitation probabilities across tongue length classes.")}
knitr::include_graphics("../output/brm_pvis00_plot1.png")
```

```{r fig4, echo = FALSE,  fig.cap = paste("Posterior conditional mean visitation probability by Kugler morphotype and bumble bee species. The tongue length class of each species is color-coded. Point-interval plots depict 66\\% and 95\\% CIs. Note that the range of the x-axis has been rescaled for each panel to emphasize the comparison of visitation probabilities across bumble bee species")}
knitr::include_graphics("../output/brm_pvis02_plot1.png")
```

## Aggregate diet similarity

Principal coordinates analysis (Model 4) suggested weak but significant effects of tongue length on the diet similarity while also highlighting the particular floral genera that accounted most strongly for the dietary separation of bumble bee species. Axis 1 accounted for 24% of total diet variance and primarily separated long- and medium-tongued species from short-tongued species (Figure 4). Long- and medium-tongued species scored low on Axis one and were associated with the lip flowers (group 5) *Ajuga*, *Rhinanthus*, and *Stachys* and flag flowers (group 6) *Anthyllis*, *Lotus*, and *Trifolium*. Short tongued species scored high on Axis one and were associated with the bell flowers (group 3) *Campanula*, *Erica*, and *Geum* and the head flowers (group 7) *Carduus*, *Centaurea*, and *Phyteuma*. Axis 2 accounted for 19% of total variance and primarily explained variation within short-tongued species. Short-tongued species scoring high on Axis 2 were associated with *Campanula* while those scoring low on Axis 2 were associated with *Erica*. Short-tongued species scoring close to zero on Axis 2 also scored lower on Axis 1 (closer to medium- and long-tongued species) and were associated with head flowers. Tongue length class explained 22% of diet variance (Adonis permutation test, F = 1.68, p = 0.005). While short-tongued bees appeared most broadly spread in ordination space, there were no significant differences in dispersion across tongue length classes (F = 0.26, p = 0.62).

```{r fig5, echo = FALSE,  fig.cap = paste("Principal coordinates ordination of bumble bee visitation pooled across all samples. To avoid overplotting, only plant genera with an absolute score of at least 0.22 on one or both axes are plotted.")}
knitr::include_graphics("../output/pcoa_aggregate.png")
```

## Site-level diet overlap (Models 5 and 6)

Site-level analysis of diet overlap revealed strong differences in the magnitude of diet overlap at the nest levels of floral morphotype, floral family, and floral genus, as well as consistent patterns in response to elevation. At the site level, however, bumble bees belonging to the same tongue length class did not have significantly more similar diets than those belonging to different tongue length classes.

Morphotype-level diet overlap was nearly double genus-level diet overlap throughout the whole elevation gradient, with family-level diet overlap taking intermediate values. The fixed intercepts of morphotype- and genus-level diet overlap each differed significantly (in opposite directions) from the reference level of family-level diet overlap (p << 0.001). The conditional smooths of each level were significantly different from flat lines (p < 0.001) and generally mirrored one another in shape **(Figure 8)**. Diet overlap was highest at the low end of the elevation gradient, decreased steeply from 600-1200 m, and leveled off or declined weakly for the remainder of the gradient, though morphotype-level diet overlap exhibited a second peak just above 1500 m that was not evident in genus- or family-level diet overlap. The Model 5 explained 78.4% of total deviance (adjusted R^2^ = 0.743).

```{r fig8, echo = FALSE,  fig.cap = paste("Niche overlap by level through elevation. Points show empirical niche overlap values, lines show GAM smooths, and shaded bands show the standard errors of the mean")}
knitr::include_graphics("../output/niche_gam00_plot1.png")
```

Whether a species pair belonged to the same or different tongue length classes had no effect on genus-level niche overlap, either in terms of intercept (p = 0.18) or shape of response to elevation **(Figure 9)**. Model 6 explained 36.5% of total deviance (adjusted R^2^ = 0.316).

```{r fig9, echo = FALSE,  fig.cap = paste("Niche overlap by tongue length class difference through elevation. Lines show GAM smooths and shaded bands show standard errors of the mean")}
knitr::include_graphics("../output/niche_gam20_plot1.png")
```

# DISCUSSION

While bumble bees are, indeed, remarkably generalized in their foraging, they are highly discriminating generalists, dramatically over-representing certain floral taxa and morphotypes and ignoring others. In our study system, lip, flag, and head flowers dominated overall visitation probability when visitation was pooled across all bumble bee species, while bell, disc, and stalk-disc flowers were underrepresented. Patterns of pooled visitation are, of course, driven disproportionately by the behavior of the most abundant bumble bee species, which in our system were *B. pascuorum*, *B. pratorum*, *B. soroeensis*, *B. terrestris/lucorum*, and *B. wurflenii* (Sponsler et al., submitted). Visitation to lip and flag flowers was driven especially by the medium-tongued *B. pascuorum*, the most abundant bumble bee in our study system, and secondarily by the medium-tongued *B. wurflenii*. Visitation to head flowers was driven mainly by the short-tongued *B. soroeensis*.

Tongue length class was an important predictor of preference for some but not all floral morphotypes. For disc and stalk-disc flowers, visitation rates were uniformly low across all tongue length classes. The opposite was true of lip flowers, which were visited at high rates by all tongue length classes. Bell flowers and head flowers were preferred by short-tongued bees, consistent with the shallow corollas of these morphotypes. Conversely, funnel flowers, with their deep corollas were preferred by long-tongued bees, though this effect was driven disproportionately by *B. gerstaeckeri*, which specializes on the funnel flowers of the genus *Aconitum* [@Ponchau2006-af]. Flag flowers were visited at the highest rate by medium-tongued species, and especially by *B. pascuorum* and *B. mucidus*.

While the variation we observed in floral preference across tongue length classes largely comports with the established relationship between tongue length and corolla depth, variation in preference *within* tongue length classes echoes the findings of @Goulson2008-zy in demonstrating the insufficiency of simple morphological matching in explaining bumble bee foraging patterns. Perhaps the most striking divergence of floral preference within tongue length class was the divisive effect of bell flowers on short-tongued bumble bees. *B. pyrenaeus*, *B. soroeensis*, *B. pratorum*, and *B. jonellus* visited bell flowers at high rates, while *B. terrestris/lucorum*, *B. psithyrus*, and *B. lapidarius* visited them relatively rarely. A similar split among the short-tongued species occurred with respect to head flowers, which were visited at high rates by *B. soroeensis* and *B. psithyrus* but at relatively low rates by other short-tongued species. There are no obvious morphological distinctions within short-tongued bumble bees that would account for these divergent affinities, and ecological drivers such as differences in spatial or temporal activity patterns are more likely explanations [@Sowig1989-ld; @Schneller2014-ry]. Nectar robbing is another mechanism by which the relationship between tongue length and corolla depth might be relaxed, but its influence in our data is unclear. Of the bumble bees in our study system, *B. terrestris/lucorum* and especially *B. wurflenii* are known to engage frequently in nectar robbing [@Brian1954-cy]. In the case of *B. wurflenii*, however, its morphotype affinities did not differ markedly from those of other medium-tongued species, and it may be that the condition of medium-tongue length alone confers enough to foraging flexibility in our study system to dampen the divergent effects of nectar robbing. Nectar robbing might, however, account for the deviation of *B. terrestris/lucorum* from most other short-tongued species with respect to flag flowers. With the exception of *B. lapidarius*, all short-tongued species besides *B. terrestris/lucorum* eschewed flag flowers, while *B. terrestris/lucorum* visited them at rates comparable to medium- and long-tongued species. Indeed, B. terrestris/lucorum was, on the whole, an idiosyncratic forager, also demonstrating a relatively high affinity for the disc flowers that were largely ignored by other species. Together, these species-level idiosyncracies of preference within tongue length classes suggest a possible answer to the question of how bumble bees of similar tongue length can coexists, namely that factors other than tongue length are evidently at play in bumble bees' morphological affinities.

With respect to diet overlap among bumble bees, our finding that bumble bee diet ordination produces a primary axis corresponding to tongue length (and associated with affinity for legumes) and a secondary axis describing the dispersion of short-tongued species is consistent with studies in Poland [@Goulson2008-oe], France [@Durieux2000-uk], and England [@Goulson2005-nz] that analyzed bumble bee diet by similar methods, suggesting that this may be a general pattern among bumble bee communities, at least within temperate Europe. Perhaps a more significant finding, though, is that community-mean diet overlap at the level of floral morphotype was consistently nearly twice as high as diet overlap at the level of floral genus. From a numerical perspective, this result is unsurprising simply because each morphotype comprises many genera. From an ecological perspective, though, this pattern suggests another answer to the question of how bumble bee species of similar tongue length can coexists. If tongue length primarily constrains bumble bee foraging at the coarse level of floral morphotype, taxonomic diversity within morphotypes creates space for bumble bees with similar tongue lengths to avoid actual dietary overlap, thus mitigating interspecific competition. Importantly, this suggests that taxonomic diversity *within* floral morphotypes may as important as morphotype diversity *per se* and overall floral abundance in facilitating the coexistence of bumble bee species, an insight which may be relevant to applied bumble bee conservation.

With respect to elevation, community mean niche overlap was highest at the lowest sites and declined sharply over the transition from valley floor to mid-slope. If this pattern of niche overlap moves upslope in response to climate change, it might result in increased competition potential at higher elevations, which could threaten high-elevation bumble bee species that currently exist under conditions of relatively low niche overlap. In our system, though, it must be noted that the lowest elevation sites are also the ones most influences by human land use, and it is possible that the elevation pattern evident in our data is driven by factors other than elevational climate variation.

<!-- If not tongue length, what else could explain the variation in floral morphotype preference witin tongue length classes? -->

# CONCLUSION

The mystery of bumble bee species coexistence hinges on the relationship between species-level floral preferences, trait-matching, and the emergent phenomenon of community-level patterns of diet overlap. While tongue length is a relatively strong predictor of species-level floral preferences, it is at best a weak predictor of diet overlap. This is due in part to divergent preference of bumble bees with similar tongue lengths toward certain floral morphotypes. Perhaps more importantly, though, diet overlap at the level of floral genus tended to be about half as strong as diet overlap at the level of floral morphotype, indicating that bumble bees with similar morphological affinities, driven in part by tongue length compatibility, subdivide floral morphotypes at the genus- and species-level, reducing the potential for competition. This suggests that taxonomic floral diversity may be as important as overall floral abundance in enabling the coexistence of bumble bee species. Something about elevation...

As discussed earlier, the interpretation of the patterns in our data hinge strongly on the extent to which interspecific competition actually occurs in our study system. We have argued that the assumption of minimal competition is reasonable in the context of our study, owing to the rich floral resources and harsh abiotic conditions of the Berchtesgadener Alps, together with the manifest co-occurrence of species with similar tongue length. This leads to the interpretation of abundance-conditioned visitation rates as *preferences* and diet overlap as *potential competition*. To the extent that interspecific competition does occur in our system, though, visitation rates would represent floral preference distorted by the pressure of competition toward niche partitioning, rendering diet overlap an underestimate of potential competition. While we maintain, based on our knowledge of our study system, that the first interpretation is more likely, both interpretations are plausible and must be kept in mind.

<!-- We analyze community mean niche overlap as an indication of diffuse competition [@Pianka1974-qk; MacArthur1972-qz] -->

# ACKNOWLEDGEMENTS

We thank D.J. McNeil for helpful conversations during data analysis.

# REFERENCES

<!-- ```{r write_citations, cache=FALSE, include=FALSE} -->
<!-- write.bibtex(file = "knitcitations.bib") -->
<!-- ``` -->

<div id ="refs"></div>