### 2021-03-15 code/ms1_abundance.Rmd
refined GAM analysis

---
title: "Abundance/diversity GAM"
output: html_document
---

# 1. Prep work
## 1.1 Prepare environment
```{r setup, include=FALSE, echo=FALSE}
library(mgcv)
library(mgcViz)
library(tidymv)
library(vegan)
library(ggvegan)
library(ggrepel)
library(ggforce)
library(tidyverse)
library(lubridate)
```

## 1.2 Load site and trait data
```{r, include=FALSE, echo=FALSE}
site_data <- read_csv("../data/processed_data/site_data.csv") %>%
  dplyr::select(-c(slope.calc, slope.est, elev.min, elev.max, temp.mean)) %>% # drop these variables
  mutate(elev.class2 = factor(elev.class2, levels = c("low", "mid", "high")),
         site = factor(site),
         tree_line = factor(tree_line)) # turn elev.class into an ordered factor and site into a regular factor

# BB traits
bb_traits <- read_csv("../data/processed_data/bb_traits.csv")

# FL traits
## family-level taxonomy
fl_tax <- read_csv("../data/processed_data/floral_tax.csv")

## Kugler morphotypes (accessed through Bioflor)
fl_ktype <- read_csv("../data/processed_data/floral_k_type.csv")

## Flower colors (accessed through BioFlor)
fl_color <- read_csv("../data/processed_data/floral_color.csv")

## diameter and corolla depth data from Gita Benadi
fl_morph <- read_csv("../data/processed_data/flower_morphology_GitaBenadi.csv") %>% 
  dplyr::select(SampleID, plant.family = Family, plant.sp = Species, 
                Unit, Flowers_per_unit, diam = Diameter, LN) %>%
  group_by(plant.sp) %>%
  summarize(diam = mean(diam), LN = mean(LN))

## verbal description of Kugler morphotypes
fl_ktype_key <- read_delim("../data/processed_data/kugler_key.tsv", delim = "\t", col_names = FALSE) %>%
  dplyr::select(k.type = 1, description = 2)

## join all floral trait tables
fl_traits <- fl_tax %>%
  full_join(fl_ktype) %>%
  full_join(fl_color) %>%
  full_join(fl_morph) %>%
  dplyr::select(plant.sp, plant.genus, plant.family, k.type, k.type.s, k.type.ss, color, diam, LN)
```

## 1.3 Load network and survey data
```{r message=FALSE}
### BB-FL visitation data
net <- read_csv("../data/processed_data/network.csv") %>%
  filter(caste != "male",
         !bb.sp %in% c("humi", "hypn", "lapi")) %>% # too rare
  dplyr::select(site, date, bb.sp, plant.sp) # drop unused columns

### Floral survey data
survey <- read_csv("../data/processed_data/floral_survey.csv") %>%
  semi_join(net, by = "plant.sp") %>% # Consider only visited plant species (anytime anywhere, not specifically for a given site*date)
  dplyr::select(site, date, plant.sp, flower.cover) %>%
  group_by(plant.sp) %>%
  mutate(present = case_when(
    max(flower.cover) > 0 ~ TRUE,
    max(flower.cover) == 0 ~ FALSE
  )) %>%
  filter(present == TRUE) # Ignore plants that never had any flower cover

vis_sp <- net %>%
  left_join(fl_tax) %>%
  select(plant.sp, plant.genus, plant.family) %>%
  unique()

surv_sp <- survey %>%
  left_join(fl_tax) %>%
  select(plant.sp, plant.genus, plant.family) %>%
  unique()

missing <- anti_join(vis_sp, surv_sp)
```

## 1.4 Calculate FL and BB abundance / diversity
```{r echo=FALSE, include=TRUE, message=FALSE}
# Per-transect BB abundance
## full dataset
bb_abund_sp <- net %>%
  dplyr::select(site, date, bb.sp) %>%
  group_by(site, date, bb.sp) %>% # 
  summarize(bb.abund = n()) %>% # get per transect bb abundance
  ungroup() %>%
  complete(site, date, bb.sp, fill = list(bb.abund = 0)) %>% # complete implicit absences
  semi_join(net, by = c("site", "date")) %>% # drop site-dates that did not exist in the data set
  left_join(site_data) %>% # add site data
  mutate(year = factor(year(date)),
         yday = yday(date),
         bb.sp = factor(bb.sp),
         management = factor(management),
         site = factor(site),
         transect = factor(transect)) %>%
  left_join(bb_traits) %>%
  mutate(tongue = factor(pbl.w.class2))

### Per-transect floral abund by species
fl_abund_sp <- survey %>%
  group_by(site, date, plant.sp) %>% # group by plant.sp*transect
  summarize(fl.abund = sum(flower.cover)) %>% # get plant.sp*transect total flower cover; this step really shouldn't be required, but in a few cases, there is more than one abundance entry for the same species on a given transect; not sure how that happened, but it's in the original data that Fabrice sent me; alternatively, I could drop the dups or average them. We'll see what Kalli and Fabrice have to say.
  group_by(site, date) %>%
  mutate(prop.fl.abund = fl.abund/sum(fl.abund)) %>% # get the proportional flower cover for each species.
  ungroup() %>%
  #full_join(dplyr::select(net, site, date, plant.sp), by = c("site", "date", "plant.sp")) %>% # add the handful of species occurring in the visitation network but missing from the survey data
  complete(site, date, plant.sp, fill = list(fl.abund = 0, prop.fl.abund = 0)) %>%
  semi_join(survey, by = c("site", "date")) %>%  # only include site-dates that actually occurred
  unique() %>%
  mutate(year = factor(year(date))) %>%
  left_join(fl_traits)
```

## 1.5 Calculate binary "preference" for each plant genus by each bumble bee species; score species 0 or 1 based on whether they were visited
```{r message=FALSE}
weights <- fl_abund_sp %>%
  select(plant.sp, plant.genus) %>%
  unique() %>%
  left_join(select(net, bb.sp, plant.sp)) %>%
  select(plant.genus, bb.sp) %>%
  distinct() %>%
  mutate(visited = rep(1, n())) %>%
  pivot_wider(values_from = visited, names_from = bb.sp, values_fill = 0) %>%
  pivot_longer(cols = 2:14, values_to = "visited", names_to = "bb.sp")

weighted_fl_abund <- fl_abund_sp %>%
  group_by(site, date, plant.genus) %>%
  summarize(flower.cover = sum(fl.abund)) %>%
  left_join(weights) %>%
  mutate(weighted.flower.cover = flower.cover*visited) %>%
  group_by(site, date, bb.sp) %>%
  summarize(total.weighted.flower.cover = sum(weighted.flower.cover)) %>%
  left_join(site_data) %>%
  select(site, date, elev.mean, bb.sp, total.weighted.flower.cover) %>%
  mutate(yday = yday(date),
         bb.sp = factor(bb.sp),
         site = factor(site),
         year = factor(year(date)))
```

# 2. Analyses

## 2.1 BB elevation distribution
```{r}
bb_range <- net %>%
  left_join(site_data) %>%
  group_by(site, date, elev.mean, tree_line, bb.sp) %>%
  summarize(abund = n()) %>%
  group_by(site, elev.mean, tree_line, bb.sp) %>%
  summarize(abund = max(abund)) %>%
  filter(abund > 0) %>%
  group_by(bb.sp) %>%
  mutate(elev.floor = min(elev.mean),
         elev.ceiling = max(elev.mean),
         elev.range = elev.ceiling - elev.floor,
         elev.med = elev.floor + ((elev.ceiling - elev.floor)/2))

ggplot(bb_range, aes(reorder(bb.sp, elev.med), elev.mean)) +
  geom_line(size = 2, alpha = 0.25) +
  geom_point(aes(size = abund, color = tree_line), alpha = 0.5) +
  theme_light(14) +
  scale_size_continuous(name = "Peak abundance",
                        breaks = c(1, 10, 20, 40, 80)) +
  scale_color_discrete(name = "Tree line") +
  xlab("Bumble bee species") +
  ylab("Elevation (m)") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

ggsave("../output/bb_range_ms1.pdf")
ggsave("../output/bb_range_ms1.png")

# What happens if I plot max abundance by elevation?

bb_range2 <- net %>%
  left_join(site_data) %>%
  group_by(site, date, elev.mean, tree_line, bb.sp) %>%
  summarize(abund = n()) %>%
  group_by(site, elev.mean, tree_line, bb.sp) %>%
  summarize(abund = max(abund))

ggplot(filter(bb_range2, bb.sp != "humi"), aes(elev.mean, abund)) +
  geom_point() +
  geom_smooth(method = "gam", formula = y ~ s(x, k = 6)) +
  facet_wrap(~bb.sp, scales = "free_y", ncol = 4) +
  geom_vline(xintercept = 1500, color = "red", linetype = "dashed")

# Overall bumble bee abundance
bb_abund_sum <- bb_abund_sp %>%
  group_by(site, bb.sp) %>%
  summarize(cum.abund = sum(bb.abund),
            mean.abund = mean(bb.abund)) %>%
  group_by(site) %>%
  mutate(prop.abund = cum.abund/sum(cum.abund)) %>%
  group_by(bb.sp) %>%
  mutate(mean.prop.abund = mean(prop.abund))

ggplot(bb_abund_sum, aes(site, prop.abund, fill = bb.sp)) +
  geom_col()

ggplot(bb_abund_sum, aes(bb.sp, mean.abund)) +
  geom_boxplot()

ggplot(bb_abund_sum, aes(reorder(bb.sp, -mean.prop.abund), prop.abund)) +
  geom_boxplot(fill = "gray85") +
  theme_light() +
  ylab("Proportional abundance") +
  xlab("Bumble bee species")

ggsave("../output/bb_abundsum.pdf")
ggsave("../output/bb_abundsum.png")
```

## 2.2. Aggregate analysis of max BB abundance and mean FL abundance  ~ elevation

### 2.2.1 Calculate max and mean abundances within site and year
```{r}
# Get mean total weighted floral abundance by BB sp
weighted_fl_abund_mean <- weighted_fl_abund %>%
  group_by(site, bb.sp, year) %>%
  summarize(mean.fl.abund = mean(total.weighted.flower.cover),
            max.fl.abund = max(total.weighted.flower.cover)) %>%
  left_join(site_data)

# Get max BB abundance by sp
bb_abund_sp_max <- bb_abund_sp %>%
  group_by(site, elev.mean, year, bb.sp) %>%
  summarize(max.bb.abund = max(bb.abund),
            mean.bb.abund = mean(bb.abund))

# Join
abund_sum <- full_join(weighted_fl_abund_mean, bb_abund_sp_max) %>%  
  unite(sp.year, c(bb.sp, year), remove = FALSE) %>%
  mutate(site = factor(site),
         bb.sp = factor(bb.sp),
         sp.year = factor(sp.year)) %>%
  unite(bb.sp.year, c(bb.sp, year), sep = "_", remove = FALSE) %>%
  mutate(bb.sp.year = factor(bb.sp.year))
```

### 2.2.2 Preliminary visualizations
```{r}
## mean floral abundance
ggplot(abund_sum, aes(elev.mean, mean.fl.abund, color = year)) +
  geom_point(alpha = 0.5) +
  #geom_smooth(method = "gam", formula = y ~ s(x, k = 8), se = FALSE) +
  facet_wrap(~bb.sp, scales = "free_y") +
  geom_vline(xintercept = 1500)

## log(mean floral abundance)
ggplot(abund_sum, aes(elev.mean, log(mean.fl.abund), color = year)) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = "gam", formula = y ~ s(x, k = 8), se = FALSE) +
  facet_wrap(~bb.sp) +
  geom_vline(xintercept = 1500)

## Max floral abundance)
ggplot(abund_sum, aes(elev.mean, max.fl.abund, color = year)) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = "gam", formula = y ~ s(x, k = 8), se = FALSE) +
  facet_wrap(~bb.sp) +
  geom_vline(xintercept = 1500)

## Max BB abundance
ggplot(abund_sum, aes(elev.mean, max.bb.abund, color = year)) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = "gam", formula = y ~ s(x, k = 8),  se = FALSE) +
  facet_wrap(~bb.sp, scales = "free_y") +
  geom_vline(xintercept = 1500)

## Max  abundance ~ mean BB abundance, to see how interchangeable they are.
## They are almost perfectly correlated, and I think max abund is more informative
ggplot(abund_sum, aes(mean.bb.abund, max.bb.abund, color = year)) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = "gam", formula = y ~ s(x, k = 8),  se = FALSE) +
  facet_wrap(~bb.sp, scales = "free_y")

## Are max FL and mean FL similarly interchangeable? Yes.
ggplot(abund_sum, aes(mean.fl.abund, max.fl.abund, color = year)) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = "gam", formula = y ~ s(x, k = 8),  se = FALSE) +
  facet_wrap(~bb.sp, scales = "free_y")

## Max BB abundance ~ mean FL abundance
ggplot(abund_sum, aes(log(mean.fl.abund), max.bb.abund)) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = "gam", formula = y ~ s(x, k = 8),  se = FALSE) +
  facet_wrap(~bb.sp, scales = "free_y")
```

### 2.2.3 Abundance models
I'm choosing to use site as a random intercept effect in these models because I think that is the correct way to specify a model involving repeated observations (in this case, bumble bee species) within site. The effect of this decision is that the elevation smoother is less influenced by individual sites and converges more strongly toward the mean. This creates less exciting curves, but I think it is a more principled way to model the data and less prone to overfitting. I can have my cake and eat it too, though, by plotting the actual data over the GAM fit so that the strong patterns of certain sites are visible. Then I can talk about the conditional elevation effect and the importance of certain peaks that are likely to be idiosyncrasies of my system (i.e. it would be overfitting to call them elevation effects).

#### 2.2.3.1 Bumble bees
```{r}
# Call model
gam_max_bb_abund_00 <- gam(max.bb.abund ~ 
                             s(bb.sp, bs = "re") +
                             s(bb.sp.year, bs = "re") + # fixed intercept for bb.sp:year
                             s(site, bs = "re") + # random intercept for site
                             s(elev.mean, by = bb.sp, k = 8), # smoother for elev by BB sp.
                           method = "REML",
                           family = "tw", # Tweedie distribution performs well
                           select = TRUE,
                           data = abund_sum) %>% getViz()

gam_max_bb_abund_00_ilink <- family(gam_max_bb_abund_00)$linkinv # get inverse link function

check.gamViz(gam_max_bb_abund_00) # evaluate model
gam_max_bb_abund_00_sum <- summary(gam_max_bb_abund_00) # summarize model
print(plot(gam_max_bb_abund_00, allTerms = TRUE), pages = 1) # preliminary visualization

# Plot smooths and year effect on the link scale
## Individual plots
gers_abund <- plot(sm(gam_max_bb_abund_00, 4)) + l_fitLine() + l_ciLine() + 
  labs(x = "Elevation (m.s.l.)", y = "Peak abundance", title = "gers")
hort_abund <- plot(sm(gam_max_bb_abund_00, 5)) + l_fitLine() + l_ciLine() + 
  labs(x = "Elevation (m.s.l.)", y = "Peak abundance", title = "hort")
jone_abund <- plot(sm(gam_max_bb_abund_00, 6)) + l_fitLine() + l_ciLine() + 
  labs(x = "Elevation (m.s.l.)", y = "Peak abundance", title = "jone")
mend_abund <- plot(sm(gam_max_bb_abund_00, 7)) + l_fitLine() + l_ciLine() + 
  labs(x = "Elevation (m.s.l.)", y = "Peak abundance", title = "mend")
mont_abund <- plot(sm(gam_max_bb_abund_00, 8)) + l_fitLine() + l_ciLine() + 
  labs(x = "Elevation (m.s.l.)", y = "Peak abundance", title = "mont")
muci_abund <- plot(sm(gam_max_bb_abund_00, 9)) + l_fitLine() + l_ciLine() + 
  labs(x = "Elevation (m.s.l.)", y = "Peak abundance", title = "muci")
pasc_abund <- plot(sm(gam_max_bb_abund_00, 10)) + l_fitLine() + l_ciLine() + 
  labs(x = "Elevation (m.s.l.)", y = "Peak abundance", title = "pasc")
prat_abund <- plot(sm(gam_max_bb_abund_00, 11)) + l_fitLine() + l_ciLine() + 
  labs(x = "Elevation (m.s.l.)", y = "Peak abundance", title = "prat")
psit_abund <- plot(sm(gam_max_bb_abund_00, 12)) + l_fitLine() + l_ciLine() + 
  labs(x = "Elevation (m.s.l.)", y = "Peak abundance", title = "psit")
pyre_abund <- plot(sm(gam_max_bb_abund_00, 13)) + l_fitLine() + l_ciLine() + 
  labs(x = "Elevation (m.s.l.)", y = "Peak abundance", title = "pyre")
soro_abund <- plot(sm(gam_max_bb_abund_00, 14)) + l_fitLine() + l_ciLine() + 
  labs(x = "Elevation (m.s.l.)", y = "Peak abundance", title = "soro")
telu_abund <- plot(sm(gam_max_bb_abund_00, 15)) + l_fitLine() + l_ciLine() + 
  labs(x = "Elevation (m.s.l.)", y = "Peak abundance", title = "telu")
wurf_abund <- plot(sm(gam_max_bb_abund_00, 16)) + l_fitLine() + l_ciLine() + 
  labs(x = "Elevation (m.s.l.)", y = "Peak abundance", title = "wurf")

## Arrange on grid plot
grid_bb_abund <- gridPrint(gers_abund, hort_abund, jone_abund, 
                           mend_abund, mont_abund, muci_abund, 
                           pasc_abund, prat_abund, psit_abund, 
                           pyre_abund, soro_abund, telu_abund, 
                           wurf_abund, year_effect)

pdf("../output/gam_max_bb_abund_grid.pdf")
plot(grid_bb_abund)
dev.off()

# Plot model predictions on response scale and add data as points
## Extract and prepare predictions
gam_max_bb_abund_00_pred <- predict.gam(gam_max_bb_abund_00, # get model predictions
                                        type = "terms") %>% 
  as.data.frame() %>% # convert to data frame
  as_tibble() %>% # convert to tibble
  select(sm.bb.sp = 1, sm.bb.sp.year = 2, sm.site = 3, gers = 4, # rename
         hort = 5, jone = 6, mend = 7, mont = 8, 
         muci = 9, pasc = 10, prat = 11, psit = 12, 
         pyre = 13, soro = 14, telu = 15, wurf = 16) %>%
  mutate(fit = # get cumulative fit and se, ignoring site effect 
           sm.bb.sp + sm.bb.sp.year + gers + 
           hort + jone + mend + mont + 
           muci + pasc + prat + psit + 
           pyre + soro + telu + wurf 
           ) %>%
  select(fit) %>%
  mutate(fit.trans = gam_max_bb_abund_00_ilink(fit)) %>%
  bind_cols(abund_sum)

## Plot
ggplot(gam_max_bb_abund_00_pred, aes(elev.mean, fit.trans, color = year)) +
  # geom_ribbon(aes(elev.mean, fit.trans, 
  #                 ymax = fit.plus.trans, ymin = fit.min.trans,
  #                 fill = year), alpha = 0.2, inherit.aes = FALSE) +
  geom_line(size = 1.5, alpha = 0.5) +
  geom_point(aes(elev.mean, max.bb.abund, color = year), inherit.aes = FALSE, alpha = 0.25) +
  facet_wrap(~bb.sp, scales = "free_y") +
  geom_vline(xintercept = 1500, linetype = "dashed") +
  theme_light() +
  labs(x = "Elevation (m.s.l.)", 
       y = bquote("Peak abundance"),
       color = "Year",
       fill = "Year")

ggsave("../output/gam_max_bb_abund.pdf", width = 6.5, height = 4.5)
ggsave("../output/gam_max_bb_abund.png", width = 6.5, height = 4.5)
```

#### 2.2.3.2 Floral resources
```{r}
gam_mean_fl_abund_00 <- gam(mean.fl.abund ~ 
                             s(site, bs = "re") + # random intercept for site
                             s(elev.mean, year, bs = "fs", k = 8) +
                             s(elev.mean, bb.sp, bs = "fs", k = 8),
                           method = "REML",
                           family = "tw",
                           select = TRUE,
                           data = abund_sum) %>% getViz()

gam_mean_fl_abund_00_ilink <- family(gam_mean_fl_abund_00)$linkinv # get inverse link function

check.gamViz(gam_mean_fl_abund_00) # evaluate model
gam_mean_fl_abund_00_sum <- summary(gam_mean_fl_abund_00) # summarize model
print(plot(gam_mean_fl_abund_00, allTerms = TRUE), pages = 1) # preliminary visualization

# Plot on the link scale
elev_effect <- plot(sm(gam_mean_fl_abund_00, 2)) + l_fitLine(alpha = 0.6, size = 1)
bb.sp_effect <- plot(sm(gam_mean_fl_abund_00, 3)) + l_fitLine(alpha = 0.6, size = 1)

grid_fl_abund <- gridPrint(elev_effect, bb.sp_effect)

pdf("../output/gam_mean_fl_abund_grid.pdf")
plot(grid_fl_abund)
dev.off()

# Plot model predictions on response scale and add data as points
## Extract and prepare predictions
gam_mean_fl_abund_00_pred <- predict.gam(gam_mean_fl_abund_00, 
                                         type = "terms", 
                                         se.fit = TRUE) %>% # get model predictions and se
  as.data.frame() %>% # convert to data frame
  as_tibble() %>% # convert to tibble
  select(sm.site = 1, sm.year = 2, sm.bb.sp = 3, # rename
         se.site = 4, se.year = 5, se.bb.sp = 6) %>%
  mutate(fit = sm.year, # get fit and se ignoring site and species
         se = se.year) %>% 
  select(fit, se) %>%
  mutate(fit.plus = fit + se, # upper CI bound
         fit.min = fit - se, # lower CI bound
         fit.trans = gam_mean_fl_abund_00_ilink(fit), # apply inverse link function
         fit.plus.trans = gam_mean_fl_abund_00_ilink(fit.plus),
         fit.min.trans = gam_mean_fl_abund_00_ilink(fit.min)) %>%
  bind_cols(na.omit(abund_sum)) %>% # the na.omit() drops the rows where floral surveying data were missing
  select(fit.trans, fit.plus.trans, fit.min.trans, 
         se, elev.mean, year, mean.fl.abund) %>%
  distinct()

ggplot(gam_mean_fl_abund_00_pred, aes(elev.mean, fit.trans, color = year)) +
  # geom_ribbon(aes(elev.mean, fit.trans, 
  #                 ymax = fit.plus.trans, ymin = fit.min.trans,
  #                 fill = year), alpha = 0.2, inherit.aes = FALSE) +
  geom_line(size = 1.5, alpha = 0.5) +
  geom_point(aes(elev.mean, mean.fl.abund, color = year), inherit.aes = FALSE, alpha = 0.15) +
  geom_vline(xintercept = 1500, linetype = "dashed") +
  theme_light() +
  labs(x = "Elevation (m.s.l.)", 
       y = bquote("Mean flower cover" ~ (m^2)),
       color = "Year",
       fill = "Year")

ggsave("../output/gam_mean_fl_abund.pdf")
ggsave("../output/gam_mean_fl_abund.png")

ggplot(gam_mean_fl_abund_00_pred, aes(elev.mean, fit.trans, color = year)) +
  # geom_ribbon(aes(elev.mean, fit.trans, 
  #                 ymax = fit.plus.trans, ymin = fit.min.trans,
  #                 fill = year), alpha = 0.2, inherit.aes = FALSE) +
  geom_line(size = 1.5, alpha = 0.5) +
  geom_point(aes(elev.mean, mean.fl.abund, color = year), inherit.aes = FALSE, alpha = 0.15) +
  geom_vline(xintercept = 1500, linetype = "dashed") +
  theme_light() +
  labs(x = "Elevation (m.s.l.)", 
       y = bquote("Mean flower cover" ~ (m^2)),
       color = "Year",
       fill = "Year") +
  ylim(c(0, 10))

ggsave("../output/gam_mean_fl_abund_sub.pdf")
ggsave("../output/gam_mean_fl_abund_sub.png")

ggplot(gam_mean_fl_abund_00_pred, aes(elev.mean, fit.trans, color = year)) +
  # geom_ribbon(aes(elev.mean, fit.trans, 
  #                 ymax = fit.plus.trans, ymin = fit.min.trans,
  #                 fill = year), alpha = 0.2, inherit.aes = FALSE) +
  geom_line(size = 1.5, alpha = 0.5) +
  geom_point(aes(elev.mean, mean.fl.abund, color = year), inherit.aes = FALSE, alpha = 0.15) +
  geom_vline(xintercept = 1500, linetype = "dashed") +
  theme_light() +
  labs(x = "Elevation (m.s.l.)", 
       y = bquote("Mean flower cover" ~ (m^2)),
       color = "Year",
       fill = "Year") +
  facet_zoom(ylim = c(0, 10), 
             show.area = FALSE,
             zoom.size = 1)

ggsave("../output/gam_mean_fl_abund_zoom.pdf", width = 6.5, height = 4.5)
ggsave("../output/gam_mean_fl_abund_zoom.png", width = 6.5, height = 4.5)
```

## 2.3 Elevation x time analysis

### 2.3.1 Floral abundance
```{r message=FALSE, warning=FALSE}
# First, let's model overall floral abundance without parsing it out by which species were visited by which bumble bee species
fl_abund_pool <- fl_abund_sp %>%
  group_by(site, date, year) %>%
  summarize(fl.abund = sum(fl.abund)) %>%
  mutate(site = factor(site),
         yday = yday(date)) %>%
  left_join(site_data)

ggplot(fl_abund_pool, aes(fl.abund)) +
  geom_histogram()

fl_abund.qp <- bam(fl.abund ~ 
                      s(site, bs = "re") +
                      s(year, bs = "re") +
                      te(yday, elev.mean,
                         bs = c("gp", "tp"),
                         k = c(8, 8)),
                    data = fl_abund_pool,
                    family = "quasipoisson",
                    discrete = TRUE,
                    select = TRUE,
                    method = "fREML") %>% getViz()

check.gamViz(fl_abund.qp)
summary(fl_abund.qp)
plot(fl_abund.qp)

# Now, by bumble bee species
weighted_fl_abund.qp <- bam(total.weighted.flower.cover ~ 
                              s(bb.sp, bs = "re") +
                              s(site, bs = "re") +
                              s(year, bs = "re") +
                              te(yday, elev.mean,
                                 by = bb.sp,
                                 bs = c("gp", "tp"),
                                 k = c(5, 5)),
                            data = filter(weighted_fl_abund, bb.sp != "humi"),
                            family = "quasipoisson",
                            discrete = TRUE,
                            select = TRUE,
                            method = "fREML") %>% getViz(nsim = 500)

check.gamViz(weighted_fl_abund.qp)
sum_fl.qp <- summary(weighted_fl_abund.qp)
print(plot(weighted_fl_abund.qp, allTerms = TRUE), pages = 4)
check2D(weighted_fl_abund.qp, x1 = "yday", x2 = "elev.mean")

gers_fl_abund <- plot(sm(weighted_fl_abund.qp, 4))
hort_fl_abund <- plot(sm(weighted_fl_abund.qp, 5))
hypn_fl_abund <- plot(sm(weighted_fl_abund.qp, 6))
jone_fl_abund <- plot(sm(weighted_fl_abund.qp, 7))
lapi_fl_abund <- plot(sm(weighted_fl_abund.qp, 8))
mend_fl_abund <- plot(sm(weighted_fl_abund.qp, 9))
mont_fl_abund <- plot(sm(weighted_fl_abund.qp, 10))
muci_fl_abund <- plot(sm(weighted_fl_abund.qp, 11))
pasc_fl_abund <- plot(sm(weighted_fl_abund.qp, 12))
prat_fl_abund <- plot(sm(weighted_fl_abund.qp, 13))
psit_fl_abund <- plot(sm(weighted_fl_abund.qp, 14))
pyre_fl_abund <- plot(sm(weighted_fl_abund.qp, 15))
soro_fl_abund <- plot(sm(weighted_fl_abund.qp, 16))
telu_fl_abund <- plot(sm(weighted_fl_abund.qp, 17))
wurf_fl_abund <- plot(sm(weighted_fl_abund.qp, 18))

pdf("../output/gam_FL.pdf")
grid_plot_fl_abun <- gridPrint(gers_fl_abund + 
                                  l_fitRaster() + 
                                  l_fitContour() + 
                                  labs(title = "gers", x = NULL, y = NULL) +
                                  guides(fill=FALSE),
                               hort_fl_abund + 
                                 l_fitRaster() + 
                                 l_fitContour() + 
                                 labs(title = "hort", x = NULL, y = NULL) +
                                 guides(fill=FALSE),
                               hypn_fl_abund + 
                                 l_fitRaster() + 
                                 l_fitContour() + 
                                 labs(title = "hypn", x = NULL, y = NULL) +
                                 guides(fill=FALSE),
                               jone_fl_abund + 
                                 l_fitRaster() + 
                                 l_fitContour() + 
                                 labs(title = "jone", x = NULL, y = NULL) +
                                 guides(fill=FALSE),
                               lapi_fl_abund + 
                                 l_fitRaster() + 
                                 l_fitContour() + 
                                 labs(title = "lapi", x = NULL, y = NULL) +
                                 guides(fill=FALSE),
                               mend_fl_abund + 
                                 l_fitRaster() + 
                                 l_fitContour() + 
                                 labs(title = "mend", x = NULL, y = NULL) +
                                 guides(fill=FALSE),
                               mont_fl_abund + 
                                 l_fitRaster() + 
                                 l_fitContour() + 
                                 labs(title = "mont", x = NULL, y = NULL) +
                                 guides(fill=FALSE),
                               muci_fl_abund + 
                                 l_fitRaster() + 
                                 l_fitContour() + 
                                 labs(title = "muci", x = NULL, y = NULL) +
                                 guides(fill=FALSE),
                               pasc_fl_abund + 
                                 l_fitRaster() + 
                                 l_fitContour() + 
                                 labs(title = "pasc", x = NULL, y = NULL) +
                                 guides(fill=FALSE),
                               prat_fl_abund + 
                                 l_fitRaster() + 
                                 l_fitContour() + 
                                 labs(title = "prat", x = NULL, y = NULL) +
                                 guides(fill=FALSE),
                               psit_fl_abund + 
                                 l_fitRaster() + 
                                 l_fitContour() + 
                                 labs(title = "psit", x = NULL, y = NULL) +
                                 guides(fill=FALSE),
                               pyre_fl_abund + 
                                 l_fitRaster() + 
                                 l_fitContour() + 
                                 labs(title = "pyre", x = NULL, y = NULL) +
                                 guides(fill=FALSE),
                               soro_fl_abund + 
                                 l_fitRaster() + 
                                 l_fitContour() + 
                                 labs(title = "soro", x = NULL, y = NULL) +
                                 guides(fill=FALSE),
                               telu_fl_abund + 
                                 l_fitRaster() + 
                                 l_fitContour() + 
                                 labs(title = "telu", x = NULL, y = NULL) +
                                 guides(fill=FALSE),
                               wurf_fl_abund + 
                                 l_fitRaster() + 
                                 l_fitContour() + 
                                 labs(title = "wurf", x = NULL, y = NULL) +
                                 guides(fill=FALSE),
                               ncol = 4)
dev.off()
```

### 2.3.2 Bumble bee abundance
```{r message=FALSE}
# Zero-inflation may be an issue; a simple Poisson model is badly overdispersed
ggplot(bb_abund_sp, aes(bb.abund)) +
  geom_histogram() +
  facet_wrap(~bb.sp)

# Quasipoisson (winner)
bb_abund.qp <- bam(bb.abund ~ 
                     s(bb.sp.ord, bs = "re") + # center species
                     s(year, bs = "re") + # random intercept for year
                     s(site, bs = "re") + # random intercept to avoid pseudoreplicating elevation
                     te(yday, elev.mean, # tensor product smooth of elevation and time
                        by = bb.sp.ord, # by bb.sp
                        bs= c("gp", "tp"), # gaussian procees for time, thinplate spline for elevation
                        k=c(5, 5), 
                        m=2),
                   data = filter(bb_abund_sp, bb.sp != "humi"), # humi is just too rare
                   discrete = TRUE,
                   family = "quasipoisson",
                   select = TRUE,
                   method = "fREML") %>% getViz(nsim = 500)

check.gamViz(bb_abund.qp)
sum.qp <- summary(bb_abund.qp)
print(plot(bb_abund.qp, allTerms = TRUE), pages = 4)
check2D(bb_abund.qp, x1 = "yday", x2 = "elev.mean")

telu_abund <- plot(sm(bb_abund.qp, 4))
pasc_abund <- plot(sm(bb_abund.qp, 5))
prat_abund <- plot(sm(bb_abund.qp, 6))
soro_abund <- plot(sm(bb_abund.qp, 7))
wurf_abund <- plot(sm(bb_abund.qp, 8))
hort_abund <- plot(sm(bb_abund.qp, 9))
jone_abund <- plot(sm(bb_abund.qp, 10))
psit_abund <- plot(sm(bb_abund.qp, 11))
pyre_abund <- plot(sm(bb_abund.qp, 12))
mend_abund <- plot(sm(bb_abund.qp, 13))
mont_abund <- plot(sm(bb_abund.qp, 14))
gers_abund <- plot(sm(bb_abund.qp, 15))
muci_abund <- plot(sm(bb_abund.qp, 16))
hypn_abund <- plot(sm(bb_abund.qp, 17))
lapi_abund <- plot(sm(bb_abund.qp, 18))

pdf("../output/gam_BB.pdf")
grid_plot_fl_abun <- gridPrint(gers_abund + 
                                  l_fitRaster() + 
                                  l_fitContour() + 
                                  labs(title = "gers", x = NULL, y = NULL) +
                                  guides(fill=FALSE),
                               hort_abund + 
                                 l_fitRaster() + 
                                 l_fitContour() + 
                                 labs(title = "hort", x = NULL, y = NULL) +
                                 guides(fill=FALSE),
                               hypn_abund + 
                                 l_fitRaster() + 
                                 l_fitContour() + 
                                 labs(title = "hypn", x = NULL, y = NULL) +
                                 guides(fill=FALSE),
                               jone_abund + 
                                 l_fitRaster() + 
                                 l_fitContour() + 
                                 labs(title = "jone", x = NULL, y = NULL) +
                                 guides(fill=FALSE),
                               lapi_abund + 
                                 l_fitRaster() + 
                                 l_fitContour() + 
                                 labs(title = "lapi", x = NULL, y = NULL) +
                                 guides(fill=FALSE),
                               mend_abund + 
                                 l_fitRaster() + 
                                 l_fitContour() + 
                                 labs(title = "mend", x = NULL, y = NULL) +
                                 guides(fill=FALSE),
                               mont_abund + 
                                 l_fitRaster() + 
                                 l_fitContour() + 
                                 labs(title = "mont", x = NULL, y = NULL) +
                                 guides(fill=FALSE),
                               muci_abund + 
                                 l_fitRaster() + 
                                 l_fitContour() + 
                                 labs(title = "muci", x = NULL, y = NULL) +
                                 guides(fill=FALSE),
                               pasc_abund + 
                                 l_fitRaster() + 
                                 l_fitContour() + 
                                 labs(title = "pasc", x = NULL, y = NULL) +
                                 guides(fill=FALSE),
                               prat_abund + 
                                 l_fitRaster() + 
                                 l_fitContour() + 
                                 labs(title = "prat", x = NULL, y = NULL) +
                                 guides(fill=FALSE),
                               psit_abund + 
                                 l_fitRaster() + 
                                 l_fitContour() + 
                                 labs(title = "psit", x = NULL, y = NULL) +
                                 guides(fill=FALSE),
                               pyre_abund + 
                                 l_fitRaster() + 
                                 l_fitContour() + 
                                 labs(title = "pyre", x = NULL, y = NULL) +
                                 guides(fill=FALSE),
                               soro_abund + 
                                 l_fitRaster() + 
                                 l_fitContour() + 
                                 labs(title = "soro", x = NULL, y = NULL) +
                                 guides(fill=FALSE),
                               telu_abund + 
                                 l_fitRaster() + 
                                 l_fitContour() + 
                                 labs(title = "telu", x = NULL, y = NULL) +
                                 guides(fill=FALSE),
                               wurf_abund + 
                                 l_fitRaster() + 
                                 l_fitContour() + 
                                 labs(title = "wurf", x = NULL, y = NULL) +
                                 guides(fill=FALSE),
                               ncol = 4)
dev.off()
```