### 2021-03-16 code/ms1_abundance.Rmd
simplified GAM visualization and opted to drop the site random effect

---
title: "Abundance/diversity GAM"
output: html_document
---

# 1. Prep work
## 1.1 Prepare environment
```{r setup, include=FALSE, echo=FALSE}
library(mgcv)
library(mgcViz)
library(patchwork)
library(tidymv)
library(vegan)
library(ggvegan)
library(ggrepel)
library(ggforce)
library(tidyverse)
library(lubridate)
```

## 1.2 Load site and trait data
```{r, include=FALSE, echo=FALSE}
site_data <- read_csv("../data/processed_data/site_data.csv") %>%
  dplyr::select(-c(slope.calc, slope.est, elev.min, elev.max, temp.mean)) %>% # drop these variables
  mutate(elev.class2 = factor(elev.class2, levels = c("low", "mid", "high")),
         site = factor(site),
         tree_line = factor(tree_line)) # turn elev.class into an ordered factor and site into a regular factor

# BB traits
bb_traits <- read_csv("../data/processed_data/bb_traits.csv")

# FL traits
## family-level taxonomy
fl_tax <- read_csv("../data/processed_data/floral_tax.csv")

## Kugler morphotypes (accessed through Bioflor)
fl_ktype <- read_csv("../data/processed_data/floral_k_type.csv")

## Flower colors (accessed through BioFlor)
fl_color <- read_csv("../data/processed_data/floral_color.csv")

## diameter and corolla depth data from Gita Benadi
fl_morph <- read_csv("../data/processed_data/flower_morphology_GitaBenadi.csv") %>% 
  dplyr::select(SampleID, plant.family = Family, plant.sp = Species, 
                Unit, Flowers_per_unit, diam = Diameter, LN) %>%
  group_by(plant.sp) %>%
  summarize(diam = mean(diam), LN = mean(LN))

## verbal description of Kugler morphotypes
fl_ktype_key <- read_delim("../data/processed_data/kugler_key.tsv", delim = "\t", col_names = FALSE) %>%
  dplyr::select(k.type = 1, description = 2)

## join all floral trait tables
fl_traits <- fl_tax %>%
  full_join(fl_ktype) %>%
  full_join(fl_color) %>%
  full_join(fl_morph) %>%
  dplyr::select(plant.sp, plant.genus, plant.family, k.type, k.type.s, k.type.ss, color, diam, LN)
```

## 1.3 Load network and survey data
```{r message=FALSE}
### BB-FL visitation data
net <- read_csv("../data/processed_data/network.csv") %>%
  filter(caste != "male",
         !bb.sp %in% c("humi", "hypn", "lapi")) %>% # too rare
  dplyr::select(site, date, bb.sp, plant.sp) # drop unused columns

### Floral survey data
survey <- read_csv("../data/processed_data/floral_survey.csv") %>%
  semi_join(net, by = "plant.sp") %>% # Consider only visited plant species (anytime anywhere, not specifically for a given site*date)
  dplyr::select(site, date, plant.sp, flower.cover) %>%
  group_by(plant.sp) %>%
  mutate(present = case_when(
    max(flower.cover) > 0 ~ TRUE,
    max(flower.cover) == 0 ~ FALSE
  )) %>%
  filter(present == TRUE) # Ignore plants that never had any flower cover

vis_sp <- net %>%
  left_join(fl_tax) %>%
  select(plant.sp, plant.genus, plant.family) %>%
  unique()

surv_sp <- survey %>%
  left_join(fl_tax) %>%
  select(plant.sp, plant.genus, plant.family) %>%
  unique()

missing <- anti_join(vis_sp, surv_sp)
```

## 1.4 Calculate FL and BB abundance / diversity
```{r echo=FALSE, include=TRUE, message=FALSE}
# Per-transect BB abundance
## full dataset
bb_abund_sp <- net %>%
  dplyr::select(site, date, bb.sp) %>%
  group_by(site, date, bb.sp) %>% # 
  summarize(bb.abund = n()) %>% # get per transect bb abundance
  ungroup() %>%
  complete(site, date, bb.sp, fill = list(bb.abund = 0)) %>% # complete implicit absences
  semi_join(net, by = c("site", "date")) %>% # drop site-dates that did not exist in the data set
  left_join(site_data) %>% # add site data
  mutate(year = factor(year(date)),
         yday = yday(date),
         bb.sp = factor(bb.sp),
         management = factor(management),
         site = factor(site),
         transect = factor(transect)) %>%
  left_join(bb_traits) %>%
  mutate(tongue = factor(pbl.w.class2))

### Per-transect floral abund by species
fl_abund_sp <- survey %>%
  group_by(site, date, plant.sp) %>% # group by plant.sp*transect
  summarize(fl.abund = sum(flower.cover)) %>% # get plant.sp*transect total flower cover; this step really shouldn't be required, but in a few cases, there is more than one abundance entry for the same species on a given transect; not sure how that happened, but it's in the original data that Fabrice sent me; alternatively, I could drop the dups or average them. We'll see what Kalli and Fabrice have to say.
  group_by(site, date) %>%
  mutate(prop.fl.abund = fl.abund/sum(fl.abund)) %>% # get the proportional flower cover for each species.
  ungroup() %>%
  #full_join(dplyr::select(net, site, date, plant.sp), by = c("site", "date", "plant.sp")) %>% # add the handful of species occurring in the visitation network but missing from the survey data
  complete(site, date, plant.sp, fill = list(fl.abund = 0, prop.fl.abund = 0)) %>%
  semi_join(survey, by = c("site", "date")) %>%  # only include site-dates that actually occurred
  unique() %>%
  mutate(year = factor(year(date))) %>%
  left_join(fl_traits)
```

## 1.5 Calculate binary "preference" for each plant genus by each bumble bee species; score species 0 or 1 based on whether they were visited
```{r message=FALSE}
weights <- fl_abund_sp %>%
  select(plant.sp, plant.genus) %>%
  unique() %>%
  left_join(select(net, bb.sp, plant.sp)) %>%
  select(plant.genus, bb.sp) %>%
  distinct() %>%
  mutate(visited = rep(1, n())) %>%
  pivot_wider(values_from = visited, names_from = bb.sp, values_fill = 0) %>%
  pivot_longer(cols = 2:14, values_to = "visited", names_to = "bb.sp")

weighted_fl_abund <- fl_abund_sp %>%
  group_by(site, date, plant.genus) %>%
  summarize(flower.cover = sum(fl.abund)) %>%
  left_join(weights) %>%
  mutate(weighted.flower.cover = flower.cover*visited) %>%
  group_by(site, date, bb.sp) %>%
  summarize(total.weighted.flower.cover = sum(weighted.flower.cover)) %>%
  left_join(site_data) %>%
  select(site, date, elev.mean, bb.sp, total.weighted.flower.cover) %>%
  mutate(yday = yday(date),
         bb.sp = factor(bb.sp),
         site = factor(site),
         year = factor(year(date)))
```

# 2. Analyses

## 2.1 BB elevation distribution
```{r}
bb_range <- net %>%
  left_join(site_data) %>%
  group_by(site, date, elev.mean, tree_line, bb.sp) %>%
  summarize(abund = n()) %>%
  group_by(site, elev.mean, tree_line, bb.sp) %>%
  summarize(abund = max(abund)) %>%
  filter(abund > 0) %>%
  group_by(bb.sp) %>%
  mutate(elev.floor = min(elev.mean),
         elev.ceiling = max(elev.mean),
         elev.range = elev.ceiling - elev.floor,
         elev.med = elev.floor + ((elev.ceiling - elev.floor)/2))

ggplot(bb_range, aes(reorder(bb.sp, elev.med), elev.mean)) +
  geom_line(size = 2, alpha = 0.25) +
  geom_point(aes(size = abund, color = tree_line), alpha = 0.5) +
  theme_light(14) +
  scale_size_continuous(name = "Peak abundance",
                        breaks = c(1, 10, 20, 40, 80)) +
  scale_color_discrete(name = "Tree line") +
  xlab("Bumble bee species") +
  ylab("Elevation (m)") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

ggsave("../output/bb_range_ms1.pdf")
ggsave("../output/bb_range_ms1.png")

# What happens if I plot max abundance by elevation?

bb_range2 <- net %>%
  left_join(site_data) %>%
  group_by(site, date, elev.mean, tree_line, bb.sp) %>%
  summarize(abund = n()) %>%
  group_by(site, elev.mean, tree_line, bb.sp) %>%
  summarize(abund = max(abund))

ggplot(filter(bb_range2, bb.sp != "humi"), aes(elev.mean, abund)) +
  geom_point() +
  geom_smooth(method = "gam", formula = y ~ s(x, k = 6)) +
  facet_wrap(~bb.sp, scales = "free_y", ncol = 4) +
  geom_vline(xintercept = 1500, color = "red", linetype = "dashed")

# Overall bumble bee abundance
bb_abund_sum <- bb_abund_sp %>%
  group_by(site, bb.sp) %>%
  summarize(cum.abund = sum(bb.abund),
            mean.abund = mean(bb.abund)) %>%
  group_by(site) %>%
  mutate(prop.abund = cum.abund/sum(cum.abund)) %>%
  group_by(bb.sp) %>%
  mutate(mean.prop.abund = mean(prop.abund))

ggplot(bb_abund_sum, aes(site, prop.abund, fill = bb.sp)) +
  geom_col()

ggplot(bb_abund_sum, aes(bb.sp, mean.abund)) +
  geom_boxplot()

ggplot(bb_abund_sum, aes(reorder(bb.sp, -mean.prop.abund), prop.abund)) +
  geom_boxplot(fill = "gray85") +
  theme_light() +
  ylab("Proportional abundance") +
  xlab("Bumble bee species")

ggsave("../output/bb_abundsum.pdf")
ggsave("../output/bb_abundsum.png")
```

## 2.2. Aggregate analysis of max BB abundance and mean FL abundance  ~ elevation

### 2.2.1 Calculate max and mean abundances within site and year
```{r}
# Get mean total weighted floral abundance by BB sp
weighted_fl_abund_mean <- weighted_fl_abund %>%
  group_by(site, bb.sp, year) %>%
  summarize(mean.fl.abund = mean(total.weighted.flower.cover),
            max.fl.abund = max(total.weighted.flower.cover))

# Get max BB abundance by sp
bb_abund_sp_max <- bb_abund_sp %>%
  group_by(site, elev.mean, year, bb.sp) %>%
  summarize(max.bb.abund = max(bb.abund),
            mean.bb.abund = mean(bb.abund))

# Join
abund_sum <- full_join(weighted_fl_abund_mean, bb_abund_sp_max) %>%  
  unite(sp.year, c(bb.sp, year), remove = FALSE) %>%
  mutate(site = factor(site),
         bb.sp = factor(bb.sp),
         sp.year = factor(sp.year)) %>%
  unite(bb.sp.year, c(bb.sp, year), sep = "_", remove = FALSE) %>%
  mutate(bb.sp.year = factor(bb.sp.year)) %>%
  left_join(site_data)
```

### 2.2.2 Preliminary visualizations
```{r}
## mean floral abundance
ggplot(abund_sum, aes(elev.mean, mean.fl.abund, color = year)) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = "gam", formula = y ~ s(x, k = 8), se = FALSE) +
  #facet_wrap(~bb.sp, scales = "free_y") +
  geom_vline(xintercept = 1500)

## log(mean floral abundance)
ggplot(abund_sum, aes(elev.mean, log(mean.fl.abund), color = year)) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = "gam", formula = y ~ s(x, k = 8), se = FALSE) +
  facet_wrap(~bb.sp) +
  geom_vline(xintercept = 1500)

## Max floral abundance)
ggplot(abund_sum, aes(elev.mean, max.fl.abund, color = year)) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = "gam", formula = y ~ s(x, k = 8), se = FALSE) +
  facet_wrap(~bb.sp) +
  geom_vline(xintercept = 1500)

## Max BB abundance
ggplot(abund_sum, aes(elev.mean, max.bb.abund, color = year)) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = "gam", formula = y ~ s(x, k = 8),  se = FALSE) +
  facet_wrap(~bb.sp, scales = "free_y") +
  geom_vline(xintercept = 1500)

## Max  abundance ~ mean BB abundance, to see how interchangeable they are.
## They are almost perfectly correlated, and I think max abund is more informative
ggplot(abund_sum, aes(mean.bb.abund, max.bb.abund, color = year)) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = "gam", formula = y ~ s(x, k = 8),  se = FALSE) +
  facet_wrap(~bb.sp, scales = "free_y")

## Are max FL and mean FL similarly interchangeable? Yes.
ggplot(abund_sum, aes(mean.fl.abund, max.fl.abund, color = year)) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = "gam", formula = y ~ s(x, k = 8),  se = FALSE) +
  facet_wrap(~bb.sp, scales = "free_y")

## Max BB abundance ~ mean FL abundance
ggplot(abund_sum, aes(log(mean.fl.abund), max.bb.abund)) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = "gam", formula = y ~ s(x, k = 8),  se = FALSE) +
  facet_wrap(~bb.sp, scales = "free_y")
```

### 2.2.3 Abundance models
I'm choosing to use site as a random intercept effect in these models because I think that is the correct way to specify a model involving repeated observations (in this case, bumble bee species) within site. The effect of this decision is that the elevation smoother is less influenced by individual sites and converges more strongly toward the mean. This creates less exciting curves, but I think it is a more principled way to model the data and less prone to overfitting. I can have my cake and eat it too, though, by plotting the actual data over the GAM fit so that the strong patterns of certain sites are visible. Then I can talk about the conditional elevation effect and the importance of certain peaks that are likely to be idiosyncrasies of my system (i.e. it would be overfitting to call them elevation effects).

#### 2.2.3.1 Bumble bees
```{r}
# Call model
gam_max_bb_abund_00 <- gam(max.bb.abund ~ 
                             s(bb.sp, bs = "re") +
                             s(bb.sp.year, bs = "re") + # fixed intercept for bb.sp:year
                             s(elev.mean, by = bb.sp, k = 8), # smoother for elev by BB sp.
                           method = "REML",
                           family = "tw", # Tweedie distribution performs well
                           select = TRUE,
                           data = abund_sum) %>% getViz()

gam_max_bb_abund_00_ilink <- family(gam_max_bb_abund_00)$linkinv # get inverse link function

check.gamViz(gam_max_bb_abund_00) # evaluate model
gam_max_bb_abund_00_sum <- summary(gam_max_bb_abund_00) # summarize model
print(plot(gam_max_bb_abund_00, allTerms = TRUE), pages = 1) # preliminary visualization

# Plot smooths and year effect on the link scale
## Individual plots
gers_abund <- plot(sm(gam_max_bb_abund_00, 3)) + l_fitLine() + l_ciLine() + 
  labs(x = NULL, y = NULL, title = "gers")
hort_abund <- plot(sm(gam_max_bb_abund_00, 4)) + l_fitLine() + l_ciLine() + 
  labs(x = NULL, y = NULL, title = "hort")
jone_abund <- plot(sm(gam_max_bb_abund_00, 5)) + l_fitLine() + l_ciLine() + 
  labs(x = NULL, y = NULL, title = "jone")
mend_abund <- plot(sm(gam_max_bb_abund_00, 6)) + l_fitLine() + l_ciLine() + 
  labs(x = NULL, y = NULL, title = "mend")
mont_abund <- plot(sm(gam_max_bb_abund_00, 7)) + l_fitLine() + l_ciLine() + 
  labs(x = NULL, y = NULL, title = "mont")
muci_abund <- plot(sm(gam_max_bb_abund_00, 8)) + l_fitLine() + l_ciLine() + 
  labs(x = NULL, y = NULL, title = "muci")
pasc_abund <- plot(sm(gam_max_bb_abund_00, 9)) + l_fitLine() + l_ciLine() + 
  labs(x = NULL, y = NULL, title = "pasc")
prat_abund <- plot(sm(gam_max_bb_abund_00, 10)) + l_fitLine() + l_ciLine() + 
  labs(x = NULL, y = NULL, title = "prat")
psit_abund <- plot(sm(gam_max_bb_abund_00, 11)) + l_fitLine() + l_ciLine() + 
  labs(x = NULL, y = NULL, title = "psit")
pyre_abund <- plot(sm(gam_max_bb_abund_00, 12)) + l_fitLine() + l_ciLine() + 
  labs(x = NULL, y = NULL, title = "pyre")
soro_abund <- plot(sm(gam_max_bb_abund_00, 13)) + l_fitLine() + l_ciLine() + 
  labs(x = NULL, y = NULL, title = "soro")
telu_abund <- plot(sm(gam_max_bb_abund_00, 14)) + l_fitLine() + l_ciLine() + 
  labs(x = NULL, y = NULL, title = "telu")
wurf_abund <- plot(sm(gam_max_bb_abund_00, 15)) + l_fitLine() + l_ciLine() + 
  labs(x = NULL, y = NULL, title = "wurf")

## Arrange significant smooths on grid plot
grid_bb_abund <- gridPrint(gers_abund, mend_abund, mont_abund, muci_abund, 
                           pasc_abund, prat_abund, psit_abund, pyre_abund, 
                           soro_abund, telu_abund, wurf_abund, ncol = 4)

pdf("../output/gam_max_bb_abund_grid.pdf", width = 8, height = 6)
plot(grid_bb_abund)
dev.off()

png("../output/gam_max_bb_abund_grid.png", res = 300, width = 8, height = 6, units = "in")
plot(grid_bb_abund)
dev.off()
```

#### 2.2.3.2 Floral resources
```{r}
gam_mean_fl_abund_00 <- gam(mean.fl.abund ~
                              s(elev.mean, bb.sp, bs = "fs", k = 8) +
                              s(elev.mean, year, bs = "fs", k = 8),
                           method = "REML",
                           family = "tw",
                           select = TRUE,
                           data = abund_sum) %>% getViz()

check.gamViz(gam_mean_fl_abund_00) # evaluate model
gam_mean_fl_abund_00_sum <- summary(gam_mean_fl_abund_00) # summarize model
print(plot(gam_mean_fl_abund_00, allTerms = TRUE), pages = 1) # preliminary visualization

# Plot on the link scale
elev_effect <- plot(sm(gam_mean_fl_abund_00, 2)) + 
  l_fitLine(alpha = 0.6, size = 1) + 
  labs(x = "Elevation (m.s.l.)", y = "Conditional effect (link scale)", color = "Year")

bb.sp_effect <- plot(sm(gam_mean_fl_abund_00, 1)) +
  l_fitLine(alpha = 0.6, size = 1) + 
  labs(x = "Elevation (m.s.l.)", y = NULL, color = "Species")

grid_fl_abund <- gridPrint(elev_effect, bb.sp_effect, ncol = 2)

pdf("../output/gam_mean_fl_abund_grid.pdf", width = 7, height = 3.5)
plot(grid_fl_abund)
dev.off()

png("../output/gam_mean_fl_abund_grid.png", res = 300, units = "in", width = 7, height = 3.5)
plot(grid_fl_abund)
dev.off()
```