### 2021-03-17 code/ms1_abundance.Rmd
update

---
title: "Abundance/diversity GAM"
output: html_document
---

# 1. Prep work
## 1.1 Prepare environment
```{r setup, include=FALSE, echo=FALSE}
library(mgcv)
library(mgcViz)
library(gratia)
library(patchwork)
library(tidymv)
library(vegan)
library(ggvegan)
library(ggrepel)
library(ggforce)
library(tidyverse)
library(lubridate)
```

## 1.2 Load site and trait data
```{r, include=FALSE, echo=FALSE}
site_data <- read_csv("../data/processed_data/site_data.csv") %>%
  dplyr::select(-c(slope.calc, slope.est, elev.min, elev.max, temp.mean)) %>% # drop these variables
  mutate(elev.class2 = factor(elev.class2, levels = c("low", "mid", "high")),
         site = factor(site),
         tree_line = factor(tree_line)) # turn elev.class into an ordered factor and site into a regular factor

# BB traits
bb_traits <- read_csv("../data/processed_data/bb_traits.csv")

# FL traits
## family-level taxonomy
fl_tax <- read_csv("../data/processed_data/floral_tax.csv")

## Kugler morphotypes (accessed through Bioflor)
fl_ktype <- read_csv("../data/processed_data/floral_k_type.csv")

## Flower colors (accessed through BioFlor)
fl_color <- read_csv("../data/processed_data/floral_color.csv")

## diameter and corolla depth data from Gita Benadi
fl_morph <- read_csv("../data/processed_data/flower_morphology_GitaBenadi.csv") %>% 
  dplyr::select(SampleID, plant.family = Family, plant.sp = Species, 
                Unit, Flowers_per_unit, diam = Diameter, LN) %>%
  group_by(plant.sp) %>%
  summarize(diam = mean(diam), LN = mean(LN))

## verbal description of Kugler morphotypes
fl_ktype_key <- read_delim("../data/processed_data/kugler_key.tsv", delim = "\t", col_names = FALSE) %>%
  dplyr::select(k.type = 1, description = 2)

## join all floral trait tables
fl_traits <- fl_tax %>%
  full_join(fl_ktype) %>%
  full_join(fl_color) %>%
  full_join(fl_morph) %>%
  dplyr::select(plant.sp, plant.genus, plant.family, k.type, k.type.s, k.type.ss, color, diam, LN)
```

## 1.3 Load network and survey data
```{r message=FALSE}
### BB-FL visitation data
net <- read_csv("../data/processed_data/network.csv") %>%
  filter(caste != "male",
         !bb.sp %in% c("humi", "hypn", "lapi")) %>% # too rare
  dplyr::select(site, date, bb.sp, plant.sp) # drop unused columns

### Floral survey data
survey <- read_csv("../data/processed_data/floral_survey.csv") %>%
  semi_join(net, by = "plant.sp") %>% # Consider only visited plant species (anytime anywhere, not specifically for a given site*date)
  dplyr::select(site, date, plant.sp, flower.cover) %>%
  group_by(plant.sp) %>%
  mutate(present = case_when(
    max(flower.cover) > 0 ~ TRUE,
    max(flower.cover) == 0 ~ FALSE
  )) %>%
  filter(present == TRUE) # Ignore plants that never had any flower cover

vis_sp <- net %>%
  left_join(fl_tax) %>%
  select(plant.sp, plant.genus, plant.family) %>%
  unique()

surv_sp <- survey %>%
  left_join(fl_tax) %>%
  select(plant.sp, plant.genus, plant.family) %>%
  unique()

missing <- anti_join(vis_sp, surv_sp)
```

## 1.4 Calculate FL and BB abundance / diversity
```{r echo=FALSE, include=TRUE, message=FALSE}
# Per-transect BB abundance
## full dataset
bb_abund_sp <- net %>%
  dplyr::select(site, date, bb.sp) %>%
  group_by(site, date, bb.sp) %>% # 
  summarize(bb.abund = n()) %>% # get per transect bb abundance
  ungroup() %>%
  complete(site, date, bb.sp, fill = list(bb.abund = 0)) %>% # complete implicit absences
  semi_join(net, by = c("site", "date")) %>% # drop site-dates that did not exist in the data set
  left_join(site_data) %>% # add site data
  mutate(year = factor(year(date)),
         yday = yday(date),
         bb.sp = factor(bb.sp),
         management = factor(management),
         site = factor(site),
         transect = factor(transect)) %>%
  left_join(bb_traits) %>%
  mutate(tongue = factor(pbl.w.class2))

### Per-transect floral abund by species
fl_abund_sp <- survey %>%
  group_by(site, date, plant.sp) %>% # group by plant.sp*transect
  summarize(fl.abund = sum(flower.cover)) %>% # get plant.sp*transect total flower cover; this step really shouldn't be required, but in a few cases, there is more than one abundance entry for the same species on a given transect; not sure how that happened, but it's in the original data that Fabrice sent me; alternatively, I could drop the dups or average them. We'll see what Kalli and Fabrice have to say.
  group_by(site, date) %>%
  mutate(prop.fl.abund = fl.abund/sum(fl.abund)) %>% # get the proportional flower cover for each species.
  ungroup() %>%
  #full_join(dplyr::select(net, site, date, plant.sp), by = c("site", "date", "plant.sp")) %>% # add the handful of species occurring in the visitation network but missing from the survey data
  complete(site, date, plant.sp, fill = list(fl.abund = 0, prop.fl.abund = 0)) %>%
  semi_join(survey, by = c("site", "date")) %>%  # only include site-dates that actually occurred
  unique() %>%
  mutate(year = factor(year(date))) %>%
  left_join(fl_traits)
```

## 1.5 Calculate binary "preference" for each plant genus by each bumble bee species; score species 0 or 1 based on whether they were visited
```{r message=FALSE}
weights <- fl_abund_sp %>%
  select(plant.sp, plant.genus) %>%
  unique() %>%
  left_join(select(net, bb.sp, plant.sp)) %>%
  select(plant.genus, bb.sp) %>%
  distinct() %>%
  mutate(visited = rep(1, n())) %>%
  pivot_wider(values_from = visited, names_from = bb.sp, values_fill = 0) %>%
  pivot_longer(cols = 2:14, values_to = "visited", names_to = "bb.sp")

weighted_flfl_abund <- fl_abund_sp %>%
  group_by(site, date, plant.genus) %>%
  summarize(flower.cover = sum(fl.abund)) %>%
  left_join(weights) %>%
  mutate(weighted.flower.cover = flower.cover*visited) %>%
  group_by(site, date, bb.sp) %>%
  summarize(total.weighted.flower.cover = sum(weighted.flower.cover)) %>%
  left_join(site_data) %>%
  select(site, date, elev.mean, bb.sp, total.weighted.flower.cover) %>%
  mutate(yday = yday(date),
         bb.sp = factor(bb.sp),
         site = factor(site),
         year = factor(year(date)))
```

# 2. Analyses

## 2.1 BB elevation distribution
```{r}
bb_range <- net %>%
  left_join(site_data) %>%
  group_by(site, date, elev.mean, tree_line, bb.sp) %>%
  summarize(abund = n()) %>%
  group_by(site, elev.mean, tree_line, bb.sp) %>%
  summarize(abund = max(abund)) %>%
  filter(abund > 0) %>%
  group_by(bb.sp) %>%
  mutate(elev.floor = min(elev.mean),
         elev.ceiling = max(elev.mean),
         elev.range = elev.ceiling - elev.floor,
         elev.med = elev.floor + ((elev.ceiling - elev.floor)/2))

ggplot(bb_range, aes(reorder(bb.sp, elev.med), elev.mean)) +
  geom_line(size = 2, alpha = 0.25) +
  geom_point(aes(size = abund, color = tree_line), alpha = 0.5) +
  theme_light(14) +
  scale_size_continuous(name = "Peak abundance",
                        breaks = c(1, 10, 20, 40, 80)) +
  scale_color_discrete(name = "Tree line") +
  xlab("Bumble bee species") +
  ylab("Elevation (m)") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

ggsave("../output/bb_range_ms1.pdf")
ggsave("../output/bb_range_ms1.png")

# What happens if I plot max abundance by elevation?

bb_range2 <- net %>%
  left_join(site_data) %>%
  group_by(site, date, elev.mean, tree_line, bb.sp) %>%
  summarize(abund = n()) %>%
  group_by(site, elev.mean, tree_line, bb.sp) %>%
  summarize(abund = max(abund))

ggplot(filter(bb_range2, bb.sp != "humi"), aes(elev.mean, abund)) +
  geom_point() +
  geom_smooth(method = "gam", formula = y ~ s(x, k = 6)) +
  facet_wrap(~bb.sp, scales = "free_y", ncol = 4) +
  geom_vline(xintercept = 1500, color = "red", linetype = "dashed")

# Overall bumble bee abundance
bb_abund_sum <- bb_abund_sp %>%
  group_by(site, bb.sp) %>%
  summarize(cum.abund = sum(bb.abund),
            mean.abund = mean(bb.abund)) %>%
  group_by(site) %>%
  mutate(prop.abund = cum.abund/sum(cum.abund)) %>%
  group_by(bb.sp) %>%
  mutate(mean.prop.abund = mean(prop.abund))

ggplot(bb_abund_sum, aes(site, prop.abund, fill = bb.sp)) +
  geom_col()

ggplot(bb_abund_sum, aes(bb.sp, mean.abund)) +
  geom_boxplot()

ggplot(bb_abund_sum, aes(reorder(bb.sp, -mean.prop.abund), prop.abund)) +
  geom_boxplot(fill = "gray85") +
  theme_light() +
  ylab("Proportional abundance") +
  xlab("Bumble bee species")

ggsave("../output/bb_abundsum.pdf")
ggsave("../output/bb_abundsum.png")
```

## 2.2. Aggregate analysis of max BB abundance and mean FL abundance  ~ elevation

### 2.2.1 Calculate max and mean abundances within site and year
```{r}
# Get mean total weighted floral abundance by BB sp
weighted_fl_abund_mean <- weighted_flfl_abund %>%
  group_by(site, bb.sp, year) %>%
  summarize(mean.fl.abund = mean(total.weighted.flower.cover),
            max.fl.abund = max(total.weighted.flower.cover))

# Get max BB abundance by sp
bb_abund_sp_max <- bb_abund_sp %>%
  group_by(site, elev.mean, year, bb.sp) %>%
  summarize(max.bb.abund = max(bb.abund),
            mean.bb.abund = mean(bb.abund))

# Join
abund_sum <- full_join(weighted_fl_abund_mean, bb_abund_sp_max) %>%  
  unite(sp.year, c(bb.sp, year), remove = FALSE) %>%
  mutate(site = factor(site),
         bb.sp = factor(bb.sp),
         sp.year = factor(sp.year)) %>%
  unite(bb.sp.year, c(bb.sp, year), sep = "_", remove = FALSE) %>%
  mutate(bb.sp.year = factor(bb.sp.year)) %>%
  left_join(site_data)
```

### 2.2.2 Call models
```{r}
# BB
gam_max_bb_abund_00 <- gam(max.bb.abund ~ 
                             s(bb.sp, bs = "re") + # random intercept for bb.sp
                             s(bb.sp.year, bs = "re") + # random intercept for bb.sp:year
                             s(elev.mean, by = bb.sp, k = 8), # smoother for elev by bb.sp.
                           method = "REML",
                           family = "tw", # Tweedie distribution performs well
                           select = TRUE,
                           data = abund_sum) %>% getViz()

# FL
gam_mean_fl_abund_00 <- gam(mean.fl.abund ~
                              year +
                              s(bb.sp, bs = "re") +
                              s(elev.mean, by = bb.sp, k = 8),
                           method = "REML",
                           family = "tw",
                           select = TRUE,
                           data = abund_sum) %>% getViz()
```

### 2.2.3 Evaluate model
```{r}
# BB
check.gamViz(gam_max_bb_abund_00) # evaluate model
gam_max_bb_abund_00_sum <- summary(gam_max_bb_abund_00) # summarize model
print(plot(gam_max_bb_abund_00, allTerms = TRUE), pages = 1) # preliminary visualization

# FL
check.gamViz(gam_mean_fl_abund_00) # evaluate model
gam_mean_fl_abund_00_sum <- summary(gam_mean_fl_abund_00) # summarize model
print(plot(gam_mean_fl_abund_00, allTerms = TRUE), pages = 1) # preliminary visualization
```

### 2.2.4 Visualize

#### 2.2.4.1 BB
```{r}
# Plot smooths and year effect on the link scale

## Individual plots
gers_abund <- plot(sm(gam_max_bb_abund_00, 3)) + l_fitLine() + l_ciLine() + 
  labs(x = NULL, y = NULL, title = "gers")
hort_abund <- plot(sm(gam_max_bb_abund_00, 4)) + l_fitLine() + l_ciLine() + 
  labs(x = NULL, y = NULL, title = "hort")
jone_abund <- plot(sm(gam_max_bb_abund_00, 5)) + l_fitLine() + l_ciLine() + 
  labs(x = NULL, y = NULL, title = "jone")
mend_abund <- plot(sm(gam_max_bb_abund_00, 6)) + l_fitLine() + l_ciLine() + 
  labs(x = NULL, y = NULL, title = "mend")
mont_abund <- plot(sm(gam_max_bb_abund_00, 7)) + l_fitLine() + l_ciLine() + 
  labs(x = NULL, y = NULL, title = "mont")
muci_abund <- plot(sm(gam_max_bb_abund_00, 8)) + l_fitLine() + l_ciLine() + 
  labs(x = NULL, y = NULL, title = "muci")
pasc_abund <- plot(sm(gam_max_bb_abund_00, 9)) + l_fitLine() + l_ciLine() + 
  labs(x = NULL, y = NULL, title = "pasc")
prat_abund <- plot(sm(gam_max_bb_abund_00, 10)) + l_fitLine() + l_ciLine() + 
  labs(x = NULL, y = NULL, title = "prat")
psit_abund <- plot(sm(gam_max_bb_abund_00, 11)) + l_fitLine() + l_ciLine() + 
  labs(x = NULL, y = NULL, title = "psit")
pyre_abund <- plot(sm(gam_max_bb_abund_00, 12)) + l_fitLine() + l_ciLine() + 
  labs(x = NULL, y = NULL, title = "pyre")
soro_abund <- plot(sm(gam_max_bb_abund_00, 13)) + l_fitLine() + l_ciLine() + 
  labs(x = NULL, y = NULL, title = "soro")
telu_abund <- plot(sm(gam_max_bb_abund_00, 14)) + l_fitLine() + l_ciLine() + 
  labs(x = NULL, y = NULL, title = "telu")
wurf_abund <- plot(sm(gam_max_bb_abund_00, 15)) + l_fitLine() + l_ciLine() + 
  labs(x = NULL, y = NULL, title = "wurf")

## Arrange smooths on grid plot

grid_bb_abund <- gridPrint(gers_abund, hort_abund, jone_abund, mend_abund, mont_abund, muci_abund, 
                           pasc_abund, prat_abund, psit_abund, pyre_abund, 
                           soro_abund, telu_abund, wurf_abund, ncol = 4)

pdf("../output/gam_bb_abund_grid.pdf", width = 8, height = 6)
plot(grid_bb_abund)
dev.off()

png("../output/gam_bb_abund_grid.png", res = 300, width = 8, height = 6, units = "in")
plot(grid_bb_abund)
dev.off()

# Extract predictions and plot on response scale

gam_max_bb_abund_00_ilink <- family(gam_max_bb_abund_00)$linkinv # get inverse link function

gam_max_bb_abund_00_pred <- predict.gam(gam_max_bb_abund_00, # get model predictions
                                        type = "terms") %>% 
  as.data.frame() %>% # convert to data frame
  as_tibble() %>% # convert to tibble
  select(bb.sp = 1, bb.sp.year = 2, gers = 3, # rename
         hort = 4, jone = 5, mend = 6, mont = 7, 
         muci = 8, pasc = 9, prat = 10, psit = 11, 
         pyre = 12, soro = 13, telu = 14, wurf = 15) %>%
  mutate(fit = # get cumulative fit
           bb.sp + bb.sp.year + gers + 
           hort + jone + mend + mont + 
           muci + pasc + prat + psit + 
           pyre + soro + telu + wurf 
           ) %>%
  select(fit) %>%
  mutate(fit.trans = gam_max_bb_abund_00_ilink(fit)) %>% # back-transform to response scale
  bind_cols(abund_sum)

ggplot(gam_max_bb_abund_00_pred, aes(elev.mean, fit.trans, linetype = year)) + # map year to linetype
  geom_point(aes(elev.mean, max.bb.abund, color = tree_line), inherit.aes = FALSE, alpha = 0.5) + # add original data and map tree line to color
  geom_line(size = 0.5, alpha = 0.5) +
  facet_wrap(~bb.sp, scales = "free_y") +
  theme_light() +
  labs(x = "Elevation (m.s.l.)", 
       y = bquote("Peak abundance"),
       color = "Tree line",
       shape = "Year",
       linetype = "Year")

ggsave("../output/gam_bb_abund.pdf")
ggsave("../output/gam_bb_abund.png")
```

#### 2.2.4.2 FL
```{r}
# Plot on the link scale

## Individual plots

gers_fl_abund <- plot(sm(gam_mean_fl_abund_00, 2)) + l_fitLine() + l_ciLine() + 
  labs(x = NULL, y = NULL, title = "gers")
hort_fl_abund <- plot(sm(gam_mean_fl_abund_00, 3)) + l_fitLine() + l_ciLine() + 
  labs(x = NULL, y = NULL, title = "hort")
jone_fl_abund <- plot(sm(gam_mean_fl_abund_00, 4)) + l_fitLine() + l_ciLine() + 
  labs(x = NULL, y = NULL, title = "jone")
mend_fl_abund <- plot(sm(gam_mean_fl_abund_00, 5)) + l_fitLine() + l_ciLine() + 
  labs(x = NULL, y = NULL, title = "mend")
mont_fl_abund <- plot(sm(gam_mean_fl_abund_00, 6)) + l_fitLine() + l_ciLine() + 
  labs(x = NULL, y = NULL, title = "mont")
muci_fl_abund <- plot(sm(gam_mean_fl_abund_00, 7)) + l_fitLine() + l_ciLine() + 
  labs(x = NULL, y = NULL, title = "muci")
pasc_fl_abund <- plot(sm(gam_mean_fl_abund_00, 8)) + l_fitLine() + l_ciLine() + 
  labs(x = NULL, y = NULL, title = "pasc")
prat_fl_abund <- plot(sm(gam_mean_fl_abund_00, 9)) + l_fitLine() + l_ciLine() + 
  labs(x = NULL, y = NULL, title = "prat")
psit_fl_abund <- plot(sm(gam_mean_fl_abund_00, 10)) + l_fitLine() + l_ciLine() + 
  labs(x = NULL, y = NULL, title = "psit")
pyre_fl_abund <- plot(sm(gam_mean_fl_abund_00, 11)) + l_fitLine() + l_ciLine() + 
  labs(x = NULL, y = NULL, title = "pyre")
soro_fl_abund <- plot(sm(gam_mean_fl_abund_00, 12)) + l_fitLine() + l_ciLine() + 
  labs(x = NULL, y = NULL, title = "soro")
telu_fl_abund <- plot(sm(gam_mean_fl_abund_00, 13)) + l_fitLine() + l_ciLine() + 
  labs(x = NULL, y = NULL, title = "telu")
wurf_fl_abund <- plot(sm(gam_mean_fl_abund_00, 14)) + l_fitLine() + l_ciLine() + 
  labs(x = NULL, y = NULL, title = "wurf")
year_fl_effect <- plot(pterm(gam_mean_fl_abund_00, 1)) + l_fitPoints() + l_ciBar() + 
  labs(x = NULL, y = NULL, title = "Year")

## Arrange significant smooths on grid plot

grid_fl_abund <- gridPrint(gers_fl_abund, hort_fl_abund, jone_fl_abund, mend_fl_abund, mont_fl_abund, muci_fl_abund, 
                           pasc_fl_abund, prat_fl_abund, psit_fl_abund, pyre_fl_abund, 
                           soro_fl_abund, telu_fl_abund, wurf_fl_abund, year_fl_effect, ncol = 4)

pdf("../output/gam_fl_abund_grid.pdf", width = 8, height = 6)
plot(grid_fl_abund)
dev.off()

png("../output/gam_fl_abund_grid.png", res = 300, width = 8, height = 6, units = "in")
plot(grid_fl_abund)
dev.off()

# Extract predictions and plot on response scale

gam_mean_fl_abund_00_ilink <- family(gam_mean_fl_abund_00)$linkinv # get inverse link function to transform to the response scale

gam_mean_fl_abund_00_pred <- predict.gam(gam_mean_fl_abund_00, # get model predictions
                                        type = "terms") %>% 
  as.data.frame() %>% # convert to data frame
  as_tibble() %>% # convert to tibble
  select(year = 1, bb.sp = 2, gers = 3, # rename
         hort = 4, jone = 5, mend = 6, mont = 7, 
         muci = 8, pasc = 9, prat = 10, psit = 11, 
         pyre = 12, soro = 13, telu = 14, wurf = 15) %>%
  mutate(fit = # get cumulative fit
           year + bb.sp + gers + 
           hort + jone + mend + mont + 
           muci + pasc + prat + psit + 
           pyre + soro + telu + wurf 
           ) %>%
  select(fit) %>%
  mutate(fit.trans = gam_max_bb_abund_00_ilink(fit)) %>% # back-transform to response scale
  bind_cols(na.omit(abund_sum)) # the na.omit() drops the rows where floral surveying data were missing

ggplot(gam_mean_fl_abund_00_pred, aes(elev.mean, fit.trans, linetype = year)) + # map year to linetype
  geom_point(aes(elev.mean, mean.fl.abund, color = tree_line), inherit.aes = FALSE, alpha = 0.5) + # add original data and map tree line to color
  geom_line(size = 0.5, alpha = 0.5) +
  facet_wrap(~bb.sp, scales = "free_y") +
  theme_light() +
  labs(x = "Elevation (m.s.l.)", 
       y = bquote("Peak abundance"),
       color = "Tree line",
       shape = "Year",
       linetype = "Year")

ggsave("../output/gam_fl_abund.pdf")
ggsave("../output/gam_fl_abund.png")
```