### 2021-03-26 ms_2/ms_2_v2.Rmd
almost done ms2...

---
title: "Potential competition in alpine bumble bees is mitigated by imperfect trait-matching and subpartitioning within floral morphotypes"

author: Douglas B. Sponsler^1^*, Fabrice Requier^2^, Katharina Kallnik^1^, Alice Claßen^1^, A. Fabienne Maihoff^1^, Ingolf Steffan-Dewenter^1^

# date: `r Sys.Date()`

output:
  bookdown::word_document2:
    fig_caption: yes
    fig_width: 7
    fig_height: 7
    df_print: kable
    keep_md: yes
    number_sections: no
    toc: no
    reference_docx: Word_template.docx

bibliography: 
  - references.bib
#  - knitcitations.bib
csl: ecology.csl
# csl: https://www.zotero.org/styles/ecology

---

^1^ Department of Animal Ecology and Tropical Biology, Biocenter, University of Würzburg, Würzburg, Germany

^2^ Université Paris-Saclay, CNRS, IRD, UMR Évolution, Génomes, Comportement et Écologie, 91198, Gif-sur-Yvette, France

`*` Corresponding author: douglas.sponsler@uni-wuerzburg.de

# ABSTRACT

Species coexistence is constrained in part by competition over shared food sources. Bumble bees are classic model organisms for the study of competition and coexistence, and this topic has taken on new urgency as climate-induced range shifts are expected to reshuffle competitive relationships between bumble bee species, particularly along the climatic gradients of the mountain slopes that host the world's richest bumble bee communities. We surveyed floral abundance and recorded bumble bee floral visitation patterns along a ~1400 m elevation gradient in the German Alps. Using these data, we first ask how floral preferences compare across bumble bee species and to what extent these patterns can be explained by trait-matching between bumble bee tongue length and floral morphotype. We then investigate the patterns of diet overlap between bumble bee species at the nested levels of floral morphotype, floral family, and floral genus, and model diet overlap as a function bumble bee tongue length and elevation. We find that while tongue length is a significant predictor of bumble bees' floral morphotype preferences, species with similar tongue lengths exhibit divergent preferences with respect to certain flora morphotypes. This indicates that factors other than tongue length are involved in flower choice, which might partly explain the puzzling coexistence of bumble bee species with similar tongue lengths. Moreover, genus-level diet overlap was approximately one half that of morphotype-level diet overlap, indicating that even when bumble bees exhibit similar biases at the level of floral morphotype, partly due to the constraints of tongue length, floral diversity enables the mitigation of competition via the subpartitioning within floral morphotypes. This suggests that floral species richness may be as important as overall floral abundance in avoiding the competitive exclusion of bumble bee species. Diet overlap was highest at the low end of our elevation gradient, declining sharply over the transition from valley floor to mid-slope, then leveling off or declining weakly for the remainder of the gradient. If diet overlap is interpreted as an index of potential competition, upslope shifts in bumble bees and flora could generate more competitive conditions at high elevations in cases where floral resources are limiting.

*Keywords*: foraging, trait matching, pollinator, *Bombus*, alpine

```{r setup, include=FALSE, cache=FALSE, message = FALSE}

library("knitr")

### Chunk options: see http://yihui.name/knitr/options/ ###
## Text results
opts_chunk$set(echo = FALSE, warning = TRUE, message = TRUE)

## Code decoration
opts_chunk$set(tidy = TRUE, comment = NA, highlight = TRUE)

## Cache
# opts_chunk$set(cache = 2, cache.path = "output/cache/")

## Plots
opts_chunk$set(fig.path = "../output/")

```

```{r knitcitations, cache = FALSE}
library(knitcitations)
cleanbib()   
cite_options(citation_format = "pandoc")
```

# INTRODUCTION

The role of competition and niche partitioning in regulating community assembly has long been a central question in ecology, but it has taken on new urgency on account of anthropogenic climate and land use change, which threaten to disrupt existing patterns of species interactions and niche relationships [@Tylianakis2008-fi]. These effects are likely to be especially pronounced in mountain ecosystems, where steep elevational temperature gradients facilitate species range shifts in response to climate change, and socioeconomic factors are driving rapid changes in historical land use patterns [@Becker2007-cg; @Loffler2011-xo; @Klein2019-dq]

Bumble bees (Hymenoptera: *Bombus* spp.) have become classic model organisms for the study of competition and species coexistence due to their broad and overlapping distributions and their gross similarity in morphology, behavior, and dietary niche [@Plowright1984-lc; @Goulson2008-zy]. In this role, however, they have proven somewhat enigmatic. On the one hand, there is evidence that resource limitation can be a significant constraint on local bumble communities [@Heinrich1976-wc; @Bowers1985-df], and certain bumble bee communities have been shown to be overdispersed in terms of tongue length or phylogeny, suggesting that competition functions as a community assembly process [@Inouye1977-kl; @Pellissier2013-pk]. Nevertheless, bumble bee communities with an apparent deficit of trait (or phylogenetic) dispersion are commonly observed [@Ranta1981-vx; @Goulson2008-zy; Iserbyt2008-jn]. To account for this inconsistency, @Ranta1981-vx proposed that bumble bees communities are governed by a spatiotemporally heterogeneous mosaic of competitive conditions, such that competitive inequalities between species are sufficiently variable to allow more species to coexist than predicted by trait-based competition models. A complementary explanation, based on the core/satellite species hypothesis [@Hanski1982-hs], is that competitive forces are significant only among regionally abundant and widespread “core species”, and the effects of competition are more pronounced in the composition of regional species pools than in that of local assemblages [@Hanski1982-ie]. In some cases, competition for food may simply be precluded by a saturating abundance of floral resources, implying that populations are limited by other factors, such as abiotic stressors, nest site availability, or predator/pathogen pressure [@Ranta1981-vx; @Williams1989-vr; Iserbyt2008-jn].

All else held equal, species coexistence is favored when intraspecific competition exceeds interspecific competition, and this stabilizing inequality is favored to the extent that species differ in dietary niche [@Abrams1977-zp; @Chesson2000-aw]. Bumble bees are exceptionally versatile foragers whose ability to solve the puzzles posed by complex flower morphologies [Plowright1984-lc], or indeed by human investigators [@Alem2016-rn; @Loukola2017-bt], renders almost any flower potentially accessible. Nevertheless, the flexibility of bumble bee foraging is tuned to floral reward, which depends on both the intrinsic reward of a given flower and the costs associated with exploiting it. These costs are closely related to the degree of morphological fittedness between bumble bees and their floral hosts, which consists principally in the matching of bumble bee tongue length to floral corolla depth. Bumble bee species can differ markedly in tongue length, and floral handling time increases when tongue length and corolla depth are poorly matched [@Inouye1980-mv; @Harder1983-bk], which leads to an association between tongue length and floral choice in bumble bee communities [e.g. @Heinrich1976-wc; @Teras1976-pi; @Inouye1978-qi; @Pyke1982-gn]. This association between tongue length and floral niche is imperfect, though, and there can be considerable overlap in floral resource use between bumble bees of different tongue lengths [e.g. @Ranta1980-ok; @Goulson2008-oe].

While niche overlap cannot serve as a metric of competition *per se* [@Abrams1980-am], it is often used as an index of *potential* competition, particularly in the context of dietary studies [e.g. @French1985-eo; @Steffan-Dewenter2000-ll; @Jung2015-cw], based on the reasoning that if and when food becomes limiting, competition can be expected to vary in proportion to the dietary overlap of co-occurring species [@Pianka1972-oi]. Our study is set in the Berchtesgadener Alps of southeast Germany, a landscape of flower-rich mountain meadows embedded in a matrix of coniferous forest. This system is one of abundant floral resources but harsh abiotic conditions, so it is likely that competition for floral resources is low, or at least transient, and that bumble bee populations are limited instead by the shortness of the growing season, as has been suggested in the similar environment of the Eyne Valley of France [@Iserbyt2008-jn]. Yet while *realized* competition between bumble bees in our system may be low, the question of *potential* competition is highly relevant because mountain ecosystems are currently subject to at least two powerful drivers of directed change. First, warming climates have already caused measurable upslope shifts in bumble bees and their floral hosts [@Ploquin2013-da; @Kerr2015-dg; @Pyke2016-ll; @Fourcade2019-ct; @Soroye2020-zb; @Marshall2020-pc], which could introduce novel competitive interactions between bumble bee species that experience increased sympatry. Second, socioeconomic factors have led to a dramatic decline in traditional alpine grazing management, particularly in the European Alps, resulting in widespread pasture abandonment and reforestation of mountain meadows [@Tappeiner2006-jp; @Tasser2007-qo]. The regression of meadows to forest entails a massive loss of floral resource availability, and even early stages of succession can cause declines in floral abundance and diversity, with corresponding effects on bumble bees [@Iserbyt2015-jc; @Walcher2017-kw; @Suzuki-Ohno2020-jv]. At the same time, more accessible areas at the bases of mountains are undergoing agricultural intensification that displaces natural vegetation and favors grasses over forbs [@Tasser2002-om]. Caught between adverse land use changes above and below and subject to climate-induced range shifts, mountain bumble bees are facing a future of reduced floral resource availability and altered sympatry, and potential competition is likely to become realized according to patterns of niche overlap between species.

Thus, the still-obscure mechanisms of bumble bee coexistence and niche partitioning remain a salient problem from a basic ecological perspective, but they have also become timely questions in the context of applied conservation. The aim of our study is to synthesize the phenomena of floral preference, trait-matching, and dietary niche overlap in a way that accounts for the coexistence of bumble bee species with similar tongue lengths and sheds light on the possible effects of habitat loss and climate-driven range shifts on the future of competition between bumble bee species in mountain habitats. We first ask how floral preferences compare across bumble bee species and to what extent these patterns can be explained by trait-matching between bumble bee tongue length and floral morphotype. We then investigate the patterns of dietary niche overlap between bumble bee species that emerge from their respective floral preferences, comparing niche overlap at the nested levels of floral morphotype, floral family, and floral genus. Finally, we ask how niche overlap varies with elevation and discuss the implications of this variation for bumble bee conservation in an era of global change.

# METHODS

## Field system

Field sampling was conducted 2010-2012 in Berchtesgaden National Park (47.55°N, 12.92°E), located in the Northern Limestone Alps of southeast Germany. The landscape is composed of mountain pastures in a coniferous forest matrix. We selected 25 study sites (60 x 60 m) on pastures at elevations ranging from 641-2032 m above sea level. Fourteen of these pastures are extensively grazed by cattle or sheep, three are mowed for hay production, and eight have been abandoned throughout the last century and are no longer subject to any human management.

Sampling consisted of repeated visits (hereafter, “visitation samples”) to each study site at approximately weekly intervals. Sampling was only conducted during periods without rain and with an air temperature of at least 6°C. Bumble bees were recorded during one 50-minute transect walk per visit, and bumblebees observed on a given flower were counted as floral visitors. Queens were identified to species level in the field to avoid destructively sampling colony foundresses, but workers and males were collected and frozen for later identification in the laboratory. Species were identified following @Amiet1996-jz. While male bumble bees was recorded, we chose to analyze only the pooled visitation of by queens and workers. In conjunction with bumble bee observations, we estimated the flower cover of each herbaceous or shrubby plant species within each 60 x 60 m study plot to the nearest 0.01 m^2^ and identified the plant species according to @Lauber2007-ha and @Oberndorfer2001-ro.

While per-sample effort was standardized across all years, the total number of samples varied: 153 in 2010, 212 in 2011, and 539 in 2012. The start of field sampling each year was determined by snowmelt, so lower sites were sampled as early as April while higher sites could not be sampled until June or July. In 2010 and 2012, this resulted in a temporally staggered pattern of sampling, but a combination of early snowmelt and a late start to field activities in 2011 resulted in sampling that was approximately aligned in time across the whole elevation gradient. Due to these annual differences in sampling intensity and duration, we opted to pool data across years rather than include year effects in our models, which would multiply model complexity and be difficult to interpret. This means, however, that our analysis cannot detect year-to-year variation in floral preferences and niche overlap, and main effects cannot be distinguished from annual anomalies. With only three years of data, though, such annual inferences would be of little value, and we deemed them not worth the cost of added model complexity.

The samples analyzed in this paper have been previously analyzed (Sponsler et al., submitted), but the analyses presented here are non-overlapping.

## Data sources

### Floral morphology

Rather than considering various floral morphological traits independently, we relied on the synthesis of @Kugler1970-gb, who developed a typology consisting of ten primary morphotypes and nested subdivision of each. Kugler’s typology was informed by laboratory studies of bumble bee flower choice [@Kugler1943-uy], and there is a precedent for analyzing bumble bee visitation in terms of Kugler’s morphotypes [@Farnberger1996-xo; @Neumayer1999-il; @Schneller2014-ry].

Kugler morphotype classifications of the plants in our study system were accessed via the BIOFLOR database [@Klotz2002-uc]. Morphotypes were simplified to the primary classes numbered 0-10: 0 = not applicable (includes wind-pollinated flora such as grasses), 1 = disc- and bowl-shaped flowers, 2 = funnel flowers, 3 = bell-shaped flowers, 4 = stalk-disc flowers, 5 = lip flowers, 6 = flag flowers, 7 = flower heads, 8 = spike flowers, 9 = brush flowers, 10 = trap flowers **(Figure 1)**. In a few cases, a plant species in our data set was not found in the BIOFLOR database. In most cases, morphotypes are constant at the genus level, and missing morphotypes were added based on the morphotypes found for congeners. In the case of the genus *Phyteuma*, we followed @Neumayer1999-il and classified it as a group 7 flower head. Morphotypes 9 and 10 were extremely rare in our study system and were therefore omitted from all analyses. Plants belonging to morphotype 0 were typically not recorded during floral surveys due to their lack of conspicuous floral displays, and they were rarely visited by bumble bees, so this morphotype was also omitted from all analyses.

### Bumble bee tongue length

Bumble bee tongue-length data (for the worker caste) were obtained from the literature values compiled and summarized by @Arbetman2017-xj. Values for *B. gerstaeckeri* and *B. mendax*, which were not included in @Arbetman2017-xj, were obtained from @Obeso1992-yp and @Durieux2000-uk, respectively. We then binned bumble bee species into three tongue-length classes: short (< 7.5 mm), medium (7.5—10.0 mm), and long (>= 10.0 mm). Short-tongued species included *B. jonellus*, *B. lapidarius*, *B. pratorum*, *B. psithyrus*-group, *B. pyrenaeus*, and *B. terrestris/lucorum*. Medium-tongued species included *B. humilis*, *B. hypnorum*, *B. monticola*, *B. mucidus*, *B. pascuorum*, and *B. wurflenii*. Long-tongued species included *B. gerstaeckeri*, *B. hortorum*, and *B. mendax*.

## Data processing

Prior to analysis, we aligned each visitation sample with a corresponding floral survey sample. In most cases, floral surveying was conducted jointly with bumble bee observation for each site and date. Occasionally, though, floral surveying could not be completed on the same day as bumble bee observation, resulting in a small temporal offset between visitation samples and floral survey samples. In these cases, visitation samples were paired with the nearest available floral survey data within a maximum distance of 7 days. Visitation samples lacking corresponding floral survey data within 7 days were omitted from further analysis, which resulted in dropping 40 site-dates out of a total of 895, leaving 855 site-dates included in our analyses. In a few cases, two consecutive visitation samples at a given site shared the same nearest floral survey sample; in these cases, the visitation samples were first pooled, then aligned to their shared nearest floral survey sample. In rare cases, a given floral species was recorded during visitation sampling but missed during floral surveying, and these species were retroactively assigned the minimum floral area value of 0.01 m^2^.

We defined floral preference as the relative rate of visitation between floral morphotypes after conditioning on the neutral effects of abundance (see model descriptions below for details). Visitation data were extremely right-skewed because the majority of available flora received zero visits during any given sampling period. To facilitate model fitting, we created a binary version of visitation rate in which visited flora were assigned to 1 and unvisited flora to 0. We then modeled binary visitation using generalized models with Bernoulli distribution families. The conditional effects inferred from these models must, therefore, be interpreted as effects on visitation probability rather than visitation frequency.

We calculated niche overlap with the goal of focusing on elevational patterns, sacrificing temporal resolution to pool observations within site. First, we calculated each bumble bee species' proportional visitation frequency at the nested levels of floral morphotype, plant family, and plant genus. Then, we calculated, for each bumble bee species, the mean proportional visitation rate to each morphotype/family/genus across all sampling dates within each site. Using these species-level visitation matrices, we calculated pairwise niche overlap as 1 - Horn-Morisita distance (calculated separately for morphotype-, family-, and genus-level matrices) for all species pairs within each site. This approach was chosen for its regularizing effect on the highly variable "snapshot" visitation data generated by our individual transect walks, but it entails the disadvantage of conflating niche overlap due to similar floral preference and niche overlap due to similar temporal (i.e. seasonal) activity patterns. We opted to analyze niche overlap only down to the level of floral genus in order to dampen the possible effects of false non-detection of floral species due to undersampling of the diets of rare bumble bee species [@Williams2005-yp].

## Statistical models

All analyses were conducted in `R` [@RCoreTeam-2021aa]. Data handling and visualization were performed with the `tidyverse` suite [@Wickham2019-aa]. All `R` code used in this analysis, together with full model output and diagnostics, is available in Appendix X.

### Model 1: Comparing overall visitation probability across floral morphotypes

We first compared overall visitation probability across floral morphotypes. In this model, visitation for each sample was pooled across bumble bee species and within floral morphotypes prior to binarization. Visitation probability was then modeled using multilevel Bayesian regression with a Bernoulli distribution family (logit link function). Models were implemented in Stan [@Stan2020aa] via the package `brms` [@Burkner2017-aa], and visualizations were prepared with the packages `tidybayes` [@Kay2020-aa] and `ggdist` [@Kay2021-aa]. *Bumble bee abundance* and *percent flower cover* (per morphotype) were included as a global interaction term, thus conditioning visitation probability on abundance and enabling the inference of preference independent of the neutral effects of abundance. *Floral morphotype* was specified as a both a varying intercept term and a varying slope term with respect to *percent flower cover*. *Site* was included as a varying intercept term to avoid overfitting due to pseudoreplication owing to the repeated observations within each site. We set a weakly informative (i.e. regularizing) Normal(0, 1.5) prior on the β-coefficients of the fixed effects [@Lemoine2019-xm] and used the default priors provided by `brms` version 2.14.4 for all other model parameters.

### Model 2: Visitation probability as a function of floral morphotype and bumble bee tongue length

To explore the functional underpinnings of bumble foraging preference, we analyzed visitation probability as a function of the interaction between floral morphology and bumble bee tongue length class. For this analysis, we pooled visitation within floral morphotype but not across bumble bee species. Then we specified a Bayesian model of the same form as Model 1, but this time included a varying slope and intercept term for *floral morphotype* : *bumble bee tongue length class* and *floral morphotype* : *bumble bee species*. In this model, the relationship between *floral morphotype* and *bumble bee tongue length class* was the focus of inference, and the *bumble bee species* effect was added to account for the grouping of observations within species.

### Model 3: Visitation probability as a function of floral morphotype and bumble bee species

To focus on species-level morphotype preferences, we removed the tongue length effect from Model 2 and re-fit the model considering only bumble bee species.

### Model 4: Aggregate diet ordination

To visualize the overall diet similarity between bumble bee species, we pooled visitation data across all samples within bumble bee species and plant genus, then performed a principal coordinates analysis (PCoA) based on Horn-Morisita distance metric. To test the significance of tongue length class as a predictor of diet similarity, we performed a permutation test. Importantly, this approach treats the whole study system in aggregate, thus conflating the causal effects of floral preference, seasonal activity patterns, and site occupancy patterns in the generation of overall diet similarity. In this sense, it can be understood as a model of potential diet overlap given the aggregate foraging patterns of the bumble bee species in our system, with the understanding that realized niche overlap at particular locations or times may vary. PCoA and permutation tests were performed with the R package `vegan` [@Oksanen_2019], and visualizations were prepared with the packages `ggplot2` [@Wickham2016-aa] and `ggvegan` [@Simpson2019aa]

### Model 5: Niche overlap by level and elevation

To investigate elevational patterns of *within-site* niche overlap, we modeled community-mean niche overlap as a function of level (morphotype, family, genus) and elevation using hierarchical generalized additive models (HGAMs). Community mean niche overlap, which can be interpreted as an index of potential *diffuse* competition [@Pianka1974-qk; @MacArthur1972-qz], was calculated by taking the mean niche overlap (within level) across all species pairs for each site. Level was included as both a fixed intercept effect and a smoothing factor in conjunction with elevation, following the type "I" model form described by @Pedersen2019-kf. Management was initially included as a fixed intercept effect, but its influence was insiginificant, so it was dropped from the final model. To account for the proportional nature of our data, we used a beta distribution family with a logit link function. Models were fit with the R package `mgcv` [@Wood_2017] and visualized with the packages `mgcViz` [@Fasiolo2018aa] and `tidymv` [@Coretta-aa2021].

### Model 6: Genus-level niche overlap by tongue length and elevation

To focus on genus-level niche overlap as a function of tongue length class, we generated a binary tongue-difference variable for each species pair: "same" for species belonging to the same tongue length class and "different" for species belonging to different tongue length classes. We then took the mean genus-level niche overlap across all species pairs within tongue-difference class and site. Then, we constructed an HGAM of genus-level niche overlap using tongue-difference as both a fixed intercept effect and a smoothing factor in conjunction with elevation, again following the type "I" model form @Pedersen2019-kf.

# RESULTS

## Summary of visitation and floral survey data

We recorded a total of 12,918 bumble- bee-flower interactions, including 16 bumble bee species (with *Bombus terrestric/lucorum* and *Bombus psithyrus* species lumped into species groups), 163 plant species (110 genera, 37 families), and 736 unique bumble- bee-plant interaction pairs. *B. pascuorum*, *B. pratorum*, *B. soroeensis*, *B. terrestris-lucorum*, and *B. wurflenii* were the most abundant species, and they occurred at all sites. *B. hortorum*, *B. jonellus*,  *B. monticola*, and *B. psithyrus* occurred at moderate abundance. The remaining species --- *B. gerstaeckeri*, *B. hypnorum*, *B. lapidarius*, *B. mendax*, *B. mucidus*, and *B. pyrenaeus* --- were relatively rare. *B. humilis* was recorded only three times and only in one year, so we omitted it from all analyses. Floral surveying yielded a total of 352 plant species (191 genera, 52 families), of which 155 (103 genera, 35 families) observed to be visited by bumble bees. For a more full analysis of bumble bee and floral abundance patterns, see Sponsler at al. (submitted).

## Floral preference (Models 1-3)

Our visitation models (Models 1-3) revealed strong morphotype preferences across the aggregate bumble bee community and marked differences in floral preference between bumble bee species. These differences in preference were partly explained by differences in tongue length, but certain floral morphotypes generated divergent preference patterns among bumble bee species with similar tongue lengths, and there was substantial overlap in floral preference between bumble bees of different tongue length.

Lip and flag flowers were the most preferred morphotypes overall, with similarly high visitation probabilities of 0.35 (95% CI = 0.27–0.44) and 0.31 (95% CI = 0.24–0.40), respectively **(Figure 1)**. In second place, head flowers had a visitation probability of 0.18 (95% CI = 0.13–0.25). Bell flowers were a distant third with a visitation probability of 0.09 (95% CI = 0.06–0.13), followed by stalk-disc, funnel, and disc flowers with visitation probabilities of 0.06 (95% CI = 0.04–0.08), 0.04 (95% CI = 0.03–0.06), and 0.04 (95% CI = 0.02–0.06), respectively. Model 1 explained approximately 37% of the total variance (97.5% CI = 35–38%).

```{r fig1, echo = FALSE,  fig.cap = paste("Posterior conditional mean visitation probability by floral morphotype. Point-interval plots depict 66\\% and 95\\% CIs. Each floral morphotype is exemplified by a representative species common in our study system. Illustrations are modifications of public domain works obtained from www.plantillustrations.org. For details of original sources, see Appendix X.")}
knitr::include_graphics("../output/brm_pvis_01_plot1_mod.png")
```

Disc and stalk-disc flowers were visited at similarly low rates across tongue length classes, and lip flowers were visited at similarly high rates across tongue length classes **(Figure 2)**. There was considerable in visitation rates between individual species, though, including those belonging to the same tongue length class **(Figure 3)**. The short-tongued *B. terrestris/lucorum*, for example, stands out in its preference for disc flowers, and the medium-tongued *B. monticola* in its preference for stalk-disc flowers. While lip flowers were visited at relatively high rates by almost all species, the medium-tongued *B. pascuorum* exhibited an especially high preference for this morphotype, while the medium-tongued *B. mucidus* ignored lip flowers almost entirely.

In contrast, visitation of funnel, bell, flag, and head flowers differed significantly both across tongue length classes **(Figure 2)** and between individual species **(Figure 3)**. Funnel flowers were preferred by long-tongued species, though effect was driven mainly by the *Aconitum* specialist *B. gerstaeckeri* [@Ponchau2006-af; @Schneller2014-ry]. Bell flowers were visited at a higher rate by short-tongued species than by long-tongued species, with medium-tongued species exhibiting intermediate visitation rates. At the species-level, bell flowers were most preferred by the short-tongued species *B. jonellus*, *B. pyrenaeus*, and *B. soroeensis*, and by the medium-tongued species *B. hypnorum* and *B. monticola*. Flag flowers were visited at a higher rate by medium-tongued species than by short-tongued species, with long-tongued species exhibiting intermediate visitation rates. Flag flowers generated strong separation in visitation rate at the species level, receiving very high visitation from the medium-tongued *B. pascuorum* and *B. mucidus*, moderately high visitation from the short-tongued *B. terrestris/lucorum* and *B. lapidarius*, the medium-tongued *B. monticola*, and the long-tongued *B. mendax* and *B. hortorum*, and low visitation by the short-tongued *B. soroeensis*, *B. pyrenaeus*, *B. psithyrus*, *B. pratorum*, and *B. jonellus*, the medium-tongued *B. hypnorum*, and the long-tongued *B. gerstaeckeri*. Models 2 and 3 both explained approximately 31% of total variance (97.5% CI = 30–32%).

```{r fig2, echo = FALSE,  fig.cap = paste("Posterior conditional mean visitation probability parsed by Kugler morphotype and bumble bee tongue length class (Model 2). Point-interval plots depict 66\\% and 95\\% CIs. Letters indicate group differences based on evidence ratio at alpha = 0.05. Note that the range of the x-axis has been rescaled for each panel to emphasize the comparison of visitation probabilities across tongue length classes within floral morphotypes.")}
knitr::include_graphics("../output/brm_pvis00_plot1_mod2.png")
```

```{r fig3, echo = FALSE,  fig.cap = paste("Posterior conditional mean visitation probability by Kugler morphotype and bumble bee species (Model 3). The tongue length class of each species is color-coded. Point-interval plots depict 66\\% and 95\\% CIs. Note that the range of the x-axis has been rescaled for each panel to emphasize the comparison of visitation probabilities across bumble bee species within floral morphptypes.")}
knitr::include_graphics("../output/brm_pvis_02_plot1_mod.png")
```

## Aggregate diet similarity

Principal coordinates analysis (Model 4) suggested weak but significant effects of tongue length on the diet similarity while also highlighting the particular floral genera that accounted most strongly for the dietary separation of bumble bee species. Axis 1 accounted for 23% of total diet variance and primarily separated long- and medium-tongued species from short-tongued species **(Figure 4)**. Long- and medium-tongued species scored low on Axis one and were associated with the lip flowers *Rhinanthus*, and *Stachys* and flag flowers *Anthyllis*, *Lotus*, and *Trifolium*. Short tongued species scored high on Axis one and were associated with the bell flowers *Campanula* and *Erica*, and the head flower *Centaurea*. Axis 2 accounted for 19% of total variance and primarily explained variation within short-tongued species. Short-tongued species scoring high on Axis 2 were associated with *Campanula* while those scoring low on Axis 2 were associated with *Erica*. Short-tongued species scoring close to zero on Axis 2 also scored lower on Axis 1 (closer to medium- and long-tongued species) and were associated with head flowers. Both the lip flower *Ajuga* and the head flower *Centaurea* scored near the middle of the ordination, reflecting their broad appeal across tongue length classes. Tongue length class explained 24% of diet variance (Adonis permutation test, F = 1.79, p = 0.005). While short-tongued bees appeared most broadly spread in ordination space, there were no significant differences in dispersion across tongue length classes (F = 0.42, p = 0.66).

```{r fig4, echo = FALSE,  fig.cap = paste("Principal coordinates ordination of bumble bee visitation pooled across all samples. To avoid overplotting, only plant genera with an absolute score of at least 0.20 on one or both axes are plotted.")}
knitr::include_graphics("../output/pcoa_aggregate_mod.png")
```

## Site-level diet overlap (Models 5 and 6)

Site-level analysis of diet overlap revealed strong differences in the magnitude of diet overlap at the nested levels of floral morphotype, floral family, and floral genus, as well as consistent patterns in response to elevation. At the site level, however, bumble bees belonging to the same tongue length class did not have significantly more similar diets than those belonging to different tongue length classes.

Morphotype-level diet overlap was nearly double genus-level diet overlap throughout the whole elevation gradient, with family-level diet overlap taking intermediate values. The fixed intercepts of morphotype- and genus-level diet overlap each differed significantly (in opposite directions) from the reference level of family-level diet overlap (p << 0.001). The conditional smooths of each level were significantly different from flat lines (p < 0.001) and generally mirrored one another in shape **(Figure 8)**. Diet overlap was highest at the low end of the elevation gradient, decreased steeply from 600-1200 m, and leveled off or declined weakly for the remainder of the gradient, though morphotype-level diet overlap exhibited a second peak just above 1500 m that was not evident in genus- or family-level diet overlap. Model 5 explained 74.2% of total deviance (adjusted R^2^ = 0.70). Whether a species pair belonged to the same or different tongue length classes had no effect on genus-level niche overlap, either in terms of intercept (p = 0.12) or shape of response to elevation **(Figure 9)**. Model 6 explained 40.0% of total deviance (adjusted R^2^ = 0.35).

```{r fig5, echo = FALSE,  fig.cap = paste("Niche overlap by level through elevation. Points show empirical niche overlap values, lines show GAM smooths, and shaded bands show 95\\% confidence intervals.")}
knitr::include_graphics("../output/niche_gam_00_plot1.png")
```

```{r fig6, echo = FALSE,  fig.cap = paste("Niche overlap by tongue length class difference through elevation. Lines show GAM smooths and shaded bands show 95\\% confidence intervals.")}
knitr::include_graphics("../output/niche_gam_20_plot1.png")
```

# DISCUSSION

Any discussion of competition and niche partitioning among *Bombus* species must begin with the observation that bumble bees are unique among insect pollinators in the degree to which their foraging combines contrasting extremes of bias and flexibility. Coexistence between bumble bee species with similar tongue lengths hinges on the relationships between species-level floral preferences, trait-matching, and the emergent phenomenon of diet similarity. In this study, we clarify these relationships with respect to bumble bees and wild flowers interacting along a mountain slope elevation gradient. In addition to addressing the question of bumble bee species coexistence, our results suggest how patterns of competition may respond to climate and land use change along mountain slopes, an issue of immediate concern for the conservation of bumble bees and their floral mutualists [@Sirois-Delisle2018-vi; @Soroye2020-zb].

## Floral preference is only partly explained by the relationship between bumble bee tongue length and floral morphotype

 While biases can be explained partly in terms of morphological matching between tongue length and floral morphotype

 The rule is flexibility, and floral morphotypes do not constitute hard boundaries to bumble bee flower choice. Nevertheless, we demonstrate that for certain floral morphotypes --- funnel flower, bell flowers, flag flowers, and head flower --- visitation rate is affected strongly by tongue length.

In our study, lip, flag, and head flowers were by far the most strongly preferred when visitation was pooled across all bumble bee species, while disc, funnel, bell, and stalk-disc flowers were comparatively underrepresented. These findings are broadly consistent with the aggregate visitation patterns previously reported by @Schneller2014-ry in the central Alps at Vorarlberg and Neumayer1999-il ~50 km south of our study system, suggesting that our findings are at least regionally generalizeable. When considering these patterns from the perspective of pollination services to mountain flora, however, it is important to note that while bumble bees are the dominant Hymenopteran pollinators in our study system, Syrphid flies are also abundance flower visitors whose visitation is mainly restricted to the open flower morphotypes (e.g. disc flowers) that receive less attention from bumble bees [@Neumayer1999-il]. In the context of our research question, we have emphasized the differences in tongue length between bumble bee species, but all bumble bees have relatively long tongues compared to the overall pollinator community, and the process of niche partitioning *among* bumble bee species is nested within the broader process of niche partitioning *between* bumble bees and other flower visitors.

While the variation we observed in floral preference across tongue length classes largely comports with the established relationship between tongue length and corolla depth, variation in preference *within* tongue length classes agrees with studies demonstrating the insufficiency of simple morphological matching in explaining bumble bee foraging patterns [e.g. @Ranta1980-ok; @Goulson2008-oe]. Perhaps the most striking divergence of floral preference within tongue length class was the divisive effect of bell flowers on short-tongued bumble bees. *B. pyrenaeus*, *B. soroeensis*, *B. pratorum*, and *B. jonellus* visited bell flowers at high rates, while *B. terrestris/lucorum*, *B. psithyrus*, and *B. lapidarius* visited them relatively rarely. A similar split among the short-tongued species occurred with respect to head flowers, which were visited at high rates by *B. soroeensis* and *B. psithyrus* but at relatively low rates by other short-tongued species. There are no obvious morphological distinctions within short-tongued bumble bees that would account for these divergent affinities, and ecological drivers such as differences in spatial or temporal activity patterns are more likely explanations [@Sowig1989-ld; @Schneller2014-ry]. Nectar robbing is another mechanism by which the relationship between tongue length and corolla depth might be relaxed, but its influence in our data is unclear. Of the bumble bees in our study system, *B. terrestris/lucorum* and especially *B. wurflenii* are known to engage frequently in nectar robbing [@Brian1954-cy]. In the case of *B. wurflenii*, however, its morphotype affinities did not differ markedly from those of other medium-tongued species, and it may be that the condition of medium-tongue length alone confers enough to foraging flexibility in our study system to dampen the divergent effects of nectar robbing. Nectar robbing might, however, account for the deviation of *B. terrestris/lucorum* from most other short-tongued species with respect to flag flowers. With the exception of *B. lapidarius*, all short-tongued species besides *B. terrestris/lucorum* eschewed flag flowers, while *B. terrestris/lucorum* visited them at rates comparable to medium- and long-tongued species. Indeed, *B. terrestris/lucorum* was, on the whole, an eclectic forager, also demonstrating a relatively high affinity for the disc flowers that were largely ignored by other species. It should be noted, though, that an unusually high abundance of *B. terrestris/lucorum* queens was observed in 2012, but nest founding success was evidently low, leading to low *B. terrestris/lucorum* abundance later in the season, an anomaly that could account for the unusual foraging patterns of this species group in the pooled data. Together, these species-level idiosyncracies of preference within tongue length classes underscore the fact that factors other than the matching of tongue length and floral morphology --- such as affinities based on floral color, plant growth form, spatial distribution of floral patches, or habitat context [@Brian1957-dg; Harder1985-ls; @Sowig1989-ld] --- are at play in bumble bees' floral preferences, which may partly explain the coexistence of bumble bees of similar tongue lengths.

<!-- - Explain that traits such as color (really especially color) probably explain much of the variance in preference not explained by morphology [@Neumayer1999-il].  -->

## The influence of tongue length on floral preference does not carry on to genus-level niche overlap, and floral diversity within morphotypes relaxes potential competition between bumble bee species of similar tongue length

With respect to diet overlap among bumble bees, our finding that bumble bee diet ordination produces a primary axis corresponding to tongue length (and associated with affinity for legumes) and a secondary axis describing the dispersion of short-tongued species is consistent with studies in Poland [@Goulson2008-oe], France [@Durieux2000-uk], and England [@Goulson2005-nz] that analyzed bumble bee diet by similar methods, suggesting that this may be a general pattern among bumble bee communities, at least within temperate Europe. Perhaps a more significant finding, though, is that community-mean diet overlap at the level of floral morphotype was consistently nearly twice as high as diet overlap at the level of floral genus. From a numerical perspective, this result is unsurprising simply because each morphotype comprises many genera. From an ecological perspective, though, this pattern suggests another answer to the question of how bumble bee species of similar tongue length can coexists. If tongue length primarily constrains bumble bee foraging at the coarse level of floral morphotype, floral diversity *within* morphotypes creates space for bumble bees with similar tongue lengths to avoid actual dietary overlap, thus mitigating interspecific competition. Importantly, this suggests that diversity within floral morphotypes may as important as morphotype diversity *per se* and overall floral abundance in facilitating the coexistence of bumble bee species, an insight which may have implications for the design of floral enhancements aimed at bumble bee conservation.

<!-- @Durieux2000-uk similarly ofund that while tongue length is correlated with floral preference, it is not correlated with diet similarity. But we did better stats! -->

## Climate, land use, and the future of bumble bee niche relations

Community mean diet overlap was highest at the lowest sites and declined sharply over the transition from valley floor to mid-slope. If this pattern of niche overlap moves upslope in response to climate change, it might result in increased competition potential at higher elevations, which could threaten high-elevation bumble bee species that currently exist under conditions of relatively low niche overlap. In our system, though, it must be noted that the lowest elevation sites are also the ones most influenced by human land use, and it is possible that the elevation pattern evident in our data is driven by factors other than elevational climate variation.

<!-- As discussed earlier, the interpretation of floral visitation rates and diet overlap hinge strongly on the extent to which interspecific competition actually occurs in our study system. We have argued that the assumption of minimal competition is reasonable in the context of the Berchtesgadener Alps, owing to their rich floral resources and restrictive abiotic conditions, together with the manifest co-occurrence of bumble bee species with similar tongue lengths. This leads to the interpretation of abundance-conditioned visitation rates as *preferences* and the interpretation of diet overlap as *potential competition*. To the extent that interspecific competition *does* occur in our system, though, visitation rates would represent floral preference distorted by the pressure of competition toward niche partitioning, rendering diet overlap an underestimate of potential competition. While we maintain, based on our knowledge of our study system, that the first interpretation is more likely, both interpretations are plausible and must be kept in mind. Given our assumption of negligible interspecific competition, however, it might be questioned whether the inference of *potential* competition is ecologically relevant, since only realized competition can cause competitive exclusion. In answer to this question, we point out that while the assumption of negligible interspecific competition is necessary for the inferences we make concerning floral preference and potential competition, the processes of (1) imperfect trait-matching bewteen bumble bee tongue length and floral morphotype and (2) taxonomic subpartitioning of floral morphotypes that we infer in our analysis still obtain under conditions of realized interspecific competition, even if their inference under those conditions would be obscured by the confounding effects of competitive niche partitioning. Indeed, conditions of realized competition likely occur in all bumble bee communities, including those of our study system, but these conditions may be spatially and/or temporally localized [@Ranta1981-vx], and perhaps associated with extreme climatic events that cause dramatic reductions in floral resource avaulability [@Rasmont2012-zq; @Thomson2016-gq; @Phillips2018-me]. This calls to mind the insight of MacArthur1972-qz, that competition need not be frequent to have lasting ecological impacts. Potential competition in the form of dietary overlap is perhaps the best indication of the effects that future instances of realized competition will have, and it is likely that the "ghosts of competition past" [@Connell1980-dn] haunt the floral preferences exhibited by bumble bees today. -->

<!-- Empirical measurements of niche overlap are, however, inherently ambiguous with respect to competition.  -->

<!-- Focusing on *potential* competition does not, however, escape the equivocity of niche overlap, since an observed pattern of niche overlap could just as plausibly be interpreted as adaptive niche partitioning in response to *realized* competition [@Pianka1974-qk]. This creates a predicament in which niche overlap can only be interpreted as an index of potential competition when realized competition is assumed to be negligible, an assumption that can only be justified on the basis of additional knowledge of the study system under consideration.  -->

<!-- - Explain that traits such as color (really especially color) probably explain much of the variance in preference not explained by morphology.  -->

<!-- Bumble bees are cold-adapted species [@Heinrich1994-cv] associated mainly with mountain habitats, where they serve as the principal pollinators of many alpine plants [@Williams1998-xi; @Goulson2010-lo]. The introduction or extinction of species from bumble bee communities, with corresponding addition and release of competitive pressure, can alter the floral visitation patterns of bumble bees and the pollination rates of their host plants [@Brosi2013-mu; @Ishii2013-sm].  -->

<!-- Lip and head flowers have the highest nectar reward per inflorescence [@Neumayer1999-il]. -->
<!-- Interestingly, @Farnberger1996-xo in a localized study near Wien found rather different results. Lip flowers were still found to be the favorite overall, and head flowers were popular, too, but flag flowers were of secondary importance, and bell flowers were the most visited morphotype for humi, pasc, and prat. Pasc almost totally ignored bell flowers in our study. -->

# ACKNOWLEDGMENTS

We thank D.J. McNeil for helpful conversations during data analysis. O. Mitesser kindly provided a high-performance computing environment to expedite Bayesian model fitting.

# REFERENCES

<!-- ```{r write_citations, cache=FALSE, include=FALSE} -->
<!-- write.bibtex(file = "knitcitations.bib") -->
<!-- ``` -->

<div id ="refs"></div>