### 2021-04-08 Dryad_ms1/README.Rmd
finalizing for submission

---
title: "Abundance and beta-diversity of bumble bees and wildflowers in the Berchtesgadener Alps"

author: Katharina Kallnik^1^, Johanna Sieger^3^, Fabrice Requier^2^, Douglas B. Sponsler^1^*,  Alice Claßen^1^, A. Fabienne Maihoff^1^,  Ingolf Steffan-Dewenter^1^

output: pdf_document
---

^1^ Department of Animal Ecology and Tropical Biology, Biocenter, University of Würzburg, Würzburg, Germany

^2^ Université Paris-Saclay, CNRS, IRD, UMR Évolution, Génomes, Comportement et Écologie, 91198, Gif-sur-Yvette, France

^3^ Naturpark Frankenhöhe, Ansbach, Germany

`*` Corresponding author: douglas.sponsler@uni-wuerzburg.de

<!-- # GENERAL INFORMATION -->

<!-- ## Timeframe of data collection -->

<!-- ## Location of data collection -->

<!-- ## Funding sources -->

# DATA FILES

## 1. `floral_tax.csv`

### Description: Supplies family-level taxonomy for the floral taxa in our study.

### Columns:
  1. `plant.sp` = plant species
  2. `plant.genus` = plant genus
  3. `plant.family` = plant family

## 2. `network.csv`

### Description: Bumble- bee-wildflower visitation data used in all analyses of (1) bumble bee abundance, (2) bumble bee beta-diversity, and (3) interaction beta-diversity.

### Columns:
  1. `year` = year
  2. `dayofyear` = day of year since January 1
  3. `site` = site name
  4. `snowcover` = whether site was snow-covered
  10. `plant.sp` = full plant species name
  11. `plant.genus` = plant genus name
  12. `flower cover` = flower cover in m^2^ to the nearest 0.01 m^2^
  12. `day` = day of month
  13. `month` = month
  14. `date` = date

## 3. `floral_survey.csv`

### Description: Floral survey data used in all analysis of (1) floral abundance and (2) floral beta-diversity.

### Columns:
  1. `year` = year
  2. `dayofyear` = day of year since January 1
  3. `site` = site name
  4. `trap.time` = time interaction was observed
  5. `caste` = bumble bee caste (queen, male, worker, NA = parasitic *Psithryus*)
  6. `pollen` = whether the bee had visible pollen loads
  7. `bb.sp` = abbreviated bumble bee species name
  8. `bb.sp.lat` = full bumble bee species name
  9. `plant.sp.abb` = abbreviated plant species name
  10. `plant.sp` = full plant species name
  11. `plant.genus` = plant genus name
  12. `day` = day of month
  13. `month` = month
  14. `date` = date

## 4. `site_data.csv`

### Description: Site data including elevation, management, and geographic coordinates.

### Columns:
 1. `site` = site name
 2. `elev.class` = elevation category (oben, mitte, unten)
 3. `management` = site management (mowing, grasing, none)
 4. `temp.mean` = mean temperature recorded with iButton devices
 5. `elev.mean` = mean elevation
 6. `transect` = transect name
 7. `slope.calc` = slope calculated from min and max elevation
 8. `slope.est` = estimated slope
 9. `elev.min` = minimum elevation within site
 10. `elev.max` = maximum elevation within site
 11. `lat` = latitude (decimal degrees)
 12. `lon` = longitude (decimal degrees)
 13. `elev.class2` = alternative elevation binning
 14. `tree_line` = whether a site was above or below the tree line; only for sites included in analysis (hence NAs)

# CODE FILES

## 1. `ms1_abundance.Rmd`

### Description: R script for running abundance analyses.

## 2. `ms1_beta_diversity.Rmd`

### Description: R script for running beta-diversity analyses.