### 2021-04-14 code/ms2_ktype_preference_nullmodel_analysis.Rmd

---
title: "visitation_bias"
author: "Doug Sponsler"
date: "12/14/2020"
output: html_document
---
# Preparations
## Packages 
```{r, warning=FALSE, include=FALSE, echo = FALSE}
library(bipartite)
library(econullnetr)
library(patchwork)
library(mgcv)
library(mgcViz)
library(tidymv)
library(tidyverse)
library(lubridate)
```

## Load data
```{r message=FALSE, warning=FALSE, include=FALSE, echo = FALSE}
prop_visits_ktype <- read_csv("../data/processed_data/prop_visits_ktype.csv") %>%
  mutate(ktype = factor(ktype), # recode variables
         tongue = factor(tongue),
         visits = as.integer(visits))

prop_visits_ktype_pooled <- read_csv("../data/processed_data/prop_visits_ktype_pool.csv") %>%
  mutate(ktype = factor(ktype), # recode variables
         visits = as.integer(visits))

bb_traits <- read_csv("../data/processed_data/bb_traits.csv") %>%
  select(bb.sp, pbl.w, tongue = pbl.w.class2)

site_data <- read_csv("../data/processed_data/site_data.csv") %>%
  dplyr::select(-c(slope.calc, slope.est, elev.min, elev.max, temp.mean, elev.class, elev.class2)) %>% 
  mutate(site = factor(site),
         tree_line = factor(tree_line))
```

# Null model by species
```{r}
# The "consumers" table input to generate_null_net()
con <- prop_visits_ktype %>%
  unite(id, c(site, date)) %>%
  select(id, bb.sp, ktype, visits) %>%
  pivot_wider(values_from = visits, names_from = ktype, values_fill = 0) %>%
  select(-c("0", "10", "9"))

# The "resources" table input to generate_null_net()
res <- prop_visits_ktype %>%
  unite(id, c(site, date)) %>%
  select(id, ktype, perc.flower.cover) %>%
  distinct() %>%
  pivot_wider(values_from = perc.flower.cover, names_from = ktype, values_fill = 0) %>%
  select(-c("0", "10", "9"))

# Generate null model
null1 <- generate_null_net(consumers = con[, 2:9],
                           resources = res[, 2:8],
                           sims = 5000, # raised from default of 1000 (takes a while to run)
                           data.type = "counts",
                           c.samples = con[, 1], # parsed by sample
                           r.samples = res[, 1])

#saveRDS(null1, file = "../output/nullmodel_ktype.rds")
null1 <- readRDS("../output/nullmodel_ktype.rds")

# Tabulate output and generate tests
null_tab1 <- test_interactions(null1, signif.level = 0.99) %>% # make it a bit more conservative to account for multiple comparisons
  as_tibble() %>%
  select(bb.sp = Consumer, ktype = Resource, everything()) %>%
  group_by(bb.sp) %>%
  mutate(total.abund = sum(Observed), # total abundance per species
         preference = Observed - Null, # a working definition of preference
         preference.upper99 = Observed - Upper.99.CL, 
         preference.lower99 = Observed - Lower.99.CL,

         # If preference equals the difference between # observed visits and the null model prediction, then dividing preference by the total abundance of each species yields a standardized preference. Right? 
         preference.std = preference/total.abund, 
         Observed.std = Observed/total.abund,
         Upper.99.CL.std = Upper.99.CL/total.abund,
         Lower.99.CL.std = Lower.99.CL/total.abund) %>%
  left_join(bb_traits) %>%
  ungroup() %>%
  mutate(bb.sp = fct_reorder(factor(bb.sp), pbl.w))
```

## Visualize network
```{r}
plot_bipartite(null1, signif.level = 0.99)
```

## Visualize species model
```{r}
# Standardized observed vs. null predictions
## Species-wise
ggplot(null_tab1, aes(Observed.std, ktype)) +
  geom_segment(aes(y = ktype, yend = ktype, 
                   x = Lower.99.CL.std, xend = Upper.99.CL.std), 
               color = "gray40", inherit.aes = FALSE) +
  geom_point(aes(color = Test)) +
  facet_wrap(~bb.sp) +
  scale_y_discrete(limits = rev) +
  scale_color_manual(values = c("gray40", "#00B6EB", "#F8766D")) +
  labs(x = "Standardized visits", y = "Floral morphotype", color = "Test (p < 0.01)") +
  theme_light()

ggsave("../output/null_model_spwise_preference.pdf", height = 6, width = 7)
ggsave("../output/null_model_spwise_preference.png", height = 6, width = 7)

## Morphotype-wise
ggplot(null_tab1, aes(Observed.std, bb.sp)) +
  geom_segment(aes(y = bb.sp, yend = bb.sp, 
                   x = Lower.99.CL.std, xend = Upper.99.CL.std), 
               color = "gray40", inherit.aes = FALSE) +
  geom_point(aes(color = Test)) +
  facet_wrap(~ktype) +
  scale_y_discrete(limits = rev) +
  scale_color_manual(values = c("gray40", "#00B6EB", "#F8766D")) +
  labs(x = "Standardized visits", y = "Bumble bee species", color = "Test (p < 0.01)") +
  theme_light()

ggsave("../output/null_model_ktypewise_preference.pdf", height = 6, width = 7)
ggsave("../output/null_model_ktypewise_preference.png", height = 6, width = 7)

# Standardized preference ~ morphotype
ggplot(null_tab1, aes(ktype, preference.std)) +
  #geom_violin(fill = "#F8766D", color = "#F8766D") +
  geom_violin(fill = "gray70", color = "gray70") +
  geom_boxplot(fill = NA, color = "black") +
  theme_light(14) +
  labs(x = "Floral morphotype", 
       y = "Standardized preference")

ggsave("../output/null_model_ktype_preference.pdf", height = 6, width = 7)
ggsave("../output/null_model_ktype_preference.png", height = 6, width = 7)

# Standardized preference ~ morphotype * tongue class
ggplot(null_tab1, aes(ktype, preference.std)) +
  #geom_violin(fill = "#F8766D", color = "#F8766D") +
  geom_violin(fill = "gray70", color = "gray70") +
  geom_boxplot(fill = NA, color = "black") +
  theme_light(14) +
  labs(x = "Floral morphotype", 
       y = "Standardized preference") +
  facet_wrap(~tongue)

ggsave("../output/null_model_ktypeXtongueclass_preference.pdf", height = 6, width = 7)
ggsave("../output/null_model_ktypeXtongueclass_preference.png", height = 6, width = 7)

# Standardized preference ~ tongue length
ggplot(null_tab1, aes(pbl.w, preference.std)) +
  geom_point(aes(color = Test)) +
  geom_smooth(method = "gam", formula = y ~ s(x, k = 5), color = "gray20", size = 0.5) +
  facet_wrap(~ktype, scales = "free") +
  theme_light()
```

## Model standardized preference ~ tongue length
```{r}
gam1 <- gam(preference.std ~ 
              #s(ktype, bs = "re") + 
              ktype + 
              s(pbl.w, by = ktype, k = 5), 
            data = null_tab1) %>% getViz()

summary(gam1)
check.gamViz(gam1)
plot.gamViz(gam1, allTerms = TRUE)

# Extract predictions
gam1_pred <- predict_gam(gam1) %>%
  mutate(fit.plus.se = fit + (2*se.fit),
         fit.minus.se = fit - (2*se.fit))

# Plot
ggplot(gam1_pred, aes(pbl.w, fit)) +
  geom_line() + # fit line
  geom_ribbon(aes(ymin = fit.minus.se, ymax = fit.plus.se), # Uncertainty
              fill = "black", alpha = 0.2) +
  geom_point(data = null_tab1, # Original data
             aes(pbl.w, preference.std, color = Test), 
             inherit.aes = FALSE) +
  facet_wrap(~ktype) +
  labs(x = "Tongue length (mm)", y = "Standardized preference", 
       color = "Test (p < 0.01)") +
  theme_light()

ggsave("../output/null_model_ktypeXtonguelength_preference.pdf", height = 6, width = 7)
ggsave("../output/null_model_ktypeXtonguelength_preference.png", height = 6, width = 7)
```

## Model standardized preference by morphotype
```{r}
library(brms)

x <- seq(-2, 2, by = .1)
z <- seq(-20, 20, by = .1)

# This seems like a good place to start
y <- (dnorm(x, mean = 0, sd = 0.5))
plot(x, y, type = "l")

ktype_brm <- brm(preference.std ~ ktype,
                 prior = c(prior(normal(0, 0.5), class = b)),
                 iter = 2000,
                 chains = 4,
                 data = null_tab1)

conditional_effects(ktype_brm)

```

# Null model by site
```{r}
# Nest and join consumer and resource data frames
con_nest <- prop_visits_ktype %>%
  select(site, date, bb.sp, ktype, visits) %>%
  pivot_wider(values_from = visits, names_from = ktype, values_fill = 0) %>%
  select(-c("0", "10", "9")) %>%
  group_by(site) %>%
  nest() %>%
  rename(bees = data)

res_nest <- prop_visits_ktype %>%
  select(site, date, ktype, perc.flower.cover) %>%
  distinct() %>%
  pivot_wider(values_from = perc.flower.cover, names_from = ktype, values_fill = 0) %>%
  select(-c("0", "10", "9")) %>%
  group_by(site) %>%
  nest() %>%
  rename(plants = data)

conres <- full_join(con_nest, res_nest)

# Create a big beautiful table of data and models
sitewise_null_models <- conres %>%

  # Generate null models
  mutate(null.model = map2(bees, plants, function(x, y) 
    generate_null_net(consumers = x[, 2:9],
                      resources = y[, 2:8],
                      sims = 5000,
                      data.type = "counts",
                      c.samples = x[, 1],
                      r.samples = y[, 1]))) %>%

  # Tabulate results (with test)
  mutate(null.model.tab = map(null.model, function(x) 
    test_interactions(x, signif.level = 0.99) %>%
      as_tibble() %>%
      mutate(preference = Observed - Null,
             preference.upper99 = Observed - Upper.99.CL,
             preference.lower99 = Observed - Lower.99.CL) %>%
      select(bb.sp = Consumer, ktype = Resource, everything()))) %>%

  # Unnest and add site data and bumble bee trait data
  unnest(null.model.tab) %>%
  left_join(site_data) %>%
  left_join(bb_traits) %>%

  # Calculate preference/total abundance per species
  group_by(site, bb.sp) %>%
  mutate(total.abund = sum(Observed)) %>%
  ungroup() %>%
  mutate(preference.std = preference/total.abund)

write_rds(sitewise_null_models, "../output/nullmodel_ktype_sitewise.rds")
sitewise_null_models <- readRDS("../output/nullmodel_ktype_sitewise.rds")
```

## Visualize site model
```{r}
ggplot(sitewise_null_models, aes(elev.mean, preference, color = ktype)) +
  geom_point() +
  geom_smooth(method = "gam", formula = y ~ s(x, k = 5)) +
  facet_wrap(~bb.sp, scales = "free")

ggplot(sitewise_null_models, aes(elev.mean, SES, color = ktype)) +
  geom_point() +
  geom_smooth(method = "gam", formula = y ~ s(x, k = 5)) +
  facet_wrap(~bb.sp, scales = "free")

ggplot(sitewise_null_models, aes(elev.mean, preference.std, color = ktype)) +
  geom_point() +
  geom_smooth(method = "gam", formula = y ~ s(x, k = 5)) +
  facet_wrap(~bb.sp, scales = "free")
```