### 2021-06-09 code/ms2_nullmodels.Rmd

---
title: "visitation_bias"
author: "Doug Sponsler"
date: "12/14/2020"
output: html_document
---

# Preparations

## Packages 
```{r, warning=FALSE, include=FALSE, echo = FALSE}
library(econullnetr)
library(patchwork)
library(mgcv)
library(mgcViz)
library(tidymv)
library(tidytext)
library(tidyverse)
library(lubridate)
library(broom)
library(vegan)
library(ggvegan)
library(ggrepel)

Sys.setlocale(locale = "en_US.UTF-8")
```

## Load data
```{r message=FALSE, warning=FALSE, include=FALSE, echo = FALSE}
#visit_cover_ktype <- readRDS("../data/processed_data/visit_cover_ktype.rds")

visitation <- readRDS("../data/processed_data/visitation.rds")

survey <- read_csv("../data/processed_data/floral_survey.csv")

flower_cover_ktype <- readRDS("../data/processed_data/flower_cover_ktype.rds")

bb_traits <- read_csv("../data/processed_data/bb_traits.csv") %>%
  select(bb.sp, pbl.w, tongue = pbl.w.class2)

site_data <- read_csv("../data/processed_data/site_data.csv") %>%
  dplyr::select(-c(slope.calc, slope.est, elev.min, elev.max, temp.mean, elev.class, elev.class2)) %>% 
  mutate(site = factor(site),
         tree_line = factor(tree_line))

fl_traits <- readRDS("../data/processed_data/fl_traits.rds")

# For setting up the HPC
# 
# bb_traits <- read_csv("./bb_traits.csv") %>%
#   select(bb.sp, pbl.w, tongue = pbl.w.class2)
# 
# site_data <- read_csv("./site_data.csv") %>%
#   dplyr::select(-c(slope.calc, slope.est, elev.min, elev.max, temp.mean, elev.class, elev.class2)) %>%
#   mutate(site = factor(site),
#          tree_line = factor(tree_line))
```

# 1. Introducing Kugler's morphotypes

## 1.1 Summarize visitation at genus level
```{r}
morphotypes <- visitation %>%
  group_by(plant.genus, ktype) %>%
  mutate(visits = sum(visited)) %>%
  mutate(ktype = case_when(
    ktype == 0 ~ "null (0)",
    ktype == 1 ~ "disc (1)",
    ktype == 2 ~ "funnel (2)",
    ktype == 3 ~ "bell (3)",
    ktype == 4 ~ "stalk-disc (4)",
    ktype == 5 ~ "lip (5)",
    ktype == 6 ~ "flag (6)",
    ktype == 7 ~ "head (7)",
    ktype == 9 ~ "brush (9)"
  )) %>%
  mutate(ktype = factor(ktype, levels = c("null (0)", "disc (1)", "funnel (2)", 
                                          "bell (3)", "stalk-disc (4)", "lip (5)", 
                                          "flag (6)", "head (7)", "brush (9)"))) %>%
  ungroup() %>%
  mutate(plant.genus = reorder_within(plant.genus, -visits, ktype))

ggplot(morphotypes, aes(plant.genus)) +
  geom_bar() +
  facet_wrap(~ktype, scales = "free") +
  labs(y = "Visits", x = "Genus") +
  theme_light(14) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.25, size = 7),
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank()) +
  scale_x_reordered()

ggsave("../ms_2/figures/morphotype_genera.pdf", height = 8, width = 9)
```

## 
1.2 Summarize visitation at family level
```{r}
morphotypes_fam <- visitation %>%
  group_by(plant.family, ktype) %>%
  mutate(visits = sum(visited)) %>%
  mutate(ktype = case_when(
    ktype == 0 ~ "null (0)",
    ktype == 1 ~ "disc (1)",
    ktype == 2 ~ "funnel (2)",
    ktype == 3 ~ "bell (3)",
    ktype == 4 ~ "stalk-disc (4)",
    ktype == 5 ~ "lip (5)",
    ktype == 6 ~ "flag (6)",
    ktype == 7 ~ "head (7)",
    ktype == 9 ~ "brush (9)"
  )) %>%
  mutate(ktype = factor(ktype, levels = c("null (0)", "disc (1)", "funnel (2)", 
                                          "bell (3)", "stalk-disc (4)", "lip (5)", 
                                          "flag (6)", "head (7)", "brush (9)"))) %>%
  ungroup() %>%
  mutate(plant.family = reorder_within(plant.family, -visits, ktype))

ggplot(morphotypes_fam, aes(plant.family)) +
  geom_bar() +
  facet_wrap(~ktype, scales = "free") +
  labs(y = "Visits", x = "Plant family") +
  theme_light(14) +
  theme(axis.text.x = element_text(angle = 45, 
                                   hjust = 1, 
                                   vjust = 1, 
                                   size = 10),
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank()) +
  scale_x_reordered()

ggsave("../ms_2/figures/morphotype_families.pdf", height = 9, width = 9)

```

# 2. Null model 1: bb.sp x ktype, fully pooled
```{r}
# The "consumers" table input to generate_null_net()
con <- visitation %>%
  #filter(visited == 1) %>%
  mutate(individual.id = row_number()) %>%
  select(sample.id, individual.id, bb.sp, ktype, visited) %>%
  group_by(sample.id) %>%
  arrange(ktype) %>%
  pivot_wider(values_from = visited, names_from = ktype, values_fill = 0) %>%
  arrange(sample.id) %>%
  ungroup() %>%
  as.data.frame()

# The "resources" table input to generate_null_net()
res <- flower_cover_ktype %>%
  select(sample.id, ktype, cover) %>%
  group_by(sample.id) %>%
  arrange(ktype) %>%
  distinct() %>%
  pivot_wider(values_from = cover, names_from = ktype, values_fill = 0) %>%
  arrange(sample.id) %>%
  ungroup() %>%
  droplevels() %>%
  as.data.frame()

# Generate null model
null1 <- generate_null_net(consumers = con[, 3:12],
                           resources = res[, 2:10],
                           sims = 500, # raised from default of 100 (takes a while to run)
                           data.type = "names",
                           # c.samples = con$sample.id, # parsed by sample
                           # r.samples = res$sample.id,
                           c.samples = con[,1], # parsed by sample
                           r.samples = res[,1],
                           summary.type = "mean") # summarizing with the mean instead of the default sum standardizes across our bumble bee species that vary in total abundance,

saveRDS(null1, file = "../output/null1.rds")
null1 <- readRDS("../output/null1.rds")

# Tabulate output and generate tests
null_tab1 <- test_interactions(null1, signif.level = 0.99) %>% # make it a bit more conservative to account for multiple comparisons
  as_tibble() %>%
  select(bb.sp = Consumer, ktype = Resource, everything()) %>%
  group_by(bb.sp) %>%
  mutate(preference = Observed - Null, # a working definition of preference
         preference.upper99 = Observed - Upper.99.CL, 
         preference.lower99 = Observed - Lower.99.CL) %>%
  left_join(bb_traits) %>%
  ungroup() %>%
  mutate(bb.sp = fct_reorder(factor(bb.sp), pbl.w)) %>%
  mutate(ktype = case_when(
    ktype == 0 ~ "null",
    ktype == 1 ~ "disc",
    ktype == 2 ~ "funnel",
    ktype == 3 ~ "bell",
    ktype == 4 ~ "stalk-disc",
    ktype == 5 ~ "lip",
    ktype == 6 ~ "flag",
    ktype == 7 ~ "head",
    ktype == 9 ~ "brush"
  )) %>%
  mutate(ktype = factor(ktype, levels = c("null", "disc", "funnel", 
                                          "bell", "stalk-disc", "lip", 
                                          "flag", "head", "brush"))) %>%
  filter(!ktype %in% c("null", "brush")) %>% # type 0 were not reliably quantified (and rarely visited) and type 9 was vanishingly rare
  droplevels()
```

## 2.1 Observed vs. null predictions
```{r}
# Species-wise
ggplot(null_tab1, aes(Observed, ktype)) +
  geom_segment(aes(y = ktype, yend = ktype, 
                   x = Lower.99.CL, xend = Upper.99.CL), 
               color = "gray40", inherit.aes = FALSE) +
  geom_point(aes(color = Test)) +
  facet_wrap(~fct_rev(bb.sp)) +
  scale_y_discrete(limits = rev) +
  scale_color_manual(values = c("gray40", "#F8766D", "#00B6EB")) +
  labs(x = "Mean proportional visitation", y = "Floral morphotype", color = "Test (p < 0.01)") +
  theme_light()

ggsave("../ms_2/figures/null_model_spwise_preference.pdf", height = 6, width = 7)
ggsave("../ms_2/figures/null_model_spwise_preference.png", height = 6, width = 7)

## Morphotype-wise
ggplot(null_tab1, aes(Observed, bb.sp)) +
  geom_segment(aes(y = bb.sp, yend = bb.sp, 
                   x = Lower.99.CL, xend = Upper.99.CL), 
               color = "gray40", inherit.aes = FALSE) +
  geom_point(aes(color = Test)) +
  facet_wrap(~ktype, ncol = 4) +
  scale_color_manual(values = c("gray40", "#F8766D", "#00B6EB")) +
  labs(x = "Mean proportional visitation", y = "Bumble bee species", color = "Test (p < 0.01)") +
  theme_light()

ggsave("../ms_2/figures/null_model_ktypewise_preference.pdf", height = 5, width = 7)
ggsave("../ms_2/figures/null_model_ktypewise_preference.png", height = 5, width = 7)
```

## 2.2 Overall bias by morphotype
```{r}
ggplot(null_tab1, aes(ktype, preference)) +
  #geom_violin(fill = "gray70", color = "white", draw_quantiles = c(0.25, 0.5, 0.75)) +
  geom_boxplot(fill = NA, color = "black", outlier.shape = 4, size = 0.35) +
  geom_jitter(aes(color = Test), width = 0.3, alpha = 0.6) +
  theme_light(14) +
  scale_color_manual(values = c("gray40", "#F8766D", "#00B6EB")) +
  labs(x = "Floral morphotype",
       y = "Bias")

ggsave("../ms_2/figures/null_model_ktype_preference.pdf", height = 6, width = 7)
ggsave("../ms_2/figures/null_model_ktype_preference.png", height = 6, width = 7)
```

## 2.3 Bias by morphotype ~ tongue length
### 2.3.1 Preliminary visualization
```{r}
# Standardized preference ~ morphotype * tongue length
ggplot(null_tab1, aes(pbl.w, preference)) +
  geom_point(aes(color = Test)) +
  geom_smooth(method = "gam", formula = y ~ s(x, k = 4)) +
  scale_color_manual(values = c("gray40", "#F8766D", "#00B6EB")) +
  facet_wrap(~ktype) +
  labs(x = "Tongue length (mm)", y = "Preference", 
       color = "Test (p < 0.01)") +
  theme_light()
```

### 2.3.2 Model
```{r}
gam1 <- gam(preference ~ 
              ktype + 
              s(pbl.w, by = ktype, k = 5), 
            data = null_tab1) %>% getViz()

summary(gam1)
check.gamViz(gam1)
#plot.gamViz(gam1, allTerms = TRUE)

# Extract predictions
gam1_pred <- predict_gam(gam1) %>%
  mutate(fit.plus.se = fit + (2*se.fit),
         fit.minus.se = fit - (2*se.fit))

# Plot
ggplot(gam1_pred, aes(pbl.w, fit)) +
  geom_line() + # fit line
  geom_ribbon(aes(ymin = fit.minus.se, ymax = fit.plus.se), # Uncertainty
              fill = "black", alpha = 0.2) +
  geom_point(data = null_tab1, # Original data
             aes(pbl.w, preference, color = bb.sp), 
             inherit.aes = FALSE, alpha = 0.5) +
  #scale_color_manual(values = c("gray40", "#F8766D", "#00B6EB")) +
  facet_wrap(~ktype) +
  labs(x = "Tongue length (mm)", y = "Bias", 
       color = "Species") +
  theme_light()

ggsave("../ms_2/figures/null_model_ktypeXtonguelength_preference.pdf", height = 6, width = 7)
ggsave("../ms_2/figures/null_model_ktypeXtonguelength_preference.png", height = 6, width = 7)
```

## 2.4 Ordinate preferences

### 2.4.1 With all species
```{r}
# Bias matrix
null1_prefs <- null_tab1 %>%
  select(bb.sp, ktype, preference) %>%
  pivot_wider(names_from = ktype, values_from = preference) %>%
  column_to_rownames("bb.sp")

# Call PCA
pca_pref <- rda(null1_prefs)

# Extract tongue length data
tongues <- bb_traits %>%
  semi_join(null_tab1) %>%
  arrange(bb.sp)

# Ordisurf on tongue length
ordisurf_pref <- ordisurf(pca_pref, tongues$pbl.w) 
summary(ordisurf_pref)

# Get variance explained
pcoa_pref_pc1 <- round(pca_pref$CA$eig[[1]] / sum(pca_pref$CA$eig)*100)
pcoa_pref_pc2 <- round(pca_pref$CA$eig[[2]] / sum(pca_pref$CA$eig)*100)

# Fortify whole model
pca_pref.fort <- fortify(pca_pref, axes = 1:2) %>%
  # for plotting, I prefer ktype to be numerically coded
  mutate(Label = recode(Label,
                        "disc" = "1",
                        "funnel" = "2",
                        "bell" = "3",
                        "stalk-disc" = "4",
                        "lip" = "5",
                        "flag" = "6",
                        "head" = "7"))

# Fortify plants
pca_pref.fort1 <- fortify(pca_pref, axes = 1:2) %>%
  filter(Score == "species") %>%
  rename(ktype = Label) %>%
  # for plotting, I prefer ktype to be numerically coded
  mutate(ktype = recode(ktype,
                        "disc" = "1",
                        "funnel" = "2",
                        "bell" = "3",
                        "stalk-disc" = "4",
                        "lip" = "5",
                        "flag" = "6",
                        "head" = "7"))

# Fortify bees
pca_pref.fort2 <- fortify(pca_pref, axes = 1:2) %>%
  filter(Score == "sites") %>%
  rename(bb.sp = Label) %>%
  left_join(bb_traits)

# Plot ordisurf

## Plotting function based on https://oliviarata.wordpress.com/2014/07/17/ordinations-in-ggplot2-v2-ordisurf/
surfplot <- function(x, y, min_score = 0, pca1, pca2) {

  grid <- x$grid # extracts the ordisurf object
  grid_exp <- expand.grid(x = grid$x, y = grid$y) # get x and ys
  grid_exp$z <- as.vector(grid$z) # unravel the matrix for the z scores

  contours <- data.frame(na.omit(grid_exp)) %>% # gets rid of the nas
    as_tibble() %>%
    rename(PC1 = x, PC2 = y)

  scores <- y %>% 
    filter((abs(PC1) > min_score | abs(PC2) > min_score) | Score == "BB.sp") %>%
    mutate(max.score = max(PC1, PC2))

  ggplot(scores, aes(PC1, PC2, pch = Score)) +
    geom_contour(data = contours, 
                 aes(PC1, PC2, z = z, color = stat(level)),
                 size = 2, alpha = 0.4, inherit.aes = FALSE) +
    #geom_point(size = 2) +
    geom_text_repel(aes(PC1, PC2, label = Label), key_glyph = "rect") +
    theme_light() +
    scale_color_viridis_c(name = "Tongue length") +
    xlab(paste("PC1 ", "(", pca1, "%", ")", sep = "")) +
    ylab(paste("PC2 ", "(", pca2, "%", ")", sep = "")) +
    coord_fixed() # Gavin Simpson says this is **really** important
}

## Plot
bias_ord <- surfplot(ordisurf_pref, pca_pref.fort, 
         pca1 = pcoa_pref_pc1, pca2 = pcoa_pref_pc2)

ggsave("../ms_2/figures/preference_ordination.png", bias_ord, height = 5.5, width = 8)
ggsave("../ms_2/figures/preference_ordination.pdf", bias_ord, height = 5.5, width = 8)
```

### 2.4.2 Without B. gers, since it's a true specialist
```{r}
# Call PCA
pca_pref2 <- null1_prefs %>%
  slice(-1) %>% # drop B. gers
  rda()

# Extract tongue length data
tongues2 <- bb_traits %>%
  semi_join(null_tab1) %>%
  arrange(bb.sp) %>%
  filter(bb.sp != "gers")

# Ordisurf on tongue length
ordisurf_pref2 <- ordisurf(pca_pref2, tongues2$pbl.w) 
summary(ordisurf_pref2)

# Get variance explained
pcoa_pref2_pc1 <- round(pca_pref2$CA$eig[[1]] / sum(pca_pref2$CA$eig)*100)
pcoa_pref2_pc2 <- round(pca_pref2$CA$eig[[2]] / sum(pca_pref2$CA$eig)*100)

# Fortify whole model
pca_pref.fort2 <- fortify(pca_pref2, axes = 1:2) %>%
  # for plotting, I prefer ktype to be numerically coded
  mutate(Label = recode(Label,
                        "disc" = "1",
                        "funnel" = "2",
                        "bell" = "3",
                        "stalk-disc" = "4",
                        "lip" = "5",
                        "flag" = "6",
                        "head" = "7"))

# Fortify plants
pca_pref2.fort1 <- fortify(pca_pref2, axes = 1:2) %>%
  filter(Score == "species") %>%
  rename(ktype = Label) %>%
  # for plotting, I prefer ktype to be numerically coded
  mutate(ktype = recode(ktype,
                        "disc" = "1",
                        "funnel" = "2",
                        "bell" = "3",
                        "stalk-disc" = "4",
                        "lip" = "5",
                        "flag" = "6",
                        "head" = "7"))

# Fortify bees
pca_pref2.fort2 <- fortify(pca_pref2, axes = 1:2) %>%
  filter(Score == "sites") %>%
  rename(bb.sp = Label) %>%
  left_join(bb_traits)

# Plot ordisurf
surfplot(ordisurf_pref2, pca_pref.fort2, 
         pca1 = pcoa_pref2_pc1, pca2 = pcoa_pref2_pc2)
ggsave("../ms_2/figures/preference_ordination2.png")
ggsave("../ms_2/figures/preference_ordination2.pdf", width = 7, height = 7)
```

### 2.4.5 Cluster preferences
```{r}
clust1 <- null1_prefs %>%
  vegdist(method = "euclidean") %>%
  hclust(method = "complete")

pdf("../ms_2/figures/preference_clust.pdf", width = 5, height = 4)
plot(clust1, main = NULL, xlab = NULL, ylab = NULL)
dev.off()

clust2 <- null1_prefs %>%
  slice(-1) %>% # drop B. gers
  vegdist(method = "euclidean") %>%
  hclust(method = "complete")

pdf("../ms_2/figures/preference_clust2.pdf")
plot(clust2)
dev.off()
```

# 3. Null model 2: bb.sp x ktype, by site
```{r}
# Nest and join consumer and resource data frames
con2 <- visitation %>%
  mutate(individual.id = row_number()) %>%
  separate(sample.id, c("site", "date"), sep = "_") %>%
  select(site, date, individual.id, bb.sp, ktype, visited) %>%
  group_by(site, date) %>%
  arrange(ktype) %>%
  pivot_wider(values_from = visited, names_from = ktype, values_fill = 0) %>%
  group_by(site) %>%
  nest() %>%
  rename(bees = data)

res2 <- flower_cover_ktype %>%
  select(sample.id, ktype, cover) %>%
  separate(sample.id, c("site", "date"), sep = "_") %>%
  group_by(site, date) %>%
  arrange(ktype) %>%
  distinct() %>%
  pivot_wider(values_from = cover, names_from = ktype, values_fill = 0.01) %>%
  group_by(site) %>%
  nest() %>%
  rename(plants = data)

conres2 <- full_join(con2, res2)

# Create a big beautiful table of data and models (run on HPC!)
null2 <- conres2 %>%

  # Generate null models
  mutate(null.model = map2(bees, plants, function(x, y)
    generate_null_net(consumers = x[, 3:12],
                      resources = y[, 2:10],
                      sims = 1000, # raised from default of 100
                      data.type = "names",
                      c.samples = x[, 1],
                      r.samples = y[, 1],
                      summary.type = "mean")))

saveRDS(null2, "../output/null2.rds")
null2 <- readRDS("../output/null2.rds")

# Tabulate results (with test)
null_tab2 <- null2 %>%
  mutate(null.model.tab = map(null.model, function(x)
    test_interactions(x, signif.level = 0.99) %>%
      as_tibble() %>%
      mutate(preference = Observed - Null,
             preference.upper99 = Observed - Upper.99.CL,
             preference.lower99 = Observed - Lower.99.CL) %>%
      select(bb.sp = Consumer, ktype = Resource, everything()))) %>%

  # Unnest and add site data and bumble bee trait data
  unnest(null.model.tab) %>%
  select(site, bb.sp, ktype, Observed, Null, Lower.99.CL, Upper.99.CL, Test, 
         SES, preference, preference.upper99, preference.lower99) %>%
  left_join(site_data) %>%
  left_join(bb_traits) %>%
  group_by(site, bb.sp) %>%
  mutate(total.observed = sum(Observed)) %>%
  filter(total.observed > 0) %>% # for each species, drop sites where the species did not occur
  unite(bb.ktype, c(bb.sp, ktype), remove = FALSE) %>%
  mutate(ktype = factor(ktype),
         bb.sp = factor(bb.sp),
         bb.ktype = factor(bb.ktype)) %>%
  mutate(bb.sp = factor(bb.sp, levels = c("jone", "pyre", "soro", "psit", # order by ascending tongue length
                                          "telu", "lapi", "prat", "mont",
                                          "wurf", "muci", "pasc", "hypn",
                                          "mend", "hort", "gers"))) %>%
  mutate(ktype = case_when(
    ktype == 0 ~ "null",
    ktype == 1 ~ "disc",
    ktype == 2 ~ "funnel",
    ktype == 3 ~ "bell",
    ktype == 4 ~ "stalk-disc",
    ktype == 5 ~ "lip",
    ktype == 6 ~ "flag",
    ktype == 7 ~ "head",
    ktype == 9 ~ "brush"
  )) %>%
  mutate(ktype = factor(ktype, levels = c("null", "disc", "funnel", 
                                          "bell", "stalk-disc", "lip", 
                                          "flag", "head", "brush"))) %>%
  filter(!ktype %in% c("null", "brush")) %>% # type 0 were not reliably quantified (and rarely visited) and type 9 was vanishingly rare
  droplevels()
```

## 3.1 Lability of preference: Do BB species preferences for morphotypes vary with elevation?
### 3.1.1 Preliminary visualization
```{r}
ggplot(null_tab2, aes(elev.mean, preference)) +
  geom_point(aes(color = Test)) +
  geom_smooth(method = "gam", formula = y ~ s(x, k = 4), color = "black") +
  scale_color_manual(values = c("gray40", "#F8766D", "#00B6EB")) +
  facet_grid(bb.sp ~ ktype, scales = "free") +
  theme_light()
```

### 3.1.2 Model
```{r}
gam2 <- bam(preference ~
              s(bb.ktype, bs = "re") +
              s(elev.mean, by = bb.ktype, k = 4),
            method = "fREML",
            discrete = TRUE,
            select = TRUE,
            data = null_tab2) %>% getViz()

print(plot.gamViz(gam2, allTerms = TRUE), pages = 6)
check.gamViz(gam2)
summary(gam2, re.test = FALSE)

# Extract significant smooths
gam2_sig_smooths <- tidy(gam2) %>%
  filter(p.value < 0.05 & term != "s(bb.ktype)") %>%
  mutate(bb.ktype = str_extract(term, ".{6}$")) %>%
  select(bb.ktype)

# Extract predictions
gam2_pred <- predict_gam(gam2) %>%
  mutate(fit.plus.se = fit + (2*se.fit),
         fit.minus.se = fit - (2*se.fit)) %>%
  semi_join(gam2_sig_smooths) %>%
  separate(bb.ktype, c("bb.sp", "ktype"))

# Grab relevant subset of original data
null_tab2_sub <- null_tab2 %>%
  semi_join(gam2_sig_smooths)

# Plot
ggplot(gam2_pred, aes(elev.mean, fit)) +
  geom_line() + # fit line
  geom_point(data = null_tab2_sub, # Original data
             aes(elev.mean, preference, color = Test),
             inherit.aes = FALSE) +
  geom_ribbon(aes(ymin = fit.minus.se, ymax = fit.plus.se), # Uncertainty
              fill = "black", alpha = 0.2) +
  scale_color_manual(values = c("gray40", "#F8766D", "#00B6EB")) +
  facet_grid(bb.sp ~ ktype, scales = "free") +
  labs(x = "Elevation (m.s.l.)", y = "Standardized preference",
       color = "Test (p < 0.01)") +
  theme_light()
```

# 4. Null model 3: bb.sp x ktype, by site-date
```{r}
# Nest and join consumer and resource data frames

# For HPC
# library(tidyverse)
# library(econullnetr)
# flower_cover_ktype <- readRDS("./flower_cover_ktype.rds")
# visitation <- readRDS("./visitation.rds")

# The "consumers" table input to generate_null_net()
con3 <- visitation %>%
  mutate(individual.id = row_number()) %>%
  select(sample.id, individual.id, bb.sp, ktype, visited) %>%
  group_by(sample.id) %>%
  arrange(ktype) %>%
  pivot_wider(values_from = visited, names_from = ktype, values_fill = 0) %>%
  group_by(sample.id) %>%
  nest() %>%
  rename(bees = data)

# The "resources" table input to generate_null_net()
res3 <- flower_cover_ktype %>%
  select(sample.id, ktype, cover) %>%
  group_by(sample.id) %>%
  arrange(ktype) %>%
  distinct() %>%
  pivot_wider(values_from = cover, names_from = ktype, values_fill = 0) %>%
  group_by(sample.id) %>%
  nest() %>%
  rename(plants = data)

conres3 <- full_join(con3, res3) %>%
  separate(sample.id, c("site", "date"), sep = "_")

# Create a big beautiful table of data and models (run on HPC!)
# This will probably take 8 hours to run
null3 <- conres3 %>%

  # Generate null models
  mutate(null.model = map2(bees, plants, function(x, y)
    generate_null_net(consumers = x[, 2:11],
                      resources = y[, 1:9],
                      sims = 500, # raised from default of 100
                      data.type = "names",
                      prog.count = TRUE,
                      summary.type = "mean")))

#saveRDS(null3, "../output/null3.rds") # hpc

null3 <- readRDS("../output/null3.rds")

# Tabulate results (with test)
null_tab3 <- null3 %>%
  mutate(null.model.tab = map(null.model, function(x)
    test_interactions(x, signif.level = 0.95) %>%
      as_tibble() %>%
      mutate(preference = Observed - Null,
             preference.upper95 = Observed - Upper.95.CL,
             preference.lower95 = Observed - Lower.95.CL) %>%
      select(bb.sp = Consumer, ktype = Resource, everything()))) %>%

  # Unnest and add site data and bumble bee trait data
  unnest(null.model.tab) %>%
  select(site, date, bb.sp, ktype, Observed, Null, Lower.95.CL, Upper.95.CL, Test, 
         SES, preference, preference.upper95, preference.lower95) %>%
  left_join(site_data) %>%
  left_join(bb_traits) %>%
  group_by(site, date, bb.sp) %>%
  mutate(total.observed = sum(Observed)) %>% #There are a few cases in which total.observed does not equal unity. This is because bees occasionally visited plants in ktypes 0, 9, or 10, which we dropped from the data prior to null model calculation but after calculating proportional visitation. This is as it should be.
  ungroup() %>%
  filter(total.observed > 0) %>% # for each species, drop sites where the species did not occur
  filter(Observed != 0 | Null != 0) %>% # when both equal zero, it's probably because the ktype had zero abundance; in any case, it's meaningless to retain these, and they just cause zero-inflation; I checked, and every time Null = 0, Observed = 0 also. 
  unite(bb.ktype, c(bb.sp, ktype), remove = FALSE) %>%
  mutate(ktype = factor(ktype),
         site = factor(site),
         yday = yday(date),
         year = factor(year(date)),
         bb.ktype = factor(bb.ktype),
         bb.sp = factor(bb.sp, levels = c("jone", "pyre", "soro", "psit", # order by ascending tongue length
                                          "telu", "lapi", "prat", "mont",
                                          "wurf", "muci", "pasc", "hypn",
                                          "mend", "hort", "gers"))) %>%
  mutate(ktype = case_when(
    ktype == 0 ~ "null",
    ktype == 1 ~ "disc",
    ktype == 2 ~ "funnel",
    ktype == 3 ~ "bell",
    ktype == 4 ~ "stalk-disc",
    ktype == 5 ~ "lip",
    ktype == 6 ~ "flag",
    ktype == 7 ~ "head",
    ktype == 9 ~ "brush"
  )) %>%
  mutate(ktype = factor(ktype, levels = c("null", "disc", "funnel", 
                                          "bell", "stalk-disc", "lip", 
                                          "flag", "head", "brush"))) %>%
  filter(!ktype %in% c("null", "brush")) %>% # type 0 were not reliably quantified (and rarely visited) and type 9 was vanishingly rare
  droplevels()

ggplot(null_tab3, aes(ktype, preference)) +
  #geom_violin(fill = "gray70", color = "white", draw_quantiles = c(0.25, 0.5, 0.75)) +
  geom_boxplot(fill = NA, color = "black", outlier.shape = 4, size = 0.35) +
  geom_jitter(aes(color = Test), width = 0.3, alpha = 0.6) +
  theme_light(14) +
  scale_color_manual(values = c("gray40", "#F8766D", "#00B6EB")) +
  labs(x = "Floral morphotype",
       y = "Bias")
```

## 4.1 Model preference lability

### 4.1.1 Model 1 (without abundance covariate)
#### 4.1.1.1 Call model
```{r}
gam3.1 <- bam(preference ~
                bb.ktype +
                s(site, by = bb.ktype, bs = "re") +
                s(year, by = bb.ktype, bs = "re") +
                te(yday, elev.mean, by = bb.ktype),
              family = "gaussian",
              method = "fREML",
              discrete = TRUE,
              data = filter(null_tab3, 
                            bb.sp %in% c("hort", "prat", "telu", 
                                         "wurf", "soro", "pasc") &
                              ktype %in% c("bell", "lip", "flag", "head"))) %>% 
                getViz()

#check.gamViz(gam3.1)
#gam3.1_sum <- summary(gam3.1, re.test = FALSE)
#plot.gamViz(gam3.1, allTerms = TRUE)
```

#### 4.1.1.2 Plot model
```{r}
# Tabulate model summary
gam3.1_s_table <- gam3.1_sum$s.table %>%
  as.data.frame() %>%
  rownames_to_column("term") %>%
  filter(str_detect(term, "te\\(")) %>%
  mutate(term = str_extract(term, ".{6}$")) %>%
  separate(term, c("bb.sp", "ktype")) %>%
  mutate(ktype = case_when(
    ktype == 3 ~ "bell",
    ktype == 5 ~ "lip",
    ktype == 6 ~ "flag",
    ktype == 7 ~ "head"
  )) %>%
  mutate(pval = signif(`p-value`, digits = 3)) %>%
  select(bb.sp, ktype, edf, pval) %>%
  as_tibble() %>%
  mutate(ktype = factor(ktype, levels = c("bell", "lip", "flag", "head")))

# Extract plotting data.
hort_3_gam3.1 <- plot(sm(gam3.1, 49))$ggObj[[1]] %>%
  mutate(bb.sp = rep("hort", n()),
         ktype = rep("bell", n()))

hort_5_gam3.1 <- plot(sm(gam3.1, 50))$ggObj[[1]] %>%
  mutate(bb.sp = rep("hort", n()),
         ktype = rep("lip", n()))

hort_6_gam3.1 <- plot(sm(gam3.1, 51))$ggObj[[1]] %>%
  mutate(bb.sp = rep("hort", n()),
         ktype = rep("flag", n()))

hort_7_gam3.1 <- plot(sm(gam3.1, 52))$ggObj[[1]] %>%
  mutate(bb.sp = rep("hort", n()),
         ktype = rep("head", n()))

pasc_3_gam3.1 <- plot(sm(gam3.1, 53))$ggObj[[1]] %>%
  mutate(bb.sp = rep("pasc", n()),
         ktype = rep("bell", n()))

pasc_5_gam3.1 <- plot(sm(gam3.1, 54))$ggObj[[1]] %>%
  mutate(bb.sp = rep("pasc", n()),
         ktype = rep("lip", n()))

pasc_6_gam3.1 <- plot(sm(gam3.1, 55))$ggObj[[1]] %>%
  mutate(bb.sp = rep("pasc", n()),
         ktype = rep("flag", n()))

pasc_7_gam3.1 <- plot(sm(gam3.1, 56))$ggObj[[1]] %>%
  mutate(bb.sp = rep("pasc", n()),
         ktype = rep("head", n()))

prat_3_gam3.1 <- plot(sm(gam3.1, 57))$ggObj[[1]] %>%
  mutate(bb.sp = rep("prat", n()),
         ktype = rep("bell", n()))

prat_5_gam3.1 <- plot(sm(gam3.1, 58))$ggObj[[1]] %>%
  mutate(bb.sp = rep("prat", n()),
         ktype = rep("lip", n()))

prat_6_gam3.1 <- plot(sm(gam3.1, 59))$ggObj[[1]] %>%
  mutate(bb.sp = rep("prat", n()),
         ktype = rep("flag", n()))

prat_7_gam3.1 <- plot(sm(gam3.1, 60))$ggObj[[1]] %>%
  mutate(bb.sp = rep("prat", n()),
         ktype = rep("head", n()))

soro_3_gam3.1 <- plot(sm(gam3.1, 61))$ggObj[[1]] %>%
  mutate(bb.sp = rep("soro", n()),
         ktype = rep("bell", n()))

soro_5_gam3.1 <- plot(sm(gam3.1, 62))$ggObj[[1]] %>%
  mutate(bb.sp = rep("soro", n()),
         ktype = rep("lip", n()))

soro_6_gam3.1 <- plot(sm(gam3.1, 63))$ggObj[[1]] %>%
  mutate(bb.sp = rep("soro", n()),
         ktype = rep("flag", n()))

soro_7_gam3.1 <- plot(sm(gam3.1, 64))$ggObj[[1]] %>%
  mutate(bb.sp = rep("soro", n()),
         ktype = rep("head", n()))

telu_3_gam3.1 <- plot(sm(gam3.1, 65))$ggObj[[1]] %>%
  mutate(bb.sp = rep("telu", n()),
         ktype = rep("bell", n()))

telu_5_gam3.1 <- plot(sm(gam3.1, 66))$ggObj[[1]] %>%
  mutate(bb.sp = rep("telu", n()),
         ktype = rep("lip", n()))

telu_6_gam3.1 <- plot(sm(gam3.1, 67))$ggObj[[1]] %>%
  mutate(bb.sp = rep("telu", n()),
         ktype = rep("flag", n()))

telu_7_gam3.1 <- plot(sm(gam3.1, 68))$ggObj[[1]] %>%
  mutate(bb.sp = rep("telu", n()),
         ktype = rep("head", n()))

wurf_3_gam3.1 <- plot(sm(gam3.1, 69))$ggObj[[1]] %>%
  mutate(bb.sp = rep("wurf", n()),
         ktype = rep("bell", n()))

wurf_5_gam3.1 <- plot(sm(gam3.1, 70))$ggObj[[1]] %>%
  mutate(bb.sp = rep("wurf", n()),
         ktype = rep("lip", n()))

wurf_6_gam3.1 <- plot(sm(gam3.1, 71))$ggObj[[1]] %>%
  mutate(bb.sp = rep("wurf", n()),
         ktype = rep("flag", n()))

wurf_7_gam3.1 <- plot(sm(gam3.1, 72))$ggObj[[1]] %>%
  mutate(bb.sp = rep("wurf", n()),
         ktype = rep("head", n()))

plotting_data__gam3.1 <- bind_rows(hort_3_gam3.1, hort_5_gam3.1, hort_6_gam3.1, hort_7_gam3.1,
                           pasc_3_gam3.1, pasc_5_gam3.1, pasc_6_gam3.1, pasc_7_gam3.1,
                           prat_3_gam3.1, prat_5_gam3.1, prat_6_gam3.1, prat_7_gam3.1,
                           soro_3_gam3.1, soro_5_gam3.1, soro_6_gam3.1, soro_7_gam3.1,
                           telu_3_gam3.1, telu_5_gam3.1, telu_6_gam3.1, telu_7_gam3.1,
                           wurf_3_gam3.1, wurf_5_gam3.1, wurf_6_gam3.1, wurf_7_gam3.1) %>%
  group_by(bb.sp, ktype) %>%
  mutate(z.norm = scale(z, center = FALSE),
         ktype = factor(ktype, levels = c("bell", "lip", "flag", "head")),
         date = as_date(x)) %>%
  ungroup() %>%
  left_join(bb_traits) %>%
  left_join(gam3.1_s_table) %>%
  select(z, z.norm, date, x, elevation = y, se, 
         bb.sp, ktype, pbl.w, edf, pval) %>%
  mutate(sig = if_else(pval < 0.05, TRUE, FALSE))

ggplot(plotting_data__gam3.1, aes(date, elevation, z = z.norm, fill = z.norm)) +
  geom_raster() +
  geom_contour(color = "black",
               size = 0.1,
               bins = 40,
               aes(linetype = sig)) + # by fixing bins, we scale contours to each plot while letting the color ramp represent differences in scale across plots; I think this is a good compromise
  facet_grid(reorder(bb.sp, -pbl.w) ~ ktype) +
  scale_linetype_manual(values = c("dotted", "solid"))+
  scale_fill_viridis_c(na.value = "transparent") +
  labs(x = NULL, y = "Elevation (m)", fill = "Normalized\neffect", linetype = "p < 0.05") +
  theme_light(12) +
  theme(panel.grid = element_blank(),
        axis.text.x = element_text(angle = 90, vjust = 0.5))

ggsave("../ms_2/figures/preference_lability_gam.pdf", height = 5, width = 6.5)
ggsave("../ms_2/figures/preference_lability_gam.png", height = 5, width = 6.5)
```

### 4.2 Any residue of abundance?
Not just a residue. There is a strict mathematical constraint relating proportional abundance to inferred bias. A rare plant ca be strongly preferred but never strong avoided, while an abundant plant can be strongly avoided but never strongly preferred. Or, put more specifically, as relative abundance approaches zero, minimum bias approachs 0 and maximum bias approach 1. As relative abundance approaches 1, minimum bias approach -1 and maximum bias approaches zero. 
```{r}
ggplot(filter(null_tab3_join,
              bb.sp %in% c("hort", "prat", "telu", 
                           "wurf", "soro", "pasc") &
                ktype %in% c("bell", "lip", "flag", "head")), aes(prop.cover, preference)) +
  geom_point(alpha = 0.2) +
  geom_smooth(method = "lm") +
  facet_grid(bb.sp~ktype) +
  ylim(c(-1, 1))

ggplot(filter(null_tab3_join,
              bb.sp %in% c("hort", "prat", "telu", 
                           "wurf", "soro", "pasc") &
                ktype %in% c("bell", "lip", "flag", "head")), aes(cover, preference)) +
  geom_point(alpha = 0.2) +
  geom_smooth(method = "lm") +
  facet_grid(bb.sp~ktype, scales = "free") +
  ylim(c(-1, 1))

#########################################################
# Can we wiggle out of this problem via transformation? #
#########################################################
null_tab3.norm <- null_tab3_join %>%
  mutate(preference.norm = case_when(
    preference == 0 ~ 0,
    preference > 0 ~ preference/(1 - prop.cover),
    preference < 0 ~ (preference/(0 - prop.cover)) * -1
  ))

ggplot(filter(null_tab3.norm,
              bb.sp %in% c("hort", "prat", "telu", 
                           "wurf", "soro", "pasc") &
                ktype %in% c("bell", "lip", "flag", "head")), 
       aes(prop.cover, preference.norm)) +
  geom_point(alpha = 0.2) +
  geom_smooth(method = "lm") +
  facet_grid(bb.sp~ktype) +
  ylim(c(-1, 1))

```

### 4.3 A binomial model: significant preference 0/1
One way to get around this problem is to model bias as a binary response, either significantly positive or not. 
#### 4.3.1 Call model
```{r}
null_tab3_bin <- null_tab3 %>%
  mutate(preference.bin = as.integer(if_else(Test == "Stronger", 1, 0)))

gam3.2 <- bam(preference.bin ~
                bb.ktype +
                s(site, by = bb.ktype, bs = "re") +
                s(year, by = bb.ktype, bs = "re") +
                te(yday, elev.mean, by = bb.ktype),
              family = "binomial",
              method = "fREML",
              discrete = TRUE,
              data = filter(null_tab3_bin, 
                            bb.sp %in% c("hort", "prat", "telu", 
                                         "wurf", "soro", "pasc") &
                              ktype %in% c("bell", "lip", "flag", "head"))) %>% 
                getViz()

check.gamViz(gam3.2)
gam3.2_sum <- summary(gam3.2, re.test = FALSE)
print(plot.gamViz(gam3.2, allTerms = TRUE), pages = 4)
```

#### 4.3.2 Plot model
```{r}
# Tabulate model summary
gam3.2_s_table <- gam3.2_sum$s.table %>%
  as.data.frame() %>%
  rownames_to_column("term") %>%
  filter(str_detect(term, "te\\(")) %>%
  mutate(term = str_extract(term, ".{6}$")) %>%
  separate(term, c("bb.sp", "ktype")) %>%
  mutate(ktype = case_when(
    ktype == 3 ~ "bell",
    ktype == 5 ~ "lip",
    ktype == 6 ~ "flag",
    ktype == 7 ~ "head"
  )) %>%
  mutate(pval = signif(`p-value`, digits = 3)) %>%
  select(bb.sp, ktype, edf, pval) %>%
  as_tibble() %>%
  mutate(ktype = factor(ktype, levels = c("bell", "lip", "flag", "head")))

# Extract plotting data.
hort_3 <- plot(sm(gam3.2, 49))$ggObj[[1]] %>%
  mutate(bb.sp = rep("hort", n()),
         ktype = rep("bell", n()))

hort_5 <- plot(sm(gam3.2, 50))$ggObj[[1]] %>%
  mutate(bb.sp = rep("hort", n()),
         ktype = rep("lip", n()))

hort_6 <- plot(sm(gam3.2, 51))$ggObj[[1]] %>%
  mutate(bb.sp = rep("hort", n()),
         ktype = rep("flag", n()))

hort_7 <- plot(sm(gam3.2, 52))$ggObj[[1]] %>%
  mutate(bb.sp = rep("hort", n()),
         ktype = rep("head", n()))

pasc_3 <- plot(sm(gam3.2, 53))$ggObj[[1]] %>%
  mutate(bb.sp = rep("pasc", n()),
         ktype = rep("bell", n()))

pasc_5 <- plot(sm(gam3.2, 54))$ggObj[[1]] %>%
  mutate(bb.sp = rep("pasc", n()),
         ktype = rep("lip", n()))

pasc_6 <- plot(sm(gam3.2, 55))$ggObj[[1]] %>%
  mutate(bb.sp = rep("pasc", n()),
         ktype = rep("flag", n()))

pasc_7 <- plot(sm(gam3.2, 56))$ggObj[[1]] %>%
  mutate(bb.sp = rep("pasc", n()),
         ktype = rep("head", n()))

prat_3 <- plot(sm(gam3.2, 57))$ggObj[[1]] %>%
  mutate(bb.sp = rep("prat", n()),
         ktype = rep("bell", n()))

prat_5 <- plot(sm(gam3.2, 58))$ggObj[[1]] %>%
  mutate(bb.sp = rep("prat", n()),
         ktype = rep("lip", n()))

prat_6 <- plot(sm(gam3.2, 59))$ggObj[[1]] %>%
  mutate(bb.sp = rep("prat", n()),
         ktype = rep("flag", n()))

prat_7 <- plot(sm(gam3.2, 60))$ggObj[[1]] %>%
  mutate(bb.sp = rep("prat", n()),
         ktype = rep("head", n()))

soro_3 <- plot(sm(gam3.2, 61))$ggObj[[1]] %>%
  mutate(bb.sp = rep("soro", n()),
         ktype = rep("bell", n()))

soro_5 <- plot(sm(gam3.2, 62))$ggObj[[1]] %>%
  mutate(bb.sp = rep("soro", n()),
         ktype = rep("lip", n()))

soro_6 <- plot(sm(gam3.2, 63))$ggObj[[1]] %>%
  mutate(bb.sp = rep("soro", n()),
         ktype = rep("flag", n()))

soro_7 <- plot(sm(gam3.2, 64))$ggObj[[1]] %>%
  mutate(bb.sp = rep("soro", n()),
         ktype = rep("head", n()))

telu_3 <- plot(sm(gam3.2, 65))$ggObj[[1]] %>%
  mutate(bb.sp = rep("telu", n()),
         ktype = rep("bell", n()))

telu_5 <- plot(sm(gam3.2, 66))$ggObj[[1]] %>%
  mutate(bb.sp = rep("telu", n()),
         ktype = rep("lip", n()))

telu_6 <- plot(sm(gam3.2, 67))$ggObj[[1]] %>%
  mutate(bb.sp = rep("telu", n()),
         ktype = rep("flag", n()))

telu_7 <- plot(sm(gam3.2, 68))$ggObj[[1]] %>%
  mutate(bb.sp = rep("telu", n()),
         ktype = rep("head", n()))

wurf_3 <- plot(sm(gam3.2, 69))$ggObj[[1]] %>%
  mutate(bb.sp = rep("wurf", n()),
         ktype = rep("bell", n()))

wurf_5 <- plot(sm(gam3.2, 70))$ggObj[[1]] %>%
  mutate(bb.sp = rep("wurf", n()),
         ktype = rep("lip", n()))

wurf_6 <- plot(sm(gam3.2, 71))$ggObj[[1]] %>%
  mutate(bb.sp = rep("wurf", n()),
         ktype = rep("flag", n()))

wurf_7 <- plot(sm(gam3.2, 72))$ggObj[[1]] %>%
  mutate(bb.sp = rep("wurf", n()),
         ktype = rep("head", n()))

plotting_data <- bind_rows(hort_3, hort_5, hort_6, hort_7,
                           pasc_3, pasc_5, pasc_6, pasc_7,
                           prat_3, prat_5, prat_6, prat_7,
                           soro_3, soro_5, soro_6, soro_7,
                           telu_3, telu_5, telu_6, telu_7,
                           wurf_3, wurf_5, wurf_6, wurf_7) %>%
  group_by(bb.sp, ktype) %>%
  mutate(z.norm = scale(z, center = FALSE),
         ktype = factor(ktype, levels = c("bell", "lip", "flag", "head")),
         date = as_date(x)) %>%
  ungroup() %>%
  left_join(bb_traits) %>%
  left_join(gam3.2_s_table) %>%
  select(z, z.norm, date, x, elevation = y, se, 
         bb.sp, ktype, pbl.w, edf, pval) %>%
  mutate(sig = if_else(pval < 0.05, TRUE, FALSE))

ggplot(plotting_data, aes(date, elevation, z = z.norm, fill = z.norm)) +
  geom_raster() +
  geom_contour(color = "black",
               size = 0.1,
               bins = 40,
               aes(linetype = sig)) + # by fixing bins, we scale contours to each plot while letting the color ramp represent differences in scale across plots; I think this is a good compromise
  facet_grid(reorder(bb.sp, -pbl.w) ~ ktype) +
  scale_linetype_manual(values = c("dotted", "solid"))+
  scale_fill_viridis_c(na.value = "transparent") +
  labs(x = NULL, y = "Elevation (m)", fill = "Normalized\neffect", linetype = "p < 0.05") +
  theme_light(12) +
  theme(panel.grid = element_blank(),
        axis.text.x = element_text(angle = 90, vjust = 0.5))
```

### 4.4 A normalized model: proportion of maximum/minimum preference
#### 4.4.1 Call model
```{r}
gam3.3 <- bam(preference.norm ~
                bb.ktype +
                s(site, by = bb.ktype, bs = "re") +
                s(year, by = bb.ktype, bs = "re") +
                te(yday, elev.mean, by = bb.ktype),
              family = "gaussian",
              method = "fREML",
              discrete = TRUE,
              data = filter(null_tab3.norm, 
                            bb.sp %in% c("hort", "prat", "telu", 
                                         "wurf", "soro", "pasc") &
                              ktype %in% c("bell", "lip", "flag", "head"))) %>% 
                getViz()

check.gamViz(gam3.3)
gam3.3_sum <- summary(gam3.3, re.test = FALSE)
print(plot.gamViz(gam3.3, allTerms = TRUE), pages = 4)
```

#### 4.4.2 Plot model
```{r}
# Tabulate model summary
gam3.3_s_table <- gam3.3_sum$s.table %>%
  as.data.frame() %>%
  rownames_to_column("term") %>%
  filter(str_detect(term, "te\\(")) %>%
  mutate(term = str_extract(term, ".{6}$")) %>%
  separate(term, c("bb.sp", "ktype")) %>%
  mutate(ktype = case_when(
    ktype == 3 ~ "bell",
    ktype == 5 ~ "lip",
    ktype == 6 ~ "flag",
    ktype == 7 ~ "head"
  )) %>%
  mutate(pval = signif(`p-value`, digits = 3)) %>%
  select(bb.sp, ktype, edf, pval) %>%
  as_tibble() %>%
  mutate(ktype = factor(ktype, levels = c("bell", "lip", "flag", "head")))

# Extract plotting data.
hort_3 <- plot(sm(gam3.3, 49))$ggObj[[1]] %>%
  mutate(bb.sp = rep("hort", n()),
         ktype = rep("bell", n()))

hort_5 <- plot(sm(gam3.3, 50))$ggObj[[1]] %>%
  mutate(bb.sp = rep("hort", n()),
         ktype = rep("lip", n()))

hort_6 <- plot(sm(gam3.3, 51))$ggObj[[1]] %>%
  mutate(bb.sp = rep("hort", n()),
         ktype = rep("flag", n()))

hort_7 <- plot(sm(gam3.3, 52))$ggObj[[1]] %>%
  mutate(bb.sp = rep("hort", n()),
         ktype = rep("head", n()))

pasc_3 <- plot(sm(gam3.3, 53))$ggObj[[1]] %>%
  mutate(bb.sp = rep("pasc", n()),
         ktype = rep("bell", n()))

pasc_5 <- plot(sm(gam3.3, 54))$ggObj[[1]] %>%
  mutate(bb.sp = rep("pasc", n()),
         ktype = rep("lip", n()))

pasc_6 <- plot(sm(gam3.3, 55))$ggObj[[1]] %>%
  mutate(bb.sp = rep("pasc", n()),
         ktype = rep("flag", n()))

pasc_7 <- plot(sm(gam3.3, 56))$ggObj[[1]] %>%
  mutate(bb.sp = rep("pasc", n()),
         ktype = rep("head", n()))

prat_3 <- plot(sm(gam3.3, 57))$ggObj[[1]] %>%
  mutate(bb.sp = rep("prat", n()),
         ktype = rep("bell", n()))

prat_5 <- plot(sm(gam3.3, 58))$ggObj[[1]] %>%
  mutate(bb.sp = rep("prat", n()),
         ktype = rep("lip", n()))

prat_6 <- plot(sm(gam3.3, 59))$ggObj[[1]] %>%
  mutate(bb.sp = rep("prat", n()),
         ktype = rep("flag", n()))

prat_7 <- plot(sm(gam3.3, 60))$ggObj[[1]] %>%
  mutate(bb.sp = rep("prat", n()),
         ktype = rep("head", n()))

soro_3 <- plot(sm(gam3.3, 61))$ggObj[[1]] %>%
  mutate(bb.sp = rep("soro", n()),
         ktype = rep("bell", n()))

soro_5 <- plot(sm(gam3.3, 62))$ggObj[[1]] %>%
  mutate(bb.sp = rep("soro", n()),
         ktype = rep("lip", n()))

soro_6 <- plot(sm(gam3.3, 63))$ggObj[[1]] %>%
  mutate(bb.sp = rep("soro", n()),
         ktype = rep("flag", n()))

soro_7 <- plot(sm(gam3.3, 64))$ggObj[[1]] %>%
  mutate(bb.sp = rep("soro", n()),
         ktype = rep("head", n()))

telu_3 <- plot(sm(gam3.3, 65))$ggObj[[1]] %>%
  mutate(bb.sp = rep("telu", n()),
         ktype = rep("bell", n()))

telu_5 <- plot(sm(gam3.3, 66))$ggObj[[1]] %>%
  mutate(bb.sp = rep("telu", n()),
         ktype = rep("lip", n()))

telu_6 <- plot(sm(gam3.3, 67))$ggObj[[1]] %>%
  mutate(bb.sp = rep("telu", n()),
         ktype = rep("flag", n()))

telu_7 <- plot(sm(gam3.3, 68))$ggObj[[1]] %>%
  mutate(bb.sp = rep("telu", n()),
         ktype = rep("head", n()))

wurf_3 <- plot(sm(gam3.3, 69))$ggObj[[1]] %>%
  mutate(bb.sp = rep("wurf", n()),
         ktype = rep("bell", n()))

wurf_5 <- plot(sm(gam3.3, 70))$ggObj[[1]] %>%
  mutate(bb.sp = rep("wurf", n()),
         ktype = rep("lip", n()))

wurf_6 <- plot(sm(gam3.3, 71))$ggObj[[1]] %>%
  mutate(bb.sp = rep("wurf", n()),
         ktype = rep("flag", n()))

wurf_7 <- plot(sm(gam3.3, 72))$ggObj[[1]] %>%
  mutate(bb.sp = rep("wurf", n()),
         ktype = rep("head", n()))

plotting_data <- bind_rows(hort_3, hort_5, hort_6, hort_7,
                           pasc_3, pasc_5, pasc_6, pasc_7,
                           prat_3, prat_5, prat_6, prat_7,
                           soro_3, soro_5, soro_6, soro_7,
                           telu_3, telu_5, telu_6, telu_7,
                           wurf_3, wurf_5, wurf_6, wurf_7) %>%
  group_by(bb.sp, ktype) %>%
  mutate(z.norm = scale(z, center = FALSE),
         ktype = factor(ktype, levels = c("bell", "lip", "flag", "head")),
         date = as_date(x)) %>%
  ungroup() %>%
  left_join(bb_traits) %>%
  left_join(gam3.3_s_table) %>%
  select(z, z.norm, date, x, elevation = y, se, 
         bb.sp, ktype, pbl.w, edf, pval) %>%
  mutate(sig = if_else(pval < 0.05, TRUE, FALSE))

ggplot(plotting_data, aes(date, elevation, z = z.norm, fill = z.norm)) +
  geom_raster() +
  geom_contour(color = "black",
               size = 0.1,
               bins = 40,
               aes(linetype = sig)) + # by fixing bins, we scale contours to each plot while letting the color ramp represent differences in scale across plots; I think this is a good compromise
  facet_grid(reorder(bb.sp, -pbl.w) ~ ktype) +
  scale_linetype_manual(values = c("dotted", "solid"))+
  scale_fill_viridis_c(na.value = "transparent") +
  labs(x = NULL, y = "Elevation (m)", fill = "Normalized\neffect", linetype = "p < 0.05") +
  theme_light(12) +
  theme(panel.grid = element_blank(),
        axis.text.x = element_text(angle = 90, vjust = 0.5))
```

### 4.1.2 Model 2 (with prop_abund as covariate)
```{r}
# With abundance covariate (interacting with bb-ktype pair)

## Add abundance data
abundance <- flower_cover_ktype %>%
  separate(sample.id, c("site", "date"), sep = "_") %>%
  group_by(site, date) %>%
  mutate(prop.cover = cover/sum(cover)) %>%
  mutate(ktype = case_when(
    ktype == 0 ~ "null",
    ktype == 1 ~ "disc",
    ktype == 2 ~ "funnel",
    ktype == 3 ~ "bell",
    ktype == 4 ~ "stalk-disc",
    ktype == 5 ~ "lip",
    ktype == 6 ~ "flag",
    ktype == 7 ~ "head",
    ktype == 9 ~ "brush"
  )) %>%
  mutate(ktype = factor(ktype))

null_tab3_join <- null_tab3 %>%
  left_join(abundance) %>%
  mutate(site = factor(site),
         year = factor(year(date)))

## Model with proportional cover data
gam3.2 <- bam(preference ~
                bb.ktype +
                #prop.cover*bb.ktype +
                cover:bb.ktype +
                s(site, by = bb.ktype, bs = "re") +
                s(year, by = bb.ktype, bs = "re") +
                te(yday, elev.mean, by = bb.ktype),
              family = "gaussian",
              method = "fREML",
              discrete = TRUE,
              data = filter(null_tab3_join, 
                            bb.sp %in% c("hort", "prat", "telu", 
                                         "wurf", "soro", "pasc") &
                              ktype %in% c("bell", "lip", "flag", "head"))) %>% 
                getViz()

#check.gamViz(gam3.2)
#gam3.2_corr_sum <- summary(gam3.2, re.test = FALSE)
#print(plot.gamViz(gam3.2, allTerms = TRUE), pages = 4)
```

Since we have a markedly negative effect of abundance on preference, due to the fact that an avoided resource will seem more avoided when it is abundant than when it is rare, even if the cognitive evaluation of the resource does not vary, perhaps it would be worth adding abundance as a covariate to the preference lability model. We do this in model 3.2.

### 4.1.3 Year-specific models
```{r}
# Year-specific models
gam3_2010 <- bam(preference ~
              bb.ktype +
              s(site, by = bb.ktype, bs = "re") +
              te(yday, elev.mean, by = bb.ktype),
            family = "gaussian",
            method = "fREML",
            discrete = TRUE,
            data = filter(null_tab3, 
                          year == 2010 &
                            bb.sp %in% c("hort", "prat", "telu", 
                                       "wurf", "soro", "pasc") &
                            ktype %in% c("3", "5", "6", "7"))) %>% 
              getViz()

check.gamViz(gam3_2010)
gam3_2010_sum <- summary(gam3_2010)
plot.gamViz(gam3_2010, allTerms = TRUE)

gam3_2011 <- bam(preference ~
              bb.ktype +
              s(site, by = bb.ktype, bs = "re") +
              te(yday, elev.mean, by = bb.ktype),
            family = "gaussian",
            method = "fREML",
            discrete = TRUE,
            data = filter(null_tab3, 
                          year == 2011 &
                            bb.sp %in% c("hort", "prat", "telu", 
                                       "wurf", "soro", "pasc") &
                            ktype %in% c("3", "5", "6", "7"))) %>% 
              getViz()

check.gamViz(gam3_2011)
gam3_2011_sum <- summary(gam3_2011)
plot.gamViz(gam3_2011, allTerms = TRUE)

gam3_2012 <- bam(preference ~
              bb.ktype +
              s(site, by = bb.ktype, bs = "re") +
              te(yday, elev.mean, by = bb.ktype),
            family = "gaussian",
            method = "fREML",
            discrete = TRUE,
            data = filter(null_tab3, 
                          year == 2012 &
                            bb.sp %in% c("hort", "prat", "telu", 
                                       "wurf", "soro", "pasc") &
                            ktype %in% c("3", "5", "6", "7"))) %>% 
              getViz()

check.gamViz(gam3_2012)
gam3_2012_sum <- summary(gam3_2012)
plot.gamViz(gam3_2012, allTerms = TRUE)
```

```{r}

gam3_corr_s_table <- gam3_corr_sum$s.table %>%
  as.data.frame() %>%
  rownames_to_column("term") %>%
  filter(str_detect(term, "te\\(")) %>%
  mutate(term = str_extract(term, ".{6}$")) %>%
  separate(term, c("bb.sp", "ktype")) %>%
  mutate(ktype = case_when(
    ktype == 3 ~ "bell",
    ktype == 5 ~ "lip",
    ktype == 6 ~ "flag",
    ktype == 7 ~ "head"
  )) %>%
  mutate(pval = signif(`p-value`, digits = 3)) %>%
  select(bb.sp, ktype, edf, pval) %>%
  as_tibble() %>%
  mutate(ktype = factor(ktype, levels = c("bell", "lip", "flag", "head")))

# Extract plotting data.

hort_3_corr <- plot(sm(gam3_corr, 49))$ggObj[[1]] %>%
  mutate(bb.sp = rep("hort", n()),
         ktype = rep("bell", n()))

hort_5_corr <- plot(sm(gam3_corr, 50))$ggObj[[1]] %>%
  mutate(bb.sp = rep("hort", n()),
         ktype = rep("lip", n()))

hort_6_corr <- plot(sm(gam3_corr, 51))$ggObj[[1]] %>%
  mutate(bb.sp = rep("hort", n()),
         ktype = rep("flag", n()))

hort_7_corr <- plot(sm(gam3_corr, 52))$ggObj[[1]] %>%
  mutate(bb.sp = rep("hort", n()),
         ktype = rep("head", n()))

pasc_3_corr <- plot(sm(gam3_corr, 53))$ggObj[[1]] %>%
  mutate(bb.sp = rep("pasc", n()),
         ktype = rep("bell", n()))

pasc_5_corr <- plot(sm(gam3_corr, 54))$ggObj[[1]] %>%
  mutate(bb.sp = rep("pasc", n()),
         ktype = rep("lip", n()))

pasc_6_corr <- plot(sm(gam3_corr, 55))$ggObj[[1]] %>%
  mutate(bb.sp = rep("pasc", n()),
         ktype = rep("flag", n()))

pasc_7_corr <- plot(sm(gam3_corr, 56))$ggObj[[1]] %>%
  mutate(bb.sp = rep("pasc", n()),
         ktype = rep("head", n()))

prat_3_corr <- plot(sm(gam3_corr, 57))$ggObj[[1]] %>%
  mutate(bb.sp = rep("prat", n()),
         ktype = rep("bell", n()))

prat_5_corr <- plot(sm(gam3_corr, 58))$ggObj[[1]] %>%
  mutate(bb.sp = rep("prat", n()),
         ktype = rep("lip", n()))

prat_6_corr <- plot(sm(gam3_corr, 59))$ggObj[[1]] %>%
  mutate(bb.sp = rep("prat", n()),
         ktype = rep("flag", n()))

prat_7_corr <- plot(sm(gam3_corr, 60))$ggObj[[1]] %>%
  mutate(bb.sp = rep("prat", n()),
         ktype = rep("head", n()))

soro_3_corr <- plot(sm(gam3_corr, 61))$ggObj[[1]] %>%
  mutate(bb.sp = rep("soro", n()),
         ktype = rep("bell", n()))

soro_5_corr <- plot(sm(gam3_corr, 62))$ggObj[[1]] %>%
  mutate(bb.sp = rep("soro", n()),
         ktype = rep("lip", n()))

soro_6_corr <- plot(sm(gam3_corr, 63))$ggObj[[1]] %>%
  mutate(bb.sp = rep("soro", n()),
         ktype = rep("flag", n()))

soro_7_corr <- plot(sm(gam3_corr, 64))$ggObj[[1]] %>%
  mutate(bb.sp = rep("soro", n()),
         ktype = rep("head", n()))

telu_3_corr <- plot(sm(gam3_corr, 65))$ggObj[[1]] %>%
  mutate(bb.sp = rep("telu", n()),
         ktype = rep("bell", n()))

telu_5_corr <- plot(sm(gam3_corr, 66))$ggObj[[1]] %>%
  mutate(bb.sp = rep("telu", n()),
         ktype = rep("lip", n()))

telu_6_corr <- plot(sm(gam3_corr, 67))$ggObj[[1]] %>%
  mutate(bb.sp = rep("telu", n()),
         ktype = rep("flag", n()))

telu_7_corr <- plot(sm(gam3_corr, 68))$ggObj[[1]] %>%
  mutate(bb.sp = rep("telu", n()),
         ktype = rep("head", n()))

wurf_3_corr <- plot(sm(gam3_corr, 69))$ggObj[[1]] %>%
  mutate(bb.sp = rep("wurf", n()),
         ktype = rep("bell", n()))

wurf_5_corr <- plot(sm(gam3_corr, 70))$ggObj[[1]] %>%
  mutate(bb.sp = rep("wurf", n()),
         ktype = rep("lip", n()))

wurf_6_corr <- plot(sm(gam3_corr, 71))$ggObj[[1]] %>%
  mutate(bb.sp = rep("wurf", n()),
         ktype = rep("flag", n()))

wurf_7_corr <- plot(sm(gam3_corr, 72))$ggObj[[1]] %>%
  mutate(bb.sp = rep("wurf", n()),
         ktype = rep("head", n()))

plotting_data_corr <- bind_rows(hort_3_corr, hort_5_corr, hort_6_corr, hort_7_corr,
                           pasc_3_corr, pasc_5_corr, pasc_6_corr, pasc_7_corr,
                           prat_3_corr, prat_5_corr, prat_6_corr, prat_7_corr,
                           soro_3_corr, soro_5_corr, soro_6_corr, soro_7_corr,
                           telu_3_corr, telu_5_corr, telu_6_corr, telu_7_corr,
                           wurf_3_corr, wurf_5_corr, wurf_6_corr, wurf_7_corr) %>%
  group_by(bb.sp, ktype) %>%
  mutate(z.norm = scale(z, center = FALSE),
         ktype = factor(ktype, levels = c("bell", "lip", "flag", "head")),
         date = as_date(x)) %>%
  ungroup() %>%
  left_join(bb_traits) %>%
  left_join(gam3_corr_s_table) %>%
  select(z, z.norm, date, x, elevation = y, se, 
         bb.sp, ktype, pbl.w, edf, pval) %>%
  filter(pval < 0.05)

ggplot(plotting_data_corr, aes(date, elevation, z = z.norm, fill = z.norm)) +
  geom_raster() +
  geom_contour(color = "black",
               size = 0.1,
               bins = 40) + # by fixing bins, we scale contours to each plot while letting the color ramp represent differences in scale across plots; I think this is a good compromise
  facet_grid(reorder(bb.sp, -pbl.w) ~ ktype) +
  scale_fill_viridis_c(na.value = "transparent") +
  labs(x = NULL, y = "Elevation (m)", fill = "Normalized\neffect") +
  theme_light(12) +
  theme(panel.grid = element_blank(),
        axis.text.x = element_text(angle = 90, vjust = 0.5))

ggsave("../output/preference_lability_gam_corr.pdf", height = 5, width = 6.5)
```

# 5 Raw visitation rates
Null model analysis is potentially problematic. It is still wroth performing and presenting, but we should compare it to patterns in raw visitation to understand "preference" more completely. The problem is well-illustrated with an anecdote. Suppose I am shopping for Molly, and I need to find gluten-free flour. In a typical grocery store, gluten-free flour is hard to come by, so my selection of it would appear to be a strong preference --- after all, I passed up so many other options to get one of the few options I most preferred. If I went to a fancy healthfood store, though, gluten-free flours would probably make up a large proportion of available options, and my decision would seem consistent with a null model of random flour selection. In both cases, I was completely committed to choosing a gluten-free flour; I preferred it perfectly. But in one case, it looks like preference, and in the other it doesn't. This same problem applies to the use of null models to infer preference in bumble bees. If, for example, B. mucidus  is committed to foraging on flag flowers in all floristic contexts, the only patterns of preference we will find are artifacts of relative abundance, created precisely by the technique intended to correct for such artifacts. I think the best approach is to present both a null-model based analysis and a raw visitation analysis and try to say something about which is a more appropriate representation of bumble bee floral selection. Are they more like Balmford's students in a cafeteria deciding whether to opt for a vegetarian meal, or are they more like Molly shopping for g-free flour? Neither null model residuals --- which we will call "bias" --- nor raw visitation values can be interpreted unambiguously as "preference". "Preference" lies latent behind these observable patterns.

## 5.1 Overall visitation rates
```{r}
# Get mean proportional visitation by morphotype*bb.sp
visitation_prop <- visitation %>% 
  group_by(sample.id, bb.sp, ktype) %>%
  summarize(visits = n()) %>%
  group_by(sample.id, bb.sp) %>%
  mutate(prop.visits = visits/sum(visits)) %>%
  select(sample.id, bb.sp, ktype, prop.visits) %>%
  group_by(bb.sp) %>%
  nest() %>%
  mutate(fill = map(data, function(x) full_join(x, flower_cover_ktype) %>% replace_na(replace = list(prop.visits = 0)))) %>%
  unnest(fill) %>%
  select(-data) %>%
  group_by(bb.sp, ktype) %>%
  summarize(mean.prop.visits = mean(prop.visits)) %>%
  filter(ktype != 10) %>%
   mutate(ktype = case_when(
    ktype == 0 ~ "null",
    ktype == 1 ~ "disc",
    ktype == 2 ~ "funnel",
    ktype == 3 ~ "bell",
    ktype == 4 ~ "stalk-disc",
    ktype == 5 ~ "lip",
    ktype == 6 ~ "flag",
    ktype == 7 ~ "head",
    ktype == 9 ~ "brush"
  )) %>%
  mutate(ktype = factor(ktype, levels = c("null", "disc", "funnel", 
                                          "bell", "stalk-disc", "lip", 
                                          "flag", "head", "brush"))) %>%
  filter(!ktype %in% c("null", "brush")) %>% # type 0 were not reliably quantified (and rarely visited) and type 9 was vanishingly rare
  droplevels() %>%
  left_join(bb_traits)

null_model_overall <- ggplot(null_tab1, aes(ktype, preference)) +
  #geom_violin(fill = "gray70", color = "white", draw_quantiles = c(0.25, 0.5, 0.75)) +
  geom_boxplot(fill = NA, color = "black", outlier.shape = 4, size = 0.35) +
  geom_jitter(aes(color = Test), width = 0.3, alpha = 0.6) +
  theme_light(10) +
  scale_color_manual(values = c("gray40", "#F8766D", "#00B6EB")) +
  labs(x = "Floral morphotype",
       y = "Bias")

raw_visits_overall <- ggplot(visitation_prop, aes(ktype, mean.prop.visits)) +
  geom_boxplot(fill = NA, color = "black", outlier.shape = 4, size = 0.35) +
  geom_jitter(width = 0.3, alpha = 0.5, color = "gray40") +
  theme_light(10) +
  labs(x = "Floral morphotype",
       y = "Mean proportional visitation")

ggsave("../ms_2/figures/raw_visitation_ktype.pdf", height = 6, width = 7)
ggsave("../ms_2/figures/raw_visitation_ktype.png", height = 6, width = 7)

overall_ktype <- (null_model_overall | raw_visits_overall)  +
  plot_annotation(tag_levels = 'A')

ggsave("../ms_2/figures/overall_ktype.pdf", overall_ktype, height = 4.5, width = 9.5)
ggsave("../ms_2/figures/overall_ktype.png", overall_ktype, height = 4.5, width = 9.5)
```

## 5.2 Visitation ~ tongue length
```{r}
gam1b <- gam(mean.prop.visits ~ 
              ktype + 
              s(pbl.w, by = ktype, k = 5),
             family = betar(eps = 0.001),
            data = visitation_prop) %>% getViz()

summary(gam1b)
check.gamViz(gam1b)
#plot.gamViz(gam1, allTerms = TRUE)

# Extract predictions
inv_logit <- function(x) {
  exp(x)/(1+exp(x))
}

gam1b_pred <- predict_gam(gam1b) %>%
  mutate(fit.plus.se = inv_logit(fit + (2*se.fit)),
         fit.minus.se = inv_logit(fit - (2*se.fit)),
         fit = inv_logit(fit))

raw_visits_pbl_GAM <- ggplot(gam1b_pred, aes(pbl.w, fit)) +
  geom_line() + # fit line
  geom_ribbon(aes(ymin = fit.minus.se, ymax = fit.plus.se), # Uncertainty
              fill = "black", alpha = 0.2) +
  geom_point(data = visitation_prop, # Original data
             aes(pbl.w, mean.prop.visits, color = reorder(bb.sp, pbl.w)), 
             inherit.aes = FALSE, alpha = 0.5) +
  facet_wrap(~ktype) +
  labs(x = "Tongue length (mm)", y = "Mean proportional visitation", 
       color = "Species") +
  scale_x_continuous(labels = scales::number_format(accuracy = 1)) +
  scale_y_continuous(labels = scales::number_format(accuracy = 0.01)) +
  theme_light(10)

null_model_pbl_GAM <- ggplot(gam1_pred, aes(pbl.w, fit)) +
  geom_line() + # fit line
  geom_ribbon(aes(ymin = fit.minus.se, ymax = fit.plus.se), # Uncertainty
              fill = "black", alpha = 0.2) +
  geom_point(data = null_tab1, # Original data
             aes(pbl.w, preference, color = bb.sp), 
             inherit.aes = FALSE, alpha = 0.5) +
  #scale_color_manual(values = c("gray40", "#F8766D", "#00B6EB")) +
  facet_wrap(~ktype) +
  labs(x = "Tongue length (mm)", y = "Bias", 
       color = "Species") +
  scale_x_continuous(labels = scales::number_format(accuracy = 1)) +
  scale_y_continuous(labels = scales::number_format(accuracy = 0.01)) +
  theme_light(10)

pbl_GAM_plot <- (null_model_pbl_GAM | raw_visits_pbl_GAM) + 
  plot_layout(guides = "collect") +
  plot_annotation(tag_levels = 'A')

ggsave("../ms_2/figures/ktypeXtonguelength_raw_visitation.pdf", pbl_GAM_plot, height = 5, width = 8.5)
ggsave("../ms_2/figures/ktypeXtonguelength_raw_visitation.png", pbl_GAM_plot, height = 5, width =8.5)
```

## 5.3 Ordination of raw visitation
```{r}
# Visitation matrix
uncorrected_visits <- visitation_prop %>%
  select(bb.sp, ktype, mean.prop.visits) %>%
  pivot_wider(names_from = ktype, values_from = mean.prop.visits) %>%
  column_to_rownames("bb.sp")

# Call PCA
pcoa_visits <- capscale(uncorrected_visits ~ 1 , distance = "horn")

# Extract tongue length data
tongues2 <- bb_traits %>%
  semi_join(visitation_prop) %>%
  arrange(bb.sp)

# Ordisurf on tongue length
ordisurf_visits <- ordisurf(pca_visits, tongues2$pbl.w) 
summary(ordisurf_visits)

# Get variance explained
pcoa_visits_mds1 <- round(pcoa_visits$CA$eig[[1]] / sum(pcoa_visits$CA$eig)*100)
pcoa_visits_mds2 <- round(pcoa_visits$CA$eig[[2]] / sum(pcoa_visits$CA$eig)*100)

# Fortify whole model
pcoa_visits.fort <- fortify(pcoa_visits, axes = 1:2) %>%
  # for plotting, I prefer ktype to be numerically coded
  mutate(Label = recode(Label,
                        "disc" = "1",
                        "funnel" = "2",
                        "bell" = "3",
                        "stalk-disc" = "4",
                        "lip" = "5",
                        "flag" = "6",
                        "head" = "7"))

# Fortify plants
pcoa_visits.fort1 <- fortify(pcoa_visits, axes = 1:2) %>%
  filter(Score == "species") %>%
  rename(ktype = Label) %>%
  # for plotting, I prefer ktype to be numerically coded
  mutate(ktype = recode(ktype,
                        "disc" = "1",
                        "funnel" = "2",
                        "bell" = "3",
                        "stalk-disc" = "4",
                        "lip" = "5",
                        "flag" = "6",
                        "head" = "7"))

# Fortify bees
pcoa_visits.fort2 <- fortify(pcoa_visits, axes = 1:2) %>%
  filter(Score == "sites") %>%
  rename(bb.sp = Label) %>%
  left_join(bb_traits)

# Plot ordisurf

## Plotting function based on https://oliviarata.wordpress.com/2014/07/17/ordinations-in-ggplot2-v2-ordisurf/
surfplot2 <- function(x, y, min_score = 0, pca1, pca2) {

  grid <- x$grid # extracts the ordisurf object
  grid_exp <- expand.grid(x = grid$x, y = grid$y) # get x and ys
  grid_exp$z <- as.vector(grid$z) # unravel the matrix for the z scores

  contours <- data.frame(na.omit(grid_exp)) %>% # gets rid of the nas
    as_tibble() %>%
    rename(MDS1 = x, MDS2 = y)

  scores <- y %>% 
    filter((abs(MDS1) > min_score | abs(MDS2) > min_score) | Score == "BB.sp") %>%
    mutate(max.score = max(MDS1, MDS2))

  ggplot(scores, aes(MDS1, MDS2, pch = Score)) +
    geom_contour(data = contours, 
                 aes(MDS1, MDS2, z = z, color = stat(level)),
                 size = 2, alpha = 0.4, inherit.aes = FALSE) +
    #geom_point(size = 2) +
    geom_text_repel(aes(MDS1, MDS2, label = Label), key_glyph = "rect") +
    theme_light() +
    scale_color_viridis_c(name = "Tongue length") +
    xlab(paste("MDS1 ", "(", pca1, "%", ")", sep = "")) +
    ylab(paste("MDS2 ", "(", pca2, "%", ")", sep = "")) +
    coord_fixed() # Gavin Simpson says this is **really** important
}

## Plot
raw_visitation_ord <- surfplot2(ordisurf_visits, pcoa_visits.fort, 
         pca1 = pcoa_visits_mds1, pca2 = pcoa_visits_mds2)

ordination_plot <- (bias_ord | raw_visitation_ord) +
  plot_layout(guides = "collect") +
  plot_annotation(tag_levels = 'A')

ggsave("../ms_2/figures/paired_ordination.png", ordination_plot, height = 6, width = 9.5)
ggsave("../ms_2/figures/paired_ordination.pdf", ordination_plot,  height = 6, width = 9.5)
```

## 5.4 Raw visitation ~ ktype x elevation x time
### 5.4.1 Call model
Binomial model works better than beta because of all the ones an zeros.
```{r}
visit_cover_sitedate <- visitation %>% 
  group_by(sample.id, bb.sp, ktype) %>%
  summarize(visits = n()) %>%
  group_by(sample.id, bb.sp) %>%
  mutate(prop.visits = visits/sum(visits)) %>%
  select(sample.id, bb.sp, ktype, prop.visits) %>%
  group_by(bb.sp) %>%
  nest() %>%
  mutate(fill = map(data, function(x) full_join(x, flower_cover_ktype) %>% replace_na(replace = list(prop.visits = 0)))) %>%
  unnest(fill) %>%
  select(-data) %>%
  filter(!ktype %in% c("0", "9")) %>% # type 0 were not reliably quantified (and rarely visited) and type 9 was vanishingly rare
  droplevels() %>%
  left_join(bb_traits) %>%
  separate(sample.id, c("site", "date"), sep = "_") %>%
  left_join(site_data) %>%
  unite(bb.ktype, c(bb.sp, ktype), remove = FALSE) %>%
  mutate(date = ymd(date),
         yday = yday(date),
         bb.ktype = factor(bb.ktype),
         site = factor(site),
         year = year(date))

# Binomial model
gam5 <- bam(prop.visits ~
              bb.ktype +
              s(site, by = bb.ktype, bs = "re") +
              s(year, by = bb.ktype, bs = "re") +
              te(yday, elev.mean, by = bb.ktype),
            #family = betar(eps = 0.001),
            family = "binomial",
            method = "fREML",
            discrete = TRUE,
            data = filter(visit_cover_sitedate, 
                          bb.sp %in% c("pasc", "prat", "telu", "hort", "soro", "wurf") &
                                         ktype %in% c("3", "5", "6", "7"))) %>% 
  getViz()

check.gamViz(gam5)
gam5_sum <- summary(gam5, re.test = FALSE) 
plot.gamViz(gam5)

# 
# # Beta model
# gam5b <- bam(prop.visits ~
#               bb.ktype +
#               s(site, by = bb.ktype, bs = "re") +
#               s(year, by = bb.ktype, bs = "re") +
#               te(yday, elev.mean, by = bb.ktype),
#             #family = betar(eps = 0.001),
#             family = betar(),
#             #family = "binomial",
#             method = "fREML",
#             discrete = TRUE,
#             data = filter(visit_cover_sitedate, 
#                           bb.sp %in% c("pasc", "prat", "telu", "hort", "soro", "wurf") &
#                                          ktype %in% c("bell", "lip", "flag", "head"))) %>% 
#   getViz()
# 
# check.gamViz(gam5b)
# summary(gam5b, re.test = FALSE) 
# plot.gamViz(gam5b)
```

### 5.4.2 Plot model
```{r}
# Tabulate model summary
gam5_s_table <- gam5_sum$s.table %>%
  as.data.frame() %>%
  rownames_to_column("term") %>%
  filter(str_detect(term, "te\\(")) %>%
  mutate(term = str_extract(term, ".{6}$")) %>%
  separate(term, c("bb.sp", "ktype")) %>%
  mutate(pval = signif(`p-value`, digits = 3)) %>%
  select(bb.sp, ktype, edf, pval) %>%
  as_tibble() %>%
  mutate(ktype = case_when(
    ktype == 3 ~ "bell",
    ktype == 5 ~ "lip",
    ktype == 6 ~ "flag",
    ktype == 7 ~ "head"  )) %>%
  mutate(ktype = factor(ktype, levels = c("bell", "lip", "flag", "head")))

# Extract plotting data.
hort_3_gam5 <- plot(sm(gam5, 49))$ggObj[[1]] %>%
  mutate(bb.sp = rep("hort", n()),
         ktype = rep("bell", n()))

hort_5_gam5 <- plot(sm(gam5, 50))$ggObj[[1]] %>%
  mutate(bb.sp = rep("hort", n()),
         ktype = rep("lip", n()))

hort_6_gam5 <- plot(sm(gam5, 51))$ggObj[[1]] %>%
  mutate(bb.sp = rep("hort", n()),
         ktype = rep("flag", n()))

hort_7_gam5 <- plot(sm(gam5, 52))$ggObj[[1]] %>%
  mutate(bb.sp = rep("hort", n()),
         ktype = rep("head", n()))

pasc_3_gam5 <- plot(sm(gam5, 53))$ggObj[[1]] %>%
  mutate(bb.sp = rep("pasc", n()),
         ktype = rep("bell", n()))

pasc_5_gam5 <- plot(sm(gam5, 54))$ggObj[[1]] %>%
  mutate(bb.sp = rep("pasc", n()),
         ktype = rep("lip", n()))

pasc_6_gam5 <- plot(sm(gam5, 55))$ggObj[[1]] %>%
  mutate(bb.sp = rep("pasc", n()),
         ktype = rep("flag", n()))

pasc_7_gam5 <- plot(sm(gam5, 56))$ggObj[[1]] %>%
  mutate(bb.sp = rep("pasc", n()),
         ktype = rep("head", n()))

prat_3_gam5 <- plot(sm(gam5, 57))$ggObj[[1]] %>%
  mutate(bb.sp = rep("prat", n()),
         ktype = rep("bell", n()))

prat_5_gam5 <- plot(sm(gam5, 58))$ggObj[[1]] %>%
  mutate(bb.sp = rep("prat", n()),
         ktype = rep("lip", n()))

prat_6_gam5 <- plot(sm(gam5, 59))$ggObj[[1]] %>%
  mutate(bb.sp = rep("prat", n()),
         ktype = rep("flag", n()))

prat_7_gam5 <- plot(sm(gam5, 60))$ggObj[[1]] %>%
  mutate(bb.sp = rep("prat", n()),
         ktype = rep("head", n()))

soro_3_gam5 <- plot(sm(gam5, 61))$ggObj[[1]] %>%
  mutate(bb.sp = rep("soro", n()),
         ktype = rep("bell", n()))

soro_5_gam5 <- plot(sm(gam5, 62))$ggObj[[1]] %>%
  mutate(bb.sp = rep("soro", n()),
         ktype = rep("lip", n()))

soro_6_gam5 <- plot(sm(gam5, 63))$ggObj[[1]] %>%
  mutate(bb.sp = rep("soro", n()),
         ktype = rep("flag", n()))

soro_7_gam5 <- plot(sm(gam5, 64))$ggObj[[1]] %>%
  mutate(bb.sp = rep("soro", n()),
         ktype = rep("head", n()))

telu_3_gam5 <- plot(sm(gam5, 65))$ggObj[[1]] %>%
  mutate(bb.sp = rep("telu", n()),
         ktype = rep("bell", n()))

telu_5_gam5 <- plot(sm(gam5, 66))$ggObj[[1]] %>%
  mutate(bb.sp = rep("telu", n()),
         ktype = rep("lip", n()))

telu_6_gam5 <- plot(sm(gam5, 67))$ggObj[[1]] %>%
  mutate(bb.sp = rep("telu", n()),
         ktype = rep("flag", n()))

telu_7_gam5 <- plot(sm(gam5, 68))$ggObj[[1]] %>%
  mutate(bb.sp = rep("telu", n()),
         ktype = rep("head", n()))

wurf_3_gam5 <- plot(sm(gam5, 69))$ggObj[[1]] %>%
  mutate(bb.sp = rep("wurf", n()),
         ktype = rep("bell", n()))

wurf_5_gam5 <- plot(sm(gam5, 70))$ggObj[[1]] %>%
  mutate(bb.sp = rep("wurf", n()),
         ktype = rep("lip", n()))

wurf_6_gam5 <- plot(sm(gam5, 71))$ggObj[[1]] %>%
  mutate(bb.sp = rep("wurf", n()),
         ktype = rep("flag", n()))

wurf_7_gam5 <- plot(sm(gam5, 72))$ggObj[[1]] %>%
  mutate(bb.sp = rep("wurf", n()),
         ktype = rep("head", n()))

plotting_data_gam5 <- bind_rows(hort_3_gam5, hort_5_gam5, hort_6_gam5, hort_7_gam5,
                           pasc_3_gam5, pasc_5_gam5, pasc_6_gam5, pasc_7_gam5,
                           prat_3_gam5, prat_5_gam5, prat_6_gam5, prat_7_gam5,
                           soro_3_gam5, soro_5_gam5, soro_6_gam5, soro_7_gam5,
                           telu_3_gam5, telu_5_gam5, telu_6_gam5, telu_7_gam5,
                           wurf_3_gam5, wurf_5_gam5, wurf_6_gam5, wurf_7_gam5) %>%
  group_by(bb.sp, ktype) %>%
  mutate(z.norm = scale(z, center = FALSE),
         ktype = factor(ktype, levels = c("bell", "lip", "flag", "head")),
         date = as_date(x)) %>%
  ungroup() %>%
  left_join(bb_traits) %>%
  left_join(gam5_s_table) %>%
  select(z, z.norm, date, x, elevation = y, se, 
         bb.sp, ktype, pbl.w, edf, pval) %>%
  mutate(sig = if_else(pval < 0.05, TRUE, FALSE))

ggplot(plotting_data_gam5, aes(date, elevation, z = z.norm, fill = z.norm)) +
  geom_raster() +
  geom_contour(color = "black",
               size = 0.1,
               bins = 40,
               aes(linetype = sig)) + # by fixing bins, we scale contours to each plot while letting the color ramp represent differences in scale across plots; I think this is a good compromise
  facet_grid(reorder(bb.sp, -pbl.w) ~ ktype) +
  scale_linetype_manual(values = c("dotted", "solid"))+
  scale_fill_viridis_c(na.value = "transparent") +
  labs(x = NULL, y = "Elevation (m)", fill = "Normalized\neffect", linetype = "p < 0.05") +
  theme_light(12) +
  theme(panel.grid = element_blank(),
        axis.text.x = element_text(angle = 90, vjust = 0.5))

ggsave("../ms_2/figures/raw_visitation_gam.pdf", height = 5, width = 6.5)
ggsave("../ms_2/figures/raw_visitation_gam.png", height = 5, width = 6.5)
```

## 5.5 Raw flower cover
Beta seems to work a bit better for the flower cover, probably because it does not contain many ones or any zeros.

### 5.5.1 Call model
```{r}
flower_cover_ktype.prop <- flower_cover_ktype %>%
  group_by(sample.id) %>%
  mutate(prop.cover = cover/sum(cover)) %>%
  separate(sample.id, c("site", "date"), sep = "_") %>%
  mutate(date = ymd(date),
         yday = yday(date),
         ktype = factor(ktype),
         site = factor(site),
         year = year(date)) %>%
  filter(!ktype %in% c("10", "0", "9")) %>% # type 0 were not reliably quantified (and rarely visited) and type 9 was vanishingly rare
  droplevels() %>%
  left_join(site_data)

gam6 <- bam(prop.cover ~
              ktype +
              s(site, by = ktype, bs = "re") +
              s(year, by = ktype, bs = "re") +
              te(yday, elev.mean, by = ktype),
            #family = betar(eps = 0.001),
            family = "binomial",
            method = "fREML",
            discrete = TRUE,
            data = filter(flower_cover_ktype.prop, 
                            ktype %in% c("3", "5", "6", "7"))) %>% 
  getViz()

check.gamViz(gam6)
gam6_sum <- summary(gam6, re.test = FALSE) 
plot.gamViz(gam6)

gam6b <- bam(prop.cover ~
              ktype +
              s(site, by = ktype, bs = "re") +
              s(year, by = ktype, bs = "re") +
              te(yday, elev.mean, by = ktype),
            family = betar(),
            #family = "binomial",
            method = "fREML",
            discrete = TRUE,
            data = filter(flower_cover_ktype.prop, 
                            ktype %in% c("3", "5", "6", "7"))) %>% 
  getViz()

check.gamViz(gam6b)
gam6b_sum <- summary(gam6b, re.test = FALSE) 
plot.gamViz(gam6b)
```

### 5.5.2 Plot model
```{r}
# Tabulate model summary
gam6b_s_table <- gam6b_sum$s.table %>%
  as.data.frame() %>%
  rownames_to_column("term") %>%
  filter(str_detect(term, "te\\(")) %>%
  mutate(term = str_extract(term, ".{1}$")) %>%
  rename(ktype = term) %>%
  mutate(pval = signif(`p-value`, digits = 3)) %>%
  select(ktype, edf, pval) %>%
  as_tibble() %>%
  mutate(ktype = case_when(
    ktype == 3 ~ "bell",
    ktype == 5 ~ "lip",
    ktype == 6 ~ "flag",
    ktype == 7 ~ "head"  )) %>%
  mutate(ktype = factor(ktype, levels = c("bell", "lip", "flag", "head")))

# Extract plotting data.
ktype_3_gam6b <- plot(sm(gam6b, 9))$ggObj[[1]] %>%
  mutate(ktype = rep("bell", n()))

ktype_5_gam6b <- plot(sm(gam6b, 10))$ggObj[[1]] %>%
  mutate(ktype = rep("lip", n()))

ktype_6_gam6b <- plot(sm(gam6b, 11))$ggObj[[1]] %>%
  mutate(ktype = rep("flag", n()))

ktype_7_gam6b <- plot(sm(gam6b, 12))$ggObj[[1]] %>%
  mutate(ktype = rep("head", n()))

plotting_data_gam6b <- bind_rows(ktype_3_gam6b, ktype_5_gam6b, ktype_6_gam6b, ktype_7_gam6b) %>%
  group_by(ktype) %>%
  mutate(z.norm = scale(z, center = FALSE),
         ktype = factor(ktype, levels = c("bell", "lip", "flag", "head")),
         date = as_date(x)) %>%
  ungroup() %>%
  left_join(gam6b_s_table) %>%
  select(z, z.norm, date, x, elevation = y, se, 
         ktype, edf, pval) %>%
  mutate(sig = if_else(pval < 0.05, TRUE, FALSE))

ggplot(plotting_data_gam6b, aes(date, elevation, z = z.norm, fill = z.norm)) +
  geom_raster() +
  geom_contour(color = "black",
               size = 0.1,
               bins = 40) + # by fixing bins, we scale contours to each plot while letting the color ramp represent differences in scale across plots; I think this is a good compromise
  facet_wrap(~ktype) +
  scale_fill_viridis_c(na.value = "transparent") +
  labs(x = NULL, y = "Elevation (m)", fill = "Normalized\neffect", linetype = "p < 0.05") +
  theme_light(12) +
  theme(panel.grid = element_blank(),
        axis.text.x = element_text(angle = 90, vjust = 0.5))

ggsave("../ms_2/figures/flower_cover_gam.pdf", height = 5, width = 6.5)
ggsave("../ms_2/figures/flower_cover_gam.png", height = 5, width = 6.5)
```