### 2021-06-09 ms_2/ms_2.Rmd

---
title: "Floral selectivity in mountain bumble bees and a test of the 'target use hypothesis'"

author: Douglas B. Sponsler^1^*, Katharina Kallnik^1^, Fabrice Requier^2^, Alice Claßen^1^, A. Fabienne Maihoff^1^, Johanna Sieger^3^, Ingolf Steffan-Dewenter^1^

# date: `r Sys.Date()`

output:
  bookdown::word_document2:
    fig_caption: yes
    fig_width: 7
    fig_height: 7
    df_print: kable
    keep_md: yes
    number_sections: no
    toc: no
    reference_docx: Word_template.docx

bibliography: 
  - references.bib
#  - knitcitations.bib
csl: ecology.csl
# csl: https://www.zotero.org/styles/ecology

---

^1^ Department of Animal Ecology and Tropical Biology, Biocenter, University of Würzburg, Würzburg, Germany

^2^ Université Paris-Saclay, CNRS, IRD, UMR Évolution, Génomes, Comportement et Écologie, 91198, Gif-sur-Yvette, France

^3^ Naturpark Frankenhöhe, Ansbach, Germany

`*` Corresponding author: douglas.sponsler@uni-wuerzburg.de

# ABSTRACT

Alternative title: Mountain bumble bees exhibit strong but flexible foraging biases through temporal and elevational gradients of floristic turnover

The process of resource selection by foraging animals has been a central theme of ecological research since the mid 20th century, particularly in the context of resource partitioning among sympatric species. Bumble bees as model system... Field studies have shown that floral choice in bumble bees is associated with the morphological fittedness between bumble bee tongue length to floral corolla depth. Morphological fittedness, however, is only a soft constraint on resource selection; bumble bees also tune their floral choices to learned patterns of reward, and they exhibit remarkable intelligence in learning the handling skills required to exploit difficult flower morphologies.

Using a dataset consisting of more than 13,000 observations collected along a 1400 m elevation gradient over three consecutive years, we describe how bumble bee floral selection is constrained by morphology yet flexibly tuned to local conditions in elevation and time, with considerable idiosyncrasy in floral selection across species with similar tongue length.  More results...

Significance ...

*Keywords*: foraging, trait matching, pollinator, *Bombus*, alpine

```{r setup, include=FALSE, cache=FALSE, message = FALSE}

library("knitr")

### Chunk options: see http://yihui.name/knitr/options/ ###
## Text results
opts_chunk$set(echo = FALSE, warning = TRUE, message = TRUE)

## Code decoration
opts_chunk$set(tidy = TRUE, comment = NA, highlight = TRUE)

## Cache
# opts_chunk$set(cache = 2, cache.path = "ms_2/figures/cache/")

## Plots
opts_chunk$set(fig.path = "../ms_2/figures/")

```

```{r knitcitations, cache = FALSE}
library(knitcitations)
cleanbib()   
cite_options(citation_format = "pandoc")
```

# 1. INTRODUCTION

The process of resource selection by animals has been an enduring theme of ecological research [@Johnson1980-co; @Beyer2010-vr]. Methodological difficulties notwithstanding, the study of animals' resource choices is one of the central tasks of community ecology, where the selection and partitioning of resources are instrumental in community assembly, species coexistence, and ecosystem function [@Chesson2000-aw]. Recently, questions of resource selection in ecological communities have taken on new urgency on account of anthropogenic climate and land use change, which threaten to disrupt existing patterns of species interactions and niche relationships [@Tylianakis2008-fi].

Bumble bees (Hymentoptera: *Bombus* spp.) and their floral hosts provide a diverse yet tractable system in which to explore resource selection. Bumble bees are exceptionally versatile foragers whose ability to solve the puzzles posed by complex flower morphologies [@Plowright1984-lc] --- or, indeed, by human investigators [@Alem2016-rn; @Loukola2017-bt] --- renders almost any flower potentially accessible. Nevertheless, the flexibility of bumble bee foraging is tuned to floral reward, which depends on both the intrinsic reward of a given flower and the costs of exploiting it. These costs are closely related to the degree of morphological fittedness between bumble bees and their floral hosts, which consists principally in the matching of bumble bee tongue length to floral corolla depth. Bumble bee species can differ markedly in tongue length, and floral handling time increases when tongue length and corolla depth are poorly matched [@Inouye1980-mv; @Harder1983-bk], leading to a well-documented association between tongue length and floral choice in bumble bee communities [e.g. @Brian1954-cy; @Heinrich1976-wc; @Teras1976-pi; @Inouye1978-qi; @Pyke1982-gn]. This association between tongue length and floral niche is imperfect, though, and there can be considerable overlap in floral resource use between bumble bees of different tongue lengths [e.g. @Ranta1980-ok; @Goulson2008-oe]. Moreover, bumble bee floral choice is also influenced by ecological modulators, such as the spatial distribution of floral patches or competition with other bumble bees [@Ishii2013-sm; @Plowright1984-lc; @Sowig1989-ld].

Understanding resource selection in bumble bee communities is more than just an exercise in ecological theory. Bumble bees are sensitive to climate and land use change [@Goulson2008-oe; @Soroye2020-zb], especially in the mountain ecosystems that host the world's richest bumble bee communities. Warming climates have already caused measurable upslope shifts in mountain bumble bees and their floral hosts [@Ploquin2013-da; @Kerr2015-dg; @Pyke2016-ll; @Fourcade2019-ct; @Soroye2020-zb; @Marshall2020-pc], which could introduce novel competitive interactions between bumble bee species that experience increased sympatry. At the same time, socioeconomic factors have led to a dramatic decline in traditional alpine grazing management, particularly in the European Alps, resulting in widespread pasture abandonment and reforestation of mountain meadows [@Tappeiner2006-jp; @Tasser2007-qo]. The regression of meadows to forest entails a massive loss of floral resource availability, and even early stages of forest succession can cause declines in floral abundance and diversity, with corresponding effects on bumble bees [@Iserbyt2015-jc; @Walcher2017-kw; @Suzuki-Ohno2020-jv]. In contrast, more accessible areas at the bases of mountains are undergoing agricultural intensification that displaces natural vegetation and favors grasses over forbs [@Tasser2002-om]. Caught between adverse land use changes above and below and subject to climate-induced range shifts, mountain bumble bees are facing a future of reduced floral resource availability and altered sympatry, and the ability of bumble species to coexist under these changes will likely depend significantly on how they select and partition floral resources.

In this study, we analyze a data set consisting of ~13,000 interactions between bumble bees and wildflowers recorded over 3 years along a 1400 m elevation gradient. Using these data in conjunction with complementary floral survey data, we explore the *selectivity* of the bumble bee community at the level of floral morphotype [@Kugler1970-gb].

**This is a good place to get our terminological ducks in a row**

We opt for the term *selectivity*, defined as the  (*sensu* @Beyer2010-vr) rather than the ... defined in this way, selectivity has the same conceptual meaning (though different mathematical derivation) as @Bluthgen2006-if *discrimination*.

We first compare the aggregate preferences of the bumble bee community across floral morphotypes and then model the preferences for each morphotype as a function of bumble bee tongue length. Then, by calculating preferences based on individual samples distributed through elevation and time, we model the variation of bumble bee preferences through temporal and elevational axes of floristic turnover, revealing the lability of bumble bee floral preferences. The balance of flexibility and constraint revealed in these analyses **Figure out how to finish this...**

Our investigation also affords an opportunity to revisit an open question regarding the role of resource abundance as a driver of flower selection in bumble bees. In studies of resource selection, it is common to pose an abundance-only hypothesis in which foraging is random and selection merely a function of frequency of encounter, resulting in resource use patterns that mirror resource abundance patterns. For bumble bees, which are known to forage non-randomly, the abundance-only hypothesis can serve as a null-model against which observed use patterns can be compared to infer *selectivity*, i.e. resource use that is either positively or negatively disproportionate to resource abundance. It has also be proposed, however --- in what has been termed the *target use hypothesis* --- that bumble bees might select resources irrespective of abundance [@De_Valpine2013-be; @Harmon-Threatt2017-bz]. This hypothesis, while obviously subject to boundary conditions of floral availability, is consistent with the observation that bumble bees sometimes persist in visiting favored flower species even when these resources are experimentally reduced to low abundance [@Heinrich1976-vb], and it was supported in a study comparing competing models of flower selection in *Bombus vosnesenskii* [@Harmon-Threatt2017-bz]. If and when the *target use hypothesis* holds, uncorrected visitation rates are a better indication of preference than the residuals of an abundance-only null model, since the subtraction of presumed neutral effects of abundance when those effects do not, in fact, exist actually creates new abundance-dependent artifacts and obscures the true process of resource selection [@Harmon-Threatt2017-bz]. On the other hand, failing to account for neutral effects of abundance when they do obtain similarly obscures the selection process underpinning foraging patterns [@Vazquez2003-qa; @Bluthgen2008-rp]. In our study, we take advantage of the fact that there exists are well-established mechanistic basis for bumble bee flower choice, namely the compatibility between floral morphology and bumble bee tongue length. Taking this mechanism as a reference, we evaluate models of preference with and without abundance effects by comparing the explanatory power of bumble bee tongue length with respect to the preferences inferred under each model.

*"[. . .] preference  is  the  ratio  of  the  use  of  a  habitat  over  its availability, conditional on the availability of all habitats to the study animals." [@Aarts2008-fo]*

One way we could handle our language and caveats is this. The kind of null model approach we used *does not* condition visitation to each focal resource on the abundances of all other resources, only on the proportional abundance of the focal resource. Following @Aarts2008-fo, we cannot then speak of *preference*. Instead, what we have is weaker concept of *selectivity* defined by Beyer2010-vr. We can, therefore, use the word *selectivity* after explaining its relation to preference in the strict sense. I do not want to go back and fit some crazy hierarchical model that conditions on the abundance of all morphotypes.

# 2. METHODS

## 2.1 Field sampling

Field sampling was conducted 2010-2012 in Berchtesgaden National Park (47.55°N, 12.92°E), located in the Northern Limestone Alps of southeast Germany. The landscape is composed of mountain pastures in a coniferous forest matrix. We selected 25 study sites (60 x 60 m) on pastures at elevations ranging from 641-2032 m above sea level. Fourteen of these pastures are extensively grazed by cattle or sheep, three are mowed for hay production, and eight have been abandoned throughout the last century and are no longer subject to any human management.

Sampling consisted of repeated visits (hereafter, “visitation samples”) to each study site at approximately weekly intervals. Sampling was only conducted during periods without rain and with an air temperature of at least 6°C. Bumble bees were recorded during one 50-minute transect walk per visit, and bumblebees observed on a given flower were counted as floral visitors. Queens were identified to species level in the field to avoid destructively sampling colony foundresses, but workers and males were collected and frozen for later identification in the laboratory. Species were identified following @Amiet1996-jz. While male bumble bees was recorded, we chose to analyze only the pooled visitation of by queens and workers. In conjunction with bumble bee observations, we estimated the flower cover of each herbaceous or shrubby plant species within each 60 x 60 m study plot to the nearest 0.01 m^2^ and identified the plant species according to @Lauber2007-ha and @Oberndorfer2001-ro.

While per-sample effort was standardized across all years, the total number of samples varied: 153 in 2010, 212 in 2011, and 539 in 2012. The start of field sampling each year was determined by snowmelt, so lower sites were sampled as early as April while higher sites could not be sampled until June or July. In 2010 and 2012, this resulted in a temporally staggered pattern of sampling, but a combination of early snowmelt and a late start to field activities in 2011 resulted in sampling that was approximately aligned in time across the whole elevation gradient. Due to these annual differences in sampling intensity and duration, we opted to pool data across years rather than include year effects in our models, which would multiply model complexity and be difficult to interpret. This means, however, that our analysis cannot detect year-to-year variation in floral preferences and niche overlap, and main effects cannot be distinguished from annual anomalies. With only three years of data, though, such annual inferences would be of little value, and we deemed them not worth the cost of added model complexity.

The samples analyzed in this paper have been previously analyzed (Sponsler et al., submitted), but the analyses presented here are non-overlapping.

## 2.2 External data sources

### 2.2.1 Floral morphology

Rather than considering various floral morphological traits independently, we relied on the synthesis of @Kugler1970-gb, who developed a typology consisting of ten primary morphotypes and nested subdivision of each. Kugler’s typology was informed by laboratory studies of bumble bee flower choice [@Kugler1943-uy], and there is a precedent for analyzing bumble bee visitation in terms of Kugler’s morphotypes [@Farnberger1996-xo; @Neumayer1999-il; @Schneller2014-ry].

Kugler morphotype classifications of the plants in our study system were accessed via the BIOFLOR database [@Klotz2002-uc]. Morphotypes were simplified to the primary classes numbered 0-10: 0 = null (includes wind-pollinated flora such as grasses), 1 = disc- and bowl-shaped flowers, 2 = funnel flowers, 3 = bell-shaped flowers, 4 = stalk-disc flowers, 5 = lip flowers, 6 = flag flowers, 7 = flower heads, 8 = spike flowers, 9 = brush flowers, 10 = trap flowers **(Figure 1)**. In a few cases, a plant species in our data set was not found in the BIOFLOR database. In most cases, morphotypes are constant at the genus level, and missing morphotypes were added based on the morphotypes found for congeners. In the case of the genus *Phyteuma*, we followed @Neumayer1999-il and classified it as a group 7 flower head. Morphotypes 9 and 10 were extremely rare in our study system and morphotype 8 was completely absent, so these morphotypes were omitted from all analyses. Plants belonging to morphotype 0 were typically not recorded during floral surveys due to their lack of conspicuous floral displays, and they were rarely visited by bumble bees, so this morphotype was also omitted from all analyses.

### 2.2.2 Bumble bee tongue length

Bumble bee tongue-length data (for the worker caste) were obtained from the literature values compiled and summarized by @Arbetman2017-xj. Values for *B. gerstaeckeri* and *B. mendax*, which were not included in @Arbetman2017-xj, were obtained from @Obeso1992-yp and @Durieux2000-uk, respectively.

## 2.3 Data processing

Prior to analysis, we aligned each visitation sample with a corresponding floral survey sample. In most cases, floral surveying was conducted jointly with bumble bee observation for each site and date. Occasionally, though, floral surveying could not be completed on the same day as bumble bee observation, resulting in a small temporal offset between visitation samples and floral survey samples. In these cases, visitation samples were paired with the nearest available floral survey data within a maximum distance of 7 days. Visitation samples lacking corresponding floral survey data within 7 days were omitted from further analysis, which resulted in dropping 40 site-dates out of a total of 895, leaving 855 site-dates included in our analyses. In a few cases, two consecutive visitation samples at a given site shared the same nearest floral survey sample; in these cases, the visitation samples were first pooled, then aligned to their shared nearest floral survey sample. In rare cases, a given floral species was recorded during visitation sampling but missed during floral surveying, and these species were retroactively assigned the minimum floral area value of 0.01 m^2^.

Having thus aligned visitation data and flower cover data, we summed visitation and floral cover across species within each floral morphotype. This yielded, for each of our 855 aligned samples, a cumulative value of flower cover for each floral morphotype and a cumulative value of visitation for each bumble- bee-morphotype pair.

## 2.4 Analysis of resource selection

**Start this by working out my terminology. Are we saying preference or bias? And what about the distinction between preference, use, and selection? See @Johnson1980-co and @Beyer2010-vr as authorities on terminology. Whatever I choose, I then need to go revise the labels of the figures accordingly. I think we could make an argument for sticking wit the word "bias" to describe the inferences of our null model analysis, since we are allowing that the raw visitation patterns may be as good or better a measure of "preference". We don't actually find that, of course, but I think we should avoid calling the null model inferences "preference" as though it were a foregone conclusion that they represent preference while the uncorrected visitation rates do not. But preference, alas, is the word the litertaure seems to have settled on, so when I discuss my conclusions, I should talk about preference, comparing preference as inferred from bias and preference as inferred from uncorrected visitation rates.**

*We will use the word "preference" in the phenomenological sense of @Beyer2010-vr to denote a disproportionality between resource availability and resource use while remaining agnostic as to the behavioral basis of this pattern. Or perhaps we will use the word bias but follow the same definition, since "bias" does not bear the teleological connotations of "preference". Double check that this is really how @Beyer2010-vr define preference. If so, it differs a bit from the more original definition of .*

A key challenge when interpreting interaction networks is to disentangle directed processes, such as a focal species' preference or avoidance of certain interaction partners, from the neutral effects of relative abundance [@Vazquez2003-qa; @Bluthgen2008-rp]. For example, in the context of a bumble bee floral visitation network such as ours, abundant floral taxa should be expected to receive more visitation than rare taxa simply because they are more frequently encountered, and it would clearly be a mistake to interpret this as meaningful "preference" or "specialization" on the part of the bumble bees. One way to control for the neutral effects of relative abundance is to compare observed visitation patterns to the simulated visitation patterns of a null model in which plants are visited randomly in proportion to their relative abundance [@Vazquez2003-qa; @Bluthgen2008-rp]. When constructing null models of interaction networks that are distributed through space and/or time, though, it is critical to respect the sampling unit of individual sites and dates, since only within this site-date sampling unit does floral visitation data correspond unequivocally to floral abundance data. Naive aggregation across sites and dates introduces forbidden links and decouples visitation and abundance [@Vaughan2018-gj].

We constructed null models following @Vaughan2018-gj. First, resource selection is simulated for individual consumers --- in our case, individual bumble bees --- by according to a resource selection process in which the probability of a resource being chosen is proportional to its relative abundance [@Agusti2003-tt]. By simulating the visitation of individual consumers rather than species in aggregate, the abundance of each resource, and therefore its probability of visitation, can vary across site-dates, thus preserving the within-sample correspondence between visitation and abundance. Simulated interactions are then summarized for each consumer species by taking the mean number of interactions with each resource across all individuals of the consumer species and all samples of the dataset. As this process is iterated, the result is a distribution of visitation rates of each consumer species to each resource under the null assumption that resource selection is determined only by resource abundance. The visitation distributions of the null model can then be compared to observed visitation rates to infer positive or negative bias of each consumer toward each resource. Since the mean of the null distribution can be interpreted as the expected value under null assumptions, the difference between the mean null value and the observed value can be interpreted as a measure of positive or negative bias [@Vaughan2018-gj]. The statistical significance of this bias can be estimated as the proportion of simulated null values more extreme than the observed value.

We performed two rounds of null model analysis targeting different ecological questions. First, we followed the procedure described above, simulating our entire dataset in one modeling process to infer aggregate pattern of bias between bumble bee species and floral morphotypes. Then, we repeated the null model analysis for each site-date individually so that the bias between each bumble bee species and floral morphotype could be modeled as a variable through elevation and time. For both analyses, we used 500 iterations to achieve a robust sampling distribution, and we considered observed visitation rates to indicate significant bias if they fell outside the interval containing 99% of simulated values.

Using the aggregate null model results, we modeled the biases toward each floral morphotype as a function of bumble bee tongue length using hierarchical generalized additive modeling (HGAM) with floral morphotype as a "by" smoothing factor [@Pedersen2019-kf]. We then compared morphotype biases across bumble bee species by performing a principal components analysis and mapped bumble bee tongue length onto the ordination space by fitting a 2-dimensional GAM surface to the first two principal components.

With the sample-wise null model results, we modeled the variation of bias between bumble bee species and floral morphotypes through elevation and time. For this analysis, we included only the six most abundant and widespread bumble bee species in our study system (*B. hortorum*, *B. pascuorum*, *B. pratorum*, *B. soroeensis*, *B. terrestris-lucorum*, and *B. wurflenii*) to ensure sufficient data points through elevation and time, and we modeled their biases only toward the 4 most visited floral morphotypes (bell, lip, flag, and head flowers). Biases for each bumble- bee-morphotype pair were modeled together in a single HGAM with a tensor-product interaction smooth between elevation and time and bumble- bee-morphotype pair as a "by" smoothing factor.

## 2.5 Software

All data processing and analyses were conducted in `R` [@RCoreTeam-2021aa]. Data handling and visualization were performed with the `tidyverse` suite [@Wickham2019-aa]. Null models were constructed using `econullnetr` [@Vaughan2018-gj]. GAM analyses were performed with `mgcv` [@Wood_2017] and visualized with `mgcViz` [@Fasiolo2018aa] and `tidymv` [@Coretta2021-aa]. PCA was performed with the `vegan` [@Oksanen_2019] and visualized using `ggplot2` [@Wickham2016-aa] and `ggvegan` [@Simpson2019aa]. Annotated R code is available in the Appendix X.

# 3. RESULTS

## 3.1 Sampling overview

```{r fig1, echo = FALSE,  fig.cap = paste("Visited families within each floral morphotype. Barplots show the total number of visits to each family across all bumble bee species and samples (note the rescaling of the y-axis across panels). Each floral morphotype is exemplified by a representative species common in our study system: 0 = Plantago lanceolata, 1 = Potentilla erecta, 2 = Gentiana asclepiadea, 3 = Campanula glomerata, 4 = Silene nutans, 5 = Ajuga reptans, 6 = Lotus corniculatus, 7 = Centaurea jacea, 8 = Salix retusa. Illustrations are modifications of public domain works obtained from www.plantillustrations.org. For details of original sources, see Appendix X.")}
knitr::include_graphics("../ms_2/figures/morphotype_families_mod.png")
```

## 3.2 Selection bias by species

```{r fig2, echo = FALSE,  fig.cap = paste("Empirical visitation rates compared to null model simulations for each bumble bee species and floral morphotype. Gray bars depict 99\\% intervals of null model simulations and points depict observed visitation rates. Points are color-coded according to where they fall with respect to the 99\\% interval of the null model: red = above (positive bias), blue = below (negative bias), gray = within (nonsignificant bias).")}
knitr::include_graphics("../ms_2/figures/null_model_spwise_preference.png")
```

## 3.3 Overall selection bias by morphotype

```{r fig3, echo = FALSE,  fig.cap = paste("Overall selection bias (A) and uncorrected visitation rate (B) of the bumble bee community toward each floral morphotype. Each point represents the bias/visitation of one bumble bee species toward the corresponding floral morphotype. The color of each point in panel A, retained from Figure 2, indicates whether a given observed bias is significantly (p < 0.01) positive or negative with respect to null model simulations.")}
knitr::include_graphics("../ms_2/figures/overall_ktype.png")
```

## 3.4 Tongue length as a predictor of biases

```{r fig4, echo = FALSE,  fig.cap = paste("Selection biases (A) and uncorrected visitation rates (B) for each morphotype regressed against bumble bee tongue length. Each point represents the bias/visitation of one bumble bee species toward the corresponding floral morphotype. Curves depict fitted GAM smooths with 95\\% confidence bands. Points are color-coded by species, and species are presented in the legend in order of increasing tongue length.")}
knitr::include_graphics("../ms_2/figures/ktypeXtonguelength_preference.png")
```

```{r fig5, echo = FALSE,  fig.cap = paste("Ordination biplots of bumble bee species according to their selection biases (A) and uncorrected visitation rates (B) for each floral morphotype. Floral morphotypes are abbreviated with their numerical codes and illustrated as in Figure 1. Colored bands depict a 2-dimensional GAM smooth representing the relationship between the first two principal components and bumble bee tongue length. Morphotype illustrations are as described in Figure 1.")}
knitr::include_graphics("../ms_2/figures/paired_ordination_mod.png")
```

## 3.5 Lability of bias through elevation and time

```{r fig6, echo = FALSE,  fig.cap = paste("Variation in bias through elevation and time modeled as HGAM surfaces. Color ramps depict z-normalized effects to allow the range of effect size to be compared across panels, but contour lines are set at the same fixed density across all panels to emphasize the within-panel patterns. Significant (p < 0.05) smooths are depicted with solid contour lines and nonsignificant smooths with dashed contour lines.")}
knitr::include_graphics("../ms_2/figures/preference_lability_gam.png")
```
```{r fig7, echo = FALSE,  fig.cap = paste("Variation in uncorrected visitation rates through elevation and time modeled as HGAM surfaces. Color ramps depict z-normalized effects to allow the range of effect size to be compared across panels, but contour lines are set at the same fixed density across all panels to emphasize the within-panel patterns. Significant (p < 0.05) smooths are depicted with solid contour lines and nonsignificant smooths with dashed contour lines.")}
knitr::include_graphics("../ms_2/figures/raw_visitation_gam.png")
```

```{r fig8, echo = FALSE,  fig.cap = paste("Variation in flower cover through elevation and time modeled as HGAM surfaces. Color ramps depict z-normalized effects to allow the range of effect size to be compared across panels, but contour lines are set at the same fixed density across all panels to emphasize the within-panel patterns. Significant (p < 0.05) smooths are depicted with solid contour lines and nonsignificant smooths with dashed contour lines.")}
knitr::include_graphics("../ms_2/figures/flower_cover_gam.png")
```

# 4. DISCUSSION

So, we would conclude that (A) abundance matters but (B) preference matters, too. We find support for a middle ground between the abundance-only and preference-only models. Need to acknowledge the potential for nonlinear "functional responses" of use to abundance [@Mysterud1998-rs].

## 4. Caveats:

Inference of preference is sensitive to the scale at which availability is estimate [@Beyer2010-vr]. My floral survey data estimates availability more or less at the meadow scale, right? But I don't know what is available outside the meadow. This problem is also discussed in [@Harmon-Threatt2017-bz], specifically in the context of bees and flowers.

The sensitivity of "preference" to availability [@Johnson1980-co; @Beyer2010-vr].

*First, the estimates ofpreference  are  not  directly  interpretable  as  reflectingthe  intrinsic  behavioural  motivations  of  the  animal.Positive  (or  negative)  preference  for  a  given  habitatimplies only that the habitat is used more (or less) fre-quently than expected by chance,conditional upon thedefined sample of availability.  Even  high  levels  of  useof a habitat can be associated with negative coefficientsif that habitat is common, and, conversely, low levelsof  use  can  be  associated  with  positive  coefficientsif the habitat is rare. The sign and magnitude of thecoefficient   cannot   be   interpreted   with   respect   tothe   behavioural   decision   the   animal   is   making.* [@Beyer2010-vr].

*The second consequence is that estimates of prefer-ence  based  on  different  samples  of  availability  (e.g.Design III inManlyet al.2002) are not directly com-parable.  The  sign  and  magnitude  of  the  estimate  ofpreference  is  a  function  of  both  the  sample  of  useand the sample of availability. This creates an identifia-bility  problem  because  preference  can  change  as  afunction of changing use and/or changing availability.If   the   sample   of   availability   differs   between   twoestimates of preference, it is meaningless to draw infer-ences based on the difference in preference. Differentpreferences can arise from the same selection process,and   similar   preferences   can   arise   from   differentselection processes.* [@Beyer2010-vr].

This is exactly the problem I am facing. I think, though, I can present my "preference lability analysis" together with the raw visitation and availability data, discuss this limitation explicitly, and let the reader decide what to make of it. This is also discussed very clearly, and in the context of bumblebees, in @Harmon-Threatt2017-bz.

*Given that preference is conditional on avail-ability,  that  availability  is  likely  to vary  in  space  andthat preference is likely to change as a function of avail-ability, there is little theoretical basis for believing thatpreferences estimated in one area will be good predic-tors  of  preference  in  others.* [@Beyer2010-vr].

In answer to this, I can point to the largely agreeing results of other bumble bee studies.

*Models of habitat preference provide a phenomenolo-gical  description  of  habitat  use,  conditional  on  thesample of availability. Although some of the simplify-ing  assumptions  upon  which  these  models  are  often based  may  be  biologically  unrealistic,  they  can  stillhave predictive power (Boyceet al.2002;Wienset al.2008). The degree, however, to which simple modelsof habitat preference help us to develop a mechanisticunderstanding of animal movement and distribution islimited. Mechanistic models are required to advanceour  biological  understanding  of  the  behavioural  pro-cesses  that  result  in  animal  movement,  distributionand habitat use.*

Tongue length is my mechanistic model. And I think the observation that the mechanistic model better explains use-availability bias than raw selection indicates that the use-availability approach better captures the way bumble bees make foraging decisions.

Need to acknowledge the potential for nonlinear "functional responses" of use to abundance [@Mysterud1998-rs].

*The problem of conflating preference and niche partitioning:*

Our study is set in the Berchtesgadener Alps of southeast Germany, a landscape of flower-rich mountain meadows embedded in a matrix of coniferous forest. This system is one of abundant floral resources but harsh abiotic conditions, so it is likely that competition for floral resources is low, or at least transient, and that bumble bee populations are limited instead by the shortness of the growing season, as has been suggested in the similar environment of the Eyne Valley of France [@Iserbyt2008-jn]. Yet while *realized* competition between bumble bees in our system may be low, the question of *potential* competition is highly relevant because mountain ecosystems are currently subject to at least two powerful drivers of directed change.

Resource usage is typically studied as a phenomenological level, with questions of mechanism relegated to speculations regarding the latent process of resource selection that lurks behind realized use patterns [@Beyer2010-vr]. In our study, we evaluate the target use hypothesis by reference to a known mechanism of floral selection, namely bumble bee tongue length.

# 5. CONCLUSIONS

# ACKNOWLEDGMENTS

We thank D.J. McNeil for helpful conversations during data analysis. O. Mitesser kindly provided a high-performance computing environment to expedite null modeling.

# REFERENCES

<!-- ```{r write_citations, cache=FALSE, include=FALSE} -->
<!-- write.bibtex(file = "knitcitations.bib") -->
<!-- ``` -->

<div id ="refs"></div>