### 2021-06-11 ms_2/ms_2.Rmd

---
title: "Floral selectivity in mountain bumble bees and a test of the 'target use hypothesis'"

author: Douglas B. Sponsler^1^*, Katharina Kallnik^1^, Fabrice Requier^2^, Alice Claßen^1^, A. Fabienne Maihoff^1^, Johanna Sieger^3^, Ingolf Steffan-Dewenter^1^

# date: `r Sys.Date()`

output:
  bookdown::word_document2:
    fig_caption: yes
    fig_width: 7
    fig_height: 7
    df_print: kable
    keep_md: yes
    number_sections: no
    toc: no
    reference_docx: Word_template.docx

bibliography: 
  - references.bib
#  - knitcitations.bib
csl: ecology.csl
# csl: https://www.zotero.org/styles/ecology

---

^1^ Department of Animal Ecology and Tropical Biology, Biocenter, University of Würzburg, Würzburg, Germany

^2^ Université Paris-Saclay, CNRS, IRD, UMR Évolution, Génomes, Comportement et Écologie, 91198, Gif-sur-Yvette, France

^3^ Naturpark Frankenhöhe, Ansbach, Germany

`*` Corresponding author: douglas.sponsler@uni-wuerzburg.de

# ABSTRACT

Understanding patterns of resource use by animals, and the underlying selection processes that govern these patterns, can clarify how ecological communities have assembled, how they currently function, and how they are likely to respond to future perturbations. Communities of bumble bees and their floral hosts provide a tractable and compelling system in which to study questions of resource use, and the answers to these questions have immediate relevance to the conservation of bumble bees and their mutualists under the forcings of climate and land use change, particularly in mountain ecosystems.

Using a dataset consisting of ~13,000 observations collected along a 1400 m elevation gradient over three consecutive years, we

Significance ...

*Keywords*: foraging, trait matching, pollinator, *Bombus*, alpine

```{r setup, include=FALSE, cache=FALSE, message = FALSE}

library("knitr")

### Chunk options: see http://yihui.name/knitr/options/ ###
## Text results
opts_chunk$set(echo = FALSE, warning = TRUE, message = TRUE)

## Code decoration
opts_chunk$set(tidy = TRUE, comment = NA, highlight = TRUE)

## Cache
# opts_chunk$set(cache = 2, cache.path = "ms_2/figures/cache/")

## Plots
opts_chunk$set(fig.path = "../ms_2/figures/")

```

```{r knitcitations, cache = FALSE}
library(knitcitations)
cleanbib()   
cite_options(citation_format = "pandoc")
```

# 1. INTRODUCTION

The process of resource selection by animals has been an enduring theme of ecological research [@Johnson1980-co; @Beyer2010-vr]. Methodological difficulties notwithstanding, the study of animals' resource choices is one of the central tasks of community ecology, where the selection and partitioning of resources are instrumental in community assembly, species coexistence, and ecosystem function [@Chesson2000-aw; @Bascompte2007-ls]. Recently, questions of resource selection in ecological communities have taken on new urgency on account of anthropogenic climate and land use change, which threaten to disrupt existing patterns of species interactions and niche relationships [@Tylianakis2008-fi].

Bumble bees (Hymentoptera: *Bombus* spp.) and their floral hosts provide a diverse yet tractable system in which to explore resource selection. Bumble bees are exceptionally versatile florivores whose ability to solve the puzzles posed by complex flower morphologies [@Plowright1984-lc] --- or, indeed, by human investigators [@Alem2016-rn; @Loukola2017-bt] --- renders almost any flower potentially accessible. Nevertheless, the flexibility of bumble bee foraging is tuned to floral reward, which depends on both the intrinsic reward of a given flower and the costs of exploiting it. These costs are closely related to the degree of morphological fittedness between bumble bees and their floral hosts, which consists principally in the matching of bumble bee tongue length to floral corolla depth. Bumble bee species can differ markedly in tongue length, and floral handling time increases when tongue length and corolla depth are poorly matched [@Inouye1980-mv; @Harder1983-bk], leading to a well-documented association between tongue length and floral choice in bumble bee communities [e.g. @Brian1954-cy; @Heinrich1976-wc; @Teras1976-pi; @Inouye1978-qi; @Pyke1982-gn]. This association between tongue length and floral niche is imperfect, though, and there can be both convergence in floral resource use between bumble bees of different tongue length and divergence in floral resource use between bumble bees of similar tongue length [e.g. @Ranta1980-ok; @Goulson2008-oe]. Moreover, bumble bee floral choice is also influenced by ecological modulators, such as the spatial distribution of floral patches or competition with other bumble bees [@Ishii2013-sm; @Plowright1984-lc; @Sowig1989-ld].

Floral choice in bumble bee communities can serve as a case study in the general phenomenon of animal resource selection, but it is also a matter of practical importance. Bumble bees are sensitive to climate and land use change [@Goulson2008-oe; @Soroye2020-zb], especially in the mountain ecosystems that host the world's richest bumble bee communities. Warming climates have already caused measurable upslope shifts in mountain bumble bees and their floral hosts [@Ploquin2013-da; @Kerr2015-dg; @Pyke2016-ll; @Fourcade2019-ct; @Soroye2020-zb; @Marshall2020-pc], which could introduce novel competitive interactions between bumble bee species that experience increased sympatry. At the same time, socioeconomic factors have led to a dramatic decline in traditional alpine grazing management, particularly in the European Alps, resulting in widespread pasture abandonment and reforestation of mountain meadows [@Tappeiner2006-jp; @Tasser2007-qo]. The regression of meadows to forest entails a massive loss of floral resource availability, and even early stages of forest succession can cause declines in floral abundance and diversity, with corresponding effects on bumble bees [@Iserbyt2015-jc; @Walcher2017-kw; @Suzuki-Ohno2020-jv]. In contrast, more accessible areas at the bases of mountains are undergoing agricultural intensification that displaces natural vegetation and favors grasses over forbs [@Tasser2002-om]. Caught between adverse land use changes above and below and subject to climate-induced range shifts, mountain bumble bees are facing a future of reduced floral resource availability and altered sympatry. The ability of bumble bee species to coexist under these changes will likely depend significantly on how they select and partition floral resources, and analogous considerations apply to their floral mutualists, many of which depend principally on bumble bee visitation for pollination service [@Goulson2010-lo].

In this study, we analyze a data set consisting of ~13,000 interactions between bumble bees and wildflowers recorded over 3 years along a 1400 m elevation gradient. Using these data in conjunction with complementary floral survey data, we explore the *selectivity* of the bumble bee community at the level of floral morphotype [@Kugler1970-gb]. Following @Beyer2010-vr, we define *selectivity* simply as any disproportionality --- positive or negative --- between the relative abundance of a resource and its rate of use. This concept approaches the idea of resource *preference*, but inference of the latter, in an absolute sense, requires either experimental control of all resource abundances [@Johnson1980-co] or statistical techniques that condition visitation simultaneously on all resource abundances [@Aarts2008-fo]. Given the scale and complexity of our observational study, we opt for simpler analytical approaches and restrict our inference to the phenomenon of selectivity, which can be understood as a metric of *relative* preference for a focal resource given unmodeled interactions with preferences for alternative resources. We deem these interactions beyond the scope of our study, but acknowledge that this requires us to remain partly agnostic as to the process underlying the empirical selectivity we describe.

The first objective of our study is to describe and explain the selectivity of the bumble bee community with respect to floral morphotype. Using null model analysis to remove neutral effects of resource abundance, we compare the selectivity of the bumble bee community across floral morphotypes and then model the species-wise selectivity for each morphotype as a function of bumble bee tongue length.

We then revisit an open question regarding the role of resource abundance as a driver of flower selection in bumble bees. In studies of resource selection, it is common to pose an abundance-only hypothesis in which foraging is random and selection merely a function of frequency of encounter, resulting in resource use patterns that mirror resource abundance patterns. For bumble bees, which are known to forage non-randomly, the abundance-only hypothesis can serve as a null-model against which observed use patterns can be compared to infer selectivity (often called "preference", following a *sensu lato* usage of the term) [@Vaughan2018-gj]. It has also be proposed, however --- in what has been termed the *target use hypothesis* --- that bumble bees might select resources irrespective of abundance [@De_Valpine2013-be; @Harmon-Threatt2017-bz]. This hypothesis, while obviously subject to boundary conditions of floral availability, is consistent with the anecdotal observation that bumble bees sometimes persist in visiting favored flower species even when these resources are experimentally reduced to low abundance [@Heinrich1976-vb], and it was supported in a study comparing competing models of flower selection in *Bombus vosnesenskii* [@Harmon-Threatt2017-bz]. If and when the *target use hypothesis* holds, raw visitation rates (i.e. raw for abundance) are a better index of preference (*sensu stricto*) than selectivity (i.e. the residuals of an abundance-only null model), since the subtraction of presumed neutral effects of abundance when those effects do not, in fact, exist actually creates new abundance-dependent artifacts and obscures the true process of resource selection [@Harmon-Threatt2017-bz]. On the other hand, failing to account for neutral effects of abundance when they do obtain similarly obscures the selection process underpinning foraging patterns [@Vazquez2003-qa; @Bluthgen2008-rp]. In our study, we take advantage of the fact that there exists a well-established mechanistic basis for bumble bee flower choice, namely the compatibility between floral morphology and bumble bee tongue length. Taking this mechanism as a reference and exploiting the tongue-length diversity of the bumble bee species in our study system, we evaluate models of preference with and without abundance effects by comparing the explanatory power of bumble bee tongue length with respect to the preferences inferred under each model.

Finally, we model the variation of selectivity, raw visitation, and floral abundance through time and elevation. This analysis reveals the elevational and phenological dynamics of the underlying floral community as well as the lability of bumble bee foraging behavior. Exploring these patterns is the first step toward understanding how climate and land use change might distort the temporal and elevational structure of bumble- bee-wildflower mutualisms in mountain ecosystems.

# 2. METHODS

## 2.1 Field sampling

The samples analyzed in this paper have been previously analyzed (Sponsler et al., submitted), but the analyses presented here are non-overlapping. Field sampling was conducted 2010-2012 in Berchtesgaden National Park (47.55°N, 12.92°E), located in the Northern Limestone Alps of southeast Germany. The landscape is composed of a coniferous forest matrix embedded with pastures either currently or historically maintained by extensive (i.e. low intensity) agricultural management. We selected a total of 25 study sites (60 x 60 m) located on pastures along a ~1400 m elevation gradient ranging from 641-2032 m above sea level (m.s.l.). Fourteen of these pastures are grazed by cattle or sheep, three are mowed for hay production, and eight have been abandoned throughout the last century and are no longer subject to any human management.

Sampling of bumble- bee-wildflower interactions consisted of repeated visits (hereafter, “visitation samples”) to each study site at approximately weekly intervals. Sampling was only conducted during periods without rain and with an air temperature of at least 6°C. Bumble bees were recorded during one 50-minute transect walk per visit, and bumblebees observed on a given flower were counted as floral visitors. Queens were identified to species level in the field to avoid destructively sampling colony foundresses, but workers and males were collected and identified in the lab following @Amiet1996-jz. While male bumble bees were recorded, we chose to analyze only the pooled visitation of queens and workers. In conjunction with bumble bee observations, we estimated the flower cover of each herbaceous or shrubby plant species within each 60 x 60 m study plot to the nearest 0.01 m^2^ (hereafter "floral survey samples") and identified the plant species according to @Lauber2007-ha and @Oberndorfer2001-ro. While per-sample effort was standardized across all years, the total number of samples varied: 153 in 2010, 212 in 2011, and 539 in 2012. The start of field sampling each year was determined by snowmelt, so lower sites were sampled as early as April while higher sites could not be sampled until June or July. In 2010 and 2012, this resulted in a temporally staggered pattern of sampling, but a combination of early snowmelt and a late start to field activities in 2011 resulted in sampling that was approximately aligned in time across the whole elevation gradient. See Appendix X for a full description of sampling patterns.

## 2.2 External data sources

### 2.2.1 Floral morphology

Rather than considering various floral morphological traits independently, we relied on the synthesis of @Kugler1970-gb, who developed a typology consisting of ten primary morphotypes and nested subdivision of each. Kugler’s typology was informed by laboratory studies of bumble bee flower choice [@Kugler1943-uy], and there is a precedent for analyzing bumble bee visitation in terms of Kugler’s morphotypes [@Farnberger1996-xo; @Neumayer1999-il; @Schneller2014-ry].

Kugler morphotype classifications of the plants in our study system were accessed via the BIOFLOR database [@Klotz2002-uc]. Morphotypes were simplified to the primary classes numbered 0-10: 0 = null (includes wind-pollinated flora such as grasses), 1 = disc- and bowl-shaped flowers, 2 = funnel flowers, 3 = bell-shaped flowers, 4 = stalk-disc flowers, 5 = lip flowers, 6 = flag flowers, 7 = flower heads, 8 = spike flowers, 9 = brush flowers, 10 = trap flowers **(Figure 1)**. In a few cases, a plant species in our data set was not found in the BIOFLOR database. In most cases, morphotypes are constant at the genus level, and missing morphotypes were added based on the morphotypes found for congeners. In the case of the genus *Phyteuma*, we followed @Neumayer1999-il and classified it as a group 7 flower head. Morphotypes 9 and 10 were extremely rare in our study system and morphotype 8 was completely absent, so these morphotypes were omitted from all analyses. Plants belonging to morphotype 0 were typically not recorded during floral surveys due to their lack of conspicuous floral displays, and they were rarely visited by bumble bees, so this morphotype was also omitted from all analyses.

### 2.2.2 Bumble bee tongue length

Bumble bee tongue-length data (for the worker caste) were obtained from the literature values compiled and summarized by @Arbetman2017-xj. Values for *B. gerstaeckeri* and *B. mendax*, which were not included in @Arbetman2017-xj, were obtained from @Obeso1992-yp and @Durieux2000-uk, respectively.

## 2.3 Data processing

Prior to analysis, we aligned each visitation sample with a corresponding floral survey sample. In most cases, floral surveying was conducted jointly with bumble bee observation for each site and date. Occasionally, though, floral surveying could not be completed on the same day as bumble bee observation, resulting in a small temporal offset between visitation samples and floral survey samples. In these cases, visitation samples were paired with the nearest available floral survey data within a maximum distance of 7 days. Visitation samples lacking corresponding floral survey data within 7 days were omitted from further analysis, which resulted in dropping 40 site-dates out of a total of 895, leaving 855 site-dates included in our analyses. In a few cases, two consecutive visitation samples at a given site shared the same nearest floral survey sample; in these cases, the visitation samples were first pooled, then aligned to their shared nearest floral survey sample. In rare cases, a given floral species was recorded during visitation sampling but missed during floral surveying, and these species were retroactively assigned the minimum floral area value of 0.01 m^2^.

Having thus aligned visitation data and flower cover data, we summed visitation and floral cover across species within each floral morphotype. This yielded, for each of our 855 aligned samples, a cumulative value of flower cover for each floral morphotype and a cumulative value of visitation for each bumble- bee-morphotype pair.

## 2.4 Analysis of resource selection

To quantify selectivity, we compared the proportional visitation of each morphotype (by each bumble bee species) with the rate of visitation expected under the null-hypothesis that visitation of a given resource is purely a function of its relative abundance [@Vazquez2003-qa; @Bluthgen2008-rp]. When conducting a null model analysis of interaction networks that are distributed through space and/or time, it is critical to respect the sampling unit of individual sites and dates, since only within this site-date sampling unit does floral visitation data correspond unequivocally to floral abundance data. Naive aggregation across sites and dates introduces forbidden links and decouples visitation and abundance [@Vaughan2018-gj]. We constructed null models following @Vaughan2018-gj. First, resource selection is simulated for individual consumers --- in our case, individual bumble bees --- according to a resource selection process in which the probability of a resource being chosen is proportional to its relative abundance [@Agusti2003-tt]. By simulating the visitation of individual consumers rather than species in aggregate, the abundance of each resource, and therefore its probability of visitation, can vary across site-dates, thus preserving the within-sample correspondence between visitation and abundance. Simulated interactions are then summarized for each consumer species by taking the mean number of interactions with each resource across all individuals of the consumer species and all samples of the dataset. As this process is iterated, the result is a distribution of visitation rates of each consumer species to each resource under the null assumption that resource selection is determined only by resource abundance. The visitation distributions of the null model can then be compared to observed visitation rates to infer positive or negative selectivity of each consumer toward each resource. Since the mean of the null distribution can be interpreted as the expected value under null assumptions, the difference between the mean null value and the observed value can be interpreted as a magnitude of selectivity [@Vaughan2018-gj]. The statistical significance of this selectivity can be estimated as the proportion of simulated null values more extreme than the observed value.

We followed the procedure described above, simulating our entire dataset in one modeling process to infer aggregate pattern of selectivity between bumble bee species and floral morphotypes. We used 500 iterations to achieve a robust sampling distribution, and we considered observed visitation rates to indicate significant selectivity if they fell outside the interval containing 99% of simulated values. Then, using the results of this analysis, we modeled selectivity toward each floral morphotype as a function of bumble bee tongue length using hierarchical generalized additive modeling (HGAM), with floral morphotype specified as a smoothing factor [@Pedersen2019-kf]. Finally, we compared morphotype selectivity across bumble bee species by performing a principal components analysis on the bumble bee species x floral morphotype selectivity matrix, and we mapped bumble bee tongue length onto the ordination space by fitting a 2-dimensional GAM surface to the first two principal components.

As a test of the *target use hypothesis*, we repeated the HGAM and ordination analyses described above using raw mean visitation rates rather than selectivity. For the ordination of raw mean visitation rates, we used principal coordinates analysis based on the Horn-Morisita distance rather than euclidean principal components analysis because the former is better suited for proportional composition data [@Borcard2018-gb].

To explore the temporal and elevational patterns in our data, we repeated the null model analysis described above but this time for each site-date individually, so that the selectivity of each bumble bee species toward each floral morphotype could be modeled as a variable through elevation and time. For this analysis, we included only the six most abundant and widespread bumble bee species in our study system (*B. hortorum*, *B. pascuorum*, *B. pratorum*, *B. soroeensis*, *B. terrestris-lucorum*, and *B. wurflenii*) to ensure sufficient data points through elevation and time, and we modeled their selectivity only toward the 4 most visited floral morphotypes (bell, lip, flag, and head flowers). Selectivity for each bumble- bee-morphotype pair (hereafter "bb-fl pair") was modeled together in a single HGAM with a tensor-product interaction smooth between elevation and time, bb-fl pair as a smoothing factor and fixed intercept effect, and the interactions site * bb-fl pair and year * bb.fl pair as random intercept effects.

An inherent problem with modeling sample-wise variation in selectivity is that the bounds of possible selectivity are strictly dependent on floral abundance [@@Beyer2010-vr]. A morphotype present at very low relative abundance would, under the null model, be expected to receive very low visitation, which bounds potential negative selectivity near zero but allows positive selectivity to approach 1. Conversely, a morphotype present at very high relative abundance would receive very high visitation under the null model, and its potential positive selectivity would be bounded near zero and its potential negative selectivity near -1. A morphotype present at exactly 50% relative abundance would have selectivity values bounded at exactly (-0.5, 0.5). This clearly creates a degree of unavoidable confoundedness between selectivity and floral abundance. With this limitation in mind, we fit HGAMs of the same form to raw visitation rates and total flower cover of each floral morphotype, and we present the results together to be interpreted jointly.

## 2.5 Software

All data processing and analyses were conducted in `R` [@RCoreTeam-2021aa]. Data handling and visualization were performed with the `tidyverse` suite [@Wickham2019-aa]. Null models were constructed using `econullnetr` [@Vaughan2018-gj]. GAM analyses were performed with `mgcv` [@Wood_2017] and visualized with `mgcViz` [@Fasiolo2018aa] and `tidymv` [@Coretta2021-aa]. PCA was performed with the `vegan` [@Oksanen_2019] and visualized using `ggplot2` [@Wickham2016-aa] and `ggvegan` [@Simpson2019aa]. Annotated R code is available in the Appendix X.

# 3. RESULTS

## 3.1 Sampling overview

A total of 12,918 bumble- bee-flower interactions were included in our final data set, including 16 bumble bee species (with *Bombus terrestric/lucorum* and *Bombus psithyrus* species lumped into species groups), 163 plant species (110 genera, 37 families), and 736 unique bumble- bee-plant interaction pairs. *B. pascuorum* ("pasc"), *B. pratorum* ("prat"), *B. soroeensis* ("soro"), *B. terrestris-lucorum* ("telu"), and *B. wurflenii* ("wurf") were the most abundant species, and they occurred at all sites. *B. hortorum* ("hort"), *B. jonellus* ("jone"),  *B. monticola* ("mont"), and *B. psithyrus* ("psit") occurred at moderate abundance. The remaining species --- *B. gerstaeckeri* ("gers"), *B. hypnorum* ("hypn"), *B. lapidarius* ("lapi"), *B. mendax* ("mend"), *B. mucidus* ("muci"), and *B. pyrenaeus* ("pyre") --- were relatively rare. *B. humilis* ("humi") was recorded only three times and only in one year, so we omitted it from all analyses.

Floral surveying yielded a total of 352 plant species (191 genera, 52 families), of which 155 (103 genera, 35 families) were observed to be visited by bumble bees. Visited floral species included all morphotypes except spike (8) and trap (10) flowers **(Figure 1)**. Null (0) and brush (9) morphotypes were, however, excluded from analysis for the reasons given above.

<!-- Check again and make sure all the numbers given here are still correct. -->

```{r fig1, echo = FALSE,  fig.cap = paste("Visited floral morphotypes and their constituent plant families. Note that some families span more than on morphotype. Barplots show the total number of visits to each family across all bumble bee species and samples (note the rescaling of the y-axis across panels). Each floral morphotype is illustrated with a representative species common in our study system: 0 = Plantago lanceolata, 1 = Potentilla erecta, 2 = Gentiana asclepiadea, 3 = Campanula glomerata, 4 = Silene nutans, 5 = Ajuga reptans, 6 = Lotus corniculatus, 7 = Centaurea jacea, 8 = Salix retusa. Illustrations are modifications of public domain works obtained from www.plantillustrations.org. For details of original sources, see Appendix X.")}
knitr::include_graphics("../ms_2/figures/morphotype_families_mod.png")
```

## 3.2 Aggregate selectivity and raw visitation

Selectivity was generally strong, with the visitation rates of most bumblebee species to most floral morphotypes falling significantly above or below the 99% interval of the null model **(Figure 2)**. Nonsignificant selectivity was observed mainly among the rarer bumblebee species because their small samples sizes lead to wide null model confidence intervals.

```{r fig2, echo = FALSE,  fig.cap = paste("Empirical visitation rates compared to null model simulations for each bumble bee species and floral morphotype. Gray bars depict 99\\% intervals of null model simulations and points depict observed visitation rates. Points are color-coded according to where they fall with respect to the 99\\% interval of the null model: red = above (positive selectivity), blue = below (negative selectivity), gray = within (nonsignificant selectivity). The difference between the observed value and the middle of the null model interval can be interpreted as magnitude of selectivity.")}
knitr::include_graphics("../ms_2/figures/null_model_spwise_preference.png")
```

When floral morphotypes are compared to one another in terms of selectivity, clear patterns emerge **(Figure 3A)**. All species exhibited negative selectivity toward disc flowers, while most species exhibited positive selectivity toward bell, lip, and flag flowers. Selectivity toward funnel and stalk-disc flowers tended to be neutral (with one strong positive outlier --- *B. gerstaeckeri* --- for funnel flowers), reflecting the fact that these morphotypes tended both to be rarely visited and to occur at low abundance. Selectivity toward head flowers varied widely from strongly positive to strongly negative.

Differences between floral morphotypes in terms of raw mean visitation rates, reflecting the target use hypothesis, were less pronounced **(Figure 3B)**. Bell, lip, flag, and head flowers were visited at high rates by one or more bumble bee species, but median raw mean visitation rates did not differ strongly across morphotypes.

<!-- I could do some sort of linear mixed effects model on this I suppose, but I think it would violate assumptions due to the non-independence of selectivity (selectivity for one thing implies selectivity against others). I've already sort of done stats on this with the null model. -->

<!-- Before we finalize this, check to make sure the raw visitation rates were really prepared correctly and are not being deflated by a bunch of zeros that should'nt be there. -->

```{r fig3, echo = FALSE,  fig.cap = paste("Selectivity (A) and raw visitation rate (B) of the bumble bee species toward each floral morphotype. Each point represents the selectivity/visitation of one bumble bee species toward the corresponding floral morphotype. The color of each point in panel A, retained from Figure 2, indicates whether a given observed bias is significantly (p < 0.01) positive or negative with respect to null model simulations.")}
knitr::include_graphics("../ms_2/figures/overall_ktype.png")
```

## 3.4 Tongue length as a predictor of floral choice

Tongue length was a strong predictor of selectivity but not of raw mean visitation rates **(Figure 4)**. Selectivity toward funnel, bell, lip, flag, and head flowers exhibited significant GAM curves in response bumble bee tongue length, and the model explained 77% of total deviance (adj. R^2^ = 0.728). For funnel and lip flowers, the response of selectivity to tongue length was positive and approximately linear. Selectivity toward bell flowers and head flowers exhibited approximately quadratic responses with selectivity peaking among short-tongued species and declining sharply with increasing tongue length before leveling off. Selectivity toward flag flowers exhibited a unimodal peak among medium-tongued species. In contrast, raw mean visitation rates were significantly predicted by tongue length only for head flowers, which exhibited a weakly quadratic decline in selectivity from a peak among short-tongued species. The raw mean visitation rate model explained only 25% of total deviance (adj. R^2^ = 0.053).

```{r fig4, echo = FALSE,  fig.cap = paste("Selectivity (A) and raw visitation rates (B) for each morphotype regressed against bumble bee tongue length. Each point represents the selectivity/visitation of one bumble bee species toward the corresponding floral morphotype. Curves depict fitted GAM smooths with 95\\% confidence bands. Solid lines indicate curves differing significantly (p < 0.05) from a flat line while dashed lines indicated nonsigificant curves. Points are color-coded by species, and species are presented in the legend in order of increasing tongue length.")}
knitr::include_graphics("../ms_2/figures/ktypeXtonguelength_paired.png")
```

Interestingly, selectivity and raw mean visitation rates produced broadly similar ordinations, each of which captured more than 80% of total variance **(Figure 5)**. In both ordinations, GAM smooths on first two axes also explained most of the variance in tongue length (selectivity mode: deviance explained = 94.7% (adj. R^2^ = 0.893), p < 0.001; raw visitation model: deviance explained = 90.8% (adj. R^2^ = 0.859), p < 0.001). We stress, however, that the two ordinations are not directly comparable in statistical terms since they involve different underlying ordination techniques, necessitated by the fact that selectivity is bounded on the interval of (-1, 1) while mean visitation rate is bounded on the interval (0, 1). We present the ordinations primarily for the sake of visualizing the similarity between bumble bee species in terms of selectivity and raw mean visitation rates toward the floral morphotypes considered.

```{r fig5, echo = FALSE,  fig.cap = paste("Ordination biplots of bumble bee species according to their selectivity (A) and raw visitation rates (B) for each floral morphotype. Floral morphotypes are abbreviated with their numerical codes and illustrated as in Figure 1. Colored bands depict a 2-dimensional GAM smooth representing the relationship between the first two principal components and bumble bee tongue length. Morphotype illustrations are as described in Figure 1.")}
knitr::include_graphics("../ms_2/figures/paired_ordination_mod.png")
```

## 3.5 Variation of floral abundance, raw visitation, and selectivity through elevation and time

The proportional abundance of the most visited floral morphotypes --- bell, lip, flag, and head flowers --- was strongly patterned in time and elevation **(Figure 6)**. Bell flowers peaked in proportional abundance early in the season and at mid-elevation, and lip flowers peaked late in the season and at low elevation. Flag flowers peaked in mid-season and exhibited two distinct elevational peaks, the lower at around 1000 m.s.l and the higher at around 1800 m.s.l. Head flowers exhibited a mid-season peak at low elevation and a broad late season peak at around 1600 m.s.l. All smooths were significant (p << 0.0001), and the model explained 64.3% of total deviance (adj. R^2^ = 0.554).

```{r fig6, echo = FALSE,  fig.cap = paste("Variation in proportional flower cover through elevation and modeled as HGAM surfaces. Color ramps depict z-normalized effects to allow the range of effect size to be compared across panels, but contour lines are set at the same fixed density across all panels to emphasize the within-panel patterns. Significant (p < 0.05) smooths are depicted with solid contour lines and nonsignificant smooths with dashed contour lines.")}
knitr::include_graphics("../ms_2/figures/flower_cover_gam.png")
```

Patterns of raw visitation **(Figure 7)** generally mirrored patterns of proportional abundance **(Figure 6)** for each morphotype. Given the expected dependency between floral abundance and raw visitation, this is not surprising. There were, however, some notable deviations. For example, *B. soroeensis* exhibited a strong peak of visitation to bell flowers late in the year at the highest sites, outside the peak abundance of bell flowers. *B. pratorum* was unique in not exhibiting high visitation rate to lip flowers during their peak abundance late in the season and at low-elevation. Bumble bee species differed in the degree to which they exploited the mid- and high-elevation peaks of flag flower abundance; *B. hortorum* and *B. wurflenii* visited flag flowers heavily at high elevation, while *B. pascuorum* --- which also visited flag flowers heavily --- tended rather to exploit the low-elevation peak of abundance. The remaining species --- which, based on selectivity analysis, appear to have a low affinity for flag flowers **(Figure 2)** --- exhibited diffuse or nonsignficiant patterns of visitation rates that appear uncorrelated with the abundance of flag flowers.

Smooths were significant (p < 0.05) for all bb-fl pairs except hort-bell, hort-head, pasc-bell, telu-head, and soro-flag. The model explained 33.3% of total deviance (adj. R^2^ = 0.272).

```{r fig7, echo = FALSE,  fig.cap = paste("Variation in raw visitation rates through elevation and time modeled as HGAM surfaces. Color ramps depict z-normalized effects to allow the range of effect size to be compared across panels, but contour lines are set at the same fixed density across all panels to emphasize the within-panel patterns. Significant (p < 0.05) smooths are depicted with solid contour lines and nonsignificant smooths with dashed contour lines.")}
knitr::include_graphics("../ms_2/figures/raw_visitation_gam.png")
```

Patterns of selectivity through elevation and time were more complex than those of raw visitation and, by design, more often differed from patterns of floral abundance **(Figure 8)**. In some cases, though, patterns of selectivity mirrored patterns of floral availability, suggesting that bumble bees may exhibit functional responses [@Mysterud1998-rs] to floral abundance not entirely accounted for by null models that assume a linear proportionality between abundance and visitation.

One of the more salient patterns is that positive selection for flag flowers was strongly associated with high elevation, except among the species that tended to avoid flag flowers (*B. pratorum* and *B. soroeensis*) **(Figure 2)**. It is also apparent that, as a rule, selectivity is more dynamic where it is also higher overall. Bell flowers, which were positively selected overall by the short-tongued bumble bee species *B. pratorum*, "*B. terrestris-lucorum*, and *B. soroeensis* **(Figure 2)**, also exhibited the most complex patterns of selectivity in elevation and time for these bumble bee species. In contrast, the long and medium-tongued species *B. hortorum*, *B. pascuorum*, and *B. wurflenii*, which exhibited neutral or weakly positive overall selection toward bell flowers **(Figure 2)**, also show weak patterns of variation in selectivity toward bell flowers. Nearly the reverse pattern is evident for flag flowers, reflecting their association with longer-tongued bumble bee species. This relationship between the complexity of selectivity dynamics and overall selectivity can be formalized by plotting the effective degrees of freedom against the selectivity HGAM surface on the corresponding overall selectivity between each **(Figure 9)**.

Smooths were significant (p < 0.05) for all bb-fl pairs except hort-head, prat-flag, telu-head, and soro-flag. The model explained 36.4% of total deviance (adj. R^2^ = 0.33).

```{r fig8, echo = FALSE,  fig.cap = paste("Variation in selectivity through elevation and time modeled as HGAM surfaces. Color ramps depict z-normalized effects to allow the range of effect size to be compared across panels, but contour lines are set at the same fixed density across all panels to emphasize the within-panel patterns. Significant (p < 0.05) smooths are depicted with solid contour lines and nonsignificant smooths with dashed contour lines.")}
knitr::include_graphics("../ms_2/figures/preference_lability_gam.png")
```

```{r fig9, echo = FALSE,  fig.cap = paste("Complexity of selectivity dynamics through elevation and time as a function of overall selectivity. Linear best-fit lines are shown in black with gray bands indicating 95% confidence intervals.")}
knitr::include_graphics("../ms_2/figures/edf_selectivity.png")
```

# 4. DISCUSSION

Understanding patterns of resource use by animals, and the underlying selection processes that govern these patterns, can clarify how ecological communities have assembled, how they currently function, and how they are likely to respond to future perturbations. Communities of bumble bees and their floral hosts provide a tractable and compelling system in which to study questions of resource use, and the answers to these questions have immediate relevance to the conservation of bumble bees and their mutualists under the forcings of climate and land use change, particularly in mountain ecosystems.

## 4.1 Floral abundance, floral selection, and the *target use hypothesis*

It is well-known that bumble bees forage non-randomly accordingly to morphological constraints, innate preferences, and experienced rewards [e.g. Kugler1943-uy; @Brian1954-cy; @Inouye1978-qi; @Raine2007-nq; @Maharaj2019-wm]. In this regard, the results of our null model analysis are as unsurprising as they are unambiguous; bumble bee resource selection almost never represents flora in unaltered proportion to their abundance. The *target use hypothesis* [@De_Valpine2013-be; @Harmon-Threatt2017-bz], however, extends this idea to the perfect antithesis of the random foraging model, positing that bumble bee foraging is a process of *pure* preference, completely irrespective of resource abundance.

Given the well-established mechanistic link between bumble bee tongue length and floral selection [e.g. @Brian1954-cy; @Heinrich1976-wc; @Teras1976-pi; @Inouye1978-qi; @Pyke1982-gn], our finding that bumble bee tongue length explained selectivity (i.e. null model residuals) much better than it explained raw visitation rates indicates that, in our study system, abundance and preference function jointly in bumble bee floral choice. This contradicts the *target use hypothesis* as a general rule of bumble bee foraging. It does not, however, mean that the *target use hypothesis* does not obtain under certain circumstances, such as when a highly preferred resource varies within a nonlimiting rage of availability. Moreover, our results suggest something akin to the *target use hypothesis* but formulated negatively with respect to resources that are *not selected*. Visitation rates to disc flowers in our system were very low despite the fact that they often occurred at high abundance, suggesting that while preference seems to interact with resource abundance, avoidance may not. In general, though, our findings support the hypothesis that bumble bee resource selection involves both the neutral effects of abundance and the directed effect of preference, which implies that analytical techniques designed to parse these effects, such as the null model analysis we employ, should be favored in studies of bumble bee foraging. Accordingly, we shall focus primarily on our calculations of selectivity, rather than raw visitation rates, as we interpret the other results of our study.

## Patterns of selectivity

We found that positive selection was focused on 3 floral morphotypes: bell, lip, and flag flowers. These results are broadly consistent with the aggregate visitation patterns previously reported by @Schneller2014-ry in the central Alps at Vorarlberg and Neumayer1999-il in the Austrian Alps ~55 km south of our study system, suggesting that our findings are at least regionally generalizeable. When considering these patterns from the perspective of pollination services to mountain flora, it is important to note that while bumble bees are the dominant Hymenopteran pollinators in our study system, flies (especially Syrphid hoverflies) are also abundant flower visitors in mountain ecosystems, and their visitation is mainly restricted to the open flower morphotypes (e.g. disc flowers) that receive less attention from bumble bees [@Neumayer1999-il]. In the context of our research question, we have emphasized the differences in tongue length between bumble bee species, but all bumble bees have relatively long tongues compared to the overall pollinator community, and it is not surprising that their foraging tends to be focused on the deeper flower morphotypes. The process of floral resource partitioning *among* bumble bee species is nested within the broader process of floral partitioning *between* bumble bees and other flower visitors.

The strong explanatory power of tongue length with respect to selectivity in our study corroborates the established link between tongue length and floral choice. *I think it does so more strongly than most other field studyies I've read.* It is important, though, to consider the species that appear as outliers with respect to the smooth dependency of selectivity on tongue length. Outliers worth talking about: telu-disc(+), muci-lip(-), muci-flag(+), soro/psit-head(+).

## Similarity of selectivity across species, potential for competition

While competition as a phenomenon arises from realized patterns of resource use, not the latent process of selection, insight into the process of selection may be a better predictor of competition. Use patterns alone are, in a sense, overfit to the observations at hand.

## Variability of floral abundance, visitation, and selectivity through elevation and time; its relevance to climate and land use change

Elevation-time patterns of visitation closely mirrored those of floral abundance, but the former were more diffuse; this may reflect the reluctance of bumble bees to switch from familiar floral taxa even when they become rarer [@Heinrich1976-vb] or less rewarding [I think I remember reading this somewhere...but there is also a paper showing that bumble bees switch more readily than honey bees].

The late high elevation peak of visitation and selectivity between soro and bell flowers is probably driven by soro's peculiar affinity for Campanula. Talk about idiosyncracies not capture dby morphology alone. importance of other traits, particularly color.

## Caveats:

The problem of queens vs. workers... Probably the best way to fend off this object would be to re-run all our analyses with only workers and see if any thing changes qualitatively. This could all go in the SI.

Need to acknowledge the potential for nonlinear "functional responses" of use to abundance [@Mysterud1998-rs].

Inference of preference is sensitive to the scale at which availability is estimate [@Beyer2010-vr]. My floral survey data estimates availability more or less at the meadow scale, right? But I don't know what is available outside the meadow. This problem is also discussed in [@Harmon-Threatt2017-bz], specifically in the context of bees and flowers.

The sensitivity of "preference" to availability [@Johnson1980-co; @Beyer2010-vr].

*First, the estimates ofpreference  are  not  directly  interpretable  as  reflectingthe  intrinsic  behavioural  motivations  of  the  animal.Positive  (or  negative)  preference  for  a  given  habitatimplies only that the habitat is used more (or less) fre-quently than expected by chance,conditional upon thedefined sample of availability.  Even  high  levels  of  useof a habitat can be associated with negative coefficientsif that habitat is common, and, conversely, low levelsof  use  can  be  associated  with  positive  coefficientsif the habitat is rare. The sign and magnitude of thecoefficient   cannot   be   interpreted   with   respect   tothe   behavioural   decision   the   animal   is   making.* [@Beyer2010-vr].

*The second consequence is that estimates of prefer-ence  based  on  different  samples  of  availability  (e.g.Design III inManlyet al.2002) are not directly com-parable.  The  sign  and  magnitude  of  the  estimate  ofpreference  is  a  function  of  both  the  sample  of  useand the sample of availability. This creates an identifia-bility  problem  because  preference  can  change  as  afunction of changing use and/or changing availability.If   the   sample   of   availability   differs   between   twoestimates of preference, it is meaningless to draw infer-ences based on the difference in preference. Differentpreferences can arise from the same selection process,and   similar   preferences   can   arise   from   differentselection processes.* [@Beyer2010-vr].

This is exactly the problem I am facing. I think, though, I can present my "preference lability analysis" together with the raw visitation and availability data, discuss this limitation explicitly, and let the reader decide what to make of it. This is also discussed very clearly, and in the context of bumblebees, in @Harmon-Threatt2017-bz.

*Given that preference is conditional on avail-ability,  that  availability  is  likely  to vary  in  space  andthat preference is likely to change as a function of avail-ability, there is little theoretical basis for believing thatpreferences estimated in one area will be good predic-tors  of  preference  in  others.* [@Beyer2010-vr].

In answer to this, I can point to the largely agreeing results of other bumble bee studies.

*Models of habitat preference provide a phenomenolo-gical  description  of  habitat  use,  conditional  on  thesample of availability. Although some of the simplify-ing  assumptions  upon  which  these  models  are  often based  may  be  biologically  unrealistic,  they  can  stillhave predictive power (Boyceet al.2002;Wienset al.2008). The degree, however, to which simple modelsof habitat preference help us to develop a mechanisticunderstanding of animal movement and distribution islimited. Mechanistic models are required to advanceour  biological  understanding  of  the  behavioural  pro-cesses  that  result  in  animal  movement,  distributionand habitat use.*

Tongue length is my mechanistic model. And I think the observation that the mechanistic model better explains use-availability bias than raw selection indicates that the use-availability approach better captures the way bumble bees make foraging decisions.

Need to acknowledge the potential for nonlinear "functional responses" of use to abundance [@Mysterud1998-rs].

*The problem of conflating preference and niche partitioning:*

Our study is set in the Berchtesgadener Alps of southeast Germany, a landscape of flower-rich mountain meadows embedded in a matrix of coniferous forest. This system is one of abundant floral resources but harsh abiotic conditions, so it is likely that competition for floral resources is low, or at least transient, and that bumble bee populations are limited instead by the shortness of the growing season, as has been suggested in the similar environment of the Eyne Valley of France [@Iserbyt2008-jn]. Yet while *realized* competition between bumble bees in our system may be low, the question of *potential* competition is highly relevant because mountain ecosystems are currently subject to at least two powerful drivers of directed change.

Resource usage is typically studied as a phenomenological level, with questions of mechanism relegated to speculations regarding the latent process of resource selection that lurks behind realized use patterns [@Beyer2010-vr]. In our study, we evaluate the target use hypothesis by reference to a known mechanism of floral selection, namely bumble bee tongue length.

# 5. CONCLUSIONS

## What we did and why

In this study, we investigated patterns of floral resource selection in a mountain bumble bee community with the goals of (1) identifying the most selected floral morphotypes, (2) evaluating the role of floral abundance in driving floral selection, (3)

# ACKNOWLEDGMENTS

We thank D.J. McNeil for helpful conversations during data analysis. O. Mitesser kindly provided a high-performance computing environment to expedite null modeling.

# REFERENCES

<!-- ```{r write_citations, cache=FALSE, include=FALSE} -->
<!-- write.bibtex(file = "knitcitations.bib") -->
<!-- ``` -->

<div id ="refs"></div>