### 2021-07-15 ms_2/ms_2_supp_methods.Rmd

---
title: "Floral preferences of mountain bumble bees are driven by trait matching but labile through elevation and time"

author: Douglas B. Sponsler^1^*, Katharina Kallnik^1^, Fabrice Requier^2^, Alice Classen^1^, A. Fabienne Maihoff^1^, Johanna Sieger^3^, Ingolf Steffan-Dewenter^1^

# date: `r Sys.Date()`

output:
  bookdown::word_document2:
    fig_caption: yes
    fig_width: 7
    fig_height: 7
    df_print: kable
    keep_md: yes
    number_sections: no
    toc: no
    reference_docx: Word_template.docx

bibliography: 
  - references.bib
#  - knitcitations.bib
csl: ecology.csl
# csl: https://www.zotero.org/styles/ecology

---

^1^ Department of Animal Ecology and Tropical Biology, Biocenter, University of Würzburg, Würzburg, Germany

^2^ Université Paris-Saclay, CNRS, IRD, UMR Évolution, Génomes, Comportement et Écologie, 91198, Gif-sur-Yvette, France

^3^ Naturpark Frankenhöhe, Ansbach, Germany

`*` Corresponding author: douglas.sponsler@uni-wuerzburg.de

# ABSTRACT

Patterns of resource use by animals can clarify how ecological communities have assembled in the past, how they currently function, and how they are likely to respond to future perturbations. Using ~13,000 records of bumble- bee-wildflower interactions along a 1400 m elevation gradient, we find that the selection of floral morphotypes by bumble bees is driven both by the directed process of trait-matching and the neutral effects of floral abundance. Patterns of floral preference are, moreover, labile through time and elevation, which may be an adaptation to the extreme spatiotemporal floristic turnover of mountain ecosystems. The flexibility of bumble bee foraging will be tested in coming years as climate and land use change drive rapid changes in floral resource distribution in time and space, particularly in the mountain ecosystems that host the world's richest communities of bumble bees and their floral mutualists.

*Keywords*: foraging, trait matching, pollinator, *Bombus*, alpine, niche, resource selection

```{r setup, include=FALSE, cache=FALSE, message = FALSE}

library("knitr")

### Chunk options: see http://yihui.name/knitr/options/ ###
## Text results
opts_chunk$set(echo = FALSE, warning = TRUE, message = TRUE)

## Code decoration
opts_chunk$set(tidy = TRUE, comment = NA, highlight = TRUE)

## Cache
# opts_chunk$set(cache = 2, cache.path = "ms_2/figures/cache/")

## Plots
opts_chunk$set(fig.path = "../ms_2/figures/")

```

```{r knitcitations, cache = FALSE}
library(knitcitations)
cleanbib()   
cite_options(citation_format = "pandoc")
```

Prior to analysis, we aligned each visitation sample with a corresponding floral survey sample. In most cases, floral surveying was conducted jointly with bumble bee observation for each site and date. Occasionally, though, floral surveying could not be completed on the same day as bumble bee observation, resulting in a small temporal offset between visitation samples and floral survey samples. In these cases, visitation samples were paired with the nearest available floral survey data within a maximum distance of 7 days. Visitation samples lacking corresponding floral survey data within 7 days were omitted from further analysis, which resulted in dropping 40 site-dates out of a total of 895, leaving 855 site-dates included in our analyses. In a few cases, two consecutive visitation samples at a given site shared the same nearest floral survey sample; in these cases, the visitation samples were first pooled, then aligned to their shared nearest floral survey sample. In rare cases, a given floral species was recorded during visitation sampling but missed during floral surveying, and these species were retroactively assigned the minimum floral area value of 0.01 m^2^.

Having thus aligned visitation data and flower cover data, we summed visitation and floral cover across species within each floral morphotype. This yielded, for each of our 855 aligned samples, a cumulative value of flower cover for each floral morphotype and a cumulative value of visitation for each bumble- bee-morphotype pair.

We used 500 iterations to achieve a robust sampling distribution, and we considered observed visitation rates to indicate significant selectivity if they fell outside the interval containing 99% of simulated values.

In addition to the qualifying remarks above, it is important to acknowledge some additional limitations of our methods. First, the inference of selection as use-relative-to-abundance is sensitive to exactly *how* abundance is estimated [@Johnson1980-co; @Beyer2010-vr]. Of particular concern in our study is the problem of spatial scale. By necessity, we measured floral abundance and visitation along transects *within* our study sites. Bumble bee foraging range, however, is much larger than our study sites, and it is possible that the relative abundances of floral morphotypes outside our study sites but within foraging range differ from our within-site estimates. Moreover, because bumble bees can forage across multiple meadow patches in a mountain forest matrix [@Mola2020-yc], there exists a prior selective process --- Johnson's (1980) "second-order selection" --- in which bumble bees choose among available foraging areas before selecting from resources within an area. Finally, to account for the neutral effects of abundance on floral use, we used a simple null model in which the probability of a resource being selected was assumed to be linearly proportional to the relative abundance of the resource. It is possible, however, for resource use to exhibit nonlinear "functional responses" @Mysterud1998-rs to resource abundance, which, relative to a linear null model, would give biased estimates of preference.

Patterns of raw visitation **(Figure 7)** generally mirrored patterns of proportional abundance **(Figure 6)** for each morphotype. Given the expected dependency between floral abundance and raw visitation, this is not surprising. There were, however, some notable deviations. For example, *B. soroeensis* exhibited a strong peak of visitation to bell flowers late in the year at the highest sites, outside the peak abundance of bell flowers. *B. pratorum* was unique in not exhibiting high visitation rate to lip flowers during their peak abundance late in the season and at low-elevation. Bumble bee species differed in the degree to which they exploited the mid- and high-elevation peaks of flag flower abundance; *B. hortorum* and *B. wurflenii* visited flag flowers heavily at high elevation, while *B. pascuorum* --- which also visited flag flowers heavily --- tended rather to exploit the low-elevation peak of abundance. The remaining species --- which, based on preference analysis, appear to have a low affinity for flag flowers **(Figure 2)** --- exhibited diffuse or nonsignficiant patterns of visitation rates that appear uncorrelated with the abundance of flag flowers.

Smooths were significant (p < 0.05) for all bb-fl pairs except hort-bell, hort-head, pasc-bell, telu-head, and soro-flag. The model explained 33.3% of total deviance (adj. R^2^ = 0.272).

```{r fig7, echo = FALSE,  fig.cap = paste("Variation in raw visitation rates through elevation and time modeled as HGAM surfaces. Color ramps depict z-normalized effects to allow the range of effect size to be compared across panels, but contour lines are set at the same fixed density across all panels to emphasize the within-panel patterns. Significant (p < 0.05) smooths are depicted with solid contour lines and nonsignificant smooths with dashed contour lines.")}
knitr::include_graphics("../ms_2/figures/raw_visitation_gam.png")
```

# REFERENCES

<!-- ```{r write_citations, cache=FALSE, include=FALSE} -->
<!-- write.bibtex(file = "knitcitations.bib") -->
<!-- ``` -->

<div id ="refs"></div>