### 2021-09-02 code/ms1_abundance_v2.Rmd
beginning ms1 revision

---
title: "Abundance/diversity GAM"
output: html_document
---

# 1. Prepare environment
## 1.1 Load packages
```{r setup, include=FALSE, echo=FALSE}
library(mgcv)
library(mgcViz)
library(patchwork)
library(tidymv)
library(vegan)
library(tidyverse)
library(lubridate)
```

## 1.2 Load site data and floral taxonomy data
```{r, include=FALSE, echo=FALSE}
site_data <- read_csv("../data/processed_data/site_data.csv") %>%
  dplyr::select(-c(slope.calc, slope.est, elev.min, elev.max, temp.mean, elev.class, elev.class2)) %>% 
  mutate(site = factor(site),
         tree_line = factor(tree_line))

# FL traits
## family-level taxonomy
fl_tax <- read_csv("../data/processed_data/floral_tax.csv") 
```

## 1.3 Load network and survey data
```{r message=FALSE}
### BB-FL visitation data
# net <- read_csv("../data/processed_data/network.csv") %>%
#   filter(caste != "male",
#          !bb.sp %in% c("humi", "hypn", "lapi")) %>% # too rare
#   dplyr::select(site, date, bb.sp, plant.sp) # drop unused columns

net <- read_csv("../data/processed_data/network.csv") %>%
  filter(caste != "male") %>% 
  dplyr::select(site, date, bb.sp, plant.sp) # drop unused columns

### Floral survey data
survey <- read_csv("../data/processed_data/floral_survey.csv") %>%
  semi_join(net, by = "plant.sp") %>% # Consider only visited plant species (anytime anywhere, not specifically for a given site*date)
  dplyr::select(site, date, plant.sp, flower.cover) %>%
  group_by(plant.sp) %>%
  mutate(present = case_when(
    max(flower.cover) > 0 ~ TRUE,
    max(flower.cover) == 0 ~ FALSE
  )) %>%
  filter(present == TRUE) # Ignore plants that never had any flower cover

### Identify species that were visited but not recorded in the survey
vis_sp <- net %>%
  left_join(fl_tax) %>%
  select(plant.sp, plant.genus, plant.family) %>%
  unique()

surv_sp <- survey %>%
  left_join(fl_tax) %>%
  select(plant.sp, plant.genus, plant.family) %>%
  unique()

missing <- anti_join(vis_sp, surv_sp)
```

# 2. Process data
## 2.1 Calculate transect-wise FL and BB abundance
```{r echo=FALSE, include=TRUE, message=FALSE}
# Survey transects
survey_transects_site <- survey %>%
  group_by(site) %>%
  summarize(survey.transects = length(unique(date)))

# BB transects
net_transects_site <- net %>%
  group_by(site) %>%
  summarize(net.transects = length(unique(date)))

# Per-transect BB abundance
bb_abund_sp <- net %>%
  dplyr::select(site, date, bb.sp) %>%
  group_by(site, date, bb.sp) %>% # 
  summarize(bb.abund = n()) %>% # get per transect bb abundance
  ungroup() %>%
  complete(site, date, bb.sp, fill = list(bb.abund = 0)) %>% # complete implicit absences
  semi_join(net, by = c("site", "date")) %>% # drop site-dates that did not exist in the data set
  left_join(site_data) %>% # add site data
  mutate(year = factor(year(date)),
         yday = yday(date),
         bb.sp = factor(bb.sp),
         management = factor(management),
         site = factor(site),
         transect = factor(transect))

### Per-transect floral abund by species
fl_abund_sp <- survey %>%
  group_by(site, date, plant.sp) %>% # group by plant.sp*transect
  summarize(fl.abund = sum(flower.cover)) %>% # get plant.sp*transect total flower cover; this step really shouldn't be required, but in a few cases, there is more than one abundance entry for the same species on a given transect; not sure how that happened, but it's in the original data; alternatively, I could drop the dups or average them. 
  group_by(site, date) %>%
  mutate(prop.fl.abund = fl.abund/sum(fl.abund)) %>% # get the proportional flower cover for each species.
  ungroup() %>%
  complete(site, date, plant.sp, fill = list(fl.abund = 0, prop.fl.abund = 0)) %>%
  semi_join(survey, by = c("site", "date")) %>%  # only include site-dates that actually occurred
  unique() %>%
  mutate(year = factor(year(date))) %>%
  left_join(fl_tax)

### Bumble bee richness per site
bb_rich_site <- net %>%
  group_by(site) %>%
  summarize(bb.rich = length(unique(bb.sp))) %>%
  left_join(site_data) %>% # add site data
  mutate(management = factor(management),
         site = factor(site),
         transect = factor(transect))

### Peak and mean BB abundance per site
bb_abund_site <- bb_abund_sp %>%
  group_by(site, date) %>%
  summarize(bb.total.abund = sum(bb.abund)) %>%
  group_by(site) %>%
  summarize(bb.peakabund = max(bb.total.abund)) %>%
  left_join(site_data) %>% # add site data
  mutate(management = factor(management),
         site = factor(site),
         transect = factor(transect))

### Floral richness per site
fl_rich_site <- survey %>%
  group_by(site) %>%
  summarize(fl.rich = length(unique(plant.sp))) %>%
  left_join(site_data) %>% # add site data
  mutate(management = factor(management),
         site = factor(site),
         transect = factor(transect))

### Peak floral abundance per site
fl_abund_site <- survey %>%
  group_by(site, date) %>%
  summarize(fl.abund = sum(flower.cover)) %>%
  group_by(site) %>%
  summarize(fl.peakabund = max(fl.abund),
            fl.meanabund = mean(fl.abund),
            fl.medabund = median(fl.abund)) %>%
  left_join(site_data) %>% # add site data
  mutate(management = factor(management),
         site = factor(site),
         transect = factor(transect))

### Interaction richness per site
interaction_rich_site <- net %>%
  group_by(site) %>%
  unite(interaction, c(bb.sp, plant.sp), sep = "_") %>%
  summarize(interaction.rich = length(unique(interaction))) %>%
  left_join(site_data) %>% # add site data
  mutate(management = factor(management),
         site = factor(site),
         transect = factor(transect))

rich_abund_collated <- bb_rich_site %>% 
  full_join(fl_rich_site) %>%
  full_join(bb_abund_site) %>% 
  full_join(fl_abund_site) %>%
  full_join(interaction_rich_site)

rich_abund_collated.long <- rich_abund_collated %>%
  pivot_longer(cols = c(bb.rich, fl.rich, bb.peakabund, 
                        fl.peakabund, fl.meanabund, 
                        fl.medabund, interaction.rich)) %>%
  separate(name, c("group", "metric"), sep = "\\.") %>%
  group_by(group, metric) %>%
  mutate(z = scale(value))

ggplot(rich_abund_collated.long, aes(elev.mean, z, color = group)) +
  geom_point() +
  geom_smooth(method = "gam", formula = y ~ s(x, k = 5)) +
  facet_wrap(~metric) +
  theme_light()

# Exploring some bivariate relations

ggplot(rich_abund_collated, aes(fl.rich, bb.rich)) +
  geom_point() +
  geom_smooth(method = "lm")

ggplot(rich_abund_collated, aes(fl.peakabund, fl.rich)) +
  geom_point() +
  geom_smooth(method = "lm")

ggplot(rich_abund_collated, aes(fl.peakabund, bb.peakabund)) +
  geom_point() +
  geom_smooth(method = "lm")

ggplot(rich_abund_collated, aes(fl.peakabund, bb.rich)) +
  geom_point() +
  geom_smooth(method = "lm") 
```

## 2.2 Summarize

# 2. Analyses

## 2.1 BB elevation distribution
```{r}
bb_range <- net %>%
  left_join(site_data) %>%
  group_by(site, date, elev.mean, tree_line, bb.sp) %>%
  summarize(abund = n()) %>%
  group_by(site, elev.mean, tree_line, bb.sp) %>%
  summarize(abund = max(abund)) %>%
  filter(abund > 0) %>%
  group_by(bb.sp) %>%
  mutate(elev.floor = min(elev.mean),
         elev.ceiling = max(elev.mean),
         elev.range = elev.ceiling - elev.floor,
         elev.med = elev.floor + ((elev.ceiling - elev.floor)/2))

ggplot(bb_range, aes(reorder(bb.sp, elev.med), elev.mean)) +
  geom_line(size = 2, alpha = 0.25) +
  geom_point(aes(size = abund, color = tree_line), alpha = 0.5) +
  theme_light(14) +
  scale_size_continuous(name = "Peak abundance",
                        breaks = c(1, 10, 20, 40, 80)) +
  scale_color_discrete(name = "Tree line") +
  xlab("Bumble bee species") +
  ylab("Elevation (m)") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

ggsave("../output/bb_range_ms1.pdf")
ggsave("../output/bb_range_ms1.png")

# Overall bumble bee abundance
bb_abund_sum <- bb_abund_sp %>%
  group_by(site, bb.sp) %>%
  summarize(cum.abund = sum(bb.abund),
            mean.abund = mean(bb.abund)) %>%
  group_by(site) %>%
  mutate(prop.abund = cum.abund/sum(cum.abund)) %>%
  group_by(bb.sp) %>%
  mutate(mean.prop.abund = mean(prop.abund))

ggplot(bb_abund_sum, aes(reorder(bb.sp, -mean.prop.abund), prop.abund)) +
  geom_boxplot(fill = "gray85") +
  theme_light() +
  ylab("Proportional abundance") +
  xlab("Bumble bee species")

ggsave("../output/bb_abundsum.pdf")
ggsave("../output/bb_abundsum.png")
```

## 2.2. Aggregate analysis of max BB abundance and mean FL abundance  ~ elevation

### 2.2.1 Calculate max and mean abundances within site and year
```{r}
# Get mean total weighted floral abundance by BB sp
weighted_fl_abund_mean <- weighted_fl_abund %>%
  group_by(site, bb.sp, year) %>%
  summarize(mean.fl.abund = mean(total.weighted.flower.cover),
            max.fl.abund = max(total.weighted.flower.cover))

# Get max BB abundance by sp
bb_abund_sp_max <- bb_abund_sp %>%
  group_by(site, elev.mean, year, bb.sp) %>%
  summarize(max.bb.abund = max(bb.abund),
            mean.bb.abund = mean(bb.abund))

# Join
abund_sum <- full_join(weighted_fl_abund_mean, bb_abund_sp_max) %>%  
  unite(sp.year, c(bb.sp, year), remove = FALSE) %>%
  mutate(site = factor(site),
         bb.sp = factor(bb.sp),
         sp.year = factor(sp.year)) %>%
  unite(bb.sp.year, c(bb.sp, year), sep = "_", remove = FALSE) %>%
  mutate(bb.sp.year = factor(bb.sp.year)) %>%
  left_join(site_data) %>%
  mutate(management = factor(management))
```

### 2.2.2 Call models
```{r}
# BB
gam_max_bb_abund_01 <- gam(max.bb.abund ~ 
                             management +
                             s(bb.sp, bs = "re") + 
                             s(bb.sp.year, bs = "re") + 
                             s(elev.mean, by = bb.sp, k = 8), 
                           method = "REML",
                           family = "nb",
                           select = TRUE,
                           data = abund_sum) %>% getViz()

saveRDS(gam_max_bb_abund_01, "../output/gam_max_bb_abund_01.rds")

# FL
gam_mean_fl_abund_01 <- gam(mean.fl.abund ~
                              management +
                              year +
                              s(bb.sp, bs = "re") +
                              s(elev.mean, by = bb.sp, k = 8),
                           method = "REML",
                           family = "tw",
                           select = TRUE,
                           data = abund_sum) %>% getViz()

saveRDS(gam_mean_fl_abund_01, "../output/gam_mean_fl_abund_01.rds")
```

### 2.2.3 Evaluate models
```{r}
# BB
gam_max_bb_abund_01_check <- check.gamViz(gam_max_bb_abund_01) # evaluate model
gam_max_bb_abund_01_sum <- summary(gam_max_bb_abund_01) # summarize model
print(plot(gam_max_bb_abund_01, allTerms = TRUE), pages = 1) # preliminary visualization

# FL
gam_mean_fl_abund_01_check <- check.gamViz(gam_mean_fl_abund_01) # evaluate model
gam_mean_fl_abund_01_sum <- summary(gam_mean_fl_abund_01) # summarize model
print(plot(gam_mean_fl_abund_01, allTerms = TRUE), pages = 1) # preliminary visualization

```

### 2.2.4 Visualize

#### 2.2.4.1 BB
```{r}
# Plot on the link scale
## Individual plots
gers_abund <- plot(sm(gam_max_bb_abund_01, 3)) + l_fitLine() + l_ciLine() + 
  labs(x = NULL, y = NULL, title = "gers") +
  l_rug(length = unit(0.15, "cm"))
hort_abund <- plot(sm(gam_max_bb_abund_01, 4)) + l_fitLine() + l_ciLine() + 
  labs(x = NULL, y = NULL, title = "hort") +
  l_rug(length = unit(0.15, "cm"))
jone_abund <- plot(sm(gam_max_bb_abund_01, 5)) + l_fitLine() + l_ciLine() + 
  labs(x = NULL, y = NULL, title = "jone") +
  l_rug(length = unit(0.15, "cm"))
mend_abund <- plot(sm(gam_max_bb_abund_01, 6)) + l_fitLine() + l_ciLine() + 
  labs(x = NULL, y = NULL, title = "mend") +
  l_rug(length = unit(0.15, "cm"))
mont_abund <- plot(sm(gam_max_bb_abund_01, 7)) + l_fitLine() + l_ciLine() + 
  labs(x = NULL, y = NULL, title = "mont") +
  l_rug(length = unit(0.15, "cm"))
muci_abund <- plot(sm(gam_max_bb_abund_01, 8)) + l_fitLine() + l_ciLine() + 
  labs(x = NULL, y = NULL, title = "muci") +
  l_rug(length = unit(0.15, "cm"))
pasc_abund <- plot(sm(gam_max_bb_abund_01, 9)) + l_fitLine() + l_ciLine() + 
  labs(x = NULL, y = NULL, title = "pasc") +
  l_rug(length = unit(0.15, "cm"))
prat_abund <- plot(sm(gam_max_bb_abund_01, 10)) + l_fitLine() + l_ciLine() + 
  labs(x = NULL, y = NULL, title = "prat") +
  l_rug(length = unit(0.15, "cm"))
psit_abund <- plot(sm(gam_max_bb_abund_01, 11)) + l_fitLine() + l_ciLine() + 
  labs(x = NULL, y = NULL, title = "psit") +
  l_rug(length = unit(0.15, "cm"))
pyre_abund <- plot(sm(gam_max_bb_abund_01, 12)) + l_fitLine() + l_ciLine() + 
  labs(x = NULL, y = NULL, title = "pyre") +
  l_rug(length = unit(0.15, "cm"))
soro_abund <- plot(sm(gam_max_bb_abund_01, 13)) + l_fitLine() + l_ciLine() + 
  labs(x = NULL, y = NULL, title = "soro") +
  l_rug(length = unit(0.15, "cm"))
telu_abund <- plot(sm(gam_max_bb_abund_01, 14)) + l_fitLine() + l_ciLine() + 
  labs(x = NULL, y = NULL, title = "telu") +
  l_rug(length = unit(0.15, "cm"))
wurf_abund <- plot(sm(gam_max_bb_abund_01, 15)) + l_fitLine() + l_ciLine() + 
  labs(x = NULL, y = NULL, title = "wurf") +
  l_rug(length = unit(0.15, "cm"))
mgmt_bb_effect <- plot(pterm(gam_max_bb_abund_01, 1)) + l_fitPoints() + l_ciBar() + 
  labs(x = NULL, y = NULL, title = "Management")

## Arrange smooths on grid plot

grid_bb_abund <- gridPrint(gers_abund, hort_abund, jone_abund, mend_abund, mont_abund, muci_abund, 
                           pasc_abund, prat_abund, psit_abund, pyre_abund, 
                           soro_abund, telu_abund, wurf_abund, mgmt_bb_effect, ncol = 4)

pdf("../output/gam_bb_abund_grid.pdf", width = 8, height = 6)
plot(grid_bb_abund)
dev.off()

png("../output/gam_bb_abund_grid.png", res = 300, width = 8, height = 6, units = "in")
plot(grid_bb_abund)
dev.off()

# 
# 
# # Extract predictions and plot on response scale
# 
# gam_max_bb_abund_01_ilink <- family(gam_max_bb_abund_01)$linkinv # get inverse link function
# 
# gam_max_bb_abund_01_pred <- predict.gam(gam_max_bb_abund_01, # get model predictions
#                                         type = "terms") %>% 
#   as.data.frame() %>% # convert to data frame
#   as_tibble() %>% # convert to tibble
#   dplyr::select(mgmt = 1, bb.sp = 2, bb.sp.year = 3, gers = 4, # rename
#          hort = 5, jone = 6, mend = 7, mont = 8, 
#          muci = 9, pasc = 10, prat = 11, psit = 12, 
#          pyre = 13, soro = 14, telu = 15, wurf = 16) %>%
#   mutate(fit = # get cumulative fit
#            mgmt + bb.sp + bb.sp.year + gers + 
#            hort + jone + mend + mont + 
#            muci + pasc + prat + psit + 
#            pyre + soro + telu + wurf 
#            ) %>%
#   dplyr::select(fit) %>%
#   mutate(fit.trans = gam_max_bb_abund_01_ilink(fit)) %>% # back-transform to response scale
#   bind_cols(abund_sum)
# 
# ggplot(gam_max_bb_abund_01_pred, aes(elev.mean, fit.trans, linetype = year)) + # map year to linetype
#   geom_point(aes(elev.mean, max.bb.abund, shape = tree_line, color = management), 
#              inherit.aes = FALSE, alpha = 0.5, size = 2) + # add original data and map tree line to color
#   #geom_line(size = 0.5, alpha = 0.5) +
#   facet_wrap(~bb.sp, scales = "free_y") +
#   theme_light() +
#   labs(x = "Elevation (m.s.l.)", 
#        y = bquote("Peak bumble bee abundance"),
#        color = "Tree line",
#        shape = "Year",
#        linetype = "Year")
# 
# ggsave("../output/gam_bb_abund.pdf")
# ggsave("../output/gam_bb_abund.png")
```

#### 2.2.4.2 FL
```{r}
# Plot on the link scale

## Individual plots

gers_fl_abund <- plot(sm(gam_mean_fl_abund_01, 2)) + l_fitLine() + l_ciLine() + 
  labs(x = NULL, y = NULL, title = "gers") +
  l_rug(length = unit(0.15, "cm"))
hort_fl_abund <- plot(sm(gam_mean_fl_abund_01, 3)) + l_fitLine() + l_ciLine() + 
  labs(x = NULL, y = NULL, title = "hort") +
  l_rug(length = unit(0.15, "cm"))
jone_fl_abund <- plot(sm(gam_mean_fl_abund_01, 4)) + l_fitLine() + l_ciLine() + 
  labs(x = NULL, y = NULL, title = "jone") +
  l_rug(length = unit(0.15, "cm"))
mend_fl_abund <- plot(sm(gam_mean_fl_abund_01, 5)) + l_fitLine() + l_ciLine() + 
  labs(x = NULL, y = NULL, title = "mend") +
  l_rug(length = unit(0.15, "cm"))
mont_fl_abund <- plot(sm(gam_mean_fl_abund_01, 6)) + l_fitLine() + l_ciLine() + 
  labs(x = NULL, y = NULL, title = "mont") +
  l_rug(length = unit(0.15, "cm"))
muci_fl_abund <- plot(sm(gam_mean_fl_abund_01, 7)) + l_fitLine() + l_ciLine() + 
  labs(x = NULL, y = NULL, title = "muci") +
  l_rug(length = unit(0.15, "cm"))
pasc_fl_abund <- plot(sm(gam_mean_fl_abund_01, 8)) + l_fitLine() + l_ciLine() + 
  labs(x = NULL, y = NULL, title = "pasc") +
  l_rug(length = unit(0.15, "cm"))
prat_fl_abund <- plot(sm(gam_mean_fl_abund_01, 9)) + l_fitLine() + l_ciLine() + 
  labs(x = NULL, y = NULL, title = "prat") +
  l_rug(length = unit(0.15, "cm"))
psit_fl_abund <- plot(sm(gam_mean_fl_abund_01, 10)) + l_fitLine() + l_ciLine() + 
  labs(x = NULL, y = NULL, title = "psit") +
  l_rug(length = unit(0.15, "cm"))
pyre_fl_abund <- plot(sm(gam_mean_fl_abund_01, 11)) + l_fitLine() + l_ciLine() + 
  labs(x = NULL, y = NULL, title = "pyre") +
  l_rug(length = unit(0.15, "cm"))
soro_fl_abund <- plot(sm(gam_mean_fl_abund_01, 12)) + l_fitLine() + l_ciLine() + 
  labs(x = NULL, y = NULL, title = "soro") +
  l_rug(length = unit(0.15, "cm"))
telu_fl_abund <- plot(sm(gam_mean_fl_abund_01, 13)) + l_fitLine() + l_ciLine() + 
  labs(x = NULL, y = NULL, title = "telu") +
  l_rug(length = unit(0.15, "cm"))
wurf_fl_abund <- plot(sm(gam_mean_fl_abund_01, 14)) + l_fitLine() + l_ciLine() + 
  labs(x = NULL, y = NULL, title = "wurf") +
  l_rug(length = unit(0.15, "cm"))
mgmt_fl_effect <- plot(pterm(gam_mean_fl_abund_01, 1)) + l_fitPoints() + l_ciBar() + 
  labs(x = NULL, y = NULL, title = "Management")
year_fl_effect <- plot(pterm(gam_mean_fl_abund_01, 2)) + l_fitPoints() + l_ciBar() + 
  labs(x = NULL, y = NULL, title = "Year")

## Arrange significant smooths on grid plot

grid_fl_abund <- gridPrint(gers_fl_abund, hort_fl_abund, jone_fl_abund, mend_fl_abund, mont_fl_abund, muci_fl_abund, 
                           pasc_fl_abund, prat_fl_abund, psit_fl_abund, pyre_fl_abund, 
                           soro_fl_abund, telu_fl_abund, wurf_fl_abund, year_fl_effect, mgmt_fl_effect, ncol = 4)

pdf("../output/gam_fl_abund_grid.pdf", width = 8, height = 6)
plot(grid_fl_abund)
dev.off()

png("../output/gam_fl_abund_grid.png", res = 300, width = 8, height = 6, units = "in")
plot(grid_fl_abund)
dev.off()

# 
# 
# # Extract predictions and plot on response scale
# 
# gam_mean_fl_abund_01_ilink <- family(gam_mean_fl_abund_01)$linkinv # get inverse link function to transform to the response scale
# 
# gam_mean_fl_abund_01_pred <- predict.gam(gam_mean_fl_abund_01, # get model predictions
#                                         type = "terms") %>% 
#   as.data.frame() %>% # convert to data frame
#   as_tibble() %>% # convert to tibble
#   dplyr::select(mgmt = 1, year = 2, bb.sp = 3, gers = 4, # rename
#          hort = 5, jone = 6, mend = 7, mont = 8, 
#          muci = 9, pasc = 10, prat = 11, psit = 12, 
#          pyre = 13, soro = 14, telu = 15, wurf = 16) %>%
#   mutate(fit = # get cumulative fit
#            mgmt + year + bb.sp + gers + 
#            hort + jone + mend + mont + 
#            muci + pasc + prat + psit + 
#            pyre + soro + telu + wurf 
#            ) %>%
#   dplyr::select(fit) %>%
#   mutate(fit.trans = gam_max_bb_abund_01_ilink(fit)) %>% # back-transform to response scale
#   bind_cols(na.omit(abund_sum)) # the na.omit() drops the rows where floral surveying data were missing
# 
# ggplot(gam_mean_fl_abund_01_pred, aes(elev.mean, fit.trans, linetype = year)) + # map year to linetype
#   geom_point(aes(elev.mean, mean.fl.abund, color = management), inherit.aes = FALSE, alpha = 0.5) + # add original data and map tree line to color
#   #geom_line(size = 0.5, alpha = 0.5) +
#   facet_wrap(~bb.sp, scales = "free_y") +
#   theme_light() +
#   labs(x = "Elevation (m.s.l.)", 
#        y = bquote("Peak floral abundance"),
#        color = "Tree line",
#        shape = "Year",
#        linetype = "Year")
# 
# ggsave("../output/gam_fl_abund.pdf")
# ggsave("../output/gam_fl_abund.png")
```

## 2.3 Mean total floral cover (not broken down by bumble bee species) ~ elevation * time
```{r}
fl_abund_total <- fl_abund_sp %>%
  group_by(site, date) %>%
  summarize(fl.abund = sum(fl.abund)) %>%
  left_join(site_data) %>%
  mutate(yday = yday(date),
         year = factor(year(date)),
         site = factor(site),
         management = factor(management))

bb_abund_total <- net %>%
  group_by(site, date) %>%
  summarize(bb.abund = n()) %>%
  left_join(site_data) %>%
  mutate(yday = yday(date),
         year = factor(year(date)),
         site = factor(site),
         management = factor(management))

fl_rich_total <- fl_abund_sp %>%
  filter(fl.abund > 0) %>%
  group_by(site, date) %>%
  summarize(fl.rich = length(unique(plant.sp))) %>%
  left_join(site_data) %>%
  mutate(yday = yday(date),
         year = factor(year(date)),
         site = factor(site),
         management = factor(management))

bb_rich_total <- net %>%
  group_by(site, date) %>%
  summarize(bb.rich = length(unique(bb.sp))) %>%
  left_join(site_data) %>%
  mutate(yday = yday(date),
         year = factor(year(date)),
         site = factor(site),
         management = factor(management))

fl_rich_total <- fl_abund_sp %>%
  filter(fl.abund > 0) %>%
  group_by(site, date) %>%
  summarize(fl.rich = length(unique(plant.sp))) %>%
  left_join(site_data) %>%
  mutate(yday = yday(date),
         year = factor(year(date)),
         site = factor(site),
         management = factor(management))

ggplot(fl_abund_total, aes(fl.abund)) +
  geom_histogram() +
  facet_wrap(~site)

ggplot(fl_rich_total, aes(fl.rich)) +
  geom_histogram() +
  facet_wrap(~site)

### Abundance
total_fl_abund_GAM <- gam(log(fl.abund + 0.01) ~ 
                             year +
                             management +
                             s(site, bs = "re") +
                             te(yday, elev.mean, by = year, bs=c("gp","tp")),
                           family = "gaussian",
                           method = "REML",
                           data = fl_abund_total) %>% getViz()

check.gamViz(total_fl_abund_GAM)
summary(total_fl_abund_GAM)
plot.gamViz(total_fl_abund_GAM, allTerms = TRUE)

total_bb_abund_GAM <- gam(bb.abund ~ 
                             year +
                             management +
                             s(site, bs = "re") +
                             te(yday, elev.mean, by = year, bs=c("gp","tp")),
                           family = "nb",
                           method = "REML",
                           data = bb_abund_total) %>% getViz()

check.gamViz(total_bb_abund_GAM)
summary(total_bb_abund_GAM)
plot.gamViz(total_bb_abund_GAM, allTerms = TRUE)

### Richness

total_fl_rich_GAM <- gam(fl.rich ~ 
                            year +
                            management +
                            s(site, bs = "re") +
                            te(yday, elev.mean, by = year, bs=c("gp","tp")),
                          family = "poisson",
                          method = "REML",
                          data = fl_rich_total) %>% getViz()

check.gamViz(total_fl_rich_GAM)
summary(total_fl_rich_GAM)
plot.gamViz(total_fl_rich_GAM, allTerms = TRUE)

total_bb_rich_GAM <- gam(bb.rich ~ 
                            year +
                            management +
                            s(site, bs = "re") +
                            te(yday, elev.mean, by = year, bs=c("gp","tp")),
                          family = "nb",
                          method = "REML",
                          data = bb_rich_total) %>% getViz()

check.gamViz(total_bb_rich_GAM)
summary(total_bb_rich_GAM)
plot.gamViz(total_bb_rich_GAM, allTerms = TRUE)
```