### 2021-09-02 ms_1/Appendix_S1.Rmd

---
title: "The tree line ecotone punctuates the abundance and β-diversity of plants, pollinators, and their interactions"
subtitle: "Appendix S1: Analytical workflow"
author: 
  - Douglas B. Sponsler^[douglas.sponsler@uni-wuerzburg.de]
  - Katharina Kallnik 
  - Fabrice Requier 
  - Alice Classen 
  - A. Fabienne Maihoff
  - Johanna Sieger
  - Ingolf Steffan-Dewenter
output: 
  pdf_document:
    toc: true
    toc_depth: 3
    latex_engine: xelatex
#mainfont: Arial
---

# Packages
```{r message=FALSE, warning=FALSE}
library(mgcViz)
library(lubridate)
library(tidyverse)
```

# Sampling patterns
## Load sampling data
```{r message=FALSE, warning=FALSE}
### BB-FL visitation data
visitation_sampling <- read_csv("../data/processed_data/network.csv") %>%
  filter(caste != "male") %>%
  dplyr::select(site, date, bb.sp, plant.sp) # drop unused columns

### Floral survey data
floral_surveying <- read_csv("../data/processed_data/floral_survey.csv") %>%
  dplyr::select(site, date, plant.sp, flower.cover) %>%
  group_by(plant.sp) %>%
  mutate(present = case_when(
    max(flower.cover) > 0 ~ TRUE,
    max(flower.cover) == 0 ~ FALSE
  )) %>%
  filter(present == TRUE) # Ignore plants that never had any flower cover

site_data <- read_csv("../data/processed_data/site_data.csv") %>%
  dplyr::select(-c(slope.calc, slope.est, elev.min, elev.max, temp.mean, elev.class, elev.class2)) %>% 
  mutate(site = factor(site),
         tree_line = factor(tree_line))
```

## Process sampling data
```{r message=FALSE, warning=FALSE}
bb_obs_per_year <- visitation_sampling %>%
  mutate(year = year(date)) %>%
  group_by(year) %>%
  summarize(obs = n())

bb_obs_per_sample <- visitation_sampling %>%
  mutate(year = factor(year(date))) %>%
  group_by(year, site, date) %>%
  summarize(obs = n()) %>%
  left_join(site_data) %>%
  mutate(date =as_date(yday(date)))

bb_samples_per_year <- visitation_sampling %>%
  mutate(year = year(date)) %>%
  select(year, site, date) %>%
  unique() %>%
  group_by(year) %>%
  summarize(samples = n())

bb_summary <- full_join(bb_obs_per_year, bb_samples_per_year) %>%
  mutate(obs.per.sample = obs/samples)
```

## Visualize
```{r message=FALSE, warning=FALSE}
ggplot(visitation_sampling, aes(x = "x")) +
  geom_bar(fill = "gray60") +
  ylab("Observations") +
  facet_wrap(~year(date)) +
  theme_light(12) +
  theme(axis.title.x = element_blank(),
        axis.text.x = element_blank(),
        axis.ticks.x = element_blank()) +
  ggtitle("Bumble bee observations per study year")

ggplot(bb_obs_per_sample, aes(x = "x")) +
  geom_bar(fill = "gray60") +
  ylab("Samples") +
  facet_wrap(~year) +
  theme_light(12) +
  theme(axis.title.x = element_blank(),
        axis.text.x = element_blank(),
        axis.ticks.x = element_blank()) +
  ggtitle("Samplings (site*day) per year")

ggplot(bb_obs_per_sample, aes(x = "x", y = obs)) +
  geom_violin(draw_quantiles = 0.5, fill = "gray60", color = NA) +
  geom_boxplot(fill = NA) +
  #geom_density(outline.type = "full", position = "fill") +
  ylab("Observations/sample") +
  #xlab(NULL) +
  facet_wrap(~year) +
  theme_light(12) +
  theme(axis.title.x = element_blank(),
        axis.text.x = element_blank(),
        axis.ticks.x = element_blank()) +
  ggtitle("Bumble bee observations per sample")

ggplot(bb_obs_per_sample, aes(as_date(yday(date)), elev.mean/10)) +
  geom_point(size = 1, aes(alpha = obs)) +
  facet_wrap(~year) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_alpha_continuous(name = "Observations") +
  scale_x_date(date_labels = "%b", breaks = as_date(c("1970-04-01", "1970-05-01", "1970-06-01",
                                                    "1970-07-01", "1970-08-01", "1970-09-01"))) +
  theme_light(12) +
  labs(x = NULL,
       y = "Elevation (m x 10)") +
  ggtitle("Bumble bee observations through elevation and time")
```

# Bumble bee abundance
```{r message=FALSE, warning=FALSE}
bb_range <- visitation_sampling %>%
  left_join(site_data) %>%
  group_by(site, date, elev.mean, tree_line, bb.sp) %>%
  summarize(abund = n()) %>%
  group_by(site, elev.mean, tree_line, bb.sp) %>%
  summarize(abund = max(abund)) %>%
  filter(abund > 0) %>%
  group_by(bb.sp) %>%
  mutate(elev.floor = min(elev.mean),
         elev.ceiling = max(elev.mean),
         elev.range = elev.ceiling - elev.floor,
         elev.med = elev.floor + ((elev.ceiling - elev.floor)/2))

ggplot(bb_range, aes(reorder(bb.sp, elev.med), elev.mean)) +
  geom_line(size = 2, alpha = 0.25) +
  geom_point(aes(size = abund, color = tree_line), alpha = 0.5) +
  theme_light(14) +
  scale_size_continuous(name = "Peak abundance",
                        breaks = c(1, 10, 20, 40, 80)) +
  scale_color_discrete(name = "Tree line") +
  xlab("Bumble bee species") +
  ylab("Elevation (m)") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  ggtitle("Range and peak abundance of each bumble bee sp.")

# Overall bumble bee abundance
bb_abund_sp <- visitation_sampling %>%
  dplyr::select(site, date, bb.sp) %>%
  group_by(site, date, bb.sp) %>% # 
  summarize(bb.abund = n()) %>% # get per transect bb abundance
  ungroup() %>%
  complete(site, date, bb.sp, fill = list(bb.abund = 0)) %>% # complete implicit absences
  semi_join(visitation_sampling, by = c("site", "date")) %>% # drop site-dates that did not exist in the data set
  left_join(site_data) %>% # add site data
  mutate(year = factor(year(date)),
         yday = yday(date),
         bb.sp = factor(bb.sp),
         management = factor(management),
         site = factor(site),
         transect = factor(transect))

bb_abund_sum <- bb_abund_sp %>%
  group_by(site, bb.sp) %>%
  summarize(cum.abund = sum(bb.abund),
            mean.abund = mean(bb.abund)) %>%
  group_by(site) %>%
  mutate(prop.abund = cum.abund/sum(cum.abund)) %>%
  group_by(bb.sp) %>%
  mutate(mean.prop.abund = mean(prop.abund))

ggplot(bb_abund_sum, aes(reorder(bb.sp, -mean.prop.abund), prop.abund)) +
  geom_boxplot(fill = "gray85") +
  theme_light() +
  ylab("Proportional abundance") +
  xlab("Bumble bee species") +
  ggtitle("Relative abundances of bumble bee species")
```

# GAM diagnostics

## Bumble bee abundance GAM
```{r message=FALSE, warning=FALSE}
bb_abund_gam <- readRDS("../output/gam_max_bb_abund_01.rds")
summary(bb_abund_gam)
check.gamViz(bb_abund_gam)
```

## Floral abundance GAM
```{r message=FALSE, warning=FALSE}
fl_abund_gam <- readRDS("../output/gam_mean_fl_abund_01.rds")
summary(fl_abund_gam)
check.gamViz(fl_abund_gam)
```