### 2021-09-02 ms_2/ms_2_oikos_part1.Rmd

---
title: "Floral preferences of mountain bumble bees are constrained by trait matching but flexible through elevation and season"
header-includes:
   - \usepackage{lineno}\linenumbers
   - \usepackage{setspace}\doublespacing
output: 
  pdf_document: 
    keep_tex: yes
---

**Douglas Sponsler** (corresponding author)  
University of Würzburg  
Department of Animal Ecology and Tropical Biology  
Biozentrum der Universität Würzburg C-018  
Am Hubland  
97074 Würzburg, Germany  
Email: douglas.sponsler@uni-wuerzburg.de  
Telephone: +49 0176 363 20273

**Katharina Kallnik**  
University of Würzburg  
Department of Animal Ecology and Tropical Biology  
97074 Würzburg, Germany  
Email: katharina.kallnik@uni-wuerzburg.de

**Fabrice Requier**  
Université Paris-Saclay  
Évolution, Génomes, Comportement et Écologie  
91198 Gif-sur-Yvette, France  
Email: fabrice.requier@egce.cnrs-gif.fr

**Alice Classen**  
University of Würzburg  
Department of Animal Ecology and Tropical Biology  
97074 Würzburg, Germany  
Email: alice.classen@uni-wuerzburg.de

**A. Fabienne Maihoff**  
University of Würzburg  
Department of Animal Ecology and Tropical Biology  
97074 Würzburg, Germany  
Email: anne_fabienne.maihoff@uni-wuerzburg.de

**Johanna Sieger**  
Naturpark Frankenhöhe  
91522 Ansbach, Germany   
Email: sieger@lpv-mfr.de

**Ingolf Steffan-Dewenter**  
University of Würzburg  
Department of Animal Ecology and Tropical Biology  
97074 Würzburg, Germany  
Email: ingolf.steffan@uni-wuerzburg.de

\

**AUTHORSHIP STATEMENT:** KK and ISD conceived the study. KK, JS, and ISD designed and carried out field work. DS analyzed the data and wrote the paper with substantial contributions from FR, AC, AFM, KK, and ISD.

**DATA ACCESSIBILITY:** All data used in this study have been archived on DRYAD (https://datadryad.org/stash/dataset/doi:10.5061/dryad.s1rn8pk7f) and will be available upon article publication.

**ARTICLE TYPE:** Research

**RUNNING TITLE:** Bumble bee floral preferences

**SECTION LENGTHS:**  
Abstract: 141 words
Main text: 4979 words
References: 61  
Figures: 6