### 2021-09-22 code/ms1_rich_abund.Rmd

---
title: "Abundance/diversity GAM"
output: html_document
---

# 1. Prepare environment
## 1.1 Load packages
```{r setup, include=FALSE, echo=FALSE}
library(mgcv)
library(mgcViz)
library(magick)
library(patchwork)
library(tidymv)
library(vegan)
library(gratia)
library(tidyverse)
library(broom)
library(lubridate)
library(cowplot)
```

## 1.2 Load site data and floral taxonomy data
```{r, include=FALSE, echo=FALSE}
site_data <- read_csv("../data/processed_data/site_data.csv") %>%
  dplyr::select(-c(slope.calc, slope.est, elev.min, elev.max, 
                   temp.mean, elev.class, elev.class2)) %>% 
  mutate(site = factor(site),
         tree_line = factor(tree_line))

# FL traits
## family-level taxonomy
fl_tax <- read_csv("../data/processed_data/floral_tax.csv") 
```

## 1.3 Load network and survey data
```{r message=FALSE}
### BB-FL visitation data
net <- read_csv("../data/processed_data/network.csv") %>%
  filter(caste != "male") %>% 
  dplyr::select(site, date, bb.sp, plant.sp) # drop unused columns

### Floral survey data
survey <- read_csv("../data/processed_data/floral_survey.csv") %>%
  # Consider only visited plant species (anytime anywhere, not specifically for a given site*date)
  semi_join(net, by = "plant.sp") %>% 
  dplyr::select(site, date, plant.sp, flower.cover) %>%
  group_by(plant.sp) %>%
  mutate(present = case_when(
    max(flower.cover) > 0 ~ TRUE,
    max(flower.cover) == 0 ~ FALSE
  )) %>%
  filter(present == TRUE) # Ignore plants that never had any flower cover
```

## 1.4 Load images
```{r}
BB <- image_read("../ms_1/figures/phylopic_BB.png")
FL <- image_read("../ms_1/figures/phylopic_FL.png")
BBxFL <- image_read("../ms_1/figures/phylopic_BBxFL.png")
```

## 1.5 Define functions
```{r}
gam_tab1 <-function(x1, x2, x3, 
                    name1, name2, name3) {

  tab1.1 <- x1 %>%
    tidy(parametric = TRUE, conf.int = TRUE) %>%
    mutate(class = rep("parametric", n()),
           estimate = round(estimate, 3),
           statistic = round(statistic, 3),
           p.value = round(p.value, 3),
           model = rep(name1, n())
           )

  tab1.2 <- x1 %>%
    tidy(parametric = FALSE) %>%
    mutate(class = rep("smooth", n()),
           edf = round(edf, 2),
           statistic = round(statistic, 3),
           p.value = round(p.value, 3),
           model = rep(name1, n()))

  tab2.1 <- x2 %>%
    tidy(parametric = TRUE, conf.int = TRUE) %>%
    mutate(class = rep("parametric", n()),
           estimate = round(estimate, 3),
           statistic = round(statistic, 3),
           p.value = round(p.value, 3),
           model = rep(name2, n())
           )

  tab2.2 <- x2 %>%
    tidy(parametric = FALSE) %>%
    mutate(class = rep("smooth", n()),
           edf = round(edf, 2),
           statistic = round(statistic, 3),
           p.value = round(p.value, 3),
           model = rep(name2, n())
           )

  tab3.1 <- x3 %>%
    tidy(parametric = TRUE, conf.int = TRUE) %>%
    mutate(class = rep("parametric", n()),
           estimate = round(estimate, 3),
           statistic = round(statistic, 3),
           p.value = round(p.value, 3),
           model = rep(name3, n())
           )

  tab3.2 <- x3 %>%
    tidy(parametric = FALSE) %>%
    mutate(class = rep("smooth", n()),
           edf = round(edf, 2),
           statistic = round(statistic, 3),
           p.value = round(p.value, 3),
           model = rep(name3, n())
           )

  out <- bind_rows(tab1.1, tab1.2, tab2.1, tab2.2, tab3.1, tab3.2) %>%
    select(model, class, term, estimate, edf, statistic, p.value) %>%
    mutate(term = case_when(
      term == "(Intercept)" ~ "Intercept",
      term == "mgmtnone" ~ "mgmt:none",
      term == "mgmtmowing" ~ "mgmt:mowing",
      term == "year2011" ~ "year:2011",
      term == "year2012" ~ "year:2012",
      term == "s(site)" ~ "s(site)",
      term == "s(net.transects)" ~ "s(samples)",
      term == "s(survey.transects)" ~ "s(samples)",
      term == "s(elev.mean)" ~ "s(elevation)",
      term == "s(elev.mean):year2010" ~ "s(elevation:2010)",
      term == "s(elev.mean):year2011" ~ "s(elevation:2011)",
      term == "s(elev.mean):year2012" ~ "s(elevation:2012)"),
      sig = case_when(
        p.value < 0.001 ~ "***",
        p.value >= 0.001 & p.value < 0.01 ~ "**",
        p.value >= 0.01 & p.value < 0.05 ~ "*",
        p.value >= 0.05 ~ ""
      ),
      p.value = ifelse(p.value < 0.001, "<0.001", p.value),
      statistic = ifelse(statistic < 0.001, "<0.001", statistic),
      edf = ifelse(edf < 0.001, "<0.001", edf)) %>%
    flextable() %>%
    merge_v(j = c("model", "class")) %>%
    valign(j = c("model", "class"), valign = "top") %>%
    autofit()
}

gam_tab2 <-function(x1, x2, name1, name2) {

  tab1.1 <- x1 %>%
    tidy(parametric = TRUE, conf.int = TRUE) %>%
    mutate(class = rep("parametric", n()),
           estimate = round(estimate, 3),
           statistic = round(statistic, 3),
           p.value = round(p.value, 3),
           model = rep(name1, n())
           )

  tab1.2 <- x1 %>%
    tidy(parametric = FALSE) %>%
    mutate(class = rep("smooth", n()),
           edf = round(edf, 2),
           statistic = round(statistic, 3),
           p.value = round(p.value, 3),
           model = rep(name1, n()))

  tab2.1 <- x2 %>%
    tidy(parametric = TRUE, conf.int = TRUE) %>%
    mutate(class = rep("parametric", n()),
           estimate = round(estimate, 3),
           statistic = round(statistic, 3),
           p.value = round(p.value, 3),
           model = rep(name2, n())
           )

  tab2.2 <- x2 %>%
    tidy(parametric = FALSE) %>%
    mutate(class = rep("smooth", n()),
           edf = round(edf, 2),
           statistic = round(statistic, 3),
           p.value = round(p.value, 3),
           model = rep(name2, n())
           )

  out <- bind_rows(tab1.1, tab1.2, tab2.1, tab2.2) %>%
    select(model, class, term, estimate, edf, statistic, p.value) %>%
    mutate(term = case_when(
      term == "(Intercept)" ~ "Intercept",
      term == "mgmtnone" ~ "mgmt:none",
      term == "mgmtmowing" ~ "mgmt:mowing",
      term == "year2011" ~ "year:2011",
      term == "year2012" ~ "year:2012",
      term == "s(site)" ~ "s(site)",
      term == "s(net.transects)" ~ "s(samples)",
      term == "s(survey.transects)" ~ "s(samples)",
      term == "s(elev.mean)" ~ "s(elevation)",
      term == "s(elev.mean):year2010" ~ "s(elevation:2010)",
      term == "s(elev.mean):year2011" ~ "s(elevation:2011)",
      term == "s(elev.mean):year2012" ~ "s(elevation:2012)"),
      sig = case_when(
        p.value < 0.001 ~ "***",
        p.value >= 0.001 & p.value < 0.01 ~ "**",
        p.value >= 0.01 & p.value < 0.05 ~ "*",
        p.value >= 0.05 ~ ""
      ),
      p.value = ifelse(p.value < 0.001, "<0.001", p.value),
      statistic = ifelse(statistic < 0.001, "<0.001", statistic),
      edf = ifelse(edf < 0.001, "<0.001", edf)) %>%
    flextable() %>%
    merge_v(j = c("model", "class")) %>%
    valign(j = c("model", "class"), valign = "top") %>%
    autofit()
}
```

## 1.6 Identify plant species that were visited but not recorded in the survey
```{r message=FALSE}
vis_sp <- net %>%
  left_join(fl_tax) %>%
  select(plant.sp, plant.genus, plant.family) %>%
  unique()

surv_sp <- survey %>%
  left_join(fl_tax) %>%
  select(plant.sp, plant.genus, plant.family) %>%
  unique()

missing <- anti_join(vis_sp, surv_sp)
```

# 2. Process data
## 2.1 Calculate transect-wise FL and BB abundance and richness
```{r echo=FALSE, include=TRUE, message=FALSE}
# Survey transects per site
survey_transects_siteyear <- survey %>%
  mutate(year = factor(year(date)),
         site = factor(site)) %>%
  group_by(site, year) %>%
  summarize(survey.transects = length(unique(date)))

# BB transects per site
net_transects_siteyear <- net %>%
  mutate(year = factor(year(date)),
         site = factor(site)) %>%
  group_by(site, year) %>%
  summarize(net.transects = length(unique(date)))

# Per-transect BB abundance ~ sp
bb_abund_sp <- net %>%
  dplyr::select(site, date, bb.sp) %>%
  group_by(site, date, bb.sp) %>% # 
  summarize(bb.abund = n()) %>% # get per transect bb abundance
  ungroup() %>%
  complete(site, date, bb.sp, fill = list(bb.abund = 0)) %>% # complete implicit absences
  semi_join(net, by = c("site", "date")) %>% # drop site-dates that did not exist in the data set
  left_join(site_data) %>% # add site data
  mutate(year = factor(year(date)),
         yday = yday(date),
         bb.sp = factor(bb.sp),
         mgmt = factor(management),
         site = factor(site),
         transect = factor(transect))

### Total BB abundance per transect
bb_abund <- bb_abund_sp %>%
  group_by(site, date) %>%
  summarize(bb.abund = sum(bb.abund)) %>%
  left_join(site_data) %>%
  mutate(site = factor(site),
         yday = yday(date),
         year = factor(year(date)),
         mgmt = factor(management)) %>%
  dplyr::select(-management)

### Per-transect floral abund by species
fl_abund_sp <- survey %>%
  group_by(site, date, plant.sp) %>% # group by plant.sp*transect
  summarize(fl.abund = sum(flower.cover)) %>% # get plant.sp*transect total flower cover; this step really shouldn't be required, but in a few cases, there is more than one abundance entry for the same species on a given transect; not sure how that happened, but it's in the original data; alternatively, I could drop the dups or average them. 
  group_by(site, date) %>%
  mutate(prop.fl.abund = fl.abund/sum(fl.abund)) %>% # get the proportional flower cover for each species.
  ungroup() %>%
  complete(site, date, plant.sp, fill = list(fl.abund = 0, prop.fl.abund = 0)) %>%
  semi_join(survey, by = c("site", "date")) %>%  # only include site-dates that actually occurred
  unique() %>%
  mutate(year = factor(year(date))) %>%
  left_join(fl_tax)

### Total floral abundance per transect
fl_abund <- fl_abund_sp %>%
  group_by(site, date) %>%
  summarize(fl.abund = sum(fl.abund)) %>%
  left_join(site_data) %>%
  mutate(site = factor(site),
         yday = yday(date),
         year = factor(year(date)),
         mgmt = factor(management)) %>%
  dplyr::select(-management) %>%
  mutate(year2010 = ifelse(year == "2010", 1, 0),
         year2011 = ifelse(year == "2011", 1, 0),
         year2012 = ifelse(year == "2012", 1, 0))
```

## 2.2 Calculate aggregate abundance and richness by site-year
```{r}
### Bumble bee richness per site-year
bb_rich_siteyear <- net %>%
  mutate(year = factor(year(date))) %>%
  group_by(site, year) %>%
  summarize(bb.rich = length(unique(bb.sp))) %>%
  left_join(site_data) %>% # add site data
  mutate(mgmt = factor(management),
         site = factor(site),
         transect = factor(transect)) %>%
  select(-management) %>%
  left_join(net_transects_siteyear)

### Peak BB abundance per site-year
bb_abund_siteyear <- bb_abund %>%
  group_by(site, year, elev.mean, mgmt) %>%
  summarize(bb.peakabund = max(bb.abund))

### Peak BB abundance by species per site-year
bb_abund_sp_siteyear <- bb_abund_sp %>%
  group_by(site, year, elev.mean, mgmt, bb.sp) %>%
  summarize(bb.peakabund = max(bb.abund))

### Floral richness per site-year
fl_rich_siteyear <- survey %>%
  mutate(year = factor(year(date))) %>%
  group_by(site, year) %>%
  summarize(fl.rich = length(unique(plant.sp))) %>%
  left_join(site_data) %>% # add site data
  mutate(mgmt = factor(management),
         site = factor(site),
         transect = factor(transect)) %>%
  select(-management) %>%
  left_join(survey_transects_siteyear)

### Peak floral abundance per site
fl_abund_siteyear <- fl_abund %>%
  group_by(site, year, elev.mean, mgmt) %>%
  summarize(fl.peakabund = max(fl.abund))

### Interaction richness per site
interaction_rich_siteyear <- net %>%
  mutate(year = factor(year(date))) %>%
  group_by(site, year) %>%
  unite(interaction, c(bb.sp, plant.sp), sep = "_") %>%
  summarize(interaction.rich = length(unique(interaction))) %>%
  left_join(site_data) %>% # add site data
  mutate(mgmt = factor(management),
         site = factor(site),
         transect = factor(transect)) %>%
  select(-management) %>%
  left_join(net_transects_siteyear)

### Get everything into a single data frame
rich_abund_siteyear <- bb_rich_siteyear %>% 
  full_join(fl_rich_siteyear) %>%
  full_join(bb_abund_siteyear) %>% 
  full_join(fl_abund_siteyear) %>%
  full_join(interaction_rich_siteyear)

### Long version for plotting
rich_abund_siteyear.long <- rich_abund_siteyear %>%
  pivot_longer(cols = c(bb.rich, fl.rich, bb.peakabund, 
                        fl.peakabund, interaction.rich)) %>%
  separate(name, c("group", "metric"), sep = "\\.") %>%
  group_by(group, metric) %>%
  mutate(z = scale(value))
```

# 3. Analyses

## 3.1 Preliminary visualization
```{r eval=FALSE, include=FALSE}
ggplot(rich_abund_siteyear.long, aes(elev.mean, z, color = group)) +
  geom_point() +
  geom_smooth(method = "gam", formula = y ~ s(x, k = 5)) +
  facet_grid(year~metric) +
  theme_classic()

ggplot(rich_abund_site.long, aes(z, fill = group)) +
  geom_histogram() +
  facet_wrap(~metric) +
  theme_classic()

# Exploring some bivariate relations

ggplot(rich_abund_site, aes(fl.rich, bb.rich)) +
  geom_point() +
  geom_smooth(method = "lm")

ggplot(rich_abund_site, aes(fl.peakabund, fl.rich)) +
  geom_point() +
  geom_smooth(method = "lm")

ggplot(rich_abund_site, aes(fl.peakabund, bb.peakabund)) +
  geom_point() +
  geom_smooth(method = "lm")

ggplot(rich_abund_site, aes(fl.peakabund, bb.rich)) +
  geom_point() +
  geom_smooth(method = "lm")

ggplot(rich_abund_site, aes(survey.transects, net.transects)) +
  geom_point() +
  geom_smooth(method = "lm") 
```

## 3.2 Model richness ~ elevation

Our interest in all these models is the global effect of elevation, but we need to account for the annual structure of our data. We do this by including year as a smoothing factor (interacting with elevation) that can deviate from the global effect of elevation. This represents the type GI model structure from Pedersen et al. (2019). What we will focus on is the "G" part of this model, i.e. the global elevation smoother, but I will include the year*elevation smooths in the Appendix.

It was a bit tricky to decide what distribution family to use for these models. The residuals are approximately gaussian, but richness data is count data, so theoretically we should use a zero-bounded distribution. Poisson would be a good start, but it is badly overdispersed in our models. Quasipoisson fixes that, and it gives results that are essentially identical to what we get with gaussian models. For some reason, a negative binomial model does not fix the dispersion problem.

The quasipoisson family also works well for the abundance models, and it can handle the continuous (i.e. non-integer) floral abundance data. So, we will stick with the quasipoisson family for all these models.

### 3.2.1 BB richness
```{r}
# Sanity check visualization
ggplot(bb_rich_siteyear, aes(elev.mean, bb.rich)) +
  geom_point() +
  geom_smooth(method = "gam", formula = y ~ s(x, k = 5))  +
  facet_wrap(~year, ncol = 1) +
  theme_classic()

# Call model
gam1.1 <- gam(bb.rich ~ 
                mgmt +
                year +
                s(site, bs = "re") +
                s(net.transects) +
                s(elev.mean) +
                s(elev.mean, by = year, k = 3, m = 1),
              method = "REML",
              family = "quasipoisson",
              #select = TRUE,
              data = bb_rich_siteyear)

# Summary and model diagnostics
summary(gam1.1)
check.gamViz(getViz(gam1.1))

# Plot all conditional effects
draw(gam1.1, 
     residuals = TRUE) & theme_classic()

# Extract elevation effect
gam1.1_elev <- evaluate_smooth(gam1.1, smooth = "s(elev.mean)") %>%
  filter(is.na(by_variable)) %>% # we want only the global effect
  mutate(upper = est + (2*se),
         lower = est - (2*se))

# Extract year effect
gam1.1_year <- evaluate_parametric_term(gam1.1, term = "year") %>%
  mutate(upper = partial + (2*se),
         lower = partial - (2*se))

# Extract managment effect
gam1.1_mgmt <- evaluate_parametric_term(gam1.1, term = "mgmt") %>%
  mutate(upper = partial + (2*se),
         lower = partial - (2*se))

# Plot elevation effect
gam1.1_plot1 <- ggplot(gam1.1_elev, aes(elev.mean, est)) +
  geom_line() +
  geom_ribbon(aes(ymin = lower, ymax = upper), # Uncertainty
              fill = "black", alpha = 0.2) +
  # geom_rug(data = semi_join(site_data, net), 
  #          aes(elev.mean, color = tree_line), 
  #          inherit.aes = FALSE) +
  theme_classic() +
  labs(y = "Conditional effect", x = "Elevation [m]", color = "Tree line")

# Plot year effect
gam1.1_plot2 <- ggplot(gam1.1_year, aes(value, partial)) +
  geom_errorbar(aes(ymin = lower, ymax = upper), width = 0.2) +
  geom_point() +
  theme_classic() +
  labs(y = NULL, x = NULL)

# Plot management effect
gam1.1_plot3 <- ggplot(gam1.1_mgmt, aes(value, partial)) +
  geom_errorbar(aes(ymin = lower, ymax = upper), width = 0.2) +
  geom_point() +
  theme_classic() +
  labs(y = NULL, x = NULL)

# Assemble
gam1.1_plot <- (gam1.1_plot1 | (gam1.1_plot2 / gam1.1_plot3)) 
```

### 3.2.2 FL richness
```{r}
# Sanity check visualization
ggplot(fl_rich_siteyear, aes(elev.mean, fl.rich)) +
  geom_point() +
  geom_smooth(method = "gam", formula = y ~ s(x, k = 5))  +
  facet_wrap(~year, ncol = 1) +
  theme_classic()

# Call model
gam2.1 <- gam(fl.rich ~ 
                mgmt +
                year +
                s(site, bs = "re") +
                s(survey.transects) +
                s(elev.mean) +
                s(elev.mean, by = year, k = 3, m = 1),
              method = "REML",
              family = "quasipoisson",
              #select = TRUE,
              data = fl_rich_siteyear)

# Summary and model diagnostics
summary(gam2.1)
check.gamViz(getViz(gam2.1))

# Plot all conditional effects
draw(gam2.1, 
     residuals = TRUE) & theme_classic()

# Extract elevation effect
gam2.1_elev <- evaluate_smooth(gam2.1, smooth = "elev") %>%
  filter(is.na(by_variable)) %>% # we want only the global effect
  mutate(upper = est + (2*se),
         lower = est - (2*se))

# Extract year effect
gam2.1_year <- evaluate_parametric_term(gam2.1, term = "year") %>%
  mutate(upper = partial + (2*se),
         lower = partial - (2*se))

# Extract managment effect
gam2.1_mgmt <- evaluate_parametric_term(gam2.1, term = "mgmt") %>%
  mutate(upper = partial + (2*se),
         lower = partial - (2*se))

# Plot elevation effect
gam2.1_plot1 <- ggplot(gam2.1_elev, aes(elev.mean, est)) +
  geom_line() +
  geom_ribbon(aes(ymin = lower, ymax = upper), # Uncertainty
              fill = "black", alpha = 0.2) +
  # geom_rug(data = semi_join(site_data, net), 
  #          aes(elev.mean, color = tree_line), 
  #          inherit.aes = FALSE) +
  theme_classic() +
  labs(y = "Conditional effect", x = "Elevation [m]", color = "Tree line")

# Plot year effect
gam2.1_plot2 <- ggplot(gam2.1_year, aes(value, partial)) +
  geom_errorbar(aes(ymin = lower, ymax = upper), width = 0.2) +
  geom_point() +
  theme_classic() +
  labs(y = NULL, x = NULL)

# Plot management effect
gam2.1_plot3 <- ggplot(gam2.1_mgmt, aes(value, partial)) +
  geom_errorbar(aes(ymin = lower, ymax = upper), width = 0.2) +
  geom_point() +
  theme_classic() +
  labs(y = NULL, x = NULL)

# Assemble
gam2.1_plot <- gam2.1_plot1 | (gam2.1_plot2 / gam2.1_plot3)
```

### 3.2.3 Interaction richness
```{r}
# Sanity check visualization
ggplot(interaction_rich_siteyear, aes(elev.mean, interaction.rich)) +
  geom_point() +
  geom_smooth(method = "gam", formula = y ~ s(x, k = 5))  +
  facet_wrap(~year, ncol = 1) +
  theme_classic()

# Call model
gam3.1 <- gam(interaction.rich ~ 
                mgmt +
                year +
                s(site, bs = "re") +
                s(net.transects) +
                s(elev.mean) +
                s(elev.mean, by = year, k = 3, m = 1),
              method = "REML",
              family = "quasipoisson",
              #select = TRUE,
              data = interaction_rich_siteyear)

# Summary and model diagnostics
summary(gam3.1)
check.gamViz(getViz(gam3.1))

# Plot all conditional effects
draw(gam3.1, 
     residuals = TRUE) & theme_classic()

# Extract elevation effect
gam3.1_elev <- evaluate_smooth(gam3.1, smooth = "s(elev.mean)") %>%
  filter(is.na(by_variable)) %>% # we want only the global effect
  mutate(upper = est + (2*se),
         lower = est - (2*se))

# Extract year effect
gam3.1_year <- evaluate_parametric_term(gam3.1, term = "year") %>%
  mutate(upper = partial + (2*se),
         lower = partial - (2*se))

# Extract managment effect
gam3.1_mgmt <- evaluate_parametric_term(gam3.1, term = "mgmt") %>%
  mutate(upper = partial + (2*se),
         lower = partial - (2*se))

# Plot elevation effect
gam3.1_plot1 <- ggplot(gam3.1_elev, aes(elev.mean, est)) +
  geom_line() +
  geom_ribbon(aes(ymin = lower, ymax = upper), # Uncertainty
              fill = "black", alpha = 0.2) +
  # geom_rug(data = semi_join(site_data, net), 
  #          aes(elev.mean, color = tree_line), 
  #          inherit.aes = FALSE) +
  theme_classic() +
  labs(y = "Conditional effect", x = "Elevation [m]", color = "Tree line")

# Plot year effect
gam3.1_plot2 <- ggplot(gam3.1_year, aes(value, partial)) +
  geom_errorbar(aes(ymin = lower, ymax = upper), width = 0.2) +
  geom_point() +
  theme_classic() +
  labs(y = NULL, x = NULL)

# Plot management effect
gam3.1_plot3 <- ggplot(gam3.1_mgmt, aes(value, partial)) +
  geom_errorbar(aes(ymin = lower, ymax = upper), width = 0.2) +
  geom_point() +
  theme_classic() +
  labs(y = NULL, x = NULL)

# Assemble
gam3.1_plot <- gam3.1_plot1 | (gam3.1_plot2 / gam3.1_plot3)

# Tabulate
library(flextable)
as_flextable(gam3.1)
```

### 3.1.4 Compile plots
```{r}
gam_plot_richness_ <- ( gam1.1_plot / gam2.1_plot / gam3.1_plot ) +
  #plot_annotation(tag_levels = 'A') +
  plot_layout(nrow = 3)

gam_plot_richness <- ggdraw() +
  draw_plot(gam_plot_richness_) +
  draw_image(BB, scale = 0.15, x = -0.18, y = 0.28) +
  draw_image(FL, scale = 0.125, x = -0.18, y = -0.025) +
  draw_image(BBxFL, scale = 0.25, x = -0.18, y = -0.38)

ggsave("../ms_1/figures/fig3.pdf", gam_plot_richness, width = 6.5, height = 8.5)
```

## 3.3 Model peak abundance ~ elevation
We will use the same model structure that we used for the richness models.

### 3.3.1 BB abundance
```{r}
ggplot(bb_abund_siteyear, aes(elev.mean, bb.peakabund)) +
  geom_point() +
  geom_smooth(method = "gam", formula = y ~ s(x, k = 10))  +
  facet_wrap(~year, ncol = 1) +
  theme_classic()

# Call model
gam4.1 <- gam(bb.peakabund ~ 
                mgmt +
                year +
                s(site, bs = "re") +
                s(elev.mean) +
                s(elev.mean, by = year, k = 3, m = 2),
              method = "REML",
              family = "quasipoisson",
              data = bb_abund_siteyear)

# Summary and model diagnostics
summary(gam4.1)
check.gamViz(getViz(gam4.1))

# Plot all conditional effects
draw(gam4.1, 
     residuals = TRUE) & theme_classic()

# Extract elevation effect
gam4.1_elev <- evaluate_smooth(gam4.1, smooth = "s(elev.mean)") %>%
  filter(is.na(by_variable)) %>% # we want only the global effect
  mutate(upper = est + (2*se),
         lower = est - (2*se))

# Extract year effect
gam4.1_year <- evaluate_parametric_term(gam4.1, term = "year") %>%
  mutate(upper = partial + (2*se),
         lower = partial - (2*se))

# Extract managment effect
gam4.1_mgmt <- evaluate_parametric_term(gam4.1, term = "mgmt") %>%
  mutate(upper = partial + (2*se),
         lower = partial - (2*se))

# Plot elevation effect
gam4.1_plot1 <- ggplot(gam4.1_elev, aes(elev.mean, est)) +
  geom_line(linetype = "dashed") +
  geom_ribbon(aes(ymin = lower, ymax = upper), # Uncertainty
              fill = "black", alpha = 0.2) +
  # geom_rug(data = semi_join(site_data, net), 
  #          aes(elev.mean, color = tree_line), 
  #          inherit.aes = FALSE) +
  theme_classic() +
  labs(y = "Conditional effect", x = "Elevation [m]", color = "Tree line")

# Plot year effect
gam4.1_plot2 <- ggplot(gam4.1_year, aes(value, partial)) +
  geom_errorbar(aes(ymin = lower, ymax = upper), width = 0.2) +
  geom_point() +
  theme_classic() +
  labs(y = NULL, x = NULL)

# Plot management effect
gam4.1_plot3 <- ggplot(gam4.1_mgmt, aes(value, partial)) +
  geom_errorbar(aes(ymin = lower, ymax = upper), width = 0.2) +
  geom_point() +
  theme_classic() +
  labs(y = NULL, x = NULL)

# Assemble
gam4.1_plot <- gam4.1_plot1 | (gam4.1_plot2 / gam4.1_plot3)
```

### 3.3.2 FL abundance
```{r}
ggplot(fl_abund_siteyear, aes(elev.mean, fl.peakabund)) +
  geom_point() +
  geom_smooth(method = "gam", formula = y ~ s(x, k = 5))  +
  facet_wrap(~year, ncol = 1) +
  theme_classic()

# Call model
gam5.1 <- gam(fl.peakabund ~ 
                mgmt +
                year +
                s(site, bs = "re") +
                s(elev.mean, k = 5) +
                s(elev.mean, by = year, k = 3, m = 2),
              method = "REML",
              family = "quasipoisson",
              data = fl_abund_siteyear)

# Summary and model diagnostics
summary(gam5.1)
check.gamViz(getViz(gam5.1))

# Plot all conditional effects
draw(gam5.1, 
     residuals = TRUE) & theme_classic()

# Extract elevation effect
gam5.1_elev <- evaluate_smooth(gam5.1, smooth = "s(elev.mean)") %>%
  filter(is.na(by_variable)) %>% # we want only the global effect
  mutate(upper = est + (2*se),
         lower = est - (2*se))

# Extract year effect
gam5.1_year <- evaluate_parametric_term(gam5.1, term = "year") %>%
  mutate(upper = partial + (2*se),
         lower = partial - (2*se))

# Extract managment effect
gam5.1_mgmt <- evaluate_parametric_term(gam5.1, term = "mgmt") %>%
  mutate(upper = partial + (2*se),
         lower = partial - (2*se))

# Plot elevation effect
gam5.1_plot1 <- ggplot(gam5.1_elev, aes(elev.mean, est)) +
  geom_line(linetype = "solid") +
  geom_ribbon(aes(ymin = lower, ymax = upper), # Uncertainty
              fill = "black", alpha = 0.2) +
  # geom_rug(data = semi_join(site_data, net), 
  #          aes(elev.mean, color = tree_line), 
  #          inherit.aes = FALSE) +
  theme_classic() +
  labs(y = "Conditional effect", x = "Elevation [m]", color = "Tree line")

# Plot year effect
gam5.1_plot2 <- ggplot(gam5.1_year, aes(value, partial)) +
  geom_errorbar(aes(ymin = lower, ymax = upper), width = 0.2) +
  geom_point() +
  theme_classic() +
  labs(y = NULL, x = NULL)

# Plot management effect
gam5.1_plot3 <- ggplot(gam5.1_mgmt, aes(value, partial)) +
  geom_errorbar(aes(ymin = lower, ymax = upper), width = 0.2) +
  geom_point() +
  theme_classic() +
  labs(y = NULL, x = NULL)

# Assemble
gam5.1_plot <- gam5.1_plot1 | (gam5.1_plot2 / gam5.1_plot3)
```

### 3.3.3 Compile plots
```{r}
gam_plot_abundance_ <- ( gam4.1_plot / gam5.1_plot) +
  #plot_annotation(tag_levels = 'A') +
  plot_layout(nrow = 2)

gam_plot_abundance <- ggdraw() +
  draw_plot(gam_plot_abundance_) +
  draw_image(BB, scale = 0.15, x = -0.18, y = 0.16) +
  draw_image(FL, scale = 0.2, x = -0.18, y = -0.3)

ggsave("../ms_1/figures/fig4.pdf", gam_plot_abundance, width = 6.5, height = 5.5)
```

### 3.3.4 BB sp-level abundance ~ elevation
For this model, the negative-binomial family works best. Maybe it handles zeros better than the quasipoisson.
```{r}
ggplot(bb_abund_sp_siteyear, aes(elev.mean, bb.peakabund, color = year)) +
  geom_point() +
  geom_smooth() +
  facet_wrap(~bb.sp, scales = "free")

gam6.1 <- gam(bb.peakabund ~ 
                year +
                mgmt +
                s(site, bs = "re") +
                s(bb.sp, bs = "re") +
                s(elev.mean, by = bb.sp),
              method = "REML",
              family = "nb",
              select = FALSE, # I want to visualize even the non-significant smooths
              data = bb_abund_sp_siteyear)

# Summary and model diagnostics
summary(gam6.1)
check.gamViz(getViz(gam6.1))

# Plot all conditional effects
draw(gam6.1, 
     residuals = FALSE) & theme_classic()

# Extract p-values
gam6.1_pval <- gam6.1 %>%
  tidy()

# Extract elevation effect
gam6.1_elev <- evaluate_smooth(gam6.1, smooth = "s(elev.mean)") %>%
  mutate(upper = est + (2*se),
         lower = est - (2*se)) %>%
  group_by(bb.sp) %>%
  mutate(diff = last(est) - first(est)) %>%
  left_join(gam6.1_pval, by = c("smooth" = "term")) %>%
  mutate(sig = factor(ifelse(p.value < 0.05, "p < 0.05", "ns")))

# Plot elevation effect
gam6.1_plot1 <- ggplot(gam6.1_elev, aes(elev.mean, est)) +
  geom_line(aes(linetype = fct_rev(sig))) +
  geom_ribbon(aes(ymin = lower, ymax = upper, fill = diff), # Uncertainty
              alpha = 0.5) +
  scale_fill_gradient2(low = "#4dac26", mid = "gray75", high = "#d01c8b") +
  theme_light(12) +
  labs(y = "Conditional effect", x = "Elevation [m]", fill = "Slope",
       linetype = "Significance") +
  facet_wrap(~reorder(bb.sp, -diff)) +
  theme()

ggsave("../ms_1/figures/fig5.pdf", gam6.1_plot1, height = 6.5, width = 6.5)
```

### 3.3.5 Compile tabulated summaries
```{r}
gam_tab.rich <- gam_tab1(x1 = gam1.1, x2 = gam2.1, x3 = gam3.1, 
                   name1 = "Bumble bee richness", 
                   name2 = "Floral richness", 
                   name3 = "Interaction richness")

save_as_image(gam_tab.rich, "../ms_1/figures/table1.png")

gam_tab.abund <- gam_tab2(x1 = gam4.1, x2 = gam5.1,
                   name1 = "Bumble bee abundance", 
                   name2 = "Floral abundance")

save_as_image(gam_tab.abund, "../ms_1/figures/table2.png")

gam_tab.abund_sp1 <- gam6.1  %>%
  tidy(parametric = TRUE) %>%
  mutate(class = rep("parametric", n()),
         estimate = round(estimate, 3),
         statistic = round(statistic, 3),
         p.value = round(p.value, 3)
         )

gam_tab.abund_sp2 <- gam6.1 %>%
  tidy(parametric = FALSE) %>%
  mutate(class = rep("smooth", n()),
         edf = round(edf, 2),
         statistic = round(statistic, 3),
         p.value = round(p.value, 3))

gam_tab.abund_sp <- bind_rows(gam_tab.abund_sp1, gam_tab.abund_sp2) %>%
    select(class, term, estimate, edf, statistic, p.value) %>%
    mutate(term = case_when(
      term == "(Intercept)" ~ "Intercept",
      term == "mgmtnone" ~ "mgmt:none",
      term == "mgmtmowing" ~ "mgmt:mowing",
      term == "year2011" ~ "year:2011",
      term == "year2012" ~ "year:2012",
      term == "s(site)" ~ "s(site)",
      term == "s(bb.sp)" ~ "s(species)",
      term == "s(elev.mean):bb.spgers" ~ "s(elevation:gers)",
      term == "s(elev.mean):bb.sphort" ~ "s(elevation:hort)",
      term == "s(elev.mean):bb.sphumi" ~ "s(elevation:humi)",
      term == "s(elev.mean):bb.sphypn" ~ "s(elevation:hypn)",
      term == "s(elev.mean):bb.spjone" ~ "s(elevation:jone)",
      term == "s(elev.mean):bb.splapi" ~ "s(elevation:lapi)",
      term == "s(elev.mean):bb.spmend" ~ "s(elevation:mend)",
      term == "s(elev.mean):bb.spmont" ~ "s(elevation:mont)",
      term == "s(elev.mean):bb.spmuci" ~ "s(elevation:muci)",
      term == "s(elev.mean):bb.sppasc" ~ "s(elevation:pasc)",
      term == "s(elev.mean):bb.spprat" ~ "s(elevation:prat)",
      term == "s(elev.mean):bb.sppsit" ~ "s(elevation:psit)",
      term == "s(elev.mean):bb.sppyre" ~ "s(elevation:pyre)",
      term == "s(elev.mean):bb.spsoro" ~ "s(elevation:soro)",
      term == "s(elev.mean):bb.sptelu" ~ "s(elevation:telu)",
      term == "s(elev.mean):bb.spwurf" ~ "s(elevation:wurf)"),
      sig = case_when(
        p.value < 0.001 ~ "***",
        p.value >= 0.001 & p.value < 0.01 ~ "**",
        p.value >= 0.01 & p.value < 0.05 ~ "*",
        p.value >= 0.05 ~ ""
      ),
      p.value = ifelse(p.value < 0.001, "<0.001", p.value),
      statistic = ifelse(statistic < 0.001, "<0.001", statistic),
      edf = ifelse(edf < 0.001, "<0.001", edf)) %>%
    flextable() %>%
    merge_v(j = "class") %>%
    valign(j = "class", valign = "top") %>%
    autofit()

save_as_image(gam_tab.abund_sp, "../ms_1/figures/table3.png")
```

### 3.3.6 BB elevation distribution
```{r}
bb_range <- net %>%
  left_join(site_data) %>%
  group_by(site, date, elev.mean, tree_line, bb.sp) %>%
  summarize(abund = n()) %>%
  group_by(site, elev.mean, tree_line, bb.sp) %>%
  summarize(abund = max(abund)) %>%
  filter(abund > 0) %>%
  group_by(bb.sp) %>%
  mutate(elev.floor = min(elev.mean),
         elev.ceiling = max(elev.mean),
         elev.range = elev.ceiling - elev.floor,
         elev.med = elev.floor + ((elev.ceiling - elev.floor)/2))

ggplot(bb_range, aes(reorder(bb.sp, elev.med), elev.mean)) +
  geom_line(size = 2, alpha = 0.25) +
  geom_point(aes(size = abund), alpha = 0.5) +
  theme_classic(14) +
  scale_size_continuous(name = "Peak abundance",
                        breaks = c(1, 10, 20, 40, 80)) +
  scale_color_discrete(name = "Tree line") +
  xlab("Bumble bee species") +
  ylab("Elevation (m)") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

ggsave("../output/bb_range_ms1.pdf")
ggsave("../output/bb_range_ms1.png")

# Compare bumble bee species in terms of proportional abundance (unit = site)
bb_abund_sum <- bb_abund_sp %>%
  group_by(site, bb.sp) %>%
  summarize(cum.abund = sum(bb.abund),
            mean.abund = mean(bb.abund)) %>%
  group_by(site) %>%
  mutate(prop.abund = cum.abund/sum(cum.abund)) %>%
  group_by(bb.sp) %>%
  mutate(mean.prop.abund = mean(prop.abund))

ggplot(bb_abund_sum, aes(reorder(bb.sp, -mean.prop.abund), prop.abund)) +
  geom_boxplot(fill = "gray85") +
  theme_classic() +
  ylab("Proportional abundance") +
  xlab("Bumble bee species")

ggsave("../output/bb_abundsum.pdf")
ggsave("../output/bb_abundsum.png")
```

## 3.4 Model sample-wise abundance ~ elevation * time
### 3.4.1 BB abundance
```{r}
gam7.1 <- gam(bb.abund ~ 
                year +
                mgmt +
                s(site, bs = "re") +
                te(yday, elev.mean, bs=c("gp","tp")) +
                te(yday, elev.mean, by = year, bs = c("tp", "tp"), m = 1),
              family = "nb",
              method = "REML",
              select = TRUE,
              data = bb_abund)

check.gamViz(getViz(gam7.1))
summary(gam7.1)

draw(gam7.1, 
     parametric = TRUE, 
     residuals = TRUE) & theme_classic()
```

### 3.4.2 FL abundance
```{r}
gam8.1 <- gam(fl.abund ~ 
                year +
                mgmt +
                s(site, bs = "re") +
                te(yday, elev.mean, bs=c("gp","tp")) +
                te(yday, elev.mean, by = year, bs = c("tp", "tp"), m = 1),
              family = "quasipoisson",
              method = "REML",
              select = TRUE,
              data = fl_abund)

check.gamViz(getViz(gam8.1))
summary(gam8.1)

draw(gam8.1, 
     parametric = TRUE, 
     residuals = TRUE) & theme_classic()
```