### 2021-09-22 ms_1/ms_revision1.Rmd

---
title: "The elevational structure of a mountain mutualism: modes and thresholds in the responses of bumble bees, wildflowers, and their interactions"

author: Douglas B. Sponsler^1^*, Fabrice Requier^2^, Katharina Kallnik^1^, Alice Claßen^1^, A. Fabienne Maihoff^1^, Johanna Sieger^3^, Ingolf Steffan-Dewenter^1^

# date: `r Sys.Date()`

output:
  bookdown::word_document2:
    fig_caption: yes
    fig_width: 7
    fig_height: 7
    df_print: kable
    keep_md: yes
    number_sections: no
    toc: no
    reference_docx: Word_template.docx

bibliography: 
  - references.bib
#  - knitcitations.bib
csl: ecology.csl
# csl: https://www.zotero.org/styles/ecology

---

^1^ Department of Animal Ecology and Tropical Biology, Biocenter, University of Würzburg, Würzburg, Germany

^2^ Université Paris-Saclay, CNRS, IRD, UMR Évolution, Génomes, Comportement et Écologie, 91198, Gif-sur-Yvette, France

^3^ Naturpark Frankenhöhe, Ansbach, Germany

`*` Corresponding author: douglas.sponsler@uni-wuerzburg.de

# Open Research Statement

Data are provided as private-for-peer review. This study also uses novel code. All data and code required to reproduce our analyses are archived on Dryad: https://datadryad.org/stash/share/M_4Yt9VVXTRm36FSfwwGzUs2OTXP9Fv52jjETdSu6ys

\newpage

# ABSTRACT

*Keywords*: climate, beta-diversity, elevation gradient, *Bombus*, bumble bee, mountain ecology, pollination network

```{r setup, include=FALSE, cache=FALSE, message = FALSE}

library("knitr")

### Chunk options: see http://yihui.name/knitr/options/ ###
## Text results
opts_chunk$set(echo = FALSE, warning = TRUE, message = TRUE)

## Code decoration
opts_chunk$set(tidy = TRUE, comment = NA, highlight = TRUE)

## Cache
# opts_chunk$set(cache = 2, cache.path = "ms_2/figures/cache/")

## Plots
opts_chunk$set(fig.path = "../ms_2/figures/")

```

```{r knitcitations, cache = FALSE}
library(knitcitations)
cleanbib()   
cite_options(citation_format = "pandoc")
```

# INTRODUCTION

Environmental gradients regulate the assembly of biological communities, and they have likely played a central role in generating and maintaining Earth's biological diversity [@Hutchinson1953-mo]. Patterns of species- and community-level responses to environmental gradients can shed light on the underlying mechanisms that govern community assembly and function [e.g. @Beals1969-in]. In addition to its basic significance, such insight is important for anticipating and managing the effects of climate and land use change, which can modify existing environmental gradients and reorganize the biological communities that inhabit them.

Mountain slopes generate steep environmental gradients over short geographical distances, facilitating the assembly of biological communities with high species richness and turnover [@Rahbek2019-sw]. Perhaps the most salient environmental condition that changes along a mountain elevation gradient is air temperature, which declines according to a linear lapse rate of approximately -0.6°C per 100 m [@Rolland2003-my], but mountain slopes also generate variation in other conditions, including precipitation, atmospheric pressure, atmospheric turbulence, substrate characteristics, and exposure to solar radiation [@Korner1995-mh; @Hodkinson2005-af].

Ecologists have long viewed mountain slopes as natural laboratories for the study of community assembly, and particular attention has been given to elevational patterns of species richness and composition. Species richness along mountain slopes tends to exhibit either a mid-elevation peak or a monotonic decline [@Rahbek1995-oj; @Guo2013-tu], though the exact shape of these responses can vary considerably across taxonomic groups [e.g. @Peters2016-hf]. Species composition along mountain slopes, best studied in plant communities, can exhibit a range of patterns from smooth gradations to relatively abrupt transitions, with steeper slopes tending to exhibit more discrete patterns of species turnover [@Beals1969-in]. For example, many mountains have a well-defined tree line ecotone, where plant communities exhibit accelerated species turnover and declining diversity as multiple abiotic conditions become more extreme away from the shelter of tree canopy [@Slatyer1992-ir; @Becker2007-cg; @Descombes2017-id; though see @Odland1999-vx].

<!-- In most cases, plant species composition along a mountain slope varies more or less continuously but with nonlinearities or thresholds of species turnover, suggesting an interaction of processes more complex than a simple response to the linear decline of temperature with elevation [@Beals1969-in].  -->

<!-- The peaking of richness at mid-elevations might reflect intermediate environmental conditions that favor species richness, but this pattern has also been attributed to the *mid-domain effect*, which predicts mid-elevation peaks as a geometric effect of bounded ranges independent of environmental conditions [@Colwell2000-qf]. Similarly, monotonic declines in richness with elevation could be explained by abiotic conditions becoming more stringent with elevation, but they are also consistent with a species-area effect driven by the geometric loss of surface area as a mountain's topography converges to a summit [@Rahbek1997-hj; though see @Elsen2015-gp].  -->

<!-- When discrete transitions in species composition occur along a continuous environmental gradient, they can be interpreted as points where the fitness inequality between competing species reverses, implying that interspecific competition is a primary mechanism of community assembly [@Hutchinson1953-mo].  -->

While elevational patterns in overall species richness and composition are informative, the functional significance of mountain slopes can be explored more deeply by focusing on the responses of interacting guilds of species. For example, studies of plants and insect herbivores along mountain elevation gradients have revealed a positive correlation between elevation and herbivore niche-breadth, suggesting an extension of the classical *latitude-niche-breadth hypothesis* to the analogous case of elevational gradients [@Rasmann2018-gl]. Similarly, plant-pollinator interactions along mountain slopes have been found to become less specialized as elevation increases, which may reflect a reciprocal relationship between the processes of competition and environmental filtering in community assembly [@Hoiss2015-gv; @Classen2020-nh]. A unifying theme of such studies is the understanding of *species interactions* as ecological entities *per se*, dependent upon but not reducible to the species-level richness and composition of the interacting guilds.

Bumble bees (Hymenoptera: *Bombus* spp.) and their floral hosts provide an excellent system in which to study elevational patterns at both the species- and interaction-level. Bumble bees are believed to have arisen in the mountains of Asia during the cooling climate of the Eocene-Oligocene boundary, and their subsequent spread and diversification along mountain corridors and into lowland habitats was likely driven by alternating range expansion during cooling periods and retreat to higher elevations during warming periods [@Hines2008-er; @Stewart2010-jn; @Martinet2018-rb]. Today, they remain characteristically cold-adapted organisms [@Heinrich1994-cv], with peak abundance and diversity in mountain ranges and northern latitudes [@Williams1998-xi]. In these habitats, they are often the dominant guild of flower-visiting insects and the principal pollinators of entomophilous flora [@Goulson2010-lo]. As in the past, so now are mountain habitats critical theaters of ecological change for bumble bees. Warming climates are expected to cause habitat loss and potentially novel competitive pressures [@Marshall2020-pc; @Soroye2020-zb; though see @Guzman2021-my]. At the same time, socioeconomic drivers are causing the reforestation of pastures on mountain slopes and the intensification of agriculture in lowlands and valleys, both of which entail a loss of floral resources for bumble bees [@Tasser2002-om; @Tappeiner2006-jp; @Tasser2007-qo; @Walcher2017-kw].

In this paper, we present a study based on ~13,000 interactions between bumble bees (Hymenoptera: *Bombus*) and their floral hosts recorded along a 1400 m elevation gradient, complemented by corresponding surveys of floral abundance. Using this dataset, we investigate elevational responses of richness, abundance, and β-diversity (i.e. species turnover), both at the level of species (bumble bees and flowering plants, respectively) and at the level of interactions (bumble- bee-flower pairs). Our results reveal nonlinear and, in some cases, contrasting elevational patterns, underscoring the fact that mountain slopes are complex ecological systems where multiple processes interact to produce biological responses. These patterns shed light on the basic ecology of the mountain mutualism between bumble bees and wildflowers, and they suggest how this mutualism might be reshaped by ongoing processes of climate and land use change.

# METHODS

## Field system

The study was conducted from 2010-2012 in Berchtesgaden National Park, located in the Northern Limestone Alps of southeast Germany (47.55°N, 12.92°E). The landscape is composed of mountain pastures surrounded mainly by coniferous forest. We selected 25 study sites (60 x 60 m) on mountain pastures at elevations ranging from 641-2032 meters above sea level (m.s.l.) **(Figure 1)**. Fourteen of these pastures are extensively grazed by cattle or sheep, 3 are mowed for hay production, and 8 have been abandoned throughout the last century and are no longer subject to any human management. Study sites were classified as lying either above or below the tree line based on an examination of Google Earth imagery corroborated by field experience at our study sites. Eight sites were located above the tree line, which generally fell at an elevation of around 1500 m.s.l., consistent with the previous descriptions of the tree line in the Berchtesgadener Alps [@Kostler1970-ya; @Mayer1970-zn].

Sampling in each year of the study consisted of repeated visits to each site at approximately weekly intervals. Samples were only collected during periods without rain when the air temperature was at least 6°C. During each visit, interactions between bumble bee and flowers were recorded during a 50-minute transect walk. Bumblebees observed on a given flower were counted as floral visitors. Bumble bee queens were identified to species level in the field, while workers and males were stored in individually labeled tubes in the freezer for later identification in the laboratory according to [@Amiet1996-jz]. Floral visitation by male bumble bees was recorded during visitation sampling, but we chose to analyze only visitation by queens and workers. Males are often found resting on flowers, particularly late in the year, so their presence on a flower cannot necessarily be interpreted as a foraging visit. In conjunction with visitation observations, we estimated the flower cover of each herbaceous or shrubby plant species within each 60 x 60 m study plot to the nearest 0.01 m^2^. Plant species identification followed @Lauber2007-ha and @Oberndorfer2001-ro.

Although per-sample effort was standardized across all years, the total number of samples (i.e. site-dates) and bumble bee observations varied across years **(Figure 2A-C)**. Since the start of field sampling each year was determined by snowmelt, lower sites were sampled as early as April while higher sites could not be sampled until June or July. This resulted in a temporally staggered pattern of sampling in 2010 and 2012, while in 2011 a combination of early snowmelt and a late start to field activities resulted in sampling that was approximately aligned in time across the whole elevation gradient **(Figure 2D)**.

## Data analysis

### Data processing

We quantified bumble bee abundance as the total number of recorded floral visits per bumble bee species per site-date. We treated the cryptic *Bombus terrestris/lucorum* species pair as a single morphospecies, since these two species cannot be visually distinguished, and we pooled representatives of the parasitic *Psithyrus* subgenus --- *Bombus barbutellus*, *B. bohemicus*, *B. campestris*, *B. flavidus*, *B. quadricolor*, and *B. sylvestris* --- into the species group *B. psithryus*.  *Bombus hypnorum*, *B. humilis*, and *B. lapidarius* were extremely rare in our samples, so we omitted them from abundance analyses but retained them for β-diversity analyses.

To focus our analysis on the subset if flowering plant species that are relevant to bumble bee foraging, we omitted from analysis all plant species that received no visits from bumble bees during the course of our study. Eight plant species --- *Rubus idaeus*, *Rosa canina*, *Juniperus communis*, *Larix decidua*, *Salix* sp., *Caltha palustris*, *Pulmonaria officinalis*, and *Rheum barbarum* --- were recorded as visited during interaction sampling but were missed during floral surveying. Each accounted for no more than 4 visits in total over the three years of our study, so they were also omitted from analysis.

### Elevational patterns of richness and abundance

We analyzed the richness and abundance of bumble bees, flowering plants, and their interactions, respectively, using hierarchical generalized additive models (HGAMs) [@Wood2011-pg; @Pedersen2019-kf]. To focus on elevational patterns, we collapsed within-year variation of richness and abundance at each site into annual summary values of cumulative richness and peak abundance. We then modeled these summary values, respectively, as smooth responses to elevation. Following the "GI" model form described by @Pedersen2019-kf, we specified both a global smoother and year-specific smoothers, with the latter understood as "random" deviations from the global effect. In each model, we also included `year` and `management` as fixed intercept effects, and we included `site` as a random intercept effect to avoid pseudoreplication due to repeated sampling of sites across years. In the richness models, we also added `number of transects` as a smoother to account for the effect of sampling intensity on observed richness. For all models, we used the quasipoisson (log-link) distribution family.

To explore species-level elevational responses in bumble bees, we fit a model with a species-wise elevation smoother and no global smoother, corresponding to the "I" model form of @Pedersen2019-kf. `Year` and `management` were included as fixed intercept effects and `site` and `species` were included as random intercept effects. For this model, we used the negative binomial (log link) distribution family because it yielded a better residual distribution than the quasipoisson family.

To facilitate clear interpretation with respect to our focal questions, the models above focus on elevational patterns of abundance by summarizing across seasonal time. For those interested in within-season abundance patterns and their interaction with elevation, more complex tensor-product GAMs are included in Appendix 1.

### Species and interaction β-diversity

For our analyses of species and interaction β-diversity, we pooled all samples by site irrespective of year. After pooling, we calculated β-diversity between all pairs of sites for bumble bees, flowering plants, and theirs interaction, respectively, using the binary Jaccard index. We then partitioned total interaction β-diversity into its components of species turnover (changes in species composition) and interaction rewiring (interaction turnover despite species constancy) [@Novotny2009-my; @Poisot2012-fk].

As an initial analysis, we visualized the relationship between site-wise elevation difference and each metric of β-diversity, fitting each response with a binomial regression smooth. To investigate the relationship between elevation and β-diversity more deeply, we performed a second analysis using generalized dissimilarity modeling (GDM) [@Ferrier2007-ji] to analyze the elevational variation in the β-diversity of bumble bees, flora, and their (unpartitioned) interactions, respectively. In addition to estimating the relationship between differences in a response variable and differences in a predictor variable, GDM captures the slope of this relationship over the range of the predictor variable, revealing potential variation in the amount of change in the response induced by a given change in the predictor [@Ferrier2007-ji]. This enabled us to ask whether the relationship between β-diversity and elevation exhibits thresholds or other nonlinearities. To control for potential confounding effects of sampling intensity, sampling days per site was included as a covariate. We also included the geographic distances between sites as a covariate to account for potential spatial autocorrelation. The importance of the variables in each model were assessed by a permutation test [@Fitzpatrick-2021aa].

### Software

All analyses were conducted in `R` [@RCoreTeam-2021aa]. Data handling and visualization were performed with the `tidyverse` suite [@Wickham2019-aa]. GAM analyses were performed with packages `mgcv` [@Wood_2017], `mgcViz` [@Fasiolo2018aa], and `gratia` [@Simpson2021-aa]. β-diversity was calculated with the package `vegan` [@Oksanen_2019] and visualized using the packages `ggplot2` [@Wickham2016-aa] and `ggvegan` [@Simpson2019aa]. Calculation and partitioning of interaction β-diversity were performed with the package `bipartite` [@Dormann2008-aa; Dormann2009-aa]. GDM analysis of β-diversity was performed with the package `gdm` [@Fitzpatrick-2021aa]. Data and code to reproduce these analyses are archived on Dryad: https://datadryad.org/stash/share/M_4Yt9VVXTRm36FSfwwGzUs2OTXP9Fv52jjETdSu6ys.

# RESULTS

**Check to make sure this summary is still true**

## Summary of visitation and floral survey data

We recorded a total of 12,918 bumble- bee-flower interactions (excluding males) over the three years of our study. The metaweb across all sites and dates consisted of 16 bumble bee species, 163 plant species (110 genera, 37 families), and 736 unique bumble- bee-plant interaction pairs. *Bombus pascuorum* ("pasc"), *B. pratorum* ("prat"), *B. soroeensis* ("soro"), *B. terrestris/lucorum* ("telu"), and *B. wurflenii* ("wurf") accounted for the bulk of overall bumble bee abundance, while *B. hortorum* ("hort"), *B. jonellus* ("jone"), *B. psithyrus* ("psit"), *B. monticola* ("mont"), *B. mendax* ("mend"), *B. mucidus* ("muci"), *B. pyrenaeus* ("pyre"), and *B. gerstaeckeri* ("gers") were present at moderate to low abundance. *Bombus hypnorum*, *B. humilis*, and *B. lapidarius* were recorded only sporadically. See Appendix S1 for a summary of species-wise abundances.

Floral surveying yielded a total of 352 plant species (192 genera and 52 families). Of these, 163 species (111 genera and 37 families) were observed to be visited by bumble bees.

## Elevational patterns of richness and abundance

The response of bumble bee richness to elevation exhibited an approximately linear increase up to 1600-1700 m and then approximately leveled off **(Figure 3A)**. Neither mowing nor abandonment had a significant effect on abundance relative to the reference management level of grazing (p > 0.05), but richness was significantly lower in 2012 than in the reference year of 2010 (p = 0.04) **(Table 1)**. In contrast, the response of floral richness exhibited a clear unimodal pattern, peaking between 1300-1500 m **(Figure 3B)**. Management had no significant effect on floral richness (P > 0.1), but floral richness was significantly lower in 2011 (p < 0.001) and 2012 (p = 0.03) than in 2010. The response of interaction richness to elevation was intermediate to that of bumble bee and floral richness, peaking around 1500 m and leveling off at higher elevation **(Figure 3C)**. Management did not have a significant effect on interaction richness (p > 0.1), but interaction richness was significantly higher in 2011 (p < 0.05) and lower in 2012 (p < 0.001) relative to the reference level of 2010.

Cumulative bumble bee abundance did not respond significantly to elevation (p > 0.1) or management (p > 0.1), but bumble bee abundance was significantly higher in 2011 than in the reference year of 2010 (p = 0.03) **(Figure 4A; Table 2)**. Floral abundance, however, peaked around 1000 and exhibited an approximately quadratic decline with increasing elevation (p = 0.003). Floral abundance was higher in abandoned meadows than in grazed meadows (p = 0.002) and lower in 2012 than in 2010 (p = 0.02).

At the species level, bumble bees exhibited variable and sometimes divergent responses to elevation **(Figure 5)**. *Bombus pyrenaeus*, *B. mendax*, *B. monticola*, and *B. mucidus* exhibited approximately linear increases with elevation (p < 0.05). In contrast, *Bombus psithyrus* and *B. pascuorum* exhibited approximately linear decreases with elevation (p < 0.001). *Bombus pratorum* was unique in exhibiting a strongly unimodal response to elevation peaking around 1600 m (p < 0.001). The remaining species did not respond significantly to elevation (p > 0.05).

## Species and interaction β-diversity

Floral β-diversity among our study sites was very high overall and responded steeply to elevation difference between sites **(Figure 3A)**. Even sites at similar elevation exhibited ~40% species turnover, and the most vertically distant sites (elevation difference > 1250 m) differed by more than 85%. Bumble bee β-diversity was, in comparison to floral β-diversity, both lower overall and less responsive to elevational difference between sites **(Figure 3A)**. Sites at similar elevation exhibited ~25% species turnover, and species turnover between the most vertically distant sites remained less than 50%.

Total interaction β-diversity was >75% between sites at similar elevation and approached perfect dissimilarity in the most vertically distant sites **(Figure 3B)**. Partitioning revealed that total interaction β-diversity was driven primarily by species turnover, which accounted for ~50% of total β-diversity for sites at similar elevation and ~90% of total β-diversity for the most vertically distant sites. Species turnover consisted mainly of turnover in the floral community **(Figure 3C)**, though the joint turnover of plants and bumble bees accounted for ~25% of total species turnover between the most vertically distant sites. Interaction rewiring was most significant between sites at similar elevation, where it accounted for ~30% of total β-diversity, but its share of total β-diversity declined to ~10% in the most vertically distant sites **(Figure 3B)**.

Bumble bees, flora, and their interactions exhibited marked differences in their rates of β-diversity in response to elevation. **(Figure 4)**. Bumble bee turnover was relatively high from ~600-1100 m, briefly leveled off, and then accelerated again around 1250 m. Floral Turnover was high from ~600-1000 m, gradually leveled off from ~1000-1500 m, and then accelerated again around the tree line ecotone at ~1500 m. Interaction turnover mirrored the high rates of bumble bee and floral β-diversity from ~600-1000 m, but it exhibited a markedly different pattern above 1000 m, with relatively low turnover between 1000-1750 m and then a sharp acceleration from ~1750-2000 m. The conditional effect of elevation was significant (p < 0.001) in all models, and its importance was 4-20 time higher that of sampling intensity and geographic distance **(Table 3)**. Nevertheless, geographic distance was a significant predictor in all models (p < 0.001), and sampling intensity was a significant predictor in the bumble bee (p = 0.004) and interaction models (p < 0.001). Our GDMs explained 31% of total deviance in bumble bee β-diversity, 73% of floral β-diversity, and 71% of interactions β-diversity.

# DISCUSSION

## Big picture

The core finding of our study, evident in each of our analyses, is that a linear elevation gradient can generate complex and nonlinear biological responses in a mountain bumble- bee-wildflower community in terms of richness, abundance and β-diversity, and at the level of both species and interactions.

The positive response of bumble bee richness to elevation sets bumble bees apart from most other taxa [@Rahbek1995-oj; @Guo2013-tu; @Peters2016-hf], underscoring their exceptional adaptedness to mountain environments. On higher mountains, the increase of bumble bee richness with elevation eventually gives way to a unimodal response, but peak richness is often found above 3000 m [e.g. @Williams1991-zu; @Streinzer2019-bz], ~1000 m above the highest peaks in the Berchtesgadener Alps. Thus, in our study system, the positive response of bumble bee richness to elevation encompasses the entire mountain slope, in agreement with the findings of @Minachilis2020-xd along the comparable elevation gradient (300---2700 m) of Mt. Olympus. Importantly, the positive response of bumble bee richness stands in sharp contrast to the monotonic decline of overall bee richness with elevation in the Berchtesgadener Alps [@Hoiss2012-sn], highlighting the special importance of bumble bees as pollinators where few other bees persist.

While the response of overall bumble bee abundance to elevation was approximately flat, this aggregate pattern concealed divergent species-level responses. The strongly positive responses of *B. mendax*, *B. monticola*, *B. mucidus*, *B. pyrenaeus* are consistent with the known affinity of these species for high-elevation and/or high latitude habitats [@Rasmont2010-ci]. At the other extreme, *B. pascuorum* and *B. psithyrus* exhibited a strongly negative responses to elevation. In the case of *B. pascuorum*, this pattern is likely driven by its exceptionally long nesting cycle, which is a liability under the short growing seasons associated with higher elevations [@Neumayer2009-gx]. Growing season length may also limit the abundance of *B. psithyrus* at high elevation, since cuckoo bumble bees emerge later than the queens of their host species and must successfully usurp host nests before they can begin to raise brood [@Goulson2010-lo]. *Bombus pratorum* was unique in exhibiting a strongly unimodal abundance pattern with a well-defined peak between 1500-1600 m, . @Iserbyt2008-jn describe a similar abundance pattern of *B. pratorum* in the Eyne Valley of the Pyrenees.

The roughly sigmoidal pattern of β-diversity rate in bumble bees across our elevation gradient, revealed by our GDM analysis, reflects the reciprocal abundance patterns of high- and low-elevation species described above as well as the occurrence of certain rarer species only at middle-elevation sites. It is important to note, though, that bumble bee turnover was low over all, and 8 out of the 16 species spanned the entire elevation gradient (Appendix 1).

Floral β-diversity also exhibited a sigmoidal pattern, but its inflection point clearly corresponds to the tree line ecotone at ~1500 m. This floristic importance of the tree line has been noted in previous work [e.g. @Pellissier2010-kg; @Descombes2017-id], and our study corroborates these findings specifically for the subset of flora visited by bumble bees. Interestingly, the tree line does not constitute a discrete transition to a new equilibrium, but rather a threshold at which turnover begins a process of acceleration that holds for the remainder of the elevation gradient. While the most obvious effect of the tree line is the relatively discrete transition from forest to scrub/grassland, the tree line also initiates a process of angiosperm species attrition that proceeds as a rate of about 40 species per 100 m [@Korner2003-sc], and the pattern of increasing turnover above the tree line mirrors the patterns of decreasing species richness **(Figure 3)** and floral abundance **(Figure 4)**. Thus, the bumble bees foraging above the tree line must cope with a floral community that not only turns over rapidly in composition but also declines steeply in diversity and abundance.

We saw a noteworthy contrast between the relative stability of the bumble bee community and the extreme dynamism of the floral community. The dominance of total interaction β-diversity by floral species turnover in our system is consistent with the findings of @Simanonok2014-ou in the Rocky Mountains of North America, and it is perhaps unsurprising given the exceptional cold-hardiness of bumble bees and the sensitivity of plants to temperature gradients. Nevertheless, it highlights the behavioral challenge that bumble bees have presumably faced for the entirety of their evolutionary history as mountain florivores: how to forage efficiently in a diverse floral community that exhibits rapid spatial and temporal turnover. A fascinating question that could not be answered by our sampling approach is whether mountain bumble bees adaptively forage up- or downslope, as suggested by @Lundberg1980-qk. Bumble bees have large foraging ranges and have been shown both to track resources through space and time [@Devoto2014-ay] and to cross forest matrix to reach patches of foraging habitat [@Mola2020-yc]. Assuming an average slope of 20° and a foraging range of 1 km, a bumble bee could travel up- or downslope by more than 340 m, thus spanning an elevation band nearly 700 m wide. Such 3-dimensional foraging would enable bumble bees both to exploit the elevational turnover of floral species and to track preferred species through their elevationally staggered phenology, the latter constituting a sort of physiological time-travel [@Van_Straalen1983-df].

<!-- Above the tree line, we found that flora resource availability declines sharply as flower-rich grassland and shrubland transitions to sparse alpine scree. The observation that the upslope advancement of alpine tree lines often lags behind climate warming [@Dullinger2004-qu] suggests that high elevation bumble bees could move upslope faster than their floral resources and become increasingly confined to resource-poor habitat, though the opposite pattern could also occur if the climate-tracking of bumble bees is limited by other factors, such as nest site availability [@Marshall2020-pc].  -->

<!-- Several studies have found that the introduction or removal of bumble bee species from a community can have strong effects on the fitness of other species and on patterns of floral visitation [@Nagamitsu2010-wi; @Brosi2013-mu; @Ishii2013-sm]. If elevational range shifts introduce new competitive pressures via the encroachment of low-elevation species into the historic ranges of high-elevation species . . . -->

<!-- The outcomes of range shifts in mountain bumble bee and floral communities have obvious significance from a conservation perspective [@Soroye2020-zb], but they also provide a model system in which to study the basic process of community assembly under climate change [@Hoiss2012-sn; @Classen2015-jc]. -->

# ACKNOWLEDGEMENTS

We thank the National Park Berchtesgaden for their support and the permission to work in the national park, and we thank the owners of the study sites outside the national park for access to their land. The field work of this study was performed within the framework of the joint research center FORKAST as part of the ‘Bavarian Climate Programme 2020'. D.J. McNeil provided helpful discussion of data analyses. M. Fitzpatrick kindly corresponded regarding GDM techniques.

# REFERENCES

<div id ="refs"></div>

\newpage

# TABLES

```{r fig8, echo = FALSE}
knitr::include_graphics("../ms_1/figures/table1.png")
```

**Table 1:** Tabular output of richness GAMs. Parametric model terms are summarized with an estimated intercept effect ("estimate") while smooth terms are summarized with their effective degrees of freedom ("edf"), the latter representing the complexity (i.e. "wiggliness") of the fitted spline. Elevation:year smoothers represent year-specific deviations from the global elevation smooths calculated across years.

\newpage

```{r fig9, echo = FALSE}
knitr::include_graphics("../ms_1/figures/table2.png")
```

**Table 2:** Tabular output of abundance GAMs. Parametric model terms are summarized with an estimated intercept effect ("estimate") while smooth terms are summarized with their effective degrees of freedom ("edf"), the latter representing the complexity (i.e. "wiggliness") of the fitted spline. Elevation:year smoothers represent year-specific deviations from the global elevation smooths calculated across years.

\newpage

```{r fig10, echo = FALSE}
knitr::include_graphics("../ms_1/figures/table3.png")
```

**Table 2:** Tabular output of species-level abundance GAMs. Parametric model terms are summarized with an estimated intercept effect ("estimate") while smooth terms are summarized with their effective degrees of freedom ("edf"), the latter representing the complexity (i.e. "wiggliness") of the fitted spline.

\newpage

```{r fig11, echo = FALSE}
knitr::include_graphics("../ms_1/figures/table4.png")
```

**Table 3:** Summary of GDM results for each model (bumble bees, flora, and interactions, respectively). Variable importance represents the percent of explained deviance lost when the focal variable is permuted, and it is mapped to a color ramp to aid interpretation. The p-value of each variable (Var.pval) is based on comparing the deviance explained by the unpermuted model (Model.dev.exp) to the distribution of deviance explained when the focal variable is permuted. Similarly, the model p-value (Model.pval) is based on a comparison between the deviance explained by the unpermuted model and the distribution of deviance explained when the entire predictor matrix is permuted.

\newpage

# FIGURE LEGENDS

**Figure 1:** Study sites in the Berchtesgadener Alps plotted over Google Earth imagery (https://earth.google.com/). Sites above the tree line are depecited in red and sites below the tree line in yellow. The elevation of each site is given in meters above sea level.

**Figure 2:** Total bumble bee observations per year (A), total samples per year (B), bumble bee observations per sample per year (C), and the distribution of samples through elevation and time (D). Each sample represents a standardized transect walk for a given site and date, each observation represents a recorded interaction between a bumble bee and flower.

**Figure 3:** Conditional effects of elevation, year, and management on the richness of bumble bee species (A), floral species (B), and bumble- bee-flower interactions (C). Effects are plotted on the link scale, and shaded bands depict 95\\% confidence intervals. Bumble bee and plant illustrations were obtained from http://phylopic.org/; see Appendix 1 for details and licensing.

**Figure 4:** Conditional effects of elevation, year, and management on the peak abundance of bumble bees (A) and flower cover (B). Effects are plotted on the link scale, and shaded bands depict 95\\% confidence intervals. The dashed line for the bumble bee abundance smooth indicates that this smooth did not differ significantly from a flat line (p > 0.1).

**Figure 5:** Conditional effects of elevation on the peak abundance of each bumble bee species. Effects are plotted on the link scale, and shaded bands depict 95\\% confidence intervals. Confidence bands are colored to indicated the linear slope of the response curve, and species are plotted in order of decreasing slope.

**Figure 6:** Species-level (A) and interaction-level (B, C) β-diversity plotted against elevational difference. Each point represents the β-diversity (or β-diversity partition) between a pair of sites, and lines represent the overall relationship between β-diversity and elevation difference using binomial regression smooths. Interaction β-diversity (B, C) is partitioned using Poisot et al.’s (2012) notation: WN = unpartitioned β-diversity, ST = β-diversity due to species turnover, OS =  β-diversity due to interaction rewiring, ST.h = β-diversity due to species turnover in the higher trophic level (bumble bees), ST.l = β-diversity due to species turnover in the lower trophic level (plants), and ST.lh =  β-diversity due to joint species turnover in higher and lower trophic levels.

**Figure 7:** GDM splines of species- and interaction-level β-diversity in response to elevation. The maximum height of each spline represents the conditional (i.e. partial) effect of elevation on β-diversity, and the shape of each spline represents to the rate of species turnover as it varies along the elevation gradient, with steeper parts of the curve indicating regions of the gradient over which species/interaction turnover is more rapid. Rug plots depict the locations of sites along the elevation axis and are color-coded to indicate whether each site fell above or below the tree line.

\newpage

# FIGURES

**Figure 1**

```{r fig1, echo = FALSE}
knitr::include_graphics("../ms_1/figures/fig1.png")
```

\newpage

**Figure 2**

```{r fig2, echo = FALSE}
knitr::include_graphics("../ms_1/figures/fig2.pdf")
```

\newpage

**Figure 3**

```{r fig3, echo = FALSE}
knitr::include_graphics("../ms_1/figures/fig3.pdf")
```

\newpage

**Figure 4**

```{r fig4, echo = FALSE}
knitr::include_graphics("../ms_1/figures/fig4.pdf")
```

\newpage

**Figure5**

```{r fig5, echo = FALSE}
knitr::include_graphics("../ms_1/figures/fig5.pdf")
```

\newpage

**Figure6**

```{r fig6, echo = FALSE}
knitr::include_graphics("../ms_1/figures/fig6.pdf")
```

\newpage

**Figure7**

```{r fig7, echo = FALSE}
knitr::include_graphics("../ms_1/figures/fig7.pdf")
```

<!-- ```{r write_citations, cache=FALSE, include=FALSE} -->
<!-- write.bibtex(file = "knitcitations.bib") -->
<!-- ``` -->